### 2010-10-28 smurf_paper.tex
Initial outline of paper

%------------------------------------------------------------------------------
% SMURF Paper
%------------------------------------------------------------------------------

\documentclass[useAMS,usenatbib,usegraphicx,nofootinbib]{mn2e}

\usepackage{amsmath}
\usepackage{url}
\usepackage{natbib}
\usepackage{rotating}

% --- Some user defined macros ------------------------------------------------
\newcommand{\apj}{\rm ApJ}
\newcommand{\apjl}{\rm ApJL}
\newcommand{\apjs}{\rm ApJS}
\newcommand{\aaps}{\rm A$\&$AS}
\newcommand{\aap}{\rm A$\&$A}
\newcommand{\aapr}{\rm A$\&$AR}
\newcommand{\mnras}{\rm MNRAS}
\newcommand{\aj}{\rm Astron. J.}
\newcommand{\araa}{\rm ARAA}
\newcommand{\nat}{\rm Nature}
\newcommand{\pasj}{\rm PASJ}
\newcommand{\pasp}{\rm Publ. Astron. Soc. Pac.}
\newcommand{\ASP}{\rm ASP COnference Series}
\newcommand{\CASP}{\rm Comm. Astrophys. Space Phys.}
\newcommand{\astroph}{\rm astro-ph/}

\newcommand{\snr}{SNR}

\newcommand{\scuba}{SCUBA-2}

\def\lsim{\mathrel{\lower2.5pt\vbox{\lineskip=0pt\baselineskip=0pt
           \hbox{$<$}\hbox{$\sim$}}}}

 \def\gsim{\mathrel{\lower2.5pt\vbox{\lineskip=0pt\baselineskip=0pt
           \hbox{$>$}\hbox{$\sim$}}}}

% ----------------------------------------------------------------------------

\title[SMURF: an iterative map-maker for SCUBA-2]{The Sub-Millimetre User
Reduction Facility: an iterative map-maker for SCUBA-2}

\author[Edward~L.~Chapin~et~al.]{
  \parbox[t]{\textwidth}{
    Edward~L.~Chapin$^{1}$\thanks{E-mail:~echapin@phas.ubc.ca},
    David~S.~Berry$^{2}$,
    Andrew~G.~Gibb$^{1}$,
    Tim~Jenness$^{2}$,
    Douglas~Scott$^{1}$
  }
  \\
  \\
  $^{1}$Dept. of Physics \& Astronomy, University of British Columbia,
  6224 Agricultural Road, Vancouver, B.C. V6T 1Z1, Canada\\
  $^{2}$JointAstronomy Centre, 660 N. A‘oh ̄ k ̄ Place, University Park, Hilo, Hawaii 96720, USA}

\begin{document}

\label{firstpage}

\maketitle

\begin{abstract}
  We describe the Sub-Millimetre User Reduction Facility (SMURF), an
  iterative map-maker that was developed for the Submillimetre Common
  User Bolometer Array 2 (SCUBA-2).
\end{abstract}

\begin{keywords}
bla, bla, bla
\end{keywords}

%------------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
%------------------------------------------------------------------------------

The Submillimetre Common User Bolometer Array 2
\citep[\scuba,][]{holland2006} is a new instrument for the the 15-m
James Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
can simultaneously image the sky in two broad bands centered over 450
and 850\,\micron. Ultimately, the instrument will also be equipped
with a fourier transform spectrometer (FTS), and a polarimeter
(POL). This paper describes the Submillimetre User Reduction Facility,
SMURF, a software package for reducing the imaging data, with an
emphasis on data taken during the SCUBA-2 Shared Risk Observing period
(S2SRO) which took place from February to April of 2010. The reduction
of FTS and POL data will be described at a later date once these
additional instruments have been commissioned.

Over the last twenty years, the submillimetre band (defined here to be
200--1000\,\micron) has revolutionized several important areas of
astrophysics: helping to characterize the early stages of
star-formation by identifying the dense, cold regions in molecular
clouds where stars may eventually form; discovering through blind
surveys the locations and surface density of a class of massive
star-forming galaxies in the early Universe, now referred to as
submillimetre galaxies, or SMGs; and finding debris disks around
nearby stars, the early stages of planet formation.

The most common source of emission at these wavelengths in the
Universe is the thermal emission from cool dust.

Styles of map-making: maximum likelihood
\citep[e.g.,][]{patanchon2008}; iterative -- see \citet{johnstone2000}
implementation of \citet{wright1996} to remove chop,
\citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA
analysis \citep[e.g.][]{laurent2005,scott2008}.

Since \scuba\ will ultimately have nearly 10,000 working bolometers,
considerably larger than any other existing ground, or space-spaced
bolometer cameras, SMURF has been designed to use a less accurate,
though faster and less memory-intensive iterative approach that
attempts to model and remove most of the correlated signal components,
and then regrid the residuals signals. This method is closest in
spirit to the Comprehensive Reduction Utility for SHARC-2
\citep[CRUSH,][]{kovacs2008}.

%------------------------------------------------------------------------------
\section{Data Properties}
\label{sec:data}
%------------------------------------------------------------------------------

\subsection{How \scuba\ takes measurements}

A brief summary of the instrument hitting on topics that will be
discussed later: TES operation; base temperature and pixel heaters to
compensate for sky variations; measuring flatfields; sample rate; scan
speed to move astronomical sources of interest to higher frequencies
in the time series.

\subsection{\scuba\ shared-risk observations}

This paper will be focussed on S2SRO: only one subarray at each
wavelength; fridge oscillations; working detectors.

%------------------------------------------------------------------------------
\section{The SMURF Algorithm}
\label{sec:data}
%------------------------------------------------------------------------------

Describe the bolometer signal model, and with the help of a flow-chart
show the basic procedure that SMURF follows.

\subsection{Data cleaning}
We basically need to correct steps and apply the flatfield correction.

\subsection{Correlated low-frequency noise}
At the end of the day, we iteratively fit and remove the common-mode
signal, and then any residual noise is removed with an iterative
high-pass filter. Here are the components that we know about:

\subsubsection{Fridge oscillations}
Talk about the fridge oscillation and how there might be residuals
correlated with things like the pattern of G or Tc.

\subsubsection{Sky noise}
Do we even see sky noise? Can show some results for the combined 450
and 850\,\micron\ joint PCA analysis.

\subsubsection{Magnetic field pickup}
Show that it occasionally appears in the dark squids.

\subsection{Noise estimation}

Measure a white noise level from the power spectrum for each detector
to use as a weight. In the map, use weighted standard error to
estimate noise. This is accurate for a single pixel, but there is
still spatially correlated noise in the maps which we can show with
some figures.

\subsection{Model degeneracies}

Introduce the topic of model degeneracies here, but defer their
solution to the example sections.

%------------------------------------------------------------------------------
\section{Examples}
\label{sec:data}
%------------------------------------------------------------------------------

\subsection{Point source}
Reduce a calibrator. Describe the use of a zero mask around the
source. Compare with \citet{wiebe2009} who used the same method to
constrain a maximum-likelihood map with poor cross-linking.

\subsection{Deep blind survey}
Heavy initial filtering, and only iteratively solve for map and
common-mode with edge constraints.

\subsection{Bright extended emission}
The hardest thing to do with SCUBA-2 at the moment. Try to generate a
source/background mask to assist the map solution based on thresholds
in \snr.

%------------------------------------------------------------------------------
\section{Summary and future work}
\label{sec:summary}
%------------------------------------------------------------------------------

\scuba\ is being re-commissioned. Maybe we will have some initial data
on fridge performance? If both it, and sky noise are relatively
well-behaved, might be able to restrict filtering to lower frequencies
and get larger-scale structures.

Larger array footprint also helps.

%------------------------------------------------------------------------------
\section{Acknowledgements}
%------------------------------------------------------------------------------

%------------------------------------------------------------------------------
\bibliographystyle{mn2e}
\bibliography{mn-jour,refs}
%------------------------------------------------------------------------------

%------------------------------------------------------------------------------
\appendix
\section[]{This is an appendix}
%------------------------------------------------------------------------------

\end{document}

### 2010-11-17 smurf_paper.tex
A few more notes and references

@@
 \citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA
-analysis \citep[e.g.][]{laurent2005,scott2008}.
+analysis \citep[e.g.][]{laurent2005,scott2008,aguirre2010}.
 
@@
 
+Show some time series.
+
+Show power spectra.
+
+Show cross-spectra to emphasize how correlated things are?
+
 %------------------------------------------------------------------------------
@@
 
+We express the signal observed by the $i$th bolometer as a function of time,
+\begin{equation}
+B_i(t) = E(t)[ A(t) + N(t) ] + C(t)
+\end{equation}
+where $A(t)$ is the time-varying astronomical signal produced by
+scanning the telescope across sources, $C(t)$ is a common signal
+observed by all of the detectors, $E(t)$ is the time-varying
+extinction, which is a function of the telescope elevation and
+observing conditions, and $N(t)$ represents all other noise terms. The
+common-mode, $C(t)$ has a strong
+
 \subsection{Data cleaning}
-We basically need to correct steps and apply the flatfield correction.
+
+\subsubsection{Step correction}
+
+\subsubsection{Flatfields using heater ramps}
+
+\subsubsection{Flatfields using the common-mode signal}
 
@@
 \section{Examples}
-\label{sec:data}
+\label{sec:examples}
 %------------------------------------------------------------------------------

### 2010-11-24 smurf_paper.tex
add first figure and some description of how scuba-2 works

--- 2010-10-28 smurf_paper.tex
@@
 
-\documentclass[useAMS,usenatbib,usegraphicx,nofootinbib]{mn2e}
+\documentclass[useAMS,usenatbib,nofootinbib]{mn2e}
+
+% this is needed to get pdftex to work properly!
+\usepackage[pdftex]{graphicx}
 
@@
 % --- Some user defined macros ------------------------------------------------
+
+% journals
 \newcommand{\apj}{\rm ApJ}
@@
 
+% common acronyms etc.
 \newcommand{\snr}{SNR}
-
 \newcommand{\scuba}{SCUBA-2}
 
+% these are approximately less-than and greater-than symbols
 \def\lsim{\mathrel{\lower2.5pt\vbox{\lineskip=0pt\baselineskip=0pt
@@
   6224 Agricultural Road, Vancouver, B.C. V6T 1Z1, Canada\\
-  $^{2}$JointAstronomy Centre, 660 N. A‘oh ̄ k ̄ Place, University Park, Hilo, Hawaii 96720, USA}
+  $^{2}$JointAstronomy Centre, 660 N. A`oh\={o}k\={u} Place, University
+  Park, Hilo, Hawaii 96720, USA}
 
@@
 \begin{abstract}
-  We describe the Sub-Millimetre User Reduction Facility (SMURF), an
-  iterative map-maker that was developed for the Submillimetre Common
-  User Bolometer Array 2 (SCUBA-2).
+  We describe the properties of data from the Submillimetre Common
+  User Bolometer Array 2 (SCUBA-2), and the production of maps using
+  the Sub-Millimetre User Reduction Facility (SMURF).
 \end{abstract}
@@
 \begin{keywords}
-bla, bla, bla
+chicken, chicken: chicken, chicken: chicken -- chicken
 \end{keywords}
@@
 and 850\,\micron. Ultimately, the instrument will also be equipped
-with a fourier transform spectrometer (FTS), and a polarimeter
+with a Fourier transform spectrometer (FTS), and a polarimeter
 (POL). This paper describes the Submillimetre User Reduction Facility,
@@
 
-The most common source of emission at these wavelengths in the
-Universe is the thermal emission from cool dust.
-
 Styles of map-making: maximum likelihood
@@
 implementation of \citet{wright1996} to remove chop,
-\citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA
-analysis \citep[e.g.][]{laurent2005,scott2008}.
+\citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA analysis
+\citep[e.g.][]{laurent2005,scott2008,aguirre2010}.
 
@@
 attempts to model and remove most of the correlated signal components,
-and then regrid the residuals signals. This method is closest in
+and then regrid the residuals signals. This method is similar in
 spirit to the Comprehensive Reduction Utility for SHARC-2
-\citep[CRUSH,][]{kovacs2008}.
+\citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
+Bolocam Galactic Plane Survey \citep{aguirre2010}.
 
@@
 
-A brief summary of the instrument hitting on topics that will be
-discussed later: TES operation; base temperature and pixel heaters to
-compensate for sky variations; measuring flatfields; sample rate; scan
-speed to move astronomical sources of interest to higher frequencies
-in the time series.
+Before describing the data-reduction algorithm in detail, we summarize
+the basic strategy used to make measurements with SCUBA-2.
+
+The first generation of submm bolometer cameras used NTD germanium
+thermistors (e.g., SCUBA, MAMBO, BOLOCAM, AzTEC, LABOCA...).
+Generally speaking, these devices are operated using a current bias;
+changing thermal conditions alter their impedence, resulting in a
+changing Voltage that can be measured. The bolometers are cooled so
+that they are {\em background limited} -- the dominant noise term is
+the optical load from sources outside of the cryostat (emissivity from
+telescope, atmospheric fluctuations, and astronomical sources).
+
+In contrast, SCUBA-2 belongs to a new generation of cameras that uses
+superconducting transition-edge sensors (TESs, like ACT, SPT,
+MUSTANG). These devices are held at the critical temperature, $T_C$,
+below which the material becomes a superconductor, and above which it
+becomes a normal conductor. Within this transition, tiny temperature
+fluctuations can produce huge changes in impedence. The devices are
+operated with a voltage bias, so that thermal fluctuations produce
+changing currents, which in turn generate changing local magnetic
+fields. Chains of superconducting quantum interference devices
+(SQUIDs) are used to amplify these signals, ultimately providing a
+larger changing current that may be digitized. Like the NTD
+thermistor-based cameras, SCUBA-2 is cooled in order to be background
+limited. However, a superior over-all sensitivity on a per-bolometer
+basis is achieved through its improved responsivity (similar to an
+improved quantum efficiency for a CCD).
+
+The load on the detectors can vary substatially, both from thermal
+sources within and outside the SCUBA-2 cryostat. In order to operate
+the camera across a broad dynamic range of conditions, it is therefore
+necessary to thermally control the focal plane. This is achieved by
+providing a fridge {\em base temperature}, $T_B$ that is well-below
+$T_C$, and then using heaters that are connected to each bolometer to
+place the array in to transition.
+
+The focal plane is divided into four quadrants, each filled with 32
+column $\times$ 40 row sub-arrays labelled s4a--s4d at 450\,\micron,
+and s8a--s8d at 850\,\micron. To avoid independent bias and readout
+wires to each bolometer, each array is multiplexed by column. There is
+an additional row of dark SQUIDs with no TESs or thermal absorber used
+to monitor non-thermal noise sources.
+
+The camera has been designed so that the absolute response of each
+detector can be established by ramping currents through the individual
+pixel heaters to produce Joule heating power, $P_\mathrm{J} =
+\epsilon I^2R$, where $I$ is the known current, $R$ is the heater
+resistance, and $\epsilon$ is the efficiency with which power is
+transferred to the bolometer. In order to calibrate the total factor
+$\epsilon R$, SCUBA-2 posesses an internal flatfield source.
+
+For further detail see...
 
@@
 This paper will be focussed on S2SRO: only one subarray at each
-wavelength; fridge oscillations; working detectors.
+wavelength; fridge oscillations; working detectors (and variation from
+setup-to-setup).
+
+Show some time series.
+
+Show power spectra.
+
+Show cross-spectra to emphasize how correlated things are?
+
+\begin{figure}
+\centering
+\includegraphics[width=0.99\linewidth]{bolos_point_mix.pdf}
+\caption{A comparison between single bolometer time series in each
+  band (top two plots), with the mixing chamber temperature and
+  azimuth/elevation pointing offsets. There is a strong correlation
+  between the bolometers and the roughly $\sim$30\,s oscillation in
+  the fridge, but not with the telescope motion.}
+\label{fig:bolos_mix}
+\end{figure}
 
@@
 
+We express the signal observed by the $i$th bolometer as a function of time,
+\begin{equation}
+B_i(t) = G_i[E_i(t) A_i(t) + N_i(t)]
+\end{equation}
+where $A(t)$ is the time-varying signal produced by scanning the
+telescope across astronomical sources, $E(t)$ is the time-varying
+extinction, which is a function of the telescope elevation and
+atmospheric conditions, and $N_i(t)$ represents sources of noise. The
+two terms in square brackets, as written, have units of power
+delivered to the detectors (pW). The scale factor $G_i$ converts this
+effective power to the digitized units recorded on disk (DAC).
+
+In practice, $G_i$ is further factored into two components: $G_i = R_i
+D$. The responsivity, $R_i$, gives the change in current as a function
+of power fluctuations (A\,pW$^{-1}$). There is then a final (and
+constant across all bolometers) conversion factor from current to DAC
+units, $D$ (DAC/A).
+
 \subsection{Data cleaning}
-We basically need to correct steps and apply the flatfield correction.
+
+\subsubsection{Step correction}
+
+\subsubsection{Flatfields}
+\label{sec:flatfields}
+
+\begin{itemize}
+
+\item Heater ramps
+
+\item common-mode signal amplitude
+
+\item astronomical sources
+
+\item flat-field source
+
+\end{itemize}
 

### 2010-11-24 smurf_paper.tex
update notes
more detailed outline and added information about input data files

@@
 
+\textbf{Guideline for figures:}
+
+\begin{itemize}
+\item minimize excess whitespace around the figure to optimize usage of space
+\item ensure a decent line thickness
+\item use a Times font for labels with a thickness that resembles the
+  text in the body. The font size should be the same as the main body
+  text or larger
+\item avoid using colour unless it is necessary for emphasis
+\end{itemize}
+
+\textbf{start intro here...}
+
 The Submillimetre Common User Bolometer Array 2
@@
 
+My thought is that this section would give the basics about how
+SCUBA-2 works, and then more specific details for the S2SRO
+observations.
+
+%--------------------------------------------------
 \subsection{How \scuba\ takes measurements}
+%--------------------------------------------------
 
@@
 superconducting transition-edge sensors (TESs, like ACT, SPT,
-MUSTANG). These devices are held at the critical temperature, $T_C$,
-below which the material becomes a superconductor, and above which it
-becomes a normal conductor. Within this transition, tiny temperature
-fluctuations can produce huge changes in impedence. The devices are
-operated with a voltage bias, so that thermal fluctuations produce
-changing currents, which in turn generate changing local magnetic
-fields. Chains of superconducting quantum interference devices
-(SQUIDs) are used to amplify these signals, ultimately providing a
-larger changing current that may be digitized. Like the NTD
-thermistor-based cameras, SCUBA-2 is cooled in order to be background
-limited. However, a superior over-all sensitivity on a per-bolometer
-basis is achieved through its improved responsivity (similar to an
-improved quantum efficiency for a CCD).
+MUSTANG). These devices are held at the critical temperature,
+$T_\mathrm{C}$, below which the material becomes a superconductor, and
+above which it becomes a normal conductor. Within this transition,
+tiny temperature fluctuations can produce huge changes in
+impedence. The devices are operated with a voltage bias, so that
+thermal fluctuations produce changing currents, which in turn generate
+changing local magnetic fields. Chains of superconducting quantum
+interference devices (SQUIDs) are used to amplify these signals,
+ultimately providing a larger changing current that may be
+digitized. Like the NTD thermistor-based cameras, SCUBA-2 is cooled in
+order to be background limited. However, a superior over-all
+sensitivity on a per-bolometer basis is achieved through its improved
+responsivity (similar to an improved quantum efficiency for a CCD).
 
@@
 necessary to thermally control the focal plane. This is achieved by
-providing a fridge {\em base temperature}, $T_B$ that is well-below
-$T_C$, and then using heaters that are connected to each bolometer to
-place the array in to transition.
+providing a fridge {\em base temperature}, $T_\mathrm{B}$ that is
+well-below $T_\mathrm{C}$, and then using heaters that are connected
+to each bolometer to place the array in to transition.
 
@@
 
+%-------------------------------------------------
 \subsection{\scuba\ shared-risk observations}
+%-------------------------------------------------
 
@@
   between the bolometers and the roughly $\sim$30\,s oscillation in
-  the fridge, but not with the telescope motion.}
+  the fridge, but not with the telescope motion. Also note that the
+  total power in this primary signal is similar at both 450 and
+  850\,\micron, pointing to an internal rather than an external
+  source.}
 \label{fig:bolos_mix}
@@
 
-%------------------------------------------------------------------------------
-\section{The SMURF Algorithm}
-\label{sec:data}
-%------------------------------------------------------------------------------
+\subsubsection{Fridge oscillations}
+As shown in Fig.~\ref{fig:bolos_mix}, the dominant signal in both
+bands is clearly an oscillation in the fridge base temperature as
+evidenced by the low-resolution mixing chamber thermometer. The total
+power delivered to the bolometers is comparable at each wavelength,
+further suggesting a source internal to the cryostat, rather than an
+external optical load. Finally, there is no obvious correspondance
+with either the azimuth or elevation motion.
 
-Describe the bolometer signal model, and with the help of a flow-chart
-show the basic procedure that SMURF follows.
-
-We express the signal observed by the $i$th bolometer as a function of time,
-\begin{equation}
-B_i(t) = G_i[E_i(t) A_i(t) + N_i(t)]
-\end{equation}
-where $A(t)$ is the time-varying signal produced by scanning the
-telescope across astronomical sources, $E(t)$ is the time-varying
-extinction, which is a function of the telescope elevation and
-atmospheric conditions, and $N_i(t)$ represents sources of noise. The
-two terms in square brackets, as written, have units of power
-delivered to the detectors (pW). The scale factor $G_i$ converts this
-effective power to the digitized units recorded on disk (DAC).
-
-In practice, $G_i$ is further factored into two components: $G_i = R_i
-D$. The responsivity, $R_i$, gives the change in current as a function
-of power fluctuations (A\,pW$^{-1}$). There is then a final (and
-constant across all bolometers) conversion factor from current to DAC
-units, $D$ (DAC/A).
-
-\subsection{Data cleaning}
-
-\subsubsection{Step correction}
-
-\subsubsection{Flatfields}
-\label{sec:flatfields}
-
-\begin{itemize}
-
-\item Heater ramps
-
-\item common-mode signal amplitude
-
-\item astronomical sources
-
-\item flat-field source
-
-\end{itemize}
-
-\subsection{Correlated low-frequency noise}
-At the end of the day, we iteratively fit and remove the common-mode
-signal, and then any residual noise is removed with an iterative
-high-pass filter. Here are the components that we know about:
-
-\subsubsection{Fridge oscillations}
-Talk about the fridge oscillation and how there might be residuals
-correlated with things like the pattern of G or Tc.
+Show what's left over when we remove the common-mode, and the
+correlation of the resulting signal with patterns in $G$ and
+$T_\mathrm{C}$ across the array.
 
@@
 
+\subsubsection{Line features}
+
+The power spectra also show some narrower line features (many
+correlated across the array). Even though we largely ignore them worth
+some mention.
+
+%------------------------------------------------------------------------------
+\section{The SMURF Algorithm}
+\label{sec:algorithm}
+%------------------------------------------------------------------------------
+
+Describe the bolometer signal model, and with the help of a flow-chart
+show the basic procedure that SMURF follows.
+
+We express the signal observed by the $i$th bolometer as a function of time,
+\begin{equation}
+\mathbf{b}_i(t) = g_i[\mathbf{e}_i(t) \mathbf{a}_i(t) + \mathbf{n}_i(t)]
+\end{equation}
+where $\mathbf{a}(t)$ is the time-varying signal produced by scanning
+the telescope across astronomical sources, $\mathbf{e}(t)$ is the
+time-varying extinction, which is a function of the telescope
+elevation and atmospheric conditions, and $\mathbf{n}_i(t)$ represents
+sources of noise. The two terms in square brackets, as written, have
+units of power delivered to the detectors (pW). The scale factor $g_i$
+converts this effective power to the digitized units recorded on disk
+(DAC), and in this formulation is assumed to be constant in time.
+
+In practice, $g_i$ is further factored into two components: $g_i = r_i
+d$. The responsivity, $r_i$, gives the change in current as a function
+of power fluctuations (A\,pW$^{-1}$). There is then a final (and
+constant across all bolometers) conversion factor from current to DAC
+units, $d$ (DAC/A).
+
+As for the noise, $\mathbf{n}_i(t)$, we break it down into several
+components,
+%
+\begin{equation}
+  \mathbf{n}_i(t) = \mathbf{n}^\mathrm{w}_i(t) + \mathbf{n}^\mathrm{f}_i(t) +
+  \mathbf{n}^\mathrm{c}(t),
+\end{equation}
+%
+where $\mathbf{n}^\mathrm{w}_i(t)$ is uncorrelated white (thermal)
+noise, $\mathbf{n}^\mathrm{f}_i(t)$ is low-frequency noise that is
+\emph{not} correlated from bolometer-to-bolometer, and
+$\mathbf{n}^\mathrm{c}(t)$ is a (predominantly low-frequency)
+correlated or \emph{common-mode} component.
+
+The primary goal of map-making is to model and remove
+$\mathbf{n}^\mathrm{f}_i$ and $\mathbf{c}^\mathrm{f}_i$ from the
+time-series. By re-gridding the remaining data we can then hope to
+approach the theoretical noise limit in the map.
+
+%-------------------------------------------------
+\subsection{Model components}
+%-------------------------------------------------
+
+Give details about each model component, roughly in the order they are
+calculated.
+
+\subsection{Common-mode estimation}
+
+DSB
+
+We estimate the common-mode, remove it from the time series, use it to
+flag bad bits of data, check for time-varying bolometer
+responsivities... what doesn't it do?!
+
+\subsection{Iterative filtering}
+
+EC
+
+The FFT filter can do high/low-pass filtering. We can also do notches
+and whitening.
+
+TJ
+
+We also have the SMO median filter that's useful for very short scans
+because edge effects don't matter so much.
+
 \subsection{Noise estimation}
@@
 Measure a white noise level from the power spectrum for each detector
-to use as a weight. In the map, use weighted standard error to
-estimate noise. This is accurate for a single pixel, but there is
-still spatially correlated noise in the maps which we can show with
-some figures.
+to use as a weight.
+
+\subsection{Map estimation}
+
+In the map, use weighted standard error to estimate noise. This is
+accurate for a single pixel, but there is still spatially correlated
+noise in the maps which we can show with some figures.
 
@@
 solution to the example sections.
+
+%-------------------------------------------------
+\subsection{Data cleaning}
+%-------------------------------------------------
+
+Since we can do most of this stuff either as a pre-processing step or
+iteratively, I thinks this goes in a generic-sounding subsection. This
+can go here, or perhaps earlier?
+
+\subsubsection{Time-domain de-spiking}
+
+DSB
+
+\subsubsection{Step correction}
+
+DSB
+
+\subsubsection{Gap filling / apodization}
+
+DSB
+
+Discuss gap filling, ``new'' window function method, also apodization
+and start/end polynomial continuity stuff. Should reference CMB
+literature (namely Stompor et al.).
+
+\subsubsection{Flatfields}
+\label{sec:flatfields}
+
+TJ
+
+Discuss heater ramps, normalizing off the common-mode signal
+amplitude, astronomical sources, and the internal flatfield source.
+
+\subsubsection{Additional bad bolometer rejection}
+
+TJ
+
+Measuring the noise power spectra and throwing away outliers. Data
+flagged bad by the DA system.
+
+\subsubsection{Map-based despiking}
+
+EC
 
@@
 source/background mask to assist the map solution based on thresholds
-in \snr.
+in \snr. Compare with BOLOCAM stuff.
 
@@
 Larger array footprint also helps.
+
+Talk about potential of Pascale et al. method to get closer to
+max-likelihood map solution (handle maps at constant scan angles with
+weights in Fourier space separately).
 

### 2010-11-24 smurf_paper.tex
add flowchart to describe iterative map-maker

@@
 \newcommand{\scuba}{SCUBA-2}
+\newcommand{\rms}{RMS}
 
@@
 \item use a Times font for labels with a thickness that resembles the
-  text in the body. The font size should be the same as the main body
+  text in the body. The font size should be similar to the main body
   text or larger
@@
 \centering
-\includegraphics[width=0.99\linewidth]{bolos_point_mix.pdf}
+\includegraphics[width=\linewidth]{bolos_point_mix.pdf}
 \caption{A comparison between single bolometer time series in each
@@
   azimuth/elevation pointing offsets. There is a strong correlation
-  between the bolometers and the roughly $\sim$30\,s oscillation in
+  between the bolometers and the roughly $\sim$25\,s oscillation in
   the fridge, but not with the telescope motion. Also note that the
@@
 approach the theoretical noise limit in the map.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.75\linewidth]{dimm.pdf}
+\caption{Typical map-making algorithm. Raw data (stored in multiple
+  files) are read and concatenated into continuous time series. A
+  pre-processing stage repairs DC steps, applies a flatfield
+  correction, identifies and removes the noisiest bolometers, and
+  flags large spikes. The iterative portion then begins: estimating
+  and removing the common-mode signal; applying the extinction
+  correction; high-pass filtering to remove residual independent
+  low-frequency noise; estimating the map by re-gridding the data; and
+  finally measuring the noise in the residual time series. If the
+  \rms\ has converged, the final map is written to disk. Otherwise any
+  multiplicative factors that may have been applied to the data are
+  inverted (e.g., extinction correction and relative gains measured
+  during the common-mode calculation), and each additive component is
+  added back into the time series and re-estimated in turn.}
+\label{fig:dimm}
+\end{figure}
 

### 2010-11-24 smurf_paper.tex
Add POL-2 and FTS-2 references to para 1 and use ISO date order.

@@
 and 850\,\micron. Ultimately, the instrument will also be equipped
-with a Fourier transform spectrometer (FTS), and a polarimeter
-(POL). This paper describes the Submillimetre User Reduction Facility,
+with a Fourier transform spectrometer \citep[FTS-2,][]{gom2010}, and a polarimeter
+\cite[POL-2,][]{bastien2005}. This paper describes the Submillimetre User Reduction Facility,
 SMURF, a software package for reducing the imaging data, with an
-emphasis on data taken during the SCUBA-2 Shared Risk Observing period
-(S2SRO) which took place from February to April of 2010. The reduction
-of FTS and POL data will be described at a later date once these
+emphasis on data taken during the SCUBA-2 Shared-Risks Observing period
+(S2SRO) which took place from 2010 February 11 to 2010 March 25. The reduction
+of FTS-2 and POL-2 data will be described at a later date once these
 additional instruments have been commissioned.

### 2010-11-25 smurf_paper.tex
Use unix line endings
So this changes every line in the file...
A test commit - remove a duplicated "the"
Correct symbol used to refer to common mode noise following eqn 2

@@
 The Submillimetre Common User Bolometer Array 2
-\citep[\scuba,][]{holland2006} is a new instrument for the the 15-m
+\citep[\scuba,][]{holland2006} is a new instrument for the 15-m
 James Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
@@
 The primary goal of map-making is to model and remove
-$\mathbf{n}^\mathrm{f}_i$ and $\mathbf{c}^\mathrm{f}_i$ from the
+$\mathbf{n}^\mathrm{f}_i$ and $\mathbf{n}^\mathrm{c}$ from the
 time-series. By re-gridding the remaining data we can then hope to

### 2010-11-25 smurf_paper.tex
SMO can be mean or median
more outline stuff + added stompor reference

@@
 
+\subsubsection{Scan strategy}
+
 %-------------------------------------------------
@@
 
-Describe the bolometer signal model, and with the help of a flow-chart
+\begin{itemize}
+
+\item Describe the bolometer signal model, and with the help of a flow-chart
 show the basic procedure that SMURF follows.
+
+\item Discuss SMURF configurability
+
+\item Some sort of proof/demonstration showing why the algorithm works
+
+\item Discuss performance (execution time, memory, disk usage etc.)
+
+\item Where do you get SMURF/starlink, usage of NDF
+
+\item Talk about pipeline?
+
+\end{itemize}
 
@@
   pre-processing stage repairs DC steps, applies a flatfield
-  correction, identifies and removes the noisiest bolometers, and
-  flags large spikes. The iterative portion then begins: estimating
-  and removing the common-mode signal; applying the extinction
-  correction; high-pass filtering to remove residual independent
-  low-frequency noise; estimating the map by re-gridding the data; and
-  finally measuring the noise in the residual time series. If the
-  \rms\ has converged, the final map is written to disk. Otherwise any
-  multiplicative factors that may have been applied to the data are
-  inverted (e.g., extinction correction and relative gains measured
-  during the common-mode calculation), and each additive component is
-  added back into the time series and re-estimated in turn.}
+  correction, and identifies and removes the noisiest bolometers. The
+  iterative portion then begins: estimating and removing the
+  common-mode signal; applying the extinction correction; high-pass
+  filtering to remove residual independent low-frequency noise;
+  estimating the map by re-gridding the data; and finally measuring
+  the noise in the residual time series. If the \rms\ has converged,
+  the final map is written to disk. Otherwise any multiplicative
+  factors that may have been applied to the data are inverted (e.g.,
+  extinction correction and relative gains measured during the
+  common-mode calculation), and each additive component is added back
+  into the time series and re-estimated in turn.}
 \label{fig:dimm}
@@
 
-We also have the SMO median filter that's useful for very short scans
-because edge effects don't matter so much.
+We also have the SMO filter that's useful for very short scans
+because edge effects don't matter so much. Two sorts - median which is
+more robust, or mean which is more accurate.
 
@@
 and start/end polynomial continuity stuff. Should reference CMB
-literature (namely Stompor et al.).
+literature -- in particular look at \citet{stompor2002}.
 

### 2010-11-25 data/prepdata.sh
Add shell script to generate input data for plots from raw data

@@ -1,0 +1,45 @@
+#!/bin/tcsh
+#+
+#  Script for preparing data for plots. Currently set up to run on octopus,
+#  but you can fiddle dataroot which is the path to a tree resembling
+# /jcmtdata/raw/scuba2 at JAC
+#
+
+set dataroot = /scuba2
+set msg = quiet
+
+# the following outputs are used by intro.pro ---------------------------------
+
+echo "*** Processing 20091214_00015 ***"
+
+echo "    Concatenating..."
+
+$SMURF_DIR/sc2concat "$dataroot/s4a/20091214/00015/s4a20091214_00015_000[23]" \
+    s4a20091214_00015_con msg_filter=$msg
+
+$SMURF_DIR/sc2concat "$dataroot/s8d/20091214/00015/s8d20091214_00015_000[23]" \
+    s8d20091214_00015_con msg_filter=$msg
+
+echo "    Cleaning..."
+
+$SMURF_DIR/sc2clean s4a20091214_00015_con s4a20091214_00015_con_clean \
+    config=^$STARLINK_DIR/share/smurf/dimmconfig.lis msg_filter=$msg
+
+$SMURF_DIR/sc2clean s8d20091214_00015_con s8d20091214_00015_con_clean \
+    config=^$STARLINK_DIR/share/smurf/dimmconfig.lis msg_filter=$msg
+
+echo "    Convert to FITS..."
+
+if( -e s4a20091214_00015_con_clean.fits ) then
+  rm s4a20091214_00015_con_clean.fits
+endif
+
+$CONVERT_DIR/ndf2fits s4a20091214_00015_con_clean \
+    s4a20091214_00015_con_clean.fits msg_filter=$msg
+
+if( -e s8d20091214_00015_con_clean.fits ) then
+  rm s8d20091214_00015_con_clean.fits
+endif
+
+$CONVERT_DIR/ndf2fits s8d20091214_00015_con_clean \
+    s8d20091214_00015_con_clean.fits msg_filter=$msg

### 2010-11-25 data/prepdata.sh
also need to get the jcmtstate

@@ -43,3 +43,7 @@ endif
 
 $CONVERT_DIR/ndf2fits s8d20091214_00015_con_clean \
     s8d20091214_00015_con_clean.fits msg_filter=$msg
+
+echo "    Getting JCMTState..."
+
+$SMURF_DIR/jcmtstate2cat s4a20091214_00015_con > state_20091214_00015.tst

### 2010-11-25 data/prepdata.sh
change first plot to example with no source, and make plot of power spectra

@@ -1,49 +1,0 @@
-#!/bin/tcsh
-#+
-#  Script for preparing data for plots. Currently set up to run on octopus,
-#  but you can fiddle dataroot which is the path to a tree resembling
-# /jcmtdata/raw/scuba2 at JAC
-#
-
-set dataroot = /scuba2
-set msg = quiet
-
-# the following outputs are used by intro.pro ---------------------------------
-
-echo "*** Processing 20091214_00015 ***"
-
-echo "    Concatenating..."
-
-$SMURF_DIR/sc2concat "$dataroot/s4a/20091214/00015/s4a20091214_00015_000[23]" \
-    s4a20091214_00015_con msg_filter=$msg
-
-$SMURF_DIR/sc2concat "$dataroot/s8d/20091214/00015/s8d20091214_00015_000[23]" \
-    s8d20091214_00015_con msg_filter=$msg
-
-echo "    Cleaning..."
-
-$SMURF_DIR/sc2clean s4a20091214_00015_con s4a20091214_00015_con_clean \
-    config=^$STARLINK_DIR/share/smurf/dimmconfig.lis msg_filter=$msg
-
-$SMURF_DIR/sc2clean s8d20091214_00015_con s8d20091214_00015_con_clean \
-    config=^$STARLINK_DIR/share/smurf/dimmconfig.lis msg_filter=$msg
-
-echo "    Convert to FITS..."
-
-if( -e s4a20091214_00015_con_clean.fits ) then
-  rm s4a20091214_00015_con_clean.fits
-endif
-
-$CONVERT_DIR/ndf2fits s4a20091214_00015_con_clean \
-    s4a20091214_00015_con_clean.fits msg_filter=$msg
-
-if( -e s8d20091214_00015_con_clean.fits ) then
-  rm s8d20091214_00015_con_clean.fits
-endif
-
-$CONVERT_DIR/ndf2fits s8d20091214_00015_con_clean \
-    s8d20091214_00015_con_clean.fits msg_filter=$msg
-
-echo "    Getting JCMTState..."
-
-$SMURF_DIR/jcmtstate2cat s4a20091214_00015_con > state_20091214_00015.tst

### 2010-12-01 smurf_paper.tex
add power spectra and principal component analysis plots
detailed captions for PSD and PCA figures

@@
   850\,\micron, pointing to an internal rather than an external
-  source.}
+  source. The grey signals overplotted in the top-two panels show the
+  residual time-series after removing the common-mode signal. This
+  operation shows: (i) that most of the low-frequency signal is common
+  to all of the bolometers; and (ii) the non-correlated, and
+  predominantly white noise at 450\,\micron\ is significantly larger
+  than at 850\,\micron.}
 \label{fig:bolos_mix}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{pspec.pdf}
+\caption{Bolometer power spectral densities. Four representative
+  bolometers have been selected from each sub-array, and the
+  flat-fielded (but otherwise raw) PSDs are shown as coloured dotted
+  lines (the blue signals are for the same time series as those shown
+  in Fig.~\ref{fig:bolos_mix}). Horizontal dotted lines at $5 \times
+  10^{-8}$ and $5 \times 10^{-9}$\,pW\,Hz$^{-1}$ at 450 and
+  850\,\micron\ respectively, give the approximate white-noise
+  levels. There are broad line features in the PSDs at both
+  wavelengths above $\sim$35\,Hz. From $\sim$50\,Hz up to the Nyquist
+  frequency, the gradual roll-off of the anti-aliasing filter is also
+  evident.  At lower frequencies, the bolometer signals exhibit clear
+  $1/f$ knees at approximately 2 and 4\,Hz at 450 and
+  850\,\micron. The solid black lines are the PSDs of the common-mode
+  signals at each wavelength, and the solid coloured lines show the
+  PSDs of the bolometers once the common-mode is removed. These
+  residual signals have significantly lower $1/f$ knees, approximately
+  0.4 and 2\,Hz at 450 and 850\,\micron. Finally, the dashed black
+  lines show the spectral shape produced by a point source, assuming
+  that the telescope is slewing at 200\,arcsec\,sec$^{-1}$ (typical
+  for S2SRO); for reference, the top horizontal axis shows the
+  conversion from frequency to angular scale. We can see that the line
+  features may add significant noise on the scale of the PSF. We also
+  see that, for this scan velocity, the low-frequency noise begins to
+  dominate on scales of approximately 8.3 and 1.7\,arcmin; these scales
+  are comparable to the array footprint.}
+\label{fig:pspec}
 \end{figure}
@@
 some mention.
+
+\begin{figure*}
+\centering
+\includegraphics[width=\linewidth]{pca.pdf}
+\caption{The first six modes from a principal component analysis,
+  ordered by decreasing significance, of the combined 450 and
+  850\,\micron\ time-series bolometer data. For each component, the
+  top plot shows the time series of the eigenvector, normalized by its
+  \rms. The bottom coloured panels show the eigenvalues for the
+  bolometers at each wavelength (the amplitude of the eigenvector in
+  the time series). For reference, both the mean, $\bar{\lambda}$, and
+  \rms, $\lambda_\mathrm{rms}$, eigenvalues for the bolometers in each
+  subarray are also shown. The first two components have comparable
+  amplitudes, and dominate by at least a factor of 10 any other
+  components of the time series. We suspect that they are caused by
+  atmospheric variations (expected to be brighter at 450\,\micron),
+  and fridge oscillations (expected to be comparable at each
+  wavelength). However, the two eigenvectors do not necessarily map
+  directly to these physical sources under this interpretation; rather
+  they are two \emph{different} linear combinations of the underlying
+  physical signals that give rise to statistically uncorrelated
+  components. The third component accurately reproduces an apparent
+  data-acquisition glitch that appears only at 450\,\micron\ (see
+  Section ???). The fourth component is a ...\,Hz line feature only
+  observed in column 15 of the 450\,\micron\ subarray. The fifth
+  component shows a series of steps and glitches that appear
+  predominantly along the bottom-left edge of the 850\,\micron\
+  subarray where the bolometers are only tenuously locked. Finally,
+  the sixth component is degenerating into a less significant, and
+  more randomly scattered signal across the focal plane. Subsequent
+  components follow this trend, and would generally be kept in a
+  PCA-cleaning procedure.}
+
+\label{fig:pca}
+\end{figure*}
 

### 2010-12-02 smurf_paper.tex
flesh-out text describing S2SRO data including time series and PSDs

@@
   We describe the properties of data from the Submillimetre Common
-  User Bolometer Array 2 (SCUBA-2), and the production of maps using
-  the Sub-Millimetre User Reduction Facility (SMURF).
+  User Bolometer Array 2 (SCUBA-2) taken during the Shared Risk
+  Observing period from 2010 February 11 to 2010 March 25, and the
+  production of maps using the Sub-Millimetre User Reduction Facility
+  (SMURF). Our iterative approach is similar to those used to reduce
+  data from other ground-based submillimetre telescopes in recent
+  years, and recovers angular scales up to the size of the array
+  footprint (approximately 2\,arcmin). The method is also able to
+  achieve the theoretical white-noise limit of the instrument for
+  point-source studies.
 \end{abstract}
@@
 
-The load on the detectors can vary substatially, both from thermal
+The load on the detectors can vary substantially, both from thermal
 sources within and outside the SCUBA-2 cryostat. In order to operate
@@
 
-For further detail see...
+In general, bolometers have two predominant noise components: (i) the
+effectively white thermal noise produced by the phonon and optical
+loading on the absorbers; and (ii) and longer-timescale drift
+resulting from slow variations in the mean load (e.g. fridge,
+telescope, or atmospheric changes), and also the readout
+electronics. As such, the spectrum corresponding to a bolometer
+time-series exhibits a ``$1/f$'' shape below some knee frequency,
+$f_\mathrm{knee}$, and is relatively flat at higher frequencies. Other
+smaller effects may include line features in the spectrum (a common
+culprit is pickup from AC mains, or microphonic signals induced by
+mechanical vibrations), and finally a roll-off towards the Nyquist
+frequency caused by the anti-aliasing filter that precedes
+digitization.
 
-\subsubsection{Scan strategy}
+In order to observe astronomical sources, it is clearly advantageous
+that the signals appear in the region of the spectrum that offers the
+highest \snr, namely frequencies above $f_\mathrm{knee}$. This effect
+is achieved by modulating the signal; the original SCUBA primarily
+used the JCMT chopping secondary to place sources into the spectrum at
+about $\sim$10\,Hz. The problem with this strategy was that SCUBA was
+ultimately limited to recovering information on angular scales smaller
+than the chop throw (a maximum of a couple of arcmin)... cite
+somebody. The primary approach to map-making with SCUBA-2 is instead
+to slew the entire telescope at such a rate that angular scales of
+interest are crossed on time scales shorter than $1/f_\mathrm{knee}$,
+similar to other ground-based instruments (MAMBO, BOLOCAM, SHARC-II,
+AzTEC, LABOCA,...). The desire to scan rapidly has two basic
+constraints: (i) the telescope cannot physically move faster that
+several hundred arcsec\,sec$^{-1}$ without requiring unacceptably
+large accelerations to change direction, and (ii) bolometers are
+designed with a thermal time constant that balance response with
+sensitivity. Taking these constraints into consideration, SCUBA-2 was
+ultimately designed to work at a maximum scan velocity of
+600\,arcsec\,sec$^{-1}$. The data acquisition system has a target
+sample rate of 200\,Hz, which at that scan velocity results in a
+sample every 3.0\,arcsec, or approximately one third of the
+450\,\micron\ diffraction-limited full-width at half-maximum (FWHM) --
+a typical rule-of-thumb for adequately sampling a Gaussian point
+spread function (PSF). Similarly, the SCUBA-2 bolometers are designed
+to have a time constant of 2 ms \textbf{check that number!} in order
+to avoid distorting PSF-sized sources at this maximum scan rate.
+
+The scan strategy is designed, primarily, to mitigate the effects of
+$1/f$ drift. The simplest scan one might use is a raster pattern, or a
+series of parallel passes across this field. However, such a scan
+yields extremely poor performace for structures that are transverse to
+the scan direction, since a long time will have ellapsed between each
+parallel step, and the signal will therefore be totally dominated by
+the $1/f$ noise. In addition, other noise features may appear in the
+data periodically. With such a regularly spaced periodic scan
+strategy, these noise features can easily conspire with the scan
+pattern to produce similarly periodic noise structures in the map, or
+\emph{scan-synchronous} noise. It has therefore been recognized for
+some time that scan strategies that visit each region of the map on as
+many different time scales, and scan angles as possible offer the best
+protection against large-scale noise (...cite some people...). As a
+tradeoff between truly random scan patterns and feasibility of
+implementation taking into consideration both mechanical and
+scheduling constraints, SCUBA-2 presently has three primary scan
+modes: PONG (cite Borys et al.), Lissajous, and Daisy... (more
+citations? Kovacs et al. has a nice description... also CMB
+literature).
 
@@
 
-This paper will be focussed on S2SRO: only one subarray at each
-wavelength; fridge oscillations; working detectors (and variation from
-setup-to-setup).
+The S2SRO period, as mentioned above, took place during February and
+March 2010. For the purpose of this paper, we will consider data taken
+as early as 2009 December 3, at which point the instrument was being
+commissioned in a similar configuration to that of the S2SRO
+period. At this time, each of the 450 and 850\,\micron\ focal planes
+were populated with single subarrays, s4a and s8d respectively. There
+were two main mechanical problems to contend with during this
+period. First, the fridge was not able to cool the focal plane to the
+desired base temperature of ... which affected stability, particularly
+at 850\,\micron. Second, there were problems with the subarray
+fabrication process which led to large numbers of dead pixels, as well
+as large-scale variations in $T_\mathrm{C}$ and the thermal
+conductivity, $G$. These latter issues made it difficult to
+simulataneously lock all of the bolometers, and in addition many of
+the bolometers that were locked did not achieve their minimum
+theoretical noise performance, since it is not possible to adjust the
+individual temperature biases and place each device into the optimal
+region of the superconducting transition.
 
-Show some time series.
+While there were moderate variations from night-to-night during the
+observing period, there were typically 600 to 900 working detectors in
+each of the s4a and s8d subarrays. Fig.~\ref{fig:sensitivities} shows
+representative responsivity maps for the two subarrays from the middle
+of S2SRO based on the response to ramping the bolometer heaters.
 
-Show power spectra.
+\begin{figure}
+\vspace{2cm}
+\textbf{This will be a figure showing s4a and s8d responsivity maps}
+\vspace{2cm}
+\caption{The response of each bolometer to known input heater power
+  for a typical night during S2SRO. The s4a subarray had a
+  particularly large variation in $T_\mathrm{C}$ towards the
+  bottom-right which resulted in lower responsivity, and
+  correspondingly larger noise. The s8d array also suffered
+  significant variations across the focal plane, in particular
+  removing larger number of useful detectors from the bottom-left.}
+\label{fig:sensitivities}
+\end{figure}
 
-Show cross-spectra to emphasize how correlated things are?
+\subsubsection{Time series}
+
+In Fig.~\ref{fig:bolos_mix} we show sample time series from single
+bolometers in each of the subarrays, as well as variations in the
+mixing chamber temperature (as a proxy for the focal-plane
+temperature), and the telescope pointing during a PONG scan.
 
@@
 \end{figure}
+
+Both bolometers share significant long-timescale structure
+($\gsim10$\,s) that appears to be related to variations in the fridge
+base temperature, although the similarity is clearly greater at
+850\,\micron. In this particular case, the total power in the
+fluctuations at 450\,\micron, from $-0.04$\,pW to $+0.08$, are larger
+than the $-0.03$\,pW to $+0.05$\,pW fluctuations at 850\,\micron. Such
+behaviour might be expected if there is a comparable varying thermal
+load from the fridge at each wavelength, but a larger contribution of
+astmospheric variations through the 450\,\micron\ bandpass filters. We
+also note that there is no obvious strong correlation between the
+low-frequency signal structure at either wavelength at the telescope
+motion.
+
+The low-frequency signal component of the bolometer signals is also
+highly correlated amongst bolometers in the same subarray. We have
+calculate a common-mode signal, $\mathbf{c}(t)$ as the straight
+average time series of all the working bolometers. We then fit the
+amplitude of $\mathbf{c}(t)$ at each wavelength to the time series
+shown in Fig.~\ref{fig:bolos_mix} and remove it, yielding the grey
+residual signals. These residuals are quite flat, although still with
+noticeable variations. The white noise is also apparent, and larger at
+450\,\micron\ as one would expect from the larger loading compared to
+850\,\micron.
+
+\subsubsection{Power spectral densities}
+
+Next, we produce power spectral density (PSD) plots for several
+representative bolometers in Fig.~\ref{fig:pspec}. To produce this
+figure, we follow the convention that the PSD as a function of
+frequency, $\mathbf{P}(f)$, is normalized such that the integral over
+frequency gives the same variance as the time-series variance across
+the full time series. In other words, given a bolometer signal
+$\mathbf{b}(t)$,
+%
+\begin{equation}
+\int \mathbf{P}(f) df = \langle\mathbf{b}^2(t)\rangle.
+\end{equation}
+%
+However, we only show the PSD up to the Nyquist frequency, so the
+conversion from variance measured in the frequency domain to the time
+domain requires and additional multiplcation by 2; hence the units are
+pW$^2$\,Hz$^{-1}$ rather than pW$^2$\,s.
 
@@
 
-\subsubsection{Fridge oscillations}
-As shown in Fig.~\ref{fig:bolos_mix}, the dominant signal in both
-bands is clearly an oscillation in the fridge base temperature as
-evidenced by the low-resolution mixing chamber thermometer. The total
-power delivered to the bolometers is comparable at each wavelength,
-further suggesting a source internal to the cryostat, rather than an
-external optical load. Finally, there is no obvious correspondance
-with either the azimuth or elevation motion.
+The dotted coloured lines show the PSDs for raw, though flatfielded
+data. At each wavelength, there is a clear $1/f$ knee at a few Hz,
+followed by a predominantly white spectrum punctuated by line features
+above $\gsim 30$\,Hz, and finally roll-off caused by the anti-aliasing
+filter above $\gsim 70$\,Hz. As indicated in the previous section, the
+correlation between the low-frequency components of the different
+bolometer signals is large. The solid black lines in
+Fig.~\ref{fig:pspec} indicate the PSDs of the common-modes
+$\mathbf{c}(t)$ at each wavelength, which reproduce most of the
+low-frequency structures, as well as the high-frequency line
+features. The $\mathbf{c}(t)$ otherwise drop substantially below the
+individual bolometer PSDs at high-frequency as expected if the
+bolometers are dominated by un-correlated white noise. Finally, we
+note that the amplitudes and slopes of the $\mathbf{c}(t)$ are
+similar at both 450 and 850\,\micron. However, since the
+450\,\micron\ white noise is larger than at 850\,\micron\ (shown
+approximately by the horizontal dotted lines), the $1/f$ knee occurs
+at a \emph{lower} frequency at 450\,\micron.
 
-Show what's left over when we remove the common-mode, and the
-correlation of the resulting signal with patterns in $G$ and
-$T_\mathrm{C}$ across the array.
+Next, the common-mode signals are fit to each bolometer time series
+and removed as in Fig.~\ref{fig:bolos_mix}, and the resulting PSDs are
+shown as solid coloured lines. In this example, the residual signals
+have $1/f$ knees significantly lower than in the raw PSDs.
+
+For reference, the top horizontal axis has been converted to angular
+scale assuming a scan velocity of 200\,arcsec\,sec$^{-1}$ which was
+typical of the S2SRO observations. The power spectra of the PSFs in
+each band (arbitrary normalized to peak values of $10^{-6}$ and
+$10^{-7}$\,pW$^2$\,Hz$^{-1}$ at 450 and 850\,\micron\ respectively)
+are also showed as dashed lines for this assumed scan velocity showing
+that the smallest features resolvable by the telescope may be slightly
+affected by the excess noise in the line features. At the lower
+frequency end, the $1/f$ noise dominates at scales $\gsim 2$\,arcmin
+in the raw data and $\gsim 10$\,arcmin in the common-mode cleaned data
+at 450\,\micron, and at scales $\gsim 1$\,arscmin in the raw data and
+$\gsim 2$\,arcmin in the common-mode cleaned data at 850\,\micron.
+
+\subsubsection{Principal component analysis}
+
+\begin{figure*}
+\centering
+\includegraphics[width=\linewidth]{pca.pdf}
+\caption{The first six modes from a principal component analysis of
+  the combined 450 and 850\,\micron\ time-series bolometer data.}
+\label{fig:pca}
+\end{figure*}
 
@@
   PCA-cleaning procedure.}
-
 \label{fig:pca}

### 2010-12-03 smurf_paper.tex
add definition of \prd for Stompor et al.
describe PCA analysis

@@
 \newcommand{\pasp}{\rm Publ. Astron. Soc. Pac.}
-\newcommand{\ASP}{\rm ASP COnference Series}
+\newcommand{\prd}{\rm PhRvD}
+\newcommand{\ASP}{\rm ASP Conference Series}
 \newcommand{\CASP}{\rm Comm. Astrophys. Space Phys.}
@@
 many different time scales, and scan angles as possible offer the best
-protection against large-scale noise (...cite some people...). As a
-tradeoff between truly random scan patterns and feasibility of
-implementation taking into consideration both mechanical and
-scheduling constraints, SCUBA-2 presently has three primary scan
-modes: PONG (cite Borys et al.), Lissajous, and Daisy... (more
-citations? Kovacs et al. has a nice description... also CMB
-literature).
+protection against large-scale noise (...cite some people...); such
+strategies are said to offer good \emph{cross-linking}. As a tradeoff
+between truly random scan patterns and feasibility of implementation
+taking into consideration both mechanical and scheduling constraints,
+SCUBA-2 presently has three primary scan modes: PONG (cite Borys et
+al.), Lissajous, and Daisy... (more citations? Kovacs et al. has a
+nice description... also CMB literature).
 
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=\linewidth]{pca.pdf}
-\caption{The first six modes from a principal component analysis of
-  the combined 450 and 850\,\micron\ time-series bolometer data.}
-\label{fig:pca}
-\end{figure*}
+A method that has been employed by teams analyzing BOLOCAM and AzTEC
+data to identify and remove low-frequency correlated noise is
+Principal Component Analysis
+\citep[PCA,][]{laurent2005,scott2008}. While we have not (yet)
+included ``PCA cleaning'' as a standard feature of SMURF, we have
+performed a preliminary analysis in IDL of some SCUBA-2 data to gain
+further insight into the nature of the low-frequency noise. The basic
+method is as follows: (i) the covariance matrix is built up for all
+pairs $(i,j)$ of the $N$ bolometer time series,
+$\langle\mathbf{b}_i(t),\mathbf{b}_j\rangle$; and (ii) an eigenvalue
+decomposition is used to identify a new set of statistically
+uncorrelated eigenvectors, $\mathbf{\xi}_i(t)$ (i.e. whose covariance
+matrix is diagonal), such that each of the bolometer time series is a
+linear combination of the eigenvectors, or \emph{components}
+i.e. $\mathbf{b}_i(t) = \bar{\mathbf{\xi}} \mathbf{\lambda}_i^T$,
+where each row of the matrix $\bar{\mathbf{\xi}}$ is an eigenvector,
+and $\mathbf{\lambda}_i^T$ is a column vector containing the
+corresponding eigenvalues. For convenience, the $\mathbf{\xi}_i$ are
+normalized by their \rms, so that the amplitude of each component is
+stored in the eigenvalues. In the earlier analyses mentioned, the
+low-frequency noise is assumed to be encapsulated in those components
+with the largest eigenvalues. Removing the projection of the time
+series along those components has been used successfully to reduce
+$1/f$ noise while retaining most of the signal in point-like sources.
 
-\subsubsection{Sky noise}
-Do we even see sky noise? Can show some results for the combined 450
-and 850\,\micron\ joint PCA analysis.
-
-\subsubsection{Magnetic field pickup}
-Show that it occasionally appears in the dark squids.
-
-\subsubsection{Line features}
-
-The power spectra also show some narrower line features (many
-correlated across the array). Even though we largely ignore them worth
-some mention.
+One unique feature of the SCUBA-2 data is that one can perform a PCA
+analysis of both the 450 and 850\,\micron\ data simultaneously. The
+first 6 most significant modes of our analysis are shown in
+Fig.~\ref{fig:pca}, with the eigenvectors in the top panels, and maps
+of the eigenvalues in each subarray in the bottom panels.
 
@@
 \end{figure*}
+
+The majority of correlated signal at both wavelengths is clearly
+broken down into the first two plotted components. The first dominates
+at 450\,micron, whereas the second dominates at 850\,micron. For
+reference, the eigenvalues of these two modes are in the range
+$\sim1$--4\,pW, whereas the remaining modes have eigenvalues
+$\lsim0.3$\,pW. As mentioned earlier, one simple model for the
+correlated signal is a simple combination of oscillations in base
+temperature (which would be similar in each band), with a varying
+atmospheric signal (which is expected to be stronger at
+450\,\micron). Indeed, the fainter component 2 most closely resembles
+the mixing chamber temperature variations (Fig.~\ref{fig:bolos_mix}),
+whereas component 1 is stronger at 450\,\micron\ and does not resemble
+the fridge signal as closely.
+
+This physical interpretation of the eigenvectors should, however, be
+taken with a grain of salt, since the PCA analysis is really a
+statistical black box. Suppose most of the low frequency noise really
+could be decomposed into two components, the fridge signal
+$\mathbf{F}(t)$, and an atmospheric signal $\mathbf{A}(t)$. The two
+eigenvectors obtained from the PCA analysis are simply linear
+combinations of $\mathbf{F}(t)$ and $\mathbf{A}(t)$ that are
+statistically uncorrelated. Depending on the particular realizations
+of these two noise sources, this decomposition may, or may not give
+rise to eigenvectors that resemble the two underlying physical
+sources. Regardless, similar analyses of a few data sets during S2SRO
+give comparable results; most of the signal at both 450\,\micron\ and
+850\,\micron\ appears to be produced by two dominant underlying
+mechanisms.
+
+\subsubsection{Magnetic field pickup}
+Show that it occasionally appears in the dark squids.
 

### 2010-12-03 smurf_paper.tex
description for later PCA modes, and intro to SMURF algorithm section

@@
 
+The subsequent components identified by the PCA analysis in
+Fig.~\ref{fig:pca} vary wildly. Component 3, though not obvious from
+its eigenvector plot, is almost a pure ramping waveform that only
+appears in the s4a data, and produces the broad line featuress
+centered over 45\,Hz and 70\,Hz in the 450\,\micron\ PSDs in
+Fig.~\ref{fig:psd}. In other data sets, this third component has also
+often instead exhibited a regularly-spaced series of spikes with
+heights quantized into several amplitude families (again, only in the
+s4a array). In both cases, the shape and phase of this signal is the
+same in each bolometer, although the amplitude (and sign) varies, as
+evidenced by the large random scatter in the s4a eigenvalue maps. It
+is thought that these features are some sort of data acquisition
+glitch as the spikes (when they appear) exhibit the expected point
+response function of the anti-aliasing filter. Similarly, component 4
+also appears to be a data acquisition glitch in the s4a array,
+although along a single column, and with a peak at a lower frequency
+of 40\,Hz, and a second feature right at the Nyquist frequency.
+
+The last two components in Fig.~\ref{fig:pca} have an entirely
+different character. The amplitudes of both components are greatest in
+low-responsivity regions of the two subarrays (lower-right at
+450\,\micron\ and bottom-left at 850\,\micron). Bolometers in these
+regions of the subarray are probably only tenuously locked, and are
+more sensitive to changes in the loading, possibly resulting in the
+step and spikes that are seen in component 5. Component 6 is clearly
+related to component 5, though with the absence of the steps and
+spikes, so that together they seem to represent a continuum of shapes
+that are linear combinations of the spikes and steps with the slower
+drift evident only in component 6.
+
 \subsubsection{Magnetic field pickup}
@@
 %------------------------------------------------------------------------------
+
+Given predictions for the performance of a fully-functioning SCUBA-2,
+and the observed properties of the first-installed subarrays during
+commissioning in late 2009 as described in the previous section, SMURF
+was designed with two primary goals in mind: (i) reduced maps should
+be scientifically useful and of ``near publication quality'' with
+minimal user interaction; and (ii) the execution time should not
+appreciably exceed the amount of time required to conduct
+observations. By attaining these goals it would be possible to conduct
+real-time reduction at the telescope to offer feedback to observers,
+and it would also be possible to maintain a central repository of
+(nearly) publishable data products.
+
+There are two general families of existing map-making strategies that
+were considered. The first are \emph{maximum likelihood} methods that
+were originally pioneered for the analysis of Cosmic Microwave
+Background (CMB) maps. Here, the pixels in the map are considered to
+be free parameters in a model-fitting problem, and the data are the
+raw time-series. By estimating the full noise covariance matrix for
+the data, it is then possible to estimate the maximum-likelihood
+map. While direct inversion solutions to this problem are untenable,
+some recent efforts to optimize these algorithms have enabled the
+reduction of smaller data sets in a reasonable amount of time. While
+maximum likelihood maps are, theoretically, the best maps that can be
+made from a noise point of view, the required computing resources and
+execution time would not have been feasible for the proposed real-time
+data reduction and data archiving model.
 

### 2011-03-09 smurf_paper.tex
overview of map-making and draft of FLT and AST

@@
 \subsubsection{Principal component analysis}
+\label{sec:pca}
 
@@
 some recent efforts to optimize these algorithms have enabled the
-reduction of smaller data sets in a reasonable amount of time. While
-maximum likelihood maps are, theoretically, the best maps that can be
-made from a noise point of view, the required computing resources and
-execution time would not have been feasible for the proposed real-time
-data reduction and data archiving model.
+reduction of smaller data sets in a reasonable amount of time
+\citep[e.g.][]{patanchon2008}. While maximum likelihood maps are,
+theoretically, the best maps that can be made from a noise point of
+view, the required computing resources and execution time would not
+have been feasible for the proposed real-time data reduction and data
+archiving model.
 
@@
 $\mathbf{n}^\mathrm{c}(t)$ is a (predominantly low-frequency)
-correlated or \emph{common-mode} component.
-
-The primary goal of map-making is to model and remove
-$\mathbf{n}^\mathrm{f}_i$ and $\mathbf{n}^\mathrm{c}$ from the
-time-series. By re-gridding the remaining data we can then hope to
-approach the theoretical noise limit in the map.
+correlated or \emph{common-mode} component. The primary goal of
+map-making is to model and remove $\mathbf{n}^\mathrm{f}_i$ and
+$\mathbf{n}^\mathrm{c}$ from the time-series. By re-gridding the
+remaining data we can then hope to approach the theoretical noise
+limit in the map. This procedure is summarized in Fig.~\ref{fig:dimm}.
 
@@
 
+First, the data files are read from disk, concatenated into contiguous
+arrays in memory, and run through a pre-processing stage that performs
+several operations including spike and sudden level step removal to
+enforce continuity in the time-series. Then, the iterative process
+within the dashed box begins. It is generally assumed that the
+dominant noise component is the common-mode, which is presumably
+dominated by large variations in the fridge temperature and sky noise
+(as argued in Section~\ref{sec:pca}). This signal is estimated as the
+average signal observed by all of the bolometers at each time step,
+and then removed from all of the bolometers. Once this large signal
+has been removed, we apply the extinction correction, which is
+determined completely from external measurements. Next, residual noise
+that differse from bolometer-to-bolometer (primarily low-frequency) is
+removed using an FFT-based filter. The resulting ``clean'' time-series
+are then re-gridded to produce a map estimate using nearest-neighbour
+sampling. Since each map pixel averages many bolometer samples
+together, it is fundamentally much less noisy than the bolometer
+time-series. We can therefore project the map back into the time
+domain, and create relatively noiseless estimates of what each
+bolometer would have seen in the absence of noise sources, and remove
+that from the time-series. This final step in the iteration gives a
+residual signal (with low-frequency and astronomical noise components
+removed) that may be used to measured the white noise level in each
+detector, and also to estimate an approximate $\chi^2$ for the model
+to track convergence. As we will discuss in the next sections, each
+signal component estimate is biased slightly from the presence of
+signal from other components, so the entire process is
+iterated. However, before the next iteration begins, we first undo the
+multiplicative factors introduced by the extinction correction, and
+also relative gains that may optionally be estimated in the
+common-mode estimation phase.
+
+In practice, the map-maker is highly configurable, both in terms of
+the order in which model components are fit (and whether they are fit
+at all), and the details of \emph{how} they are fit. In the following
+sections we describe these details.
+
 %-------------------------------------------------
@@
 
-Give details about each model component, roughly in the order they are
-calculated.
+In addition to specifying when each model component is calculated
+(normally an attempt is made to order them by decreasing amplitude in
+the time-series), there are a number of parameters that control their
+operation.
 
-\subsection{Common-mode estimation}
+\subsection{\texttt{COM,GAI}: common-mode estimation}
 
@@
 flag bad bits of data, check for time-varying bolometer
-responsivities... what doesn't it do?!
+responsivities... what doesn't it do?! Also mention the optional
+presence of GAI.
 
-\subsection{Iterative filtering}
+\subsection{\texttt{FLT}: fourier transform filter}
 
-EC
+This model takes the FFT of the bolometer data, and can apply both
+high- and low-pass filters, as well as notch filteres, at hard
+frequency edges specified by the user. Alternatively, the frequency
+edges of the filters may be defined in terms of an \emph{angular
+  scale}, but converted into a frequency through knowledge of the mean
+telescope slew speed. The time-series may optionally have apodization
+applied before the transform to avoid ringing (primarily caused by
+wrap-around discontinuities at the ends of the time-series). The
+signal that is \emph{removed} from the time-series by this process are
+stored in the \texttt{FLT} container array.
 
-The FFT filter can do high/low-pass filtering. We can also do notches
-and whitening.
+\subsection{\texttt{SMO}: time-domain smoothing filter}
 
@@
 
-\subsection{Noise estimation}
+\subsection{\texttt{AST}: map estimation}
 
-Measure a white noise level from the power spectrum for each detector
-to use as a weight.
+Map estimation is accomplished using a simple nearest-neighbour
+resampling of the data onto a pre-defined map grid. For the $i$th map
+pixel $\mathbf{m}(x_i,y_i)$, the brightness is estimated as the
+weighted average of the bolometer data samples $\mathbf{b}_j$ that
+land within that pixel (from any bolometer or point in time),
+%
+\begin{equation}
+  \mathbf{m}(x_i,y_i) = \frac{\sum_j \mathbf{w}_j \mathbf{b}_j }
+                             { \sum_j \mathbf{w}_j } .
+\end{equation}
+%
+For the initial iteration the weights $\mathbf{w}_j$ are set to 1, but
+subsequently they are set to $1/\sigma_j^2$, the estimated inverse
+variance expected from the bolometer white noise levels as discussed
+in Section~\ref{sec:noi}. This weighting scheme is sensible provided
+that the bolometer data have no correlated (i.e., low-frequency)
+noise.
 
-\subsection{Map estimation}
+In addition, a variance map $\mathbf{v}(x_j,y_j)$ is estimated. The
+default behaviour is to estimate this weighted error on the mean given
+the scatter in the weighted samples. This is accomplished by dividing
+the \emph{biased weighted sample variance} by the number of samples
+that went into the average (akin to the formula for standard error on
+the mean, but accounting for weights),
+%
+\begin{equation}
+\mathbf{v}(x_j,y_j) = \frac{\sum_j \mathbf{w}_j
+                            \sum_j \mathbf{w}_j \mathbf{b}_j -
+                            \left( \sum_j \mathbf{w}_j \mathbf{b}_j \right)^2 }
+                           { N \left( \sum_j \mathbf{w}_j \right)^2 },
+\end{equation}
+%
+where $N_j$ is the total number of bolometer samples that land in the
+pixel. We decided not to use the \emph{unbiased} estimator since in
+practice it would require accumulating an additional array of values
+at every map pixel, and only results in a small difference where there
+are less than $\sim10$ samples per pixel (a situation that is almost
+never encountered in a SCUBA-2 map except in the edge pixels).
 
-In the map, use weighted standard error to estimate noise. This is
-accurate for a single pixel, but there is still spatially correlated
-noise in the maps which we can show with some figures.
+Finally, once the map estimation is complete, the map is projected
+into the time domain (the signal that would be produced in each
+bolometer by the signal represented by the map) and removed.
+
+\subsubsection{Map-based despiking}
+
+In addition, the \texttt{AST} model can use the map and its variance
+to provide additional despiking of the time-series. Unlike the
+time-domain despiker described ???, this calculation uses the scatter
+in the population of samples used to estimate the map pixel value
+(from different times and bolometers) to reject outliers. This method
+is more robust against false-positive detections of bright/compact
+astronomical sources since transient features in the time-series are
+unlikely to occur by chance whenever a bolometer crosses a specific
+location on the sky, whereas real astronomical sources are at fixed
+spatial locations.
+
+The estimated variance, $\mathbf{v_p}(x_i,y_i)$, of the normalized
+weighted samples that land in the $i$th map pixel is simply the biased
+weighted sample variance (i.e., the variance map value multiplied by
+the number of samples)
+%
+\begin{equation}
+  \mathbf{v_p}(x_i,y_i) = N_i \mathbf{v}(x_i,y_i).
+\end{equation}
+
+In order to compare the weighted \emph{differences} between the
+samples and the map values, [$\mathbf{b}_j - \mathbf{m}(x_i,y_i)$] to
+$\mathbf{v_p}$, they must be scaled appropriately. We define a
+normalized difference, $\mathbf{d}_j$, in such a way that the variance
+of this new variable gives the weighted sample variance of the
+underlying data points,
+%
+\begin{eqnarray}
+  \frac{\sum_j \mathbf{d}_j^2}{N} &=&
+  \frac{\sum_j \mathbf{w}_j [\mathbf{b}_j - \mathbf{m}(x_i,y_i)]^2}
+       {\sum_k \mathbf{w}_k} \\
+   \Rightarrow \mathbf{d}_j^2 &=& \frac{N \mathbf{w}_j [\mathbf{b}_j -
+       \mathbf{m}(x_i,y_i)]^2}{ \sum_k \mathbf{w}_k} .
+\end{eqnarray}
+%
+The map-based despiker flags those $\mathbf{d}_j$ that are further
+than some threshold number of standard deviations,
+$\sqrt{\mathbf{v_p}}$, away from zero so that they are not used in
+subsequent iterations.
+
+\subsection{\texttt{NOI}: noise estimation}
+\label{sec:noi}
+
+The primary purpose of the noise component, \texttt{NOI} is to measure
+the noise in each bolometer residual once all of the other models have
+been fit and removed. This noise may then be used to estimate weights
+for the data in subsequent iterations.
 
@@
 can go here, or perhaps earlier?
+
+\subsubsection{Down-sampling}
+
+EC
+
+\subsubsection{Whitening filter}
 

### 2011-03-10 smurf_paper.tex
add description of NOI, and re-organize model component section

@@
 \subsubsection{Power spectral densities}
+\label{sec:psd}
 
@@
 \begin{equation}
-\int \mathbf{P}(f) df = \langle\mathbf{b}^2(t)\rangle.
+\label{eq:psd}
+\langle\mathbf{b}^2(t)\rangle = \int \mathbf{P}(f) df .
 \end{equation}
@@
 centered over 45\,Hz and 70\,Hz in the 450\,\micron\ PSDs in
-Fig.~\ref{fig:psd}. In other data sets, this third component has also
-often instead exhibited a regularly-spaced series of spikes with
+Fig.~\ref{fig:pspec}. In other data sets, this third component has
+also often instead exhibited a regularly-spaced series of spikes with
 heights quantized into several amplitude families (again, only in the
@@
 the order in which model components are fit (and whether they are fit
-at all), and the details of \emph{how} they are fit. In the following
-sections we describe these details.
+at all), and the details of \emph{how} they are fit. A summary of the
+model components is given in the next section.
 
@@
 \subsection{Model components}
+\label{sec:components}
 %-------------------------------------------------
@@
 the time-series), there are a number of parameters that control their
-operation.
-
-\subsection{\texttt{COM,GAI}: common-mode estimation}
+operation. Table~\ref{tab:components} list the model components and
+the typical order in which they are calculated. The following sections
+describe their implementation in detail.
+
+\begin{table}
+  \caption{Summary of the model components that can be fit to
+    \scuba\ time-series data with SMURF. Only the first group of
+    models are typically fit to the data (\texttt{COM}--\texttt{NOI})
+    in the indicated order. The remaining models
+    (\texttt{DKS}--\texttt{PLN}) have only been included for
+    completeness}
+  \vspace{0.2cm}
+  \centering
+  \begin{tabular}{c|l}
+    \hline
+    Model & Description \\
+    \hline
+    \texttt{COM} & remove common-mode signal \\
+    \texttt{GAI} & common-mode scaled to each bolometer \\
+    \texttt{EXT} & extinction correction \\
+    \texttt{FLT} & Fourier transform filter \\
+    \texttt{AST} & map estimate of astronomical signal \\
+    \texttt{NOI} & noise estimation \\
+    \hline
+    \texttt{DKS} & dark squid cleaning along columns \\
+    \texttt{SMO} & time-domain smoothing filter \\
+    \texttt{PLN} & 2-dimensional time-varying plane removal \\
+    \hline
+    \end{tabular}
+  \label{tab:components}
+\end{table}
+
+\subsubsection{\texttt{COM,GAI}: common-mode estimation}
+\label{sec:comgai}
 
@@
 
-\subsection{\texttt{FLT}: fourier transform filter}
+\subsubsection{\texttt{EXT}: extinction correction}
+\label{sec:ext}
+
+AGG
+
+This is actually only calculated once before the map solution begins
+using external opacity monitors (WVM, $\tau_\mathrm{CSO}$,
+skydip). Summarize the measurements, order of preference and something
+about the calibration for the SCUBA-2 filters.
+
+\subsubsection{\texttt{FLT}: fourier transform filter}
+\label{sec:flt}
 
@@
 
-\subsection{\texttt{SMO}: time-domain smoothing filter}
-
-TJ
-
-We also have the SMO filter that's useful for very short scans
-because edge effects don't matter so much. Two sorts - median which is
-more robust, or mean which is more accurate.
-
-\subsection{\texttt{AST}: map estimation}
+\subsubsection{\texttt{AST}: map estimation}
+\label{sec:ast}
 
@@
 
+In addition to map estimation, the \texttt{AST} model can also be used
+to perform map-based despiking of the time-series
+(Section~\ref{sec:mapdespike}), and to apply constraints to the map to
+improve convergences (Section~???).
+
+\subsubsection{\texttt{NOI}: noise estimation}
+\label{sec:noi}
+
+The primary purpose of the noise component, \texttt{NOI} is to measure
+the noise in each bolometer residual once all of the other models have
+been fit and removed. This noise may then be used to estimate weights
+for the data in subsequent iterations.
+
+As in Section~\ref{sec:psd} the FFT of each bolometer time series is
+taken, and the PSD calculated. An average white noise level is then
+measured from 2 to 10\,Hz, a relatively clean region that lies above
+the $1/f$ knee, but below the high-frequency line features for most of
+the bolometer data taken during S2SRO (Fig.~\ref{fig:pspec}). Taking
+this constant level for $\mathbf{P}(f)$ we then calculate the expected
+variance of the time-series (in approximately 200\,Hz samples) using
+Eq.~\ref{eq:psd}. If the bolometer noise were produced purely by
+uncorrelated thermal sources (i.e. no other long-scale drifts), with
+no high-frequency line features, and without the attenuation at even
+higher frequencies by the anti-aliasing filter, this is the
+theoretical noise limit of the detectors. These measured variances are
+stored, and then used in subsequent iterations to weight the data
+points when calculating the map estimate (Section~\ref{sec:ast}).
+
+Since noise estimation is usually calculated as the final step, it is
+the cleanest data for which some cleaning operations may be
+performed. For this reason \texttt{NOI} can optionally use the
+time-domain DC step fixer (Section~\ref{sec:steps}) and spike
+detection (Section~\ref{sec:timedespike}).
+
+\subsubsection{\texttt{DKS}: dark squid cleaning along columns}
+\label{sec:dks}
+
+Columns share significant magnetic field pickup which is measured by
+the dark squids. This model has been included although it makes little
+improvement, or injects noise since the magnetic field pickup noise is
+generally sub-dominant.
+
+\subsubsection{\texttt{SMO}: time-domain smoothing filter}
+\label{sec:smo}
+
+TJ
+
+We also have the SMO filter that's useful for very short scans
+because edge effects don't matter so much. Two sorts - median which is
+more robust, or mean which is more accurate.
+
+\subsubsection{\texttt{PLN}: 2d time-varying sky noise plane removal}
+\label{sec:pln}
+
+TJ
+
+We can remove a plane at each timeslice but it doesn't seem to improve
+things. Perhaps refer to other papers that claim to have seen coherent
+2d structures in the sky noise?
+
+\subsection{Model degeneracies}
+
+Introduce the topic of model degeneracies here, but defer their
+solution to the example sections.
+
+%-------------------------------------------------
+\subsection{Data cleaning}
+%-------------------------------------------------
+
+Since we can do most of this stuff either as a pre-processing step or
+iteratively, I thinks this goes in a generic-sounding subsection. This
+can go here, or perhaps earlier?
+
+\subsubsection{Down-sampling}
+
+EC
+
+\subsubsection{Whitening filter}
+
+EC
+
+\subsubsection{Time-domain de-spiking}
+\label{sec:timedespike}
+
+DSB
+
+\subsubsection{Step correction}
+\label{sec:steps}
+
+DSB
+
+\subsubsection{Gap filling / apodization}
+\label{sec:gaps}
+
+DSB
+
+Discuss gap filling, ``new'' window function method, also apodization
+and start/end polynomial continuity stuff. Should reference CMB
+literature -- in particular look at \citet{stompor2002}.
+
+\subsubsection{Flatfields}
+\label{sec:flatfields}
+
+TJ
+
+Discuss heater ramps, normalizing off the common-mode signal
+amplitude, astronomical sources, and the internal flatfield source.
+
+\subsubsection{Additional bad bolometer rejection}
+
+TJ
+
+Measuring the noise power spectra and throwing away outliers. Data
+flagged bad by the DA system.
+
 \subsubsection{Map-based despiking}
-
-In addition, the \texttt{AST} model can use the map and its variance
-to provide additional despiking of the time-series. Unlike the
-time-domain despiker described ???, this calculation uses the scatter
-in the population of samples used to estimate the map pixel value
-(from different times and bolometers) to reject outliers. This method
-is more robust against false-positive detections of bright/compact
-astronomical sources since transient features in the time-series are
-unlikely to occur by chance whenever a bolometer crosses a specific
-location on the sky, whereas real astronomical sources are at fixed
-spatial locations.
+\label{sec:mapdespike}
+
+It is possible to use the map estimate (Section~\ref{sec:ast}) to
+provide additional despiking of the time-series. Unlike the
+time-domain despiker (Section~\ref{sec:timedespike}), this calculation
+uses the scatter in the population of samples used to estimate the map
+pixel values (from different times and bolometers) to reject
+outliers. This method is more robust against false-positive detections
+of bright/compact astronomical sources since transient features in the
+time-series are unlikely to occur by chance whenever a bolometer
+crosses a specific location on the sky, whereas real astronomical
+sources are at fixed spatial locations.
 
@@
 
-\subsection{\texttt{NOI}: noise estimation}
-\label{sec:noi}
-
-The primary purpose of the noise component, \texttt{NOI} is to measure
-the noise in each bolometer residual once all of the other models have
-been fit and removed. This noise may then be used to estimate weights
-for the data in subsequent iterations.
-
-\subsection{Model degeneracies}
-
-Introduce the topic of model degeneracies here, but defer their
-solution to the example sections.
-
-%-------------------------------------------------
-\subsection{Data cleaning}
-%-------------------------------------------------
-
-Since we can do most of this stuff either as a pre-processing step or
-iteratively, I thinks this goes in a generic-sounding subsection. This
-can go here, or perhaps earlier?
-
-\subsubsection{Down-sampling}
-
-EC
-
-\subsubsection{Whitening filter}
-
-\subsubsection{Time-domain de-spiking}
-
-DSB
-
-\subsubsection{Step correction}
-
-DSB
-
-\subsubsection{Gap filling / apodization}
-
-DSB
-
-Discuss gap filling, ``new'' window function method, also apodization
-and start/end polynomial continuity stuff. Should reference CMB
-literature -- in particular look at \citet{stompor2002}.
-
-\subsubsection{Flatfields}
-\label{sec:flatfields}
-
-TJ
-
-Discuss heater ramps, normalizing off the common-mode signal
-amplitude, astronomical sources, and the internal flatfield source.
-
-\subsubsection{Additional bad bolometer rejection}
-
-TJ
-
-Measuring the noise power spectra and throwing away outliers. Data
-flagged bad by the DA system.
-
-\subsubsection{Map-based despiking}
-
-EC
-
 %------------------------------------------------------------------------------

### 2011-03-11 smurf_paper.tex
add section about magnetic field pickup

@@
 \subsubsection{Magnetic field pickup}
-Show that it occasionally appears in the dark squids.
+\label{sec:magpickup}
+
+An additional noise source that appears to be significant in only a
+small subset of the S2SRO data is magnetic field pickup. Since the
+bolometer signals are ultimately detected through the amplification of
+magnetic fields, any additional changing fields within the cryostat
+will also be detected.
+
+An example of an observation where pickup appears to be significant is
+shown in Fig.~\ref{fig:magpickup}. The time-series for two
+450\,\micron\ bolometers in the same column (not flatfielded) show
+that there is a strong signal with a similar shape, but opposite
+signs. This behaviour is seen across the entire array. The dark squid
+signal for the same columns exhibits a similar shape and
+amplitude. Since the dark squid has no thermal absorber or TES
+attached to it, this observed signal is not likely to be optical or
+thermal in nature (although there can be some crosstalk with the
+bolometers...). It is also for this reason that the higher-frequency
+noise that appears in the bolometer data (thermal) does not appear in
+the dark squid, leading to a significantly higher \snr. The sign of
+such a signal is expected to be random in the bolometer time-series
+(ask detector person for an explanation).
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{magpickup.pdf}
+\caption{Evidence for significant magnetic field pickup for
+  observation 20100228\_00016.  The top panel shows two un-flatfielded
+  (but mean-subtracted) bolometer time-series from the same column,
+  with a 200 sample boxcar smooth (approximately 1\,s), illustrating
+  that they are dominated by a similar signal with opposite signs. The
+  second panel shows the dark squid signal for the column, also
+  mean-subtracted and with the same boxcar smooth. The bottom panel
+  shows the azimuthal and elevation offsets from the map centre. For
+  reference, the mean azimuth was 171.9\,deg, and the mean elevation
+  68.0\,deg. Only the azimuthal signal appears to be correlated with
+  the dark squids and bolometer signals, which suggests a magnetic
+  field stationary with respect to the telescope dome as the source,
+  since its direction with respect to the cryostat only changes with
+  azimuthal motion.}
+\label{fig:magpickup}
+\end{figure}
+
+The telescope pointing offsets for this approximately 0.5\,deg
+diameter scan are also shown in Fig.~\ref{fig:magpickup}. Since the
+phase of the azimuth offsets from the map centre in this scan pattern
+slowly drifts with respect to the elevation offsets, it is clear that
+the bolometer and dark squid signals are detecting a noise source that
+is correlated only with the azimuthal motion and \emph{not} the
+elevation. This behaviour would be expected if if there were a strong
+magnetic field fixed with respect to the telescope dome (i.e., the
+earth's magnetic field). Since SCUBA-2 is mounted on a Nasmyth
+platform, only azimuthal motion will change the effective direction of
+such a field with respect to the cryostat. It is also worth noting
+that the absolute azimuth of the telescope was close to 180\,deg
+(i.e., nearly due-south). We suspect that this observing
+configuration, combined with the large amplitude of the scan, may have
+conspired to produce these large signals.
 

### 2011-10-21 smurf_paper.tex
edits to intro

@@
   \parbox[t]{\textwidth}{
-    Edward~L.~Chapin$^{1}$\thanks{E-mail:~echapin@phas.ubc.ca},
+    Edward~L.~Chapin$^{1,2}$\thanks{E-mail:~echapin@phas.ubc.ca},
     David~S.~Berry$^{2}$,
@@
 The Submillimetre Common User Bolometer Array 2
-\citep[\scuba,][]{holland2006} is a new instrument for the 15-m
-James Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
-can simultaneously image the sky in two broad bands centered over 450
-and 850\,\micron. Ultimately, the instrument will also be equipped
-with a Fourier transform spectrometer \citep[FTS-2,][]{gom2010}, and a polarimeter
-\cite[POL-2,][]{bastien2005}. This paper describes the Submillimetre User Reduction Facility,
-SMURF, a software package for reducing the imaging data, with an
-emphasis on data taken during the SCUBA-2 Shared-Risks Observing period
-(S2SRO) which took place from 2010 February 11 to 2010 March 25. The reduction
-of FTS-2 and POL-2 data will be described at a later date once these
-additional instruments have been commissioned.
-
-Over the last twenty years, the submillimetre band (defined here to be
-200--1000\,\micron) has revolutionized several important areas of
-astrophysics: helping to characterize the early stages of
-star-formation by identifying the dense, cold regions in molecular
-clouds where stars may eventually form; discovering through blind
-surveys the locations and surface density of a class of massive
-star-forming galaxies in the early Universe, now referred to as
-submillimetre galaxies, or SMGs; and finding debris disks around
-nearby stars, the early stages of planet formation.
-
-Styles of map-making: maximum likelihood
-\citep[e.g.,][]{patanchon2008}; iterative -- see \citet{johnstone2000}
-implementation of \citet{wright1996} to remove chop,
-\citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA analysis
+\citep[\scuba,][]{holland2006} is a new instrument for the 15-m James
+Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
+simultaneously images the sky in two broad bands centered over 450 and
+850\,\micron. The focal planes at each wavelength are populated with 4
+rectangular subarrays, consisting of $40 \times 32$ bolometers each,
+and subtend a nearly contiguous $4' \times 4'$ footprint on the sky.
+Ultimately, the instrument will also be equipped with a Fourier
+transform spectrometer \citep[FTS-2,][]{gom2010}, and a polarimeter
+\cite[POL-2,][]{bastien2005}. This paper describes the Submillimetre
+User Reduction Facility, SMURF, a software package for reducing the
+imaging data. The reduction of FTS-2 and POL-2 data will be described
+at a later date once these additional instruments have been
+commissioned. The details of the instrument design, performance, and
+its calibration are given in two companion papers: \citet{holland2011}
+and \citet{dempsey2011}.
+
+Over the last twenty years, observations in the submillimetre band
+(defined here to be 200--1000\,\micron) have helped revolutionized
+several important areas of astrophysics including: the
+characterization of early stages of star-formation by identifying the
+dense, cold regions in molecular clouds where stars may eventually
+form; discovering through blind surveys a class of massive dusty
+star-forming galaxies in the early ($z>2$) Universe, now referred to
+as submillimetre galaxies, or SMGs; and identifying debris disks
+around nearby stars, the early stages of planet formation.  With
+$\sim10000$ total detectors, SCUBA-2 is presently the largest
+submillimetre camera in the world, and the fastest wide-area
+ground-based submillimetre imager.
+
+Most submillimetre imaging cameras to date have used bolometers to
+maximize sensitivities: each pixel consists of an efficient thermal
+absorber to collect photons that have passed through a broad-band
+filter, and the signal is read-out using a thermometer. The
+high-frequency broad-band noise in modern systems, such as SCUBA-2, is
+limited by photon noise from the instrument and ambient
+backgrounds. The low-frequency noise, however, is dominated by sources
+such as the low-frequency drift in the readout electronics, and drifts
+in the mean background levels (such as the atmosphere for ground-based
+instruments). The process of making a map therefore amounts to
+separating the component of the bolometer signals that is produced by
+astronomical sources, from the non-white (and therefore correlated)
+contaminating noise components. Fortunately each has a particular
+profile: astronomical sources are typically fixed in a sky reference
+frame (to a good approximation), whereas other noise sources are often
+correlated in time, and usually from bolometer-to-bolometer. ...also
+mention scan strategies...
+
+There are numerous ways to attack this problem, most of them evolving
+out of techniques that were developed to produce maps of the
+large-scale anisitropies in the Cosmic Microwave Backgroubd.  Styles
+of map-making: maximum likelihood \citep[e.g.,][]{patanchon2008};
+iterative -- see \citet{johnstone2000} implementation of
+\citet{wright1996} to remove chop, \citet{kovacs2008} for SHARC-2
+etc.; de-correlation using PCA analysis
 \citep[e.g.][]{laurent2005,scott2008,aguirre2010}.
 
-Since \scuba\ will ultimately have nearly 10,000 working bolometers,
-considerably larger than any other existing ground, or space-spaced
-bolometer cameras, SMURF has been designed to use a less accurate,
+\scuba\ is presently unique in terms of the shear volume of data that
+it produces on a nightly basis, and the fact that it is observing a
+wide variety of science targets for multiple users (unlike dedicated
+CMB experiments, for example). For this reason, SMURF has been
+designed to use a less accurate (primarily on large angular scales),
 though faster and less memory-intensive iterative approach that
 attempts to model and remove most of the correlated signal components,
-and then regrid the residuals signals. This method is similar in
-spirit to the Comprehensive Reduction Utility for SHARC-2
+and then regrid the residual signals. This method is similar in spirit
+to the Comprehensive Reduction Utility for SHARC-2
 \citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
-Bolocam Galactic Plane Survey \citep{aguirre2010}.
+Bolocam Galactic Plane Survey \citep{aguirre2010}. While most of the
+analysis described here applies equally to all of the data taken
+to-date with \scuba, we note that the SCUBA-2 Shared-Risks Observing
+period (S2SRO) which took place from 2010 February 11 to 2010 March 25
+used only a single subarray at each of the two observing wavelengths,
+and had considerably different noise properties and poorer performance
+than the fully-commissioned instrument in 2011.
 
@@
 
-Before describing the data-reduction algorithm in detail, we summarize
-the basic strategy used to make measurements with SCUBA-2.
-
-The first generation of submm bolometer cameras used NTD germanium
-thermistors (e.g., SCUBA, MAMBO, BOLOCAM, AzTEC, LABOCA...).
-Generally speaking, these devices are operated using a current bias;
-changing thermal conditions alter their impedence, resulting in a
-changing Voltage that can be measured. The bolometers are cooled so
-that they are {\em background limited} -- the dominant noise term is
-the optical load from sources outside of the cryostat (emissivity from
-telescope, atmospheric fluctuations, and astronomical sources).
-
-In contrast, SCUBA-2 belongs to a new generation of cameras that uses
-superconducting transition-edge sensors (TESs, like ACT, SPT,
-MUSTANG). These devices are held at the critical temperature,
-$T_\mathrm{C}$, below which the material becomes a superconductor, and
-above which it becomes a normal conductor. Within this transition,
-tiny temperature fluctuations can produce huge changes in
-impedence. The devices are operated with a voltage bias, so that
-thermal fluctuations produce changing currents, which in turn generate
-changing local magnetic fields. Chains of superconducting quantum
-interference devices (SQUIDs) are used to amplify these signals,
-ultimately providing a larger changing current that may be
-digitized. Like the NTD thermistor-based cameras, SCUBA-2 is cooled in
-order to be background limited. However, a superior over-all
-sensitivity on a per-bolometer basis is achieved through its improved
-responsivity (similar to an improved quantum efficiency for a CCD).
-
-The load on the detectors can vary substantially, both from thermal
-sources within and outside the SCUBA-2 cryostat. In order to operate
-the camera across a broad dynamic range of conditions, it is therefore
-necessary to thermally control the focal plane. This is achieved by
-providing a fridge {\em base temperature}, $T_\mathrm{B}$ that is
-well-below $T_\mathrm{C}$, and then using heaters that are connected
-to each bolometer to place the array in to transition.
-
-The focal plane is divided into four quadrants, each filled with 32
-column $\times$ 40 row sub-arrays labelled s4a--s4d at 450\,\micron,
-and s8a--s8d at 850\,\micron. To avoid independent bias and readout
-wires to each bolometer, each array is multiplexed by column. There is
-an additional row of dark SQUIDs with no TESs or thermal absorber used
-to monitor non-thermal noise sources.
+While the details of how \scuba\ works is described in
+\citet{holland2011}, and its calibration in \citet{dempsey2011}, we
+summarize the basic concepts relative to map-making here.
+
+The \scuba\ bolometer temperatures are read-out using superconducting
+transition-edge sensors (TESs), with critical temperatures
+$T_\mathrm{C}$ (...typical value...). The range about $T_\mathrm{C}$
+in which the transition from a superconductor to a normal material
+occurs is extremely small (...typical value...?). If the devices are
+biased to a temperature within this transition on large time scales,
+then tiny temperature fluctuations on short time-scales will produce
+significant changes in the impedence. The entire SCUBA-2 focal plane
+is therefore held at a \emph{base temperature} somewhat below
+$T_\mathrm{C}$. The TESs then have a small electrical voltage bias,
+and individual pixel heaters (using a current bias to apply the
+desired Ohmic heating) are used to place the detectors in the
+transition. Only a small detector bias voltage is required since the
+heaters compensate for most of the variations in the base temperature,
+and other ambient loads (like the atmosphere).  Small temperature (and
+therefore impedence) fluctuations produce a changing current through
+the TESs. The resulting changing magnetic fields are then amplified
+using chains of superconducting quantum interference devices (SQUIDs),
+and the larger output current is finally digitized. Each detector has
+its own SQUID for the first-stage of the amplification, but the
+remaining stages occur within a common chain of SQUIDs for each column
+of 40 detectors. Each of the 32 rows is then read-out in sequence that
+occurs at a frequency of ...kHz?... This multiplexing scheme greatly
+reduces the number of readout and bias lines to each detector, making
+such large-format bolometer array designs feasible. It should also be
+noted that there is an additional 33rd row of SQUID reasouds that are
+not connected to TES. These ``dark SQUIDs'' are used to track
+non-thermal noise sources that are common to each column's amplifier
+chain.
+
+The focal planes at each of 450 and 850\,\micron\ are populated with
+four of these $40 \times 32$ subarrays (referred to as s4a--s4d, and
+s8a--s8d respectively), and simultaneously image approximately the
+same region of the sky by means of a beam-splitter.
 

### 2011-10-21 smurf_paper.tex
more work on intro, re-organization of early sections

@@
 out of techniques that were developed to produce maps of the
-large-scale anisitropies in the Cosmic Microwave Backgroubd.  Styles
+large-scale anisitropies in the Cosmic Microwave Background (CMB).  Styles
 of map-making: maximum likelihood \citep[e.g.,][]{patanchon2008};
@@
 CMB experiments, for example). For this reason, SMURF has been
-designed to use a less accurate (primarily on large angular scales),
-though faster and less memory-intensive iterative approach that
-attempts to model and remove most of the correlated signal components,
-and then regrid the residual signals. This method is similar in spirit
-to the Comprehensive Reduction Utility for SHARC-2
+designed to use a faster, and less memory-intensive algorithm than
+CMB-style maximum likelihood map-makers, although it is less accurate
+at reproducing large-scale structures. The basic procedure is to
+iteratively model and remove most of the bolometer-bolometer
+correlated signal components and independent low-frequency noise, and
+then regrid the residual signals. This method is similar in spirit to
+the Comprehensive Reduction Utility for SHARC-2
 \citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
@@
 %------------------------------------------------------------------------------
-\section{Data Properties}
+\section{\scuba\ data properties}
 \label{sec:data}
@@
 
-My thought is that this section would give the basics about how
-SCUBA-2 works, and then more specific details for the S2SRO
-observations.
+In this section we summarize how \scuba\ bolometers detect temperature
+variations (Section~\ref{sec:bolos}), typical \scuba\ scan strategies
+(Section~\ref{sec:scan}).
 
 %--------------------------------------------------
-\subsection{How \scuba\ takes measurements}
+\subsection{How the \scuba\ bolometers work}
+\label{sec:bolos}
 %--------------------------------------------------
@@
 is therefore held at a \emph{base temperature} somewhat below
-$T_\mathrm{C}$. The TESs then have a small electrical voltage bias,
+$T_\mathrm{C}$. The TESs then have a small electrical voltage biases,
 and individual pixel heaters (using a current bias to apply the
-desired Ohmic heating) are used to place the detectors in the
+desired Joule heating) are used to place the detectors in the
 transition. Only a small detector bias voltage is required since the
@@
 remaining stages occur within a common chain of SQUIDs for each column
-of 40 detectors. Each of the 32 rows is then read-out in sequence that
-occurs at a frequency of ...kHz?... This multiplexing scheme greatly
+of 40 detectors. Each of the 32 rows is then read-out in sequence at a
+row-visit frequency of ...kHz?... This multiplexing scheme greatly
 reduces the number of readout and bias lines to each detector, making
@@
 
-The focal planes at each of 450 and 850\,\micron\ are populated with
-four of these $40 \times 32$ subarrays (referred to as s4a--s4d, and
-s8a--s8d respectively), and simultaneously image approximately the
-same region of the sky by means of a beam-splitter.
-
-The camera has been designed so that the absolute response of each
-detector can be established by ramping currents through the individual
-pixel heaters to produce Joule heating power, $P_\mathrm{J} =
-\epsilon I^2R$, where $I$ is the known current, $R$ is the heater
-resistance, and $\epsilon$ is the efficiency with which power is
-transferred to the bolometer. In order to calibrate the total factor
-$\epsilon R$, SCUBA-2 posesses an internal flatfield source.
+The response of the output digitized current, $I$, with respect to a
+change in the power $P$ incident on the bolometer, $dI/dP$ depends on
+the efficiency of the thermal absorber, how well the TES is biased
+into the transition, and the gain of the SQUID amplifiers. Since there
+are variations in $T_\mathrm{C}$ across each subarray due to
+manufacturing difficulties, considerable effort has gone into
+dynamically applying position-dependent biases as well as an over-all
+focal plane bias to make as many detectors useable as possible
+\citep{holland2011}. In practice, the response of a given detector can
+therefore vary somewhat from observation to observation depending on
+how the bias and amplification parameters were set up. To track these
+changes, at the start of each observation, the current is ramped
+through the bolometer heaters such that $dI/dP$ can be measured. Note
+that the bolometers may have a slightly different response to heater
+power due to conductance as compared to the response to optical power,
+so additioanl efficiency corrections must also be obtained. While not
+discussed further in this paper, details may be found in
+\citet{dempsey2011}. We will refer to measurements of $dI/dP$ across
+the focal plane as the \emph{flatfield}. Fig.~\ref{fig:sensitivities}
+shows the flatfield for each focal plane.
+
+\begin{figure}
+\vspace{2cm}
+\textbf{This will be a figure showing s4a and s8d responsivity maps}
+\vspace{2cm}
+\caption{The response of each bolometer to known input heater power
+  for a typical night during S2SRO. The s4a subarray had a
+  particularly large variation in $T_\mathrm{C}$ towards the
+  bottom-right which resulted in lower responsivity, and
+  correspondingly larger noise. The s8d array also suffered
+  significant variations across the focal plane, in particular
+  removing larger number of useful detectors from the bottom-left.}
+\label{fig:sensitivities}
+\end{figure}
 
@@
 digitization.
+
+%--------------------------------------------------
+\subsection{Scan strategies}
+\label{sec:scan}
+%--------------------------------------------------
 
@@
 pattern to produce similarly periodic noise structures in the map, or
-\emph{scan-synchronous} noise. It has therefore been recognized for
-some time that scan strategies that visit each region of the map on as
-many different time scales, and scan angles as possible offer the best
-protection against large-scale noise (...cite some people...); such
-strategies are said to offer good \emph{cross-linking}. As a tradeoff
-between truly random scan patterns and feasibility of implementation
-taking into consideration both mechanical and scheduling constraints,
-SCUBA-2 presently has three primary scan modes: PONG (cite Borys et
-al.), Lissajous, and Daisy... (more citations? Kovacs et al. has a
-nice description... also CMB literature).
+\emph{scan-synchronous} noise. It has therefore been recognized that
+scan strategies that visit each region of the map on as many different
+time scales, and scan angles as possible offer the best protection
+against large-scale noise (...cite some people...); such strategies
+are said to offer good \emph{cross-linking}. As a tradeoff between
+truly random scan patterns and feasibility of implementation taking
+into consideration both mechanical and scheduling constraints, SCUBA-2
+presently has three primary scan modes: PONG (cite Borys et al.),
+Lissajous, and Daisy... (more citations? Kovacs et al. has a nice
+description... also CMB literature).
 
 %-------------------------------------------------
-\subsection{\scuba\ shared-risk observations}
+\subsection{Typical bolometer signals}
 %-------------------------------------------------
-
-The S2SRO period, as mentioned above, took place during February and
-March 2010. For the purpose of this paper, we will consider data taken
-as early as 2009 December 3, at which point the instrument was being
-commissioned in a similar configuration to that of the S2SRO
-period. At this time, each of the 450 and 850\,\micron\ focal planes
-were populated with single subarrays, s4a and s8d respectively. There
-were two main mechanical problems to contend with during this
-period. First, the fridge was not able to cool the focal plane to the
-desired base temperature of ... which affected stability, particularly
-at 850\,\micron. Second, there were problems with the subarray
-fabrication process which led to large numbers of dead pixels, as well
-as large-scale variations in $T_\mathrm{C}$ and the thermal
-conductivity, $G$. These latter issues made it difficult to
-simulataneously lock all of the bolometers, and in addition many of
-the bolometers that were locked did not achieve their minimum
-theoretical noise performance, since it is not possible to adjust the
-individual temperature biases and place each device into the optimal
-region of the superconducting transition.
-
-While there were moderate variations from night-to-night during the
-observing period, there were typically 600 to 900 working detectors in
-each of the s4a and s8d subarrays. Fig.~\ref{fig:sensitivities} shows
-representative responsivity maps for the two subarrays from the middle
-of S2SRO based on the response to ramping the bolometer heaters.
-
-\begin{figure}
-\vspace{2cm}
-\textbf{This will be a figure showing s4a and s8d responsivity maps}
-\vspace{2cm}
-\caption{The response of each bolometer to known input heater power
-  for a typical night during S2SRO. The s4a subarray had a
-  particularly large variation in $T_\mathrm{C}$ towards the
-  bottom-right which resulted in lower responsivity, and
-  correspondingly larger noise. The s8d array also suffered
-  significant variations across the focal plane, in particular
-  removing larger number of useful detectors from the bottom-left.}
-\label{fig:sensitivities}
-\end{figure}
-
-\subsubsection{Time series}
 
@@
 load from the fridge at each wavelength, but a larger contribution of
-astmospheric variations through the 450\,\micron\ bandpass filters. We
+atmospheric variations through the 450\,\micron\ bandpass filters. We
 also note that there is no obvious strong correlation between the
-low-frequency signal structure at either wavelength at the telescope
-motion.
+low-frequency signal structure, at either wavelength, with the
+telescope motion.
 
@@
 highly correlated amongst bolometers in the same subarray. We have
-calculate a common-mode signal, $\mathbf{c}(t)$ as the straight
-average time series of all the working bolometers. We then fit the
-amplitude of $\mathbf{c}(t)$ at each wavelength to the time series
-shown in Fig.~\ref{fig:bolos_mix} and remove it, yielding the grey
-residual signals. These residuals are quite flat, although still with
+calculated a common-mode signal, $\mathbf{c}(t)$, as the average time
+series of all the working bolometers. We then fit the amplitude of
+$\mathbf{c}(t)$ at each wavelength to the signals shown in
+Fig.~\ref{fig:bolos_mix} and remove it, yielding the grey residual
+signals. These residuals are quite flat, although still with
 noticeable variations. The white noise is also apparent, and larger at
-450\,\micron\ as one would expect from the larger loading compared to
-850\,\micron.
-
-\subsubsection{Power spectral densities}
-\label{sec:psd}
+450\,\micron\ as one would expect from the larger backgrounds compared
+to 850\,\micron.
 
@@
 
-The dotted coloured lines show the PSDs for raw, though flatfielded
-data. At each wavelength, there is a clear $1/f$ knee at a few Hz,
-followed by a predominantly white spectrum punctuated by line features
-above $\gsim 30$\,Hz, and finally roll-off caused by the anti-aliasing
-filter above $\gsim 70$\,Hz. As indicated in the previous section, the
-correlation between the low-frequency components of the different
-bolometer signals is large. The solid black lines in
-Fig.~\ref{fig:pspec} indicate the PSDs of the common-modes
-$\mathbf{c}(t)$ at each wavelength, which reproduce most of the
-low-frequency structures, as well as the high-frequency line
+The dotted coloured lines in Fig~\ref{fig:pspec} show the PSDs for
+raw, though flatfielded data. At each wavelength, there is a clear
+$1/f$ knee at a few Hz, followed by a predominantly white spectrum
+punctuated by line features above $\gsim 30$\,Hz, and finally roll-off
+caused by the anti-aliasing filter above $\gsim 70$\,Hz. As indicated
+in the previous section, the correlation between the low-frequency
+components of the different bolometer signals is large. The solid
+black lines in Fig.~\ref{fig:pspec} indicate the PSDs of the
+common-modes $\mathbf{c}(t)$ at each wavelength, which reproduce most
+of the low-frequency structures, as well as the high-frequency line
 features. The $\mathbf{c}(t)$ otherwise drop substantially below the
@@
 bolometers are dominated by un-correlated white noise. Finally, we
-note that the amplitudes and slopes of the $\mathbf{c}(t)$ are
-similar at both 450 and 850\,\micron. However, since the
-450\,\micron\ white noise is larger than at 850\,\micron\ (shown
-approximately by the horizontal dotted lines), the $1/f$ knee occurs
-at a \emph{lower} frequency at 450\,\micron.
+note that the amplitudes and slopes of the $\mathbf{c}(t)$ are similar
+at both 450 and 850\,\micron. However, since the 450\,\micron\ white
+noise is larger than at 850\,\micron\ (shown approximately by the
+horizontal dotted lines), the $1/f$ knee occurs at a \emph{lower}
+frequency at 450\,\micron.
 
@@
 
-\subsubsection{Principal component analysis}
+\subsection{Magnetic field pickup}
+\label{sec:magpickup}
+
+An additional noise source that appears to be significant in only a
+small subset of the S2SRO data is magnetic field pickup. Since the
+bolometer signals are ultimately detected through the amplification of
+magnetic fields, any additional changing fields within the cryostat
+will also be detected.
+
+An example of an observation where pickup appears to be significant is
+shown in Fig.~\ref{fig:magpickup}. The time-series for two
+450\,\micron\ bolometers in the same column (not flatfielded) show
+that there is a strong signal with a similar shape, but opposite
+signs. This behaviour is seen across the entire array. The dark squid
+signal for the same columns exhibits a similar shape and
+amplitude. Since the dark squid has no thermal absorber or TES
+attached to it, this observed signal is not likely to be optical or
+thermal in nature (although there can be some crosstalk with the
+bolometers...). It is also for this reason that the higher-frequency
+noise that appears in the bolometer data (thermal) does not appear in
+the dark squid, leading to a significantly higher \snr. The sign of
+such a signal is expected to be random in the bolometer time-series
+(ask detector person for an explanation).
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{magpickup.pdf}
+\caption{Evidence for significant magnetic field pickup for
+  observation 20100228\_00016.  The top panel shows two un-flatfielded
+  (but mean-subtracted) bolometer time-series from the same column,
+  with a 200 sample boxcar smooth (approximately 1\,s), illustrating
+  that they are dominated by a similar signal with opposite signs. The
+  second panel shows the dark squid signal for the column, also
+  mean-subtracted and with the same boxcar smooth. The bottom panel
+  shows the azimuthal and elevation offsets from the map centre. For
+  reference, the mean azimuth was 171.9\,deg, and the mean elevation
+  68.0\,deg. Only the azimuthal signal appears to be correlated with
+  the dark squids and bolometer signals, which suggests a magnetic
+  field stationary with respect to the telescope dome as the source,
+  since its direction with respect to the cryostat only changes with
+  azimuthal motion.}
+\label{fig:magpickup}
+\end{figure}
+
+The telescope pointing offsets for this approximately 0.5\,deg
+diameter scan are also shown in Fig.~\ref{fig:magpickup}. Since the
+phase of the azimuth offsets from the map centre in this scan pattern
+slowly drifts with respect to the elevation offsets, it is clear that
+the bolometer and dark squid signals are detecting a noise source that
+is correlated only with the azimuthal motion and \emph{not} the
+elevation. This behaviour would be expected if if there were a strong
+magnetic field fixed with respect to the telescope dome (i.e., the
+earth's magnetic field). Since SCUBA-2 is mounted on a Nasmyth
+platform, only azimuthal motion will change the effective direction of
+such a field with respect to the cryostat. It is also worth noting
+that the absolute azimuth of the telescope was close to 180\,deg
+(i.e., nearly due-south). We suspect that this observing
+configuration, combined with the large amplitude of the scan, may have
+conspired to produce these large signals.
+
+%-------------------------------------------------
+\subsection{Comparison between \scuba\ shared-risk observations and the 2011 re-commissioned instrument}
+%-------------------------------------------------
+
+The S2SRO period took place during February and March 2010. For the
+purpose of this paper, we will consider data taken as early as 2009
+December 3, at which point the instrument was being commissioned in a
+similar configuration to that of the S2SRO period. At this time, each
+of the 450 and 850\,\micron\ focal planes were populated with single
+subarrays, s4a and s8d respectively. There were two main mechanical
+problems to contend with during this period. First, the fridge was not
+able to cool the focal plane to the desired base temperature of
+... which affected stability, particularly at 850\,\micron. Second,
+there were problems with the subarray fabrication process which led to
+large numbers of dead pixels, as well as large-scale variations in
+$T_\mathrm{C}$ and the thermal conductivity, $G$. These latter issues
+made it difficult to simulataneously place all of the TESs into the
+optimal region of the superconducting transition, leading to
+significantly poorer sensitivities.
+
+While there were moderate variations from night-to-night during the
+observing period, there were typically 600 to 900 working detectors in
+each of the s4a and s8d subarrays. Fig.~\ref{fig:sensitivities} shows
+representative responsivity maps for the two subarrays from the middle
+of S2SRO based on the response to ramping the bolometer heaters.
+
+%-------------------------------------------------
+\subsection{Principal component analysis}
 \label{sec:pca}
+%-------------------------------------------------
 
@@
 
-\subsubsection{Magnetic field pickup}
-\label{sec:magpickup}
-
-An additional noise source that appears to be significant in only a
-small subset of the S2SRO data is magnetic field pickup. Since the
-bolometer signals are ultimately detected through the amplification of
-magnetic fields, any additional changing fields within the cryostat
-will also be detected.
-
-An example of an observation where pickup appears to be significant is
-shown in Fig.~\ref{fig:magpickup}. The time-series for two
-450\,\micron\ bolometers in the same column (not flatfielded) show
-that there is a strong signal with a similar shape, but opposite
-signs. This behaviour is seen across the entire array. The dark squid
-signal for the same columns exhibits a similar shape and
-amplitude. Since the dark squid has no thermal absorber or TES
-attached to it, this observed signal is not likely to be optical or
-thermal in nature (although there can be some crosstalk with the
-bolometers...). It is also for this reason that the higher-frequency
-noise that appears in the bolometer data (thermal) does not appear in
-the dark squid, leading to a significantly higher \snr. The sign of
-such a signal is expected to be random in the bolometer time-series
-(ask detector person for an explanation).
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{magpickup.pdf}
-\caption{Evidence for significant magnetic field pickup for
-  observation 20100228\_00016.  The top panel shows two un-flatfielded
-  (but mean-subtracted) bolometer time-series from the same column,
-  with a 200 sample boxcar smooth (approximately 1\,s), illustrating
-  that they are dominated by a similar signal with opposite signs. The
-  second panel shows the dark squid signal for the column, also
-  mean-subtracted and with the same boxcar smooth. The bottom panel
-  shows the azimuthal and elevation offsets from the map centre. For
-  reference, the mean azimuth was 171.9\,deg, and the mean elevation
-  68.0\,deg. Only the azimuthal signal appears to be correlated with
-  the dark squids and bolometer signals, which suggests a magnetic
-  field stationary with respect to the telescope dome as the source,
-  since its direction with respect to the cryostat only changes with
-  azimuthal motion.}
-\label{fig:magpickup}
-\end{figure}
-
-The telescope pointing offsets for this approximately 0.5\,deg
-diameter scan are also shown in Fig.~\ref{fig:magpickup}. Since the
-phase of the azimuth offsets from the map centre in this scan pattern
-slowly drifts with respect to the elevation offsets, it is clear that
-the bolometer and dark squid signals are detecting a noise source that
-is correlated only with the azimuthal motion and \emph{not} the
-elevation. This behaviour would be expected if if there were a strong
-magnetic field fixed with respect to the telescope dome (i.e., the
-earth's magnetic field). Since SCUBA-2 is mounted on a Nasmyth
-platform, only azimuthal motion will change the effective direction of
-such a field with respect to the cryostat. It is also worth noting
-that the absolute azimuth of the telescope was close to 180\,deg
-(i.e., nearly due-south). We suspect that this observing
-configuration, combined with the large amplitude of the scan, may have
-conspired to produce these large signals.
-
 %------------------------------------------------------------------------------

### 2011-10-22 smurf_paper.tex
more edits

@@
 how the bias and amplification parameters were set up. To track these
-changes, at the start of each observation, the current is ramped
-through the bolometer heaters such that $dI/dP$ can be measured. Note
-that the bolometers may have a slightly different response to heater
-power due to conductance as compared to the response to optical power,
-so additioanl efficiency corrections must also be obtained. While not
-discussed further in this paper, details may be found in
-\citet{dempsey2011}. We will refer to measurements of $dI/dP$ across
-the focal plane as the \emph{flatfield}. Fig.~\ref{fig:sensitivities}
-shows the flatfield for each focal plane.
+responsivity changes, at the start of each observation, the current is
+ramped through the bolometer heaters such that $dI/dP$ can be
+measured. Note that the bolometers may have a slightly different
+response to heater power due to conductance as compared to the
+response to optical power, so additioanl efficiency corrections must
+also be obtained. While not discussed further in this paper, details
+may be found in \citet{dempsey2011}. We will refer to measurements of
+$dI/dP$ across the focal plane as the
+\emph{flatfield}. Fig.~\ref{fig:sensitivities} shows the flatfield for
+each focal plane.
 
@@
 digitization.
+
+The final conversion of power received by the detector into
+astronomical flux units requires two more fundamental corrections: (i)
+atmospheric extinction (due to atmospheric water vapour) must be
+compensated for; and (ii) the corrected power is related to an
+astronomical flux density through the effective area of the detector,
+the spectral response of the bolometers (bandpass), the spectrum of
+the source, and additional optical efficiency factors -- a \emph{flux
+  conversion factor} (FCF). Extinction is measured, in most cases,
+using the JCMT line-of-sight water vapour radiometer (WVM) which is
+calibrated using skydips and the CSO 225\,GHz tipper. The FCF is
+determined empirically using measurements of standard sources. For
+further details on these calibration steps, see \citet{dempsey2011}.
 
@@
 %-------------------------------------------------
-\subsection{Comparison between \scuba\ shared-risk observations and the 2011 re-commissioned instrument}
+\subsection{Differences between the S2SRO and 2011 re-commissioned instrument}
 %-------------------------------------------------
@@
 A method that has been employed by teams analyzing BOLOCAM and AzTEC
-data to identify and remove low-frequency correlated noise is
-Principal Component Analysis
-\citep[PCA,][]{laurent2005,scott2008}. While we have not (yet)
-included ``PCA cleaning'' as a standard feature of SMURF, we have
-performed a preliminary analysis in IDL of some SCUBA-2 data to gain
-further insight into the nature of the low-frequency noise. The basic
-method is as follows: (i) the covariance matrix is built up for all
-pairs $(i,j)$ of the $N$ bolometer time series,
-$\langle\mathbf{b}_i(t),\mathbf{b}_j\rangle$; and (ii) an eigenvalue
-decomposition is used to identify a new set of statistically
+data to identify and remove correlated noise is Principal Component
+Analysis \citep[PCA,][]{laurent2005,scott2008}. We initially used PCA
+to gain further insight into the nature of the \scuba\ noise
+properties, but it is now also included it in SMURF as method for
+cleaning the data.
+
+The basic method is as follows: (i) a covariance matrix is built up
+for all pairs $(i,j)$ of the $N$ bolometer time series,
+$\langle\mathbf{b}_i(t),\mathbf{b}_j\rangle$; and (ii) a singular
+value decomposition is used to identify a new set of statistically
 uncorrelated eigenvectors, $\mathbf{\xi}_i(t)$ (i.e. whose covariance
@@
 and $\mathbf{\lambda}_i^T$ is a column vector containing the
-corresponding eigenvalues. For convenience, the $\mathbf{\xi}_i$ are
-normalized by their \rms, so that the amplitude of each component is
-stored in the eigenvalues. In the earlier analyses mentioned, the
-low-frequency noise is assumed to be encapsulated in those components
-with the largest eigenvalues. Removing the projection of the time
-series along those components has been used successfully to reduce
-$1/f$ noise while retaining most of the signal in point-like sources.
+corresponding eigenvalues. The $\mathbf{\xi}_i$ are normalized by
+their \rms, so that the amplitude of each component is stored in the
+eigenvalues. In the earlier analyses mentioned, the low-frequency
+noise is assumed to be encapsulated in those components with the
+largest eigenvalues. Removing the projection of the time series along
+those components has been used successfully to reduce $1/f$ noise
+while retaining most of the signal in point-like sources. There is an
+implicit assumption in this analysis that the noise properties are
+stationary.
 
@@
 
-Given predictions for the performance of a fully-functioning SCUBA-2,
-and the observed properties of the first-installed subarrays during
-commissioning in late 2009 as described in the previous section, SMURF
-was designed with two primary goals in mind: (i) reduced maps should
-be scientifically useful and of ``near publication quality'' with
-minimal user interaction; and (ii) the execution time should not
-appreciably exceed the amount of time required to conduct
-observations. By attaining these goals it would be possible to conduct
-real-time reduction at the telescope to offer feedback to observers,
+Given \scuba's data rate, SMURF was designed with two primary goals in
+mind: (i) reduced maps should be scientifically useful and of ``near
+publication quality'' with minimal user interaction; and (ii) the
+execution time should not appreciably exceed the amount of time
+required to conduct observations on a high-end multi-tcore desktop
+computer. By attaining these goals it would be possible to conduct
+real-time reductions at the telescope to offer feedback to observers,
 and it would also be possible to maintain a central repository of
@@
 
-There are two general families of existing map-making strategies that
-were considered. The first are \emph{maximum likelihood} methods that
-were originally pioneered for the analysis of Cosmic Microwave
-Background (CMB) maps. Here, the pixels in the map are considered to
-be free parameters in a model-fitting problem, and the data are the
-raw time-series. By estimating the full noise covariance matrix for
-the data, it is then possible to estimate the maximum-likelihood
-map. While direct inversion solutions to this problem are untenable,
-some recent efforts to optimize these algorithms have enabled the
-reduction of smaller data sets in a reasonable amount of time
-\citep[e.g.][]{patanchon2008}. While maximum likelihood maps are,
-theoretically, the best maps that can be made from a noise point of
-view, the required computing resources and execution time would not
-have been feasible for the proposed real-time data reduction and data
-archiving model.
+As described in the introduction, rather than pursue a slow but
+potentially more accurate \emph{maximum likelihood} method that would
+require the inversion of a complex linear problem, we instead use an
+iterative approach that estimates and removes most of the correlated
+(and in particular low-frequency) noise sources in parallel with a
+simplified map estimation. We do note, however, that recent advances
+in direct inversion methods such as \citet{patanchon2008} may in fact
+be tenable for SCUBA-2 data. A future goal of the software team is to
+investigate meshing SMURF (perhaps as a cleaning step and initial
+condition) for a SANEPIC-like algorithm.
+
+The remainder of this section gives an overview of the model that is
+assumed by SMURF for the \scuba\ data, and the typical steps used to
+estimate a map. Subsequent sections discuss problems with this model
+and additional processing steps and constraints that can be used to
+improve the solution in different cases.
 
@@
 converts this effective power to the digitized units recorded on disk
-(DAC), and in this formulation is assumed to be constant in time.
-
-In practice, $g_i$ is further factored into two components: $g_i = r_i
-d$. The responsivity, $r_i$, gives the change in current as a function
-of power fluctuations (A\,pW$^{-1}$). There is then a final (and
-constant across all bolometers) conversion factor from current to DAC
-units, $d$ (DAC/A).
-
-As for the noise, $\mathbf{n}_i(t)$, we break it down into several
+(DAC) -- the \emph{flatfield} multiplied by a digitization constant --
+and in this formulation it is assumed to be constant in time.
+
+We then express the noise, $\mathbf{n}_i(t)$, as the sum of several
 components,
@@
 map-making is to model and remove $\mathbf{n}^\mathrm{f}_i$ and
-$\mathbf{n}^\mathrm{c}$ from the time-series. By re-gridding the
+$\mathbf{n}^\mathrm{c}$ from the time-series, while re-gridding the
 remaining data we can then hope to approach the theoretical noise
@@
 has been removed, we apply the extinction correction, which is
-determined completely from external measurements. Next, residual noise
-that differse from bolometer-to-bolometer (primarily low-frequency) is
-removed using an FFT-based filter. The resulting ``clean'' time-series
-are then re-gridded to produce a map estimate using nearest-neighbour
-sampling. Since each map pixel averages many bolometer samples
-together, it is fundamentally much less noisy than the bolometer
-time-series. We can therefore project the map back into the time
-domain, and create relatively noiseless estimates of what each
+determined completely from external measurements
+\citep[see][]{dempsey2011}. Next, residual noise that differs from
+bolometer-to-bolometer (primarily low-frequency) is removed, usually
+using a high-pass filter. The resulting ``clean'' time-series are then
+re-gridded to produce a map estimate using nearest-neighbour
+sampling. Since each map pixel is calculated as the average of many
+bolometer samples, it is fundamentally much less noisy than the raw
+bolometer time-series. We can therefore project the map back into the
+time domain (i.e. scanning the array across the current estimate of
+the map), and create approximately noiseless estimates of what each
 bolometer would have seen in the absence of noise sources, and remove
@@
 
-AGG
-
-This is actually only calculated once before the map solution begins
-using external opacity monitors (WVM, $\tau_\mathrm{CSO}$,
-skydip). Summarize the measurements, order of preference and something
-about the calibration for the SCUBA-2 filters.
+The extinction correction is normally derived using the JCMT water
+vapour radiometer (WVM), and is not considered to be a free parameter
+in the solution. For details on how it is calculated see
+\citet{dempsey2011}. Note that its numerical value is calculated only
+once, and simply applied as a scale factor in the iterative solution.
 
@@
 applied before the transform to avoid ringing (primarily caused by
-wrap-around discontinuities at the ends of the time-series). The
-signal that is \emph{removed} from the time-series by this process are
-stored in the \texttt{FLT} container array.
+wrap-around discontinuities at the ends of the time-series). Finally,
+a whitening filter may also be applied in which a simple $1/f^\alpha +
+\mathrm{constant}$ is fit to the power spectrum of each bolometer, and
+then the bolometer FFT is multiplied by its inverse (though normalized
+to preserve the white noise level).  The signal that is \emph{removed}
+from the time-series by this process are stored in the \texttt{FLT}
+container array.
 
@@
 improvement, or injects noise since the magnetic field pickup noise is
-generally sub-dominant.
+generally sub-dominant. Furthermore, not every column that contains
+working bolometers also has a working dark squid. An alternative is to
+use the \texttt{TMP} model described below.
+
+\subsubsection{\texttt{TMP}: remove a scaled template}
+
+Since the \texttt{DKS} cannot generally be used due to the missing
+dark squids, an alternative is to use the telescope pointing as a
+proxy for the signal due to magnetic field pickup
+(Section~\ref{sec:magpickup}. Generally speaking the azimuth (or its
+sine) as a function of time is used as a template that is simply
+scaled and removed from each bolometer signal. Some modest
+improvements can be achieved with this model.
 
@@
 
-Introduce the topic of model degeneracies here, but defer their
-solution to the example sections.
+A fundamental problem faced by the SMURF algorithm, as described, is
+that there are degeneracies between model parameters. Thinking of each
+pixel in the output map, as well as each sample in the various noise
+model components, such as the common-mode and signal removed by the
+Fourier filtering, any degeneracy can cause the solution to
+diverge. Of particular concern are situations in which the map
+diverges.
 

### 2011-11-11 smurf_paper.tex
Frossie is a member of the core DR team

@@
     Tim~Jenness$^{2}$,
-    Douglas~Scott$^{1}$
+    Douglas~Scott$^{1}$,
+    Frossie~Economou$^2$
   }

### 2011-11-11 smurf_paper.tex
Add a couple of comments to sections.

@@
 than the fully-commissioned instrument in 2011.
+
+\textit{May want to note that the pipeline is discussed elsewhere,
+  e.g. Gibb et al 2006.}
 
@@
 
+\textit{There are also the popped bolometers that could help fill the
+gaps.}
+
 \subsubsection{\texttt{TMP}: remove a scaled template}
@@
 amplitude, astronomical sources, and the internal flatfield source.
+
+\textit{Surely flatfields are discussed in the main SCUBA-2 paper so
+don't really have any useful purpose here since they aren't
+technically part of the map-maker.}
 
@@
 
+\textit{And matched filter is discussed here or superceded by
+the 2D FFTs?}
+
 \subsection{Bright extended emission}
@@
 
+\textit{Or now an externally supplied mask.}
+
 %------------------------------------------------------------------------------
@@
 
+\textit{Presumably the standard telescope acknowledgment here plus a thank you
+to CFI for funding the DR work. Ed should probably thank both CANFAR
+and JAC for additionally funding although his JAC funding is noted in
+the affiliation.}
+
 %------------------------------------------------------------------------------

### 2011-11-16 smurf_paper.tex
Add description of COM and GAI

@@
 
-DSB
-
-We estimate the common-mode, remove it from the time series, use it to
-flag bad bits of data, check for time-varying bolometer
-responsivities... what doesn't it do?! Also mention the optional
-presence of GAI.
+﻿Fig.~\ref{fig:com} shows the time series from a selection of typical
+bolometers.\footnote{The data have been flat-fielded and each time series
+has been adjusted to a mean value of zero.} The similarity between most
+bolometers is evident, and forms the common-mode signal - assumed to be a
+consequence of variations in the atmospheric emission and fridge
+temperature. This common-mode usually dominates the astronomical signal
+for all but the brightest sources, and swamps faint extended structure.
+The purpose of the COM and GAI models is to remove this common-mode
+signal.
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{com.pdf}
+\caption{A selection of typical bolometer time series after flatfielding
+and removal of a constant baseline. It can be seen that most bolometers
+exhibit a common time variation overlayed with other features.}
+\label{fig:com}
+\end{figure}
+
+The COM model is the common-mode signal itself. It is a single time
+series that is estimated by finding the mean of all bolometers at each
+time step. Bolometer values that have been flagged as unusable for any
+reason are excluded from the mean.
+
+Since bolometers have varying sensitivities, the amplitude of the
+common-mode variations will also vary from bolometer to bolometer. So
+comparing each bolometer time stream with the common-mode signal allows
+an estimate of the bolometer sensitivity to be obtained. In practice, a
+least squares linear fit is performed between the bolometer time series
+and the common-mode to determine a gain and offset for each bolometer.
+The gain and offset for each bolometer is known as the GAI model. Each
+bolometer value can then optionally be scaled and shifted using these
+values so that all bolometers share a common (but as yet unknown)
+calibration. This provides an alternative( or additional) flatfielding
+strategy to that described in in section~\ref{sec:bolos}.
+
+An option exists to cater for time-varying sensitivities. If used, the
+least squares fits described above are performed on short blocks of
+contiguous time slices, providing multiple gain and offset values for
+each bolometer (one pair for each block of time slices). The gain and
+offset at any required time slice can then be found by interpolation
+between these values.
+
+It can be seen from Fig.~\ref{fig:com} that some bolometers depart
+radically from the common-mode, indicating some problem within the
+bolometer. Such bolometers are identified by calculating the Pearson
+coefficient of correlation between each bolometer time stream and the
+common mode. Bolometers for which the correlation coefficient is below a
+specified limit, or which have a negative gain, are flagged as bad in
+order to omit them from the final map. If the option described above for
+handling time-varying sensitivities is used, then a correlation
+coefficient can be determined for each individual block of time slices.
+This allows individual bad blocks to be rejected from a bolometer time
+stream, rather than rejecting the whole bolometer. This is beneficial for
+bolometers that follow the common-mode for most of the time but suffer
+from transient problems at specific times.
+
+The common-mode is calculated from all bolometers that have not
+previously been flagged as bad. Thus aberrant bolometers with low
+correlation coefficients will be included, polluting the common-mode. For
+this reason, an iterative approach is taken to estimating the common-mode.
+The initial common-mode estimate is formed from all remaining bolometers.
+Aberrant bolometers with low correlation coefficients are then identified
+by comparing each bolometer time stream with the common-mode. Such
+aberrant bolometers are flagged as bad, and a new estimate of the
+common-mode is produced omitting the newly flagged bolometers. This
+process is repeated until no further bolometers are rejected.
+
+Any astronomical sources that are smaller than the array size will
+contribute signal to some bolometers but not other bolometers, thus
+biassing the simple mean used to estimate the common mode. However, on
+each iteration of the map-making algorithm illustrated in
+Fig.~\ref{fig:dimm}, a large fraction of the remaining astronomical
+signal is extracted from the bolometer time series and transferred to the
+output map, resulting in subsequent estimates of the common-mode being
+more accurate.
+
+Any extended astronomical emission on a scale comparable to or larger
+than the spatial extent of the area used to estimate the common-mode will
+contribute a similar signal to all bolometers. Therefore such extended
+emission is indistinguishable from the other sources of common-mode
+signal (fridge and atmosphere variations) and will be removed by the COM
+model. This places a limit on the spatial extent of astronomical
+structure that can be recovered.
+
+For this reason, the usual practice is to estimate a single COM model by
+examining data from all four sub-arrays in each waveband, since this
+allow spatial structure on the scale of the whole focal plane to be
+recovered. However, sometimes there is evidence that the common-mode
+differs from one array to another, and so an option exists to estimate a
+separate COM model for each individual subarray, with a consequent
+lowering in the scale of spatial structure that can be recovered.
 

### 2011-11-16 smurf_paper.tex
Filled in time-domain despiking section

@@
 
-DSB
+Spikes of short duration and high amplitude are often seen in the time
+series data. If not removed, they can cause ringing in the FLT model
+(Section~\ref{sec:flt}). Two alternative approaches may be used to remove
+these spikes. This section describes the detection and removal of spikes
+within the time series of each bolometer, and
+section~\ref{sec:mapdespike} describes the detection and removal of
+spikes within the final map. In practice, map-based despiking
+(Section~\ref{sec:mapdespike}) usually gives superior results, and so
+time-domain despiking is usually switched off.
+
+\emph{Do we need some discussion of the physical causes and
+characteristics of spikes?}
+
+Each one-dimensional bolometer time series is processed independently. At
+each time slice, the median value of the current bolometer in a box
+centred on the time slice, and containing a specified number of time
+slices (typically 50), is found. If the residual between the
+time slice value and the median value is greater than some specified
+multiple (typically 10) of the local noise level, the time slice is
+flagged as a spike.
+
+If the local noise level were estimated within the same box used to
+determine the median value, a spike in the box would cause the local
+noise level to be over-estimated severely. For this reason, the local
+noise level is taken as the standard deviation of the values within a
+neighbouring box on the ``down-stream'' side of the median box (that is,
+the side that has already been checked for spikes). That is, the high end
+of the noise box is just below the low end of the median filter box. This
+introduces a slight asymmetry in the noise, but this should not matter
+unless the noise varies significantly on a time scale shorter than the
+box size.
+
+This simple algorithm is not very good at distinguishing between spikes
+and bright point sources, and so the threshold for spike detection is
+usually raised when making maps of bright point sources.
 

### 2011-11-25 smurf_paper.tex
Add first part of step correction section

@@
 
-DSB
+\begin{figure*}
+\centering
+\includegraphics[width=\linewidth]{steps1.pdf}
+\caption{Examples of steps in time series data. Steps occur with a wide
+range of heights from the large steps shown on the left to the small
+steps shown on the right. Large steps are often followed by a brief
+``over-shoot'' as shown on the left. In both plots, the black curve is
+the uncorrected time series, and the red curve is the corrected time
+series. Samples close to a step are omitted in the corrected time series.
+In the right hand plot, the blue curve is the uncorrected time series for
+a nearby bolometer. The similarity between the red and blue curves shows
+that the step correction is performing well.
+}
+\label{fig:steps1}
+\end{figure*}
+
+Sudden steps can occur in the time stream data from each bolometer.
+Fig.~\ref{fig:steps1} shows such steps in the time streams for two
+bolometers. For each bolometer, correcting these steps involves:
+
+\begin{enumerate}
+
+\item Locating each candidate step.
+
+\item Determining the widths of the candidate steps. Some steps
+occur over a single sample such as the step close to sample 5000 in the
+right-hand plot of Fig.~\ref{fig:steps1}, but others happen more
+gradually such as the step close to sample 5300. In addition, a step can
+be preceeded or followed by a short period of instability, as is visible
+at the bottom of the step in the right hand plot of
+Fig.~\ref{fig:steps1}.
+
+\item Determine the height of the step. Steps can vary enormously in height.
+For instance the step in the left plot of Fig.~\ref{fig:steps1} is
+roughly 3 pW in height, whereas the step at sample 5100 in the right hand
+plot is only 0.01 pW in height. The height of a step is taken as the
+difference in the local median value just before and just after the step.
+
+\item Correcting all subsequent data values by the step height.
+
+\item Once all steps have been corrected, a global offset is added to the
+corrected bolometer time stream to ensure that the mean value before and
+after step correction is unchanged.
+
+\end{enumerate}
+
+\paragraph{Detecting Candidate Steps:}
+The first stage in the algorithm is to search the bolometer time stream for
+contiguous blocks of samples that have unusually large mean gradients
+compared to the neighbouring samples. Each such block of samples is
+considered to be a candidate step. The starting and ending time of each
+such block is noted. Once all candidate steps have been found for a
+bolometer, the algorithm proceeds to the next stage. Candidate steps are
+found as follows:
+
+\begin{enumerate}
+
+\item Smooth the bolometer time stream using a median filter of width
+50 samples. This reduces the effects of the noise in the data. Using a
+median filter rather than a mean filter helps to preserve the sharp edges
+at each step, and is robust against spikes, etc.
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{steps2.pdf}
+\caption{A section of the right hand bolometer from  Fig.~\ref{fig:steps1}
+with the median smoothed data overlayed in red.}
+\label{fig:steps2}
+\end{figure}
+
+The red curve in Fig.~\ref{fig:steps2} shows the median smoothed data for
+the right hand bolometer from Fig.~\ref{fig:steps1}.
+
+\item At each time sample, find the difference between the two adjacent
+values in the median filtered bolometer data. This is proportional to the
+gradient of the median filtered bolometer data. Fig.~\ref{fig:steps3}
+shows these differences for the sample bolometer shown in previous
+figures. The high gradients at the three step are clearly visible at
+indices 5040, 5100 and 5280. In addition, the varying gradient of the
+underlying noise-free time series can be seen in the lower level changes
+in the difference values.
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{steps3.pdf}
+\caption{Differences between adjacent values of the median smoothed data
+in  Fig.~\ref{fig:steps2}.}
+\label{fig:steps3}
+\end{figure}
+
+\item Steps are located by looking for difference values that exceed 25
+times the RMS of the neighbouring difference value. However, the RMS of
+the differences can be heavily affected by the gradient of the underlying
+noise-free time series. If there is a strong upward or downward trend in
+the time series data, the RMS of the difference values will be higher
+potentially resulting in some steps being missed. Therefore to enhance
+the reliability of the step detection, an attempt is made to remove these
+variations before locating the steps. To do this, the differences are
+smoothed using mean filter, and the smoothed differences are removed. The
+remaining residual difference values represent the gradient due to steps,
+spikes, noise, etc. It is important that the smoothed differences are not
+polluted by the large values caused by spikes and steps. So before
+smoothing, differences of more than three times the RMS difference (taken
+over the whole time stream) are identified and flagged. The RMS is then
+re-calculated omitting the flagged values and a further two such
+rejection iterations are performed. This flags high gradient values
+caused by spikes, jumps, point sources, etc. The difference values close
+to a step are often not typical of those in the wider neighbourhood, so
+each block of contiguous flagged difference values is doubled in width.
+These flagged values are shown as black in Fig.~\ref{fig:steps4}, with
+the remaining unflagged values shown in red.
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{steps4.pdf}
+\caption{The red curve shows the data from Fig.~\ref{fig:steps3} plotted with smaller upper
+and lower limits. The blue curve shows the results of smoothing the red
+curve. Points omitted from the red curve are shown in black.}
+\label{fig:steps4}
+\end{figure}
+
+\item The differences for the remaining samples are smoothed
+using a mean filter of width 50 samples. The smoothing process fills in small
+holes in the array (up to 40 samples), but will not fill in larger holes.
+Such larger holes are filled in later by linear interpolation. The
+resulting smoothed differences (shown in blue in Fig.~\ref{fig:steps4})
+provide an estimate of the local gradient of the underlying noise-free
+median-filtered data stream.
+
+\item The smoothed differences are subtracted from the total differences
+to get the residual differences caused by noise, spikes, steps, etc.
+These are show in black in Fig.~\ref{fig:steps5}.
+
+\item The local RMS of the residual differences at each time sample is
+found by smoothing the squared residual differences using a median filter
+of width 200 samples, and then taking the square root of the smoothed values.
+These are show in green in Fig.~\ref{fig:steps5}.
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{steps5.pdf}
+\caption{The black curve shows the residuals between the red and blue curves in
+Fig.~\ref{fig:steps4}. Missing data is filled in by a combination of
+smoothing and linear interpolation. The green curve is the local RMS of the data
+over a box width of 200 samples.}
+\label{fig:steps5}
+\end{figure}
+
+\item Blocks of contiguous samples for which the residual difference exceeds
+25 times the local RMS of the residual differences are found. Any two
+such blocks that are separated by fewer than 100 samples are merged into
+a single block. Blocks are ignored if they fail any of the following
+tests:
+
+\begin{enumerate}
+
+\item The width of the block must be less than 200 samples.
+
+\item The total of all the SNR values (the absolute residual difference
+divided by the local RMS of the residual differences) in the block must be
+at least 25.
+
+\item The total of all SNR values in the block must exceed half the value of
+the largest single SNR value in the block.
+
+\end{enumerate}
+
+The second and third tests are intended to prevent spikes being treated as
+two very close steps, since such steps will have SNR values of opposite
+signs and will thus cancel out.
+
+The data in Fig.~\ref{fig:steps5} results in two blocks being defined -
+one between samples 5037 and 5102, and another between samples 5286 and
+5296. Each of these blocks define the start and end of a candidate step.
+
+\end{enumerate}
 

### 2011-11-28 smurf_paper.tex
Move step fixing details to an appendix

@@
 Sudden steps can occur in the time stream data from each bolometer.
-Fig.~\ref{fig:steps1} shows such steps in the time streams for two
-bolometers. For each bolometer, correcting these steps involves:
-
-\begin{enumerate}
-
-\item Locating each candidate step.
-
-\item Determining the widths of the candidate steps. Some steps
+
+\emph{why?}
+
+The black curves in Fig.~\ref{fig:steps1} show examples of such steps in
+the time streams for two bolometers. If not removed, these steps can
+cause severe ringing in the FLT model (Section~\ref{sec:flt}) and visible
+streaks in the final map, corresponding to the paths of individual
+bolometers over the sky.
+
+Steps occur with a wide range of heights and shapes. The ratio of step
+height to noise can vary from less than 10 to several hundred. Some steps
 occur over a single sample such as the step close to sample 5000 in the
@@
 be preceeded or followed by a short period of instability, as is visible
-at the bottom of the step in the right hand plot of
+at the bottom of the step in the left hand plot of
+Fig.~\ref{fig:steps1}. Further problems are caused by steps that occur
+close together in time, such as the large downward step followed by a
+smaller upward step close to sample 5000 in the right hand plot of
 Fig.~\ref{fig:steps1}.
 
-\item Determine the height of the step. Steps can vary enormously in height.
-For instance the step in the left plot of Fig.~\ref{fig:steps1} is
-roughly 3 pW in height, whereas the step at sample 5100 in the right hand
-plot is only 0.01 pW in height. The height of a step is taken as the
-difference in the local median value just before and just after the step.
-
-\item Correcting all subsequent data values by the step height.
-
-\item Once all steps have been corrected, a global offset is added to the
-corrected bolometer time stream to ensure that the mean value before and
-after step correction is unchanged.
+Detecting and correcting such a wide variety of steps reliably has proved
+to be a challenge. The current step-correction algorithm is described in
+detail in appendix~\ref{appendix:steps}. In outline, the following stages
+are involved in detecting and correcting steps in a single bolometer time
+stream:
+
+\begin{enumerate}
+
+\item Median smooth the whole time stream.
+\item Find the gradient of the median smoothed time series at each sample.
+\item Smooth the gradient values to determine the local mean gradient and
+subtract this local mean from the total gradient to get residual gradients.
+\item Find residual gradient values that exceed 25 times the local RMS of
+the residual gradients.
+\item Group these high residual gradients into contiguous blocks of
+samples.
+\item Merge blocks that are separated by less than 100 samples
 
@@
 
-\paragraph{Detecting Candidate Steps:}
-The first stage in the algorithm is to search the bolometer time stream for
-contiguous blocks of samples that have unusually large mean gradients
-compared to the neighbouring samples. Each such block of samples is
-considered to be a candidate step. The starting and ending time of each
-such block is noted. Once all candidate steps have been found for a
-bolometer, the algorithm proceeds to the next stage. Candidate steps are
-found as follows:
-
-\begin{enumerate}
-
-\item Smooth the bolometer time stream using a median filter of width
-50 samples. This reduces the effects of the noise in the data. Using a
-median filter rather than a mean filter helps to preserve the sharp edges
-at each step, and is robust against spikes, etc.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{steps2.pdf}
-\caption{A section of the right hand bolometer from  Fig.~\ref{fig:steps1}
-with the median smoothed data overlayed in red.}
-\label{fig:steps2}
-\end{figure}
-
-The red curve in Fig.~\ref{fig:steps2} shows the median smoothed data for
-the right hand bolometer from Fig.~\ref{fig:steps1}.
-
-\item At each time sample, find the difference between the two adjacent
-values in the median filtered bolometer data. This is proportional to the
-gradient of the median filtered bolometer data. Fig.~\ref{fig:steps3}
-shows these differences for the sample bolometer shown in previous
-figures. The high gradients at the three step are clearly visible at
-indices 5040, 5100 and 5280. In addition, the varying gradient of the
-underlying noise-free time series can be seen in the lower level changes
-in the difference values.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{steps3.pdf}
-\caption{Differences between adjacent values of the median smoothed data
-in  Fig.~\ref{fig:steps2}.}
-\label{fig:steps3}
-\end{figure}
-
-\item Steps are located by looking for difference values that exceed 25
-times the RMS of the neighbouring difference value. However, the RMS of
-the differences can be heavily affected by the gradient of the underlying
-noise-free time series. If there is a strong upward or downward trend in
-the time series data, the RMS of the difference values will be higher
-potentially resulting in some steps being missed. Therefore to enhance
-the reliability of the step detection, an attempt is made to remove these
-variations before locating the steps. To do this, the differences are
-smoothed using mean filter, and the smoothed differences are removed. The
-remaining residual difference values represent the gradient due to steps,
-spikes, noise, etc. It is important that the smoothed differences are not
-polluted by the large values caused by spikes and steps. So before
-smoothing, differences of more than three times the RMS difference (taken
-over the whole time stream) are identified and flagged. The RMS is then
-re-calculated omitting the flagged values and a further two such
-rejection iterations are performed. This flags high gradient values
-caused by spikes, jumps, point sources, etc. The difference values close
-to a step are often not typical of those in the wider neighbourhood, so
-each block of contiguous flagged difference values is doubled in width.
-These flagged values are shown as black in Fig.~\ref{fig:steps4}, with
-the remaining unflagged values shown in red.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{steps4.pdf}
-\caption{The red curve shows the data from Fig.~\ref{fig:steps3} plotted with smaller upper
-and lower limits. The blue curve shows the results of smoothing the red
-curve. Points omitted from the red curve are shown in black.}
-\label{fig:steps4}
-\end{figure}
-
-\item The differences for the remaining samples are smoothed
-using a mean filter of width 50 samples. The smoothing process fills in small
-holes in the array (up to 40 samples), but will not fill in larger holes.
-Such larger holes are filled in later by linear interpolation. The
-resulting smoothed differences (shown in blue in Fig.~\ref{fig:steps4})
-provide an estimate of the local gradient of the underlying noise-free
-median-filtered data stream.
-
-\item The smoothed differences are subtracted from the total differences
-to get the residual differences caused by noise, spikes, steps, etc.
-These are show in black in Fig.~\ref{fig:steps5}.
-
-\item The local RMS of the residual differences at each time sample is
-found by smoothing the squared residual differences using a median filter
-of width 200 samples, and then taking the square root of the smoothed values.
-These are show in green in Fig.~\ref{fig:steps5}.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{steps5.pdf}
-\caption{The black curve shows the residuals between the red and blue curves in
-Fig.~\ref{fig:steps4}. Missing data is filled in by a combination of
-smoothing and linear interpolation. The green curve is the local RMS of the data
-over a box width of 200 samples.}
-\label{fig:steps5}
-\end{figure}
-
-\item Blocks of contiguous samples for which the residual difference exceeds
-25 times the local RMS of the residual differences are found. Any two
-such blocks that are separated by fewer than 100 samples are merged into
-a single block. Blocks are ignored if they fail any of the following
-tests:
-
-\begin{enumerate}
-
-\item The width of the block must be less than 200 samples.
-
-\item The total of all the SNR values (the absolute residual difference
-divided by the local RMS of the residual differences) in the block must be
-at least 25.
-
-\item The total of all SNR values in the block must exceed half the value of
-the largest single SNR value in the block.
-
-\end{enumerate}
-
-The second and third tests are intended to prevent spikes being treated as
-two very close steps, since such steps will have SNR values of opposite
-signs and will thus cancel out.
-
-The data in Fig.~\ref{fig:steps5} results in two blocks being defined -
-one between samples 5037 and 5102, and another between samples 5286 and
-5296. Each of these blocks define the start and end of a candidate step.
-
-\end{enumerate}
+The results of step correction are shown by the red curves in
+Fig.~\ref{fig:steps1}. For comparison, the blue curve shows the
+uncorrected time stream from a nearby bolometer that does not suffer from
+steps. The agreement between the red and blue curves confirms that
+the step correction algorithm is working satisfactorily.
 
@@
 \appendix
-\section[]{This is an appendix}
+\section[]{Step correction in detail}
+
+An outline of the step correction algorithm is given in
+Section~\ref{sec:steps}. This appendix contains a more detailed
+description.
+
+Detecting and correcting steps in a single bolometer time stream involves:
+
+\begin{enumerate}
+
+\item Locating each candidate step.
+
+\item Determining the widths of the candidate steps.
+
+\item Determine the height of the step. Steps can vary enormously in height.
+For instance the step in the left plot of Fig.~\ref{fig:steps1} is
+roughly 3 pW in height, whereas the step at sample 5100 in the right hand
+plot is only 0.01 pW in height. The height of a step is taken as the
+difference in the local median value just before and just after the step.
+
+\item Correcting all subsequent data values by the step height.
+
+\item Once all steps have been corrected, a global offset is added to the
+corrected bolometer time stream to ensure that the mean value before and
+after step correction is unchanged.
+
+\end{enumerate}
+
+\subsection{Detecting candidate steps}
+The first stage in the algorithm is to search the bolometer time stream for
+contiguous blocks of samples that have unusually large mean gradients
+compared to the neighbouring samples. Each such block of samples is
+considered to be a candidate step. The starting and ending time of each
+such block is noted. Once all candidate steps have been found for a
+bolometer, the algorithm proceeds to the next stage. Candidate steps are
+found as follows:
+
+\begin{enumerate}
+
+\item Smooth the bolometer time stream using a median filter of width
+50 samples. This reduces the effects of the noise in the data. Using a
+median filter rather than a mean filter helps to preserve the sharp edges
+at each step, and is robust against spikes, etc.
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{steps2.pdf}
+\caption{A section of the right hand bolometer from  Fig.~\ref{fig:steps1}
+with the median smoothed data overlayed in red.}
+\label{fig:steps2}
+\end{figure}
+
+The red curve in Fig.~\ref{fig:steps2} shows the median smoothed data for
+the right hand bolometer from Fig.~\ref{fig:steps1}.
+
+\item At each time sample, find the difference between the two adjacent
+values in the median filtered bolometer data. This is proportional to the
+gradient of the median filtered bolometer data. Fig.~\ref{fig:steps3}
+shows these differences for the sample bolometer shown in previous
+figures. The high gradients at the three step are clearly visible at
+indices 5040, 5100 and 5280.
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{steps3.pdf}
+\caption{Differences between adjacent values of the median smoothed data
+in  Fig.~\ref{fig:steps2}.}
+\label{fig:steps3}
+\end{figure}
+
+\item Steps are located by looking for difference values that exceed 25
+times the RMS value of the nearby neighbouring differences. However, the
+local RMS of the differences includes the gradient of the underlying
+noise-free time series. So if there is a strong upward or downward trend
+in the time series data, the RMS of the difference values will be higher,
+potentially resulting in some steps being missed. For instance, at sample
+5000 the overall trend in the time series is downwards, resulting in a
+negative bias to the difference values.\footnote{This bias is not visible in
+Fig.~\ref{fig:steps3} because of the large vertical range, but can be seen in
+Fig.~\ref{fig:steps4} which has a smaller vertical scale.}
+
+Therefore to enhance the reliability of the step detection, such biases
+are removed before locating the steps. To do this, the differences are
+smoothed using a mean filter, and the smoothed differences are removed.
+The remaining residual difference values represent the gradient due to
+steps, spikes, noise, etc. It is important that the smoothed differences
+are not polluted by the large values caused by spikes and steps. So
+before smoothing, differences of more than three times the RMS difference
+(taken over the whole time stream) are identified and flagged. The RMS is
+then re-calculated omitting the flagged values and a further two such
+rejection iterations are performed. This process flags high gradient values
+caused by spikes, jumps, point sources, etc. The difference values close
+to a step are often not typical of those in the wider neighbourhood, so
+each block of contiguous flagged difference values is doubled in width.
+These flagged values are shown as black in Fig.~\ref{fig:steps4}, with
+the remaining unflagged values shown in red.
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{steps4.pdf}
+\caption{The red curve shows the data from Fig.~\ref{fig:steps3} plotted with smaller upper
+and lower limits. The blue curve shows the results of smoothing the red
+curve. Points omitted from the red curve are shown in black.}
+\label{fig:steps4}
+\end{figure}
+
+\item The differences for the remaining samples are smoothed
+using a mean filter of width 50 samples. The smoothing process fills in small
+holes in the array (up to 40 samples), but will not fill in larger holes.
+Such larger holes are filled in later by linear interpolation. The
+resulting smoothed differences (shown in blue in Fig.~\ref{fig:steps4})
+provide an estimate of the local gradient of the underlying noise-free
+median-filtered data stream.
+
+\item The smoothed differences are subtracted from the total differences
+to get the residual differences caused by noise, spikes, steps, etc.
+These are show in black in Fig.~\ref{fig:steps5}.
+
+\item The local RMS of the residual differences at each time sample is
+found by smoothing the squared residual differences using a median filter
+of width 200 samples, and then taking the square root of the smoothed values.
+These are show in green in Fig.~\ref{fig:steps5}.
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{steps5.pdf}
+\caption{The black curve shows the residuals between the red and blue curves in
+Fig.~\ref{fig:steps4}. Missing data is filled in by a combination of
+smoothing and linear interpolation. The green curve is the local RMS of the data
+over a box width of 200 samples.}
+\label{fig:steps5}
+\end{figure}
+
+\item Blocks of contiguous samples for which the residual difference exceeds
+25 times the local RMS of the residual differences are found. Any two
+such blocks that are separated by fewer than 100 samples are merged into
+a single block. Blocks are ignored if they fail any of the following
+tests:
+
+\begin{enumerate}
+
+\item The width of the block must be less than 200 samples.
+
+\item The total of all the SNR values (the absolute residual difference
+divided by the local RMS of the residual differences) in the block must be
+at least 25.
+
+\item The total of all SNR values in the block must exceed half the value of
+the largest single SNR value in the block.
+
+\end{enumerate}
+
+The second and third tests are intended to prevent spikes being treated as
+two very close steps, since such steps will have SNR values of opposite
+signs and will thus cancel out.
+
+The data in Fig.~\ref{fig:steps5} results in two blocks being defined -
+one between samples 5037 and 5102 and another between samples 5286 and
+5296. Each of these blocks define the start and end of a candidate step.
+The first block includes two visible steps, but they have been merged
+into a single candidate step due to their small separation (less than 100
+samples).
+
+\end{enumerate}
+
+\subsection{Measuring and confirming candidate steps}
+
 %------------------------------------------------------------------------------

### 2011-11-28 smurf_paper.tex
Complete steps section in main body of paper
The appendix though is not yet complete.

@@
 gradually such as the step close to sample 5300. In addition, a step can
-be preceeded or followed by a short period of instability, as is visible
+be preceded or followed by a short period of instability, as is visible
 at the bottom of the step in the left hand plot of
@@
 detail in appendix~\ref{appendix:steps}. In outline, the following stages
-are involved in detecting and correcting steps in a single bolometer time
-stream:
+are involved in detecting steps in a single bolometer time stream:
 
 \begin{enumerate}
-
 \item Median smooth the whole time stream.
@@
 \item Merge blocks that are separated by less than 100 samples
-
 \end{enumerate}
+
+The above process produces a list of candidate steps in each bolometer
+time series. Each candidate step is then verified, measured and corrected
+using the following procedure:
+
+\begin{enumerate}
+
+\item The above process can mis-interpret the edges of a bright source
+as a step. So we ignore blocks that occur close to bright sources.
+
+\item If the block passes the above test, a least squares linear fit is
+performed to the median-filtered bolometer data just before the block, and
+this fit is extrapolated to predict the data value expected at the centre
+of the block on the basis of the preceding data.
+
+\item A least squares linear fit is performed to the median-filtered
+bolometer data just after the block, and this fit is extrapolated to
+predict the data value expected at the centre of the block on the basis
+of the following data.
+
+\item The difference between these two expected data values is taken as
+the step height.
+
+\item The preceding three steps are repeated several times, each time
+including a different selection of samples in the two least squares fits.
+The mean and standard deviation of the corresponding set of step heights
+is found.
+
+\item If the mean step height is small compared to the standard deviation
+of the step heights, or compared to the noise in the bolometer data, then
+the step is ignored.
+
+\item If the above checks are passed, all subsequent bolometer samples are
+corrected by the mean step height.
+
+\item Bolometer samples within the duration of the step, and a few
+samples on either side, are flagged as unusable.
+
+\end{enumerate}
+
+Once all steps have been corrected within a bolometer time stream, a
+constant value is added to all samples in the time stream to restore its
+original mean value.
 

### 2011-11-29 smurf_paper.tex
Gap filling, apodisation and padding

@@
 
-DSB
-
-Discuss gap filling, ``new'' window function method, also apodization
-and start/end polynomial continuity stuff. Should reference CMB
-literature -- in particular look at \citet{stompor2002}.
+The evaluation of the FLT model (Section~\ref{sec:flt}) involves taking
+the Fast Fourier Transform of each bolometer time series. Data that has
+been flagged as bad for any reason (for instance, due to the presence of
+spikes, steps, or unusual common-mode signal) needs to be excluded from
+this FFT. Two method are available:
+
+\begin{itemize}
+\item Gap-filling: each contiguous block of bad data samples is filled
+with artificial data before taking the FFT. A least squares linear fit is
+performed to the 50 samples preceding the block, and a similar fit is
+performed to the 50 samples following the block. These are used to
+estimate the expected values at the start and end of the block of bad
+values. The bad values are then replaced by linear interpolation between
+the expected start and end values. Gaussian noise is added with a
+standard deviation equal to the mean of the RMS residuals in the two
+fits.
+\item Masking: each bad data value is replaced by zero, and an associated
+mask is created that holds 1.0 for each good sample in the time stream and
+0.0 for each bad sample. The masked data and the mask are then each filtered
+in the same way using FFTs, and the filtered data is finally normalised by
+dividing it by the filtered mask.
+\end{itemize}
+
+In addition to replacing bad samples before the FFT, it is also necessary
+to ensure that the data values at the start and end of the time series
+are similar. Since a Fast Fourier transform treats the data as a single
+cycle in an infinitely repeating waveform, any large difference between
+starting and ending value will effectively introduce sudden steps at the
+start and end of each cycle, causing unwanted oscillations (ringing) in
+the transform. Another consequence of the cyclic nature of the FFT is
+that features at one end of the time series can affect the filtered
+values at the other end of the time series. Two methods are available to
+avoid these two problems:
+
+\begin{itemize}
+\item Apodisation: a number of samples at the start and end of each
+bolometer time series are multiplied by a cosine function in order to
+roll the data values off smoothly to zero. The default number of samples
+modified at each end of the time series is given by half the ratio of the
+sampling frequency to the lowest retained frequency. In addition, each
+end of the time stream is padded with double this number of zeros. This
+method is illustrated in Fig.~\ref{fig:pad2}.
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{pad2.pdf}
+\caption{The black curve shows a bolometer time stream, padded
+with zeros. The red curve shows the time stream after apodisation.}
+\label{fig:pad2}
+\end{figure}
+
+\item Padding with artificial data: instead of padding with zeros, each
+time stream is padded with artificial data that connects the two ends of
+the data stream smoothly and includes Gaussian noise. No apodisation is
+performed. The number of samples of padding at each end is again equal to
+the ratio of the sampling frequency to the lowest retained frequency.
+This is illustrated in Fig.~\ref{fig:pad1}.
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{pad1.pdf}
+\caption{The black curve shows the same bolometer time stream as in
+Fig.~\ref{fig:pad2}, againpadded with zeros. The red curve shows the time stream
+after padding with artificial data.}
+\label{fig:pad1}
+\end{figure}
+
+\end{itemize}
+
+\emph{Should reference CMB literature -- in particular look at
+\citet{stompor2002}}.
 

### 2011-12-07 smurf_paper.tex
define macro \model for displaying model component names

@@
           \hbox{$>$}\hbox{$\sim$}}}}
+
+% how are model names displayed
+\newcommand{\model}[1]{\texttt{#1}}
 
@@
     \scuba\ time-series data with SMURF. Only the first group of
-    models are typically fit to the data (\texttt{COM}--\texttt{NOI})
+    models are typically fit to the data (\model{COM}--\model{NOI})
     in the indicated order. The remaining models
-    (\texttt{DKS}--\texttt{PLN}) have only been included for
+    (\model{DKS}--\model{PLN}) have only been included for
     completeness}
@@
     \hline
-    \texttt{COM} & remove common-mode signal \\
-    \texttt{GAI} & common-mode scaled to each bolometer \\
-    \texttt{EXT} & extinction correction \\
-    \texttt{FLT} & Fourier transform filter \\
-    \texttt{AST} & map estimate of astronomical signal \\
-    \texttt{NOI} & noise estimation \\
+    \model{COM} & remove common-mode signal \\
+    \model{GAI} & common-mode scaled to each bolometer \\
+    \model{EXT} & extinction correction \\
+    \model{FLT} & Fourier transform filter \\
+    \model{AST} & map estimate of astronomical signal \\
+    \model{NOI} & noise estimation \\
     \hline
-    \texttt{DKS} & dark squid cleaning along columns \\
-    \texttt{SMO} & time-domain smoothing filter \\
-    \texttt{PLN} & 2-dimensional time-varying plane removal \\
+    \model{DKS} & dark squid cleaning along columns \\
+    \model{SMO} & time-domain smoothing filter \\
+    \model{PLN} & 2-dimensional time-varying plane removal \\
     \hline
@@
 
-\subsubsection{\texttt{COM,GAI}: common-mode estimation}
+\subsubsection{\model{COM,GAI}: common-mode estimation}
 \label{sec:comgai}
@@
 
-\subsubsection{\texttt{EXT}: extinction correction}
+\subsubsection{\model{EXT}: extinction correction}
 \label{sec:ext}
@@
 
-\subsubsection{\texttt{FLT}: fourier transform filter}
+\subsubsection{\model{FLT}: fourier transform filter}
 \label{sec:flt}
@@
 to preserve the white noise level).  The signal that is \emph{removed}
-from the time-series by this process are stored in the \texttt{FLT}
+from the time-series by this process are stored in the \model{FLT}
 container array.
 
-\subsubsection{\texttt{AST}: map estimation}
+\subsubsection{\model{AST}: map estimation}
 \label{sec:ast}
@@
 
-In addition to map estimation, the \texttt{AST} model can also be used
+In addition to map estimation, the \model{AST} model can also be used
 to perform map-based despiking of the time-series
@@
 
-\subsubsection{\texttt{NOI}: noise estimation}
+\subsubsection{\model{NOI}: noise estimation}
 \label{sec:noi}
 
-The primary purpose of the noise component, \texttt{NOI} is to measure
+The primary purpose of the noise component, \model{NOI} is to measure
 the noise in each bolometer residual once all of the other models have
@@
 the cleanest data for which some cleaning operations may be
-performed. For this reason \texttt{NOI} can optionally use the
+performed. For this reason \model{NOI} can optionally use the
 time-domain DC step fixer (Section~\ref{sec:steps}) and spike
@@
 
-\subsubsection{\texttt{DKS}: dark squid cleaning along columns}
+\subsubsection{\model{DKS}: dark squid cleaning along columns}
 \label{sec:dks}
@@
 working bolometers also has a working dark squid. An alternative is to
-use the \texttt{TMP} model described below.
+use the \model{TMP} model described below.
 
@@
 
-\subsubsection{\texttt{TMP}: remove a scaled template}
-
-Since the \texttt{DKS} cannot generally be used due to the missing
+\subsubsection{\model{TMP}: remove a scaled template}
+
+Since the \model{DKS} cannot generally be used due to the missing
 dark squids, an alternative is to use the telescope pointing as a
@@
 
-\subsubsection{\texttt{SMO}: time-domain smoothing filter}
+\subsubsection{\model{SMO}: time-domain smoothing filter}
 \label{sec:smo}
@@
 
-\subsubsection{\texttt{PLN}: 2d time-varying sky noise plane removal}
+\subsubsection{\model{PLN}: 2d time-varying sky noise plane removal}
 \label{sec:pln}

### 2011-12-19 smurf_paper.tex
updates to model sections

@@
 the time-series), there are a number of parameters that control their
-operation. Table~\ref{tab:components} list the model components and
+operation. Table~\ref{tab:components} lists the model components and
 the typical order in which they are calculated. The following sections
@@
     \scuba\ time-series data with SMURF. Only the first group of
-    models are typically fit to the data (\model{COM}--\model{NOI})
-    in the indicated order. The remaining models
-    (\model{DKS}--\model{PLN}) have only been included for
+    models (in the order indicated) are typically fit to the data
+    (\model{COM}--\model{NOI}) in the indicated order. The remaining
+    models (\model{DKS}--\model{PLN}) have only been included for
     completeness}
@@
 for all but the brightest sources, and swamps faint extended structure.
-The purpose of the COM and GAI models is to remove this common-mode
+The purpose of the \model{COM} and \model{GAI} models is to remove this common-mode
 signal.
@@
 
-The COM model is the common-mode signal itself. It is a single time
+The \model{COM} model is the common-mode signal itself. It is a single time
 series that is estimated by finding the mean of all bolometers at each
@@
 common-mode variations will also vary from bolometer to bolometer. So
-comparing each bolometer time stream with the common-mode signal allows
-an estimate of the bolometer sensitivity to be obtained. In practice, a
-least squares linear fit is performed between the bolometer time series
-and the common-mode to determine a gain and offset for each bolometer.
-The gain and offset for each bolometer is known as the GAI model. Each
-bolometer value can then optionally be scaled and shifted using these
-values so that all bolometers share a common (but as yet unknown)
-calibration. This provides an alternative( or additional) flatfielding
-strategy to that described in in section~\ref{sec:bolos}.
+comparing each bolometer time stream with the common-mode signal
+allows an estimate of the bolometer sensitivity to be obtained. In
+practice, a least squares linear fit is performed between the
+bolometer time series and the common-mode to determine a gain and
+offset for each bolometer.  The gain and offset for each bolometer is
+known as the \model{GAI} model. Each bolometer value can then
+optionally be scaled and shifted using these values so that all
+bolometers share a common (but as yet unknown) calibration. This
+provides an alternative( or additional) flatfielding strategy to that
+described in in section~\ref{sec:bolos}.
 
@@
 emission is indistinguishable from the other sources of common-mode
-signal (fridge and atmosphere variations) and will be removed by the COM
+signal (fridge and atmosphere variations) and will be removed by the 
 model. This places a limit on the spatial extent of astronomical
@@
 
-For this reason, the usual practice is to estimate a single COM model by
-examining data from all four sub-arrays in each waveband, since this
-allow spatial structure on the scale of the whole focal plane to be
-recovered. However, sometimes there is evidence that the common-mode
-differs from one array to another, and so an option exists to estimate a
-separate COM model for each individual subarray, with a consequent
-lowering in the scale of spatial structure that can be recovered.
+For this reason, the usual practice is to estimate a single
+\model{COM} model by examining data from all four sub-arrays in each
+waveband, since this allow spatial structure on the scale of the whole
+focal plane to be recovered. However, sometimes there is evidence that
+the common-mode differs from one array to another, and so an option
+exists to estimate a separate \model{COM} model for each individual
+subarray, with a consequent lowering in the scale of spatial structure
+that can be recovered.
 
@@
 
-The extinction correction is normally derived using the JCMT water
-vapour radiometer (WVM), and is not considered to be a free parameter
-in the solution. For details on how it is calculated see
+The extinction correction is a multiplicative factor that is normally
+derived using the JCMT water vapour radiometer (WVM), and is not
+considered to be a free parameter in the solution. However, it is
+applied as part of the iterative solution, rather than a
+pre-processing step, since any small errors will be amplified by the
+large low-frequency drifts in the raw bolometer time-series. For
+example, if the bolometer drift is 1000 times greater than an
+astronomical source of interest, a 1 per cent error in the flatfield
+will produce stripes of order 10 times the astronomical signal
+amplitude in the final map! If \model{EXT} is applied after the bulk
+of the low-frequency noise has been removed (e.g., \model{COM,GAI}),
+then there is little potential for such small errors to affect the
+final map.  For details on how it is calculated see
 \citet{dempsey2011}. Note that its numerical value is calculated only
@@
 from the time-series by this process are stored in the \model{FLT}
-container array.
+container array. Typically this model is used purely as a high-pass
+filter to remove most of the $1/f$ noise. Low-pass filtering is
+redundant for two reasons: (i) SMURF already low-pass filters and
+re-samples to a lower sample rate as described in
+Section~\ref{sec:downsamp}; and (ii) the act of re-gridding the data
+to produce map estimates effectively low-pass filters the data at a
+frequency that corresponds to the inverse of the crossing time of a
+single map pixel. Notch filters have not been used with \scuba\ data
+particularly since line features tend to move around (i.e. a dynamic
+line-detection system would have to be developed, and care would have
+to be taken that astronomical sources are not suppressed).
 
@@
 
-Map estimation is accomplished using a simple nearest-neighbour
-resampling of the data onto a pre-defined map grid. For the $i$th map
-pixel $\mathbf{m}(x_i,y_i)$, the brightness is estimated as the
-weighted average of the bolometer data samples $\mathbf{b}_j$ that
-land within that pixel (from any bolometer or point in time),
+Map estimation is accomplished using a nearest-neighbour resampling of
+the data onto a pre-defined map grid. For the $i$th map pixel
+$\mathbf{m}(x_i,y_i)$, the brightness is estimated as the weighted
+average of the bolometer data samples $\mathbf{b}_j$ that land within
+that pixel (from any bolometer or point in time),
 %
@@
 where $N_j$ is the total number of bolometer samples that land in the
-pixel. We decided not to use the \emph{unbiased} estimator since in
+pixel. As written, this algorithm is numerically unstable if the two
+terms in the numerator are large, and of nearly the same value:
+floating point precision errors can cause the difference to be
+significantly incorrect. In practice we use the more numerically
+stable ``weighted incremental algorithm'' that calculates incremental
+differences, as described in \citet{west1979}.
+
+ We decided not to use the \emph{unbiased} estimator since in
 practice it would require accumulating an additional array of values
@@
 \subsubsection{Down-sampling}
+\label{sec:downsamp}
 

### 2011-12-22 smurf_paper.tex
smurf: add some key figures illustrating map degeneracy/constraint concepts

@@
 
+In this section three different types of data are reduced. In each
+case, the baseline reduction using defaults as described in
+Section~\ref{sec:algorithm} are used first. Then, modifications are
+described to improve each reduction.
+
 \subsection{Point source}
+In this example we produce a map of Uranus, which is a nearly
+point-like source for SCUBA-2 that is commonly used as a primary flux
+calibrator (observation 26 on 20111017). The first reduction uses the
+defaults as described in Section~\ref{sec:algorithm}, but we manually
+set the number of iterations to a fixed value of 100 to illustrate how
+the solution converges in the absence of any boundary conditions.
+
 Reduce a calibrator. Describe the use of a zero mask around the
@@
 constrain a maximum-likelihood map with poor cross-linking.
+
+\begin{figure*}
+\centering
+\includegraphics[width=\linewidth]{pointmaps.pdf}
+\caption{Multiple reductions of a 450\,\micron\ daisy scan of Uranus,
+  all scaled between -0.002\, pW (white) and +0.002\,pW (black): (a)
+  Reduction in which only common-mode subtraction is used to suppress
+  low-frequency noise, and the reduction is forced to use 2 iterations
+  (after the first iteration an estimate of the source flux is
+  removed, and the noise is measured in the residuals to obtain
+  appropriate weighting for the second and final iteration) ---
+  circular streaks are caused by independent low-frequency noise that
+  is not removed by the common-mode; (b) same as (a) but using 50
+  iterations, illustrating that there is a degeneracy between
+  large-scale structure and the common-mode (the footprint of working
+  bolometers is also shown for reference); (c) reduction in which
+  high-pass filtering above 0.775\,Hz (corresponding to a spatial
+  scale of 200\,arcsec as indicated) is applied after common-mode
+  removal, but before the map estimate --- the map-based convergence
+  test is activated and the solution halts after 10 iterations, but
+  leaving large-scale ringing due to the filter; (d) same as (c), but
+  now the region of the map beyond the white dot-dashed circle is
+  constrained to a value of zero until all but the final iteration ---
+  the map is now extremely flat, there is little attenuation of the
+  source flux, and the diffraction pattern is clearly seen.}
+\label{fig:pointmaps}
+\end{figure*}
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{degeneracy.pdf}
+\caption{Demonstration of the degeneracy between large-scale
+  structures in the map and common-mode removal, corresponding to
+  panels (a) and (b) in Fig.~\ref{fig:pointmaps}. The black and grey
+  lines in the top panel show the residual signal for a single
+  bolometer after 2 and 50 iterations respectively (they are nearly
+  identical). The green and red lines show the \emph{difference}
+  between the \model{COM} and \model{AST} (the signal produced by the
+  current map estimate for a given bolometer) model components for
+  that bolometer between iterations 2 and 50 respectively. This shows
+  that a strong signal has grown over time, and it has equal but
+  opposite signs in the two model components, so that they cancel one
+  another. The bottom panel shows the scan pattern of the telescope
+  over the same period; clearly the \model{COM}/\model{AST} model
+  degeneracy is correlated with the elevation offset, and referring to
+  Fig.~\ref{fig:pointmaps}, panel (b), this clearly corresponds to the
+  strong vertical gradient that has appeared}.
+\label{fig:degeneracy}
+\end{figure}
 

### 2012-01-11 smurf_paper.tex
add important map-making ref + downsampling section draft

@@
 out of techniques that were developed to produce maps of the
-large-scale anisitropies in the Cosmic Microwave Background (CMB).  Styles
-of map-making: maximum likelihood \citep[e.g.,][]{patanchon2008};
-iterative -- see \citet{johnstone2000} implementation of
-\citet{wright1996} to remove chop, \citet{kovacs2008} for SHARC-2
-etc.; de-correlation using PCA analysis
-\citep[e.g.][]{laurent2005,scott2008,aguirre2010}.
+large-scale anisitropies in the Cosmic Microwave Background (CMB).
+Styles of map-making: least squares \citep{janssen1992} maximum
+likelihood \citep[e.g.,][]{patanchon2008}; iterative -- see
+\citet{johnstone2000} implementation of \citet{wright1996} to remove
+chop, \citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA
+analysis \citep[e.g.][]{laurent2005,scott2008,aguirre2010}.
 
@@
 
-\subsubsection{Down-sampling}
+\subsubsection{Time-series down-sampling and map pixel size}
 \label{sec:downsamp}
 
-EC
+The highest useful frequency in the nominally 200\,Hz-sampled
+\scuba\ data is that which corresponds to the smallest angular scale
+that the instrument is sensitive to. As mentioned earlier, the usual
+rule-of-thumb for a Gaussian beam is to provide at least 3 samples for
+each FWHM, or roughly 2.5\,arcsec for the 7.5\,arcsec
+450\,\micron\ channel, and 5\,arcsec for the 14.5\,arcsec
+850\,\micron\ channel. For a typical scan speed of
+300\,arcsec\,s$^{-1}$, the maximum useful sample rate is therefore
+about 120\,Hz at 450\,\micron, and 40\,Hz at 850\,\micron. In order to
+save execution time and memory usage (both of which scale linearly
+with data volume), it is clearly advantageous to re-sample the data to
+these lower rates.
+
+Similarly, from the point of view of retaining all of the useful
+information, the size of the output map pixels need not be any smaller
+than the FWHM/3 rule. Larger pixels may be selected, but at the
+expense of retaining information at the diffraction limit of the
+telescope.
+
+The default behaviour of SMURF is to select output map pixel sizes of
+2 and 4\,arcsec at 450 and 850\,\micron, respectively (slightly
+over-sampled). The slew speed of the telescope (approximately constant
+for all standard scan modes) is then used to calculate the highest
+useful frequency for the time-series given the map pixel size,
+$f_\mathrm{max}$. The data are then re-sampled to this new rate prior
+to map-making (as well as the data cleaning steps).
+
+The method used by SMURF to down-sample simply averages together
+multiple samples from the original time-series, $x_i$ to estimate the
+lower-frequency output time-series, $y_j$. In general there will not
+be an integer number of samples from $x_i$ in each of the $y_j$, so
+fractional samples from $x_i$ are used at the $y_j$ time
+boundaries. The algorithm is fast since each sample in $x_i$ need only
+be visited once. From a spectral point of view, this boxcar average is
+equivalent to applying a sinc-function low-pass filter in frequency
+space. Such behaviour is desirable since it serves as an anti-aliasing
+filter; most non-white features above the Nyquist frequency in $y_j$
+that could contaminate the output data are largely removed.
+
+As an alternative, we also investigated an algorithm in which the FFT
+of $x_i$ is simply truncated to the target sample rate before
+transforming back to the time domain (i.e., applying a hard-edged
+low-pass filter). In practice the noise performance was
+indistinguishable, and required slightly more execution time, using
+this method compared to the boxcar method described above. It is
+therefore not currently available to users.
 

### 2012-01-21 smurf_paper.tex
write text for point sources, add scripts/figures for cosmology reduction

@@
           \hbox{$>$}\hbox{$\sim$}}}}
+
+% sinc function
+\def\sinc{\mathrm{sinc}}
 
@@
 \subsection{Point source}
-In this example we produce a map of Uranus, which is a nearly
-point-like source for SCUBA-2 that is commonly used as a primary flux
-calibrator (observation 26 on 20111017). The first reduction uses the
-defaults as described in Section~\ref{sec:algorithm}, but we manually
-set the number of iterations to a fixed value of 100 to illustrate how
-the solution converges in the absence of any boundary conditions.
-
-Reduce a calibrator. Describe the use of a zero mask around the
-source. Compare with \citet{wiebe2009} who used the same method to
-constrain a maximum-likelihood map with poor cross-linking.
+
+The accurate measurement of positions and brightnesses of point
+sources are necessary in real-time to establish telescope pointing
+offsets and focus. They are also necessary to measure the FCF
+(absolute calibration), and hence noise performance of the instrument
+in astronomical units.  In this example we produce a 450\,\micron\ map
+of Uranus (observation 26 on 20111017), which is a nearly point-like
+source for SCUBA-2 that is commonly used as a primary flux
+calibrator. The now standard CV daisy scan pattern was used, with a
+scan speed of 155 arcsec\,sec$^{-1}$. We perform several different
+reductions of the data to illustrate the purpose of different model
+components and the convergence properties of the solution
+(Fig.~\ref{fig:pointmaps}). In all cases the maps are produced on a
+grid of azimuth (horizontal) and elevation (vertical) offsets from the
+position of Uranus (the origin), using 2\,arcsec pixels.
+
+The first, simplest reduction of the data uses only the \model{COM}
+model to estimate and remove the common-mode signal in order to
+suppress low-frequency noise in the data. After \model{COM}, the
+extinction correction is applied (\model{EXT}), and an initial map is
+estimated using equal weighting for all of the detectors. This
+estimate of \model{AST} is then removed from the data, and the noise
+is measured in the residuals to estimate weights for the subsequent
+and final iteration. The resulting map after these two iterations is
+shown in Fig.~\ref{fig:pointmaps}a. While the S/N of Uranus is clearly
+large ($\sim$175 as estimated by SMURF), enabling us to see the faint
+sidelobes (circle and cross pattern), the map also has obvious
+circular streaks and other large-scale ripples. The circular streaks
+are due to the fact that \model{COM} does not account for all of the
+low-frequency noise (see Fig.~\ref{fig:pspec}), and therefore each
+bolometer leaves a trace of the circular scan pattern in the map as
+each of their baselines slowly drift independently. Most of the
+larger-scale ripples in the map, however, are probably due to
+degeneracies in the map solution.
+
+To illustrate how large-scale ripples can form (and in fact, diverge),
+the same map solution is run for 50 iterations and shown in
+Fig.~\ref{fig:pointmaps}b, now exhibiting a strong vertical
+gradient. The degeneracy is easy to understand if the time-domain
+behaviour of each model component is considered. The top panel of
+Fig.~\ref{fig:degeneracy} shows the residual signals for a single
+bolometer after 2 (black) and 50 (grey) iterations which are nearly
+identical, yet the \emph{change} in the estimated \model{COM} (green)
+and \model{AST} (red) signals between 2 and 50 iterations are
+large. However, it is also clear that the estimated \model{COM} and
+\model{AST} signals are \emph{complimentary}. In other words, the large
+change in the \model{AST} signal is cancelled by freedom in the
+\model{COM} signal to grow with opposite sign. For comparison, the
+bottom panel of Fig.~\ref{fig:degeneracy} shows the pointing solution
+for this section of data, and the shapes of the \model{AST} and
+\model{COM} signals match the elevation component, which is aligned
+with the gradient in Fig.~\ref{fig:pointmaps}b. Generically, the
+calculation of \model{COM} will remove any information on angular
+scales that are larger than the array footprint (outline shown in
+Fig.~\ref{fig:pointmaps}b for reference), meaning that the map
+solution is unconstrained on such large scales.
+
+One simple improvement is to simply restrict the low-frequency signal
+that goes into the maps using a high-pass filter. A third reduction of
+the data uses the default map-making parameters, as described in
+Section~\ref{sec:algorithm}, which adds the \model{FLT} model to
+accomplish this task immediately prior to map estimation. We set the
+filter edge based on an angular scale of 200\,arcsec, which given the
+scan speed of 155\,arcsec\,sec$^{-1}$, corresponds to a frequency of
+0.78\,Hz. Now using the automatic map-based convergence test (Section
+???), the resulting map converges after 10 iterations
+(Fig.~\ref{fig:pointmaps}c). Both the circular streaks and the
+large-scale gradient are now removed, but they have been replaced by
+an obvious, circularly-symmetric ringing pattern about Uranus. The
+reason for this ringing is due to the fact that the hard-edged
+high-pass filter in frequency space is equivalent to a $\sinc$
+function-like response in map space. Since the scan pattern is
+isotropic (scans equally at all position angles), and there is a
+bright point-like source, the result is an azimuthally-symmetric
+$\sinc$ function-like pattern in the map about Uranus. Note that due
+to the iterative nature of the solution, much of the signal in the
+source is actually removed for subsequent iterations of
+\model{FLT}, but as in the previous case, generacies between the
+map and the \model{FLT} and \model{COM} signal components can still
+lead to complicated structures.
+
+The preceding examples illustrate the need for prior constraints in
+the map solution. For calibrators (and other previously known bright,
+compact sources), a good, simple prior is to constrain the map to a
+value of zero away from the known locations of emission. In
+Fig.~\ref{fig:pointmaps}d, a map is produced in an identical manner to
+the previous example, but now setting the map explicitly to zero
+beyond a radius of 60\,arcsec from the location of Uranus (much larger
+than the FWHM of the main lobe), for all but the final iteration. In
+this case, the map converges after 22 iterations, and all of the
+previous artifacts have been effectively removed. Since the map is now
+flat away from the source, and constrained to a value of zero, it is
+now appropriate for performing aperture photometry directly, with no
+need for an additional reference annulus. This approach to constrained
+map-making is similar to that employed for poorly cross-linked scans
+of compact (though resolved) sources by \citet{wiebe2009} using BLAST
+data.
 
@@
 \caption{Multiple reductions of a 450\,\micron\ daisy scan of Uranus,
-  all scaled between -0.002\, pW (white) and +0.002\,pW (black): (a)
+  all scaled between -0.002\,pW (white) and +0.002\,pW (black): (a)
   Reduction in which only common-mode subtraction is used to suppress
@@
   circular streaks are caused by independent low-frequency noise that
-  is not removed by the common-mode; (b) same as (a) but using 50
-  iterations, illustrating that there is a degeneracy between
-  large-scale structure and the common-mode (the footprint of working
-  bolometers is also shown for reference); (c) reduction in which
-  high-pass filtering above 0.775\,Hz (corresponding to a spatial
-  scale of 200\,arcsec as indicated) is applied after common-mode
-  removal, but before the map estimate --- the map-based convergence
-  test is activated and the solution halts after 10 iterations, but
-  leaving large-scale ringing due to the filter; (d) same as (c), but
-  now the region of the map beyond the white dot-dashed circle is
+  is not removed by the common-mode (Uranus peak 0.14\,pW); (b) same
+  as (a) but using 50 iterations, illustrating that there is a
+  degeneracy between large-scale structure and the common-mode (the
+  footprint of working bolometers is also shown for reference, Uranus
+  peak 0.14\,pW); (c) reduction in which high-pass filtering above
+  0.775\,Hz (corresponding to a spatial scale of 200\,arcsec as
+  indicated) is applied after common-mode removal, but before the map
+  estimate --- the map-based convergence test is activated and the
+  solution halts after 10 iterations, but leaving large-scale ringing
+  due to the filter (Uranus peak 0.38\,pW); (d) same as (c), but now
+  the region of the map beyond the white dot-dashed circle is
   constrained to a value of zero until all but the final iteration ---
   the map is now extremely flat, there is little attenuation of the
-  source flux, and the diffraction pattern is clearly seen.}
+  source flux, and the diffraction pattern is clearly seen (Uranus
+  peak 0.38\,pW).}
 \label{fig:pointmaps}
@@
 \subsection{Deep blind survey}
-Heavy initial filtering, and only iteratively solve for map and
-common-mode with edge constraints.
-
-\textit{And matched filter is discussed here or superceded by
-the 2D FFTs?}
+
+\scuba\ surveys designed to detect extremely faint point-sources
+(e.g. high-redshift star-forming galaxies, and debris disks) need to
+obtain the best noise performance possible. The approach described
+here involves three major steps: (i) generating a map that removes
+most low-frequency noise sources; (ii) \emph{whitening} (essentially
+flattening) the map to make the noise properties of the map easier to
+deal with; and (iii) detecting point sources using a
+\emph{matched-filter}. Note that variations on this general procedure
+have been used extensively in the submm cosmology community using
+previous instruments \citep[e.g.][]{scott2008,perera2008}.  In this
+section we will reduce data of the Lockman Hole taken in 2010 during
+S2SRO as a pilot project for the Cosmology Legacy Survey. It consists
+of $\sim8.5$\,hours of data taken in 36 separate scans (average
+15\,min. each) spread over February and March, covering an area of
+about 50\,arcmin$^2$. ... full obs list ... dropped scans due to
+terrible noise properties ...
+
+The first step, map generation, is accomplished in a way similar to
+the default described in the previous section (without any boundary
+constraints)...
+
+Even though the map \emph{looks} extremely flat, it is clear that
+there is a mixture of faint astronomical sources, and what is probably
+residual low-frequency noise causing faint ripples. With uncertainties
+about what is source, and what is an artifact, the first step is to
+remove the low-frequency noise in a way that does not affect
+(appreciably) the real sources. To accomplish this goal, the map is
+run through a whitening filter that results in a flat angular noise
+power spectrum. The noise power spectrum is estimated from a
+\emph{jackknife map}: maps are produced from two independent halves of
+the total data set, and then a pixel-by-pixel difference is
+calculated. Provided that the noise in one half of the data is
+uncorrelated with that from the other half, the noise in the jackknife
+map should resemble that of the total map, though larger by a factor
+of $\sim\sqrt(2)$ depending on how closely the pixel variances in the
+two maps are matched (if one map pixel is significantly noisier, it
+will dominate the noise in the jackknife).  The astronomical signal,
+however, should be cleanly removed (provided that there are no strong
+time-varying signals, and also assuming that errors due to calibration
+between the two halves are insignificant). The approach we have taken
+to minimize systematics is to produce two halves of the jackknife
+using odd and even scan numbers (i.e., each map will contain an even
+sampling of data across the full dataset).
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{lockman_rawmaps.pdf}
+\caption{Procedure for producing maps of deep fields.}
+\label{fig:lockman_rawmaps}
+\end{figure}
+
+Since the \scuba\ scan strategy is usually isotropic (all position
+angles scanned with roughly equal weights), we make the simplifying
+assumption that the angular noise power spectrum is azimuthally
+symmetric (inspection of the 2-dimensional FFT confirms
+this). ... calculate filter as slightly-smoothed complement of
+azavpspec ... Finally, the filter is applied to the map, resulting in
+panel...
+
+An additional test that can be accomplished with the jackknife map, is
+the accuracy of the individual map-pixel variances. If the calculation
+is correction, the variance in each jackknife pixel,
+$\sigma^2_\mathrm{j}$, is simply the sum of the variances in the even
+and odd half-maps, $\sigma^2_\mathrm{e} + \sigma^2_\mathrm{o}$. If a
+S/N map is made from the jackknife (i.e., normalizing by
+$\sigma_\mathrm{j}$, the expected distribution is simply a Gaussian
+with mean zero, and unit standard deviation. A histogram of the
+central region of this map is shown in Fig. ???
+
+... fig. showing S/N histogram of the jackknife...
+
+Since the
+
+Both map-making itself, and the whitening filter, will suppress (and
+add sidelobes), to the nominal PSF for a point source.
 

### 2012-01-25 smurf_paper.tex
added in an old reference that might be relevant
more plots for cosmology section
flesh out text for cosmology section

@@
 than the fully-commissioned instrument in 2011.
+
+Another paper worth citing? \citet{weferling2002}
 
@@
 
-\subsection{Point source}
-
-The accurate measurement of positions and brightnesses of point
+\subsection{Known point source}
+\label{sec:point}
+
+The accurate measurement of positions and brightnesses of known point
 sources are necessary in real-time to establish telescope pointing
@@
 larger-scale ripples in the map, however, are probably due to
-degeneracies in the map solution.
-
-To illustrate how large-scale ripples can form (and in fact, diverge),
-the same map solution is run for 50 iterations and shown in
+degeneracies in the map solution as we discuss below.
+
+To illustrate how large-scale ripples can form (and grow), the same
+map solution is run for 50 iterations and shown in
 Fig.~\ref{fig:pointmaps}b, now exhibiting a strong vertical
@@
 large. However, it is also clear that the estimated \model{COM} and
-\model{AST} signals are \emph{complimentary}. In other words, the large
-change in the \model{AST} signal is cancelled by freedom in the
+\model{AST} signals are \emph{complimentary}. In other words, the
+large change in the \model{AST} signal is cancelled by freedom in the
 \model{COM} signal to grow with opposite sign. For comparison, the
@@
 
-One simple improvement is to simply restrict the low-frequency signal
-that goes into the maps using a high-pass filter. A third reduction of
-the data uses the default map-making parameters, as described in
-Section~\ref{sec:algorithm}, which adds the \model{FLT} model to
-accomplish this task immediately prior to map estimation. We set the
-filter edge based on an angular scale of 200\,arcsec, which given the
-scan speed of 155\,arcsec\,sec$^{-1}$, corresponds to a frequency of
-0.78\,Hz. Now using the automatic map-based convergence test (Section
-???), the resulting map converges after 10 iterations
-(Fig.~\ref{fig:pointmaps}c). Both the circular streaks and the
-large-scale gradient are now removed, but they have been replaced by
-an obvious, circularly-symmetric ringing pattern about Uranus. The
-reason for this ringing is due to the fact that the hard-edged
+One improvement is to simply restrict the low-frequency signal that
+goes into the maps by applying a high-pass filter to the bolometer
+time-series. A third reduction of the data uses the default map-making
+parameters, as described in Section~\ref{sec:algorithm}, which adds
+the \model{FLT} model to accomplish this task immediately prior to map
+estimation. We set the filter edge based on an angular scale of
+200\,arcsec, which given the scan speed of 155\,arcsec\,sec$^{-1}$,
+corresponds to a frequency of 0.78\,Hz. Now using the automatic
+map-based convergence test (Section ???), the resulting map converges
+after 10 iterations (Fig.~\ref{fig:pointmaps}c). Both the circular
+streaks and the large-scale gradient are now removed, but they have
+been replaced by an obvious, circularly-symmetric ringing pattern
+about Uranus. The reason for this ringing is that the hard-edged
 high-pass filter in frequency space is equivalent to a $\sinc$
@@
 to the iterative nature of the solution, much of the signal in the
-source is actually removed for subsequent iterations of
-\model{FLT}, but as in the previous case, generacies between the
-map and the \model{FLT} and \model{COM} signal components can still
-lead to complicated structures.
-
-The preceding examples illustrate the need for prior constraints in
-the map solution. For calibrators (and other previously known bright,
-compact sources), a good, simple prior is to constrain the map to a
-value of zero away from the known locations of emission. In
-Fig.~\ref{fig:pointmaps}d, a map is produced in an identical manner to
-the previous example, but now setting the map explicitly to zero
-beyond a radius of 60\,arcsec from the location of Uranus (much larger
-than the FWHM of the main lobe), for all but the final iteration. In
-this case, the map converges after 22 iterations, and all of the
-previous artifacts have been effectively removed. Since the map is now
-flat away from the source, and constrained to a value of zero, it is
-now appropriate for performing aperture photometry directly, with no
-need for an additional reference annulus. This approach to constrained
-map-making is similar to that employed for poorly cross-linked scans
-of compact (though resolved) sources by \citet{wiebe2009} using BLAST
-data.
+source is actually removed for subsequent iterations of \model{FLT},
+but as in the previous case, degeneracies between the map and the
+\model{FLT} and \model{COM} signal components can still lead to
+complicated structures.
+
+The preceding examples illustrate the need for constraints in the map
+solution in many situations. For calibrators (and other previously
+known bright, compact sources), a good, simple prior is to constrain
+the map to a value of zero away from the known locations of
+emission. In Fig.~\ref{fig:pointmaps}d, a map is produced in an
+identical manner to the previous example, but now setting the map
+explicitly to zero beyond a radius of 60\,arcsec from the location of
+Uranus (much larger than the FWHM of the main lobe), for all but the
+final iteration. In this case, the map converges after 22 iterations,
+and all of the previous artifacts have been effectively removed. Since
+the map is now flat away from the source, and constrained to a value
+of zero, it is appropriate for performing aperture photometry
+directly, with no need for an additional reference annulus. This
+approach to constrained map-making is similar to that employed for
+poorly cross-linked scans of compact (though resolved) sources by
+\citet{wiebe2009} using BLAST data.
 
@@
   is not removed by the common-mode (Uranus peak 0.14\,pW); (b) same
-  as (a) but using 50 iterations, illustrating that there is a
-  degeneracy between large-scale structure and the common-mode (the
-  footprint of working bolometers is also shown for reference, Uranus
-  peak 0.14\,pW); (c) reduction in which high-pass filtering above
-  0.775\,Hz (corresponding to a spatial scale of 200\,arcsec as
-  indicated) is applied after common-mode removal, but before the map
-  estimate --- the map-based convergence test is activated and the
-  solution halts after 10 iterations, but leaving large-scale ringing
-  due to the filter (Uranus peak 0.38\,pW); (d) same as (c), but now
-  the region of the map beyond the white dot-dashed circle is
-  constrained to a value of zero until all but the final iteration ---
-  the map is now extremely flat, there is little attenuation of the
-  source flux, and the diffraction pattern is clearly seen (Uranus
-  peak 0.38\,pW).}
+  as (a) but using 50 iterations, illustrating the degeneracy between
+  large-scale structure and the common-mode (the footprint of working
+  bolometers is also shown for reference, Uranus peak 0.14\,pW); (c)
+  reduction in which high-pass filtering above 0.775\,Hz
+  (corresponding to a spatial scale of 200\,arcsec as indicated) is
+  applied after common-mode removal, but before the map estimate ---
+  the map-based convergence test is activated and the solution halts
+  after 10 iterations, but leaving large-scale ringing due to the
+  filter (Uranus peak 0.38\,pW); (d) same as (c), but now the region
+  of the map beyond the white dot-dashed circle is constrained to a
+  value of zero until all but the final iteration --- the map is now
+  extremely flat, there is little attenuation of the source flux, and
+  the diffraction pattern is clearly seen (Uranus peak 0.38\,pW).}
 \label{fig:pointmaps}
@@
 \subsection{Deep blind survey}
+\label{sec:cosmo}
 
 \scuba\ surveys designed to detect extremely faint point-sources
-(e.g. high-redshift star-forming galaxies, and debris disks) need to
-obtain the best noise performance possible. The approach described
-here involves three major steps: (i) generating a map that removes
-most low-frequency noise sources; (ii) \emph{whitening} (essentially
-flattening) the map to make the noise properties of the map easier to
-deal with; and (iii) detecting point sources using a
+(e.g. high-redshift star-forming galaxies, and debris disks) are
+ideally limited by the white-noise performance of the instrument. The
+approach described here for maximizing the S/N of point-sources
+involves three major steps: (i) generating a map that removes most
+low-frequency noise sources with approximately linear response without
+prior knowledge of the location of sources; (ii) \emph{whitening}
+(essentially flattening) the map to make the noise properties of the
+map easier to deal with; and (iii) detecting point sources using a
 \emph{matched-filter}. Note that variations on this general procedure
 have been used extensively in the submm cosmology community using
-previous instruments \citep[e.g.][]{scott2008,perera2008}.  In this
-section we will reduce data of the Lockman Hole taken in 2010 during
-S2SRO as a pilot project for the Cosmology Legacy Survey. It consists
-of $\sim8.5$\,hours of data taken in 36 separate scans (average
-15\,min. each) spread over February and March, covering an area of
-about 50\,arcmin$^2$. ... full obs list ... dropped scans due to
-terrible noise properties ...
-
-The first step, map generation, is accomplished in a way similar to
-the default described in the previous section (without any boundary
-constraints)...
-
-Even though the map \emph{looks} extremely flat, it is clear that
-there is a mixture of faint astronomical sources, and what is probably
-residual low-frequency noise causing faint ripples. With uncertainties
-about what is source, and what is an artifact, the first step is to
-remove the low-frequency noise in a way that does not affect
-(appreciably) the real sources. To accomplish this goal, the map is
-run through a whitening filter that results in a flat angular noise
-power spectrum. The noise power spectrum is estimated from a
-\emph{jackknife map}: maps are produced from two independent halves of
-the total data set, and then a pixel-by-pixel difference is
-calculated. Provided that the noise in one half of the data is
-uncorrelated with that from the other half, the noise in the jackknife
-map should resemble that of the total map, though larger by a factor
-of $\sim\sqrt(2)$ depending on how closely the pixel variances in the
-two maps are matched (if one map pixel is significantly noisier, it
-will dominate the noise in the jackknife).  The astronomical signal,
-however, should be cleanly removed (provided that there are no strong
-time-varying signals, and also assuming that errors due to calibration
-between the two halves are insignificant). The approach we have taken
-to minimize systematics is to produce two halves of the jackknife
-using odd and even scan numbers (i.e., each map will contain an even
-sampling of data across the full dataset).
+previous instruments
+\citep[e.g.][]{scott2002,laurent2005,coppin2006,scott2008,perera2008,devlin2009}.
+In this section we will reduce data of the Lockman Hole taken in 2010
+during S2SRO as a pilot project for the Cosmology Legacy Survey. It
+consists of $\sim8.5$\,hours of data taken in 36 separate scans
+(average 15\,min. each) spread over February and March, covering an
+area of about 50\,arcmin$^2$. ... full obs list ... dropped scans due
+to terrible noise properties ...
+
+\begin{figure*}
+\centering
+\includegraphics[width=\linewidth]{lockman_maps.pdf}
+\caption{Reduction of a blank-field survey: the Lockman Hole. (a) raw
+  output of SMURF using high-pass filtering as a pre-processing step,
+  followed by 4 iterations using only the \model{COM} model to remove
+  residual low-frequency noise. White contours correspond to estimated
+  noise levels of 1.25, 2.5 and 5.0 times the minimum noise at the
+  centre of the map. (b) the whitened map using a filter based on the
+  angular power spectrum of the jackknife map. (c) The whitened map
+  cross-correlated with the whitened PSF to identify point
+  sources. The image shows the S/N, with 3.8-$\sigma$ peaks indicated
+  with blue circles (radius 8\,arcsec). The orange ``$+$'' give the
+  locations of radio sources from ... . Of the 10 submm peaks, 9 are
+  within this search radius of at least one radio source. (d)
+  jackknife map produced from the difference of two maps ,using the
+  even and odd scan numbers, respectively. Provided that all noise
+  sources are statistically un-correlated between the two halves of
+  the data, the map is a plausible realization of the noise without
+  contamination from astronomical sources. (e) The jackknife map
+  whitened using the same filter as that in panel (b). The whitened
+  jackknife map cross-correlated with the same whitened PSF as in
+  panel (c). Again, orange ``$+$'' and blue circles indicate radio
+  sources and 3.8-$\sigma$ peaks, respectively. Unlike panel (c),
+  there is only a single significant peak, and it is not in the
+  vicinity of a radio source.}
+\label{fig:lockman_maps}
+\end{figure*}
+
+The first step, map generation, is different from that described in
+Section~\ref{sec:point} in two key ways. Since the locations of
+sources are unknown \emph{a priori}, a map constraint is not
+employed. Large-scale diverging structures in the map must be
+mitigated, and the method used in this example (the default processing
+in SMURF) is to apply a high-pass filter to the data \emph{once} as a
+pre-processing steps. The iterative solution is then run using only
+\model{COM}, \model{EXT}, \model{AST}, and \model{NOI}. In other
+words, there is no information in the bolometer signals below some
+cutoff frequency, and residual noise is removed only using iterative
+estimation of the residual (higher-frequency) common-mode. The map is
+shown in Fig.~\ref{fig:lockman_maps}a. The map-maker has been tested
+in two ways: (i) large numbers of iterations are used to verify that
+the maps converge without the growth of large structure; and (ii)
+adding synthetic sources to the data at a range of brightnesses verify
+that the map-maker response to them is \emph{linear} (i.e. the
+relative shape and amplitude compared to the input source is
+independent of brightness). The results of this second test are shown
+by the solid and dotted lines in Fig.~\ref{fig:psf},
+respectively. Clearly the down-side of using a high-pass filter as a
+pre-processing step, and having no other map-constraints, has the
+down-side of introducing sidelobes around the main peak. ... high-pass
+filter + iterative common-mode removal takes place of PCA cleaning in
+AzTEC reduction ... probably more linear.
+
+Even though the map \emph{looks} extremely flat, there is a mixture of
+faint astronomical sources, and what is probably residual
+low-frequency noise causing faint ripples visible to the naked
+eye. Since the mixture of the two components is unknown, the first
+step is to remove the low-frequency noise in a way that does not
+affect (appreciably) the real sources. To accomplish this goal, the
+map is run through a whitening filter that ideally results in a flat
+angular noise power spectrum. The noise power spectrum is estimated
+from a \emph{jackknife map}: maps are produced from two independent
+halves of the total data set, and then a pixel-by-pixel difference is
+calculated. Specifically, the jackknife signal in a map pixel,
+$S_\mathrm{JK}$, and variance, $\sigma^2_\mathrm{JK}$ are estimated
+from the map quantities, $S_1$, $S_2$, $\sigma^2_1$, and $\sigma^2_2$
+as,
+%
+\begin{eqnarray}
+S_\mathrm{JK} = \frac{S_1 - S_2}{2} \\
+\sigma^2_\mathrm{JK} = \frac{\sigma^2_1 + \sigma^2_2}{2}
+\end{eqnarray}
+
+Provided that the noise in one half of the data is uncorrelated with
+that from the other half, the noise in the jackknife map should
+resemble that of the raw map. The astronomical signal, however, should
+be cleanly removed (provided that there are no strong time-varying
+signals, and also assuming that errors due to calibration between the
+two halves are insignificant). The approach we have taken to minimize
+systematics is to produce two halves of the jackknife using odd and
+even scan numbers (i.e., each map will contain an even sampling of
+data across the full dataset).
 
@@
 \centering
-\includegraphics[width=\linewidth]{lockman_rawmaps.pdf}
-\caption{Procedure for producing maps of deep fields.}
-\label{fig:lockman_rawmaps}
+\includegraphics[width=\linewidth]{lockman_psf.pdf}
+\caption{Slice through the angular response to an ideal Gaussian point
+  source (solid line) along the R.A. axis following map-making using
+  the default blank-field processing with SMURF (dotted line), and
+  upon application of the whitening filter (dashed line). The ``map
+  filtered'' response is produced by adding the ideal Gaussian to the
+  real data near the centre of the Lockman Hole map, and re-reducing
+  the data. The resulting dotted line gives the expected shape of a
+  source in panel (a) from Fig.~\ref{fig:lockman_maps}. Applying the
+  whitening filter (reciprocal of the solid orange line in
+  Fig.~\ref{fig:lockman_pspec}) to the dotted line gives the whitened
+  line, which is the expected shape of a source in panel (b) from
+  Fig.~\ref{fig:lockman_maps}. Finally, cross-correlating the whitened
+  map with this whitened PSF is an effective ``matched filter'' for
+  identifying point-like sources, and this smoothed map is shown in
+  panel (c) of Fig.~\ref{fig:lockman_maps}.}
+\label{fig:lockman_psf}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{lockman_pspec.pdf}
+\caption{Azimuthally-averaged angular power spectral densities for the
+  raw map output by SMURF (dashed black line), the jackknife (solid
+  orange line), and whitened (solid black line) maps (panels (a), (d),
+  and (b) from Fig.~\ref{fig:lockman_maps}, respectively. Since the
+  raw map contains spatially correlated signals on large scales both
+  due to noise and astronomical signals (low-spatial frequencies), the
+  jackknife map (difference of two approximately equal-length subsets
+  of the data) is used to generate a plausible realization of pure
+  noise. Assuming that the noise properties are isotropic, a whitening
+  filter is estimated from the reciprocal of the jackknife power
+  spectrum with only a radial dependence. The power spectrum of the
+  resulting signal map has residual power on large scales which is
+  presumably due to astronomical sources.}
+\label{fig:lockman_pspec}
 \end{figure}

### 2012-01-26 smurf_paper.tex
add histogram plot to cosmology section

@@
 independent of brightness). The results of this second test are shown
-by the solid and dotted lines in Fig.~\ref{fig:psf},
+by the solid and dotted lines in Fig.~\ref{fig:lockman_psf},
 respectively. Clearly the down-side of using a high-pass filter as a
@@
 S_\mathrm{JK} = \frac{S_1 - S_2}{2} \\
-\sigma^2_\mathrm{JK} = \frac{\sigma^2_1 + \sigma^2_2}{2}
+\sigma^2_\mathrm{JK} = \frac{\sigma^2_1 + \sigma^2_2}{4}
 \end{eqnarray}
@@
 Provided that the noise in one half of the data is uncorrelated with
-that from the other half, the noise in the jackknife map should
-resemble that of the raw map. The astronomical signal, however, should
-be cleanly removed (provided that there are no strong time-varying
-signals, and also assuming that errors due to calibration between the
-two halves are insignificant). The approach we have taken to minimize
-systematics is to produce two halves of the jackknife using odd and
-even scan numbers (i.e., each map will contain an even sampling of
-data across the full dataset).
+that from the other half, the signal in the jackknife map should
+resemble noise drawn from the same parent distribution as that of the
+real map. The astronomical signal, however, should be cleanly removed
+(provided that there are no strong time-varying signals, and also
+assuming that errors due to calibration between the two halves are
+insignificant). The approach we have taken to minimize systematics is
+to produce two halves of the jackknife using odd and even scan numbers
+(i.e., each map will contain a nearly evenly-spaced sampling of data
+across the full dataset).
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{lockman_hist.pdf}
+\caption{Histograms of S/N using pixels from the central region of
+  Lockman Hole (second contour) in Fig.~\ref{fig:lockman_maps}a. The
+  top panel shows the histograme for pixels from the whitened signal
+  and jackknife maps (Figs.~\ref{fig:lockman_maps}b,e), compared to a
+  Gaussian distribution with mean zero, standard deviation one, and an
+  area normalized to the total number of pixels -- the expected
+  distribution for a map of spatially-uncorrelated Gaussian noise. The
+  good agreement indicates that these maps are indeed dominated by
+  white noise, whose amplitude has been accurately estimated during
+  map-making. The lower panel shows the results for matched-filtered
+  signal and jackknife maps (Figs.~\ref{fig:lockman_maps}c,f). The
+  filtered jackknife map distribution is still close to the
+  expectation (Gaussian), but now the matched-filter has picked out
+  significant signal leading to the large positive tail. }
+\label{fig:lockman_hist}
+\end{figure}
+
 An additional test that can be accomplished with the jackknife map, is

### 2012-01-27 smurf_paper.tex
more text for cosmology section

@@
 involves three major steps: (i) generating a map that removes most
-low-frequency noise sources with approximately linear response without
-prior knowledge of the location of sources; (ii) \emph{whitening}
-(essentially flattening) the map to make the noise properties of the
-map easier to deal with; and (iii) detecting point sources using a
-\emph{matched-filter}. Note that variations on this general procedure
-have been used extensively in the submm cosmology community using
-previous instruments
+low-frequency noise sources with approximately linear response,
+without prior knowledge of the location of sources; (ii)
+\emph{whitening} (essentially flattening) the map to make the noise
+properties of the map easier to deal with; and (iii) detecting point
+sources using a \emph{matched-filter}. Note that variations on this
+general procedure have been used extensively in the submm cosmology
+community using previous instruments
 \citep[e.g.][]{scott2002,laurent2005,coppin2006,scott2008,perera2008,devlin2009}.
-In this section we will reduce data of the Lockman Hole taken in 2010
+In this section we will reduce scans of the Lockman Hole taken in 2010
 during S2SRO as a pilot project for the Cosmology Legacy Survey. It
@@
 (average 15\,min. each) spread over February and March, covering an
-area of about 50\,arcmin$^2$. ... full obs list ... dropped scans due
-to terrible noise properties ...
+area of about 50\,arcmin$^2$. The full list of dates and observation
+numbers includes 20100218: 63, 64, 65, 70, 71, 72, 90, 91, 92, 97, 98,
+99; 20100220: 111, 112, 113, 118, 128, 129, 130; 20100303: 61, 62, 64,
+69, 70, 72, 73, 74, 75; 20100309: 87, 88, 90, 91; and 20100311: 59,
+64, 65, 72. These observations represent about 80\% of the total data
+taken as part of the project; the remaining observations were dropped
+due to problems keeping the arrays properly locked during S2SRO, and
+were easily identified by their highly variable and erratic bolometer
+time-series (which resulted in maps full of large streaks).
 
@@
 \includegraphics[width=\linewidth]{lockman_pspec.pdf}
-\caption{Azimuthally-averaged angular power spectral densities for the
-  raw map output by SMURF (dashed black line), the jackknife (solid
-  orange line), and whitened (solid black line) maps (panels (a), (d),
-  and (b) from Fig.~\ref{fig:lockman_maps}, respectively. Since the
-  raw map contains spatially correlated signals on large scales both
-  due to noise and astronomical signals (low-spatial frequencies), the
-  jackknife map (difference of two approximately equal-length subsets
-  of the data) is used to generate a plausible realization of pure
-  noise. Assuming that the noise properties are isotropic, a whitening
-  filter is estimated from the reciprocal of the jackknife power
-  spectrum with only a radial dependence. The power spectrum of the
-  resulting signal map has residual power on large scales which is
-  presumably due to astronomical sources.}
+\caption{Radial (azimuthally-averaged) angular power spectral
+  densities for the raw map output by SMURF (dashed black line), the
+  jackknife (solid orange line), and whitened (solid black line) maps
+  (panels (a), (d), and (b) from Fig.~\ref{fig:lockman_maps},
+  respectively (considering only the square region indicated in
+  Fig.~\ref{fig:lockman_maps}d). Since the raw map contains spatially
+  correlated signals on large scales both due to noise and
+  astronomical signals (low-spatial frequencies), the jackknife map
+  (difference of two approximately equal-length subsets of the data)
+  is used to generate a plausible realization of pure noise. Assuming
+  that the noise properties are isotropic, a whitening filter is
+  estimated from the reciprocal of the jackknife power spectrum with
+  only a radial dependence. The power spectrum of the resulting signal
+  map still has residual power on large scales, which is presumably
+  due to astronomical sources.}
 \label{fig:lockman_pspec}
@@
 assumption that the angular noise power spectrum is azimuthally
-symmetric (inspection of the 2-dimensional FFT confirms
-this). ... calculate filter as slightly-smoothed complement of
-azavpspec ... Finally, the filter is applied to the map, resulting in
-panel...
+symmetric. For these data, there are no obvious anisotropic structures
+in the 2-dimensional FFT. The radial (azimuthally-averaged) angular
+power spectrum therefore encodes all of the useful information about
+the noise properties. These power spectra for the raw output of SMURF,
+and the jackknife map (transforming only the approximately uniform
+region indicated by the square in Fig.~\ref{fig:lockman_maps}d in each
+case) by the dashed black, and solid orange lines in
+Fig.~\ref{fig:lockman_pspec}, respectively. Both power spectra are
+approximately flat at spatial frequencies $\gsim 0.06$\,arcsec$^{-1}$
+(scales $\lsim16$\,arcsec), with the exception of a narrow line
+feature at $\sim0.175$\,arcsec$^{-1}$ (a scale of
+$\sim5.7$\,arcsec). A possible theory for the feature at 5.7\,arcsec
+is that it is related to the inter-bolometer spacing in the focal
+plane (which happens to be this size); due to small relative drifts in
+the baselines of adjacent bolometers, there may be faint parallel
+stripes in the map along the scan direction (the superposition of many
+scans at different angles then results in an isotropic noise
+pattern). It is not likely that this signal is due to astronomical
+sources because it appears with a nearly equal amplitude in both the
+real map and the jackknife.  At lower spatial frequencies, both the
+real map and the jackknife power spectra grow significantly, as a
+result of the more obvious large-scale undulations in
+Figs.~\ref{fig:lockman_maps}(a) and (d).
+
+In this example, to identify point sources, we first filter the maps
+to suppress the large-scale waves. We use the radial power spectrum of
+the jackknife map (orange line in Fig.~\ref{fig:lockman_pspec}) as it
+should contain no appreciable contamination from astronomical sources.
 
@@
 
+default: converges after 21 iterations
+
 %------------------------------------------------------------------------------

### 2012-04-02 smurf_paper.tex
small test commit from behind ESAC firewall

@@
 by the solid and dotted lines in Fig.~\ref{fig:lockman_psf},
-respectively. Clearly the down-side of using a high-pass filter as a
+respectively. Clearly the use of a high-pass filter as a
 pre-processing step, and having no other map-constraints, has the
-down-side of introducing sidelobes around the main peak. ... high-pass
-filter + iterative common-mode removal takes place of PCA cleaning in
-AzTEC reduction ... probably more linear.
+down-side of introducing sidelobes around the main peak. However, the
+way this filter affects point-sources is both well-known, and linear.
+
+%... high-pass
+%filter + iterative common-mode removal takes place of PCA cleaning in
+%AzTEC reduction ... probably more linear.
 

### 2012-05-14 smurf_paper.tex
Update Frossie's present address

@@
     Douglas~Scott$^{1}$,
-    Frossie~Economou$^2$
+    Frossie~Economou$^2$\thanks{Present address: National Optical
+      Astronomy Observatory, 950 N.\ Cherry Avenue, Tucson, AZ 85719, USA}
   }

### 2012-05-14 smurf_paper.tex
Update Dempsey and Holland to 2012

@@
 commissioned. The details of the instrument design, performance, and
-its calibration are given in two companion papers: \citet{holland2011}
-and \citet{dempsey2011}.
+its calibration are given in two companion papers: \citet{holland2012}
+and \citet{dempsey2012}.
 
@@
 While the details of how \scuba\ works is described in
-\citet{holland2011}, and its calibration in \citet{dempsey2011}, we
+\citet{holland2012}, and its calibration in \citet{dempsey2012}, we
 summarize the basic concepts relative to map-making here.
@@
 focal plane bias to make as many detectors useable as possible
-\citep{holland2011}. In practice, the response of a given detector can
+\citep{holland2012}. In practice, the response of a given detector can
 therefore vary somewhat from observation to observation depending on
@@
 also be obtained. While not discussed further in this paper, details
-may be found in \citet{dempsey2011}. We will refer to measurements of
+may be found in \citet{dempsey2012}. We will refer to measurements of
 $dI/dP$ across the focal plane as the
@@
 determined empirically using measurements of standard sources. For
-further details on these calibration steps, see \citet{dempsey2011}.
+further details on these calibration steps, see \citet{dempsey2012}.
 
@@
 determined completely from external measurements
-\citep[see][]{dempsey2011}. Next, residual noise that differs from
+\citep[see][]{dempsey2012}. Next, residual noise that differs from
 bolometer-to-bolometer (primarily low-frequency) is removed, usually
@@
 final map.  For details on how it is calculated see
-\citet{dempsey2011}. Note that its numerical value is calculated only
+\citet{dempsey2012}. Note that its numerical value is calculated only
 once, and simply applied as a scale factor in the iterative solution.

### 2012-05-29 smurf_paper.tex
work on intro, new section on convergence tests

@@
 % journals
+\newcommand{\ao}{\rm AO}
 \newcommand{\apj}{\rm ApJ}
@@
   \parbox[t]{\textwidth}{
-    Edward~L.~Chapin$^{1,2}$\thanks{E-mail:~echapin@phas.ubc.ca},
-    David~S.~Berry$^{2}$,
-    Andrew~G.~Gibb$^{1}$,
-    Tim~Jenness$^{2}$,
-    Douglas~Scott$^{1}$,
-    Frossie~Economou$^2$\thanks{Present address: National Optical
+    Edward~L.~Chapin$^{1,2,3}$\thanks{E-mail:~echapin@sciops.esa.int},
+    David~S.~Berry$^{3}$,
+    Andrew~G.~Gibb$^{2}$,
+    Tim~Jenness$^{3}$,
+    Douglas~Scott$^{2}$,
+    Frossie~Economou$^3$\thanks{Present address: National Optical
       Astronomy Observatory, 950 N.\ Cherry Avenue, Tucson, AZ 85719, USA}
@@
   \\
-  $^{1}$Dept. of Physics \& Astronomy, University of British Columbia,
+  $^{1}$XMM SOC, ESAC, Apartado 78, 28691 Villanueva de la Ca\~nada, Madrid,
+  Spain\\
+  $^{2}$Dept. of Physics \& Astronomy, University of British Columbia,
   6224 Agricultural Road, Vancouver, B.C. V6T 1Z1, Canada\\
-  $^{2}$JointAstronomy Centre, 660 N. A`oh\={o}k\={u} Place, University
+  $^{3}$JointAstronomy Centre, 660 N. A`oh\={o}k\={u} Place, University
   Park, Hilo, Hawaii 96720, USA}
@@
 
-\textbf{Guideline for figures:}
-
-\begin{itemize}
-\item minimize excess whitespace around the figure to optimize usage of space
-\item ensure a decent line thickness
-\item use a Times font for labels with a thickness that resembles the
-  text in the body. The font size should be similar to the main body
-  text or larger
-\item avoid using colour unless it is necessary for emphasis
-\end{itemize}
-
-\textbf{start intro here...}
+%\textbf{Guideline for figures:}
+
+%\begin{itemize}
+%\item minimize excess whitespace around the figure to optimize usage of space
+%\item ensure a decent line thickness
+%\item use a Times font for labels with a thickness that resembles the
+%  text in the body. The font size should be similar to the main body
+%  text or larger
+%\item avoid using colour unless it is necessary for emphasis
+%\end{itemize}
+
+%\textbf{start intro here...}
 
@@
 Over the last twenty years, observations in the submillimetre band
-(defined here to be 200--1000\,\micron) have helped revolutionized
-several important areas of astrophysics including: the
+(defined here to be 200--1000\,\micron) have helped revolutionize
+several important areas of astrophysics, including: the
 characterization of early stages of star-formation by identifying the
@@
 
-Most submillimetre imaging cameras to date have used bolometers to
-maximize sensitivities: each pixel consists of an efficient thermal
-absorber to collect photons that have passed through a broad-band
-filter, and the signal is read-out using a thermometer. The
-high-frequency broad-band noise in modern systems, such as SCUBA-2, is
-limited by photon noise from the instrument and ambient
-backgrounds. The low-frequency noise, however, is dominated by sources
-such as the low-frequency drift in the readout electronics, and drifts
-in the mean background levels (such as the atmosphere for ground-based
-instruments). The process of making a map therefore amounts to
-separating the component of the bolometer signals that is produced by
-astronomical sources, from the non-white (and therefore correlated)
-contaminating noise components. Fortunately each has a particular
-profile: astronomical sources are typically fixed in a sky reference
-frame (to a good approximation), whereas other noise sources are often
-correlated in time, and usually from bolometer-to-bolometer. ...also
-mention scan strategies...
-
-There are numerous ways to attack this problem, most of them evolving
-out of techniques that were developed to produce maps of the
-large-scale anisitropies in the Cosmic Microwave Background (CMB).
-Styles of map-making: least squares \citep{janssen1992} maximum
-likelihood \citep[e.g.,][]{patanchon2008}; iterative -- see
-\citet{johnstone2000} implementation of \citet{wright1996} to remove
-chop, \citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA
-analysis \citep[e.g.][]{laurent2005,scott2008,aguirre2010}.
+Most submillimetre imaging cameras, to date, have used bolometers to
+maximize sensitivities \citep[e.g.][]{griffin2002}. Each pixel
+consists of a thermal absorber to collect photons that have passed
+through broad-band filters, and the signal is read-out using a
+thermometer. The high-frequency broad-band noise in modern systems,
+such as SCUBA-2, is limited by photon noise from the instrument and
+ambient backgrounds. The low-frequency noise, however, is dominated by
+sources such as slow variations in mean backgrounds (such as thermal
+variations within the cryostat, and the atmosphere for ground-based
+cameras), and drift in the readout electronics. Since the
+low-frequency noise is largely correlated between all, or subsets, of
+the bolometers in time, it can be suppressed during map-making, since
+astronomical signals have the distinct property that they are fixed in
+a sky reference frame (assuming they are not time-varying). If
+successful, the noise in the resulting map is said to be ``background
+noise limited'', meaning that the only significant remaining noise
+sources are uncorrelated spatially in the map (the theoretical limit).
+
+There are numerous ways to attack this problem, both in terms of the
+data-collection method, and data processing. Historically, both radio
+and submillimetre telescope systems have used ``chopping'' to perform
+rapid differences between target sky locations and some reference
+(e.g., another point on the sky, internal calibrators, and different
+frequencies in the case of heterodyne systems). In fact, spatial
+chopping was used by SCUBA, the predecessor to SCUBA-2, to remove
+low-frequency drifts \citep{holland1999}. However, chopping (in
+particular spatial chopping) has the unfortunate side-effect of
+filtering out information on scales larger than the chop-throw
+\textbf{(references)}.For this reason, modern cameras like SCUBA-2 aim
+to provide a high-degree of inherent detector stability, and rapidly
+slew the telescope so that astronomical signals up to large scales are
+captured in a signal band at considerably higher frequency than the
+residual low-frequency drift.
+
+In the case of chopped data, ideally the resulting data are completely
+free of low-frequency noise, and therefore the data samples are
+statistically uncorrelated.
+
+often used various
+``chopping'' methods to remove low-frequency drifts -- rapid
+differences are made between the sky signal at a given position and an
+internal reference signal; the sky signal in two different bands for
+heterodyne systems (frequency chopping); and a different position in
+the sky \citep[a method employed by SCUBA, the predecessor of
+SCUBA-2][]{holland1999}.
+
+ many of them evolving
+out of techniques for producing maps of the large-scale anisitropies
+in the Cosmic Microwave Background (CMB)
+
+  Styles of map-making: least
+squares \citep{janssen1992} maximum likelihood
+\citep[e.g.,][]{patanchon2008}; iterative -- see \citet{johnstone2000}
+implementation of \citet{wright1996} to remove chop,
+\citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA analysis
+\citep[e.g.][]{laurent2005,scott2008,aguirre2010}.
 
@@
   \mathbf{n}^\mathrm{c}(t),
+\label{eq:noise}
 \end{equation}
@@
 \begin{equation}
+\label{eq:varmap}
 \mathbf{v}(x_j,y_j) = \frac{\sum_j \mathbf{w}_j
@@
 
-The primary purpose of the noise component, \model{NOI} is to measure
-the noise in each bolometer residual once all of the other models have
-been fit and removed. This noise may then be used to estimate weights
-for the data in subsequent iterations.
+The primary purpose of the noise component, \model{NOI}, is to measure
+the white noise of each bolometer ($\mathbf{n}^\mathrm{w}_i(t)$ in
+Eq.~\ref{eq:noise}). The measurement generally occurs once all of the
+other models have been fit and removed. This noise may then be used to
+estimate weights for the data in subsequent iterations.
 
@@
 taken, and the PSD calculated. An average white noise level is then
-measured from 2 to 10\,Hz, a relatively clean region that lies above
-the $1/f$ knee, but below the high-frequency line features for most of
-the bolometer data taken during S2SRO (Fig.~\ref{fig:pspec}). Taking
+measured from 2 to 10\,Hz, a relatively clean region of the PSD that
+tends to lie above the $1/f$ knee, but below the high-frequency line
+features for typical bolometer data (Fig.~\ref{fig:pspec}). Taking
 this constant level for $\mathbf{P}(f)$ we then calculate the expected
@@
 stored, and then used in subsequent iterations to weight the data
-points when calculating the map estimate (Section~\ref{sec:ast}).
-
-Since noise estimation is usually calculated as the final step, it is
-the cleanest data for which some cleaning operations may be
-performed. For this reason \model{NOI} can optionally use the
+points when calculating the map estimate (Section~\ref{sec:ast}). They
+are also used for convergence tests (Section~\ref{sec:converge}).
+
+Since noise estimation is usually calculated as the final step in the
+iteration, it is the cleanest data for which some cleaning operations
+may be performed. For this reason \model{NOI} can optionally use the
 time-domain DC step fixer (Section~\ref{sec:steps}) and spike
 detection (Section~\ref{sec:timedespike}).
+
+Finally, it should be noted that the default behaviour is to calculate
+the white noise levels \emph{once} within \emph{NOI}, after the second
+iteration. The reason for fixing these values is to prevent any
+potential divergence in the weight estimates with iterations, and also
+to provide a fixed reference for the convergence tests
+(Section~\ref{sec:converge}). Note that the absolute values of the
+noise, and therefore weights calculated by the model component are
+irrelevant (only their relative values). The reason is that the final
+noise in the map is measured empirically from the scatter of the data
+points that land in each pixel (Section~\ref{sec:ast}), and
+large-scale noise sources can be checked by analyzing the angular
+power spectra of jackknife maps (e.g., Section~\ref{sec:cosmo}).
 
@@
 
-\subsection{Model degeneracies}
-
-A fundamental problem faced by the SMURF algorithm, as described, is
-that there are degeneracies between model parameters. Thinking of each
-pixel in the output map, as well as each sample in the various noise
-model components, such as the common-mode and signal removed by the
-Fourier filtering, any degeneracy can cause the solution to
-diverge. Of particular concern are situations in which the map
-diverges.
+\subsection{Convergence tests and model degeneracies}
+\label{sec:converge}
+
+The map-maker will halt if the convergence criteria have been
+met. Presently there are two numerical tests that may be performed by
+SMURF, or alternatively, a fixed number of iterations may be
+specified. The first numerical test is an approximate check of the
+change in reduced chi-squared, $\chi^2_\mathrm{r}$. The standard
+deviations, $\sigma_i$, are measured for each of the residual
+bolometer signals once the modeled signal components have been
+removed. In the early iterations, the residual signals usually contain
+both white noise, and other long-timescale features. However, once the
+solution has converged, this signal should look approximately
+white. Once the white nosie, $\mathbf{n}^\mathrm{w}_i(t)$, has been
+measured from the bolometer PSDs in \model{NOI}
+(Section~\ref{sec:noi}), $\chi^2_\mathrm{r}$ is calculated as $(1/N)
+\sum_i \sigma^2_i / \mathbf{n}^\mathrm{w}_i(t)$, where $i$ runs over
+the $N$ bolometers. In other words, it is the average ratio between
+the measured bolometer variances and their theoretical white-noise
+levels. Note that this expression \emph{does not account for the
+degrees of freedom}. Clearly there are a large number of parameters in
+the model, which will ``fit-out'' some of the uncorrelated white
+noise, and potentially bias this estimate of $\chi^2_r$ low. One
+approach would be to keep track of the degrees of freedom, and
+therefore provide a correction, as in \citet{kovacs2008}. However,
+note that we are only using this value as a convergence test, and its
+value is not used in an absolute sense. Furthermore, the reference
+white noise values are calculated only once in SMURF
+(Section~\ref{sec:noi}), and the number of model parameters are fixed,
+so the value of $\chi^2_r$ should decrease monotonically if the
+solution is behaving well. The convergence criterion is met if the
+change in subsequent values of $\chi^2_r$ is smaller than the
+parameter \texttt{chitol}, which is set to $10^{-3}$ by default.
+
+It was immediately found in the analysis of SCUBA-2 data that the
+$\chi^2_r$ test described above was not a sufficient stopping
+criterion. The reason for this is that various components of the model
+are, under normal circumstances, highly degenerate. Once simple
+example is the degeneracy between large-scale astronomical structures
+(larger than the array footprint), and the common-mode rejection step
+(\model{COM}), which will be illustrated in
+Section~\ref{sec:point}. Since the flatness of the residual only
+reflects the ability of the model to fit the shape of the bolometer
+time-series, it does not necessarily correlate with convergence of the
+most important model component: the map. Indeed, examination of the
+map after large numbers of iterations (e.g., $>10$), shows the
+presence of large-scale structures that are anti-correlated with
+structures in other model components. For this reason, a second
+map-based convergence test was added.
+
+The map-based convergence statistic, $M_c$, is the average absolute
+change in the value of map pixels between subsequent iterations,
+normalized by the map pixel uncertainties (square root of
+Eq.~\ref{eq:varmap}), or
+%
+\begin{equation}
+M^j_c = \frac{1}{N} \sum_i \frac{| \mathbf{m}(x_j,y_j) -
+  \mathbf{m}(x_{j-1},y_{j-1}) |} {\sqrt{\mathbf{v}(x_j,y_j)}} ,
+\end{equation}
+%
+where $i$ runs over the $N$ map pixels, and $j$ enumerates
+iteration. The convergence criterion is met once this average change
+is smaller than the parameter \texttt{maptol}. Experimentally we found
+that a threshold of 0.05 gives good results (on average, map pixels
+change by 5\% of the estimated map RMS in subsequent iterations).
+This provides good correspondance with what we would choose ``by
+eye'', and furthermore, letting the solver run for many more
+iterations in several test cases yields insignificant differences.
 

### 2012-05-30 smurf_paper.tex
finish draft of cosmology section

@@
   We describe the properties of data from the Submillimetre Common
-  User Bolometer Array 2 (SCUBA-2) taken during the Shared Risk
+  User Bolometer Array 2 (\scuba) taken during the Shared Risk
   Observing period from 2010 February 11 to 2010 March 25, and the
@@
 around nearby stars, the early stages of planet formation.  With
-$\sim10000$ total detectors, SCUBA-2 is presently the largest
+$\sim10000$ total detectors, \scuba\ is presently the largest
 submillimetre camera in the world, and the fastest wide-area
@@
 thermometer. The high-frequency broad-band noise in modern systems,
-such as SCUBA-2, is limited by photon noise from the instrument and
+such as \scuba, is limited by photon noise from the instrument and
 ambient backgrounds. The low-frequency noise, however, is dominated by
@@
 frequencies in the case of heterodyne systems). In fact, spatial
-chopping was used by SCUBA, the predecessor to SCUBA-2, to remove
+chopping was used by SCUBA, the predecessor to \scuba, to remove
 low-frequency drifts \citep{holland1999}. However, chopping (in
@@
 filtering out information on scales larger than the chop-throw
-\textbf{(references)}.For this reason, modern cameras like SCUBA-2 aim
+\textbf{(references)}.For this reason, modern cameras like \scuba\ aim
 to provide a high-degree of inherent detector stability, and rapidly
@@
 the sky \citep[a method employed by SCUBA, the predecessor of
-SCUBA-2][]{holland1999}.
+\scuba][]{holland1999}.
 
@@
 analysis described here applies equally to all of the data taken
-to-date with \scuba, we note that the SCUBA-2 Shared-Risks Observing
+to-date with \scuba, we note that the \scuba\ Shared-Risks Observing
 period (S2SRO) which took place from 2010 February 11 to 2010 March 25
@@
 then tiny temperature fluctuations on short time-scales will produce
-significant changes in the impedence. The entire SCUBA-2 focal plane
+significant changes in the impedence. The entire \scuba\ focal plane
 is therefore held at a \emph{base temperature} somewhat below
@@
 than the chop throw (a maximum of a couple of arcmin)... cite
-somebody. The primary approach to map-making with SCUBA-2 is instead
+somebody. The primary approach to map-making with \scuba\ is instead
 to slew the entire telescope at such a rate that angular scales of
@@
 designed with a thermal time constant that balance response with
-sensitivity. Taking these constraints into consideration, SCUBA-2 was
+sensitivity. Taking these constraints into consideration, \scuba\ was
 ultimately designed to work at a maximum scan velocity of
@@
 a typical rule-of-thumb for adequately sampling a Gaussian point
-spread function (PSF). Similarly, the SCUBA-2 bolometers are designed
+spread function (PSF). Similarly, the \scuba\ bolometers are designed
 to have a time constant of 2 ms \textbf{check that number!} in order
@@
 truly random scan patterns and feasibility of implementation taking
-into consideration both mechanical and scheduling constraints, SCUBA-2
+into consideration both mechanical and scheduling constraints, \scuba
 presently has three primary scan modes: PONG (cite Borys et al.),
@@
 \subsection{Typical bolometer signals}
+\label{sec:bolosignal}
 %-------------------------------------------------
@@
 magnetic field fixed with respect to the telescope dome (i.e., the
-earth's magnetic field). Since SCUBA-2 is mounted on a Nasmyth
+earth's magnetic field). Since \scuba\ is mounted on a Nasmyth
 platform, only azimuthal motion will change the effective direction of
@@
 
-One unique feature of the SCUBA-2 data is that one can perform a PCA
+One unique feature of the \scuba\ data is that one can perform a PCA
 analysis of both the 450 and 850\,\micron\ data simultaneously. The
@@
 in direct inversion methods such as \citet{patanchon2008} may in fact
-be tenable for SCUBA-2 data. A future goal of the software team is to
+be tenable for \scuba\ data. A future goal of the software team is to
 investigate meshing SMURF (perhaps as a cleaning step and initial
@@
 are less than $\sim10$ samples per pixel (a situation that is almost
-never encountered in a SCUBA-2 map except in the edge pixels).
+never encountered in a \scuba\ map except in the edge pixels).
 
@@
 (Section~\ref{sec:mapdespike}), and to apply constraints to the map to
-improve convergences (Section~???).
+improve convergences (Section~\ref{sec:converge}).
 
@@
 
-As in Section~\ref{sec:psd} the FFT of each bolometer time series is
-taken, and the PSD calculated. An average white noise level is then
-measured from 2 to 10\,Hz, a relatively clean region of the PSD that
-tends to lie above the $1/f$ knee, but below the high-frequency line
-features for typical bolometer data (Fig.~\ref{fig:pspec}). Taking
-this constant level for $\mathbf{P}(f)$ we then calculate the expected
-variance of the time-series (in approximately 200\,Hz samples) using
+First, the bolometer PSDs are calculated as in Eq.~\ref{eq:psd}. An
+average white noise level is then measured from 2 to 10\,Hz, a
+relatively clean region of the PSD that tends to lie above the $1/f$
+knee, but below the high-frequency line features for typical bolometer
+data (Fig.~\ref{fig:pspec}). Taking this constant level for
+$\mathbf{P}(f)$ we then calculate the expected variance of the
+time-series (in approximately 200\,Hz samples) using
 Eq.~\ref{eq:psd}. If the bolometer noise were produced purely by
@@
 the $N$ bolometers. In other words, it is the average ratio between
-the measured bolometer variances and their theoretical white-noise
-levels. Note that this expression \emph{does not account for the
-degrees of freedom}. Clearly there are a large number of parameters in
-the model, which will ``fit-out'' some of the uncorrelated white
-noise, and potentially bias this estimate of $\chi^2_r$ low. One
-approach would be to keep track of the degrees of freedom, and
-therefore provide a correction, as in \citet{kovacs2008}. However,
-note that we are only using this value as a convergence test, and its
-value is not used in an absolute sense. Furthermore, the reference
-white noise values are calculated only once in SMURF
-(Section~\ref{sec:noi}), and the number of model parameters are fixed,
-so the value of $\chi^2_r$ should decrease monotonically if the
-solution is behaving well. The convergence criterion is met if the
-change in subsequent values of $\chi^2_r$ is smaller than the
-parameter \texttt{chitol}, which is set to $10^{-3}$ by default.
-
-It was immediately found in the analysis of SCUBA-2 data that the
+the measured time-series bolometer variances and their white-noise
+levels measured between 2--10\,Hz. Note that this expression
+\emph{does not account for the degrees of freedom}. Clearly there are
+a large number of parameters in the model, which will ``fit-out'' some
+of the uncorrelated white noise, and potentially bias this estimate of
+$\chi^2_r$ low. One approach would be to keep track of the degrees of
+freedom, and therefore provide a correction, as in
+\citet{kovacs2008}. However, note that we are only using this value as
+a convergence test, and its value is not used in an absolute
+sense. Furthermore, the reference white noise values are calculated
+only once in SMURF (Section~\ref{sec:noi}), and the number of model
+parameters are fixed, so the value of $\chi^2_r$ should decrease
+monotonically if the solution is behaving well. The convergence
+criterion is met if the change in subsequent values of $\chi^2_r$ is
+smaller than the parameter \texttt{chitol}, which is set to $10^{-3}$
+by default.
+
+It was immediately found in the analysis of \scuba\ data that the
 $\chi^2_r$ test described above was not a sufficient stopping
 criterion. The reason for this is that various components of the model
-are, under normal circumstances, highly degenerate. Once simple
-example is the degeneracy between large-scale astronomical structures
-(larger than the array footprint), and the common-mode rejection step
+are, under normal circumstances, highly degenerate. One simple example
+is the degeneracy between large-scale astronomical structures (larger
+than the array footprint), and the common-mode rejection step
 (\model{COM}), which will be illustrated in
@@
 
-\textit{Surely flatfields are discussed in the main SCUBA-2 paper so
+\textit{Surely flatfields are discussed in the main \scuba\ paper so
 don't really have any useful purpose here since they aren't
@@
 of Uranus (observation 26 on 20111017), which is a nearly point-like
-source for SCUBA-2 that is commonly used as a primary flux
+source for \scuba\ that is commonly used as a primary flux
 calibrator. The now standard CV daisy scan pattern was used, with a
@@
 corresponds to a frequency of 0.78\,Hz. Now using the automatic
-map-based convergence test (Section ???), the resulting map converges
-after 10 iterations (Fig.~\ref{fig:pointmaps}c). Both the circular
-streaks and the large-scale gradient are now removed, but they have
-been replaced by an obvious, circularly-symmetric ringing pattern
-about Uranus. The reason for this ringing is that the hard-edged
-high-pass filter in frequency space is equivalent to a $\sinc$
-function-like response in map space. Since the scan pattern is
+map-based convergence test (Section~\ref{sec:converge}), the resulting
+map converges after 10 iterations (Fig.~\ref{fig:pointmaps}c). Both
+the circular streaks and the large-scale gradient are now removed, but
+they have been replaced by an obvious, circularly-symmetric ringing
+pattern about Uranus. The reason for this ringing is that the
+hard-edged high-pass filter in frequency space is equivalent to a
+$\sinc$ function-like response in map space. Since the scan pattern is
 isotropic (scans equally at all position angles), and there is a
@@
 \citep[e.g.][]{scott2002,laurent2005,coppin2006,scott2008,perera2008,devlin2009}.
-In this section we will reduce scans of the Lockman Hole taken in 2010
-during S2SRO as a pilot project for the Cosmology Legacy Survey. It
-consists of $\sim8.5$\,hours of data taken in 36 separate scans
-(average 15\,min. each) spread over February and March, covering an
-area of about 50\,arcmin$^2$. The full list of dates and observation
+In this section we will reduce scans of the Lockman Hole taken during
+S2SRO as a pilot project for the Cosmology Legacy Survey. It consists
+of $\sim8.5$\,hours of data taken in 36 separate scans (average
+15\,min. each) spread over February and March 2010, covering an area
+of about 50\,arcmin$^2$. The full list of dates and observation
 numbers includes 20100218: 63, 64, 65, 70, 71, 72, 90, 91, 92, 97, 98,
@@
   angular power spectrum of the jackknife map. (c) The whitened map
-  cross-correlated with the whitened PSF to identify point
-  sources. The image shows the S/N, with 3.8-$\sigma$ peaks indicated
-  with blue circles (radius 8\,arcsec). The orange ``$+$'' give the
-  locations of radio sources from ... . Of the 10 submm peaks, 9 are
-  within this search radius of at least one radio source. (d)
-  jackknife map produced from the difference of two maps ,using the
-  even and odd scan numbers, respectively. Provided that all noise
-  sources are statistically un-correlated between the two halves of
-  the data, the map is a plausible realization of the noise without
-  contamination from astronomical sources. (e) The jackknife map
-  whitened using the same filter as that in panel (b). The whitened
-  jackknife map cross-correlated with the same whitened PSF as in
-  panel (c). Again, orange ``$+$'' and blue circles indicate radio
-  sources and 3.8-$\sigma$ peaks, respectively. Unlike panel (c),
-  there is only a single significant peak, and it is not in the
-  vicinity of a radio source.}
+  cross-correlated with the whitened PSF to identify point sources
+  [restricted to a lower-noise region, within the area denoted by the
+  second contour of panel (a)]. The image shows the S/N, with
+  3.8-$\sigma$ peaks indicated with blue circles (radius
+  8\,arcsec). The orange ``$+$'' signs show the locations of 1.4\,GHz
+  radio sources from \citet{owen2008} with $S\gsim15$\,$\mu$Jy. Of the
+  10 submm peaks, 9 are within this search radius of at least one
+  radio source. (d) jackknife map produced from the difference of two
+  maps, using the even and odd scan numbers, respectively. Provided
+  that all noise sources are statistically un-correlated between the
+  two halves of the data, the map is a plausible realization of the
+  noise without contamination from astronomical sources. (e) The
+  jackknife map whitened using the same filter as that in panel
+  (b). The whitened jackknife map cross-correlated with the same
+  whitened PSF as in panel (c). Again, orange ``$+$'' and blue circles
+  indicate radio sources and 3.8-$\sigma$ peaks, respectively. Unlike
+  panel (c), there is only a single significant peak, and it is not in
+  the vicinity of a radio source.}
 \label{fig:lockman_maps}
@@
 in SMURF) is to apply a high-pass filter to the data \emph{once} as a
-pre-processing steps. The iterative solution is then run using only
+pre-processing step. The iterative solution is then run using only
 \model{COM}, \model{EXT}, \model{AST}, and \model{NOI}. In other
 words, there is no information in the bolometer signals below some
-cutoff frequency, and residual noise is removed only using iterative
-estimation of the residual (higher-frequency) common-mode. The map is
-shown in Fig.~\ref{fig:lockman_maps}a. The map-maker has been tested
-in two ways: (i) large numbers of iterations are used to verify that
-the maps converge without the growth of large structure; and (ii)
-adding synthetic sources to the data at a range of brightnesses verify
-that the map-maker response to them is \emph{linear} (i.e. the
-relative shape and amplitude compared to the input source is
-independent of brightness). The results of this second test are shown
-by the solid and dotted lines in Fig.~\ref{fig:lockman_psf},
-respectively. Clearly the use of a high-pass filter as a
-pre-processing step, and having no other map-constraints, has the
-down-side of introducing sidelobes around the main peak. However, the
-way this filter affects point-sources is both well-known, and linear.
-
-%... high-pass
-%filter + iterative common-mode removal takes place of PCA cleaning in
-%AzTEC reduction ... probably more linear.
-
-Even though the map \emph{looks} extremely flat, there is a mixture of
-faint astronomical sources, and what is probably residual
-low-frequency noise causing faint ripples visible to the naked
-eye. Since the mixture of the two components is unknown, the first
-step is to remove the low-frequency noise in a way that does not
-affect (appreciably) the real sources. To accomplish this goal, the
-map is run through a whitening filter that ideally results in a flat
-angular noise power spectrum. The noise power spectrum is estimated
-from a \emph{jackknife map}: maps are produced from two independent
-halves of the total data set, and then a pixel-by-pixel difference is
-calculated. Specifically, the jackknife signal in a map pixel,
-$S_\mathrm{JK}$, and variance, $\sigma^2_\mathrm{JK}$ are estimated
-from the map quantities, $S_1$, $S_2$, $\sigma^2_1$, and $\sigma^2_2$
-as,
-%
-\begin{eqnarray}
-S_\mathrm{JK} = \frac{S_1 - S_2}{2} \\
-\sigma^2_\mathrm{JK} = \frac{\sigma^2_1 + \sigma^2_2}{4}
-\end{eqnarray}
-
-Provided that the noise in one half of the data is uncorrelated with
-that from the other half, the signal in the jackknife map should
-resemble noise drawn from the same parent distribution as that of the
-real map. The astronomical signal, however, should be cleanly removed
-(provided that there are no strong time-varying signals, and also
-assuming that errors due to calibration between the two halves are
-insignificant). The approach we have taken to minimize systematics is
-to produce two halves of the jackknife using odd and even scan numbers
-(i.e., each map will contain a nearly evenly-spaced sampling of data
-across the full dataset).
+cutoff frequency, and residual correlated high-frequency noise above
+the cutoff is only removed through iterative common-mode
+subtraction. The map is shown in Fig.~\ref{fig:lockman_maps}a. The
+map-maker has been tested in two ways: (i) large numbers of iterations
+are used to verify that the maps converge without the growth of large
+structure; and (ii) adding synthetic sources to the data at a range of
+brightnesses verify that the map-maker response to them is
+\emph{linear} (i.e. the relative shape and amplitude compared to the
+input source is independent of brightness). The response to a
+synthetic point source (solid line) after map-making (dotted line) is
+shown in Fig.~\ref{fig:lockman_psf}. Clearly the use of a high-pass
+filter as a pre-processing step, and having no other map-constraints,
+has the down-side of introducing sidelobes around the main
+peak. However, the way this filter affects point-sources is both
+well-known, and linear.
 
@@
 
+%... high-pass
+%filter + iterative common-mode removal takes place of PCA cleaning in
+%AzTEC reduction ... probably more linear.
+
+Even though the map looks quite flat, there is a mixture of faint
+astronomical sources, and what is probably residual low-frequency
+noise causing faint ripples visible to the naked eye. Since the
+mixture of the two components is unknown, the first step is to
+suppress the low-frequency noise, under the assumption that such
+contaminants occur randomly in time, while astronomical sources are
+(usually) constant.
+
+First, the angular power spectrum of noise is estimated from a
+\emph{jackknife map}: maps are produced from two independent halves of
+the total data set, and the jackknife signal in a map pixel,
+$S_\mathrm{JK}$, and its variance, $\sigma^2_\mathrm{JK}$ are
+estimated from the two input map fluxes, $S_1$, $S_2$, and the
+corresponding variances, $\sigma^2_1$, and $\sigma^2_2$ as,
+%
+\begin{eqnarray}
+S_\mathrm{JK} = \frac{S_1 - S_2}{2} \\
+\sigma^2_\mathrm{JK} = \frac{\sigma^2_1 + \sigma^2_2}{4}.
+\end{eqnarray}
+%
+Provided that the noise in one half of the data is uncorrelated with
+that from the other half, the signal in the jackknife map should
+resemble noise drawn from the same parent distribution as that of the
+real map. The astronomical signal, however, should be cleanly removed
+(provided that there are no strong time-varying signals, and also
+assuming that errors due to calibration between the two halves are
+insignificant). The approach we have taken to minimize systematics is
+to produce the two maps using odd and even scan numbers (i.e., each
+map will contain a nearly evenly-spaced sampling of data across the
+full dataset).
+
+Since the \scuba\ scan strategy is usually isotropic (all position
+angles scanned with roughly equal weights), we make the simplifying
+assumption that the angular noise power spectrum is azimuthally
+symmetric. For these data, there are no obvious anisotropic structures
+in the 2-dimensional FFT. The radial (azimuthally-averaged) angular
+power spectrum therefore encodes all of the useful information about
+the noise properties. These power spectra for the raw output of SMURF,
+and the jackknife map (transforming only the approximately uniform
+region indicated by the square in Fig.~\ref{fig:lockman_maps}d in each
+case) are shown by the dashed black, and solid orange lines in
+Fig.~\ref{fig:lockman_pspec}, respectively. Both power spectra are
+approximately flat at spatial frequencies $\gsim 0.06$\,arcsec$^{-1}$
+(scales $\lsim16$\,arcsec), with the exception of a line at
+$\sim0.175$\,arcsec$^{-1}$ (a scale of $\sim5.7$\,arcsec). A theory
+for this feature is that it is related to the inter-bolometer spacing
+in the focal plane (which happens to be this size): small relative
+drifts in the baselines of adjacent bolometers may produce faint
+parallel stripes in the map along the scan direction (the
+superposition of many scans at different angles then results in an
+isotropic noise pattern). It is not likely that this signal is due to
+astronomical sources because it appears with a nearly equal amplitude
+in both the real map and the jackknife.  At lower spatial frequencies,
+both the real map and the jackknife power spectra grow significantly,
+as a result of the more obvious large-scale undulations in
+Figs.~\ref{fig:lockman_maps}(a) and (d).
+
 \begin{figure}
@@
 
-Since the \scuba\ scan strategy is usually isotropic (all position
-angles scanned with roughly equal weights), we make the simplifying
-assumption that the angular noise power spectrum is azimuthally
-symmetric. For these data, there are no obvious anisotropic structures
-in the 2-dimensional FFT. The radial (azimuthally-averaged) angular
-power spectrum therefore encodes all of the useful information about
-the noise properties. These power spectra for the raw output of SMURF,
-and the jackknife map (transforming only the approximately uniform
-region indicated by the square in Fig.~\ref{fig:lockman_maps}d in each
-case) by the dashed black, and solid orange lines in
-Fig.~\ref{fig:lockman_pspec}, respectively. Both power spectra are
-approximately flat at spatial frequencies $\gsim 0.06$\,arcsec$^{-1}$
-(scales $\lsim16$\,arcsec), with the exception of a narrow line
-feature at $\sim0.175$\,arcsec$^{-1}$ (a scale of
-$\sim5.7$\,arcsec). A possible theory for the feature at 5.7\,arcsec
-is that it is related to the inter-bolometer spacing in the focal
-plane (which happens to be this size); due to small relative drifts in
-the baselines of adjacent bolometers, there may be faint parallel
-stripes in the map along the scan direction (the superposition of many
-scans at different angles then results in an isotropic noise
-pattern). It is not likely that this signal is due to astronomical
-sources because it appears with a nearly equal amplitude in both the
-real map and the jackknife.  At lower spatial frequencies, both the
-real map and the jackknife power spectra grow significantly, as a
-result of the more obvious large-scale undulations in
-Figs.~\ref{fig:lockman_maps}(a) and (d).
-
-In this example, to identify point sources, we first filter the maps
-to suppress the large-scale waves. We use the radial power spectrum of
-the jackknife map (orange line in Fig.~\ref{fig:lockman_pspec}) as it
-should contain no appreciable contamination from astronomical sources.
+To suppress noise in the map, we construct a \emph{whitening filter}
+from the reciprocal of the jackknife angular power spectrum (orange
+line in Fig.~\ref{fig:lockman_pspec}), normalized by the white-noise
+level estimated from the RMS power at angular frequencies $>
+0.1$\,arcsec$^{-1}$. The filter is applied by scaling the power
+spectrum of the map by this function, and then transforming back into
+real space. The whitened map is shown in Fig.~\ref{fig:lockman_maps}b,
+and for comparison, the jackknife map has also been whitened in
+Fig.~\ref{fig:lockman_maps}e. In both cases, the maps are both
+visually flatter than the non-whitened cases.
+
+The angular power spectrum of the whitened signal map is shown with a
+solid line in Fig.~\ref{fig:lockman_pspec}. At low angular frequencies
+($\lsim 0.5$\,arcsec$^{-1}$) there is significantly less power than
+the raw map. However, it also clearly has \emph{some} power in excess
+of the white noise level. In theory, this residual signal is produced
+by astronomical sources, although its origin cannot be determined from
+this plot alone; nor are astronomical sources readily visible in the
+map (Fig.~\ref{fig:lockman_maps}b. To identify sources, we next apply
+a \emph{matched filter} to the whitened maps.
+
+For blind, high-redshift surveys, individual sources are expected to
+be un-resolved by the \scuba\ 7.5--14.5\,arcsec FWHM beams. Under this
+assumption, cross-correlation between the map and the known PSF yields
+the maximum-likelihood flux density of supposed point-sources centered
+over every location in the resulting map \citep[an extremely
+well-known result throughout astronomy, see][]{stetson1987}. Peak
+identification in such smoothed maps have been used extensively in the
+submillimetre community as both an efficient source-detection and
+photometric measurement strategy. In terms of the angular power
+spectrum, a matched-filter may also be thought of as an optimal
+\emph{low-pass} filter, that suppresses noise on scales smaller (and
+therefore higher angular frequencies) than the beam. For the case at
+hand, the effective PSF for the whitened maps is given by the dotted
+curve in Fig.~\ref{fig:lockman_psf}. Maps smoothed by this shape are
+shown for the whitened signal, and jackknife maps in
+Figures~\ref{fig:lockman_maps}c and f respectively. Note that these
+images are plotted in SNR units, where the smooth noise maps have been
+calculated by propagating the original noise maps output by SMURF both
+through the whitening and matched filters (each of which are linear
+operations).
+
+Have real astronomical sources been detected using the matched-filter?
+For both the smoothed signal and jackknife maps, orange circles denote
+3.8-$\sigma$ peaks. While not justified here, this threshold is fairly
+typical for other ground-based submillimetre surveys in recent years
+\citep[e.g.,][]{coppin2006,perera2008,weiss2009} leading to estimated
+false-identification rates of order $\sim$5\%, and is chosen here as a
+convenient reference. In the former, 10 peaks are found, whereas there
+is only 1 in the latter. However, this test does not preclude the
+possibility that some correlated noise made it into the jackknife
+map, in which case the estimated SNRs are misleading.
+
+One simple test of the calculated noise properties is to compare the
+signal and jackknife SNR distributions with ideal Gaussians. The top
+panel of Fig.~\ref{fig:lockman_hist} shows the whitened (but not match
+filtered) signal (blue) and jackknife (histograms), along with a
+Gaussian (mean 0, standard deviation 1, and area normalized to the
+number of map pixels) as a dashed line. In this case, it is clear that
+the SNR distributions for both maps are nearly indistinguishable from
+the theoretical distribution of white noise. This result shows us
+that: (i) the whitening filter appears to have removed correlated
+large-scale noise since the jackknife map histogram is consistent with
+white noise; and (ii) any potential astronomical signals are small
+compared to the typical white noise in most map pixels (unsurprising
+given the appearance of Fig.~\ref{fig:lockman_maps}b). Next, we
+examine the SNR histograms for maps processed with the matched filter
+in the bottom panel of Fig.~\ref{fig:lockman_hist}. Again, the
+histogram of the jackknife SNR data appears consistent with pure
+noise. However, the signal map now deviates significantly from a
+Gaussian distribution, with a clear \emph{positive} tail (as one would
+expect for emitting sources). In fact, integrating the positive tails
+beyond our 3.8-$\sigma$ source-detection threshold (vertical solid
+line) yields 229 map pixels in the signal map, compared with 3 in the
+jackknife map (out of 80603 pixels in the entire region). Note that
+given the small pixel size of the map, several pixels generally
+contribute to each peak.
 
@@
   Lockman Hole (second contour) in Fig.~\ref{fig:lockman_maps}a. The
-  top panel shows the histograme for pixels from the whitened signal
+  top panel shows the histograms for pixels from the whitened signal
   and jackknife maps (Figs.~\ref{fig:lockman_maps}b,e), compared to a
@@
   expectation (Gaussian), but now the matched-filter has picked out
-  significant signal leading to the large positive tail. }
+  significant signal leading to the large positive tail. The vertical
+  solid line shows the chosen 3.8-$\sigma$ source-detection
+  threshold.}
 \label{fig:lockman_hist}
@@
 
-An additional test that can be accomplished with the jackknife map, is
-the accuracy of the individual map-pixel variances. If the calculation
-is correction, the variance in each jackknife pixel,
-$\sigma^2_\mathrm{j}$, is simply the sum of the variances in the even
-and odd half-maps, $\sigma^2_\mathrm{e} + \sigma^2_\mathrm{o}$. If a
-S/N map is made from the jackknife (i.e., normalizing by
-$\sigma_\mathrm{j}$, the expected distribution is simply a Gaussian
-with mean zero, and unit standard deviation. A histogram of the
-central region of this map is shown in Fig. ???
-
-... fig. showing S/N histogram of the jackknife...
-
-Since the
-
-Both map-making itself, and the whitening filter, will suppress (and
-add sidelobes), to the nominal PSF for a point source.
+As an additional external check, we have over-plotted orange ``$+$''
+signs at the locations of 1.4\,GHz radio sources from \citet{owen2008}
+with $S\gsim15$\,$\mu$Jy. Such radio catalogues have historically
+proven invaluable for the precise identifications of high-redshift
+submillimetre galaxies due to their relatively low surface densities
+(compare with optical catalogues, for example), and a strong
+correlation between the radio and submillimetre emission mechanisms
+\citep[e.g.][]{smail2000,pope2006,ivison2007,chapin2009b}. Taking a
+representative search radius of 8\,arcsec from these studies with
+similar SNR sources and source sizes (the same size as the blue
+circles), 9 out of 10 peaks in the smoothed signal map have potential
+radio counterparts, whereas the single peak in the smoothed jackknife
+map does not lie near any radio source.
+
+Note that for significantly deeper future SCUBA-2 maps, in which the
+RMS in a PSF-smoothed map is dominated by point-source confusion,
+rather than instrumental noise, a modified matched-filter will offer
+improved results. See Appendix~A in \citet{chapin2011}, which shows
+how to include confusion (when known \emph{a priori}) explicitly as a
+noise term in the calculation of such filters.
 
 \subsection{Bright extended emission}
-The hardest thing to do with SCUBA-2 at the moment. Try to generate a
+The hardest thing to do with \scuba\ at the moment. Try to generate a
 source/background mask to assist the map solution based on thresholds

### 2012-06-01 smurf_paper.tex
first stab at extended example m17

@@
 given the small pixel size of the map, several pixels generally
-contribute to each peak.
+contribute to each peak. Note that this procedure gives only a flavour
+of the analysis that is usually required to produce a robust source
+lists. Additional tests, along with a careful consideration of
+completeness and false-positive rates usual require a series of Monte
+Carlo simulations that are beyond the scope of this work. We direct
+the interested reader to a selection of papers on the subject:
+\citet{scott2002,coppin2006,perera2008,weiss2009} and
+\citet{chapin2011}.
 
@@
 radio counterparts, whereas the single peak in the smoothed jackknife
-map does not lie near any radio source.
-
-Note that for significantly deeper future SCUBA-2 maps, in which the
-RMS in a PSF-smoothed map is dominated by point-source confusion,
-rather than instrumental noise, a modified matched-filter will offer
-improved results. See Appendix~A in \citet{chapin2011}, which shows
-how to include confusion (when known \emph{a priori}) explicitly as a
-noise term in the calculation of such filters.
+map does not lie near any radio source. Again, a proper
+cross-identification analysis must inevitably include simulations to
+establish completeness, false-positive rates, as well as the effects
+of point source clustering and confusion. See the aforementioned
+papers and references therein for examples.
+
+For future, significantly deeper SCUBA-2 maps, in which the RMS in a
+PSF-smoothed map is dominated by point-source confusion, rather than
+instrumental noise, a modified matched-filter will offer improved
+results. See Appendix~A in \citet{chapin2011}, which shows how to
+include confusion (when known \emph{a priori}) explicitly as a noise
+term in the calculation of such filters.
 
 \subsection{Bright extended emission}
-The hardest thing to do with \scuba\ at the moment. Try to generate a
-source/background mask to assist the map solution based on thresholds
-in \snr. Compare with BOLOCAM stuff.
+
+In this final example, we analyze a map of M17 which contains bright,
+extended emission, in contrast with the earlier two examples. A
+promise of \scuba\ is to recover larger angular scales than what was
+possible with earlier-generation ground-based submillimetre
+cameras. The two main factors contributing to this expectation are:
+(i) a larger array footprint on the sky which should make the
+subtraction of common sky-emission signals more effective; and (ii)
+fast detectors (with a correspondingly fast sample rate) to enable
+higher slew-rates (up to 600\,arcsec\,s$^{-1}$) to capture as much
+large-scale structure as possible above the detector $1/f$ knees. We
+produce maps of observation 11 on 20110531 using the 850\,\micron\
+array. It is a rotating PONG convering a diameter of 0.375\,deg, with
+a scan speed of 300\,arcsec\,s$^{-1}$, and a transverse spacing of
+180\,arcsec, taking 37.5\,min to complete.
+
+\begin{figure*}
+\centering
+\includegraphics[width=\linewidth]{m17.pdf}
+\caption{An 850\,\micron\ rotating pong map of M17. Intensity is
+logarithmically scaled between -0.0003\,pw (white) and +0.01\,pW
+(black). Iteration numbers are given in the corner of each
+panel. Panels (a) and (b) show the results for a reduction using the
+default parameters (the solution halted after reaching the default
+map-based convergence criterion in 30 iterations). Panel (a) also
+depicts the array footprint (position angle indicative of the start of
+the observation), and a 300\,arcsec line shows the spatial scale
+corresponding to the default \model{FLT} high-pass filter. Similar to
+Fig.~\ref{fig:pointmaps}c, the lack of prior constraints on the map
+leads to degeneracies, and therefore ripples on scales larger than the
+array footprint (due to common-mode subtraction) and filter scale
+(especially evident around the brightest peak at the centre of the
+map). Panels (c) and (d) show the ``bright extended'' reduction, in
+which a zero-mask is created iteratively from all of the pixels that
+exceed a SNR of 5. While this region (red contour) only encompasses
+the brightest peaks early in the solution, the final iteration it
+identifies most of the bright, extended emission, and it significantly
+helps with negative ringing.}
+\label{fig:m17}
+\end{figure*}
+
+The default reduction of these data is shown in the top panels of
+Fig.~\ref{fig:m17}, after 1, 30 iterations (by this last
+iteration the map-based convergence criterion was achieved). The first
+panel also depicts the array footprint, and the angular scale
+corresponding the high-pass filter edge. Much like the reduction of a
+point source without any prior constraints
+(Fig.~\ref{fig:pointmaps}c), degeneracies on scales larger than the
+array footprint and filter produce ripples in the map.
+
+Unlike the case of a known point-source (Section~\ref{sec:point}), it
+may not be possible for the observer to define, in advance, a mask to
+define regions of source and blank sky. Indeed, for this map, much of
+the field clearly contains extended structures. Furthermore, the goal
+of such maps may be to detect \emph{previously unknown} cool, dense
+regions of the ISM that may not have appeared at other wavelengths
+(e.g. the first optically-thick cloud-collapse stages of
+star-formation). While the option does exist for the user to supply
+their own mask, a facility has been added to SMURF to generate one
+automatically by flagging pixels above some SNR threshold after each
+iteration and is referred to as the ``bright extended'' configuration.
+
+The results of this automatic masking are shown in the bottom panels
+of Fig.~\ref{fig:m17}. After the first iteration, only narrow regions
+around the brightest peaks are identified. However, as the solution
+progresses, the negative bowls around the bright sources are slowly
+reduced and the mask ``grows'' out from the brightest areas. By the
+final iteration, most of the obvious structures in the data are
+encompasses by the mask.
+
+Note that alternatives to this approach do exist for other iterative
+map-makers. For example, \citet{kovacs2008} restricts the solution to
+a small number of iterations to avoid divergence on large-scales. For
+BOLOCAM data of the Galactic Plane, \citet{aguirre2010} used a
+Maximum-Entropy filtering step to suppress large-scales. The
+zero-masking implemented in SMURF is similar in principle to the
+\citet{aguirre2010}, while being slightly conceptually
+simpler. However, since both methods are non-linear, detailed
+simulations would be required to determine the SNR-dependent transfer
+function.
+
+While the reduction in the bottom panels of Fig.~\ref{fig:m17} is
+clearly superior in at least a cosmetic sense to those in the top
+panels, it is important to quantify both the noise properties of the
+maps and their response to real structures. Similar to the previous
+section, we will use a jackknife test to estimate the noise, and then
+inject known sources into the data to observe how they are attenuated.
+
+%A second map of M17 using the same observing mode was made shortly
+%after the first, observation 13 on 20110511. We produce jackknife maps
+%from the differences between the two ... angular power spectrum of the
+%noise (should be worse at large-scale for the default map).
+
+Next, we inject simulated clouds into a relative empty portion of the
+observation 11 M17 map. It is created by filtering white noise first
+with a power-law index $alpha=$bla (a reasonable approximation for
+Galactic Cirrus... cite bla bla), and then with a 14.5\,arcsec
+Gaussian to model the effect of the SCUBA-2 optical response. Finally,
+the signal is apodized ... Compare power-spectrum before and after.
+
+One fairly arbitrary parameter in our data reduction is the choice of
+angular scale for the \model{FLT} filter. The current default of
+300\,arcsec\ was chosen ``by-eye'' as a compromise that would provides
+aesthetically reasonable results for a range of observing
+modes. However, there is a simple tradeoff that warrants
+experimentation for specific fields. By choosing a larger scale (lower
+frequency in the bolometers signals) for the filter, potentially more
+structure will be faithfully re-produced in the map (up to the scale
+of the array footprint, at which point the common-mode removal will
+dominate). However, lower the filter edge can potentially let through
+more independent low-frequency noise. We have therefore re-run the
+previous tests using two other scales for the filters, and show the
+resulting maps and angular SNR spectra in Figs ... .
 
 \textit{Or now an externally supplied mask.}
-
-default: converges after 21 iterations
 

### 2012-06-01 smurf_paper.tex
text updates for extended example

@@
 The default reduction of these data is shown in the top panels of
-Fig.~\ref{fig:m17}, after 1, 30 iterations (by this last
-iteration the map-based convergence criterion was achieved). The first
-panel also depicts the array footprint, and the angular scale
-corresponding the high-pass filter edge. Much like the reduction of a
-point source without any prior constraints
+Fig.~\ref{fig:m17}, after iterations 2 (the first map estimated after
+the noise weights have been measured) and 30 iterations (when the map
+has converged). The first panel also depicts the array footprint, and
+the angular scale corresponding the high-pass filter edge. Much like
+the reduction of a point source without any prior constraints
 (Fig.~\ref{fig:pointmaps}c), degeneracies on scales larger than the
@@
 Unlike the case of a known point-source (Section~\ref{sec:point}), it
-may not be possible for the observer to define, in advance, a mask to
-define regions of source and blank sky. Indeed, for this map, much of
-the field clearly contains extended structures. Furthermore, the goal
-of such maps may be to detect \emph{previously unknown} cool, dense
-regions of the ISM that may not have appeared at other wavelengths
-(e.g. the first optically-thick cloud-collapse stages of
-star-formation). While the option does exist for the user to supply
-their own mask, a facility has been added to SMURF to generate one
-automatically by flagging pixels above some SNR threshold after each
-iteration and is referred to as the ``bright extended'' configuration.
+may not be possible for the observer to define, in advance, a mask of
+regions containing blank sky. Indeed, for this map, much of the field
+clearly contains extended structures. Furthermore, the goal of such
+maps may be to detect \emph{previously unknown} cool, dense regions of
+the ISM that may not have appeared at other wavelengths (e.g. the
+first optically-thick cloud-collapse stages of star-formation). While
+the option does exist for the user to supply their own mask, a
+facility has been added to SMURF to generate one automatically by
+flagging pixels above some SNR threshold after each iteration as part
+of the ``bright extended'' configuration.
 
 The results of this automatic masking are shown in the bottom panels
-of Fig.~\ref{fig:m17}. After the first iteration, only narrow regions
+of Fig.~\ref{fig:m17}. After the second iteration, only narrow regions
 around the brightest peaks are identified. However, as the solution
@@
 final iteration, most of the obvious structures in the data are
-encompasses by the mask.
-
-Note that alternatives to this approach do exist for other iterative
-map-makers. For example, \citet{kovacs2008} restricts the solution to
-a small number of iterations to avoid divergence on large-scales. For
-BOLOCAM data of the Galactic Plane, \citet{aguirre2010} used a
-Maximum-Entropy filtering step to suppress large-scales. The
-zero-masking implemented in SMURF is similar in principle to the
-\citet{aguirre2010}, while being slightly conceptually
-simpler. However, since both methods are non-linear, detailed
-simulations would be required to determine the SNR-dependent transfer
-function.
+encompassed by the mask, negative bowling is significantly reduced,
+and the brightest regions are clearly more extended.
 
@@
 panels, it is important to quantify both the noise properties of the
-maps and their response to real structures. Similar to the previous
-section, we will use a jackknife test to estimate the noise, and then
-inject known sources into the data to observe how they are attenuated.
+maps and their response to real structures (the transfer
+function). Similar to the previous section, we will use a jackknife
+test to estimate the noise, and then inject known sources into the
+data to observe how they are attenuated.
+
+Maps are produced using the first and second continuous halves of the
+data. This is not an ideal situation since the noise properties may
+evolve with time (e.g., due to changing sky conditions) leading to a
+biased estimate of the parent noise distribution from the
+jackknife. Also, since the zero-masking depends on the SNR of the map,
+it will be restricted to regions approximately $\sqrt{2}$ shallower in
+these maps than for the full data set. Ideally one would have many
+full maps as in the case of the Lockman Hole data in the previous
+section, from which alternating maps could be combined. However, for
+the discussion at hand our method
 
@@
 resulting maps and angular SNR spectra in Figs ... .
+
+Note that alternatives to the zero-masking approach do exist for other
+iterative map-makers. For example, \citet{kovacs2008} restricts the
+solution to a small number of iterations to avoid divergence on
+large-scales. For BOLOCAM data of the Galactic Plane,
+\citet{aguirre2010} used a Maximum-Entropy filtering step to suppress
+large-scales. The zero-masking implemented in SMURF is similar in
+principle to the \citet{aguirre2010} method, while being slightly
+conceptually simpler. However, since both methods are non-linear,
+detailed simulations would be required to determine their
+SNR-dependent angular transfer functions.
 

### 2012-06-05 smurf_paper.tex
add in final m17 plots + some of the text for noise/transfer function

@@
 
-For future, significantly deeper SCUBA-2 maps, in which the RMS in a
-PSF-smoothed map is dominated by point-source confusion, rather than
-instrumental noise, a modified matched-filter will offer improved
-results. See Appendix~A in \citet{chapin2011}, which shows how to
-include confusion (when known \emph{a priori}) explicitly as a noise
-term in the calculation of such filters.
-
-\subsection{Bright extended emission}
-
-In this final example, we analyze a map of M17 which contains bright,
-extended emission, in contrast with the earlier two examples. A
-promise of \scuba\ is to recover larger angular scales than what was
-possible with earlier-generation ground-based submillimetre
-cameras. The two main factors contributing to this expectation are:
-(i) a larger array footprint on the sky which should make the
-subtraction of common sky-emission signals more effective; and (ii)
-fast detectors (with a correspondingly fast sample rate) to enable
-higher slew-rates (up to 600\,arcsec\,s$^{-1}$) to capture as much
-large-scale structure as possible above the detector $1/f$ knees. We
-produce maps of observation 11 on 20110531 using the 850\,\micron\
-array. It is a rotating PONG convering a diameter of 0.375\,deg, with
-a scan speed of 300\,arcsec\,s$^{-1}$, and a transverse spacing of
-180\,arcsec, taking 37.5\,min to complete.
-
 \begin{figure*}
@@
 
+For future, significantly deeper SCUBA-2 maps, in which the RMS in a
+PSF-smoothed map is dominated by point-source confusion, rather than
+instrumental noise, a modified matched-filter will offer improved
+results. See Appendix~A in \citet{chapin2011}, which shows how to
+include confusion (when known \emph{a priori}) explicitly as a noise
+term in the calculation of such filters.
+
+\subsection{Bright extended emission}
+
+In this final example, we analyze a map of M17 which contains bright,
+extended emission, in contrast with the earlier two examples. A
+promise of \scuba\ is to recover larger angular scales than what was
+possible with earlier-generation ground-based submillimetre
+cameras. The two main factors contributing to this expectation are:
+(i) a larger array footprint on the sky which should make the
+subtraction of common sky-emission signals more effective; and (ii)
+fast detectors (with a correspondingly fast sample rate) to enable
+higher slew-rates (up to 600\,arcsec\,s$^{-1}$) to capture as much
+large-scale structure as possible above the detector $1/f$ knees. We
+produce maps of observation 11 on 20110531 using the 850\,\micron\
+array. It is a rotating PONG convering a diameter of 0.375\,deg, with
+a scan speed of 300\,arcsec\,s$^{-1}$, and a transverse spacing of
+180\,arcsec, taking 37.5\,min to complete.
+
 The default reduction of these data is shown in the top panels of
@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=0.8\linewidth]{m17_jk.pdf}
+\caption{Maps of M17 used to characterize the noise properties and
+transfer function of SMURF. For each column, a different high-pass
+filter edge scale was adopted (indicated in the top panels). First
+(top) row: Average of two halves of the data analyzed independently,
+using the default configuration. The data have had added to them
+synthetic signal within a 600\,arcsec diameter circle (S.E. of M17)
+created as a realization of noise from a $P(k) \propto k^{-3}$ angular
+power spectrum multiplied by the PSF, subtracting the minimum to make
+it positive, scaling it to a similar signal range as M17 itself, and
+rolling-off the edges smoothly using half a period of a radial
+(1+cosine)/2 function across 100\,arcsec beyond the edge of the
+600\,arcsec region. Second row: jackknife maps produced from the
+differences of the maps of each half of the data that went into the
+first row. Third row: average of the two halves of the data using the
+bright extended reduction. The blue and red contours indicated the
+zero-masked regions in each case (note for the 300\,arcsec case that
+the region about M17 matches fairly closely the mask in
+Fig.~\ref{fig:m17}d). Clearly as the filter scale is increased, due to
+the high SNR of the data, larger emissions regions are (correctly)
+detected and reproduced in the map. Fourth (bottom) row: Jackknife
+maps for the bright extended reductions.}
+\label{fig:m17_jk}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.49\linewidth]{pspec_m17_default.pdf}
+\includegraphics[width=0.49\linewidth]{cor_pspec_m17_default.pdf}
+\caption{Angular power spectral densities (PSDs) for the region of the
+M17 map in Fig.~\ref{fig:m17_jk} containing synthetic data, using the
+default configuration. left: raw PSDs for the input (noiseless)
+simulated data (thick black line), the signal (average of each half)
+PSDs (thin solid lines), and noise PSDs estimated from the jackknives
+(dashed lines). Vertical dotted lines indicate the high-pass filter
+scales. The red, green, and blue colours indicate data filtered at
+150, 300, and 600\,arcsec scales. right: corrected PSDs. Since the
+input PSD is known, it is possible to measure the transfer function of
+the map-maker as the ratio between the output map signal and input
+PSDs (thick solid lines). The remaining lines are as in the left
+panel, but now corrected by the transfer function. This plot shows
+that while increasing the filter scale improves the SNR at large
+scales ($\gsim100$\,arcsec), it also reduces the SNR at smaller scales
+($\lsim100$\,arcsec).}
+\label{fig:m17_def_ps}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.49\linewidth]{pspec_m17_bright_extended.pdf}
+\includegraphics[width=0.49\linewidth]{cor_pspec_m17_bright_extended.pdf}
+\caption{Lines have same meaning as in Fig.~\ref{fig:m17_def_ps},
+except now using the bright extended configuration. This configuration
+has resulted in an improved transfer function and SNR at large angular
+scales. Also note that increasing the filter scale does not have as
+large of an impact on the small-scale noise as in the default
+configuration.}
+\label{fig:m17_bc_ps}
+\end{figure*}
+
 While the reduction in the bottom panels of Fig.~\ref{fig:m17} is
@@
 panels, it is important to quantify both the noise properties of the
-maps and their response to real structures (the transfer
-function). Similar to the previous section, we will use a jackknife
-test to estimate the noise, and then inject known sources into the
-data to observe how they are attenuated.
+maps and their response to real structures (the transfer function). We
+would also like to know these are affected by our choice of filter
+scale.  Similar to the previous section, we will use a jackknife test
+to estimate the noise, and then inject known sources into the data to
+observe how they are attenuated.
 
 Maps are produced using the first and second continuous halves of the
-data. This is not an ideal situation since the noise properties may
-evolve with time (e.g., due to changing sky conditions) leading to a
-biased estimate of the parent noise distribution from the
-jackknife. Also, since the zero-masking depends on the SNR of the map,
-it will be restricted to regions approximately $\sqrt{2}$ shallower in
-these maps than for the full data set. Ideally one would have many
-full maps as in the case of the Lockman Hole data in the previous
-section, from which alternating maps could be combined. However, for
-the discussion at hand our method
+data in Fig.~\ref{fig:m17_jk}. This is not an ideal situation since
+the noise properties may evolve with time (e.g., due to changing sky
+conditions) leading to a biased estimate of the parent noise
+distribution from the jackknife. Also, since the zero-masking depends
+on the SNR of the map, it will be restricted to regions approximately
+$\sqrt{2}$ shallower in these maps than for the full data
+set. Finally, the cross-linking (positional angles sampled) is
+similar, though not identical in the two halves. Ideally one would
+have many full maps as in the case of the Lockman Hole data in the
+previous section, from which alternating maps could be combined.
 
@@
 
-Next, we inject simulated clouds into a relative empty portion of the
-observation 11 M17 map. It is created by filtering white noise first
-with a power-law index $alpha=$bla (a reasonable approximation for
-Galactic Cirrus... cite bla bla), and then with a 14.5\,arcsec
-Gaussian to model the effect of the SCUBA-2 optical response. Finally,
-the signal is apodized ... Compare power-spectrum before and after.
+In addition to dividing the data into two halves, we also inject
+simulated clouds into a relative empty portion of the map. It is
+created by drawing a realization of noise within an 800\,arcsec
+on-a-side box, from an angular power spectrum $P(k) \propto k^{-3}$
+which is appropriate for Galactic cirrus clouds
+\citep[e.g.][]{gautier1992}. It is then filtered again with a
+14.5\,arcsec Gaussian to model the effect of the SCUBA-2 optical
+response. The RMS of these fluctuations are then normalized to 0.002
+pW so that they are comparable to the dynamic range of M17 itself. The
+minimum is subtracted so that the signal is positive. Finally, half a
+period of a (1+cosine)/2 function is used to roll-off the signal to
+zero between radii of 300 and 400 arcsec.
+
+The first row of Fig.~\ref{fig:m17_jk} shows the total signal image
+averaging the maps made of each independent half of the data, for the
+default configuration (inverse-variance weighting has been used). The
+columns show reductions using 150, 300, and 600\,arcsec filter edges
+in the reductions. The synthetic data are clearly seen as the circular
+region S.E. of M17. As the filter scale is increased, the size of the
+ripples increases accordingly. While larger structures do seem to
+appear, negative bowls are a major problem without any other map
+constraints.The second row of Fig.~\ref{fig:m17_jk} shows the
+jackknife maps. The astronomical emission is almost perfectly removed,
+except for a slight impression of M17 near the centre of the map that
+is probably due to some combination of detector gain and pointing
+drifts.
 

### 2012-06-06 smurf_paper.tex
re-made point source maps, update text for change to 100 from 50 iterations

@@
 shown in Fig.~\ref{fig:pointmaps}a. While the S/N of Uranus is clearly
-large ($\sim$175 as estimated by SMURF), enabling us to see the faint
+large ($\sim$250 as estimated by SMURF), enabling us to see the faint
 sidelobes (circle and cross pattern), the map also has obvious
@@
 bolometer leaves a trace of the circular scan pattern in the map as
-each of their baselines slowly drift independently. Most of the
-larger-scale ripples in the map, however, are probably due to
-degeneracies in the map solution as we discuss below.
+each of their baselines slowly drift independently. A significant
+contribution to the larger-scale ripples in the map, however, can be
+made by degeneracies in the map solution as we discuss below.
 
 To illustrate how large-scale ripples can form (and grow), the same
-map solution is run for 50 iterations and shown in
+map solution is run for 100 iterations and shown in
 Fig.~\ref{fig:pointmaps}b, now exhibiting a strong vertical
@@
 Fig.~\ref{fig:degeneracy} shows the residual signals for a single
-bolometer after 2 (black) and 50 (grey) iterations which are nearly
+bolometer after 2 (black) and 100 (grey) iterations which are nearly
 identical, yet the \emph{change} in the estimated \model{COM} (green)
-and \model{AST} (red) signals between 2 and 50 iterations are
+and \model{AST} (red) signals between 2 and 100 iterations are
 large. However, it is also clear that the estimated \model{COM} and
@@
 estimation. We set the filter edge based on an angular scale of
-200\,arcsec, which given the scan speed of 155\,arcsec\,sec$^{-1}$,
+200\,arcsec, which, given the scan speed of 155\,arcsec\,sec$^{-1}$,
 corresponds to a frequency of 0.78\,Hz. Now using the automatic
@@
 bright point-like source, the result is an azimuthally-symmetric
-$\sinc$ function-like pattern in the map about Uranus. Note that due
-to the iterative nature of the solution, much of the signal in the
-source is actually removed for subsequent iterations of \model{FLT},
-but as in the previous case, degeneracies between the map and the
-\model{FLT} and \model{COM} signal components can still lead to
-complicated structures.
+$\sinc$ function-like pattern in the map. Note that due to the
+iterative nature of the solution, much of the signal in the source is
+actually removed for subsequent iterations of \model{FLT}, but as in
+the previous case, degeneracies between the map and the \model{FLT}
+and \model{COM} signal components can still lead to complicated
+structures.
 
@@
 Uranus (much larger than the FWHM of the main lobe), for all but the
-final iteration. In this case, the map converges after 22 iterations,
+final iteration. In this case, the map converges after 6 iterations,
 and all of the previous artifacts have been effectively removed. Since
@@
   circular streaks are caused by independent low-frequency noise that
-  is not removed by the common-mode (Uranus peak 0.14\,pW); (b) same
+  is not removed by the common-mode (Uranus peak 0.15\,pW); (b) same
   as (a) but using 50 iterations, illustrating the degeneracy between
   large-scale structure and the common-mode (the footprint of working
-  bolometers is also shown for reference, Uranus peak 0.14\,pW); (c)
+  bolometers is also shown for reference, Uranus peak 0.15\,pW); (c)
   reduction in which high-pass filtering above 0.775\,Hz
@@
   after 10 iterations, but leaving large-scale ringing due to the
-  filter (Uranus peak 0.38\,pW); (d) same as (c), but now the region
+  filter (Uranus peak 0.27\,pW); (d) same as (c), but now the region
   of the map beyond the white dot-dashed circle is constrained to a
@@
   extremely flat, there is little attenuation of the source flux, and
-  the diffraction pattern is clearly seen (Uranus peak 0.38\,pW).}
+  the diffraction pattern is clearly seen (Uranus peak 0.27\,pW).}
 \label{fig:pointmaps}
@@
   lines in the top panel show the residual signal for a single
-  bolometer after 2 and 50 iterations respectively (they are nearly
+  bolometer after 2 and 100 iterations respectively (they are nearly
   identical). The green and red lines show the \emph{difference}
@@
   current map estimate for a given bolometer) model components for
-  that bolometer between iterations 2 and 50 respectively. This shows
+  that bolometer between iterations 2 and 100 respectively. This shows
   that a strong signal has grown over time, and it has equal but

### 2012-06-11 smurf_paper.tex
finish draft + re-made extended plots. Also minor updates throughout paper

@@
 rectangular subarrays, consisting of $40 \times 32$ bolometers each,
-and subtend a nearly contiguous $4' \times 4'$ footprint on the sky.
+and subtend a nearly contiguous $7' \times 7'$ footprint on the sky.
 Ultimately, the instrument will also be equipped with a Fourier
@@
     \model{DKS} & dark squid cleaning along columns \\
+    \model{PLN} & 2-dimensional time-varying plane removal \\
     \model{SMO} & time-domain smoothing filter \\
-    \model{PLN} & 2-dimensional time-varying plane removal \\
+    \model{TMP} & pointing as baseline template \\
     \hline
@@
 
-\subsubsection{\model{DKS}: dark squid cleaning along columns}
-\label{sec:dks}
+\subsubsection{Other experimental models}
+
+We don't use them, but we should mention them since a reviewer might
+ask us if we tried them: \model{DKS}, \model{TMP}, \model{SMO},
+\model{PLN}.
 
@@
 gaps.}
-
-\subsubsection{\model{TMP}: remove a scaled template}
 
@@
 
-\subsubsection{\model{SMO}: time-domain smoothing filter}
-\label{sec:smo}
-
-TJ
-
 We also have the SMO filter that's useful for very short scans
@@
 more robust, or mean which is more accurate.
-
-\subsubsection{\model{PLN}: 2d time-varying sky noise plane removal}
-\label{sec:pln}
-
-TJ
 
@@
 
+Another major source of divergence is correlation between \model{COM}
+and \model{FLT}. Since \model{FLT} usually consists of a high-pass
+filter following the application of \model{COM}, \model{COM} is
+completely free to grow any large-scale structure at frequencies below
+the chosen filter edge. While such structure does not appear in the
+map (as it is removed by \model{FLT}), we found that the solution
+could be made to converge significantly faster by ``re-mixing''
+\model{COM} and \model{FLT} at the start of each iteration. In
+practice, at the start of each iteration, the previous iterations of
+each model are both added back into the residual immediately prior to
+re-calculating \model{COM}. In this way, truly common-mode signals,
+even at low-frequencies, do not leak into \model{FLT}.
+
 %-------------------------------------------------
@@
 
-\subsubsection{Whitening filter}
-
-EC
+\subsubsection{Bolometer Filtering}
+
+Despite the ability of the map-maker to iterative remove many noise
+components, under some circumstances it may be desirable to filter the
+data \emph{once} during the pre-processing step. Three main filtering
+options are available:
+
+\begin{itemize}
+
+\item The most commonly-used pre-processing filter is polynomial
+subtraction. A polynomial of the requested order is fit and removed
+from each bolometer time-series. At a bare minimum, the mean is
+removed from all of the bolometers (order 0) in all reductions
+described in this paper.
+
+\item All of the filters available as part of the iterative Fourier
+Transform Filter (Section~\ref{sec:flt}) can also be applied once
+during pre-processing.
+
+\item As a complete alternative to the iterative map-making procedure,
+cleaning by principal component analysis (PCA) is available as an
+experimental pre-processing option. As described in
+Section~\ref{sec:pca}, a new set of basic vectors for the bolometer
+data are found which diagonalize their covariance matrix. The
+brightest components can then be subtracted from the data. A single
+parameter specifies the threshold on the amplitude of the eigenvalues
+to be removed, as a number of standard deviations away from the mean
+value. Subarrays are cleaned independently once for the full length of
+each continuousy chunk of data. Given the computational expense of
+this method, and initial tests which showed little improvement over
+simple high-pass filtering, this method has not yet been explored in
+detail with SCUBA-2 data.
+
+\end{itemize}
 
@@
 
-In this section three different types of data are reduced. In each
-case, the baseline reduction using defaults as described in
-Section~\ref{sec:algorithm} are used first. Then, modifications are
-described to improve each reduction.
-
+In this section three different common types of data are reduced: a
+bright point source, Uranus (Section~\ref{sec:point}); a bind survey
+of high-redshift galaxies in the Lockman Hole
+(Section~\ref{sec:cosmo}); and a map of a bright extended star-forming
+region in our Galaxy, M17 (Section~\ref{sec:extended}). In each case,
+variations on the default algorithm from Section~\ref{sec:algorithm}
+are described.
+
+%-------------------------------------------------
 \subsection{Known point source}
 \label{sec:point}
+%-------------------------------------------------
 
@@
 
+%-------------------------------------------------
 \subsection{Deep blind survey}
 \label{sec:cosmo}
+%-------------------------------------------------
 
@@
 
+%-------------------------------------------------
 \subsection{Bright extended emission}
+\label{sec:extended}
+%-------------------------------------------------
 
@@
 produce maps of observation 11 on 20110531 using the 850\,\micron\
-array. It is a rotating PONG convering a diameter of 0.375\,deg, with
-a scan speed of 300\,arcsec\,s$^{-1}$, and a transverse spacing of
+array. It is a rotating PONG covering a diameter of 0.375\,deg, with a
+scan speed of 300\,arcsec\,s$^{-1}$, and a transverse spacing of
 180\,arcsec, taking 37.5\,min to complete.
@@
 Fig.~\ref{fig:m17}, after iterations 2 (the first map estimated after
-the noise weights have been measured) and 30 iterations (when the map
+the noise weights have been measured) and 17 iterations (when the map
 has converged). The first panel also depicts the array footprint, and
-the angular scale corresponding the high-pass filter edge. Much like
-the reduction of a point source without any prior constraints
+the angular scale corresponding to the high-pass filter edge. Much
+like the reduction of a point source without any prior constraints
 (Fig.~\ref{fig:pointmaps}c), degeneracies on scales larger than the
@@
 \centering
-\includegraphics[width=0.8\linewidth]{m17_jk.pdf}
+\includegraphics[width=\linewidth]{m17_jk.pdf}
 \caption{Maps of M17 used to characterize the noise properties and
-transfer function of SMURF. For each column, a different high-pass
-filter edge scale was adopted (indicated in the top panels). First
-(top) row: Average of two halves of the data analyzed independently,
-using the default configuration. The data have had added to them
-synthetic signal within a 600\,arcsec diameter circle (S.E. of M17)
-created as a realization of noise from a $P(k) \propto k^{-3}$ angular
-power spectrum multiplied by the PSF, subtracting the minimum to make
-it positive, scaling it to a similar signal range as M17 itself, and
+transfer function of SMURF (same intensity scale as
+Fig.~\ref{fig:m17}). For each column, a different high-pass filter
+edge scale was adopted (indicated in the top panels). First (top) row:
+Average of two halves of the data analyzed independently, using the
+default configuration. The data have had added to them synthetic
+signal within a 600\,arcsec diameter circle (S.E. of M17) created as a
+realization of noise from a $P(k) \propto k^{-3}$ angular power
+spectrum multiplied by the PSF, subtracting the minimum to make it
+positive, scaling it to a similar signal range as M17 itself, and
 rolling-off the edges smoothly using half a period of a radial
@@
 (dashed lines). Vertical dotted lines indicate the high-pass filter
-scales. The red, green, and blue colours indicate data filtered at
-150, 300, and 600\,arcsec scales. right: corrected PSDs. Since the
-input PSD is known, it is possible to measure the transfer function of
-the map-maker as the ratio between the output map signal and input
-PSDs (thick solid lines). The remaining lines are as in the left
-panel, but now corrected by the transfer function. This plot shows
-that while increasing the filter scale improves the SNR at large
-scales ($\gsim100$\,arcsec), it also reduces the SNR at smaller scales
-($\lsim100$\,arcsec).}
+scales: 150\,arcsec (red); 300\,arcsec (orange); 600\,arcsec (green);
+and 900\,arcsec (blue). Since the input PSD is known, it is possible
+to measure the transfer function of the map-maker as the ratio between
+the output map signal PSDs and the input PSD (from the left panel),
+giving the thin coloured lines (linear vertical axis shown on right of
+plot). The remaining lines are as in the left panel, but now corrected
+by the transfer function. This plot shows that increasing the filter
+scale improves the SNR at intermediates scales
+($\sim$200--600\,arcsec), although the SNR worsens at smaller scales
+($\lsim$200\,arcsec).}
 \label{fig:m17_def_ps}
@@
 large of an impact on the small-scale noise as in the default
-configuration.}
-\label{fig:m17_bc_ps}
+configuration, although it is still noticeably larger up to scales of
+about $\sim$60\,arcsec.}
+\label{fig:m17_be_ps}
 \end{figure*}
@@
 maps and their response to real structures (the transfer function). We
-would also like to know these are affected by our choice of filter
+would also like to know how each are affected by our choice of filter
 scale.  Similar to the previous section, we will use a jackknife test
-to estimate the noise, and then inject known sources into the data to
-observe how they are attenuated.
+to estimate the noise, as well as injecting known sources into the
+data to observe how they are attenuated.
 
@@
 conditions) leading to a biased estimate of the parent noise
-distribution from the jackknife. Also, since the zero-masking depends
-on the SNR of the map, it will be restricted to regions approximately
-$\sqrt{2}$ shallower in these maps than for the full data
-set. Finally, the cross-linking (positional angles sampled) is
-similar, though not identical in the two halves. Ideally one would
+distribution in the complete map from the jackknife. Also, since the
+zero-masking depends on the SNR of the map, it will be restricted to
+regions approximately $\sqrt{2}$ shallower in these maps than for the
+full data set. Finally, the cross-linking (positional angles sampled)
+is similar, though not identical in the two halves. Ideally one would
 have many full maps as in the case of the Lockman Hole data in the
@@
 
-%A second map of M17 using the same observing mode was made shortly
-%after the first, observation 13 on 20110511. We produce jackknife maps
-%from the differences between the two ... angular power spectrum of the
-%noise (should be worse at large-scale for the default map).
-
-In addition to dividing the data into two halves, we also inject
-simulated clouds into a relative empty portion of the map. It is
-created by drawing a realization of noise within an 800\,arcsec
-on-a-side box, from an angular power spectrum $P(k) \propto k^{-3}$
+Since our goal in this section is to measure the response of the
+map-maker to extended structures, we inject a simulated signal with
+power at a range of scales to a relatively empty region of the map.
+It is created by drawing a realization of noise within an 800\,arcsec
+on-a-side box, with an angular power spectrum $P(k) \propto k^{-3}$
 which is appropriate for Galactic cirrus clouds
@@
 default configuration (inverse-variance weighting has been used). The
-columns show reductions using 150, 300, and 600\,arcsec filter edges
-in the reductions. The synthetic data are clearly seen as the circular
-region S.E. of M17. As the filter scale is increased, the size of the
-ripples increases accordingly. While larger structures do seem to
-appear, negative bowls are a major problem without any other map
-constraints.The second row of Fig.~\ref{fig:m17_jk} shows the
+columns show reductions using 150, 300, 600, and 900\,arcsec filter
+edges in the reductions. For reference, the largest scale that is
+completely inscribed by the array footprint is about 400\,arcsec, and
+the diagonal of the array is about 600\,arcsec (meaning that scales
+beyond 600\,arcsec are completely unconstrained when common-mode
+rejection is used).  The synthetic data are clearly seen as the
+circular region S.E. of M17. As the filter scale is increased, the
+size of the ripples increases accordingly. While larger structures do
+seem to appear, negative bowls are a major problem without any other
+map constraints.The second row of Fig.~\ref{fig:m17_jk} shows the
 jackknife maps. The astronomical emission is almost perfectly removed,
@@
 is probably due to some combination of detector gain and pointing
-drifts.
-
-One fairly arbitrary parameter in our data reduction is the choice of
-angular scale for the \model{FLT} filter. The current default of
-300\,arcsec\ was chosen ``by-eye'' as a compromise that would provides
-aesthetically reasonable results for a range of observing
-modes. However, there is a simple tradeoff that warrants
-experimentation for specific fields. By choosing a larger scale (lower
-frequency in the bolometers signals) for the filter, potentially more
-structure will be faithfully re-produced in the map (up to the scale
-of the array footprint, at which point the common-mode removal will
-dominate). However, lower the filter edge can potentially let through
-more independent low-frequency noise. We have therefore re-run the
-previous tests using two other scales for the filters, and show the
-resulting maps and angular SNR spectra in Figs ... .
+drifts. Otherwise the jackknife appears to be a plausible realization
+of noise from the same parent distribution as for the averaged maps.
+
+The third and fourth rows in Fig.~\ref{fig:m17_jk} repeat this
+exercise using the bright extended configuration, in which the SNR
+threshold of 5 is again used to constrain the map. As the filter scale
+is increased, more of the extended structure in M17 is reproduced in
+the map, as evidenced by the blue and red contours (masks generated
+from the first and second halves of the data, respectively). The
+masking does a generally good job of suppressing the largest-scale
+ripples that are produced by the default reduction. However, the noise
+away from regions of bright emission does increase noticeably (mottled
+appearance). With filter scales $\gsim600$\,arcsec, virtually the
+entire region of synthetic emission is correctly identified by the SNR
+mask and allowed to vary freely in the solution.
+
+Next, we analyze the angular power spectra densities (PSDs) of the
+maps to understand the signal and noise properties of the map-maker in
+the region of synthetic sources, as a function of filter scale. The
+left panel of Fig.~\ref{fig:m17_def_ps} shows the raw PSDs for the
+input synthetic signal (thick black line), the output map signals
+(thin solid lines), and the jackknife maps (dashed lines). Colours
+encode the filter scales used: 150\,arcsec (red); 300\,arcsec
+(orange); 600\,arcsec (green); and 900\,arcsec (blue). Note that, in
+the exception of the synthetic data, we only plot the PSDs down two
+the second-lowest spatial frequency bin of
+$1.875\times10^{-3}$\,arcsec$^{-1}$, or 533\,arcsec, due to the fact
+that the smallest is very poorly sampled, and therefore noisy. As the
+filter edge is increased, more power is measured in the map PSDs at
+larger scales. However, much of this power is clearly produced by
+noise which appears in the jackknife PSDs. We estimate the map-maker
+transfer function as the ratio between the portion of the signal PSDs
+\emph{not} produced by noise, to the input PSD, or
+%
+\begin{equation}
+G(k) = \frac{P_\mathrm{M}(k) - P_\mathrm{JK}(k)}{P_\mathrm{S}(k)},
+\end{equation}
+%
+where $k$ is the spatial frequency, and the subscripts ``M'', ``JK'',
+and ``S'' refer to the signal map, jackknife map, and synthetic map,
+respectively.
+
+The transfer functions $G(k)$ are plotted as thin solid lines in the
+right panel of Fig.~\ref{fig:m17_def_ps}. This formula produced
+extremely noisy values at large frequencies (small scales), and was
+set to a value of 1 above 0.015\,arcsec$^{-1}$, as well as any point
+in the curve where $P_\mathrm{M}(k)$ exceeded $P_\mathrm{S}(k)$
+(i.e. $G(K)$ is assumed to be $\le$1). As expected, the larger the
+scale of the filter, the lower the spatial frequency at which the map
+transfer function falls. We then correct the map and jackknife PSDs by
+dividing by $G(k)$ to produce the thick solid, and dashed lines in the
+right panel of Fig.~\ref{fig:m17_def_ps}. This shows us that, even
+though the raw noise in the left panel is lower at small scales when a
+smaller-scale filter is used (e.g. the red dashed line), once we
+correct for the transfer function, we actually win in a SNR sense at
+large scales using the larger-scale filter as you would expect.
+
+These tests are then repeated using the bright extended reduction in
+Fig.~\ref{fig:m17_be_ps}. The most obvious improvement with this
+reduction over the default reduction is that the transfer functions
+fall more slowly at large angular scales, accompanied by a slower
+increase in noise; in other words, there is greatly improved SNR at
+large angular scales (an obvious conclusion given the appearence of
+the maps in Fig.~\ref{fig:m17_jk}). In fact, using the 900\,arcsec
+filter edge, the map response is still above 80\% right out to the
+largest scale accurately measured in the PSDs, 533\,arcsec, which is
+about the largest scale that should be recoverable given the size of
+the array footprint and the fact that we use common-mode rejection.
+Another interesting feature of these reductions is that the increase
+in small-scale noise as the filter edge is increased is not as drastic
+as in the default reduction, and the SNR improves at scales
+$\gsim$60\,arcsec.
+
+One case in which the SNR is \emph{worse} using the bright extended
+reduction is using a 150\,arcsec filter. In this case the noise is
+considerably larger in the bright extended reduction as evidenced by
+the ``kink'' near 150\,arcsec. Referring to the mask contours in the
+left panel of the third row, it is clear that the map-maker has failed
+to identify much of the bright, extended emission in the region of the
+synthetic source. Each area that is not within the contours is
+constrained to zero throughout the solution, therefore suppressing
+power (and lowering the transfer function), and subsequently reducing
+the SNR of the final result. This measurement serves as a warning: the
+map-maker response is non-linear when using SNR masking. Harsh
+filtering can provide misleading results as in this example. Maps of
+\emph{faint} extended emission will also suffer considerably, as the
+structures of interest may not clear the SNR threshold for the mask.
 
 Note that alternatives to the zero-masking approach do exist for other
-iterative map-makers. For example, \citet{kovacs2008} restricts the
-solution to a small number of iterations to avoid divergence on
-large-scales. For BOLOCAM data of the Galactic Plane,
-\citet{aguirre2010} used a Maximum-Entropy filtering step to suppress
-large-scales. The zero-masking implemented in SMURF is similar in
-principle to the \citet{aguirre2010} method, while being slightly
-conceptually simpler. However, since both methods are non-linear,
-detailed simulations would be required to determine their
-SNR-dependent angular transfer functions.
-
-\textit{Or now an externally supplied mask.}
+iterative map-makers. For example, \citet{kovacs2008} typically
+restricts the solution to a small fixed number of iterations (of order
+$\sim$10), although we note that there is no guarantee that maps will
+have reasonably converged with such a strategy. For BOLOCAM data of
+the Galactic Plane, \citet{aguirre2010} used a Maximum-Entropy
+filtering step to suppress large-scales. As in the case of the
+SNR-based masking described here, Maximum-Entropy filtering is also
+non-linear, meaning that it too requires simulations to establish its
+transfer function and noise properties as a function of angular scale.
 
@@
 
-\scuba\ is being re-commissioned. Maybe we will have some initial data
-on fridge performance? If both it, and sky noise are relatively
-well-behaved, might be able to restrict filtering to lower frequencies
-and get larger-scale structures.
-
-Larger array footprint also helps.
-
-Talk about potential of Pascale et al. method to get closer to
-max-likelihood map solution (handle maps at constant scan angles with
-weights in Fourier space separately).
+\begin{enumerate}
+
+\item SMURF is an iterative map-maker that can produce SCUBA-2 maps on
+single high-end modern desktops in a reasonable amount of time.
+
+\item A simple model which removes most noise in a common-mode
+rejection step, along with high-pass filtering, results in maps that
+approach the white-noise performance of the detectors.
+
+\item The main problem to overcome with our strategy is the divergence
+of the map on large angular scales, due to the degeneracy between the
+map, common-mode, and other residual low-frequency noise. A simple
+strategy of constraining empty regions of the map to zero (either a
+user-supplied mask for known sources, or an iterative determination of
+signal above some SNR threshold) provides good constraints for compact
+objects, and bright/extended structures.
+
+\item For maps of faint point sources, a single (non-iterative)
+high-pass filter at the start of the reduction produces maps that are
+nearly white-noise limited. We illustatre how to completely whiten the
+maps using jackknife measurements of the angular noise power spectrum
+in order to reliably extract sources down to the limits of the map.
+
+\item The map-maker has an automatic stopping criterion based on the
+change in the map and $\chi^2$ between subsequent iterations, allowing
+it to adapt to a wide range of data in a pipeline setting.
+
+\item It is believed that a maximum-likelihood map-maker, such as
+SANEPIC, will provide improved maps over the iterative algorithm
+described here. However, SMURF will most likely be used as an initial
+step in such processing since it can quickly clean the bolometer
+time-series, as well as perform map-based despiking (a necessarily
+iterative step).
+
+\end{enumerate}
+
+%\scuba\ is being re-commissioned. Maybe we will have some initial data
+%on fridge performance? If both it, and sky noise are relatively
+%well-behaved, might be able to restrict filtering to lower frequencies
+%and get larger-scale structures.
+
+%Larger array footprint also helps.
+
+%Talk about potential of Pascale et al. method to get closer to
+%max-likelihood map solution (handle maps at constant scan angles with
+%weights in Fourier space separately).
 
@@
 
-\textit{Presumably the standard telescope acknowledgment here plus a thank you
-to CFI for funding the DR work. Ed should probably thank both CANFAR
-and JAC for additionally funding although his JAC funding is noted in
-the affiliation.}
+\textit{Presumably the standard telescope acknowledgment here plus a
+thank you to CFI for funding the DR work. Ed should probably thank
+both CANFAR and JAC for additionally funding although his JAC funding
+is noted in the affiliation.}
 

### 2012-06-12 smurf_paper.tex
abstract and intro work

@@
 
-\title[SMURF: an iterative map-maker for SCUBA-2]{The Sub-Millimetre User
-Reduction Facility: an iterative map-maker for SCUBA-2}
+\title[SMURF: an iterative map-maker for \scuba]{The Sub-Millimetre User
+Reduction Facility: an iterative map-maker for \scuba}
 
@@
   We describe the properties of data from the Submillimetre Common
-  User Bolometer Array 2 (\scuba) taken during the Shared Risk
-  Observing period from 2010 February 11 to 2010 March 25, and the
-  production of maps using the Sub-Millimetre User Reduction Facility
-  (SMURF). Our iterative approach is similar to those used to reduce
-  data from other ground-based submillimetre telescopes in recent
-  years, and recovers angular scales up to the size of the array
-  footprint (approximately 2\,arcmin). The method is also able to
-  achieve the theoretical white-noise limit of the instrument for
-  point-source studies.
+  User Bolometer Array 2 (\scuba), a camera for the 15-m James Clerk
+  Maxwell Telescope that provides simultaneous imaging capabilities at
+  450 and 850\,\micron, and the production of maps using the
+  Sub-Millimetre User Reduction Facility (SMURF). With nearly 5000
+  bolometers in each imaging band, and an approximately 5\,arcmin
+  field-of-view, \scuba\ is presently the largest submillimetre camera
+  in the world, and produces data at rates an order-of-magnitude
+  faster than previous instruments. We have adopted a fast, iterative
+  approach to map-making that enables data reduction on single,
+  modern, high-end desktop computers with execution times that are
+  shorter than the observing times.  SMURF is used in an automated
+  setting both at the telescope for real-time feedback to observers,
+  as well as for the production of generic science products at the
+  Canadian Astronomical Data Centre. A suite of tools are also
+  provided for interactive data analysis.  Three detailed case studies
+  are used to: (i) explore convergence properties of the map-maker
+  using simple prior constraints (Uranus -- a point source); (ii)
+  demonstrate that our strategy is capable of recovering angular
+  scales up to the size of the array footprint (approximately
+  5\,arcmin) for bright extended sources (star-forming region M17);
+  and (iii) achieve the white-noise limit of the instrument for faint
+  point-source studies (blank-field survey of the Lockman Hole).
 \end{abstract}
@@
 simultaneously images the sky in two broad bands centered over 450 and
-850\,\micron. The focal planes at each wavelength are populated with 4
-rectangular subarrays, consisting of $40 \times 32$ bolometers each,
-and subtend a nearly contiguous $7' \times 7'$ footprint on the sky.
-Ultimately, the instrument will also be equipped with a Fourier
-transform spectrometer \citep[FTS-2,][]{gom2010}, and a polarimeter
-\cite[POL-2,][]{bastien2005}. This paper describes the Submillimetre
-User Reduction Facility, SMURF, a software package for reducing the
-imaging data. The reduction of FTS-2 and POL-2 data will be described
-at a later date once these additional instruments have been
-commissioned. The details of the instrument design, performance, and
-its calibration are given in two companion papers: \citet{holland2012}
-and \citet{dempsey2012}.
+850\,\micron, with approximately 7.5 and 14.5\,arcsec full-width at
+half-maximum (FWHM) point spread functions (PSFs). The focal planes at
+each wavelength are populated with 4 rectangular subarrays, consisting
+of $40 \times 32$ bolometers each, and subtend a nearly contiguous $7'
+\times 7'$ footprint on the sky.  Ultimately, the instrument will also
+be equipped with a Fourier transform spectrometer
+\citep[FTS-2,][]{gom2010}, and a polarimeter
+\cite[POL-2,][]{bastien2005}. This paper describes the properties of
+\scuba\ data that are relevant to producing maps of the imaging data,
+and the Submillimetre User Reduction Facility, SMURF, a software
+package for perfoming the reduction. The analysis of FTS-2 and POL-2
+data will be described at a later date once these additional
+instruments have been commissioned. The details of the instrument
+design, performance, and its calibration are given in two companion
+papers: \citet{holland2012} and \citet{dempsey2012}.
 
 Over the last twenty years, observations in the submillimetre band
-(defined here to be 200--1000\,\micron) have helped revolutionize
+(defined here to be 200--1200\,\micron) have helped revolutionize
 several important areas of astrophysics, including: the
@@
 
-Most submillimetre imaging cameras, to date, have used bolometers to
-maximize sensitivities \citep[e.g.][]{griffin2002}. Each pixel
-consists of a thermal absorber to collect photons that have passed
-through broad-band filters, and the signal is read-out using a
-thermometer. The high-frequency broad-band noise in modern systems,
-such as \scuba, is limited by photon noise from the instrument and
+Submillimetre imaging cameras generally use bolometers to maximize
+sensitivity, in which an absorber collects photons that pass through
+broad-band filters, and the signal is read out-out using a thermometer
+\citep[e.g.][]{griffin2002}. The sensitivity of bolometers in a camera
+like \scuba\ are limited by photon noise from the instrument and
 ambient backgrounds. The low-frequency noise, however, is dominated by
-sources such as slow variations in mean backgrounds (such as thermal
+sources such as slow variations in mean backgrounds (e.g., thermal
 variations within the cryostat, and the atmosphere for ground-based
-cameras), and drift in the readout electronics. Since the
-low-frequency noise is largely correlated between all, or subsets, of
-the bolometers in time, it can be suppressed during map-making, since
-astronomical signals have the distinct property that they are fixed in
-a sky reference frame (assuming they are not time-varying). If
-successful, the noise in the resulting map is said to be ``background
-noise limited'', meaning that the only significant remaining noise
-sources are uncorrelated spatially in the map (the theoretical limit).
-
-There are numerous ways to attack this problem, both in terms of the
-data-collection method, and data processing. Historically, both radio
-and submillimetre telescope systems have used ``chopping'' to perform
-rapid differences between target sky locations and some reference
-(e.g., another point on the sky, internal calibrators, and different
-frequencies in the case of heterodyne systems). In fact, spatial
-chopping was used by SCUBA, the predecessor to \scuba, to remove
-low-frequency drifts \citep{holland1999}. However, chopping (in
-particular spatial chopping) has the unfortunate side-effect of
-filtering out information on scales larger than the chop-throw
-\textbf{(references)}.For this reason, modern cameras like \scuba\ aim
-to provide a high-degree of inherent detector stability, and rapidly
-slew the telescope so that astronomical signals up to large scales are
-captured in a signal band at considerably higher frequency than the
-residual low-frequency drift.
-
-In the case of chopped data, ideally the resulting data are completely
-free of low-frequency noise, and therefore the data samples are
-statistically uncorrelated.
-
-often used various
-``chopping'' methods to remove low-frequency drifts -- rapid
-differences are made between the sky signal at a given position and an
-internal reference signal; the sky signal in two different bands for
-heterodyne systems (frequency chopping); and a different position in
-the sky \citep[a method employed by SCUBA, the predecessor of
-\scuba][]{holland1999}.
-
- many of them evolving
-out of techniques for producing maps of the large-scale anisitropies
-in the Cosmic Microwave Background (CMB)
+cameras), and drift in the readout electronics. Such noise has a power
+spectrum $\propto 1/f^\alpha$ ($\alpha>0$), and the frequency at which
+it is comparable to the white noise level is called the ``$1/f$
+knee''. Since the low-frequency noise is largely correlated between
+all, or subsets, of the bolometers in time, it can be suppressed
+during map-making, since astronomical signals have the distinct
+property that they are fixed in a sky reference frame (assuming they
+are not time-varying). If successful, the noise in the resulting map
+is said to be ``background noise limited'', meaning that the map noise
+is uncorrelated spatially, and has an amplitude that scales as the
+$\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the noise-equivalent
+flux-density (the white noise level of a bolometer in 1\,s of
+integration), and $t$ is the amount of integration time in a map
+pixel.
+
+There are numerous ways to attack the problem of map production, both
+in terms of the data-collection method, and processing. The two most
+important principles to follow in terms of scan strategy are: (i) to
+modulate the astronomical signals of interest in such a way that they
+appear in the lowest-noise regions of the bolometer noise power
+spectrum, i.e., above the $1/f$ knee; and (ii) to provide good
+``cross-linking'', in which each portion of the map is sampled on a
+range of temporal scales, again, to help distinguish time-varying
+noise features from fixed astronomical sources. In the case of \scuba,
+(i) is achieved through fast-scanning of the entire telescope (up to
+600\,arcsec\,sec$^{-1}$ slew speeds); and (ii) by offering scan
+patterns that cross the sky at a wide range of position angles,
+ultimately observing each portion of the map at a number of different
+times. Such methods are now used by virtually all existing
+ground-based submillimetre cameras, in favour of ``chopping'' methods
+(where the secondary is quickly moved to modulate the signal) that
+were more appropriate for older instruments that had poorer
+low-frequency noise performance, and were only sensivitve to modest
+angular scales
+\citep[e.g.][]{glenn1998,weferling2002,laurent2005,wilson2008,kovacs2008b}.
+
+There are three general styles of map-making that are described in the
+literature for bolometer data. The best maps are produced using
+\emph{maximum likelihood} techniques, in which the map is expressed as
+a linear combination of the data using weights that are a function of
+the time-domain noise properties of the bolometers. The first good
+description of this method appears in \citet{janssen1992} for the
+production of maps from the COsmic Background Explorer
+(COBE). However, this method can be extremely computationally
+expensive, even when small numbers of bolometers are involved,
+although we note that an algorithm called ``SANEPIC'' for reducing
+maps from the Balloon-borne Large-Aperture Submillimeter Telescope
+(BLAST) involving hundreds of bolometers and correctly incorporating
+inter-bolometer noise correlations, has been implemented with
+reasonable execution speeds and memory requirements
+\citep[][]{patanchon2008}.
 
@@
 
-Another paper worth citing? \citet{weferling2002}
-
 \textit{May want to note that the pipeline is discussed elsewhere,
   e.g. Gibb et al 2006.}
+
+% Historically, both radio
+%and submillimetre telescope systems have used ``chopping'' to perform
+%rapid differences between target sky locations and some reference
+%(e.g., another point on the sky, internal calibrators, and different
+%frequencies in the case of heterodyne systems). In fact, spatial
+%chopping was used by SCUBA, the predecessor to \scuba, to remove
+%low-frequency drifts \citep{holland1999}. However, chopping (in
+%particular spatial chopping) has the unfortunate side-effect of
+%filtering out information on scales larger than the chop-throw
+%\textbf{(references)}.For this reason, modern cameras like \scuba\ aim
+%to provide a high-degree of inherent detector stability, and rapidly
+%slew the telescope so that astronomical signals up to large scales are
+%captured in a signal band at considerably higher frequency than the
+%residual low-frequency drift.
+
+%In the case of chopped data, ideally the resulting data are completely
+%free of low-frequency noise, and therefore the data samples are
+statistically uncorrelated.
+
+%often used various
+%``chopping'' methods to remove low-frequency drifts -- rapid
+%differences are made between the sky signal at a given position and an
+%internal reference signal; the sky signal in two different bands for
+%heterodyne systems (frequency chopping); and a different position in
+%the sky \citep[a method employed by SCUBA, the predecessor of
+%\scuba][]{holland1999}.
 
@@
 simple high-pass filtering, this method has not yet been explored in
-detail with SCUBA-2 data.
+detail with \scuba\ data.
 
@@
 
-For future, significantly deeper SCUBA-2 maps, in which the RMS in a
+For future, significantly deeper \scuba\ maps, in which the RMS in a
 PSF-smoothed map is dominated by point-source confusion, rather than
@@
 \citep[e.g.][]{gautier1992}. It is then filtered again with a
-14.5\,arcsec Gaussian to model the effect of the SCUBA-2 optical
+14.5\,arcsec Gaussian to model the effect of the \scuba\ optical
 response. The RMS of these fluctuations are then normalized to 0.002
@@
 
-\item SMURF is an iterative map-maker that can produce SCUBA-2 maps on
+\item SMURF is an iterative map-maker that can produce \scuba\ maps on
 single high-end modern desktops in a reasonable amount of time.

### 2012-06-13 smurf_paper.tex
more intro work

@@
 \begin{abstract}
-  We describe the properties of data from the Submillimetre Common
-  User Bolometer Array 2 (\scuba), a camera for the 15-m James Clerk
+  We describe data properties of the Submillimetre Common User
+  Bolometer Array 2 (\scuba), a camera for the 15-m James Clerk
   Maxwell Telescope that provides simultaneous imaging capabilities at
@@
 are not time-varying). If successful, the noise in the resulting map
-is said to be ``background noise limited'', meaning that the map noise
-is uncorrelated spatially, and has an amplitude that scales as the
+is said to be ``white noise limited'', meaning that the map noise is
+uncorrelated spatially, and has an amplitude that scales as the
 $\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the noise-equivalent
@@
 (i) is achieved through fast-scanning of the entire telescope (up to
-600\,arcsec\,sec$^{-1}$ slew speeds); and (ii) by offering scan
-patterns that cross the sky at a wide range of position angles,
-ultimately observing each portion of the map at a number of different
-times. Such methods are now used by virtually all existing
+600\,arcsec\,sec$^{-1}$ slew speeds), such that significant drift in
+the bolometers due to low-frequency noise occurs more slowly than the
+crossing times for the astronomical scales of interest; and (ii) by
+offering scan patterns that cross the sky at a wide range of position
+angles, ultimately observing each portion of the map at a number of
+different times. Such methods are now used by virtually all existing
 ground-based submillimetre cameras, in favour of ``chopping'' methods
@@
 angular scales
-\citep[e.g.][]{glenn1998,weferling2002,laurent2005,wilson2008,kovacs2008b}.
-
-There are three general styles of map-making that are described in the
-literature for bolometer data. The best maps are produced using
-\emph{maximum likelihood} techniques, in which the map is expressed as
-a linear combination of the data using weights that are a function of
-the time-domain noise properties of the bolometers. The first good
-description of this method appears in \citet{janssen1992} for the
-production of maps from the COsmic Background Explorer
-(COBE). However, this method can be extremely computationally
-expensive, even when small numbers of bolometers are involved,
-although we note that an algorithm called ``SANEPIC'' for reducing
-maps from the Balloon-borne Large-Aperture Submillimeter Telescope
-(BLAST) involving hundreds of bolometers and correctly incorporating
-inter-bolometer noise correlations, has been implemented with
-reasonable execution speeds and memory requirements
-\citep[][]{patanchon2008}.
-
-  Styles of map-making: least
-squares \citep{janssen1992} maximum likelihood
-\citep[e.g.,][]{patanchon2008}; iterative -- see \citet{johnstone2000}
-implementation of \citet{wright1996} to remove chop,
-\citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA analysis
-\citep[e.g.][]{laurent2005,scott2008,aguirre2010}.
+\citep[e.g.][]{glenn1998,weferling2002,wilson2008,kovacs2008b}.
+
+There are three general styles of map-making that are relevant to
+reducing bolomter data in the literature. The simplest methods involve
+some basic processing of the data to remove as much noise
+contamination as possible, e.g., using baseline removal, filters, and
+then simply re-gridding the data into a map. Such was the basic recipe
+for the baseline reduction of chopped data from SCUBA-2's predecessor
+SCUBA (reference?), and MAMBO, another camera from the same
+generation. Generally speaking, such methods are fast, although
+dependending on the science goals and noise properties of the data,
+they may not achieve the best theoretical noise performance on the
+angular scales of interest. A particular case of this method was
+successfully used for the reduction of data from BOLOCAM
+\citep[e.g.][]{laurent2005}, and its younger sibling AzTEC
+\citep[e.g.][]{scott2008}, using ``principal component analysis''
+(PCA) cleaning. A statistical black-box is used to remove the most
+correlated components of the bolometer signals, enabling the detection
+of point-sources very close to the theoretical white-noise limits of
+the detectors with reasonable computation times. However, PCA cleaning
+is not a good generic solution for producing maps of extended
+structures, since large astronomical sources produce correlated
+signals amongst many detectors, and are removed by this procedure.
+
+The best map-making strategies for recovering information on all
+angular scales are \emph{maximum likelihood} techniques, in which the
+time-series data are expressed as a sampling of the ``true'' map of
+the sky plus noise, and then the equation is inverted to estimate the
+map as some weighted linear combination of the data that minimizes the
+variance.  The first good description of this method appears in
+\citet{janssen1992} for the production of maps from the COsmic
+Background Explorer (COBE -- the description is relevant despite the
+fact that it used a differential radiometer instead of
+bolometers). Other descriptions in the CMB literature include
+\citet{tegmark1997} and \citet{stompor2002}.  The downside to such
+methods is that they can be both computationally expensive and have
+large memory requirements. While for some experimental designs fast
+iterative methods for the inversion do exist with minimal memory usage
+\citep[e.g.][]{wright1996}, for the more general map-making problem
+involving many detectors, the problem is significantly complicated
+both by the need to measure the cross power-spectra for all unique
+pairs of bolometers, and then to perform the inversion itself. The
+most promising maximum-likelihood method that may one day be applied
+to SCUBA-2 data is ``SANEPIC'', which successfully reduced maps from
+the Balloon-borne Large-Aperture Submillimeter Telescope, involving
+hundreds of bolometers, and correctly incorporating inter-bolometer
+noise correlations \citep[][]{patanchon2008}.
+
+%  Styles of map-making: least
+%squares \citep{janssen1992} maximum likelihood
+%\citep[e.g.,][]{patanchon2008}; iterative -- see \citet{johnstone2000}
+%implementation of \citet{wright1996} to remove chop,
+%\citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA analysis
+%\citep[e.g.][]{laurent2005,scott2008,aguirre2010}.
 

### 2012-06-14 smurf_paper.tex
complete draft of intro

@@
 \citep[e.g.][]{griffin2002}. The sensitivity of bolometers in a camera
-like \scuba\ are limited by photon noise from the instrument and
-ambient backgrounds. The low-frequency noise, however, is dominated by
-sources such as slow variations in mean backgrounds (e.g., thermal
-variations within the cryostat, and the atmosphere for ground-based
-cameras), and drift in the readout electronics. Such noise has a power
-spectrum $\propto 1/f^\alpha$ ($\alpha>0$), and the frequency at which
-it is comparable to the white noise level is called the ``$1/f$
-knee''. Since the low-frequency noise is largely correlated between
-all, or subsets, of the bolometers in time, it can be suppressed
-during map-making, since astronomical signals have the distinct
-property that they are fixed in a sky reference frame (assuming they
-are not time-varying). If successful, the noise in the resulting map
-is said to be ``white noise limited'', meaning that the map noise is
-uncorrelated spatially, and has an amplitude that scales as the
-$\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the noise-equivalent
-flux-density (the white noise level of a bolometer in 1\,s of
-integration), and $t$ is the amount of integration time in a map
-pixel.
+like \scuba\ are limited by white photon and phonon noise from the
+instrument and ambient backgrounds. The low-frequency noise, however,
+is dominated by sources such as slow variations in mean backgrounds
+(e.g., thermal variations within the cryostat, and the atmosphere for
+ground-based cameras), and drift in the readout electronics. Such
+noise has a power spectrum $\propto 1/f^\alpha$ ($\alpha>0$), and the
+frequency at which it is comparable to the white noise level is called
+the ``$1/f$ knee''. Since the low-frequency noise is largely
+correlated between all, or subsets, of the bolometers in time, it can
+be suppressed during map-making, since astronomical signals have the
+distinct property that they are fixed in a sky reference frame
+(assuming they are not time-varying). If successful, the noise in the
+resulting map is said to be ``white noise limited'', meaning that the
+map noise is uncorrelated spatially, and has an amplitude that scales
+as the $\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the
+noise-equivalent flux-density (the white noise level of a bolometer in
+1\,s of integration), and $t$ is the amount of integration time in a
+map pixel.
 
@@
 There are three general styles of map-making that are relevant to
-reducing bolomter data in the literature. The simplest methods involve
-some basic processing of the data to remove as much noise
-contamination as possible, e.g., using baseline removal, filters, and
-then simply re-gridding the data into a map. Such was the basic recipe
-for the baseline reduction of chopped data from SCUBA-2's predecessor
-SCUBA (reference?), and MAMBO, another camera from the same
-generation. Generally speaking, such methods are fast, although
+reducing bolomter data in the literature. The simplest ``direct
+methods'' involve some basic processing of the data to remove as much
+noise contamination as possible, e.g., using baseline removal,
+filters, and then simply re-gridding the data into a map. Such was the
+basic recipe for the baseline reduction of chopped data from SCUBA-2's
+predecessor SCUBA (reference?), and MAMBO, another camera from the
+same generation. Generally speaking, such methods are fast, although
 dependending on the science goals and noise properties of the data,
@@
 
-\scuba\ is presently unique in terms of the shear volume of data that
-it produces on a nightly basis, and the fact that it is observing a
-wide variety of science targets for multiple users (unlike dedicated
-CMB experiments, for example). For this reason, SMURF has been
-designed to use a faster, and less memory-intensive algorithm than
-CMB-style maximum likelihood map-makers, although it is less accurate
-at reproducing large-scale structures. The basic procedure is to
-iteratively model and remove most of the bolometer-bolometer
-correlated signal components and independent low-frequency noise, and
-then regrid the residual signals. This method is similar in spirit to
-the Comprehensive Reduction Utility for SHARC-2
+The third approach adopted here for \scuba\ is a compromise between
+the previous two methods. Under the assumption that a significant
+portion of the (predominantly low-frequency) non-white noise sources
+can be modeled, an iterative solution is obtained for both the
+astronomical image and the parameters of the noise model. Since the
+remaining (non-modelled) noise sources are assumed to be white, a
+single scalar noise may be calculated for all of the data points from
+a given bolometer (since the noise at any instant in time is
+uncorrelated with others, nor with the data from other bolometers),
+greatly simplifying the measurement of noise properties and the
+inversion step that is so complicated in the maximum-likelihood
+methods. In principle, it should also be possible to recover a wider
+dynamic range in angular scales than the simplest ``direct'' methods,
+with an increase in calculation time that is merely linear in the
+number of iterations needed, while using a comparable amount of
+memory. A fast iterative method for handling chopped data from CMB
+experiments was described in \citet{wright1996}, and later applied to
+chopped SCUBA data by \citet{johnstone2000}, although in each case the
+chopping renders the data essentially white without the need to model
+additional noise components. More similar to the algorithm described
+here was the iterative method for fitting and removing baseline drifts
+in SCUBA scan-map data as the astronomical image estimate improved
+(...SURF reference?...). The two closest modern relatives of SMURF are the
+Comprehensive Reduction Utility for SHARC-2
 \citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
-Bolocam Galactic Plane Survey \citep{aguirre2010}. While most of the
-analysis described here applies equally to all of the data taken
-to-date with \scuba, we note that the \scuba\ Shared-Risks Observing
-period (S2SRO) which took place from 2010 February 11 to 2010 March 25
-used only a single subarray at each of the two observing wavelengths,
-and had considerably different noise properties and poorer performance
-than the fully-commissioned instrument in 2011.
-
-\textit{May want to note that the pipeline is discussed elsewhere,
-  e.g. Gibb et al 2006.}
+Bolocam Galactic Plane Survey \citep{aguirre2010}.
+
+A reasonable model for correlated noise in \scuba\ is a single
+``common-mode'' signal seen by all of the bolometers. Iterative
+estimation and removal of this signal significantly lowers the $1/f$
+knee, without compromosing structures on angular scales larger than
+the array footprint. Residual independent drifts at lower frequencies
+are removed with an iterative high-pass filter. This strategy enables
+SMURF to reduce data faster than they are taken, on single high-end
+desktop computers, and it has successfully been used as part of a
+real-time pipeline offering feedback to observers at the
+telescope. The pipeline is also used to generate near-science grade
+products at the Canadian Astronomical Data Centre (CADC). Details of
+the pipeline design are given in \citet{gibb2005}
+
+This paper is divided in the following way. We first describe the
+properties of \scuba\ data and scan patterns in Section~\ref{sec:data}
+to motivate the data model that we have adopted. Note that while most
+of the analysis described here applies equally to all of the data
+taken to-date with \scuba, the \scuba\ Shared-Risks Observing period
+(S2SRO) which took place from 2010 February 11 to 2010 March 25 used
+only a single subarray at each of the two observing wavelengths, and
+had considerably different noise properties and poorer performance
+than the fully-commissioned instrument in 2011 (a comparison is
+provided in Section~\ref{sec:s2sro}. Next, the details of the SMURF
+algorithm are given in Section~\ref{sec:algorithm}. The paper is
+concluded in Section~\ref{sec:examples} with three detailed test cases
+that span the majority of observation types likely to be undertaken
+with \scuba, with an emphasis on the mitigation of divergence
+problems, and characterizing the output maps: (i) Uranus, a bright,
+compact source (Section~\ref{sec:point}); (ii) the Lockman Hole, a
+blind survey of faint point-like source (Section~\ref{sec:cosmo}); and
+(iii) the star-forming region M17, including bright, extended emission
+(Section~\ref{sec:extended}). All of the data analyzed in this paper
+are publicly available through the CADC SCUBA-2 raw-data queries page
+for the dates and observation numbers given in the
+text\footnote{\texttt{http://www3.cadc-ccda.hia-iha.nrc-cnrc.gc.ca/
+jcmt/search/scuba2}}. All of the analysis was performed using
+\texttt{Starlink} version ... .
+
+%\scuba\ is presently unique in terms of the shear volume of data that
+%it produces on a nightly basis, and the fact that it is observing a
+%wide variety of science targets for multiple users (unlike dedicated
+%CMB experiments, for example). For this reason, SMURF has been
+%designed to use a faster, and less memory-intensive algorithm than
+%CMB-style maximum likelihood map-makers, although it is less accurate
+%at reproducing large-scale structures. The basic procedure is to
+%iteratively model and remove most of the bolometer-bolometer
+%correlated signal components and independent low-frequency noise, and
+%then regrid the residual signals.
 
@@
 %free of low-frequency noise, and therefore the data samples are
-statistically uncorrelated.
+%statistically uncorrelated.
 
@@
 \subsection{Differences between the S2SRO and 2011 re-commissioned instrument}
+\label{sec:s2sro}
 %-------------------------------------------------

### 2012-08-10 smurf_paper.tex
re-do plots for intro (time-series and power spec), new and old s2sro data

@@
 
-\begin{figure}
-\vspace{2cm}
-\textbf{This will be a figure showing s4a and s8d responsivity maps}
-\vspace{2cm}
-\caption{The response of each bolometer to known input heater power
-  for a typical night during S2SRO. The s4a subarray had a
-  particularly large variation in $T_\mathrm{C}$ towards the
-  bottom-right which resulted in lower responsivity, and
-  correspondingly larger noise. The s8d array also suffered
-  significant variations across the focal plane, in particular
-  removing larger number of useful detectors from the bottom-left.}
-\label{fig:sensitivities}
-\end{figure}
+%\begin{figure}
+%\vspace{2cm}
+%\textbf{This will be a figure showing s4a and s8d responsivity maps}
+%\vspace{2cm}
+%\caption{The response of each bolometer to known input heater power
+%  for a typical night during S2SRO. The s4a subarray had a
+%  particularly large variation in $T_\mathrm{C}$ towards the
+%  bottom-right which resulted in lower responsivity, and
+%  correspondingly larger noise. The s8d array also suffered
+%  significant variations across the focal plane, in particular
+%  removing larger number of useful detectors from the bottom-left.}
+%\label{fig:sensitivities}
+%\end{figure}
 
@@
 
-\begin{figure}
+\begin{figure*}
 \centering
-\includegraphics[width=\linewidth]{bolos_point_mix.pdf}
+\includegraphics[width=0.49\linewidth]{bolos_point_mix_s2sro.pdf}
+\includegraphics[width=0.49\linewidth]{bolos_point_mix.pdf}
 \caption{A comparison between single bolometer time series in each
@@
 \label{fig:bolos_mix}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}
+\begin{figure*}
 \centering
-\includegraphics[width=\linewidth]{pspec.pdf}
+\includegraphics[width=0.49\linewidth]{pspec_s2sro.pdf}
+\includegraphics[width=0.49\linewidth]{pspec.pdf}
 \caption{Bolometer power spectral densities. Four representative
@@
 \label{fig:pspec}
-\end{figure}
+\end{figure*}
 

### 2012-08-14 smurf_paper.tex
Add standard JCMT acknowledgment
Needs some clarification on the words for EC CANFAR funding.

@@
 
-\textit{Presumably the standard telescope acknowledgment here plus a
-thank you to CFI for funding the DR work. Ed should probably thank
-both CANFAR and JAC for additionally funding although his JAC funding
-is noted in the affiliation.}
+The James Clerk Maxwell Telescope is operated by the Joint Astronomy
+Centre on behalf of the Science and Technology Facilities Council of
+the United Kingdom, the Netherlands Organisation for Scientific
+Research, and the National Research Council of Canada. Additional
+funds for the contruction of SCUBA-2 were provided by the Canada
+Foundation for Innovation. EC thanks CANARIE/CANFAR for additional
+funding.
+
+\textit{Is there some kind of grant number that has to be referenced
+  for Ed's CANFAR funding?}
 

### 2012-08-14 smurf_paper.tex
Fix missing label for appendix
Update references
SPIE references were missing page numbers and editor information.
Added some references to SURF and JSA into the text. Updated
a astro-ph reference that was actually published in 2009.

@@
 The Submillimetre Common User Bolometer Array 2
-\citep[\scuba,][]{holland2006} is a new instrument for the 15-m James
+\citep[\scuba,][]{2006SPIE.6275E..45H,holland2012} is a new instrument for the 15-m James
 Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
@@
 \citep[FTS-2,][]{gom2010}, and a polarimeter
-\cite[POL-2,][]{bastien2005}. This paper describes the properties of
+\cite[POL-2,][]{2011ASPC..449...68B}. This paper describes the properties of
 \scuba\ data that are relevant to producing maps of the imaging data,
 and the Submillimetre User Reduction Facility, SMURF, a software
-package for perfoming the reduction. The analysis of FTS-2 and POL-2
-data will be described at a later date once these additional
+package for perfoming the reduction written using the Starlink
+Software Environment \citep{2009ASPC..411..418J}. The analysis
+of FTS-2 and POL-2 data will be described at a later date once these additional
 instruments have been commissioned. The details of the instrument
@@
 basic recipe for the baseline reduction of chopped data from SCUBA-2's
-predecessor SCUBA (reference?), and MAMBO, another camera from the
-same generation. Generally speaking, such methods are fast, although
-dependending on the science goals and noise properties of the data,
+predecessor SCUBA \citep{1998ASPCd..145..216J,2000ASPC..216..559J},
+and MAMBO, another camera from the same generation. Generally
+speaking, such methods are fast, although dependending on the science
+goals and noise properties of the data,
 they may not achieve the best theoretical noise performance on the
@@
 telescope. The pipeline is also used to generate near-science grade
-products at the Canadian Astronomical Data Centre (CADC). Details of
-the pipeline design are given in \citet{gibb2005}
+products using the JCMT Science Archive \citep{2011ASPC..442..203E}
+hosted by the Canadian Astronomical Data Centre (CADC). Details of
+the pipeline design are given in \citet{gibb2005} and
+\citet{2008AN....329..295C}.
 
@@
 typical for other ground-based submillimetre surveys in recent years
-\citep[e.g.,][]{coppin2006,perera2008,weiss2009} leading to estimated
+\citep[e.g.,][]{coppin2006,perera2008,2009ApJ...707.1201W} leading to estimated
 false-identification rates of order $\sim$5\%, and is chosen here as a
@@
 the interested reader to a selection of papers on the subject:
-\citet{scott2002,coppin2006,perera2008,weiss2009} and
+\citet{scott2002,coppin2006,perera2008,2009ApJ...707.1201W} and
 \citet{chapin2011}.
@@
 \section[]{Step correction in detail}
+\label{appendix:steps}
 

### 2012-08-22 smurf_paper.tex
updates to most of the paper from Douglas Scott

@@
 \begin{abstract}
-  We describe data properties of the Submillimetre Common User
-  Bolometer Array 2 (\scuba), a camera for the 15-m James Clerk
-  Maxwell Telescope that provides simultaneous imaging capabilities at
-  450 and 850\,\micron, and the production of maps using the
-  Sub-Millimetre User Reduction Facility (SMURF). With nearly 5000
-  bolometers in each imaging band, and an approximately 5\,arcmin
-  field-of-view, \scuba\ is presently the largest submillimetre camera
-  in the world, and produces data at rates an order-of-magnitude
-  faster than previous instruments. We have adopted a fast, iterative
-  approach to map-making that enables data reduction on single,
-  modern, high-end desktop computers with execution times that are
-  shorter than the observing times.  SMURF is used in an automated
+  The Submillimetre Common User Bolometer Array 2 (\scuba) is an
+  instrument operating on the 15-m James Clerk Maxwell Telescope,
+  consisting of about 5000 bolometers in each of two simultaneous
+  imaging bands centered over 450 and 850\,\micron. The camera is
+  scanned across the sky and records data at a rate of 200\,Hz. This
+  data volume is much greater than previously existing submillimetre
+  cameras, and represents a significant data anlaysis challenge. We
+  describe the production of SCUBA-2 maps using the Sub-Millimetre
+  User Reduction Facility (SMURF) in which we have adopted a fast,
+  iterative approach to map-making that enables data reduction on
+  single, modern, high-end desktop computers with execution times that
+  are shorter than the observing times.  SMURF is used in an automated
   setting both at the telescope for real-time feedback to observers,
@@
   using simple prior constraints (Uranus -- a point source); (ii)
-  demonstrate that our strategy is capable of recovering angular
+  achieve the white-noise limit of the instrument for faint
+  point-source studies (blank-field survey of the Lockman Hole); and
+  (iii) demonstrate that our strategy is capable of recovering angular
   scales up to the size of the array footprint (approximately
-  5\,arcmin) for bright extended sources (star-forming region M17);
-  and (iii) achieve the white-noise limit of the instrument for faint
-  point-source studies (blank-field survey of the Lockman Hole).
+  5\,arcmin) for bright extended sources (star-forming region M17).
 \end{abstract}
@@
 The Submillimetre Common User Bolometer Array 2
-\citep[\scuba,][]{2006SPIE.6275E..45H,holland2012} is a new instrument for the 15-m James
-Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
-simultaneously images the sky in two broad bands centered over 450 and
-850\,\micron, with approximately 7.5 and 14.5\,arcsec full-width at
-half-maximum (FWHM) point spread functions (PSFs). The focal planes at
-each wavelength are populated with 4 rectangular subarrays, consisting
-of $40 \times 32$ bolometers each, and subtend a nearly contiguous $7'
-\times 7'$ footprint on the sky.  Ultimately, the instrument will also
-be equipped with a Fourier transform spectrometer
-\citep[FTS-2,][]{gom2010}, and a polarimeter
-\cite[POL-2,][]{2011ASPC..449...68B}. This paper describes the properties of
-\scuba\ data that are relevant to producing maps of the imaging data,
-and the Submillimetre User Reduction Facility, SMURF, a software
-package for perfoming the reduction written using the Starlink
-Software Environment \citep{2009ASPC..411..418J}. The analysis
-of FTS-2 and POL-2 data will be described at a later date once these additional
-instruments have been commissioned. The details of the instrument
-design, performance, and its calibration are given in two companion
-papers: \citet{holland2012} and \citet{dempsey2012}.
+\citep[\scuba,][]{2006SPIE.6275E..45H,holland2012} is a new instrument
+for the 15-m James Clerk Maxwell Telescope (JCMT) on Mauna Kea,
+Hawai'i. The camera simultaneously images the sky in two broad bands
+centered over 450 and 850\,\micron, with approximately 7.5 and
+14.5\,arcsec full-width at half-maximum (FWHM) point spread functions
+(PSFs). The focal planes at each wavelength are populated with 4
+rectangular subarrays, consisting of $40 \times 32$ bolometers each,
+and subtend a nearly contiguous 7\,arcmin $\times$ 7\,arcmin footprint
+on the sky.  The instrument is also equipped with a Fourier transform
+spectrometer \citep[FTS-2,][]{gom2010}, and a polarimeter
+\cite[POL-2,][]{2011ASPC..449...68B}. This paper describes the
+properties of \scuba\ data that are relevant to producing maps of the
+imaging data, and the Submillimetre User Reduction Facility, SMURF, a
+software package for perfoming the reduction written using the
+Starlink Software Environment \citep{2009ASPC..411..418J}. The
+analysis of FTS-2 and POL-2 data will be described at a later date
+once these additional instruments have been commissioned and
+characterized. The details of the instrument design, performance, and
+its calibration are given in two companion papers: \citet{holland2012}
+and \citet{dempsey2012}.
 
 Over the last twenty years, observations in the submillimetre band
-(defined here to be 200--1200\,\micron) have helped revolutionize
+(defined here to be 200--1200\,\micron) have helped revolutionise
 several important areas of astrophysics, including: the
-characterization of early stages of star-formation by identifying the
+characterisation of the early stages of star-formation by identifying the
 dense, cold regions in molecular clouds where stars may eventually
@@
 as submillimetre galaxies, or SMGs; and identifying debris disks
-around nearby stars, the early stages of planet formation.  With
+around nearby stars, helping us understand the early stages of planet formation.  With
 $\sim10000$ total detectors, \scuba\ is presently the largest
@@
 
-Submillimetre imaging cameras generally use bolometers to maximize
+Submillimetre imaging cameras generally use bolometers to maximise
 sensitivity, in which an absorber collects photons that pass through
-broad-band filters, and the signal is read out-out using a thermometer
+broad-band filters, and the signal is read out using a thermometer
 \citep[e.g.][]{griffin2002}. The sensitivity of bolometers in a camera
@@
 instrument and ambient backgrounds. The low-frequency noise, however,
-is dominated by sources such as slow variations in mean backgrounds
+is dominated by sources which produce slow variations in mean backgrounds
 (e.g., thermal variations within the cryostat, and the atmosphere for
-ground-based cameras), and drift in the readout electronics. Such
+ground-based cameras), and drifts in the readout electronics. Such
 noise has a power spectrum $\propto 1/f^\alpha$ ($\alpha>0$), and the
@@
 
-There are numerous ways to attack the problem of map production, both
-in terms of the data-collection method, and processing. The two most
+There are numerous ways to attack this map-making problem, both in
+terms of the data-collection method, and processing. The two most
 important principles to follow in terms of scan strategy are: (i) to
@@
 different times. Such methods are now used by virtually all existing
-ground-based submillimetre cameras, in favour of ``chopping'' methods
-(where the secondary is quickly moved to modulate the signal) that
-were more appropriate for older instruments that had poorer
-low-frequency noise performance, and were only sensivitve to modest
-angular scales
-\citep[e.g.][]{glenn1998,weferling2002,wilson2008,kovacs2008b}.
+ground-based submillimetre cameras
+\citep[e.g.][]{glenn1998,weferling2002,wilson2008,kovacs2008b}., in
+favour of ``chopping'' methods (where the secondary is quickly moved
+to modulate the signal) that were more appropriate for older
+instruments that had poorer low-frequency noise performance, and were
+only sensivitve to modest angular scales
 
@@
 methods'' involve some basic processing of the data to remove as much
-noise contamination as possible, e.g., using baseline removal,
-filters, and then simply re-gridding the data into a map. Such was the
-basic recipe for the baseline reduction of chopped data from SCUBA-2's
-predecessor SCUBA \citep{1998ASPCd..145..216J,2000ASPC..216..559J},
-and MAMBO, another camera from the same generation. Generally
-speaking, such methods are fast, although dependending on the science
-goals and noise properties of the data,
-they may not achieve the best theoretical noise performance on the
+noise contamination as possible, (e.g., using baseline removal and
+other simple filters), and then re-gridding these cleaned data into a
+map. Such was the basic recipe for the baseline reduction of chopped
+data from SCUBA-2's predecessor SCUBA
+\citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO, another
+camera from the same generation. Generally speaking, such methods are
+fast, although dependending on the science goals and noise properties
+of the data, they may not achieve the best noise performance on the
 angular scales of interest. A particular case of this method was
@@
 \citep[e.g.][]{scott2008}, using ``principal component analysis''
-(PCA) cleaning. A statistical black-box is used to remove the most
-correlated components of the bolometer signals, enabling the detection
-of point-sources very close to the theoretical white-noise limits of
-the detectors with reasonable computation times. However, PCA cleaning
-is not a good generic solution for producing maps of extended
-structures, since large astronomical sources produce correlated
-signals amongst many detectors, and are removed by this procedure.
+(PCA) cleaning. A statistical black-box removes the most correlated
+components of the bolometer signals, enabling the detection of
+point-sources very close to the theoretical white-noise limits of the
+detectors with reasonable computation times. However, PCA cleaning is
+not a good generic solution for producing maps of extended structures,
+since such sources produce correlated signals amongst many detectors,
+and are removed by this procedure.
 
@@
 bolometers). Other descriptions in the CMB literature include
-\citet{tegmark1997} and \citet{stompor2002}.  The downside to such
-methods is that they can be both computationally expensive and have
-large memory requirements. While for some experimental designs fast
-iterative methods for the inversion do exist with minimal memory usage
-\citep[e.g.][]{wright1996}, for the more general map-making problem
-involving many detectors, the problem is significantly complicated
-both by the need to measure the cross power-spectra for all unique
-pairs of bolometers, and then to perform the inversion itself. The
-most promising maximum-likelihood method that may one day be applied
-to SCUBA-2 data is ``SANEPIC'', which successfully reduced maps from
-the Balloon-borne Large-Aperture Submillimeter Telescope, involving
-hundreds of bolometers, and correctly incorporating inter-bolometer
-noise correlations \citep[][]{patanchon2008}.
+\citet{tegmark1997} and \citet{stompor2002}, and an application to
+data from SCUBA is described in \citet{borys2004}.  The downside to
+such methods is that they can be both computationally expensive and
+have large memory requirements. While for some experimental designs
+fast iterative methods for the inversion do exist with minimal memory
+usage \citep[e.g.][]{wright1996}, for the more general map-making
+problem involving many detectors, the problem is significantly
+complicated both by the need to measure the cross power-spectra for
+all unique pairs of bolometers, and then to perform the inversion
+itself. The most promising maximum-likelihood method that may one day
+be applied to SCUBA-2 data is ``SANEPIC'', which successfully reduced
+maps from the Balloon-borne Large-Aperture Submillimeter Telescope,
+involving hundreds of bolometers, and correctly incorporating
+inter-bolometer noise correlations \citep[][]{patanchon2008}.
 
@@
 In this section we summarize how \scuba\ bolometers detect temperature
-variations (Section~\ref{sec:bolos}), typical \scuba\ scan strategies
-(Section~\ref{sec:scan}).
+variations (Section~\ref{sec:bolos}), describe \scuba\ scan strategies
+(Section~\ref{sec:scan}), typical bolometers signals
+(Section~\ref{sec:bolosignal}), and the impact of magnetic field
+pickup (Section~\ref{sec:magpickup}, describe the differences between
+SCUBA-2 Shared Risk Observing data and those of the the fully
+commissioned instrument (Section~\ref{sec:s2sro}), and finally
+illustrate the use of principal component analysis
+(Section~\ref{sec:pca}).
 
@@
 
-While the details of how \scuba\ works is described in
+While the details of how \scuba\ works are described in
 \citet{holland2012}, and its calibration in \citet{dempsey2012}, we
-summarize the basic concepts relative to map-making here.
+summarize the basic concepts relevant to map-making here.
 
@@
 biased to a temperature within this transition on large time scales,
-then tiny temperature fluctuations on short time-scales will produce
+then tiny temperature fluctuations on short time scales will produce
 significant changes in the impedence. The entire \scuba\ focal plane
 is therefore held at a \emph{base temperature} somewhat below
-$T_\mathrm{C}$. The TESs then have a small electrical voltage biases,
+$T_\mathrm{C}$. The TESs then have small electrical voltage biases,
 and individual pixel heaters (using a current bias to apply the
@@
 such large-format bolometer array designs feasible. It should also be
-noted that there is an additional 33rd row of SQUID reasouds that are
+noted that there is an additional 33rd row of SQUID readouts that are
 not connected to TES. These ``dark SQUIDs'' are used to track
@@
 are variations in $T_\mathrm{C}$ across each subarray due to
-manufacturing difficulties, considerable effort has gone into
-dynamically applying position-dependent biases as well as an over-all
-focal plane bias to make as many detectors useable as possible
+manufacturing differences, considerable effort has gone into
+dynamically applying position-dependent biases as well as a total
+focal plane bias to make as many detectors usable as possible
 \citep{holland2012}. In practice, the response of a given detector can
-therefore vary somewhat from observation to observation depending on
+therefore vary somewhat from observation to observation, depending on
 how the bias and amplification parameters were set up. To track these
@@
 effectively white thermal noise produced by the phonon and optical
-loading on the absorbers; and (ii) and longer-timescale drift
+loading on the absorbers; and (ii) and longer timescale drift
 resulting from slow variations in the mean load (e.g. fridge,
@@
 atmospheric extinction (due to atmospheric water vapour) must be
-compensated for; and (ii) the corrected power is related to an
-astronomical flux density through the effective area of the detector,
-the spectral response of the bolometers (bandpass), the spectrum of
-the source, and additional optical efficiency factors -- a \emph{flux
-  conversion factor} (FCF). Extinction is measured, in most cases,
-using the JCMT line-of-sight water vapour radiometer (WVM) which is
-calibrated using skydips and the CSO 225\,GHz tipper. The FCF is
-determined empirically using measurements of standard sources. For
-further details on these calibration steps, see \citet{dempsey2012}.
+compensated for as a function of time; and (ii) the corrected power is
+related to an astronomical flux density through the effective area of
+the detector, the spectral response of the bolometers (bandpass), the
+spectrum of the source, and additional optical efficiency factors -- a
+\emph{flux conversion factor} (FCF). Extinction is measured, in most
+cases, using the JCMT line-of-sight water vapour radiometer (WVM)
+which is calibrated using sky-dips and the CSO 225\,GHz tipper. The
+FCF is determined empirically using measurements of standard
+sources. For further details on these calibration steps, see
+\citet{dempsey2012}.
 
@@
 ultimately limited to recovering information on angular scales smaller
-than the chop throw (a maximum of a couple of arcmin)... cite
-somebody. The primary approach to map-making with \scuba\ is instead
-to slew the entire telescope at such a rate that angular scales of
-interest are crossed on time scales shorter than $1/f_\mathrm{knee}$,
-similar to other ground-based instruments (MAMBO, BOLOCAM, SHARC-II,
-AzTEC, LABOCA,...). The desire to scan rapidly has two basic
-constraints: (i) the telescope cannot physically move faster that
-several hundred arcsec\,sec$^{-1}$ without requiring unacceptably
-large accelerations to change direction, and (ii) bolometers are
-designed with a thermal time constant that balance response with
-sensitivity. Taking these constraints into consideration, \scuba\ was
-ultimately designed to work at a maximum scan velocity of
-600\,arcsec\,sec$^{-1}$. The data acquisition system has a target
-sample rate of 200\,Hz, which at that scan velocity results in a
-sample every 3.0\,arcsec, or approximately one third of the
-450\,\micron\ diffraction-limited full-width at half-maximum (FWHM) --
-a typical rule-of-thumb for adequately sampling a Gaussian point
-spread function (PSF). Similarly, the \scuba\ bolometers are designed
-to have a time constant of 2 ms \textbf{check that number!} in order
-to avoid distorting PSF-sized sources at this maximum scan rate.
+than the chop throw (a maximum of several arcmin). The primary
+approach to map-making with \scuba\ is instead to slew the entire
+telescope at such a rate that angular scales of interest are crossed
+on time scales shorter than $1/f_\mathrm{knee}$, similar to other
+ground-based instruments (MAMBO, BOLOCAM, SHARC-II, AzTEC,
+LABOCA,...). The desire to scan rapidly has two basic constraints: (i)
+the telescope cannot physically move faster that several hundred
+arcsec\,sec$^{-1}$ without requiring unacceptably large accelerations
+to change direction; and (ii) bolometers are designed with a thermal
+time constant that balances response with sensitivity. Taking these
+constraints into consideration, \scuba\ was ultimately designed to
+work at a maximum scan speed of 600\,arcsec\,sec$^{-1}$. The data
+acquisition system has a target sample rate of 200\,Hz, which at that
+scan velocity results in a sample every 3.0\,arcsec, giving
+approximately 2.5 samples per 450\,\micron\ diffraction-limited
+full-width at half-maximum (FWHM) -- a typical rule-of-thumb for
+adequately sampling a Gaussian point spread function (PSF). Similarly,
+the \scuba\ bolometers are designed to have a time constant of 2 ms
+\textbf{check that number!} in order to avoid distorting PSF-sized
+sources at this maximum scan rate.
 
 The scan strategy is designed, primarily, to mitigate the effects of
-$1/f$ drift. The simplest scan one might use is a raster pattern, or a
-series of parallel passes across this field. However, such a scan
-yields extremely poor performace for structures that are transverse to
-the scan direction, since a long time will have ellapsed between each
-parallel step, and the signal will therefore be totally dominated by
-the $1/f$ noise. In addition, other noise features may appear in the
-data periodically. With such a regularly spaced periodic scan
-strategy, these noise features can easily conspire with the scan
-pattern to produce similarly periodic noise structures in the map, or
-\emph{scan-synchronous} noise. It has therefore been recognized that
-scan strategies that visit each region of the map on as many different
-time scales, and scan angles as possible offer the best protection
-against large-scale noise (...cite some people...); such strategies
-are said to offer good \emph{cross-linking}. As a tradeoff between
-truly random scan patterns and feasibility of implementation taking
-into consideration both mechanical and scheduling constraints, \scuba
+$1/f$ drift. The simplest scan one might use is a raster
+``boustrophedon'' pattern, i.e. a series of parallel passes across the
+field. However, such a scan yields extremely poor performace for
+structures that are transverse to the scan direction, since a long
+time will have ellapsed between each parallel step, and the signal
+will therefore be totally dominated by the $1/f$ noise. In addition,
+other periodic noise features may appear in the data. With such a
+regularly spaced scan strategy, these noise features can easily
+conspire with the scan pattern to produce similarly periodic noise
+structures in the map, or \emph{scan-synchronous} noise. It has
+therefore been recognized that scan strategies that visit each region
+of the map on as many different time scales, and scan angles as
+possible offer the best protection against large-scale noise
+\citep[e.g.][]{tegmark1997,kovacs2008b}; such strategies are said to
+offer good \emph{cross-linking}. As a tradeoff between truly random
+scan patterns and feasibility of implementation taking into
+consideration both mechanical and scheduling constraints, \scuba
 presently has three primary scan modes: PONG (cite Borys et al.),
-Lissajous, and Daisy... (more citations? Kovacs et al. has a nice
+LISSAJOUS, and DAISY... (more citations? Kovacs et al. has a nice
 description... also CMB literature).
@@
 850\,\micron. In this particular case, the total power in the
-fluctuations at 450\,\micron, from $-0.04$\,pW to $+0.08$, are larger
+fluctuations at 450\,\micron, from $-0.04$\,pW to $+0.08$\,pW, are larger
 than the $-0.03$\,pW to $+0.05$\,pW fluctuations at 850\,\micron. Such
@@
 
-The low-frequency signal component of the bolometer signals is also
+The low-frequency signal component of the bolometer output is also
 highly correlated amongst bolometers in the same subarray. We have
@@
 conversion from variance measured in the frequency domain to the time
-domain requires and additional multiplcation by 2; hence the units are
+domain requires an additional multiplcation by 2; hence the units are
 pW$^2$\,Hz$^{-1}$ rather than pW$^2$\,s.
@@
   10^{-8}$ and $5 \times 10^{-9}$\,pW\,Hz$^{-1}$ at 450 and
-  850\,\micron\ respectively, give the approximate white-noise
+  850\,\micron, respectively, give the approximate white-noise
   levels. There are broad line features in the PSDs at both
@@
   see that, for this scan velocity, the low-frequency noise begins to
-  dominate on scales of approximately 8.3 and 1.7\,arcmin; these scales
-  are comparable to the array footprint.}
+  dominate on scales of approximately 8.3 and 1.7\,arcmin; these
+  scales are comparable to the array footprint.}
 \label{fig:pspec}
@@
 The dotted coloured lines in Fig~\ref{fig:pspec} show the PSDs for
-raw, though flatfielded data. At each wavelength, there is a clear
+raw, though flat-fielded data. At each wavelength, there is a clear
 $1/f$ knee at a few Hz, followed by a predominantly white spectrum
-punctuated by line features above $\gsim 30$\,Hz, and finally roll-off
-caused by the anti-aliasing filter above $\gsim 70$\,Hz. As indicated
-in the previous section, the correlation between the low-frequency
-components of the different bolometer signals is large. The solid
-black lines in Fig.~\ref{fig:pspec} indicate the PSDs of the
-common-modes $\mathbf{c}(t)$ at each wavelength, which reproduce most
-of the low-frequency structures, as well as the high-frequency line
-features. The $\mathbf{c}(t)$ otherwise drop substantially below the
-individual bolometer PSDs at high-frequency as expected if the
-bolometers are dominated by un-correlated white noise. Finally, we
-note that the amplitudes and slopes of the $\mathbf{c}(t)$ are similar
-at both 450 and 850\,\micron. However, since the 450\,\micron\ white
-noise is larger than at 850\,\micron\ (shown approximately by the
-horizontal dotted lines), the $1/f$ knee occurs at a \emph{lower}
-frequency at 450\,\micron.
+punctuated by line features above $\gsim 30$\,Hz, and finally a
+roll-off caused by the anti-aliasing filter above $\gsim 70$\,Hz. As
+indicated in the previous section, the correlation between the
+low-frequency components of the different bolometer signals is
+large. The solid black lines in Fig.~\ref{fig:pspec} indicate the PSDs
+of the common-modes $\mathbf{c}(t)$ at each wavelength, which
+reproduce most of the low-frequency structure, as well as the
+high-frequency line features. The $\mathbf{c}(t)$ otherwise drop
+substantially below the individual bolometer PSDs at high-frequency,
+as expected if the bolometers are dominated by un-correlated white
+noise. Finally, we note that the amplitudes and slopes of
+$\mathbf{c}(t)$ are similar at both 450 and 850\,\micron. However,
+since the 450\,\micron\ white noise is larger than at 850\,\micron\
+(shown approximately by the horizontal dotted lines), the $1/f$ knee
+occurs at a \emph{lower} frequency at 450\,\micron.
 
@@
 
+%-------------------------------------------------
 \subsection{Magnetic field pickup}
 \label{sec:magpickup}
+%-------------------------------------------------
 
@@
 attached to it, this observed signal is not likely to be optical or
-thermal in nature (although there can be some crosstalk with the
-bolometers...). It is also for this reason that the higher-frequency
+thermal in nature (although there can be some cross-talk with the
+bolometers). It is also for this reason that the higher-frequency
 noise that appears in the bolometer data (thermal) does not appear in
-the dark squid, leading to a significantly higher \snr. The sign of
-such a signal is expected to be random in the bolometer time-series
-(ask detector person for an explanation).
+the dark squid, leading to a significantly higher \snr.
 
@@
   shows the azimuthal and elevation offsets from the map centre. For
-  reference, the mean azimuth was 171.9\,deg, and the mean elevation
-  68.0\,deg. Only the azimuthal signal appears to be correlated with
-  the dark squids and bolometer signals, which suggests a magnetic
-  field stationary with respect to the telescope dome as the source,
-  since its direction with respect to the cryostat only changes with
-  azimuthal motion.}
+  reference, the mean azimuth was 171.9$^\circ$ and the mean elevation
+  68.0$^\circ$. Only the azimuthal signal appears to be correlated
+  with the dark squids and bolometer signals, which suggests a
+  magnetic field stationary with respect to the telescope dome as the
+  source, since its direction with respect to the cryostat only
+  changes with azimuthal motion.}
 \label{fig:magpickup}
@@
 
-The telescope pointing offsets for this approximately 0.5\,deg
+The telescope pointing offsets for this approximately 0.5\,degree
 diameter scan are also shown in Fig.~\ref{fig:magpickup}. Since the
@@
 such a field with respect to the cryostat. It is also worth noting
-that the absolute azimuth of the telescope was close to 180\,deg
+that the absolute azimuth of the telescope was close to 180$^\circ$
 (i.e., nearly due-south). We suspect that this observing
@@
 of the 450 and 850\,\micron\ focal planes were populated with single
-subarrays, s4a and s8d respectively. There were two main mechanical
+subarrays, s4a and s8d, respectively. There were two main mechanical
 problems to contend with during this period. First, the fridge was not
@@
 ... which affected stability, particularly at 850\,\micron. Second,
-there were problems with the subarray fabrication process which led to
-large numbers of dead pixels, as well as large-scale variations in
+there were problems with the subarray fabrication process, which led
+to large numbers of dead pixels, as well as large-scale variations in
 $T_\mathrm{C}$ and the thermal conductivity, $G$. These latter issues
@@
 representative responsivity maps for the two subarrays from the middle
-of S2SRO based on the response to ramping the bolometer heaters.
+of S2SRO, based on the response to ramping the bolometer heaters.
 
@@
 to gain further insight into the nature of the \scuba\ noise
-properties, but it is now also included it in SMURF as method for
+properties, but it is now also included in SMURF as method for
 cleaning the data.
@@
 linear combination of the eigenvectors, or \emph{components}
-i.e. $\mathbf{b}_i(t) = \bar{\mathbf{\xi}} \mathbf{\lambda}_i^T$,
-where each row of the matrix $\bar{\mathbf{\xi}}$ is an eigenvector,
-and $\mathbf{\lambda}_i^T$ is a column vector containing the
+i.e. $\mathbf{b}_i(t) = \bar{\mathbf{\xi}}
+\mathbf{\lambda}_i^\mathrm{T}$, where each row of the matrix
+$\bar{\mathbf{\xi}}$ is an eigenvector, and
+$\mathbf{\lambda}_i^\mathrm{T}$ is a column vector containing the
 corresponding eigenvalues. The $\mathbf{\xi}_i$ are normalized by
@@
 those components has been used successfully to reduce $1/f$ noise
-while retaining most of the signal in point-like sources. There is an
-implicit assumption in this analysis that the noise properties are
-stationary.
+while retaining most of the signal in point-like sources. It is also
+worth noting that there is an implicit assumption in this analysis
+that the noise properties are stationary.
 
@@
 The majority of correlated signal at both wavelengths is clearly
-broken down into the first two plotted components. The first dominates
-at 450\,micron, whereas the second dominates at 850\,micron. For
+broken down into the first two components plotted. The first dominates
+at 450\,\micron, whereas the second dominates at 850\,\micron. For
 reference, the eigenvalues of these two modes are in the range
@@
 atmospheric signal (which is expected to be stronger at
-450\,\micron). Indeed, the fainter component 2 most closely resembles
+450\,\micron). Indeed, the fainter Component 2 most closely resembles
 the mixing chamber temperature variations (Fig.~\ref{fig:bolos_mix}),
-whereas component 1 is stronger at 450\,\micron\ and does not resemble
+whereas Component 1 is stronger at 450\,\micron\ and does not resemble
 the fridge signal as closely.
@@
 850\,\micron\ appears to be produced by two dominant underlying
-mechanisms.
+mechanisms, even if we cannot identify the mechanisms with any
+confidence.
 
@@
 its eigenvector plot, is almost a pure ramping waveform that only
-appears in the s4a data, and produces the broad line featuress
-centered over 45\,Hz and 70\,Hz in the 450\,\micron\ PSDs in
+appears in the s4a data, and produces the broad line featuress centred
+over 45\,Hz and 70\,Hz in the 450\,\micron\ PSDs in
 Fig.~\ref{fig:pspec}. In other data sets, this third component has
@@
 is thought that these features are some sort of data acquisition
-glitch as the spikes (when they appear) exhibit the expected point
-response function of the anti-aliasing filter. Similarly, component 4
+glitch, as the spikes (when they appear) exhibit the expected point
+response function of the anti-aliasing filter. Similarly, Component 4
 also appears to be a data acquisition glitch in the s4a array,
 although along a single column, and with a peak at a lower frequency
-of 40\,Hz, and a second feature right at the Nyquist frequency.
+of 40\,Hz, together with a second feature right at the Nyquist
+frequency.
 
@@
 more sensitive to changes in the loading, possibly resulting in the
-step and spikes that are seen in component 5. Component 6 is clearly
-related to component 5, though with the absence of the steps and
+step and spikes that are seen in Component 5. Component 6 is clearly
+related to Component 5, though with the absence of the steps and
 spikes, so that together they seem to represent a continuum of shapes
-that are linear combinations of the spikes and steps with the slower
-drift evident only in component 6.
+that are linear combinations of the spikes and steps, with the slower
+drift evident only in Component 6.
 
@@
 simplified map estimation. We do note, however, that recent advances
-in direct inversion methods such as \citet{patanchon2008} may in fact
-be tenable for \scuba\ data. A future goal of the software team is to
-investigate meshing SMURF (perhaps as a cleaning step and initial
-condition) for a SANEPIC-like algorithm.
+in direct inversion methods such as SANPIEC \citep{patanchon2008} may
+in fact be tenable for \scuba\ data. A future goal of the software
+team is to investigate meshing SMURF (perhaps as a cleaning step and
+initial condition) for a SANEPIC-like algorithm.
 
@@
 \begin{equation}
-\mathbf{b}_i(t) = g_i[\mathbf{e}_i(t) \mathbf{a}_i(t) + \mathbf{n}_i(t)]
+\mathbf{b}_i(t) = g_i[\mathbf{e}_i(t) \mathbf{a}_i(t) + \mathbf{n}_i(t)],
 \end{equation}
@@
 map-making is to model and remove $\mathbf{n}^\mathrm{f}_i$ and
-$\mathbf{n}^\mathrm{c}$ from the time-series, while re-gridding the
-remaining data we can then hope to approach the theoretical noise
-limit in the map. This procedure is summarized in Fig.~\ref{fig:dimm}.
+$\mathbf{n}^\mathrm{c}$ from the time-series, and then by re-gridding
+the remaining data we can hope to approach the theoretical noise limit
+in the map. This procedure is summarized in Fig.~\ref{fig:dimm}.
 
@@
 to track convergence. As we will discuss in the next sections, each
-signal component estimate is biased slightly from the presence of
+signal component estimate is biased slightly due to the presence of
 signal from other components, so the entire process is
@@
 multiplicative factors introduced by the extinction correction, and
-also relative gains that may optionally be estimated in the
+also relative gains that may optionally be determined in the
 common-mode estimation phase.
@@
 
-In addition to specifying when each model component is calculated
+In addition to specifying when each model component is computed
 (normally an attempt is made to order them by decreasing amplitude in
@@
 ﻿Fig.~\ref{fig:com} shows the time series from a selection of typical
-bolometers.\footnote{The data have been flat-fielded and each time series
-has been adjusted to a mean value of zero.} The similarity between most
-bolometers is evident, and forms the common-mode signal - assumed to be a
-consequence of variations in the atmospheric emission and fridge
-temperature. This common-mode usually dominates the astronomical signal
-for all but the brightest sources, and swamps faint extended structure.
-The purpose of the \model{COM} and \model{GAI} models is to remove this common-mode
-signal.
+bolometers.\footnote{The data have been flat-fielded and each time
+series has been adjusted to a mean value of zero.} The similarity
+between most bolometers is evident, and forms the common-mode signal
+-- assumed to be a consequence of variations in the atmospheric
+emission and fridge temperature. This common-mode usually dominates
+the astronomical signal for all but the brightest sources, and swamps
+faint extended structure.  The purpose of the \model{COM} and
+\model{GAI} models is to remove this common-mode signal.
 
@@
 \includegraphics[width=\linewidth]{com.pdf}
-\caption{A selection of typical bolometer time series after flatfielding
-and removal of a constant baseline. It can be seen that most bolometers
-exhibit a common time variation overlayed with other features.}
+\caption{A selection of typical bolometer time series after
+flatfielding and removal of a constant baseline. It can be seen that
+most bolometers exhibit a common time variation overlayed with other
+features. Empty squares indicate the locations of dead bolometers.}
 \label{fig:com}
@@
 bolometers share a common (but as yet unknown) calibration. This
-provides an alternative( or additional) flatfielding strategy to that
-described in in section~\ref{sec:bolos}.
+provides an alternative (or additional) flat-fielding strategy to that
+described in section~\ref{sec:bolos}.
 
@@
 bolometer. Such bolometers are identified by calculating the Pearson
-coefficient of correlation between each bolometer time stream and the
-common mode. Bolometers for which the correlation coefficient is below a
-specified limit, or which have a negative gain, are flagged as bad in
-order to omit them from the final map. If the option described above for
-handling time-varying sensitivities is used, then a correlation
-coefficient can be determined for each individual block of time slices.
-This allows individual bad blocks to be rejected from a bolometer time
-stream, rather than rejecting the whole bolometer. This is beneficial for
-bolometers that follow the common-mode for most of the time but suffer
-from transient problems at specific times.
+coefficient of correlation between each bolometer time-stream and the
+common mode. Bolometers for which the correlation coefficient is below
+a specified limit, or which have a negative gain, are flagged as bad
+in order to omit them from the final map. If the option described
+above for handling time-varying sensitivities is used, then a
+correlation coefficient can be determined for each individual block of
+time slices.  This allows individual bad blocks to be rejected from a
+bolometer time stream, rather than rejecting the whole bolometer. This
+is beneficial for bolometers that follow the common-mode for most of
+the time but suffer from transient problems at specific times.
 
 The common-mode is calculated from all bolometers that have not
-previously been flagged as bad. Thus aberrant bolometers with low
+previously been flagged as bad. Thus, aberrant bolometers with low
 correlation coefficients will be included, polluting the common-mode. For
@@
 \model{COM} model by examining data from all four sub-arrays in each
-waveband, since this allow spatial structure on the scale of the whole
-focal plane to be recovered. However, sometimes there is evidence that
-the common-mode differs from one array to another, and so an option
-exists to estimate a separate \model{COM} model for each individual
-subarray, with a consequent lowering in the scale of spatial structure
-that can be recovered.
+waveband, since this allows spatial structure on the scale of the
+whole focal plane to be recovered. However, sometimes there is
+evidence that the common-mode differs from one array to another, and
+so an option exists to estimate a separate \model{COM} model for each
+individual subarray, with a consequent lowering in the scale of
+spatial structure that can be recovered.
 
@@
 
-This model takes the FFT of the bolometer data, and can apply both
-high- and low-pass filters, as well as notch filteres, at hard
-frequency edges specified by the user. Alternatively, the frequency
-edges of the filters may be defined in terms of an \emph{angular
-  scale}, but converted into a frequency through knowledge of the mean
-telescope slew speed. The time-series may optionally have apodization
-applied before the transform to avoid ringing (primarily caused by
-wrap-around discontinuities at the ends of the time-series). Finally,
-a whitening filter may also be applied in which a simple $1/f^\alpha +
-\mathrm{constant}$ is fit to the power spectrum of each bolometer, and
-then the bolometer FFT is multiplied by its inverse (though normalized
-to preserve the white noise level).  The signal that is \emph{removed}
-from the time-series by this process are stored in the \model{FLT}
-container array. Typically this model is used purely as a high-pass
-filter to remove most of the $1/f$ noise. Low-pass filtering is
-redundant for two reasons: (i) SMURF already low-pass filters and
-re-samples to a lower sample rate as described in
-Section~\ref{sec:downsamp}; and (ii) the act of re-gridding the data
-to produce map estimates effectively low-pass filters the data at a
-frequency that corresponds to the inverse of the crossing time of a
-single map pixel. Notch filters have not been used with \scuba\ data
+This model takes the FFT of the bolometer time-stream data, and can
+apply both high- and low-pass filters, as well as notch filteres, at
+hard frequency edges specified by the user. Alternatively, the
+frequency edges of the filters may be defined in terms of an
+\emph{angular scale}, but converted into a frequency through knowledge
+of the mean telescope slew speed. The time-series may optionally have
+apodization applied before the transform to avoid ringing (primarily
+caused by wrap-around discontinuities at the ends of the
+time-series). Finally, a whitening filter may also be applied in which
+a simple $1/f^\alpha + \mathrm{constant}$ is fit to the power spectrum
+of each bolometer, and then the bolometer FFT is multiplied by its
+inverse (though normalized to preserve the white noise level).  The
+signals that are \emph{removed} from the time-series by this process
+are stored in the \model{FLT} container array. Typically this model is
+used purely as a high-pass filter to remove most of the residual $1/f$
+noise following common-mode removal. Low-pass filtering is redundant
+for two reasons: (i) SMURF already low-pass filters and re-samples to
+a lower sample rate, as described in Section~\ref{sec:downsamp}; and
+(ii) the act of re-gridding the data to produce map estimates
+effectively low-pass filters the data at a frequency that corresponds
+to the inverse of the crossing time of a single map pixel. Notch
+filters have not been proven to be useful with \scuba\ data,
 particularly since line features tend to move around (i.e. a dynamic
@@
 subsequently they are set to $1/\sigma_j^2$, the estimated inverse
-variance expected from the bolometer white noise levels as discussed
+variance expected from the bolometer white noise levels, as discussed
 in Section~\ref{sec:noi}. This weighting scheme is sensible provided
@@
 
-In addition, a variance map $\mathbf{v}(x_j,y_j)$ is estimated. The
-default behaviour is to estimate this weighted error on the mean given
-the scatter in the weighted samples. This is accomplished by dividing
-the \emph{biased weighted sample variance} by the number of samples
-that went into the average (akin to the formula for standard error on
-the mean, but accounting for weights),
+In addition to the signal map, a variance map $\mathbf{v}(x_j,y_j)$ is
+estimated. The default behaviour is to estimate this weighted error on
+the mean given the scatter in the weighted samples. This is
+accomplished by dividing the \emph{biased weighted sample variance} by
+the number of samples that went into the average (akin to the formula
+for standard error on the mean, but accounting for weights),
 %
@@
 
- We decided not to use the \emph{unbiased} estimator since in
-practice it would require accumulating an additional array of values
-at every map pixel, and only results in a small difference where there
-are less than $\sim10$ samples per pixel (a situation that is almost
-never encountered in a \scuba\ map except in the edge pixels).
+ We decided not to use the \emph{unbiased} estimator, since in practice
+it would require accumulating an additional array of values at every
+map pixel, and only results in a small difference where there are less
+than $\sim10$ samples per pixel (a situation that is almost never
+encountered in a \scuba\ map except in the edge pixels).
 
@@
 (Section~\ref{sec:mapdespike}), and to apply constraints to the map to
-improve convergences (Section~\ref{sec:converge}).
+improve convergence (Section~\ref{sec:converge}).
 
@@
 Since noise estimation is usually calculated as the final step in the
-iteration, it is the cleanest data for which some cleaning operations
-may be performed. For this reason \model{NOI} can optionally use the
-time-domain DC step fixer (Section~\ref{sec:steps}) and spike
-detection (Section~\ref{sec:timedespike}).
+iteration, the data at this stage have had most of the astronomical
+and other large, low-frequency noise signals removed. For this reason,
+\model{NOI} may optionally perform some cleaning options, such as the
+DC step fixer (Section~\ref{sec:steps}) and spike detection
+(Section~\ref{sec:timedespike}), which may work better with these
+cleaner time-streams.
 
 Finally, it should be noted that the default behaviour is to calculate
-the white noise levels \emph{once} within \emph{NOI}, after the second
+the white noise levels \emph{once} within \model{NOI}, after the second
 iteration. The reason for fixing these values is to prevent any
@@
 
-We don't use them, but we should mention them since a reviewer might
-ask us if we tried them: \model{DKS}, \model{TMP}, \model{SMO},
-\model{PLN}.
-
-Columns share significant magnetic field pickup which is measured by
-the dark squids. This model has been included although it makes little
-improvement, or injects noise since the magnetic field pickup noise is
-generally sub-dominant. Furthermore, not every column that contains
-working bolometers also has a working dark squid. An alternative is to
-use the \model{TMP} model described below.
-
-\textit{There are also the popped bolometers that could help fill the
-gaps.}
-
-Since the \model{DKS} cannot generally be used due to the missing
-dark squids, an alternative is to use the telescope pointing as a
-proxy for the signal due to magnetic field pickup
-(Section~\ref{sec:magpickup}. Generally speaking the azimuth (or its
-sine) as a function of time is used as a template that is simply
-scaled and removed from each bolometer signal. Some modest
-improvements can be achieved with this model.
-
-We also have the SMO filter that's useful for very short scans
-because edge effects don't matter so much. Two sorts - median which is
-more robust, or mean which is more accurate.
-
-We can remove a plane at each timeslice but it doesn't seem to improve
-things. Perhaps refer to other papers that claim to have seen coherent
-2d structures in the sky noise?
+Additional models exist as options, altough they are not generally
+used: \model{DKS}, the use of dark squids as a template for removing
+magnetic field pickup; \model{TMP} which uses the azimuth of the
+telescope also as a template for magnetic field pickup; \model{SMO},
+which provides a time-domain smoothing alternative to the FFT-based
+filter model \model{FLT}; and \model{PLN}, which fits a plane to the
+signal distribution across the focal plane at each time slice in an
+attempt to remove possible coherent atmospheric sky-noise structures.
+
+As described in Section~\ref{sec:magpickup}, wide scans can produce
+significant magnetic field pickup that tracks the azimuthal motion of
+the telescope. The \model{DKS} model uses the dark squid signals for
+each column as a template that is simply scaled (gain and offset) to
+each bolometer time-series before removal. Since not all columns have
+working dark squids, another alternative model is \model{TMP} in which
+the azimuth of the telescope itself is used as the template. In both
+cases, while the models had some success at removing the magnetic
+field pickup, it did not reduce the $1/f$ knee substantially,
+necessitating the continued use of the \model{FLT} model to remove the
+remaining low-frequency noise. Little or no difference in the final
+maps could be seen with or without the inclusion of these models, and
+so they are not generally used.
+
+The \model{SMO} model uses a rolling mean or median boxcar filter to
+calculate the low-frequency component of the bolometer signals, which
+are then removed. In other words, this is an alternative to the
+high-pass filtering for which \model{FLT} is generally used. The
+primary reason for developing this model was to make it more robust
+against ringing near the ends of the time-series (since the FFT
+assumes the signal is periodic, and there may be a large step between
+the start and finish), or residual spikes (for which the median
+filtering is particularly useful).
+
+Finally, we experimented with an alternative to \model{COM} for
+removing correlated atmospheric noise. Rather than subtracting the
+average signal at each time slice, \model{PLN} fits a plane to the
+observed signal at each instant. We found that there was no obvious
+improvement (either in terms of reducing the $1/f$ knee or reducing
+the noise in final maps). This result is basically consistent with
+those from earlier instruments, and suggests that the angular scale of
+the emitting regions in the atmosphere are un-resolved at the \scuba\
+focal plane. \citet{sayers2010}, in particular, examined several
+different ways of modelling and removing the correlated atmospheric
+noise from Bolocam data at 143 and 268\,GHz (also atop Mauna Kea),
+finding only marginal improvements by fitting a plane, or even using
+higher-order polynomials (see their Fig.~10, as well as the discussion
+and references to earlier work in Section~4.3).
 
@@
 deviations, $\sigma_i$, are measured for each of the residual
-bolometer signals once the modeled signal components have been
+bolometer signals once the modelled signal components have been
 removed. In the early iterations, the residual signals usually contain
@@
 the measured time-series bolometer variances and their white-noise
-levels measured between 2--10\,Hz. Note that this expression
+levels measured between 2 and 10\,Hz. Note that this expression
 \emph{does not account for the degrees of freedom}. Clearly there are
@@
 of the uncorrelated white noise, and potentially bias this estimate of
-$\chi^2_r$ low. One approach would be to keep track of the degrees of
-freedom, and therefore provide a correction, as in
-\citet{kovacs2008}. However, note that we are only using this value as
-a convergence test, and its value is not used in an absolute
+$\chi^2_\mathrm{r}$ low. One approach would be to keep track of the
+degrees of freedom, and therefore provide a correction, as in
+\citet{kovacs2008}. However, note that we are only using this quantity
+as a convergence test, and its value is not used in an absolute
 sense. Furthermore, the reference white noise values are calculated
 only once in SMURF (Section~\ref{sec:noi}), and the number of model
-parameters are fixed, so the value of $\chi^2_r$ should decrease
-monotonically if the solution is behaving well. The convergence
-criterion is met if the change in subsequent values of $\chi^2_r$ is
-smaller than the parameter \texttt{chitol}, which is set to $10^{-3}$
-by default.
+parameters are fixed, so the value of $\chi^2_\mathrm{r}$ should
+decrease monotonically if the solution is behaving well. The
+convergence criterion is met if the change in subsequent values of
+$\chi^2_\mathrm{r}$ is smaller than the parameter \texttt{chitol},
+which is set to $10^{-3}$ by default.
 
 It was immediately found in the analysis of \scuba\ data that the
-$\chi^2_r$ test described above was not a sufficient stopping
+$\chi^2_\mathrm{r}$ test described above i not a sufficient stopping
 criterion. The reason for this is that various components of the model
@@
 
-The map-based convergence statistic, $M_c$, is the average absolute
-change in the value of map pixels between subsequent iterations,
-normalized by the map pixel uncertainties (square root of
+The map-based convergence statistic, $M_\mathrm{c}$, is the average
+absolute change in the value of map pixels between subsequent
+iterations, normalized by the map pixel uncertainties (square root of
 Eq.~\ref{eq:varmap}), or
@@
 \begin{equation}
-M^j_c = \frac{1}{N} \sum_i \frac{| \mathbf{m}(x_j,y_j) -
+M^j_\mathrm{c} = \frac{1}{N} \sum_i \frac{| \mathbf{m}(x_j,y_j) -
   \mathbf{m}(x_{j-1},y_{j-1}) |} {\sqrt{\mathbf{v}(x_j,y_j)}} ,
@@
 %
-where $i$ runs over the $N$ map pixels, and $j$ enumerates
-iteration. The convergence criterion is met once this average change
+where $i$ runs over the $N$ map pixels, and $j$ enumerates the
+iterations. The convergence criterion is met once this average change
 is smaller than the parameter \texttt{maptol}. Experimentally we found
@@
 %-------------------------------------------------
-\subsection{Data cleaning}
+\subsection{Data preparation}
 %-------------------------------------------------
 
-Since we can do most of this stuff either as a pre-processing step or
-iteratively, I thinks this goes in a generic-sounding subsection. This
-can go here, or perhaps earlier?
+In this section we describe a number of data preparation steps that
+preceed the calculation of model components as described in the
+previous section.
 
@@
 fractional samples from $x_i$ are used at the $y_j$ time
-boundaries. The algorithm is fast since each sample in $x_i$ need only
-be visited once. From a spectral point of view, this boxcar average is
-equivalent to applying a sinc-function low-pass filter in frequency
-space. Such behaviour is desirable since it serves as an anti-aliasing
-filter; most non-white features above the Nyquist frequency in $y_j$
-that could contaminate the output data are largely removed.
+boundaries. The algorithm is fast, since each sample in $x_i$ need
+only be visited once. From a spectral point of view, this boxcar
+average is equivalent to applying a sinc-function low-pass filter in
+frequency space. Such behaviour is desirable, since it serves as an
+anti-aliasing filter; most non-white features above the Nyquist
+frequency in $y_j$ that could contaminate the output data are largely
+removed.
 
@@
 
-\subsubsection{Bolometer Filtering}
-
-Despite the ability of the map-maker to iterative remove many noise
+\subsubsection{Bolometer filtering}
+
+Despite the ability of the map-maker to iteratively remove many noise
 components, under some circumstances it may be desirable to filter the
@@
 simple high-pass filtering, this method has not yet been explored in
-detail with \scuba\ data.
+detail with \scuba\ data. However, since systematic effects seem to
+come and go, it is possible that PCA could be useful for particular
+data sets.
 
@@
 series data. If not removed, they can cause ringing in the FLT model
-(Section~\ref{sec:flt}). Two alternative approaches may be used to remove
-these spikes. This section describes the detection and removal of spikes
-within the time series of each bolometer, and
+(Section~\ref{sec:flt}). Two alternative approaches may be used to
+remove these spikes. This section describes the detection and removal
+of spikes within the time series of each bolometer, and
 section~\ref{sec:mapdespike} describes the detection and removal of
@@
 
-\emph{Do we need some discussion of the physical causes and
-characteristics of spikes?}
-
-Each one-dimensional bolometer time series is processed independently. At
-each time slice, the median value of the current bolometer in a box
-centred on the time slice, and containing a specified number of time
-slices (typically 50), is found. If the residual between the
-time slice value and the median value is greater than some specified
-multiple (typically 10) of the local noise level, the time slice is
-flagged as a spike.
+Each one-dimensional bolometer time series is processed
+independently. At each time slice, the median value of the current
+bolometer is found in a box centred on the time slice, and containing
+a specified number of time slices (typically 50). If the residual
+between the time slice value and the median value is greater than some
+specified multiple (typically 10) of the local noise level, the time
+slice is flagged as a spike.
 
@@
 noise level is taken as the standard deviation of the values within a
-neighbouring box on the ``down-stream'' side of the median box (that is,
-the side that has already been checked for spikes). That is, the high end
-of the noise box is just below the low end of the median filter box. This
-introduces a slight asymmetry in the noise, but this should not matter
-unless the noise varies significantly on a time scale shorter than the
-box size.
+neighbouring box on the ``down-stream'' side of the median box (that
+is, the side that has already been checked for spikes). In other
+words, the high end of the noise box is just below the low end of the
+median filter box. This introduces a slight asymmetry in the noise,
+but this should not matter unless the noise varies significantly on a
+time scale shorter than the box size.
 
@@
 Steps occur with a wide range of heights and shapes. The ratio of step
-height to noise can vary from less than 10 to several hundred. Some steps
-occur over a single sample such as the step close to sample 5000 in the
-right-hand plot of Fig.~\ref{fig:steps1}, but others happen more
-gradually such as the step close to sample 5300. In addition, a step can
-be preceded or followed by a short period of instability, as is visible
-at the bottom of the step in the left hand plot of
-Fig.~\ref{fig:steps1}. Further problems are caused by steps that occur
-close together in time, such as the large downward step followed by a
-smaller upward step close to sample 5000 in the right hand plot of
-Fig.~\ref{fig:steps1}.
+height to noise can vary from less than 10 to several hundred. Some
+steps occur over a single sample, such as the step close to sample
+5000 in the right-hand plot of Fig.~\ref{fig:steps1}, but others
+happen more gradually, such as the step close to sample 5300. In
+addition, a step can be preceded or followed by a short period of
+instability, as is visible at the bottom of the step in the left hand
+plot of Fig.~\ref{fig:steps1}. Further problems are caused by steps
+that occur close together in time, such as the large downward step
+followed by a smaller upward step close to sample 5000 in the right
+hand plot of Fig.~\ref{fig:steps1}.
 
@@
 to be a challenge. The current step-correction algorithm is described in
-detail in appendix~\ref{appendix:steps}. In outline, the following stages
+detail in Appendix~\ref{appendix:steps}. In outline, the following stages
 are involved in detecting steps in a single bolometer time stream:
@@
 \begin{enumerate}
-\item Median smooth the whole time stream.
-\item Find the gradient of the median smoothed time series at each sample.
-\item Smooth the gradient values to determine the local mean gradient and
-subtract this local mean from the total gradient to get residual gradients.
-\item Find residual gradient values that exceed 25 times the local RMS of
-the residual gradients.
+
+\item median smooth the whole time stream;
+
+\item find the gradient of the median smoothed time series at each
+sample;
+
+\item smooth the gradient values to determine the local mean gradient
+and subtract this local mean from the total gradient to get residual
+gradient;
+
+\item find residual gradient values that exceed 25 times the local RMS
+of the residual gradients;
+
 \item Group these high residual gradients into contiguous blocks of
-samples.
-\item Merge blocks that are separated by less than 100 samples
+samples;
+
+\item merge blocks that are separated by less than 100 samples.
 \end{enumerate}
@@
 
-\item The above process can mis-interpret the edges of a bright source
-as a step. So we ignore blocks that occur close to bright sources.
-
-\item If the block passes the above test, a least squares linear fit is
-performed to the median-filtered bolometer data just before the block, and
-this fit is extrapolated to predict the data value expected at the centre
-of the block on the basis of the preceding data.
-
-\item A least squares linear fit is performed to the median-filtered
+\item the above process can mis-interpret the edges of a bright source
+as a step, so we ignore blocks that occur close to bright sources;
+
+\item if the block passes the above test, a least squares linear fit
+is performed to the median-filtered bolometer data just before the
+block, and this fit is extrapolated to predict the data value expected
+at the centre of the block on the basis of the preceding data;
+
+\item a least squares linear fit is performed to the median-filtered
 bolometer data just after the block, and this fit is extrapolated to
-predict the data value expected at the centre of the block on the basis
-of the following data.
-
-\item The difference between these two expected data values is taken as
-the step height.
-
-\item The preceding three steps are repeated several times, each time
-including a different selection of samples in the two least squares fits.
-The mean and standard deviation of the corresponding set of step heights
-is found.
-
-\item If the mean step height is small compared to the standard deviation
+predict the data value expected at the centre of the block on the
+basis of the following data;
+
+\item the difference between these two expected data values is taken
+as the step height;
+
+\item the preceding three steps are repeated several times, each time
+including a different selection of samples in the two least squares
+fits, with the mean and standard deviation of the corresponding set of
+step heights found;
+
+\item if the mean step height is small compared to the standard deviation
 of the step heights, or compared to the noise in the bolometer data, then
-the step is ignored.
-
-\item If the above checks are passed, all subsequent bolometer samples are
-corrected by the mean step height.
-
-\item Bolometer samples within the duration of the step, and a few
+the step is ignored;
+
+\item if the above checks are passed, all subsequent bolometer samples
+are corrected by the mean step height;
+
+\item bolometer samples within the duration of the step, and a few
 samples on either side, are flagged as unusable.
@@
 
-The evaluation of the FLT model (Section~\ref{sec:flt}) involves taking
-the Fast Fourier Transform of each bolometer time series. Data that has
-been flagged as bad for any reason (for instance, due to the presence of
-spikes, steps, or unusual common-mode signal) needs to be excluded from
-this FFT. Two method are available:
+The evaluation of the \model{FLT} model (Section~\ref{sec:flt})
+involves taking the Fast Fourier Transform (FFT) of each bolometer
+time series. Data that have been flagged as bad for any reason (for
+instance, due to the presence of spikes, steps, or unusual common-mode
+signal) need to be excluded from this FFT. Two methods are available:
 
@@
 fits.
-\item Masking: each bad data value is replaced by zero, and an associated
-mask is created that holds 1.0 for each good sample in the time stream and
-0.0 for each bad sample. The masked data and the mask are then each filtered
-in the same way using FFTs, and the filtered data is finally normalised by
-dividing it by the filtered mask.
+
+\item Masking: each bad data value is replaced by zero, and an
+associated mask is created that holds 1.0 for each good sample in the
+time stream and 0.0 for each bad sample. The masked data and the mask
+are then each filtered in the same way using FFTs, and the filtered
+data are finally normalised by dividing them by the filtered mask.
 \end{itemize}
 
-In addition to replacing bad samples before the FFT, it is also necessary
-to ensure that the data values at the start and end of the time series
-are similar. Since a Fast Fourier transform treats the data as a single
-cycle in an infinitely repeating waveform, any large difference between
-starting and ending value will effectively introduce sudden steps at the
-start and end of each cycle, causing unwanted oscillations (ringing) in
-the transform. Another consequence of the cyclic nature of the FFT is
-that features at one end of the time series can affect the filtered
-values at the other end of the time series. Two methods are available to
-avoid these two problems:
+In addition to replacing bad samples before the FFT, it is also
+necessary to ensure that the data values at the start and end of the
+time series are similar. Since an FFT treats the data as a single
+cycle in an infinitely repeating waveform, any large difference
+between starting and ending value will effectively introduce sudden
+steps at the start and end of each cycle, causing unwanted
+oscillations (ringing) in the transform. Another consequence of the
+cyclic nature of the FFT is that features at one end of the time
+series can affect the filtered values at the other end of the time
+series. Two methods are available to avoid these two problems:
 
@@
 
-\item Padding with artificial data: instead of padding with zeros, each
-time stream is padded with artificial data that connects the two ends of
-the data stream smoothly and includes Gaussian noise. No apodisation is
-performed. The number of samples of padding at each end is again equal to
-the ratio of the sampling frequency to the lowest retained frequency.
-This is illustrated in Fig.~\ref{fig:pad1}.
+\item Padding with artificial data: instead of padding with zeros,
+each time stream is padded with artificial data that connects the two
+ends of the data stream smoothly and includes Gaussian noise. No
+apodisation is performed. The number of samples of padding at each end
+is again equal to the ratio of the sampling frequency to the lowest
+retained frequency.  This is illustrated in Fig.~\ref{fig:pad1}. See
+\citet{stompor2002} for a thorough discussion of this procedure within
+the context of CMB map-making.
 
@@
 \caption{The black curve shows the same bolometer time stream as in
-Fig.~\ref{fig:pad2}, againpadded with zeros. The red curve shows the time stream
-after padding with artificial data.}
+Fig.~\ref{fig:pad2}, againpadded with zeros. The red curve shows the
+time stream after padding with artificial data.}
 \label{fig:pad1}
@@
 \end{itemize}
-
-\emph{Should reference CMB literature -- in particular look at
-\citet{stompor2002}}.
 
@@
 outliers. This method is more robust against false-positive detections
-of bright/compact astronomical sources since transient features in the
+of bright/compact astronomical sources, since transient features in the
 time-series are unlikely to occur by chance whenever a bolometer
@@
 of this new variable gives the weighted sample variance of the
-underlying data points,
+underlying data points:
 %
@@
 In this section three different common types of data are reduced: a
-bright point source, Uranus (Section~\ref{sec:point}); a bind survey
+bright point source, Uranus (Section~\ref{sec:point}); a blind survey
 of high-redshift galaxies in the Lockman Hole
@@
 source for \scuba\ that is commonly used as a primary flux
-calibrator. The now standard CV daisy scan pattern was used, with a
-scan speed of 155 arcsec\,sec$^{-1}$. We perform several different
-reductions of the data to illustrate the purpose of different model
-components and the convergence properties of the solution
-(Fig.~\ref{fig:pointmaps}). In all cases the maps are produced on a
-grid of azimuth (horizontal) and elevation (vertical) offsets from the
-position of Uranus (the origin), using 2\,arcsec pixels.
+calibrator. The now standard constant-velocity daisy scan pattern was
+used (Section~\ref{sec:scan}), with a scan speed of 155
+arcsec\,sec$^{-1}$. We perform several different reductions of the
+data to illustrate the purpose of various model components and the
+convergence properties of the solution (Fig.~\ref{fig:pointmaps}). In
+all cases the maps are produced on a grid of azimuth (horizontal) and
+elevation (vertical) offsets from the position of Uranus (the origin),
+using 2\,arcsec pixels.
 
@@
 and final iteration. The resulting map after these two iterations is
-shown in Fig.~\ref{fig:pointmaps}a. While the S/N of Uranus is clearly
+shown in Fig.~\ref{fig:pointmaps}a. While the \snr\ of Uranus is clearly
 large ($\sim$250 as estimated by SMURF), enabling us to see the faint
@@
 contribution to the larger-scale ripples in the map, however, can be
-made by degeneracies in the map solution as we discuss below.
+made by degeneracies in the map solution, as we discuss below.
 
@@
 $\sinc$ function-like response in map space. Since the scan pattern is
-isotropic (scans equally at all position angles), and there is a
-bright point-like source, the result is an azimuthally-symmetric
-$\sinc$ function-like pattern in the map. Note that due to the
-iterative nature of the solution, much of the signal in the source is
-actually removed for subsequent iterations of \model{FLT}, but as in
-the previous case, degeneracies between the map and the \model{FLT}
-and \model{COM} signal components can still lead to complicated
-structures.
+fairly isotropic (scans at all position angles), and there is a bright
+point-like source at the centre, the result is an
+azimuthally-symmetric $\sinc$ function-like pattern in the map. Note
+that due to the iterative nature of the solution, much of the signal
+in the source is actually removed for subsequent iterations of
+\model{FLT}, but as in the previous case, degeneracies between the map
+and the \model{FLT} and \model{COM} signal components can still lead
+to complicated structures.
 
@@
 the map to a value of zero away from the known locations of
-emission. In Fig.~\ref{fig:pointmaps}d, a map is produced in an
+emission. In Fig.~\ref{fig:pointmaps}d, a solution is produced in an
 identical manner to the previous example, but now setting the map
@@
 \includegraphics[width=\linewidth]{pointmaps.pdf}
-\caption{Multiple reductions of a 450\,\micron\ daisy scan of Uranus,
-  all scaled between -0.002\,pW (white) and +0.002\,pW (black): (a)
-  Reduction in which only common-mode subtraction is used to suppress
-  low-frequency noise, and the reduction is forced to use 2 iterations
-  (after the first iteration an estimate of the source flux is
-  removed, and the noise is measured in the residuals to obtain
-  appropriate weighting for the second and final iteration) ---
-  circular streaks are caused by independent low-frequency noise that
-  is not removed by the common-mode (Uranus peak 0.15\,pW); (b) same
-  as (a) but using 50 iterations, illustrating the degeneracy between
-  large-scale structure and the common-mode (the footprint of working
-  bolometers is also shown for reference, Uranus peak 0.15\,pW); (c)
-  reduction in which high-pass filtering above 0.775\,Hz
-  (corresponding to a spatial scale of 200\,arcsec as indicated) is
-  applied after common-mode removal, but before the map estimate ---
-  the map-based convergence test is activated and the solution halts
-  after 10 iterations, but leaving large-scale ringing due to the
-  filter (Uranus peak 0.27\,pW); (d) same as (c), but now the region
-  of the map beyond the white dot-dashed circle is constrained to a
-  value of zero until all but the final iteration --- the map is now
-  extremely flat, there is little attenuation of the source flux, and
-  the diffraction pattern is clearly seen (Uranus peak 0.27\,pW).}
+\caption{Multiple reductions of a 450\,\micron\ constant-velocity
+  daisy scan of Uranus, all scaled between $-$0.002\,pW (white) and
+  +0.002\,pW (black). (a) Reduction in which only common-mode
+  subtraction is used to suppress low-frequency noise, and the
+  reduction is forced to use 2 iterations (after the first iteration
+  an estimate of the source flux is removed, and the noise is measured
+  in the residuals to obtain appropriate weighting for the second and
+  final iteration). Circular streaks are caused by independent
+  low-frequency noise that is not removed by the common-mode (Uranus
+  peak 0.15\,pW). (b) Same as (a) but using 50 iterations,
+  illustrating the degeneracy between large-scale structure and the
+  common-mode (the footprint of working bolometers is also shown for
+  reference, Uranus peak 0.15\,pW). (c) Reduction in which high-pass
+  filtering above 0.775\,Hz (corresponding to a spatial scale of
+  200\,arcsec as indicated) is applied after common-mode removal, but
+  before the map estimate. The map-based convergence test is activated
+  and the solution halts after 10 iterations, but leaving large-scale
+  ringing due to the filter (Uranus peak 0.27\,pW). (d) Same as (c),
+  but now the region of the map beyond the white dot-dashed circle is
+  constrained to a value of zero until all but the final iteration.
+  The map is now extremely flat, there is little attenuation of the
+  source flux, and the diffraction pattern is clearly seen (Uranus
+  peak 0.27\,pW).}
 \label{fig:pointmaps}
@@
   lines in the top panel show the residual signal for a single
-  bolometer after 2 and 100 iterations respectively (they are nearly
-  identical). The green and red lines show the \emph{difference}
-  between the \model{COM} and \model{AST} (the signal produced by the
-  current map estimate for a given bolometer) model components for
-  that bolometer between iterations 2 and 100 respectively. This shows
-  that a strong signal has grown over time, and it has equal but
-  opposite signs in the two model components, so that they cancel one
-  another. The bottom panel shows the scan pattern of the telescope
-  over the same period; clearly the \model{COM}/\model{AST} model
-  degeneracy is correlated with the elevation offset, and referring to
-  Fig.~\ref{fig:pointmaps}, panel (b), this clearly corresponds to the
-  strong vertical gradient that has appeared}.
+  bolometer after 2 and 100 iterations, respectively (they are nearly
+  identical; the black line lies beneath the grey line). The green and
+  red lines show the \emph{difference} between the \model{COM} and
+  \model{AST} (the signal produced by the current map estimate for a
+  given bolometer) model components for that bolometer between
+  iterations 2 and 100, respectively. This shows that a strong signal
+  has grown over time, and it has equal but opposite signs in the two
+  model components, so that they cancel one another. The bottom panel
+  shows the scan pattern of the telescope over the same period;
+  clearly the \model{COM}/\model{AST} model degeneracy is correlated
+  with the elevation offset, and referring to
+  Fig.~\ref{fig:pointmaps}, panel (b), this corresponds to the strong
+  vertical gradient that has appeared}.
 \label{fig:degeneracy}
@@
 ideally limited by the white-noise performance of the instrument. The
-approach described here for maximizing the S/N of point-sources
+approach described here for maximizing the \snr\ of point-sources
 involves three major steps: (i) generating a map that removes most
@@
 community using previous instruments
-\citep[e.g.][]{scott2002,laurent2005,coppin2006,scott2008,perera2008,devlin2009}.
+\citep[e.g.][]{scott2002,borys2003,laurent2005,coppin2006,scott2008,perera2008,devlin2009}.
 In this section we will reduce scans of the Lockman Hole taken during
@@
 of about 50\,arcmin$^2$. The full list of dates and observation
-numbers includes 20100218: 63, 64, 65, 70, 71, 72, 90, 91, 92, 97, 98,
-99; 20100220: 111, 112, 113, 118, 128, 129, 130; 20100303: 61, 62, 64,
-69, 70, 72, 73, 74, 75; 20100309: 87, 88, 90, 91; and 20100311: 59,
-64, 65, 72. These observations represent about 80\% of the total data
-taken as part of the project; the remaining observations were dropped
-due to problems keeping the arrays properly locked during S2SRO, and
-were easily identified by their highly variable and erratic bolometer
-time-series (which resulted in maps full of large streaks).
+numbers includes: 20100218, 63, 64, 65, 70, 71, 72, 90, 91, 92, 97,
+98, 99; 20100220, 111, 112, 113, 118, 128, 129, 130; 20100303, 61, 62,
+64, 69, 70, 72, 73, 74, 75; 20100309, 87, 88, 90, 91; and 20100311,
+59, 64, 65, 72. These observations represent about 80 per cent of the
+total data taken as part of the project; the remaining observations
+were dropped due to problems keeping the arrays properly locked during
+S2SRO, and were easily identified by their highly variable and erratic
+bolometer time-series (which resulted in maps full of large streaks).
 
@@
 \includegraphics[width=\linewidth]{lockman_maps.pdf}
-\caption{Reduction of a blank-field survey: the Lockman Hole. (a) raw
+\caption{Reduction of a blank-field survey: the Lockman Hole. (a) Raw
   output of SMURF using high-pass filtering as a pre-processing step,
@@
   noise levels of 1.25, 2.5 and 5.0 times the minimum noise at the
-  centre of the map. (b) the whitened map using a filter based on the
+  centre of the map. (b) The whitened map using a filter based on the
   angular power spectrum of the jackknife map. (c) The whitened map
@@
   [restricted to a lower-noise region, within the area denoted by the
-  second contour of panel (a)]. The image shows the S/N, with
+  second contour of panel (a)]. The image shows the \snr\, with
   3.8-$\sigma$ peaks indicated with blue circles (radius
@@
   10 submm peaks, 9 are within this search radius of at least one
-  radio source. (d) jackknife map produced from the difference of two
+  radio source. (d) Jackknife map produced from the difference of two
   maps, using the even and odd scan numbers, respectively. Provided
@@
   indicate radio sources and 3.8-$\sigma$ peaks, respectively. Unlike
-  panel (c), there is only a single significant peak, and it is not in
-  the vicinity of a radio source.}
+  panel (c), there is only a single (apparently) significant peak, and
+  it is not in the vicinity of a radio source.}
 \label{fig:lockman_maps}
@@
 
-%... high-pass
-%filter + iterative common-mode removal takes place of PCA cleaning in
-%AzTEC reduction ... probably more linear.
-
 Even though the map looks quite flat, there is a mixture of faint
 astronomical sources, and what is probably residual low-frequency
-noise causing faint ripples visible to the naked eye. Since the
+noise,q causing faint ripples visible to the naked eye. Since the
 mixture of the two components is unknown, the first step is to
@@
 \begin{eqnarray}
-S_\mathrm{JK} = \frac{S_1 - S_2}{2} \\
+S_\mathrm{JK} = \frac{S_1 - S_2}{2}, \\
 \sigma^2_\mathrm{JK} = \frac{\sigma^2_1 + \sigma^2_2}{4}.
@@
 map will contain a nearly evenly-spaced sampling of data across the
-full dataset).
+full data set).
 
@@
 (scales $\lsim16$\,arcsec), with the exception of a line at
-$\sim0.175$\,arcsec$^{-1}$ (a scale of $\sim5.7$\,arcsec). A theory
-for this feature is that it is related to the inter-bolometer spacing
-in the focal plane (which happens to be this size): small relative
-drifts in the baselines of adjacent bolometers may produce faint
-parallel stripes in the map along the scan direction (the
-superposition of many scans at different angles then results in an
-isotropic noise pattern). It is not likely that this signal is due to
-astronomical sources because it appears with a nearly equal amplitude
-in both the real map and the jackknife.  At lower spatial frequencies,
-both the real map and the jackknife power spectra grow significantly,
-as a result of the more obvious large-scale undulations in
-Figs.~\ref{fig:lockman_maps}(a) and (d).
+$\sim0.175$\,arcsec$^{-1}$ (a scale of $\sim5.7$\,arcsec). One
+possibility for this feature is that it is related to the
+inter-bolometer spacing in the focal plane (which happens to be this
+size): small relative drifts in the baselines of adjacent bolometers
+may produce faint parallel stripes in the map along the scan direction
+(the superposition of many scans at different angles then results in
+an isotropic noise pattern). It is not likely that this signal is due
+to astronomical sources because it appears with a nearly equal
+amplitude in both the real map and the jackknife.  At lower spatial
+frequencies, both the real map and the jackknife power spectra grow
+significantly, as a result of the more obvious large-scale undulations
+in Figs.~\ref{fig:lockman_maps}(a) and (d).
 
@@
   (panels (a), (d), and (b) from Fig.~\ref{fig:lockman_maps},
-  respectively (considering only the square region indicated in
-  Fig.~\ref{fig:lockman_maps}d). Since the raw map contains spatially
-  correlated signals on large scales both due to noise and
+  respectively, considering only the square region indicated in
+  Fig.~\ref{fig:lockman_maps}d. Since the raw map contains spatially
+  correlated signals on large scales, both due to noise and
   astronomical signals (low-spatial frequencies), the jackknife map
@@
 and for comparison, the jackknife map has also been whitened in
-Fig.~\ref{fig:lockman_maps}e. In both cases, the maps are both
-visually flatter than the non-whitened cases.
+Fig.~\ref{fig:lockman_maps}e. In both cases, the maps are visually
+flatter than the non-whitened cases.
 
@@
 this plot alone; nor are astronomical sources readily visible in the
-map (Fig.~\ref{fig:lockman_maps}b. To identify sources, we next apply
+map (Fig.~\ref{fig:lockman_maps}b). To identify sources, we next apply
 a \emph{matched filter} to the whitened maps.
@@
 shown for the whitened signal, and jackknife maps in
-Figures~\ref{fig:lockman_maps}c and f respectively. Note that these
-images are plotted in SNR units, where the smooth noise maps have been
-calculated by propagating the original noise maps output by SMURF both
-through the whitening and matched filters (each of which are linear
-operations).
+Figs.~\ref{fig:lockman_maps}c and f, respectively. Note that these
+images are plotted in \snr\ units, where the smooth noise maps have
+been calculated by propagating the original noise maps output by SMURF
+both through the whitening and matched filters (each of which are
+linear operations).
 
@@
 typical for other ground-based submillimetre surveys in recent years
-\citep[e.g.,][]{coppin2006,perera2008,2009ApJ...707.1201W} leading to estimated
-false-identification rates of order $\sim$5\%, and is chosen here as a
-convenient reference. In the former, 10 peaks are found, whereas there
-is only 1 in the latter. However, this test does not preclude the
-possibility that some correlated noise made it into the jackknife
-map, in which case the estimated SNRs are misleading.
+\citep[e.g.,][]{coppin2006,perera2008,2009ApJ...707.1201W} leading to
+estimated false-identification rates of order $\sim$5\%, and is chosen
+as a convenient reference. In the former, 10 peaks are found, whereas
+there is only 1 in the latter. However, this test does not preclude
+the possibility that some correlated noise made it into the jackknife
+map, in which case the estimated \snr s are misleading.
 
 One simple test of the calculated noise properties is to compare the
-signal and jackknife SNR distributions with ideal Gaussians. The top
-panel of Fig.~\ref{fig:lockman_hist} shows the whitened (but not match
-filtered) signal (blue) and jackknife (histograms), along with a
+signal and jackknife \snr\ distributions with ideal Gaussians. The top
+panel of Fig.~\ref{fig:lockman_hist} shows the whitened (but not
+match-filtered) signal (blue) and jackknife (histograms), along with a
 Gaussian (mean 0, standard deviation 1, and area normalized to the
 number of map pixels) as a dashed line. In this case, it is clear that
-the SNR distributions for both maps are nearly indistinguishable from
-the theoretical distribution of white noise. This result shows us
+the \snr\ distributions for both maps are nearly indistinguishable
+from the theoretical distribution of white noise. This result shows us
 that: (i) the whitening filter appears to have removed correlated
-large-scale noise since the jackknife map histogram is consistent with
-white noise; and (ii) any potential astronomical signals are small
-compared to the typical white noise in most map pixels (unsurprising
-given the appearance of Fig.~\ref{fig:lockman_maps}b). Next, we
-examine the SNR histograms for maps processed with the matched filter
-in the bottom panel of Fig.~\ref{fig:lockman_hist}. Again, the
-histogram of the jackknife SNR data appears consistent with pure
-noise. However, the signal map now deviates significantly from a
-Gaussian distribution, with a clear \emph{positive} tail (as one would
-expect for emitting sources). In fact, integrating the positive tails
-beyond our 3.8-$\sigma$ source-detection threshold (vertical solid
-line) yields 229 map pixels in the signal map, compared with 3 in the
-jackknife map (out of 80603 pixels in the entire region). Note that
-given the small pixel size of the map, several pixels generally
-contribute to each peak. Note that this procedure gives only a flavour
-of the analysis that is usually required to produce a robust source
-lists. Additional tests, along with a careful consideration of
-completeness and false-positive rates usual require a series of Monte
-Carlo simulations that are beyond the scope of this work. We direct
-the interested reader to a selection of papers on the subject:
+large-scale noise, since the jackknife map histogram is consistent
+with white noise; and (ii) any potential astronomical signals are
+small compared to the typical white noise in most map pixels
+(unsurprising given the appearance of
+Fig.~\ref{fig:lockman_maps}b). Next, we examine the \snr\ histograms
+for maps processed with the matched filter in the bottom panel of
+Fig.~\ref{fig:lockman_hist}. Again, the histogram of the jackknife
+\snr\ data appears consistent with pure noise. However, the signal map
+now deviates significantly from a Gaussian distribution, with a clear
+\emph{positive} tail (as one would expect for emitting sources). In
+fact, integrating the positive tails beyond our 3.8-$\sigma$
+source-detection threshold (vertical solid line) yields 229 map pixels
+in the signal map, compared with 3 in the jackknife map (out of 80603
+pixels in the entire region). Recall that given the small pixel size
+of the map, several pixels generally contribute to each peak. Note
+that this procedure gives only a flavour of the analysis that is
+usually required to produce a robust source lists. Additional tests,
+along with a careful consideration of completeness and false-positive
+rates usual require a series of Monte Carlo simulations that are
+beyond the scope of this work. We direct the interested reader to a
+selection of papers on the subject:
 \citet{scott2002,coppin2006,perera2008,2009ApJ...707.1201W} and
@@
 \includegraphics[width=\linewidth]{lockman_hist.pdf}
-\caption{Histograms of S/N using pixels from the central region of
-  Lockman Hole (second contour) in Fig.~\ref{fig:lockman_maps}a. The
-  top panel shows the histograms for pixels from the whitened signal
-  and jackknife maps (Figs.~\ref{fig:lockman_maps}b,e), compared to a
-  Gaussian distribution with mean zero, standard deviation one, and an
-  area normalized to the total number of pixels -- the expected
+\caption{Histograms of \snr\ using pixels from the central region of
+  the Lockman Hole (second contour) in
+  Fig.~\ref{fig:lockman_maps}a. The top panel shows the histograms for
+  pixels from the whitened signal and jackknife maps
+  (Figs.~\ref{fig:lockman_maps}b,e), compared to a Gaussian
+  distribution with mean zero, standard deviation one, and an area
+  normalized to the total number of pixels -- the expected
   distribution for a map of spatially-uncorrelated Gaussian noise. The
@@
   expectation (Gaussian), but now the matched-filter has picked out
-  significant signal leading to the large positive tail. The vertical
+  significant signal, leading to the large positive tail. The vertical
   solid line shows the chosen 3.8-$\sigma$ source-detection
@@
 representative search radius of 8\,arcsec from these studies with
-similar SNR sources and source sizes (the same size as the blue
+similar \snr\ sources and source sizes (the same size as the blue
 circles), 9 out of 10 peaks in the smoothed signal map have potential
@@
 \includegraphics[width=\linewidth]{m17.pdf}
-\caption{An 850\,\micron\ rotating pong map of M17. Intensity is
-logarithmically scaled between -0.0003\,pw (white) and +0.01\,pW
+\caption{An 850\,\micron\ rotating PONG map of M17. Intensity is
+logarithmically scaled between $-$0.0003\,pw (white) and +0.01\,pW
 (black). Iteration numbers are given in the corner of each
@@
 which a zero-mask is created iteratively from all of the pixels that
-exceed a SNR of 5. While this region (red contour) only encompasses
-the brightest peaks early in the solution, the final iteration it
-identifies most of the bright, extended emission, and it significantly
+exceed an \snr\ of 5. While this region (red contour) only encompasses
+the brightest peaks early in the solution, in the final iteration it
+identifies most of the bright, extended emission, and significantly
 helps with negative ringing.}
@@
 The default reduction of these data is shown in the top panels of
-Fig.~\ref{fig:m17}, after iterations 2 (the first map estimated after
+Fig.~\ref{fig:m17}, after 2 iterations (the first map estimated after
 the noise weights have been measured) and 17 iterations (when the map
@@
 regions containing blank sky. Indeed, for this map, much of the field
-clearly contains extended structures. Furthermore, the goal of such
+clearly contains extended structure. Furthermore, the goal of such
 maps may be to detect \emph{previously unknown} cool, dense regions of
@@
 facility has been added to SMURF to generate one automatically by
-flagging pixels above some SNR threshold after each iteration as part
+flagging pixels above some \snr\ threshold after each iteration as part
 of the ``bright extended'' configuration.
@@
 edge scale was adopted (indicated in the top panels). First (top) row:
-Average of two halves of the data analyzed independently, using the
+average of two halves of the data analyzed independently, using the
 default configuration. The data have had added to them synthetic
-signal within a 600\,arcsec diameter circle (S.E. of M17) created as a
-realization of noise from a $P(k) \propto k^{-3}$ angular power
+signal within a 600\,arcsec diameter circle (south of M17) created as
+a realization of noise from a $P(k) \propto k^{-3}$ angular power
 spectrum multiplied by the PSF, subtracting the minimum to make it
@@
 first row. Third row: average of the two halves of the data using the
-bright extended reduction. The blue and red contours indicated the
-zero-masked regions in each case (note for the 300\,arcsec case that
-the region about M17 matches fairly closely the mask in
-Fig.~\ref{fig:m17}d). Clearly as the filter scale is increased, due to
-the high SNR of the data, larger emissions regions are (correctly)
-detected and reproduced in the map. Fourth (bottom) row: Jackknife
-maps for the bright extended reductions.}
+bright extended reduction. The blue and red contours indicate the
+zero-masked regions for each half of the data (note for the
+300\,arcsec case that the region about M17 matches fairly closely the
+mask in Fig.~\ref{fig:m17}d). Clearly as the filter scale is
+increased, due to the high \snr\ of the data, larger emission regions
+are (correctly) detected and reproduced in the map. Fourth (bottom)
+row: Jackknife maps for the bright extended reductions.}
 \label{fig:m17_jk}
@@
 M17 map in Fig.~\ref{fig:m17_jk} containing synthetic data, using the
-default configuration. left: raw PSDs for the input (noiseless)
+default configuration. Left: raw PSDs for the input (noiseless)
 simulated data (thick black line), the signal (average of each half)
@@
 scales: 150\,arcsec (red); 300\,arcsec (orange); 600\,arcsec (green);
-and 900\,arcsec (blue). Since the input PSD is known, it is possible
-to measure the transfer function of the map-maker as the ratio between
-the output map signal PSDs and the input PSD (from the left panel),
-giving the thin coloured lines (linear vertical axis shown on right of
-plot). The remaining lines are as in the left panel, but now corrected
-by the transfer function. This plot shows that increasing the filter
-scale improves the SNR at intermediates scales
-($\sim$200--600\,arcsec), although the SNR worsens at smaller scales
-($\lsim$200\,arcsec).}
+and 900\,arcsec (blue). PSDs are also colour-coded by filter
+scale. Right: since the input PSD is known, it is possible to measure
+the transfer function of the map-maker as the ratio between the output
+map signal PSDs and the input PSD (from the left panel), giving the
+thin coloured lines (linear vertical axis shown on right of plot). The
+remaining lines are as in the left panel, but now corrected by the
+transfer function. This plot shows that increasing the filter scale
+improves the \snr\ at intermediates scales ($\sim$200--600\,arcsec),
+although the \snr\ worsens at smaller scales ($\lsim$200\,arcsec).}
 \label{fig:m17_def_ps}
@@
 except now using the bright extended configuration. This configuration
-has resulted in an improved transfer function and SNR at large angular
-scales. Also note that increasing the filter scale does not have as
-large of an impact on the small-scale noise as in the default
+has resulted in an improved transfer function and \snr\ at large
+angular scales. Also note that increasing the filter scale does not
+have as large an impact on the small-scale noise as in the default
 configuration, although it is still noticeably larger up to scales of
-about $\sim$60\,arcsec.}
+about 60\,arcsec.}
 \label{fig:m17_be_ps}
@@
 While the reduction in the bottom panels of Fig.~\ref{fig:m17} is
-clearly superior in at least a cosmetic sense to those in the top
+clearly superior, in at least a cosmetic sense to those in the top
 panels, it is important to quantify both the noise properties of the
@@
 Maps are produced using the first and second continuous halves of the
-data in Fig.~\ref{fig:m17_jk}. This is not an ideal situation since
+data in Fig.~\ref{fig:m17_jk}. This is not an ideal situation, since
 the noise properties may evolve with time (e.g., due to changing sky
@@
 distribution in the complete map from the jackknife. Also, since the
-zero-masking depends on the SNR of the map, it will be restricted to
+zero-masking depends on the \snr\ of the map, it will be restricted to
 regions approximately $\sqrt{2}$ shallower in these maps than for the
@@
 is similar, though not identical in the two halves. Ideally one would
-have many full maps as in the case of the Lockman Hole data in the
+have many full maps, as in the case of the Lockman Hole data in the
 previous section, from which alternating maps could be combined.
@@
 map-maker to extended structures, we inject a simulated signal with
-power at a range of scales to a relatively empty region of the map.
+power at a range of scales into a relatively empty region of the map.
 It is created by drawing a realization of noise within an 800\,arcsec
@@
 14.5\,arcsec Gaussian to model the effect of the \scuba\ optical
-response. The RMS of these fluctuations are then normalized to 0.002
-pW so that they are comparable to the dynamic range of M17 itself. The
-minimum is subtracted so that the signal is positive. Finally, half a
-period of a (1+cosine)/2 function is used to roll-off the signal to
-zero between radii of 300 and 400 arcsec.
+response. The RMS of these fluctuations is then normalized to 0.002 pW
+so that they are comparable to the dynamic range of M17 itself. The
+minimum is then subtracted so that the signal is positive. Finally,
+half a period of a (1+cosine)/2 function is used to roll-off the
+signal to zero between radii of 300 and 400 arcsec.
 
@@
 rejection is used).  The synthetic data are clearly seen as the
-circular region S.E. of M17. As the filter scale is increased, the
+circular region south of M17. As the filter scale is increased, the
 size of the ripples increases accordingly. While larger structures do
@@
 jackknife maps. The astronomical emission is almost perfectly removed,
-except for a slight impression of M17 near the centre of the map that
+except for a slight impression of M17 near the centre of the map, that
 is probably due to some combination of detector gain and pointing
@@
 The third and fourth rows in Fig.~\ref{fig:m17_jk} repeat this
-exercise using the bright extended configuration, in which the SNR
+exercise using the bright extended configuration, in which the \snr\
 threshold of 5 is again used to constrain the map. As the filter scale
@@
 appearance). With filter scales $\gsim600$\,arcsec, virtually the
-entire region of synthetic emission is correctly identified by the SNR
+entire region of synthetic emission is correctly identified by the \snr\
 mask and allowed to vary freely in the solution.
 
-Next, we analyze the angular power spectra densities (PSDs) of the
+Next, we analyse the angular power spectra densities (PSDs) of the
 maps to understand the signal and noise properties of the map-maker in
@@
 encode the filter scales used: 150\,arcsec (red); 300\,arcsec
-(orange); 600\,arcsec (green); and 900\,arcsec (blue). Note that, in
-the exception of the synthetic data, we only plot the PSDs down two
-the second-lowest spatial frequency bin of
+(orange); 600\,arcsec (green); and 900\,arcsec (blue). Note that, with
+the exception of the synthetic data, we only plot the PSDs down to the
+second-lowest spatial frequency bin of
 $1.875\times10^{-3}$\,arcsec$^{-1}$, or 533\,arcsec, due to the fact
@@
 The transfer functions $G(k)$ are plotted as thin solid lines in the
-right panel of Fig.~\ref{fig:m17_def_ps}. This formula produced
-extremely noisy values at large frequencies (small scales), and was
-set to a value of 1 above 0.015\,arcsec$^{-1}$, as well as any point
-in the curve where $P_\mathrm{M}(k)$ exceeded $P_\mathrm{S}(k)$
-(i.e. $G(K)$ is assumed to be $\le$1). As expected, the larger the
-scale of the filter, the lower the spatial frequency at which the map
-transfer function falls. We then correct the map and jackknife PSDs by
-dividing by $G(k)$ to produce the thick solid, and dashed lines in the
-right panel of Fig.~\ref{fig:m17_def_ps}. This shows us that, even
-though the raw noise in the left panel is lower at small scales when a
-smaller-scale filter is used (e.g. the red dashed line), once we
-correct for the transfer function, we actually win in a SNR sense at
-large scales using the larger-scale filter as you would expect.
-
-These tests are then repeated using the bright extended reduction in
+right panel of Fig.~\ref{fig:m17_def_ps}. This formula produces
+extremely noisy values at large frequencies (small scales), and we
+therefore set it to a value of 1 above 0.015\,arcsec$^{-1}$, as well
+as any point in the curve where $P_\mathrm{M}(k)$ exceeds
+$P_\mathrm{S}(k)$ (i.e. $G(K)$ is assumed to be $\le$1). As expected,
+the larger the scale of the filter, the lower the spatial frequency at
+which the map transfer function falls. We then correct the map and
+jackknife PSDs by dividing by $G(k)$ to produce the thick solid, and
+dashed lines in the right panel of Fig.~\ref{fig:m17_def_ps}. This
+shows us that, even though the raw noise in the left panel is lower at
+small scales when a smaller-scale filter is used (e.g. the red dashed
+line), once we correct for the transfer function, we actually win in a
+\snr\ sense at large scales using the larger-scale filter, as would be
+expected.
+
+These tests are then repeated using the bright extended reduction, in
 Fig.~\ref{fig:m17_be_ps}. The most obvious improvement with this
@@
 fall more slowly at large angular scales, accompanied by a slower
-increase in noise; in other words, there is greatly improved SNR at
+increase in noise; in other words, there is greatly improved \snr\ at
 large angular scales (an obvious conclusion given the appearence of
 the maps in Fig.~\ref{fig:m17_jk}). In fact, using the 900\,arcsec
-filter edge, the map response is still above 80\% right out to the
-largest scale accurately measured in the PSDs, 533\,arcsec, which is
-about the largest scale that should be recoverable given the size of
-the array footprint and the fact that we use common-mode rejection.
+filter edge, the map response is still above 80 per centre right out to
+the largest scale accurately measured in the PSDs, 533\,arcsec, which
+is about the largest scale that should be recoverable, given the size
+of the array footprint and the fact that we use common-mode rejection.
 Another interesting feature of these reductions is that the increase
 in small-scale noise as the filter edge is increased is not as drastic
-as in the default reduction, and the SNR improves at scales
+as in the default reduction, and the \snr\ improves at scales
 $\gsim$60\,arcsec.
 
-One case in which the SNR is \emph{worse} using the bright extended
-reduction is using a 150\,arcsec filter. In this case the noise is
-considerably larger in the bright extended reduction as evidenced by
-the ``kink'' near 150\,arcsec. Referring to the mask contours in the
-left panel of the third row, it is clear that the map-maker has failed
-to identify much of the bright, extended emission in the region of the
-synthetic source. Each area that is not within the contours is
+One case in which the \snr\ is \emph{worse} using the bright extended
+reduction is when using a 150\,arcsec filter. In this case the noise
+is considerably larger in the bright extended reduction as evidenced
+by the ``kink'' near 150\,arcsec. Referring to the mask contours in
+the left panel of the third row, it is clear that the map-maker has
+failed to identify much of the bright, extended emission in the region
+of the synthetic source. Each area that is not within the contours is
 constrained to zero throughout the solution, therefore suppressing
 power (and lowering the transfer function), and subsequently reducing
-the SNR of the final result. This measurement serves as a warning: the
-map-maker response is non-linear when using SNR masking. Harsh
+the \snr\ of the final result. This measurement serves as a warning:
+the map-maker response is non-linear when using \snr\ masking. Harsh
 filtering can provide misleading results as in this example. Maps of
 \emph{faint} extended emission will also suffer considerably, as the
-structures of interest may not clear the SNR threshold for the mask.
+structures of interest may not clear the \snr\ threshold for the mask.
 
@@
 filtering step to suppress large-scales. As in the case of the
-SNR-based masking described here, Maximum-Entropy filtering is also
+\snr-based masking described here, Maximum-Entropy filtering is also
 non-linear, meaning that it too requires simulations to establish its
@@
 
+This paper has described the Submillimetre User Reduction Facility
+(SMURF), which was designed to produce maps from the rapidly sampled
+$\sim10^4$ bolometers of the \scuba\ instrument. We have presented
+sufficient detail in order to be useful not only to users of \scuba,
+but for similar detector arrays in the future. The basic elements are:
+
 \begin{enumerate}
-
-\item SMURF is an iterative map-maker that can produce \scuba\ maps on
-single high-end modern desktops in a reasonable amount of time.
 
@@
 of the map on large angular scales, due to the degeneracy between the
-map, common-mode, and other residual low-frequency noise. A simple
-strategy of constraining empty regions of the map to zero (either a
-user-supplied mask for known sources, or an iterative determination of
-signal above some SNR threshold) provides good constraints for compact
-objects, and bright/extended structures.
+map, common-mode, and other residual low-frequency noise
+contributions. A simple strategy of constraining empty regions of the
+map to zero (either a user-supplied mask for known sources, or an
+iterative determination of signal above some \snr\ threshold) provides
+good constraints for compact objects, and bright/extended structures.
 
@@
 it to adapt to a wide range of data in a pipeline setting.
+
+\item Principal Component Analysis (PCA) has been used to identify
+correlated noise contributions to \scuba\ data, and it is available as
+a generic data cleaning option.
+
+\item The memory requirements and execution times are reasonable
+(possible to reduce data in less time than it takes to collect) on
+modern high-end desktop computers due to the simplicity of the SMURF
+algorithm. It is used in real-time at the telescope for feedback
+during observations, and it is also regularly produces nearly
+science-grade products at the Canadian Astronomical Data Centre.
+
+\item SMURF is highly configurable, making it easy to vary the order
+and types of models that are fit to the data. It is also easy to add
+new models to the existing infrastructure.
 
@@
 
-\item The total of all the SNR values (the absolute residual difference
+\item The total of all the \snr\ values (the absolute residual difference
 divided by the local RMS of the residual differences) in the block must be
@@
 
-\item The total of all SNR values in the block must exceed half the value of
-the largest single SNR value in the block.
+\item The total of all \snr\ values in the block must exceed half the value of
+the largest single \snr\ value in the block.
 
@@
 The second and third tests are intended to prevent spikes being treated as
-two very close steps, since such steps will have SNR values of opposite
+two very close steps, since such steps will have \snr\ values of opposite
 signs and will thus cancel out.

### 2012-08-23 smurf_paper.tex
comments from Per Friberg
These were actually from early June I think. I screwed up a
merge at some point, and had to add these in again later. Some
points are not addressed yet, so I left "Per: bla bla" comments
in a few places as a reminder

@@
 rectangular subarrays, consisting of $40 \times 32$ bolometers each,
-and subtend a nearly contiguous 7\,arcmin $\times$ 7\,arcmin footprint
-on the sky.  The instrument is also equipped with a Fourier transform
-spectrometer \citep[FTS-2,][]{gom2010}, and a polarimeter
-\cite[POL-2,][]{2011ASPC..449...68B}. This paper describes the
-properties of \scuba\ data that are relevant to producing maps of the
-imaging data, and the Submillimetre User Reduction Facility, SMURF, a
-software package for perfoming the reduction written using the
-Starlink Software Environment \citep{2009ASPC..411..418J}. The
+and together subtend a nearly 7\,arcmin $\times$ 7\,arcmin footprint
+on the sky (excluding gaps, the continuous solid angle is about
+43\,arcmin$^{2}$ per focal plane).  The instrument is also equipped
+with a Fourier transform spectrometer \citep[FTS-2,][]{gom2010}, and a
+polarimeter \cite[POL-2,][]{2011ASPC..449...68B}. This paper describes
+the properties of \scuba\ data that are relevant to producing maps of
+the imaging data, and the Submillimetre User Reduction Facility,
+SMURF, a software package for perfoming the reduction written using
+the Starlink Software Environment \citep{2009ASPC..411..418J}. The
 analysis of FTS-2 and POL-2 data will be described at a later date
@@
 transition-edge sensors (TESs), with critical temperatures
-$T_\mathrm{C}$ (...typical value...). The range about $T_\mathrm{C}$
-in which the transition from a superconductor to a normal material
-occurs is extremely small (...typical value...?). If the devices are
-biased to a temperature within this transition on large time scales,
-then tiny temperature fluctuations on short time scales will produce
-significant changes in the impedence. The entire \scuba\ focal plane
-is therefore held at a \emph{base temperature} somewhat below
-$T_\mathrm{C}$. The TESs then have small electrical voltage biases,
-and individual pixel heaters (using a current bias to apply the
-desired Joule heating) are used to place the detectors in the
-transition. Only a small detector bias voltage is required since the
-heaters compensate for most of the variations in the base temperature,
-and other ambient loads (like the atmosphere).  Small temperature (and
-therefore impedence) fluctuations produce a changing current through
-the TESs. The resulting changing magnetic fields are then amplified
-using chains of superconducting quantum interference devices (SQUIDs),
-and the larger output current is finally digitized. Each detector has
-its own SQUID for the first-stage of the amplification, but the
-remaining stages occur within a common chain of SQUIDs for each column
-of 40 detectors. Each of the 32 rows is then read-out in sequence at a
-row-visit frequency of ...kHz?... This multiplexing scheme greatly
-reduces the number of readout and bias lines to each detector, making
-such large-format bolometer array designs feasible. It should also be
-noted that there is an additional 33rd row of SQUID readouts that are
-not connected to TES. These ``dark SQUIDs'' are used to track
-non-thermal noise sources that are common to each column's amplifier
-chain.
+$T_\mathrm{C}$ approximately 140 and 200\,mK at 850 and 450\,\micron,
+respectively. The range about $T_\mathrm{C}$ in which the transition
+from a superconductor to a normal material occurs is extremely small
+(...typical value...?). The entire \scuba\ focal plane is held at a
+\emph{base temperature} somewhat below $T_\mathrm{C}$. The TESs are
+then biased up into the transition using two systems. First,
+individual pixel heaters (using a current bias to apply the desired
+Joule heating) are set to fixed values at the start of an observation
+to compensate for most of the loading from the instrument and
+atmosphere (in better weather, less compensation from the heaters is
+required). Then, an electrical voltage bias is applied. As the
+instrument observes, and there are tiny temperature fluctuations due
+to astronomical sources and atmosphere, the impedence
+changes. Thermoelectrical feedback helps to keep the TESs ``locked''
+in the transition as this occurs; an increase in temperature increases
+the impedence, and lowers the current and resulting heating
+power. Changing magnetic fields resulting from these current changes
+are then amplified using chains of superconducting quantum
+interference devices (SQUIDs), and the larger output current is
+finally digitized. Each detector has its own SQUID for the first-stage
+of the amplification, but the remaining stages occur within a common
+chain of SQUIDs for each of the 32 columns. All 40 rows are then
+read-out in sequence at a row-visit frequency of about 12\,kHz. This
+multiplexing scheme greatly reduces the number of readout and bias
+lines to each detector, making such large-format bolometer array
+designs feasible. It should also be noted that there is an additional
+41st row of SQUID readouts that are not connected to TES. These ``dark
+SQUIDs'' are used to track non-thermal noise sources that are common
+to each column's amplifier chain.
 
@@
 the efficiency of the thermal absorber, how well the TES is biased
-into the transition, and the gain of the SQUID amplifiers. Since there
-are variations in $T_\mathrm{C}$ across each subarray due to
-manufacturing differences, considerable effort has gone into
-dynamically applying position-dependent biases as well as a total
-focal plane bias to make as many detectors usable as possible
-\citep{holland2012}. In practice, the response of a given detector can
-therefore vary somewhat from observation to observation, depending on
-how the bias and amplification parameters were set up. To track these
-responsivity changes, at the start of each observation, the current is
+into the transition, and the gain of the SQUID amplifiers. In
+practice, the response of a detector can therefore vary somewhat from
+observation to observation depending on how the arrays have been
+tuned, and the observing conditions \citep[see][]{holland2012}. To
+track these changes, at the start of each observation, the current is
 ramped through the bolometer heaters such that $dI/dP$ can be
-measured. Note that the bolometers may have a slightly different
-response to heater power due to conductance as compared to the
-response to optical power, so additioanl efficiency corrections must
-also be obtained. While not discussed further in this paper, details
-may be found in \citet{dempsey2012}. We will refer to measurements of
-$dI/dP$ across the focal plane as the
-\emph{flatfield}. Fig.~\ref{fig:sensitivities} shows the flatfield for
-each focal plane.
-
-%\begin{figure}
-%\vspace{2cm}
-%\textbf{This will be a figure showing s4a and s8d responsivity maps}
-%\vspace{2cm}
-%\caption{The response of each bolometer to known input heater power
-%  for a typical night during S2SRO. The s4a subarray had a
-%  particularly large variation in $T_\mathrm{C}$ towards the
-%  bottom-right which resulted in lower responsivity, and
-%  correspondingly larger noise. The s8d array also suffered
-%  significant variations across the focal plane, in particular
-%  removing larger number of useful detectors from the bottom-left.}
-%\label{fig:sensitivities}
-%\end{figure}
+measured. We will refer to the measurements of $dI/dP$ across the
+focal plane as the \emph{flatfield} for a given observation.
 
@@
 effectively white thermal noise produced by the phonon and optical
-loading on the absorbers; and (ii) and longer timescale drift
-resulting from slow variations in the mean load (e.g. fridge,
-telescope, or atmospheric changes), and also the readout
-electronics. As such, the spectrum corresponding to a bolometer
-time-series exhibits a ``$1/f$'' shape below some knee frequency,
-$f_\mathrm{knee}$, and is relatively flat at higher frequencies. Other
-smaller effects may include line features in the spectrum (a common
-culprit is pickup from AC mains, or microphonic signals induced by
-mechanical vibrations), and finally a roll-off towards the Nyquist
-frequency caused by the anti-aliasing filter that precedes
-digitization.
+loading on the absorbers (largely independent from bolometer to
+bolometer); and (ii) and longer timescale drift resulting from slow
+variations in the mean load (e.g. fridge, telescope, or atmospheric
+changes), and also the readout electronics (often correlated amongst
+many or all of the bolometers). As such, the spectrum corresponding to
+a bolometer time-series exhibits a ``$1/f$'' shape below some knee
+frequency, $f_\mathrm{knee}$, and is relatively flat at higher
+frequencies. Other smaller effects may include line features in the
+spectrum (a common culprit is pickup from AC mains, or microphonic
+signals induced by mechanical vibrations), and finally a roll-off
+towards the Nyquist frequency caused by the anti-aliasing filter that
+precedes digitization.
 
@@
 related to an astronomical flux density through the effective area of
-the detector, the spectral response of the bolometers (bandpass), the
-spectrum of the source, and additional optical efficiency factors -- a
-\emph{flux conversion factor} (FCF). Extinction is measured, in most
-cases, using the JCMT line-of-sight water vapour radiometer (WVM)
-which is calibrated using sky-dips and the CSO 225\,GHz tipper. The
-FCF is determined empirically using measurements of standard
-sources. For further details on these calibration steps, see
-\citet{dempsey2012}.
+the detector, the spectral response of the bolometers (with bandpasses
+set by optical filters), the spectrum of the source, and the optical
+efficiencies of the dish surface and other optical components in the
+system -- a \emph{flux conversion factor} (FCF). Extinction is
+measured, in most cases, using the JCMT line-of-sight water vapour
+radiometer (WVM) which is calibrated using sky dips and the CSO
+225\,GHz tipper. The FCF is determined empirically using measurements
+of standard sources. For further details on these calibration steps,
+see \citet{dempsey2012}.
 
@@
 ultimately limited to recovering information on angular scales smaller
-than the chop throw (a maximum of several arcmin). The primary
-approach to map-making with \scuba\ is instead to slew the entire
-telescope at such a rate that angular scales of interest are crossed
-on time scales shorter than $1/f_\mathrm{knee}$, similar to other
-ground-based instruments (MAMBO, BOLOCAM, SHARC-II, AzTEC,
-LABOCA,...). The desire to scan rapidly has two basic constraints: (i)
-the telescope cannot physically move faster that several hundred
-arcsec\,sec$^{-1}$ without requiring unacceptably large accelerations
-to change direction; and (ii) bolometers are designed with a thermal
-time constant that balances response with sensitivity. Taking these
-constraints into consideration, \scuba\ was ultimately designed to
-work at a maximum scan speed of 600\,arcsec\,sec$^{-1}$. The data
-acquisition system has a target sample rate of 200\,Hz, which at that
-scan velocity results in a sample every 3.0\,arcsec, giving
-approximately 2.5 samples per 450\,\micron\ diffraction-limited
-full-width at half-maximum (FWHM) -- a typical rule-of-thumb for
-adequately sampling a Gaussian point spread function (PSF). Similarly,
-the \scuba\ bolometers are designed to have a time constant of 2 ms
-\textbf{check that number!} in order to avoid distorting PSF-sized
-sources at this maximum scan rate.
+than the chop throw \citep[up to several arcmin,][]{holland1999}.  The
+primary approach to map-making with \scuba\ is instead to slew the
+entire telescope at such a rate that angular scales of interest are
+crossed on time scales shorter than $1/f_\mathrm{knee}$, as with other
+earlier ground-based instruments that did not use a chopping secondary
+\citep[e.g.,][]{glenn1998,weferling2002,wilson2008,kovacs2008b}. The
+desire to scan rapidly has two basic constraints: (i) while the
+telescope can physically move up to about 1800 and
+3600\,arcsec\,s$^{-1}$ in azimuth and elevation, respectively, the
+acceleration limit is about 600\,arcsec\,s$^{-2}$ (in azimuth, larger
+in elevation), meaning that significant time is waisted on
+``turn-arounds'' at large scan speeds; and (ii) bolometers are
+designed with a thermal time constant that balances response with
+sensitivity. Taking these constraints into consideration, \scuba\ was
+ultimately designed to operate at a maximum scan speed of
+600\,arcsec\,sec$^{-1}$. The data acquisition system has a target
+sample rate of 200\,Hz (although in practice it is typically in the
+range 170--180\,Hz), which at that scan velocity results in a sample
+every 3.0\,arcsec, or approximately one third of the 450\,\micron\
+diffraction-limited full-width at half-maximum (FWHM) -- a typical
+rule-of-thumb for adequately sampling a Gaussian point spread function
+(PSF). Similarly, the \scuba\ bolometers are designed to have a time
+constant of $\sim$1\,ms in order to avoid distorting PSF-sized sources
+at this maximum scan rate.
 
@@
 possible offer the best protection against large-scale noise
-\citep[e.g.][]{tegmark1997,kovacs2008b}; such strategies are said to
-offer good \emph{cross-linking}. As a tradeoff between truly random
-scan patterns and feasibility of implementation taking into
-consideration both mechanical and scheduling constraints, \scuba
-presently has three primary scan modes: PONG (cite Borys et al.),
-LISSAJOUS, and DAISY... (more citations? Kovacs et al. has a nice
-description... also CMB literature).
+\citep[see discussion in][and references therein]{kovacs2008b}; such
+strategies are said to offer good \emph{cross-linking}. As a tradeoff
+between truly random scan patterns and feasibility of implementation
+taking into consideration both mechanical and scheduling constraints,
+\scuba\ presently has three primary scan modes: PONG, which covers
+rectangular areas larger than the array footprint in a closed pattern
+that ``bounces'' off the edges at 45\,degree angles (and also usually
+combined with a rotation through a number of fixed position angles to
+create a ``rotating PONG'' with even better cross-linking); Lissajous,
+in which sinusoidal functions describe the motion along orthogonal
+axes (also to cover rectangular areas -- the PONG and Lissajous modes
+are both related to similar scan modes developed at the Caltech
+Submillimetre Observatory); and a ``constant velocity daisy'' for
+small areas (of order the array footprint), in which the telescope
+moves in a circle, whose centre also slowly traces out a circle. For a
+more complete description of the \scuba\ observing modes, see
+\citet{holland2012}.
 
@@
 \label{eq:psd}
-\langle\mathbf{b}^2(t)\rangle = \int \mathbf{P}(f) df .
+\langle\mathbf{b}^2(t)\rangle = 2 \int_0^{f_\mathrm{N}} \mathbf{P}(f)
+df .
 \end{equation}
 %
-However, we only show the PSD up to the Nyquist frequency, so the
-conversion from variance measured in the frequency domain to the time
-domain requires an additional multiplcation by 2; hence the units are
-pW$^2$\,Hz$^{-1}$ rather than pW$^2$\,s.
+where we only integrate over the positive frequencies up to
+$f_\mathrm{N}$, the Nyquist frequency, and the factor of 2 accounts
+for the (equal) power that appears at negative frequencies in the
+discrete Fourier Transform. The units of the PSD written in this form
+are then pW$^2$\,Hz$^{-1}$; without the factor of 2 in front the
+convention would be to write the units as pW$^2$\,s.
 
@@
   dominate on scales of approximately 8.3 and 1.7\,arcmin; these
-  scales are comparable to the array footprint.}
+  scales are comparable to the array footprint. .7\,arcmin; these
+  scales are comparable to the array footprint. \bf{Per:The array foot
+  print can't both be 1.7 and 8.3 arc in unless you used 4 arrays at
+  450 and only one at 850?}}
 \label{fig:pspec}
@@
 as expected if the bolometers are dominated by un-correlated white
-noise. Finally, we note that the amplitudes and slopes of
+noise. Finally, we note that the amplitudes and slopes of the
 $\mathbf{c}(t)$ are similar at both 450 and 850\,\micron. However,
@@
 noise that appears in the bolometer data (thermal) does not appear in
-the dark squid, leading to a significantly higher \snr.
+the dark squid, leading to a significantly higher \snr. Due to the
+fact that the sign of the gain in each stage of SQUID amplification is
+random (although the combined gain is constrained to be negative), and
+since magnetic field pickup is only seen at the input to the second
+stage, the pickup can appear with random signs for bolometers along a
+a column, giving it a distinct signature from other common signals
+that appear with the same sign.
 
@@
 platform, only azimuthal motion will change the effective direction of
-such a field with respect to the cryostat. It is also worth noting
-that the absolute azimuth of the telescope was close to 180$^\circ$
-(i.e., nearly due-south). We suspect that this observing
-configuration, combined with the large amplitude of the scan, may have
-conspired to produce these large signals.
+such a field with respect to the cryostat. Tests have shown that, as
+in this example, large scans in azimuth generically produce pickup. In
+contrast, and as expected, changes in elevation results in pickup that
+is approximately three orders-of-magnitude smaller.
 
@@
 of the 450 and 850\,\micron\ focal planes were populated with single
-subarrays, s4a and s8d, respectively. There were two main mechanical
-problems to contend with during this period. First, the fridge was not
-able to cool the focal plane to the desired base temperature of
-... which affected stability, particularly at 850\,\micron. Second,
-there were problems with the subarray fabrication process, which led
-to large numbers of dead pixels, as well as large-scale variations in
-$T_\mathrm{C}$ and the thermal conductivity, $G$. These latter issues
-made it difficult to simulataneously place all of the TESs into the
-optimal region of the superconducting transition, leading to
-significantly poorer sensitivities.
-
-While there were moderate variations from night-to-night during the
-observing period, there were typically 600 to 900 working detectors in
-each of the s4a and s8d subarrays. Fig.~\ref{fig:sensitivities} shows
-representative responsivity maps for the two subarrays from the middle
-of S2SRO, based on the response to ramping the bolometer heaters.
+subarrays, s4a and s8d respectively. There were three main problems to
+contend with during this period. First, the fridge was not able to
+cool the focal plane to the desired base temperature (partly due to
+the heat conductance of the wires from the arrays during this period)
+which affected stability, particularly at 850\,\micron. Second, a
+non-sinusoidal, large-scale oscillation in the fridge temperature
+could easily be seen as a common signal to all of the detectors with a
+period of approximately 25\,s.  Third, there were problems with the
+subarray fabrication process which led to large numbers of dead
+pixels, as well as large-scale variations in $T_\mathrm{C}$ and the
+thermal conductivity, $G$. These latter issues made it difficult to
+simulataneously place all of the TESs into the optimal region of the
+superconducting transition, leading to significantly poorer
+sensitivities. While there were moderate variations from
+night-to-night during the observing period, there were typically 600
+to 900 working detectors in each of the s4a and s8d subarrays.
+
+Since S2SRO the fridge performance has been greatly improved using NTD
+thermometry to servo the temperature to a constant value, and with new
+wiring adequate base temperatures can now be achieved. However, the
+variability across the focal plane continues to be a problem now that
+there is a full complement of subarrays (which still includes the two
+subarrays used during S2SRO). Optimal tuning to achieve the greatest
+overall array sensitivity is described in \citet{holland2012}.
 
@@
 for all pairs $(i,j)$ of the $N$ bolometer time series,
-$\langle\mathbf{b}_i(t),\mathbf{b}_j\rangle$; and (ii) a singular
+$\langle\mathbf{b}_i(t),\mathbf{b}_j(t)\rangle$; and (ii) a singular
 value decomposition is used to identify a new set of statistically
@@
   predominantly along the bottom-left edge of the 850\,\micron\
-  subarray where the bolometers are only tenuously locked. Finally,
+  subarray where the bolometers are probably poorly biased. Finally,
   the sixth component is degenerating into a less significant, and
@@
 450\,\micron\ and bottom-left at 850\,\micron). Bolometers in these
-regions of the subarray are probably only tenuously locked, and are
+regions of the subarray are probably not optimally biased, and are
 more sensitive to changes in the loading, possibly resulting in the
@@
 simplified map estimation. We do note, however, that recent advances
-in direct inversion methods such as SANPIEC \citep{patanchon2008} may
+in direct inversion methods such as SANEPIC \citep{patanchon2008} may
 in fact be tenable for \scuba\ data. A future goal of the software
@@
 
-We express the signal observed by the $i$th bolometer as a function of time,
+We express the signal observed by the $i$th bolometer as a function of
+ time, \textbf{Per: $e_i(t)$ is written as a vector ( bold ) but I assume
+ we use it as a scalar.}
+
 \begin{equation}
@@
   common-mode calculation), and each additive component is added back
-  into the time series and re-estimated in turn.}
+  into the time series and re-estimated in turn.  {\bf Per: The data
+  flow is not clear to me. One way would be to take the ast map and
+  remove extinction and gain correction and subtract it from the raw
+  data and then simply start the iteration again. That would refine
+  the com model from ast model pollution. But the text suggest that
+  the other models also are removed from the data. Indeed this would
+  be way different models can interact diverge. This process how and
+  when this is done is not clear and there must be a number of
+  subtleties affecting the subtraction of the models.}}
 \label{fig:dimm}
@@
 
-The \model{COM} model is the common-mode signal itself. It is a single time
-series that is estimated by finding the mean of all bolometers at each
-time step. Bolometer values that have been flagged as unusable for any
-reason are excluded from the mean.
+The \model{COM} model is the common-mode signal itself. It is a single
+time series that is estimated by finding the mean of all bolometers at
+each time step. Bolometer values that have been flagged as unusable
+for any reason are excluded from the mean. {\bf Per: has a comment in
+here about stability of the gain pattern... it should be constant
+right?}.
 
@@
 is beneficial for bolometers that follow the common-mode for most of
-the time but suffer from transient problems at specific times.
+the time but suffer from transient problems at specific times. {\bf
+Per: If magnetic pickup dominates the common mode a negative gain is
+normal and should not be removed. This might be a risk at high
+elevation in very good stable weather!}
 
@@
 
- We decided not to use the \emph{unbiased} estimator, since in practice
-it would require accumulating an additional array of values at every
-map pixel, and only results in a small difference where there are less
+ We decided not to use an \emph{unbiased} estimator (e.g., the
+extension of the common $(1/N-1)\sum_j (\mathbf{b}_j -
+\bar{\mathbf{b}})^2$ estimator using weights), since in practice it
+would require accumulating an additional array of values at every map
+pixel, and only results in a small difference where there are less
 than $\sim10$ samples per pixel (a situation that is almost never
@@
 expense of retaining information at the diffraction limit of the
-telescope.
+telescope.  \textbf{ Per: loss of information?}.
 
@@
 
-Sudden steps can occur in the time stream data from each bolometer.
-
-\emph{why?}
+Sudden steps can occur in the time stream data from each bolometer,
+with the most likely cause being wraparound in the first stage squid
+amplification, the details of which are given in \citet{holland2012}
+\textbf{Per gave a technical explanation, but this is probably best in
+another paper?}.
 
@@
   good agreement indicates that these maps are indeed dominated by
-  white noise, whose amplitude has been accurately estimated during
-  map-making. The lower panel shows the results for matched-filtered
+  white noise. The lower panel shows the results for matched-filtered
   signal and jackknife maps (Figs.~\ref{fig:lockman_maps}c,f). The

### 2012-08-23 smurf_paper.tex
missed some of Douglas' comments

@@
 Foundation for Innovation. EC thanks CANARIE/CANFAR for additional
-funding.
-
-\textit{Is there some kind of grant number that has to be referenced
-  for Ed's CANFAR funding?}
+funding. Also mention NSERC.
 
@@
 
-\item Locating each candidate step.
-
-\item Determining the widths of the candidate steps.
-
-\item Determine the height of the step. Steps can vary enormously in height.
-For instance the step in the left plot of Fig.~\ref{fig:steps1} is
-roughly 3 pW in height, whereas the step at sample 5100 in the right hand
-plot is only 0.01 pW in height. The height of a step is taken as the
-difference in the local median value just before and just after the step.
-
-\item Correcting all subsequent data values by the step height.
-
-\item Once all steps have been corrected, a global offset is added to the
+\item locating each candidate step;
+
+\item determining the widths of the candidate steps;
+
+\item determining the height of the step (steps can vary enormously in
+height -- for instance the step in the left-hand panel of
+Fig.~\ref{fig:steps1} is roughly 3\,pW in height, whereas the step at
+sample 5100 in the right-hand panel is only 0.01\,pW in height -- the
+height of a step is taken as the difference in the local median value
+just before and just after the step);
+
+\item correcting all subsequent data values by the step height;
+
+\item once all steps have been corrected, a global offset is added to the
 corrected bolometer time stream to ensure that the mean value before and
@@
 bolometer, the algorithm proceeds to the next stage. Candidate steps are
-found as follows:
+found as follows.
 
@@
 \includegraphics[width=\linewidth]{steps2.pdf}
-\caption{A section of the right hand bolometer from  Fig.~\ref{fig:steps1}
-with the median smoothed data overlayed in red.}
+\caption{A section of the bolometer signal from the right-hand panel
+of Fig.~\ref{fig:steps1} with the median smoothed data overlaid in
+red.}
 \label{fig:steps2}
@@
 \includegraphics[width=\linewidth]{steps3.pdf}
-\caption{Differences between adjacent values of the median smoothed data
-in  Fig.~\ref{fig:steps2}.}
+\caption{Differences between adjacent values of the median smoothed
+data shown in Fig.~\ref{fig:steps2}.}
 \label{fig:steps3}
@@
 holes in the array (up to 40 samples), but will not fill in larger holes.
-Such larger holes are filled in later by linear interpolation. The
+Such larger gaps are filled in later by linear interpolation. The
 resulting smoothed differences (shown in blue in Fig.~\ref{fig:steps4})
@@
 
-\item Blocks of contiguous samples for which the residual difference exceeds
-25 times the local RMS of the residual differences are found. Any two
-such blocks that are separated by fewer than 100 samples are merged into
-a single block. Blocks are ignored if they fail any of the following
-tests:
+\item Blocks of contiguous samples for which the residual difference
+exceeds 25 times the local RMS of the residual differences are
+found. Any two such blocks that are separated by fewer than 100
+samples are merged into a single block. Blocks are only kept if they
+pass all of the following tests:
 
@@
 
-\item The width of the block must be less than 200 samples.
-
-\item The total of all the \snr\ values (the absolute residual difference
+\item the width of the block must be less than 200 samples;
+
+\item the total of all the \snr\ values (the absolute residual difference
 divided by the local RMS of the residual differences) in the block must be
-at least 25.
-
-\item The total of all \snr\ values in the block must exceed half the value of
+at least 25;
+
+\item the total of all \snr\ values in the block must exceed half the value of
 the largest single \snr\ value in the block.
@@
 
-The data in Fig.~\ref{fig:steps5} results in two blocks being defined -
-one between samples 5037 and 5102 and another between samples 5286 and
-5296. Each of these blocks define the start and end of a candidate step.
-The first block includes two visible steps, but they have been merged
-into a single candidate step due to their small separation (less than 100
-samples).
+The data in Fig.~\ref{fig:steps5} results in two blocks being defined,
+one between samples 5037 and 5102, and another between samples 5286
+and 5296. Each of these blocks define the start and end of a candidate
+step.  The first block includes two visible steps, but they have been
+merged into a single candidate step due to their small separation
+(less than 100 samples).
 

### 2012-08-23 smurf_paper.tex
more work on the start - mostly the example bolometer data section

@@
 
-\title[SMURF: an iterative map-maker for \scuba]{The Sub-Millimetre User
-Reduction Facility: an iterative map-maker for \scuba}
+\title[\scuba: map-making with SMURF]{\scuba: iterative map-making
+with the Sub-Millimetre User Reduction Facility}
 
@@
 other simple filters), and then re-gridding these cleaned data into a
-map. Such was the basic recipe for the baseline reduction of chopped
-data from SCUBA-2's predecessor SCUBA
+map. Such was the basic recipe for the reduction of chopped data from
+SCUBA-2's predecessor SCUBA
 \citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO, another
@@
 estimation and removal of this signal significantly lowers the $1/f$
-knee, without compromosing structures on angular scales larger than
+knee, without compromosing structures on angular scales smaller than
 the array footprint. Residual independent drifts at lower frequencies
@@
 products using the JCMT Science Archive \citep{2011ASPC..442..203E}
-hosted by the Canadian Astronomical Data Centre (CADC). Details of
-the pipeline design are given in \citet{gibb2005} and
+hosted by the Canadian Astronomical Data Centre (CADC). Details of the
+pipeline design are given in \citet{gibb2005} and
 \citet{2008AN....329..295C}.
@@
 properties of \scuba\ data and scan patterns in Section~\ref{sec:data}
-to motivate the data model that we have adopted. Note that while most
-of the analysis described here applies equally to all of the data
-taken to-date with \scuba, the \scuba\ Shared-Risks Observing period
-(S2SRO) which took place from 2010 February 11 to 2010 March 25 used
-only a single subarray at each of the two observing wavelengths, and
-had considerably different noise properties and poorer performance
-than the fully-commissioned instrument in 2011 (a comparison is
-provided in Section~\ref{sec:s2sro}. Next, the details of the SMURF
-algorithm are given in Section~\ref{sec:algorithm}. The paper is
-concluded in Section~\ref{sec:examples} with three detailed test cases
-that span the majority of observation types likely to be undertaken
-with \scuba, with an emphasis on the mitigation of divergence
-problems, and characterizing the output maps: (i) Uranus, a bright,
-compact source (Section~\ref{sec:point}); (ii) the Lockman Hole, a
-blind survey of faint point-like source (Section~\ref{sec:cosmo}); and
-(iii) the star-forming region M17, including bright, extended emission
-(Section~\ref{sec:extended}). All of the data analyzed in this paper
-are publicly available through the CADC SCUBA-2 raw-data queries page
-for the dates and observation numbers given in the
+to motivate the data model that we have adopted. Note that most of the
+analysis described here applies equally to all of the data taken
+to-date with \scuba, including during the Shared-Risks Observing
+period (S2SRO) from 2010 February 11 to 2010 March 25 (during which
+only single subarrays were available at each of the two observing
+wavelengths). A comparison between the S2SRO and fully-commissioned
+instrument is given in Section~\ref{sec:s2sro}. Next, the details of
+the SMURF algorithm are given in Section~\ref{sec:algorithm}. The
+paper is concluded in Section~\ref{sec:examples} with three detailed
+test cases that span the majority of observation types likely to be
+undertaken with \scuba, with an emphasis on the mitigation of
+divergence problems, and characterizing the output maps: (i) Uranus, a
+bright, compact source (Section~\ref{sec:point}); (ii) the Lockman
+Hole, a blind survey of faint point-like source
+(Section~\ref{sec:cosmo}); and (iii) the star-forming region M17,
+including bright, extended emission (Section~\ref{sec:extended}). All
+of the data analyzed in this paper are publicly available through the
+CADC SCUBA-2 raw-data queries page for the dates and observation
+numbers given in the
 text\footnote{\texttt{http://www3.cadc-ccda.hia-iha.nrc-cnrc.gc.ca/
@@
 lines to each detector, making such large-format bolometer array
-designs feasible. It should also be noted that there is an additional
-41st row of SQUID readouts that are not connected to TES. These ``dark
-SQUIDs'' are used to track non-thermal noise sources that are common
-to each column's amplifier chain.
+designs feasible. Such a high sample rate is unnecessary to produce
+maps, so the data are re-sampled to approximately 200\,Hz.  It should
+also be noted that there is an additional 41st row of SQUID readouts
+that are not connected to TES. These ``dark SQUIDs'' are used to track
+non-thermal noise sources that are common to each column's amplifier
+chain.
 
@@
 loading on the absorbers (largely independent from bolometer to
-bolometer); and (ii) and longer timescale drift resulting from slow
+bolometer); and (ii) longer timescale drift resulting from slow
 variations in the mean load (e.g. fridge, telescope, or atmospheric
@@
 frequencies. Other smaller effects may include line features in the
-spectrum (a common culprit is pickup from AC mains, or microphonic
-signals induced by mechanical vibrations), and finally a roll-off
-towards the Nyquist frequency caused by the anti-aliasing filter that
-precedes digitization.
+spectrum (common culprits are aliased high-frequency noise detected
+during the multiplexed readout of the arrays as described in
+Section~\ref{sec:bolos}, pickup from AC mains, or microphonic signals
+induced by mechanical vibrations), and finally a roll-off towards the
+Nyquist frequency caused by the anti-aliasing filter that precedes
+digitisation.
 
@@
 %-------------------------------------------------
-\subsection{Typical bolometer signals}
-\label{sec:bolosignal}
-%-------------------------------------------------
-
-In Fig.~\ref{fig:bolos_mix} we show sample time series from single
-bolometers in each of the subarrays, as well as variations in the
-mixing chamber temperature (as a proxy for the focal-plane
-temperature), and the telescope pointing during a PONG scan.
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.49\linewidth]{bolos_point_mix_s2sro.pdf}
-\includegraphics[width=0.49\linewidth]{bolos_point_mix.pdf}
-\caption{A comparison between single bolometer time series in each
-  band (top two plots), with the mixing chamber temperature and
-  azimuth/elevation pointing offsets. There is a strong correlation
-  between the bolometers and the roughly $\sim$25\,s oscillation in
-  the fridge, but not with the telescope motion. Also note that the
-  total power in this primary signal is similar at both 450 and
-  850\,\micron, pointing to an internal rather than an external
-  source. The grey signals overplotted in the top-two panels show the
-  residual time-series after removing the common-mode signal. This
-  operation shows: (i) that most of the low-frequency signal is common
-  to all of the bolometers; and (ii) the non-correlated, and
-  predominantly white noise at 450\,\micron\ is significantly larger
-  than at 850\,\micron.}
-\label{fig:bolos_mix}
-\end{figure*}
-
-Both bolometers share significant long-timescale structure
-($\gsim10$\,s) that appears to be related to variations in the fridge
-base temperature, although the similarity is clearly greater at
-850\,\micron. In this particular case, the total power in the
-fluctuations at 450\,\micron, from $-0.04$\,pW to $+0.08$\,pW, are larger
-than the $-0.03$\,pW to $+0.05$\,pW fluctuations at 850\,\micron. Such
-behaviour might be expected if there is a comparable varying thermal
-load from the fridge at each wavelength, but a larger contribution of
-atmospheric variations through the 450\,\micron\ bandpass filters. We
-also note that there is no obvious strong correlation between the
-low-frequency signal structure, at either wavelength, with the
-telescope motion.
-
-The low-frequency signal component of the bolometer output is also
-highly correlated amongst bolometers in the same subarray. We have
-calculated a common-mode signal, $\mathbf{c}(t)$, as the average time
-series of all the working bolometers. We then fit the amplitude of
-$\mathbf{c}(t)$ at each wavelength to the signals shown in
-Fig.~\ref{fig:bolos_mix} and remove it, yielding the grey residual
-signals. These residuals are quite flat, although still with
-noticeable variations. The white noise is also apparent, and larger at
-450\,\micron\ as one would expect from the larger backgrounds compared
-to 850\,\micron.
-
-Next, we produce power spectral density (PSD) plots for several
-representative bolometers in Fig.~\ref{fig:pspec}. To produce this
-figure, we follow the convention that the PSD as a function of
-frequency, $\mathbf{P}(f)$, is normalized such that the integral over
-frequency gives the same variance as the time-series variance across
-the full time series. In other words, given a bolometer signal
-$\mathbf{b}(t)$,
-%
-\begin{equation}
-\label{eq:psd}
-\langle\mathbf{b}^2(t)\rangle = 2 \int_0^{f_\mathrm{N}} \mathbf{P}(f)
-df .
-\end{equation}
-%
-where we only integrate over the positive frequencies up to
-$f_\mathrm{N}$, the Nyquist frequency, and the factor of 2 accounts
-for the (equal) power that appears at negative frequencies in the
-discrete Fourier Transform. The units of the PSD written in this form
-are then pW$^2$\,Hz$^{-1}$; without the factor of 2 in front the
-convention would be to write the units as pW$^2$\,s.
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.49\linewidth]{pspec_s2sro.pdf}
-\includegraphics[width=0.49\linewidth]{pspec.pdf}
-\caption{Bolometer power spectral densities. Four representative
-  bolometers have been selected from each sub-array, and the
-  flat-fielded (but otherwise raw) PSDs are shown as coloured dotted
-  lines (the blue signals are for the same time series as those shown
-  in Fig.~\ref{fig:bolos_mix}). Horizontal dotted lines at $5 \times
-  10^{-8}$ and $5 \times 10^{-9}$\,pW\,Hz$^{-1}$ at 450 and
-  850\,\micron, respectively, give the approximate white-noise
-  levels. There are broad line features in the PSDs at both
-  wavelengths above $\sim$35\,Hz. From $\sim$50\,Hz up to the Nyquist
-  frequency, the gradual roll-off of the anti-aliasing filter is also
-  evident.  At lower frequencies, the bolometer signals exhibit clear
-  $1/f$ knees at approximately 2 and 4\,Hz at 450 and
-  850\,\micron. The solid black lines are the PSDs of the common-mode
-  signals at each wavelength, and the solid coloured lines show the
-  PSDs of the bolometers once the common-mode is removed. These
-  residual signals have significantly lower $1/f$ knees, approximately
-  0.4 and 2\,Hz at 450 and 850\,\micron. Finally, the dashed black
-  lines show the spectral shape produced by a point source, assuming
-  that the telescope is slewing at 200\,arcsec\,sec$^{-1}$ (typical
-  for S2SRO); for reference, the top horizontal axis shows the
-  conversion from frequency to angular scale. We can see that the line
-  features may add significant noise on the scale of the PSF. We also
-  see that, for this scan velocity, the low-frequency noise begins to
-  dominate on scales of approximately 8.3 and 1.7\,arcmin; these
-  scales are comparable to the array footprint. .7\,arcmin; these
-  scales are comparable to the array footprint. \bf{Per:The array foot
-  print can't both be 1.7 and 8.3 arc in unless you used 4 arrays at
-  450 and only one at 850?}}
-\label{fig:pspec}
-\end{figure*}
-
-The dotted coloured lines in Fig~\ref{fig:pspec} show the PSDs for
-raw, though flat-fielded data. At each wavelength, there is a clear
-$1/f$ knee at a few Hz, followed by a predominantly white spectrum
-punctuated by line features above $\gsim 30$\,Hz, and finally a
-roll-off caused by the anti-aliasing filter above $\gsim 70$\,Hz. As
-indicated in the previous section, the correlation between the
-low-frequency components of the different bolometer signals is
-large. The solid black lines in Fig.~\ref{fig:pspec} indicate the PSDs
-of the common-modes $\mathbf{c}(t)$ at each wavelength, which
-reproduce most of the low-frequency structure, as well as the
-high-frequency line features. The $\mathbf{c}(t)$ otherwise drop
-substantially below the individual bolometer PSDs at high-frequency,
-as expected if the bolometers are dominated by un-correlated white
-noise. Finally, we note that the amplitudes and slopes of the
-$\mathbf{c}(t)$ are similar at both 450 and 850\,\micron. However,
-since the 450\,\micron\ white noise is larger than at 850\,\micron\
-(shown approximately by the horizontal dotted lines), the $1/f$ knee
-occurs at a \emph{lower} frequency at 450\,\micron.
-
-Next, the common-mode signals are fit to each bolometer time series
-and removed as in Fig.~\ref{fig:bolos_mix}, and the resulting PSDs are
-shown as solid coloured lines. In this example, the residual signals
-have $1/f$ knees significantly lower than in the raw PSDs.
-
-For reference, the top horizontal axis has been converted to angular
-scale assuming a scan velocity of 200\,arcsec\,sec$^{-1}$ which was
-typical of the S2SRO observations. The power spectra of the PSFs in
-each band (arbitrary normalized to peak values of $10^{-6}$ and
-$10^{-7}$\,pW$^2$\,Hz$^{-1}$ at 450 and 850\,\micron\ respectively)
-are also showed as dashed lines for this assumed scan velocity showing
-that the smallest features resolvable by the telescope may be slightly
-affected by the excess noise in the line features. At the lower
-frequency end, the $1/f$ noise dominates at scales $\gsim 2$\,arcmin
-in the raw data and $\gsim 10$\,arcmin in the common-mode cleaned data
-at 450\,\micron, and at scales $\gsim 1$\,arscmin in the raw data and
-$\gsim 2$\,arcmin in the common-mode cleaned data at 850\,\micron.
-
-%-------------------------------------------------
-\subsection{Magnetic field pickup}
-\label{sec:magpickup}
-%-------------------------------------------------
-
-An additional noise source that appears to be significant in only a
-small subset of the S2SRO data is magnetic field pickup. Since the
-bolometer signals are ultimately detected through the amplification of
-magnetic fields, any additional changing fields within the cryostat
-will also be detected.
-
-An example of an observation where pickup appears to be significant is
-shown in Fig.~\ref{fig:magpickup}. The time-series for two
-450\,\micron\ bolometers in the same column (not flatfielded) show
-that there is a strong signal with a similar shape, but opposite
-signs. This behaviour is seen across the entire array. The dark squid
-signal for the same columns exhibits a similar shape and
-amplitude. Since the dark squid has no thermal absorber or TES
-attached to it, this observed signal is not likely to be optical or
-thermal in nature (although there can be some cross-talk with the
-bolometers). It is also for this reason that the higher-frequency
-noise that appears in the bolometer data (thermal) does not appear in
-the dark squid, leading to a significantly higher \snr. Due to the
-fact that the sign of the gain in each stage of SQUID amplification is
-random (although the combined gain is constrained to be negative), and
-since magnetic field pickup is only seen at the input to the second
-stage, the pickup can appear with random signs for bolometers along a
-a column, giving it a distinct signature from other common signals
-that appear with the same sign.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{magpickup.pdf}
-\caption{Evidence for significant magnetic field pickup for
-  observation 20100228\_00016.  The top panel shows two un-flatfielded
-  (but mean-subtracted) bolometer time-series from the same column,
-  with a 200 sample boxcar smooth (approximately 1\,s), illustrating
-  that they are dominated by a similar signal with opposite signs. The
-  second panel shows the dark squid signal for the column, also
-  mean-subtracted and with the same boxcar smooth. The bottom panel
-  shows the azimuthal and elevation offsets from the map centre. For
-  reference, the mean azimuth was 171.9$^\circ$ and the mean elevation
-  68.0$^\circ$. Only the azimuthal signal appears to be correlated
-  with the dark squids and bolometer signals, which suggests a
-  magnetic field stationary with respect to the telescope dome as the
-  source, since its direction with respect to the cryostat only
-  changes with azimuthal motion.}
-\label{fig:magpickup}
-\end{figure}
-
-The telescope pointing offsets for this approximately 0.5\,degree
-diameter scan are also shown in Fig.~\ref{fig:magpickup}. Since the
-phase of the azimuth offsets from the map centre in this scan pattern
-slowly drifts with respect to the elevation offsets, it is clear that
-the bolometer and dark squid signals are detecting a noise source that
-is correlated only with the azimuthal motion and \emph{not} the
-elevation. This behaviour would be expected if if there were a strong
-magnetic field fixed with respect to the telescope dome (i.e., the
-earth's magnetic field). Since \scuba\ is mounted on a Nasmyth
-platform, only azimuthal motion will change the effective direction of
-such a field with respect to the cryostat. Tests have shown that, as
-in this example, large scans in azimuth generically produce pickup. In
-contrast, and as expected, changes in elevation results in pickup that
-is approximately three orders-of-magnitude smaller.
-
-%-------------------------------------------------
-\subsection{Differences between the S2SRO and 2011 re-commissioned instrument}
+\subsection{Differences between the S2SRO and 2011 fully-commissioned instrument}
 \label{sec:s2sro}
@@
 thermometry to servo the temperature to a constant value, and with new
-wiring adequate base temperatures can now be achieved. However, the
-variability across the focal plane continues to be a problem now that
-there is a full complement of subarrays (which still includes the two
-subarrays used during S2SRO). Optimal tuning to achieve the greatest
-overall array sensitivity is described in \citet{holland2012}.
+wiring, adequate base temperatures can now be achieved. However, the
+gradient in detector properties across the focal plane continues to be
+a problem now that there is a full complement of subarrays (which
+still includes the two subarrays used during S2SRO). Optimal tuning to
+achieve the greatest overall array sensitivity is described in
+\citet{holland2012}.
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.49\linewidth]{bolos_point_mix_s2sro.pdf}
+\includegraphics[width=0.49\linewidth]{bolos_point_mix.pdf}
+\caption{A comparison between single bolometer time series in each of
+  the 450 and 850\,\micron\ bands with the mixing chamber temperature
+  and azimuth/elevation pointing offsets, before and after upgrading
+  the instrument. The grey signals overplotted in the top two panels
+  show the residual time-series after removing the common-mode
+  signals.  Left: data taken during the S2SRO period, observation 29
+  on 20100313. There is a strong correlation between the bolometers
+  and the roughly $\sim$25\,s oscillation in the fridge, but not with
+  the telescope motion. Also note that the total power in this primary
+  signal is similar at both 450 and 850\,\micron, pointing to an
+  internal rather than an external source.  The nearly flat
+  common-mode subtracted signals show: (i) that most of the
+  low-frequency signal is common to all of the bolometers; and (ii)
+  the non-correlated, and predominantly white noise at 450\,\micron\
+  is significantly larger than at 850\,\micron. Right: data taken with
+  the fully-commissioned instrument, observation 38 on
+  20111112. Unlike the S2SRO data, there is no strong signal produced
+  by variations in the fridge temperature. The 450\,\micron\ data have
+  minimal correlated low-frequency noise (as evidenced by the lack of
+  a difference between the raw and common-mode subtracted signals),
+  although the 850\,\micron data show a common signal correlated with
+  the telescope motion (note that this scan has a significantly larger
+  amplitude than the S2SRO data set).}
+\label{fig:bolos_mix}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.49\linewidth]{pspec_s2sro.pdf}
+\includegraphics[width=0.49\linewidth]{pspec.pdf}
+\caption{Bolometer power spectral densities for the same two data sets
+  (before and after upgrades) used in Fig.~\ref{fig:bolos_mix}. Four
+  representative bolometers have been selected at each wavelength, and
+  the flat-fielded (but otherwise raw) PSDs are shown as coloured
+  dotted lines (the blue signals are for the same time series as those
+  shown in Fig.~\ref{fig:bolos_mix}).  The solid black lines are the
+  PSDs of the common-mode signals at each wavelength, and the solid
+  coloured lines show the PSDs of the bolometers once the common-mode
+  is removed.  Finally, the dashed black lines show the spectral shape
+  produced by a point source given the scan speeds for the two
+  observations (120\,arcsec\,s$^{-1}$ in the left-hand plot, and
+  190\,arcsec\,s$^{-1}$ in the right-hand plot); for reference, the
+  top horizontal axes shows the conversion from frequency to angular
+  scale. Horizontal dotted lines at $2 \times 10^5$ and $5 \times
+  10^4$\,pW\,Hz$^{-1}$ at 450 and 850\,\micron, respectively, are
+  provided as a visual reference for the white-noise levels. In
+  addition to $1/f$ and white noise components, and line features, all
+  of the PSDs exhibit the gradual roll-off of the anti-aliasing filter
+  just below the Nyquist frequency.  Left: for the S2SRO data there
+  are broad line features in the PSDs at both wavelengths above
+  $\sim$35\,Hz.  At lower frequencies, the bolometer signals exhibit
+  clear $1/f$ knees at approximately 2 and 4\,Hz at 450 and
+  850\,\micron. Common-mode subtraction removes most of the correlated
+  fridge oscillation signal, lowering the $1/f$ knees to approximately
+  0.4 and 2\,Hz (or angular scales of 1.7 and 8,3\,arcmin) at 450 and
+  850\,\micron, respectively. Right: data from the fully-commissioned
+  instrument tend to have lower $1/f$ knees, and fewer line
+  features. The remaining low-frequency noise, now that the fridge
+  oscillations have been removed, is considerably more independent
+  from bolometer to bolometer than in the S2SRO data (note the larger
+  spread in the dotted lines at low frequencies compared to the
+  left-hand plot). Finally, this example shows how much the noise in
+  the arrays can vary over time. Despite selecting bolometers from the
+  same subarrays in each example, these particular 850\,\micron\ data
+  from the fully-commissioned instrument are somewhat noisier.}
+\label{fig:pspec}
+\end{figure*}
+
+%-------------------------------------------------
+\subsection{Typical bolometer signals}
+\label{sec:bolosignal}
+%-------------------------------------------------
+
+In Fig.~\ref{fig:bolos_mix} we show sample time series from single
+bolometers in each of the 450 and 850\,micron\ focal planes, as well
+as variations in the mixing chamber temperature (as a proxy for the
+focal-plane temperature), and the telescope pointing, for two data
+sets, before and after the upgrades that followed S2SRO.
+
+In the S2SRO data (left panel), observation 29 on 20100313, both
+bolometers share significant long-timescale structure ($\gsim10$\,s)
+that appears to be related to variations in the fridge base
+temperature, although the similarity is clearly greater at
+850\,\micron. In this particular case, the total power in the
+fluctuations at 450\,\micron, from $-0.04$\,pW to $+0.08$\,pW, are
+larger than the $-0.03$\,pW to $+0.05$\,pW fluctuations at
+850\,\micron. Such behaviour might be expected if there is a
+comparable varying thermal load from the fridge at each wavelength,
+but a larger contribution of atmospheric variations through the
+450\,\micron\ bandpass filters. We also note that there is no obvious
+strong correlation between the low-frequency signal structure, at
+either wavelength, with the telescope motion.
+
+The low-frequency signal component of the S2SRO bolometer output is
+also highly correlated amongst bolometers in the same subarray. We
+have calculated a common-mode signal, $\mathbf{c}(t)$, as the average
+time series of all the working bolometers. We then fit the amplitude
+of $\mathbf{c}(t)$ at each wavelength to the signals shown in
+Fig.~\ref{fig:bolos_mix} and remove it, yielding the grey residual
+signals. These residuals are quite flat, although still with
+noticeable variations. The white noise is also apparent, and larger at
+450\,\micron\ as one would expect from the larger backgrounds compared
+to 850\,\micron.
+
+Data from the fully commissioned instrument, observation 38 on
+20111112, are shown in the right panel of
+Fig.~\ref{fig:bolos_mix}. Having solved the fridge oscillation
+problem, these data no longer exhibit a correlation with the mixing
+chamber (note that variations are still seen in the mixing chamber; it
+is likely that these are not observed at the focal plane itself, where
+the NTD thermometry used in the servo loop is performed). There is
+minimal correlated low-frequency noise in these 450\,\micron\ data
+(note that there is almost no difference once the common-mode has been
+subtarcted). The 850\,\micron\ channel, however, exhibits a
+significant signal that is obviously correlated with the telescope
+motion. This signal is correlated amongs many of the detectors, and
+common-mode removal yields fairly flat residuals. Part of the reason
+that this is seen here, and not in the S2SRO data set shown, is that
+the amplitude of the scan pattern is somewhat larger. Most of this
+``scan-synchronous'' noise is attributed to magnetic field pickup, as
+described in Section~\ref{sec:magpickup}.
+
+Next, we produce power spectral density (PSD) plots for several
+representative bolometers from both focal planes, using the same two
+data sets, in Fig.~\ref{fig:pspec}. To produce this figure, we follow
+the convention that the PSD as a function of frequency,
+$\mathbf{P}(f)$, is normalized such that the integral over frequency
+gives the same variance as the time-series variance across the full
+time series. In other words, given a bolometer signal $\mathbf{b}(t)$,
+%
+\begin{equation}
+\label{eq:psd}
+\langle\mathbf{b}^2(t)\rangle = 2 \int_0^{f_\mathrm{N}} \mathbf{P}(f)
+df ,
+\end{equation}
+%
+where we only integrate over the positive frequencies up to
+$f_\mathrm{N}$, the Nyquist frequency, and the factor of 2 accounts
+for the (equal) power that appears at negative frequencies in the
+discrete Fourier Transform. The units of the PSD written in this form
+are pW$^2$\,Hz$^{-1}$; without the factor of 2 in front the convention
+would be to write the units as pW$^2$\,s to distinguish the two
+normalisations.
+
+The dotted coloured lines in Fig~\ref{fig:pspec} show the PSDs for
+raw, though flat-fielded data. At each wavelength, and in both data
+sets, there are clear $1/f$ knees at frequencies ranging from roughly
+1 to 10\,Hz, followed by a predominantly white spectrum punctuated by
+line features at higher frequencies, and finally a roll-off caused by
+the anti-aliasing filter above $\gsim 70$\,Hz. As indicated in the
+previous section, the correlation between the low-frequency components
+of the different bolometer signals is large. The solid black lines in
+Fig.~\ref{fig:pspec} indicate the PSDs of the common-modes
+$\mathbf{c}(t)$ at each wavelength, which reproduce much of the
+low-frequency structure, as well as some of the higher-frequency line
+features. The $\mathbf{c}(t)$ otherwise drop substantially below the
+individual bolometer PSDs at high-frequency, as expected if the
+bolometers are dominated by un-correlated white noise. The common-mode
+signals are fit to each bolometer time series and removed as in
+Fig.~\ref{fig:bolos_mix}, and the resulting PSDs are shown as solid
+coloured lines.  For reference, the top horizontal axes have been
+converted to angular scale assuming the scan speeds of each
+observation (120\,arcsec\,s$^{-1}$ in the left-hand plot, and
+190\,arcsec\,s$^{-1}$ in the right-hand plot). The power spectra of
+unresolved point-sources are also shown as dashed lines in each band
+(arbitrarily normalized), showing that the smallest features
+resolvable by the telescope are only minimally affected by the excess
+noise in the line features.
+
+In the S2SRO data (left-hand panel of Fig.~\ref{fig:pspec}), the
+low-frequency noise is very correlated between the detectors (note the
+tight scatter in the dotted colour lines, and their resemblance to the
+common-mode), due to it being dominated by the fridge
+oscillations. Common-mode subtraction is very effective, reducing the
+$1/f$ knees in both wavelengths by about a factor of 5.
+
+Raw data from the fully-commissioned instrument (right-hand panel of
+Fig.~\ref{fig:pspec}) generally have noticeably lower $1/f$ knees
+(with the exception of some of the noisier 850\,\micron\ bolometers in
+this example) as compared with the S2SRO data. The residual
+low-frequency noise is considerably less correlated amongst detectors
+(larger spread in the dotted colour lines), leading to a less drastic
+improvement upon common-mode removal. However, since the data are less
+dominated by low-frequency noise to begin with, the signals generally
+require less low-frequency filtering to produce maps, and therefore
+retain larger-scale structures than with the S2SRO data. Furthermore,
+there are generally fewer line features in the PSDs of bolometers in
+the fully-commissioned instrument. Finally, these two data sets
+illustrate how the white noise performance (NEFD) can vary. For this
+particular data set, the 850\,\micron\ noise performance is
+substantially worse than in the example from the S2SRO
+period. Generally speaking such noise steps occur in only a subset of
+the \scuba\ subarrays. However, given the 4-fold increase in available
+working bolometers since S2SRO, and the improved low-frequency noise
+performance, \scuba\ is now a significantly faster mapping instrument,
+and preserves information up to larger angular scales.
+
+%-------------------------------------------------
+\subsection{Magnetic field pickup}
+\label{sec:magpickup}
+%-------------------------------------------------
+
+An additional noise source that appears to be significant in only a
+small subset of the data (and moreso during the S2SRO period) is
+magnetic field pickup. Since the bolometer signals are ultimately
+detected through the amplification of magnetic fields, any additional
+changing fields within the cryostat will also be detected.
+
+An example of an observation where pickup appears to be significant is
+shown in Fig.~\ref{fig:magpickup}. The time-series for two
+450\,\micron\ bolometers in the same column (not flatfielded) show
+that there is a strong signal with a similar shape, but opposite
+signs. This behaviour is seen across the entire array. The dark squid
+signal for the same columns exhibits a similar shape and
+amplitude. Since the dark squid has no thermal absorber or TES
+attached to it, this observed signal is not likely to be optical or
+thermal in nature (although there can be some cross-talk with the
+bolometers). It is also for this reason that the higher-frequency
+noise that appears in the bolometer data (thermal) does not appear in
+the dark squid, leading to a significantly higher \snr. Due to the
+fact that the sign of the gain in each stage of SQUID amplification is
+random (although the combined gain is constrained to be negative), and
+since magnetic field pickup is only seen at the input to the second
+stage, the pickup can appear with random signs for bolometers along a
+a column, giving it a distinct signature from other common signals
+that appear with the same sign.
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{magpickup.pdf}
+\caption{Evidence for significant magnetic field pickup for
+  observation 20100228\_00016.  The top panel shows two un-flatfielded
+  (but mean-subtracted) bolometer time-series from the same column,
+  with a 200 sample boxcar smooth (approximately 1\,s), illustrating
+  that they are dominated by a similar signal with opposite signs. The
+  second panel shows the dark squid signal for the column, also
+  mean-subtracted and with the same boxcar smooth. The bottom panel
+  shows the azimuthal and elevation offsets from the map centre. For
+  reference, the mean azimuth was 171.9$^\circ$ and the mean elevation
+  68.0$^\circ$. Only the azimuthal signal appears to be correlated
+  with the dark squids and bolometer signals, which suggests a
+  magnetic field stationary with respect to the telescope dome as the
+  source, since its direction with respect to the cryostat only
+  changes with azimuthal motion.}
+\label{fig:magpickup}
+\end{figure}
+
+The telescope pointing offsets for this approximately 0.5\,degree
+diameter scan are also shown in Fig.~\ref{fig:magpickup}. Since the
+phase of the azimuth offsets from the map centre in this scan pattern
+slowly drifts with respect to the elevation offsets, it is clear that
+the bolometer and dark squid signals are detecting a noise source that
+is correlated only with the azimuthal motion and \emph{not} the
+elevation. This behaviour would be expected if if there were a strong
+magnetic field fixed with respect to the telescope dome (i.e., the
+earth's magnetic field). Since \scuba\ is mounted on a Nasmyth
+platform, only azimuthal motion will change the effective direction of
+such a field with respect to the cryostat. Tests have shown that, as
+in this example, large scans in azimuth generically produce pickup. In
+contrast, and as expected, changes in elevation results in pickup that
+is approximately three orders-of-magnitude smaller.
 
@@
   components. The third component accurately reproduces an apparent
-  data-acquisition glitch that appears only at 450\,\micron\ (see
-  Section ???). The fourth component is a ...\,Hz line feature only
-  observed in column 15 of the 450\,\micron\ subarray. The fifth
-  component shows a series of steps and glitches that appear
-  predominantly along the bottom-left edge of the 850\,\micron\
-  subarray where the bolometers are probably poorly biased. Finally,
-  the sixth component is degenerating into a less significant, and
-  more randomly scattered signal across the focal plane. Subsequent
-  components follow this trend, and would generally be kept in a
-  PCA-cleaning procedure.}
+  data-acquisition glitch that appears only at 450\,\micron. The
+  fourth component is a periodic feature only observed in column 15 of
+  the 450\,\micron\ subarray. The fifth component shows a series of
+  steps and glitches that appear predominantly along the bottom-left
+  edge of the 850\,\micron\ subarray where the bolometers are poorly
+  biased. Finally, the sixth component is degenerating into a less
+  significant, and more randomly scattered signal across the focal
+  plane. Subsequent components follow this trend, and would generally
+  be kept in a PCA-cleaning procedure.}
 \label{fig:pca}
@@
 combinations of $\mathbf{F}(t)$ and $\mathbf{A}(t)$ that are
-statistically uncorrelated. Depending on the particular realizations
+statistically uncorrelated. Depending on the particular realisations
 of these two noise sources, this decomposition may, or may not give
@@
 evidenced by the large random scatter in the s4a eigenvalue maps. It
-is thought that these features are some sort of data acquisition
-glitch, as the spikes (when they appear) exhibit the expected point
-response function of the anti-aliasing filter. Similarly, Component 4
-also appears to be a data acquisition glitch in the s4a array,
-although along a single column, and with a peak at a lower frequency
-of 40\,Hz, together with a second feature right at the Nyquist
-frequency.
+is thought that these features are some sort of aliased high-frequency
+data acquisition glitch, as the spikes (when they appear) exhibit the
+expected point response function of the anti-aliasing
+filter. Similarly, Component 4 also appears to be a data acquisition
+glitch in the s4a array, although along a single column, and with a
+peak at a lower frequency of 40\,Hz, together with a second feature
+right at the Nyquist frequency.
 
@@
 the start and finish), or residual spikes (for which the median
-filtering is particularly useful).
+filtering is particularly useful). However, the gap-filling and
+continuity algorithms that we have employed (Section~\ref{sec:gaps})
+successfully mitigate these problems, and the \model{FLT} model is
+substantially faster,
 
@@
 higher-order polynomials (see their Fig.~10, as well as the discussion
-and references to earlier work in Section~4.3).
+and references to earlier work in their Section~4.3).
 
@@
 total data taken as part of the project; the remaining observations
-were dropped due to problems keeping the arrays properly locked during
+were dropped due to problems keeping the arrays properly tuned during
 S2SRO, and were easily identified by their highly variable and erratic
@@
 filtering step to suppress large-scales. As in the case of the
-\snr-based masking described here, Maximum-Entropy filtering is also
+\snr-based masking described here, Maximum-Entropy filtering is
 non-linear, meaning that it too requires simulations to establish its

### 2012-08-23 smurf_paper.tex
More bibliography fixes
Mainly to get SPIE and ASP papers into MN style
Add missing reference to iterative ORAC-DR SCUBA recipe

@@
 % journals
-\newcommand{\ao}{\rm AO}
-\newcommand{\apj}{\rm ApJ}
-\newcommand{\apjl}{\rm ApJL}
-\newcommand{\apjs}{\rm ApJS}
-\newcommand{\aaps}{\rm A$\&$AS}
-\newcommand{\aap}{\rm A$\&$A}
-\newcommand{\aapr}{\rm A$\&$AR}
-\newcommand{\mnras}{\rm MNRAS}
-\newcommand{\aj}{\rm Astron. J.}
-\newcommand{\araa}{\rm ARAA}
-\newcommand{\nat}{\rm Nature}
-\newcommand{\pasj}{\rm PASJ}
-\newcommand{\pasp}{\rm Publ. Astron. Soc. Pac.}
-\newcommand{\prd}{\rm PhRvD}
-\newcommand{\ASP}{\rm ASP Conference Series}
+\newcommand{\ao}{AO}
+\newcommand{\apj}{ApJ}
+\newcommand{\apjl}{ApJL}
+\newcommand{\apjs}{ApJS}
+\newcommand{\aaps}{A$\&$AS}
+\newcommand{\aap}{A$\&$A}
+\newcommand{\aapr}{A$\&$AR}
+\newcommand{\mnras}{MNRAS}
+\newcommand{\aj}{A J}
+\newcommand{\araa}{\rm ARA\&A}
+\newcommand{\nat}{Nat}
+\newcommand{\pasj}{PASJ}
+\newcommand{\pasp}{PASP}
+\newcommand{\prd}{PhRvD}
+\newcommand{\ASP}{ASP Conf.\ Ser.~}
 \newcommand{\CASP}{\rm Comm. Astrophys. Space Phys.}
 \newcommand{\astroph}{\rm astro-ph/}
+\newcommand{\spie}{Proc.\ SPIE~}
 
@@
 additional noise components. More similar to the algorithm described
-here was the iterative method for fitting and removing baseline drifts
+here was the iterative pipeline recipe for fitting and removing baseline drifts
 in SCUBA scan-map data as the astronomical image estimate improved
-(...SURF reference?...). The two closest modern relatives of SMURF are the
+\citet{1999ASPC..172..171J}. The two closest modern relatives of SMURF are the
 Comprehensive Reduction Utility for SHARC-2

### 2012-08-24 smurf_paper.tex
minor fiddle in intro

@@
 variations (Section~\ref{sec:bolos}), describe \scuba\ scan strategies
-(Section~\ref{sec:scan}), typical bolometers signals
-(Section~\ref{sec:bolosignal}), and the impact of magnetic field
-pickup (Section~\ref{sec:magpickup}, describe the differences between
-SCUBA-2 Shared Risk Observing data and those of the the fully
-commissioned instrument (Section~\ref{sec:s2sro}), and finally
-illustrate the use of principal component analysis
-(Section~\ref{sec:pca}).
+(Section~\ref{sec:scan}), describe the differences between SCUBA-2
+Shared Risk Observing data and those of the the fully commissioned
+instrument (Section~\ref{sec:s2sro}), typical bolometers signals
+(Section~\ref{sec:bolosignal}), the impact of magnetic field pickup
+(Section~\ref{sec:magpickup}, and finally illustrate the use of
+principal component analysis (Section~\ref{sec:pca}).
 

### 2012-08-24 smurf_paper.tex
affiliations should really be the ones that paid for SCUBA-2 work

@@
   \parbox[t]{\textwidth}{
-    Edward~L.~Chapin$^{1,2,3}$\thanks{E-mail:~echapin@sciops.esa.int},
-    David~S.~Berry$^{3}$,
-    Andrew~G.~Gibb$^{2}$,
-    Tim~Jenness$^{3}$,
-    Douglas~Scott$^{2}$,
-    Frossie~Economou$^3$\thanks{Present address: National Optical
+    Edward~L.~Chapin$^{1,2}$\thanks{E-mail:~echapin@sciops.esa.int,
+      Present address: XMM SOC, ESAC, Apartado 78, 28691 Villanueva de
+      la Ca\~nada, Madrid, Spain},
+    David~S.~Berry$^{2}$,
+    Andrew~G.~Gibb$^{1}$,
+    Tim~Jenness$^{2}$,
+    Douglas~Scott$^{1}$,
+    Frossie~Economou$^2$\thanks{Present address: National Optical
       Astronomy Observatory, 950 N.\ Cherry Avenue, Tucson, AZ 85719, USA}
@@
   \\
-  $^{1}$XMM SOC, ESAC, Apartado 78, 28691 Villanueva de la Ca\~nada, Madrid,
-  Spain\\
-  $^{2}$Dept. of Physics \& Astronomy, University of British Columbia,
+  $^{1}$Dept. of Physics \& Astronomy, University of British Columbia,
   6224 Agricultural Road, Vancouver, B.C. V6T 1Z1, Canada\\
-  $^{3}$JointAstronomy Centre, 660 N. A`oh\={o}k\={u} Place, University
+  $^{2}$JointAstronomy Centre, 660 N. A`oh\={o}k\={u} Place, University
   Park, Hilo, Hawaii 96720, USA}

### 2012-08-27 smurf_paper.tex
update typical bolometer plots + section again
The s8b data post-upgrades is noisier than the old s8d,
but I worked harder at my bolometer selection to illustrate
the features that I was interested in.

@@
 \begin{figure*}
-\centering
 \includegraphics[width=0.49\linewidth]{bolos_point_mix_s2sro.pdf}
@@
   on 20100313. There is a strong correlation between the bolometers
-  and the roughly $\sim$25\,s oscillation in the fridge, but not with
-  the telescope motion. Also note that the total power in this primary
-  signal is similar at both 450 and 850\,\micron, pointing to an
-  internal rather than an external source.  The nearly flat
+  and the roughly $\sim$25\,s oscillation in the fridge, but only a
+  minor correlation with the telescope motion. The nearly flat
   common-mode subtracted signals show: (i) that most of the
@@
   a difference between the raw and common-mode subtracted signals),
-  although the 850\,\micron data show a common signal correlated with
+  although the 850\,\micron\ data show a common signal correlated with
   the telescope motion (note that this scan has a significantly larger
-  amplitude than the S2SRO data set).}
+  amplitude than the S2SRO data set), mostly likely caused by magnetic
+  field pickup (see Section~\ref{sec:magpickup} and
+  Fig.~\ref{fig:magpickup}).}
 \label{fig:bolos_mix}
@@
   (before and after upgrades) used in Fig.~\ref{fig:bolos_mix}. Four
-  representative bolometers have been selected at each wavelength, and
-  the flat-fielded (but otherwise raw) PSDs are shown as coloured
-  dotted lines (the blue signals are for the same time series as those
-  shown in Fig.~\ref{fig:bolos_mix}).  The solid black lines are the
-  PSDs of the common-mode signals at each wavelength, and the solid
-  coloured lines show the PSDs of the bolometers once the common-mode
-  is removed.  Finally, the dashed black lines show the spectral shape
-  produced by a point source given the scan speeds for the two
-  observations (120\,arcsec\,s$^{-1}$ in the left-hand plot, and
-  190\,arcsec\,s$^{-1}$ in the right-hand plot); for reference, the
-  top horizontal axes shows the conversion from frequency to angular
-  scale. Horizontal dotted lines at $2 \times 10^5$ and $5 \times
-  10^4$\,pW\,Hz$^{-1}$ at 450 and 850\,\micron, respectively, are
-  provided as a visual reference for the white-noise levels. In
-  addition to $1/f$ and white noise components, and line features, all
-  of the PSDs exhibit the gradual roll-off of the anti-aliasing filter
-  just below the Nyquist frequency.  Left: for the S2SRO data there
-  are broad line features in the PSDs at both wavelengths above
-  $\sim$35\,Hz.  At lower frequencies, the bolometer signals exhibit
-  clear $1/f$ knees at approximately 2 and 4\,Hz at 450 and
-  850\,\micron. Common-mode subtraction removes most of the correlated
-  fridge oscillation signal, lowering the $1/f$ knees to approximately
-  0.4 and 2\,Hz (or angular scales of 1.7 and 8,3\,arcmin) at 450 and
-  850\,\micron, respectively. Right: data from the fully-commissioned
-  instrument tend to have lower $1/f$ knees, and fewer line
-  features. The remaining low-frequency noise, now that the fridge
-  oscillations have been removed, is considerably more independent
-  from bolometer to bolometer than in the S2SRO data (note the larger
-  spread in the dotted lines at low frequencies compared to the
-  left-hand plot). Finally, this example shows how much the noise in
-  the arrays can vary over time. Despite selecting bolometers from the
-  same subarrays in each example, these particular 850\,\micron\ data
-  from the fully-commissioned instrument are somewhat noisier.}
+  of the most sensitive bolometers have been selected at each
+  wavelength, and the flat-fielded (but otherwise raw) PSDs are shown
+  as coloured dotted lines (the blue signals are for the same time
+  series as those shown in Fig.~\ref{fig:bolos_mix}).  The solid black
+  lines are the PSDs of the common-mode signals at each wavelength,
+  and the solid coloured lines show the PSDs of the bolometers once
+  the common-mode is removed.  Finally, the dashed black lines show
+  the spectral shape produced by a point source given the scan speeds
+  for the two observations (120\,arcsec\,s$^{-1}$ in the left-hand
+  plot, and 190\,arcsec\,s$^{-1}$ in the right-hand plot); for
+  reference, the top horizontal axes shows the conversion from
+  frequency to angular scale. Horizontal dotted lines at $2 \times
+  10^5$ and $1 \times 10^5$\,pW\,Hz$^{-1}$ at 450 and 850\,\micron,
+  respectively, are provided as a visual reference for the white-noise
+  levels. In addition to $1/f$ and white noise components, and line
+  features, all of the PSDs exhibit the gradual roll-off of the
+  anti-aliasing filter just below the Nyquist frequency.  Left: for
+  the S2SRO data there are broad line features in the PSDs at both
+  wavelengths above $\sim$35\,Hz.  At lower frequencies, the bolometer
+  signals exhibit clear $1/f$ knees at approximately 1 and 2\,Hz at
+  450 and 850\,\micron. Common-mode subtraction removes most of the
+  correlated fridge oscillation signal, lowering the $1/f$ knees to
+  approximately 0.2 and 0.7\,Hz at 450 and 850\,\micron,
+  respectively. Right: data from the fully-commissioned instrument
+  tend to have lower $1/f$ knees, and fewer line features. The
+  remaining low-frequency noise, now that the fridge oscillations have
+  been removed, is considerably more independent from bolometer to
+  bolometer than in the S2SRO data (note the larger spread in the
+  dotted lines at low frequencies compared to the left-hand plot,
+  particularly at 450\,\micron). Finally, this example shows that the
+  white noise performance is in factor similar for the two subarrays
+  (s4a and s8b) that were used both before and after the upgrades.}
 \label{fig:pspec}
@@
 In Fig.~\ref{fig:bolos_mix} we show sample time series from single
-bolometers in each of the 450 and 850\,micron\ focal planes, as well
+bolometers in each of the 450 and 850\,\micron\ focal planes, as well
 as variations in the mixing chamber temperature (as a proxy for the
@@
 850\,\micron. In this particular case, the total power in the
-fluctuations at 450\,\micron, from $-0.04$\,pW to $+0.08$\,pW, are
-larger than the $-0.03$\,pW to $+0.05$\,pW fluctuations at
+fluctuations at 450\,\micron, are comparable to those at
 850\,\micron. Such behaviour might be expected if there is a
-comparable varying thermal load from the fridge at each wavelength,
-but a larger contribution of atmospheric variations through the
-450\,\micron\ bandpass filters. We also note that there is no obvious
-strong correlation between the low-frequency signal structure, at
-either wavelength, with the telescope motion.
+comparable varying thermal load from the fridge at each wavelength
+that dominates. We also note that there is no obvious strong
+correlation between the low-frequency signal structure, at either
+wavelength, with the telescope motion.
 
@@
 (note that there is almost no difference once the common-mode has been
-subtarcted). The 850\,\micron\ channel, however, exhibits a
+subtarcted). The 850\,\micron\ channel, however, exhibits a more
 significant signal that is obviously correlated with the telescope
-motion. This signal is correlated amongs many of the detectors, and
-common-mode removal yields fairly flat residuals. Part of the reason
+motion. It is also correlated amongs many of the detectors, and
+common-mode removal corrects it to some extent. Part of the reason
 that this is seen here, and not in the S2SRO data set shown, is that
-the amplitude of the scan pattern is somewhat larger. Most of this
+the amplitude of the scan pattern is larger. Most of this
 ``scan-synchronous'' noise is attributed to magnetic field pickup, as
@@
 
-Next, we produce power spectral density (PSD) plots for several
-representative bolometers from both focal planes, using the same two
+Next, we produce power spectral density (PSD) plots for four of the
+most sensitive bolometers from both focal planes, using the same two
 data sets, in Fig.~\ref{fig:pspec}. To produce this figure, we follow
@@
 sets, there are clear $1/f$ knees at frequencies ranging from roughly
-1 to 10\,Hz, followed by a predominantly white spectrum punctuated by
+1 to 2\,Hz, followed by a predominantly white spectrum punctuated by
 line features at higher frequencies, and finally a roll-off caused by
@@
 individual bolometer PSDs at high-frequency, as expected if the
-bolometers are dominated by un-correlated white noise. The common-mode
-signals are fit to each bolometer time series and removed as in
-Fig.~\ref{fig:bolos_mix}, and the resulting PSDs are shown as solid
-coloured lines.  For reference, the top horizontal axes have been
-converted to angular scale assuming the scan speeds of each
-observation (120\,arcsec\,s$^{-1}$ in the left-hand plot, and
-190\,arcsec\,s$^{-1}$ in the right-hand plot). The power spectra of
-unresolved point-sources are also shown as dashed lines in each band
-(arbitrarily normalized), showing that the smallest features
+bolometers are dominated by un-correlated white noise in that part of
+the spectrum. The common-mode signals are fit to each bolometer time
+series and removed as in Fig.~\ref{fig:bolos_mix}, and the resulting
+PSDs are shown as solid coloured lines.  For reference, the top
+horizontal axes have been converted to angular scale assuming the scan
+speeds of each observation (120\,arcsec\,s$^{-1}$ in the left-hand
+plot, and 190\,arcsec\,s$^{-1}$ in the right-hand plot). The power
+spectra of unresolved point-sources are also shown as dashed lines in
+each band (arbitrarily normalized), showing that the smallest features
 resolvable by the telescope are only minimally affected by the excess
-noise in the line features.
+noise in the line features at these scan speeds (this may not be the
+case at higher scan speeds).
 
@@
 Raw data from the fully-commissioned instrument (right-hand panel of
-Fig.~\ref{fig:pspec}) generally have noticeably lower $1/f$ knees
-(with the exception of some of the noisier 850\,\micron\ bolometers in
-this example) as compared with the S2SRO data. The residual
-low-frequency noise is considerably less correlated amongst detectors
-(larger spread in the dotted colour lines), leading to a less drastic
-improvement upon common-mode removal. However, since the data are less
-dominated by low-frequency noise to begin with, the signals generally
-require less low-frequency filtering to produce maps, and therefore
-retain larger-scale structures than with the S2SRO data. Furthermore,
-there are generally fewer line features in the PSDs of bolometers in
-the fully-commissioned instrument. Finally, these two data sets
-illustrate how the white noise performance (NEFD) can vary. For this
-particular data set, the 850\,\micron\ noise performance is
-substantially worse than in the example from the S2SRO
-period. Generally speaking such noise steps occur in only a subset of
-the \scuba\ subarrays. However, given the 4-fold increase in available
-working bolometers since S2SRO, and the improved low-frequency noise
-performance, \scuba\ is now a significantly faster mapping instrument,
-and preserves information up to larger angular scales.
+Fig.~\ref{fig:pspec}) generally have less significant $1/f$ noise as
+compared with the S2SRO data. The residual low-frequency noise is
+considerably less correlated amongst detectors (larger spread in the
+dotted colour lines, particularly noticeable in these 450\,\micron\
+data), leading to a less drastic improvement upon common-mode
+removal. However, since the data are less dominated by low-frequency
+noise to begin with, the signals generally require less low-frequency
+filtering to produce maps, and therefore retain larger-scale
+structures than with the S2SRO data. Furthermore, there are generally
+fewer line features in the PSDs of bolometers in the
+fully-commissioned instrument. Finally, these two data sets illustrate
+that the white noise performance (NEFD) is in fact similar before and
+the upgrades for these two subarrays (s4a at 450\,micron, and s8b at
+850\,\micron). The main improvements are reduction in correlated and
+line noise features mentioned above, and the larger number of working
+bolometers (approximately a factor of 4).
 

### 2012-08-31 smurf_paper.tex
mostly PCA section updates

@@
 
-\title[\scuba: map-making with SMURF]{\scuba: iterative map-making
+\title[\scuba: iterative map-making with SMURF]{\scuba: iterative map-making
 with the Sub-Millimetre User Reduction Facility}
@@
   scanned across the sky and records data at a rate of 200\,Hz. This
-  data volume is much greater than previously existing submillimetre
-  cameras, and represents a significant data anlaysis challenge. We
-  describe the production of SCUBA-2 maps using the Sub-Millimetre
-  User Reduction Facility (SMURF) in which we have adopted a fast,
-  iterative approach to map-making that enables data reduction on
-  single, modern, high-end desktop computers with execution times that
-  are shorter than the observing times.  SMURF is used in an automated
-  setting both at the telescope for real-time feedback to observers,
-  as well as for the production of generic science products at the
-  Canadian Astronomical Data Centre. A suite of tools are also
-  provided for interactive data analysis.  Three detailed case studies
-  are used to: (i) explore convergence properties of the map-maker
-  using simple prior constraints (Uranus -- a point source); (ii)
-  achieve the white-noise limit of the instrument for faint
-  point-source studies (blank-field survey of the Lockman Hole); and
-  (iii) demonstrate that our strategy is capable of recovering angular
-  scales up to the size of the array footprint (approximately
-  5\,arcmin) for bright extended sources (star-forming region M17).
+  data volume is much greater than that of previously existing
+  submillimetre cameras, and represents a significant data anlaysis
+  challenge. We describe the production of SCUBA-2 maps using the
+  Sub-Millimetre User Reduction Facility (SMURF) in which we have
+  adopted a fast, iterative approach to map-making that enables data
+  reduction on single, modern, high-end desktop computers with
+  execution times that are shorter than the observing times.  SMURF is
+  used in an automated setting both at the telescope for real-time
+  feedback to observers, as well as for the production of generic
+  science products at the Canadian Astronomical Data Centre. A suite
+  of tools are also provided for interactive data analysis.  Three
+  detailed case studies are used to: (i) explore convergence
+  properties of the map-maker using simple prior constraints (Uranus
+  -- a point source); (ii) achieve the white-noise limit of the
+  instrument for faint point-source studies (blank-field survey of the
+  Lockman Hole); and (iii) demonstrate that our strategy is capable of
+  recovering angular scales up to the size of the array footprint
+  (approximately 5\,arcmin) for bright extended sources (star-forming
+  region M17).
 \end{abstract}
@@
 SCUBA-2's predecessor SCUBA
-\citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO, another
-camera from the same generation. Generally speaking, such methods are
-fast, although dependending on the science goals and noise properties
-of the data, they may not achieve the best noise performance on the
-angular scales of interest. A particular case of this method was
-successfully used for the reduction of data from BOLOCAM
+\citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO
+\citep[e.g.][]{omont2001}, another camera from the same
+generation. Generally speaking, such methods are fast, although
+dependending on the science goals and noise properties of the data,
+they may not achieve the best noise performance on the angular scales
+of interest. In more recent years a method that has been popular for
+reducing data in fields of faint point sources with BOLOCAM
 \citep[e.g.][]{laurent2005}, and its younger sibling AzTEC
-\citep[e.g.][]{scott2008}, using ``principal component analysis''
-(PCA) cleaning. A statistical black-box removes the most correlated
+\citep[e.g.][]{scott2008}, is ``principal component analysis'' (PCA)
+cleaning. A statistical black-box removes the most correlated
 components of the bolometer signals, enabling the detection of
 point-sources very close to the theoretical white-noise limits of the
-detectors with reasonable computation times. However, PCA cleaning is
-not a good generic solution for producing maps of extended structures,
-since such sources produce correlated signals amongst many detectors,
-and are removed by this procedure.
-
-The best map-making strategies for recovering information on all
-angular scales are \emph{maximum likelihood} techniques, in which the
-time-series data are expressed as a sampling of the ``true'' map of
-the sky plus noise, and then the equation is inverted to estimate the
-map as some weighted linear combination of the data that minimizes the
-variance.  The first good description of this method appears in
+detectors with reasonable computation times when small numbers of
+bolometers are involved (hundreds rather than thousands of
+detectors). However, PCA cleaning is not a good generic solution for
+producing maps of extended structures, since such sources produce
+correlated signals amongst many detectors, and are removed by this
+procedure.
+
+The best existing map-making strategies for recovering information on
+all angular scales are \emph{maximum likelihood} techniques, in which
+the time-series data are expressed as a sampling of the ``true'' map
+of the sky plus noise, and then the equation is inverted to estimate
+the map as some weighted linear combination of the data that minimizes
+the variance.  The first good description of this method appears in
 \citet{janssen1992} for the production of maps from the COsmic
@@
 additional noise components. More similar to the algorithm described
-here was the iterative pipeline recipe for fitting and removing baseline drifts
-in SCUBA scan-map data as the astronomical image estimate improved
-\citet{1999ASPC..172..171J}. The two closest modern relatives of SMURF are the
-Comprehensive Reduction Utility for SHARC-2
+here was the iterative pipeline recipe for fitting and removing
+baseline drifts in SCUBA scan-map data as the astronomical image
+estimate improved \citep{1999ASPC..172..171J}. The two closest modern
+relatives of SMURF are the Comprehensive Reduction Utility for SHARC-2
 \citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
@@
 In this section we summarize how \scuba\ bolometers detect temperature
-variations (Section~\ref{sec:bolos}), describe \scuba\ scan strategies
-(Section~\ref{sec:scan}), describe the differences between SCUBA-2
-Shared Risk Observing data and those of the the fully commissioned
-instrument (Section~\ref{sec:s2sro}), typical bolometers signals
-(Section~\ref{sec:bolosignal}), the impact of magnetic field pickup
-(Section~\ref{sec:magpickup}, and finally illustrate the use of
-principal component analysis (Section~\ref{sec:pca}).
+variations (Section~\ref{sec:bolos}), give exampels of typical
+bolometers signals (Section~\ref{sec:bolosignal}), and the impact of
+magnetic field pickup (Section~\ref{sec:magpickup}), and finally
+illustrate the use of principal component analysis to explore the
+correlated noise properties of \scuba\ data (Section~\ref{sec:pca}).
 
 %--------------------------------------------------
-\subsection{How the \scuba\ bolometers work}
+\subsection{Description of \scuba}
 \label{sec:bolos}
@@
 
-The \scuba\ bolometer temperatures are read-out using superconducting
-transition-edge sensors (TESs), with critical temperatures
-$T_\mathrm{C}$ approximately 140 and 200\,mK at 850 and 450\,\micron,
-respectively. The range about $T_\mathrm{C}$ in which the transition
-from a superconductor to a normal material occurs is extremely small
-(...typical value...?). The entire \scuba\ focal plane is held at a
-\emph{base temperature} somewhat below $T_\mathrm{C}$. The TESs are
-then biased up into the transition using two systems. First,
-individual pixel heaters (using a current bias to apply the desired
-Joule heating) are set to fixed values at the start of an observation
-to compensate for most of the loading from the instrument and
-atmosphere (in better weather, less compensation from the heaters is
-required). Then, an electrical voltage bias is applied. As the
-instrument observes, and there are tiny temperature fluctuations due
-to astronomical sources and atmosphere, the impedence
-changes. Thermoelectrical feedback helps to keep the TESs ``locked''
-in the transition as this occurs; an increase in temperature increases
-the impedence, and lowers the current and resulting heating
-power. Changing magnetic fields resulting from these current changes
-are then amplified using chains of superconducting quantum
-interference devices (SQUIDs), and the larger output current is
-finally digitized. Each detector has its own SQUID for the first-stage
-of the amplification, but the remaining stages occur within a common
-chain of SQUIDs for each of the 32 columns. All 40 rows are then
-read-out in sequence at a row-visit frequency of about 12\,kHz. This
-multiplexing scheme greatly reduces the number of readout and bias
-lines to each detector, making such large-format bolometer array
-designs feasible. Such a high sample rate is unnecessary to produce
-maps, so the data are re-sampled to approximately 200\,Hz.  It should
-also be noted that there is an additional 41st row of SQUID readouts
-that are not connected to TES. These ``dark SQUIDs'' are used to track
-non-thermal noise sources that are common to each column's amplifier
-chain.
-
-The response of the output digitized current, $I$, with respect to a
-change in the power $P$ incident on the bolometer, $dI/dP$ depends on
-the efficiency of the thermal absorber, how well the TES is biased
-into the transition, and the gain of the SQUID amplifiers. In
-practice, the response of a detector can therefore vary somewhat from
-observation to observation depending on how the arrays have been
-tuned, and the observing conditions \citep[see][]{holland2012}. To
-track these changes, at the start of each observation, the current is
-ramped through the bolometer heaters such that $dI/dP$ can be
-measured. We will refer to the measurements of $dI/dP$ across the
-focal plane as the \emph{flatfield} for a given observation.
-
-In general, bolometers have two predominant noise components: (i) the
+Incoming light passes first through a beam-splitter, and then bandpass
+filters, providing simultaneous illumination and wavelength definition
+of both the 450 and 850\,\micron\ focal planes. The bolometers in each
+focal plane consist of absorbers coupled to superconducting
+transition-edge sensors (TESs) for thermometry. The focal planes are
+held at a temperature somewhat below the critical temperatures of the
+TESs by the fridge. The devices are then biased into the
+superconducting transition using a combination of individual pixel
+heaters (to compensate for most of the bulk variations in loading from
+the atmosphere and other backgrounds), and an electrical voltage
+bias. Throughout an observation the heater bias is fixed, but
+smaller-scale load variations both from the backgrounds and, to a much
+lesser extent, astronomical sources, significantly alter the TES
+impedences as their temperatures vary. Thermoelectrical feedback then
+keeps the TESs ``locked'' in the superconducting transition; an
+increase in temperature increases the impedence, and lowers the
+current, and therefore the heating resulting from the voltage
+bias. These current changes produce varying magnetic fields that are
+amplified using chains of superconducting quantum interference devices
+(SQUIDs), and the larger output currents are digitized. Each detector
+has its own SQUID for the first-stage of the amplification, but the
+remaining stages occur within a common chain of SQUIDs for each of the
+32 columns. All 40 rows are read-out in sequence at a row-visit
+frequency of about 12\,kHz. Such a high sample rate is unnecessary to
+produce maps, so the data are re-sampled to approximately 200\,Hz
+before writing to disk.  It should also be noted that there is an
+additional 41st row of SQUID readouts that are not connected to
+TESs. These ``dark SQUIDs'' track non-thermal noise sources that are
+common to each column's amplifier chain. The relationship between the
+output digitized current, $I$, and the input power, $P$, or $dI/dP$ is
+established using \emph{flatfield} observations immediately prior to
+science observations, in which the output signal is measured
+throughout a ramp of the pixel heaters (which provide a known input
+power); see Section~2.1 in \citet{dempsey2012} and Section~5.3 in
+\citet{holland2012}. Finally, the conversion to astronomical flux
+units from power involves a correction for atmospheric extinction
+\citep[primarily using the JCMT Water Vapour Monitor to track
+line-of-sight opacity variations, see Section~3 in][]{dempsey2012},
+and the application of a flux conversion factor (FCF) which is
+established from regular measurements of calibrators such as Uranus
+\citep[Section~5 in][]{dempsey2012}.
+
+The predominant noise features in \scuba\ bolometers are: (i) the
 effectively white thermal noise produced by the phonon and optical
 loading on the absorbers (largely independent from bolometer to
-bolometer); and (ii) longer timescale drift resulting from slow
+bolometer); and (ii) longer timescale drifts resulting from slow
 variations in the mean load (e.g. fridge, telescope, or atmospheric
 changes), and also the readout electronics (often correlated amongst
-many or all of the bolometers). As such, the spectrum corresponding to
-a bolometer time-series exhibits a ``$1/f$'' shape below some knee
-frequency, $f_\mathrm{knee}$, and is relatively flat at higher
-frequencies. Other smaller effects may include line features in the
-spectrum (common culprits are aliased high-frequency noise detected
-during the multiplexed readout of the arrays as described in
-Section~\ref{sec:bolos}, pickup from AC mains, or microphonic signals
-induced by mechanical vibrations), and finally a roll-off towards the
-Nyquist frequency caused by the anti-aliasing filter that precedes
-digitisation.
-
-The final conversion of power received by the detector into
-astronomical flux units requires two more fundamental corrections: (i)
-atmospheric extinction (due to atmospheric water vapour) must be
-compensated for as a function of time; and (ii) the corrected power is
-related to an astronomical flux density through the effective area of
-the detector, the spectral response of the bolometers (with bandpasses
-set by optical filters), the spectrum of the source, and the optical
-efficiencies of the dish surface and other optical components in the
-system -- a \emph{flux conversion factor} (FCF). Extinction is
-measured, in most cases, using the JCMT line-of-sight water vapour
-radiometer (WVM) which is calibrated using sky dips and the CSO
-225\,GHz tipper. The FCF is determined empirically using measurements
-of standard sources. For further details on these calibration steps,
-see \citet{dempsey2012}.
-
-%--------------------------------------------------
-\subsection{Scan strategies}
-\label{sec:scan}
-%--------------------------------------------------
-
-In order to observe astronomical sources, it is clearly advantageous
-that the signals appear in the region of the spectrum that offers the
-highest \snr, namely frequencies above $f_\mathrm{knee}$. This effect
-is achieved by modulating the signal; the original SCUBA primarily
-used the JCMT chopping secondary to place sources into the spectrum at
-about $\sim$10\,Hz. The problem with this strategy was that SCUBA was
-ultimately limited to recovering information on angular scales smaller
-than the chop throw \citep[up to several arcmin,][]{holland1999}.  The
-primary approach to map-making with \scuba\ is instead to slew the
-entire telescope at such a rate that angular scales of interest are
-crossed on time scales shorter than $1/f_\mathrm{knee}$, as with other
-earlier ground-based instruments that did not use a chopping secondary
-\citep[e.g.,][]{glenn1998,weferling2002,wilson2008,kovacs2008b}. The
-desire to scan rapidly has two basic constraints: (i) while the
-telescope can physically move up to about 1800 and
-3600\,arcsec\,s$^{-1}$ in azimuth and elevation, respectively, the
-acceleration limit is about 600\,arcsec\,s$^{-2}$ (in azimuth, larger
-in elevation), meaning that significant time is waisted on
-``turn-arounds'' at large scan speeds; and (ii) bolometers are
-designed with a thermal time constant that balances response with
-sensitivity. Taking these constraints into consideration, \scuba\ was
-ultimately designed to operate at a maximum scan speed of
-600\,arcsec\,sec$^{-1}$. The data acquisition system has a target
-sample rate of 200\,Hz (although in practice it is typically in the
-range 170--180\,Hz), which at that scan velocity results in a sample
-every 3.0\,arcsec, or approximately one third of the 450\,\micron\
-diffraction-limited full-width at half-maximum (FWHM) -- a typical
-rule-of-thumb for adequately sampling a Gaussian point spread function
-(PSF). Similarly, the \scuba\ bolometers are designed to have a time
-constant of $\sim$1\,ms in order to avoid distorting PSF-sized sources
-at this maximum scan rate.
-
-The scan strategy is designed, primarily, to mitigate the effects of
-$1/f$ drift. The simplest scan one might use is a raster
-``boustrophedon'' pattern, i.e. a series of parallel passes across the
-field. However, such a scan yields extremely poor performace for
-structures that are transverse to the scan direction, since a long
-time will have ellapsed between each parallel step, and the signal
-will therefore be totally dominated by the $1/f$ noise. In addition,
-other periodic noise features may appear in the data. With such a
-regularly spaced scan strategy, these noise features can easily
-conspire with the scan pattern to produce similarly periodic noise
+many or all of the bolometers). As such, the power spectra of
+bolometer time-series exhibit a ``$1/f^\alpha$'' ($\alpha>0$) shape
+below some knee frequency, $f_\mathrm{knee}$, and they are then
+relatively flat at higher frequencies. The spectra also include line
+features which are predominantly produced by aliased high-frequency
+noise sources detected during the multiplexed readout stage as
+described above.  Finally, an anti-aliasing filter produces roll-off
+in the spectrum at frequencies approaching Nyquist prior to the
+200\,Hz re-sampling.
+
+We note that the noise performace of \scuba\ has evolved over time. An
+initial ``\scuba\ shared-risk observing'' period (S2SRO) took place
+during February and March 2010, during which each of the 450 and
+850\,\micron\ focal planes were populated with single subarrays, s4a
+and s8d respectively. In addition to having significantly fewer
+available bolometers than the current fully-commissioned instrument
+(and therefore a mapping speed reduced by a factor of $\sim$4),
+instabilities in the fridge lead to a large periodic signal with a
+period of about 25\,s that was correlated amongst the detectors. After
+upgrading and commissioning the instrument, a servo using newly added
+NTD thermometers in the focal planes has effectively mitigated this
+problem \citep[Section~2.5 in][]{holland2012}. In addition, there were
+improvements to the magnetic shielding (reducing magnetic field pickup
+as described in Section~\ref{sec:magpickup}), as well as more
+effective removal of aliased noise sources (which has reduced both the
+presence of line features and the mean white noise level).
+
+In order to reduce the impact of noise on the final maps, it is
+essential that an appropriate scan strategy be used. First, it is
+clearly advantageous that the astronomical signals appear in the
+lowest-noise region of the power spectrum, namely frequencies above
+$f_\mathrm{knee}$. This is achieved by scanning the telescope at large
+slew rates, which are practically limited to about
+600\,arcsec\,s$^{-2}$ for the JCMT in order to have reasonable
+turn-around overheads (the amount of time spent changing direction and
+speed, over the largest elevation range possible). At this maximum
+speed there is a sample every 3.0\,arcsec, or approximately one third
+of the 450\,\micron\ diffraction-limited full-width at half-maximum
+(FWHM) -- a typical rule-of-thumb for adequately sampling a Gaussian
+point spread function (PSF). The scan \emph{patterns} are designed,
+primarily, to mitigate the effects of low-frequency $1/f$ drift. The
+simplest scan one might use is a raster ``boustrophedon'' pattern,
+i.e. a series of parallel passes across the field. However, such a
+scan yields extremely poor performace for structures that are
+transverse to the scan direction, since a long time will have elapsed
+between each parallel step. In addition, other periodic noise features
+may conspire with such a regularly-spaced pattern to produce periodic
 structures in the map, or \emph{scan-synchronous} noise.  It has
@@
 of the map on as many different time scales, and scan angles as
-possible offer the best protection against large-scale noise
+possible, offer the best protection against large-scale noise
 \citep[see discussion in][and references therein]{kovacs2008b}; such
@@
 taking into consideration both mechanical and scheduling constraints,
-\scuba\ presently has three primary scan modes: PONG, which covers
+\scuba\ presently has two primary scan modes: PONG, which covers
 rectangular areas larger than the array footprint in a closed pattern
@@
 combined with a rotation through a number of fixed position angles to
-create a ``rotating PONG'' with even better cross-linking); Lissajous,
-in which sinusoidal functions describe the motion along orthogonal
-axes (also to cover rectangular areas -- the PONG and Lissajous modes
-are both related to similar scan modes developed at the Caltech
-Submillimetre Observatory); and a ``constant velocity daisy'' for
-small areas (of order the array footprint), in which the telescope
-moves in a circle, whose centre also slowly traces out a circle. For a
-more complete description of the \scuba\ observing modes, see
-\citet{holland2012}.
-
-%-------------------------------------------------
-\subsection{Differences between the S2SRO and 2011 fully-commissioned instrument}
-\label{sec:s2sro}
-%-------------------------------------------------
-
-The S2SRO period took place during February and March 2010. For the
-purpose of this paper, we will consider data taken as early as 2009
-December 3, at which point the instrument was being commissioned in a
-similar configuration to that of the S2SRO period. At this time, each
-of the 450 and 850\,\micron\ focal planes were populated with single
-subarrays, s4a and s8d respectively. There were three main problems to
-contend with during this period. First, the fridge was not able to
-cool the focal plane to the desired base temperature (partly due to
-the heat conductance of the wires from the arrays during this period)
-which affected stability, particularly at 850\,\micron. Second, a
-non-sinusoidal, large-scale oscillation in the fridge temperature
-could easily be seen as a common signal to all of the detectors with a
-period of approximately 25\,s.  Third, there were problems with the
-subarray fabrication process which led to large numbers of dead
-pixels, as well as large-scale variations in $T_\mathrm{C}$ and the
-thermal conductivity, $G$. These latter issues made it difficult to
-simulataneously place all of the TESs into the optimal region of the
-superconducting transition, leading to significantly poorer
-sensitivities. While there were moderate variations from
-night-to-night during the observing period, there were typically 600
-to 900 working detectors in each of the s4a and s8d subarrays.
-
-Since S2SRO the fridge performance has been greatly improved using NTD
-thermometry to servo the temperature to a constant value, and with new
-wiring, adequate base temperatures can now be achieved. However, the
-gradient in detector properties across the focal plane continues to be
-a problem now that there is a full complement of subarrays (which
-still includes the two subarrays used during S2SRO). Optimal tuning to
-achieve the greatest overall array sensitivity is described in
+create a ``rotating PONG'' with even better cross-linking) which was
+based on a similar scan mode developed at the Caltech Submillimeter
+Observatory; and a ``constant velocity daisy'' (CV Daisy) for small
+areas (of order the array footprint), in which the telescope moves in
+a circle, whose centre also slowly traces out a circle. For a more
+complete description of the \scuba\ observing modes, see Section~5 in
 \citet{holland2012}.
@@
 \caption{Bolometer power spectral densities for the same two data sets
-  (before and after upgrades) used in Fig.~\ref{fig:bolos_mix}. Four
-  of the most sensitive bolometers have been selected at each
-  wavelength, and the flat-fielded (but otherwise raw) PSDs are shown
+  (before and after upgrades) used in Fig.~\ref{fig:bolos_mix}. The
+  PSDs have been boxcar smoothed with a width of 0.1\,Hz to reduce the
+  noise slightly and clarify some features. Four of the most sensitive
+  bolometers have been selected at each wavelength, and the
+  flat-fielded and step-corrected (but otherwise raw) PSDs are shown
   as coloured dotted lines (the blue signals are for the same time
@@
 significant signal that is obviously correlated with the telescope
-motion. It is also correlated amongs many of the detectors, and
+motion. It is also correlated amongst many of the detectors, and
 common-mode removal corrects it to some extent. Part of the reason
@@
 discrete Fourier Transform. The units of the PSD written in this form
-are pW$^2$\,Hz$^{-1}$; without the factor of 2 in front the convention
-would be to write the units as pW$^2$\,s to distinguish the two
-normalisations.
+are pW$^2$\,Hz$^{-1}$.
 
 The dotted coloured lines in Fig~\ref{fig:pspec} show the PSDs for
-raw, though flat-fielded data. At each wavelength, and in both data
-sets, there are clear $1/f$ knees at frequencies ranging from roughly
-1 to 2\,Hz, followed by a predominantly white spectrum punctuated by
-line features at higher frequencies, and finally a roll-off caused by
-the anti-aliasing filter above $\gsim 70$\,Hz. As indicated in the
-previous section, the correlation between the low-frequency components
-of the different bolometer signals is large. The solid black lines in
+raw, though flat-fielded and step-corrected (Section~\ref{sec:steps})
+data. At each wavelength, and in both data sets, there are clear $1/f$
+knees at frequencies ranging from roughly 1 to 2\,Hz, followed by a
+predominantly white spectrum punctuated by line features at higher
+frequencies, and finally a roll-off caused by the anti-aliasing filter
+above $\gsim 70$\,Hz. As indicated in the previous section, the
+correlation between the low-frequency components of the different
+bolometer signals is large. The solid black lines in
 Fig.~\ref{fig:pspec} indicate the PSDs of the common-modes
@@
 
-An additional noise source that appears to be significant in only a
-small subset of the data (and moreso during the S2SRO period) is
-magnetic field pickup. Since the bolometer signals are ultimately
-detected through the amplification of magnetic fields, any additional
-changing fields within the cryostat will also be detected.
-
-An example of an observation where pickup appears to be significant is
-shown in Fig.~\ref{fig:magpickup}. The time-series for two
-450\,\micron\ bolometers in the same column (not flatfielded) show
-that there is a strong signal with a similar shape, but opposite
-signs. This behaviour is seen across the entire array. The dark squid
-signal for the same columns exhibits a similar shape and
-amplitude. Since the dark squid has no thermal absorber or TES
-attached to it, this observed signal is not likely to be optical or
-thermal in nature (although there can be some cross-talk with the
-bolometers). It is also for this reason that the higher-frequency
-noise that appears in the bolometer data (thermal) does not appear in
-the dark squid, leading to a significantly higher \snr. Due to the
-fact that the sign of the gain in each stage of SQUID amplification is
-random (although the combined gain is constrained to be negative), and
-since magnetic field pickup is only seen at the input to the second
-stage, the pickup can appear with random signs for bolometers along a
-a column, giving it a distinct signature from other common signals
-that appear with the same sign.
+An additional noise source that is significant in only a small subset
+of the data (and moreso during the S2SRO period) is magnetic field
+pickup. Since the bolometer signals are ultimately detected through
+the amplification of magnetic fields, any additional changing fields
+within the cryostat will also be detected.
+
+Example data from the 450\,\micron\ subarray s4a where pickup appears
+to be significant (observation 16 on 20100228) are shown in
+Fig.~\ref{fig:magpickup}. The time-series for two bolometers in the
+same column (not flatfielded) show that there is a strong signal with
+a similar shape, but opposite signs. This behaviour is seen across the
+entire array. The dark squid signal for the same columns exhibits a
+similar shape and amplitude. Since the dark squid has no thermal
+absorber or TES attached to it, this observed signal is not likely to
+be optical or thermal in nature (although there can be some cross-talk
+with the bolometers). Due to the fact that the sign of the gain in
+each stage of SQUID amplification is random (although the combined
+gain is constrained to be negative), and since magnetic field pickup
+is only seen at the input to the second stage, the pickup can appear
+with random signs for bolometers along a a column, giving it a
+distinct signature from other common signals that appear with the same
+sign.
 
@@
 \caption{Evidence for significant magnetic field pickup for
-  observation 20100228\_00016.  The top panel shows two un-flatfielded
-  (but mean-subtracted) bolometer time-series from the same column,
-  with a 200 sample boxcar smooth (approximately 1\,s), illustrating
-  that they are dominated by a similar signal with opposite signs. The
-  second panel shows the dark squid signal for the column, also
+  observation 16 on 20100228 at 450\,\micron\ (s4a subarray).  The top
+  panel shows two un-flatfielded (but mean-subtracted and step
+  corrected) bolometer time-series from the same column, with a 200
+  sample boxcar smooth (approximately 1\,s), illustrating that they
+  are dominated by a similar signal with opposite signs. The second
+  panel shows the dark squid signal for the column, also
   mean-subtracted and with the same boxcar smooth. The bottom panel
@@
 
-A method that has been employed by teams analyzing BOLOCAM and AzTEC
-data to identify and remove correlated noise is Principal Component
-Analysis \citep[PCA,][]{laurent2005,scott2008}. We initially used PCA
-to gain further insight into the nature of the \scuba\ noise
-properties, but it is now also included in SMURF as method for
-cleaning the data.
-
-The basic method is as follows: (i) a covariance matrix is built up
-for all pairs $(i,j)$ of the $N$ bolometer time series,
-$\langle\mathbf{b}_i(t),\mathbf{b}_j(t)\rangle$; and (ii) a singular
-value decomposition is used to identify a new set of statistically
-uncorrelated eigenvectors, $\mathbf{\xi}_i(t)$ (i.e. whose covariance
-matrix is diagonal), such that each of the bolometer time series is a
-linear combination of the eigenvectors, or \emph{components}
-i.e. $\mathbf{b}_i(t) = \bar{\mathbf{\xi}}
-\mathbf{\lambda}_i^\mathrm{T}$, where each row of the matrix
-$\bar{\mathbf{\xi}}$ is an eigenvector, and
-$\mathbf{\lambda}_i^\mathrm{T}$ is a column vector containing the
-corresponding eigenvalues. The $\mathbf{\xi}_i$ are normalized by
-their \rms, so that the amplitude of each component is stored in the
-eigenvalues. In the earlier analyses mentioned, the low-frequency
-noise is assumed to be encapsulated in those components with the
-largest eigenvalues. Removing the projection of the time series along
-those components has been used successfully to reduce $1/f$ noise
-while retaining most of the signal in point-like sources. It is also
-worth noting that there is an implicit assumption in this analysis
-that the noise properties are stationary.
-
-One unique feature of the \scuba\ data is that one can perform a PCA
-analysis of both the 450 and 850\,\micron\ data simultaneously. The
-first 6 most significant modes of our analysis are shown in
-Fig.~\ref{fig:pca}, with the eigenvectors in the top panels, and maps
-of the eigenvalues in each subarray in the bottom panels.
-
 \begin{figure*}
@@
 \includegraphics[width=\linewidth]{pca.pdf}
-\caption{The first six modes from a principal component analysis,
-  ordered by decreasing significance, of the combined 450 and
-  850\,\micron\ time-series bolometer data. For each component, the
-  top plot shows the time series of the eigenvector, normalized by its
-  \rms. The bottom coloured panels show the eigenvalues for the
-  bolometers at each wavelength (the amplitude of the eigenvector in
-  the time series). For reference, both the mean, $\bar{\lambda}$, and
-  \rms, $\lambda_\mathrm{rms}$, eigenvalues for the bolometers in each
-  subarray are also shown. The first two components have comparable
-  amplitudes, and dominate by at least a factor of 10 any other
-  components of the time series. We suspect that they are caused by
-  atmospheric variations (expected to be brighter at 450\,\micron),
-  and fridge oscillations (expected to be comparable at each
-  wavelength). However, the two eigenvectors do not necessarily map
-  directly to these physical sources under this interpretation; rather
-  they are two \emph{different} linear combinations of the underlying
-  physical signals that give rise to statistically uncorrelated
-  components. The third component accurately reproduces an apparent
-  data-acquisition glitch that appears only at 450\,\micron. The
-  fourth component is a periodic feature only observed in column 15 of
-  the 450\,\micron\ subarray. The fifth component shows a series of
-  steps and glitches that appear predominantly along the bottom-left
-  edge of the 850\,\micron\ subarray where the bolometers are poorly
+\caption{The first twelve components from a principal component
+  analysis, ranked by the mean eigenvalues, of combined 450 (s4a
+  subarray) and 850\,\micron\ (s8b subarray) time-series bolometer
+  data, for the same observation as that used in the right-hand panels
+  of Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec}. For each
+  component, the top plot shows the time series of the eigenvector,
+  normalized by its \rms. The bottom coloured panels show the
+  eigenvalues for the bolometers at each wavelength (the amplitude of
+  the eigenvector in the time series). For reference, both the mean,
+  $\bar{\lambda}$, and \rms, $\lambda_\mathrm{rms}$, eigenvalues for
+  the bolometers in each subarray are also shown. The first two
+  components have comparable amplitudes, and dominate by at least a
+  factor of 10 any other components of the time series. We suspect
+  that they are caused by atmospheric variations (expected to be
+  brighter at 450\,\micron), and fridge oscillations (expected to be
+  comparable at each wavelength). However, the two eigenvectors do not
+  necessarily map directly to these physical sources under this
+  interpretation; rather they are two \emph{different} linear
+  combinations of the underlying physical signals that give rise to
+  statistically uncorrelated components. The third component
+  accurately reproduces an apparent data-acquisition glitch that
+  appears only at 450\,\micron. The fourth component is a periodic
+  feature only observed in column 15 of the 450\,\micron\
+  subarray. The fifth component shows a series of steps and glitches
+  that appear predominantly along the bottom-left edge of the
+  850\,\micron\ subarray where the bolometers are poorly
   biased. Finally, the sixth component is degenerating into a less
@@
 
-The majority of correlated signal at both wavelengths is clearly
-broken down into the first two components plotted. The first dominates
-at 450\,\micron, whereas the second dominates at 850\,\micron. For
-reference, the eigenvalues of these two modes are in the range
-$\sim1$--4\,pW, whereas the remaining modes have eigenvalues
-$\lsim0.3$\,pW. As mentioned earlier, one simple model for the
-correlated signal is a simple combination of oscillations in base
-temperature (which would be similar in each band), with a varying
-atmospheric signal (which is expected to be stronger at
-450\,\micron). Indeed, the fainter Component 2 most closely resembles
-the mixing chamber temperature variations (Fig.~\ref{fig:bolos_mix}),
-whereas Component 1 is stronger at 450\,\micron\ and does not resemble
-the fridge signal as closely.
-
-This physical interpretation of the eigenvectors should, however, be
-taken with a grain of salt, since the PCA analysis is really a
-statistical black box. Suppose most of the low frequency noise really
-could be decomposed into two components, the fridge signal
-$\mathbf{F}(t)$, and an atmospheric signal $\mathbf{A}(t)$. The two
-eigenvectors obtained from the PCA analysis are simply linear
-combinations of $\mathbf{F}(t)$ and $\mathbf{A}(t)$ that are
-statistically uncorrelated. Depending on the particular realisations
-of these two noise sources, this decomposition may, or may not give
-rise to eigenvectors that resemble the two underlying physical
-sources. Regardless, similar analyses of a few data sets during S2SRO
-give comparable results; most of the signal at both 450\,\micron\ and
-850\,\micron\ appears to be produced by two dominant underlying
-mechanisms, even if we cannot identify the mechanisms with any
-confidence.
-
-The subsequent components identified by the PCA analysis in
-Fig.~\ref{fig:pca} vary wildly. Component 3, though not obvious from
-its eigenvector plot, is almost a pure ramping waveform that only
-appears in the s4a data, and produces the broad line featuress centred
-over 45\,Hz and 70\,Hz in the 450\,\micron\ PSDs in
-Fig.~\ref{fig:pspec}. In other data sets, this third component has
-also often instead exhibited a regularly-spaced series of spikes with
-heights quantized into several amplitude families (again, only in the
-s4a array). In both cases, the shape and phase of this signal is the
-same in each bolometer, although the amplitude (and sign) varies, as
-evidenced by the large random scatter in the s4a eigenvalue maps. It
-is thought that these features are some sort of aliased high-frequency
-data acquisition glitch, as the spikes (when they appear) exhibit the
-expected point response function of the anti-aliasing
-filter. Similarly, Component 4 also appears to be a data acquisition
-glitch in the s4a array, although along a single column, and with a
-peak at a lower frequency of 40\,Hz, together with a second feature
-right at the Nyquist frequency.
-
-The last two components in Fig.~\ref{fig:pca} have an entirely
-different character. The amplitudes of both components are greatest in
-low-responsivity regions of the two subarrays (lower-right at
-450\,\micron\ and bottom-left at 850\,\micron). Bolometers in these
-regions of the subarray are probably not optimally biased, and are
-more sensitive to changes in the loading, possibly resulting in the
-step and spikes that are seen in Component 5. Component 6 is clearly
-related to Component 5, though with the absence of the steps and
-spikes, so that together they seem to represent a continuum of shapes
-that are linear combinations of the spikes and steps, with the slower
-drift evident only in Component 6.
+A method that has been used to remove correlated noise as part of the
+map-making procedure for BOLOCAM and AzTEC is Principal Component
+Analysis \citep[PCA,][]{laurent2005,scott2008}. Here we use PCA to
+explore correlated signals in \scuba\ data.
+
+The basic method is as follows: (i) a covariance matrix is built up
+for all pairs $(i,j)$ of the $N$ bolometer time series,
+$\langle\mathbf{b}_i(t),\mathbf{b}_j(t)\rangle$; and (ii) a singular
+value decomposition identifies a new set of statistically uncorrelated
+eigenvectors, $\mathbf{\xi}_i(t)$ (i.e. whose covariance matrix is
+diagonal), such that each of the bolometer time series is a linear
+combination of the eigenvectors, or \emph{components}
+i.e. $\mathbf{b}_i(t) = \bar{\mathbf{\xi}}
+\mathbf{\lambda}_i^\mathrm{T}$, where each row of the matrix
+$\bar{\mathbf{\xi}}$ is an eigenvector, and
+$\mathbf{\lambda}_i^\mathrm{T}$ is a column vector containing the
+corresponding eigenvalues. The $\mathbf{\xi}_i(t)$ are normalized by
+$[\sum_t \mathbf{\xi}_i^2(t)]^{1/2}$ such that the root mean square
+(\rms) amplitude of each component in a given bolometer signal is
+stored in the eigenvalues. In the earlier analyses mentioned, the
+low-frequency noise is assumed to be encapsulated in those components
+with the largest eigenvalues. Removing the projection of the time
+series along those components then significantly reduces $1/f$ noise
+while retaining most of the (higher-frequency) signal in point-like
+sources
+
+A novel feature of our \scuba\ analysis is that we can perform PCA
+with 450 and 850\,\micron\ data simultaneously, potentially helping us
+to differentiate thermal and optical noise signals (e.g., from the
+atmosphere) that might appear in both wavelengths, from other noise
+sources that are restricted to single subarrays (such as readout
+noise). In Fig.~\ref{fig:pca} we show the results for a combined
+analysis of the s4a and s8b subarrays for the first $\sim$100\,s of
+the same observation that was used in the right-hand
+(fully-commissioned) panels of Figs.~\ref{fig:bolos_mix} and
+\ref{fig:pspec}. The 12 most significant components are shown (ranked
+by the mean eigenvalues across all working bolometers in both
+bands). The top and middle panels for each component show the
+time-series and PSDs of the normalized eigenvectors. The bottom panels
+are maps indicating the eigenvalues (amplitudes) of the component
+across the focal plane (with the mean $\bar{\lambda}$ and and \rms,
+$\lambda_\mathrm{rms}$, of the eigenvalues for all of the bolometers
+calculated separately at each wavelength also shown).
+
+The majority of the correlated signal at both wavelengths in
+Fig.~\ref{fig:pca} is clearly produced by Components~1 and 2. In both
+cases, the eigenvector time series exhibit a roughly periodic signal
+that resembles the scan pattern in the right-hand panel of
+Fig.~\ref{fig:pca}. While these correlated signals also clearly appear
+at both wavelengths, referring to the maps of eigenvalues, they are
+considerably stronger at 850\,\micron\ (consistent with the visual
+appearance of the bolometer signals in Fig.~\ref{fig:pca}). Also note
+that the PSDs for these eigenvectors exhibit nearly a pure $1/f$
+signature with almost no line features, or a white-noise plateau
+(suggesting that these are high-\snr\ measurements of a purely
+low-frequency drift).
+
+What is the physical interpretation of these two components? First,
+the smoothly-varying pattern in the eigenvalue maps (particularly
+evident at 850\,\micron) suggests that the scan-synchronous noise
+(presumably magentic field pickup) has a different response across the
+focal plane; in this case the PCA may have identified two orthogonal
+shapes that, when mixed in different quantities, can reproduce most of
+the signal in each bolometer. It is also likely that a varying
+atmospheric contribution is also included in these eigenvectors. One
+would expect the atmosphere to be more significant at 450 than at
+850\,\micron, although it is not obviously visible due to the strong
+presence of the scan-synchronous noise.
+
+Next, Components~3, 4, 7 and 11 in Fig.~\ref{fig:pca} have a common
+character. In all cases, the PCA appears to have identified strong,
+but transient glitches that appear in only one or several bolometers,
+such as the two black bolometers near column 8, row 21 at
+850\,\micron\ in Component~3.
+
+The most interesting feature of Component~5 in Fig.~\ref{fig:pca} is a
+strong narrow line in the PSD at a frequency of about 57\,Hz
+(suggesting the 60\,Hz mains as the culprit). The eigenvalue maps show
+that it is only significant for several bolometers in the
+850\,\micron\ subarray: at column 22 row 22; several bolometers at the
+bottom of column 28; and more near the top-right of the subarray.
+
+Components~6, and 8--10 in Fig.~\ref{fig:pca} have several generic
+features in common. First, the PSDs exhibit moderate low-frequency
+$1/f$ noise. Second, there are broad-line features at frequencies
+above $\sim$30\,Hz which we believe to be aliased high-frequency
+noise, in part due to the clear correlation patterns that appear along
+columns in the eigenvalue maps at both wavelengths (e.g. columns 26
+and 29 at 450\,\micron in Component~6). Finally, these components also
+have potentially unrelated features mixed into them, like sudden
+offsets that are likely residuals due to imperfect step correction
+(e.g. $\sim$50\,s in the eigenvector time-series for Component~6; also
+see Section~\ref{sec:steps}).
+
+Finally, Component~12 in Fig.~\ref{fig:pca} is again significant in
+only a handful of detectors in the 850\,\micron\ subarray (e.g. column
+21 row 25). The time-series plot for the eigenvector exhibits a large
+positive tail of sudden excursions which gradually worsen over
+time. Bolometers that exhibit this signal are probably poorly biased.
 

### 2012-09-02 smurf_paper.tex
some work on smurf algorithm section

@@
   data, for the same observation as that used in the right-hand panels
-  of Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec}. For each
-  component, the top plot shows the time series of the eigenvector,
-  normalized by its \rms. The bottom coloured panels show the
-  eigenvalues for the bolometers at each wavelength (the amplitude of
-  the eigenvector in the time series). For reference, both the mean,
+  of Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec}. For each component
+  the top plot shows the time series of the normalized eigenvectors,
+  the middle plot its PSD, and the bottom coloured panels the
+  eigenvalues for the bolometers at each wavelength (the amplitudes of
+  the eigenvectors in the time series). For reference, both the mean,
   $\bar{\lambda}$, and \rms, $\lambda_\mathrm{rms}$, eigenvalues for
-  the bolometers in each subarray are also shown. The first two
-  components have comparable amplitudes, and dominate by at least a
-  factor of 10 any other components of the time series. We suspect
-  that they are caused by atmospheric variations (expected to be
-  brighter at 450\,\micron), and fridge oscillations (expected to be
-  comparable at each wavelength). However, the two eigenvectors do not
-  necessarily map directly to these physical sources under this
-  interpretation; rather they are two \emph{different} linear
-  combinations of the underlying physical signals that give rise to
-  statistically uncorrelated components. The third component
-  accurately reproduces an apparent data-acquisition glitch that
-  appears only at 450\,\micron. The fourth component is a periodic
-  feature only observed in column 15 of the 450\,\micron\
-  subarray. The fifth component shows a series of steps and glitches
-  that appear predominantly along the bottom-left edge of the
-  850\,\micron\ subarray where the bolometers are poorly
-  biased. Finally, the sixth component is degenerating into a less
-  significant, and more randomly scattered signal across the focal
-  plane. Subsequent components follow this trend, and would generally
-  be kept in a PCA-cleaning procedure.}
+  the bolometers in each subarray are also shown. Most of the
+  correlated $1/f$ signal is encapsulated in Components~1 and 2, and
+  appears to be dominated by scan-synchronous noise (compare with the
+  scan pattern in the right-hand panel of Fig.~\ref{fig:bolos_mix}).
+  Components~3, 4, 7 and 11 are transient glitches that only appear in
+  a small subset of the bolometers. Component~5 also appears in only a
+  few bolometers, and is dominated by a narrow line at 57\,Hz which
+  suggests that it is related to the 60\,Hz mains. Components~6 and
+  8--10 contain a mixture of broad lines above $\sim$30\,Hz (which are
+  probably aliased high-frequency noise), with some low-frequency
+  drifts. Finally, the erratic appearance of the eigenvector for
+  Component~12 is typical of a poorly-biased detector.}
 \label{fig:pca}
@@
 
-Components~6, and 8--10 in Fig.~\ref{fig:pca} have several generic
+Components~6 and 8--10 in Fig.~\ref{fig:pca} have several generic
 features in common. First, the PSDs exhibit moderate low-frequency
@@
 
+This example illustrates some of the types of correlated signals and
+patterns that PCA can identify in \scuba\ data. While there are
+typically one or two strong signal components detected, in general,
+the details can vary significantly from data set to data set, and the
+lengths of the time-series analyzed. The reason for this is probably
+due to the fact that many of the noise sources are not stationary in
+time (e.g. scan-synchronous noise which obviously depends on the scan
+pattern). It should also be clear from this example that while PCA
+offers some helpful insight into the various sources of noise, it does
+not necessarily identify clean patterns that can easily be
+modeled. For example, while there are clearly correlation patterns
+along columns (Components~6 and 8--10, particularly at 450\,\micron),
+as one might expect given the common amplification chain, the
+intensities of these high-frequency signals are random in the
+eigenvalue maps, and a common-mode signal estimated from the data for
+each column would not remove it.
+
 %------------------------------------------------------------------------------
@@
 
-The remainder of this section gives an overview of the model that is
-assumed by SMURF for the \scuba\ data, and the typical steps used to
-estimate a map. Subsequent sections discuss problems with this model
-and additional processing steps and constraints that can be used to
-improve the solution in different cases.
+In practice, the model generally adopted for low-frequency noise
+\scuba\ data is a single common-mode (average signal across all
+working bolometers in a given focal plane), with most remaining noise
+sources being removed by a high-pass filter. Many variations on this
+were attempted throughout the commissioning phase, but as the PCA in
+Secion~\ref{sec:pca} illustrates, less significant noise sources are
+both transient and particularly difficult to model.
+
+This section gives an overview of the SMURF algorithm, including data
+cleaning and other pre-processing steps (Section~\ref{sec:dataprep}),
+the iterative estimation of the map and noise model components
+(Section~\ref{sec:components}), and concludes with a discussion of
+convergence (Section~\ref{sec:converge}).
 
@@
 
-\item Describe the bolometer signal model, and with the help of a flow-chart
-show the basic procedure that SMURF follows.
-
 \item Discuss SMURF configurability
@@
 We express the signal observed by the $i$th bolometer as a function of
- time, \textbf{Per: $e_i(t)$ is written as a vector ( bold ) but I assume
- we use it as a scalar.}
-
+time,
+%
 \begin{equation}
@@
 \end{equation}
+%
 where $\mathbf{a}(t)$ is the time-varying signal produced by scanning
@@
 \subsection{Data preparation}
+\label{sec:dataprep}
 %-------------------------------------------------

### 2012-09-03 smurf_paper.tex
Update common mode section

@@
 each time step. Bolometer values that have been flagged as unusable
-for any reason are excluded from the mean. {\bf Per: has a comment in
+are excluded from the mean. {\bf Per: has a comment in
 here about stability of the gain pattern... it should be constant
@@
 
-Since bolometers have varying sensitivities, the amplitude of the
-common-mode variations will also vary from bolometer to bolometer. So
-comparing each bolometer time stream with the common-mode signal
-allows an estimate of the bolometer sensitivity to be obtained. In
-practice, a least squares linear fit is performed between the
-bolometer time series and the common-mode to determine a gain and
-offset for each bolometer.  The gain and offset for each bolometer is
-known as the \model{GAI} model. Each bolometer value can then
-optionally be scaled and shifted using these values so that all
-bolometers share a common (but as yet unknown) calibration. This
-provides an alternative (or additional) flat-fielding strategy to that
-described in section~\ref{sec:bolos}.
+Even after flat-fielding, bolometers may have slightly varying
+sensitivities and so the amplitude of the common-mode variations will
+also vary from bolometer to bolometer. Comparing each bolometer time
+stream with the common-mode signal allows an estimate of the relative
+bolometer sensitivity to be obtained. In practice, a least squares linear
+fit is performed between the bolometer time series and the common-mode to
+determine a gain and offset for each bolometer.  The gain and offset for
+each bolometer is known as the \model{GAI} model. Each bolometer value
+can then optionally be scaled and shifted using these values so that all
+bolometers share a common (but as yet unknown) calibration. This provides
+an alternative (or additional) flat-fielding strategy to that described
+in section~\ref{sec:bolos}.
 
@@
 common mode. Bolometers for which the correlation coefficient is below
-a specified limit, or which have a negative gain, are flagged as bad
-in order to omit them from the final map. If the option described
+a specified limit, or which have unusually high or low gain compared to
+the other bolometers, are flagged as bad in order to omit them from the
+final map. If the option described
 above for handling time-varying sensitivities is used, then a
@@
 
-The common-mode is calculated from all bolometers that have not
-previously been flagged as bad. Thus, aberrant bolometers with low
-correlation coefficients will be included, polluting the common-mode. For
-this reason, an iterative approach is taken to estimating the common-mode.
-The initial common-mode estimate is formed from all remaining bolometers.
-Aberrant bolometers with low correlation coefficients are then identified
-by comparing each bolometer time stream with the common-mode. Such
-aberrant bolometers are flagged as bad, and a new estimate of the
-common-mode is produced omitting the newly flagged bolometers. This
-process is repeated until no further bolometers are rejected.
+The common-mode value at each time slice is calculated as the unweighted
+mean of the values of all bolometers that have not previously been
+flagged as bad. Thus, aberrant bolometers with low correlation
+coefficients will be included, polluting the common-mode. By default, no
+attempt is made to exclude these aberrant bolometers\footnote{An option
+exists to reject unusually high or low bolometer values when calculating
+the common mode value at each time slice, but this seems to introduce
+instability into the convergence process and makes little difference to
+the final maps.}.
+
+Bolometer values that were flagged as poorly correlated with the common
+mode on the previous iteration are nevertheless included in the
+calculation of the common mode on the next iteration. If sensitivities
+are allowed to vary with time as described above, then poorly correlated
+values will be flagged in blocks of - typically of 30 seconds duration.
+Excluding such flagged blocks from subsequent estimates of the common
+mode can result in discontinuities in the common mode signal at the
+edges of these blocks. This in turn can lead to ringing in the low pass
+filter, and instabilities in the convergence process.
 

### 2012-09-03 smurf_paper.tex
more COM changes

@@
 mean of the values of all bolometers that have not previously been
-flagged as bad. Thus, aberrant bolometers with low correlation
-coefficients will be included, polluting the common-mode. By default, no
-attempt is made to exclude these aberrant bolometers\footnote{An option
-exists to reject unusually high or low bolometer values when calculating
-the common mode value at each time slice, but this seems to introduce
-instability into the convergence process and makes little difference to
-the final maps.}.
-
-Bolometer values that were flagged as poorly correlated with the common
-mode on the previous iteration are nevertheless included in the
-calculation of the common mode on the next iteration. If sensitivities
-are allowed to vary with time as described above, then poorly correlated
-values will be flagged in blocks of - typically of 30 seconds duration.
-Excluding such flagged blocks from subsequent estimates of the common
-mode can result in discontinuities in the common mode signal at the
-edges of these blocks. This in turn can lead to ringing in the low pass
-filter, and instabilities in the convergence process.
+flagged as bad for some other reason. Bolometer values that were flagged
+as bad simply because they were poorly correlated with the common-mode on
+the previous iteration are included in the new common-mode estimate. If
+such samples are excluded, there is a strong possibility of
+discontinuities appearing in the COM model at block boundaries.
+These in turn can lead to ringing in the low pass filter, and
+instabilities in the convergence process.
 

### 2012-09-03 smurf_paper.tex
Remove step fixing appendix

@@
 Detecting and correcting such a wide variety of steps reliably has proved
-to be a challenge. The current step-correction algorithm is described in
-detail in Appendix~\ref{appendix:steps}. In outline, the following stages
+to be a challenge. In outline, the following stages
 are involved in detecting steps in a single bolometer time stream:
@@
 
-%------------------------------------------------------------------------------
-\appendix
-\section[]{Step correction in detail}
-\label{appendix:steps}
-
-An outline of the step correction algorithm is given in
-Section~\ref{sec:steps}. This appendix contains a more detailed
-description.
-
-Detecting and correcting steps in a single bolometer time stream involves:
-
-\begin{enumerate}
-
-\item locating each candidate step;
-
-\item determining the widths of the candidate steps;
-
-\item determining the height of the step (steps can vary enormously in
-height -- for instance the step in the left-hand panel of
-Fig.~\ref{fig:steps1} is roughly 3\,pW in height, whereas the step at
-sample 5100 in the right-hand panel is only 0.01\,pW in height -- the
-height of a step is taken as the difference in the local median value
-just before and just after the step);
-
-\item correcting all subsequent data values by the step height;
-
-\item once all steps have been corrected, a global offset is added to the
-corrected bolometer time stream to ensure that the mean value before and
-after step correction is unchanged.
-
-\end{enumerate}
-
-\subsection{Detecting candidate steps}
-The first stage in the algorithm is to search the bolometer time stream for
-contiguous blocks of samples that have unusually large mean gradients
-compared to the neighbouring samples. Each such block of samples is
-considered to be a candidate step. The starting and ending time of each
-such block is noted. Once all candidate steps have been found for a
-bolometer, the algorithm proceeds to the next stage. Candidate steps are
-found as follows.
-
-\begin{enumerate}
-
-\item Smooth the bolometer time stream using a median filter of width
-50 samples. This reduces the effects of the noise in the data. Using a
-median filter rather than a mean filter helps to preserve the sharp edges
-at each step, and is robust against spikes, etc.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{steps2.pdf}
-\caption{A section of the bolometer signal from the right-hand panel
-of Fig.~\ref{fig:steps1} with the median smoothed data overlaid in
-red.}
-\label{fig:steps2}
-\end{figure}
-
-The red curve in Fig.~\ref{fig:steps2} shows the median smoothed data for
-the right hand bolometer from Fig.~\ref{fig:steps1}.
-
-\item At each time sample, find the difference between the two adjacent
-values in the median filtered bolometer data. This is proportional to the
-gradient of the median filtered bolometer data. Fig.~\ref{fig:steps3}
-shows these differences for the sample bolometer shown in previous
-figures. The high gradients at the three step are clearly visible at
-indices 5040, 5100 and 5280.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{steps3.pdf}
-\caption{Differences between adjacent values of the median smoothed
-data shown in Fig.~\ref{fig:steps2}.}
-\label{fig:steps3}
-\end{figure}
-
-\item Steps are located by looking for difference values that exceed 25
-times the RMS value of the nearby neighbouring differences. However, the
-local RMS of the differences includes the gradient of the underlying
-noise-free time series. So if there is a strong upward or downward trend
-in the time series data, the RMS of the difference values will be higher,
-potentially resulting in some steps being missed. For instance, at sample
-5000 the overall trend in the time series is downwards, resulting in a
-negative bias to the difference values.\footnote{This bias is not visible in
-Fig.~\ref{fig:steps3} because of the large vertical range, but can be seen in
-Fig.~\ref{fig:steps4} which has a smaller vertical scale.}
-
-Therefore to enhance the reliability of the step detection, such biases
-are removed before locating the steps. To do this, the differences are
-smoothed using a mean filter, and the smoothed differences are removed.
-The remaining residual difference values represent the gradient due to
-steps, spikes, noise, etc. It is important that the smoothed differences
-are not polluted by the large values caused by spikes and steps. So
-before smoothing, differences of more than three times the RMS difference
-(taken over the whole time stream) are identified and flagged. The RMS is
-then re-calculated omitting the flagged values and a further two such
-rejection iterations are performed. This process flags high gradient values
-caused by spikes, jumps, point sources, etc. The difference values close
-to a step are often not typical of those in the wider neighbourhood, so
-each block of contiguous flagged difference values is doubled in width.
-These flagged values are shown as black in Fig.~\ref{fig:steps4}, with
-the remaining unflagged values shown in red.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{steps4.pdf}
-\caption{The red curve shows the data from Fig.~\ref{fig:steps3} plotted with smaller upper
-and lower limits. The blue curve shows the results of smoothing the red
-curve. Points omitted from the red curve are shown in black.}
-\label{fig:steps4}
-\end{figure}
-
-\item The differences for the remaining samples are smoothed
-using a mean filter of width 50 samples. The smoothing process fills in small
-holes in the array (up to 40 samples), but will not fill in larger holes.
-Such larger gaps are filled in later by linear interpolation. The
-resulting smoothed differences (shown in blue in Fig.~\ref{fig:steps4})
-provide an estimate of the local gradient of the underlying noise-free
-median-filtered data stream.
-
-\item The smoothed differences are subtracted from the total differences
-to get the residual differences caused by noise, spikes, steps, etc.
-These are show in black in Fig.~\ref{fig:steps5}.
-
-\item The local RMS of the residual differences at each time sample is
-found by smoothing the squared residual differences using a median filter
-of width 200 samples, and then taking the square root of the smoothed values.
-These are show in green in Fig.~\ref{fig:steps5}.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{steps5.pdf}
-\caption{The black curve shows the residuals between the red and blue curves in
-Fig.~\ref{fig:steps4}. Missing data is filled in by a combination of
-smoothing and linear interpolation. The green curve is the local RMS of the data
-over a box width of 200 samples.}
-\label{fig:steps5}
-\end{figure}
-
-\item Blocks of contiguous samples for which the residual difference
-exceeds 25 times the local RMS of the residual differences are
-found. Any two such blocks that are separated by fewer than 100
-samples are merged into a single block. Blocks are only kept if they
-pass all of the following tests:
-
-\begin{enumerate}
-
-\item the width of the block must be less than 200 samples;
-
-\item the total of all the \snr\ values (the absolute residual difference
-divided by the local RMS of the residual differences) in the block must be
-at least 25;
-
-\item the total of all \snr\ values in the block must exceed half the value of
-the largest single \snr\ value in the block.
-
-\end{enumerate}
-
-The second and third tests are intended to prevent spikes being treated as
-two very close steps, since such steps will have \snr\ values of opposite
-signs and will thus cancel out.
-
-The data in Fig.~\ref{fig:steps5} results in two blocks being defined,
-one between samples 5037 and 5102, and another between samples 5286
-and 5296. Each of these blocks define the start and end of a candidate
-step.  The first block includes two visible steps, but they have been
-merged into a single candidate step due to their small separation
-(less than 100 samples).
-
-\end{enumerate}
-
-\subsection{Measuring and confirming candidate steps}
-
-%------------------------------------------------------------------------------
-
 \end{document}

### 2012-09-03 smurf_paper.tex
Remove the masking alternative to gap filling since it is not used by default

@@
 
-The evaluation of the \model{FLT} model (Section~\ref{sec:flt})
-involves taking the Fast Fourier Transform (FFT) of each bolometer
-time series. Data that have been flagged as bad for any reason (for
-instance, due to the presence of spikes, steps, or unusual common-mode
-signal) need to be excluded from this FFT. Two methods are available:
-
-\begin{itemize}
-\item Gap-filling: each contiguous block of bad data samples is filled
-with artificial data before taking the FFT. A least squares linear fit is
-performed to the 50 samples preceding the block, and a similar fit is
-performed to the 50 samples following the block. These are used to
-estimate the expected values at the start and end of the block of bad
-values. The bad values are then replaced by linear interpolation between
-the expected start and end values. Gaussian noise is added with a
-standard deviation equal to the mean of the RMS residuals in the two
-fits.
-
-\item Masking: each bad data value is replaced by zero, and an
-associated mask is created that holds 1.0 for each good sample in the
-time stream and 0.0 for each bad sample. The masked data and the mask
-are then each filtered in the same way using FFTs, and the filtered
-data are finally normalised by dividing them by the filtered mask.
-\end{itemize}
+The evaluation of the \model{FLT} model (Section~\ref{sec:flt}) involves
+taking the Fast Fourier Transform (FFT) of each bolometer time series.
+Data that have been flagged as bad for any reason (for instance, due to
+the presence of spikes, steps, or unusual common-mode signal) need to be
+excluded from this FFT. For this reason, each contiguous block of bad
+data samples is filled with artificial data before taking the FFT. A
+least squares linear fit is performed to the 50 samples preceding the
+block, and a similar fit is performed to the 50 samples following the
+block. These are used to estimate the expected values at the start and
+end of the block of bad values. The bad values are then replaced by
+linear interpolation between the expected start and end values. Gaussian
+noise is added with a standard deviation equal to the mean of the RMS
+residuals in the two fits.
 

### 2012-09-03 smurf_paper.tex
a partial commit of algo section

@@
 
-Given \scuba's data rate, SMURF was designed with two primary goals in
-mind: (i) reduced maps should be scientifically useful and of ``near
-publication quality'' with minimal user interaction; and (ii) the
-execution time should not appreciably exceed the amount of time
-required to conduct observations on a high-end multi-tcore desktop
-computer. By attaining these goals it would be possible to conduct
-real-time reductions at the telescope to offer feedback to observers,
-and it would also be possible to maintain a central repository of
-(nearly) publishable data products.
-
-As described in the introduction, rather than pursue a slow but
-potentially more accurate \emph{maximum likelihood} method that would
-require the inversion of a complex linear problem, we instead use an
-iterative approach that estimates and removes most of the correlated
-(and in particular low-frequency) noise sources in parallel with a
-simplified map estimation. We do note, however, that recent advances
-in direct inversion methods such as SANEPIC \citep{patanchon2008} may
-in fact be tenable for \scuba\ data. A future goal of the software
-team is to investigate meshing SMURF (perhaps as a cleaning step and
-initial condition) for a SANEPIC-like algorithm.
+As described in the introduction, the approach taken by SMURF to
+reducing \scuba\ data is to model (and remove) noise sources, and to
+iterate this process along with estimates of the map. This sort of
+algorithm has been used successfully in the past with instruments
+including SCUBA \citep{1999ASPC..172..171J}, SHARC-2
+\citep{kovacs2008}, and BOLOCAM \citep{aguirre2010}, and is capable of
+recovering modest angular scales up to of order the footprint of the
+array (depending, of course, on the noise in the data). The main
+benefit of this approach over CMB-style maximum-likelihood methods
+(which are generally superior at recovering structure on all scales
+available in the data) is that the computing resources are
+significantly smaller, both in terms of memory usage and execution
+time. Given the massive volume of \scuba\ data produced in a single
+observation (a rate of approximately 10 GB\,hr$^{-1}$ in each focal
+plane), iterative techniques were the only viable option for providing
+real-time feedback at the telescope, and for automatically reducing
+near-science grade maps on a timescale comensurate with the
+observation times using standard high-end desktop computers.  We do
+note, however, that recent advances in direct inversion methods such
+as SANEPIC \citep{patanchon2008} may in fact be tenable for \scuba\
+data. A future goal of the software team is to investigate meshing
+SMURF (perhaps as a cleaning step and initial condition) with a
+SANEPIC-like algorithm.
+
+Significant experimentation with different models for noise sources
+during \scuba\ commissioning required a highly flexible software
+framework with a simple (though extensive) interface for user
+configurability. Though written in C, \smurf\ uses an object-oriented
+style of handling the
+
+As we have seen in Sections~\ref{sec:bolosignal}--\ref{sec:pca}
 

### 2012-09-04 smurf_paper.tex
some more work on SMURF description

@@
 As described in the introduction, the approach taken by SMURF to
-reducing \scuba\ data is to model (and remove) noise sources, and to
-iterate this process along with estimates of the map. This sort of
+reducing \scuba\ data is to model (and remove) predominantly
+low-frequency noise sources that are correlated amongst detectors, and
+to iterate this process along with estimates of the map. This sort of
 algorithm has been used successfully in the past with instruments
@@
 during \scuba\ commissioning required a highly flexible software
-framework with a simple (though extensive) interface for user
-configurability. Though written in C, \smurf\ uses an object-oriented
-style of handling the
+framework with a simple interface for user configurability. To achieve
+this goal, while minimizing development time, it was decided to create
+SMURF as a Starlink package \citep{2009ASPC..411..418J}, which
+provides access to a large suite of libraries (including a commanding
+and messaging interface, astrometric conversions and generation of
+standard WCS information, file formats etc.). Furthermore, Starlink is
+already used extensively at the Joint Astronomy Centre (host of the
+JCMT) for many other systems, which helps with
+interoperability. Though originally written primarily in FORTRAN, many
+of the core Starlink libraries are now ported into native C, or else
+have, at least, a C interface. It was therefore decided to develop
+SMURF in C (rather than C++ which would have required adding further
+dependencies to Starlink), although the noise models are implemented
+with an object-oriented approach (all data associated with models are
+stored in structures, and there is a standard interface for method
+functions). In this way it is fairly easy to add new
+models. Parallelisation is incorporated in the most time-consuming
+low-level routines (e.g. performing Fast Fourier Transforms,
+re-gridding the data), usually handling either independent blocks of
+bolometers, or blocks of time, in each thread, depending on the nature
+of the calculation. ... something about scalability ...
 

### 2012-09-05 smurf_paper.tex
re-write of intro to section 3 done

--- 2012-09-03 smurf_paper.tex
@@
 %------------------------------------------------------------------------------
-\section{The SMURF Algorithm}
+\section{Production of maps}
 \label{sec:algorithm}
 %------------------------------------------------------------------------------
-
-As described in the introduction, the approach taken by SMURF to
-reducing \scuba\ data is to model (and remove) noise sources, and to
-iterate this process along with estimates of the map. This sort of
-algorithm has been used successfully in the past with instruments
-including SCUBA \citep{1999ASPC..172..171J}, SHARC-2
-\citep{kovacs2008}, and BOLOCAM \citep{aguirre2010}, and is capable of
-recovering modest angular scales up to of order the footprint of the
-array (depending, of course, on the noise in the data). The main
-benefit of this approach over CMB-style maximum-likelihood methods
-(which are generally superior at recovering structure on all scales
-available in the data) is that the computing resources are
-significantly smaller, both in terms of memory usage and execution
-time. Given the massive volume of \scuba\ data produced in a single
-observation (a rate of approximately 10 GB\,hr$^{-1}$ in each focal
-plane), iterative techniques were the only viable option for providing
-real-time feedback at the telescope, and for automatically reducing
-near-science grade maps on a timescale comensurate with the
-observation times using standard high-end desktop computers.  We do
-note, however, that recent advances in direct inversion methods such
-as SANEPIC \citep{patanchon2008} may in fact be tenable for \scuba\
-data. A future goal of the software team is to investigate meshing
-SMURF (perhaps as a cleaning step and initial condition) with a
-SANEPIC-like algorithm.
-
-Significant experimentation with different models for noise sources
-during \scuba\ commissioning required a highly flexible software
-framework with a simple (though extensive) interface for user
-configurability. Though written in C, \smurf\ uses an object-oriented
-style of handling the
-
-As we have seen in Sections~\ref{sec:bolosignal}--\ref{sec:pca}
-
-In practice, the model generally adopted for low-frequency noise
-\scuba\ data is a single common-mode (average signal across all
-working bolometers in a given focal plane), with most remaining noise
-sources being removed by a high-pass filter. Many variations on this
-were attempted throughout the commissioning phase, but as the PCA in
-Secion~\ref{sec:pca} illustrates, less significant noise sources are
-both transient and particularly difficult to model.
-
-This section gives an overview of the SMURF algorithm, including data
-cleaning and other pre-processing steps (Section~\ref{sec:dataprep}),
-the iterative estimation of the map and noise model components
-(Section~\ref{sec:components}), and concludes with a discussion of
-convergence (Section~\ref{sec:converge}).
-
-\begin{itemize}
-
-\item Discuss SMURF configurability
-
-\item Some sort of proof/demonstration showing why the algorithm works
-
-\item Discuss performance (execution time, memory, disk usage etc.)
-
-\item Where do you get SMURF/starlink, usage of NDF
-
-\item Talk about pipeline?
-
-\end{itemize}
-
-We express the signal observed by the $i$th bolometer as a function of
-time,
-%
-\begin{equation}
-\mathbf{b}_i(t) = g_i[\mathbf{e}_i(t) \mathbf{a}_i(t) + \mathbf{n}_i(t)],
-\end{equation}
-%
-where $\mathbf{a}(t)$ is the time-varying signal produced by scanning
-the telescope across astronomical sources, $\mathbf{e}(t)$ is the
-time-varying extinction, which is a function of the telescope
-elevation and atmospheric conditions, and $\mathbf{n}_i(t)$ represents
-sources of noise. The two terms in square brackets, as written, have
-units of power delivered to the detectors (pW). The scale factor $g_i$
-converts this effective power to the digitized units recorded on disk
-(DAC) -- the \emph{flatfield} multiplied by a digitization constant --
-and in this formulation it is assumed to be constant in time.
-
-We then express the noise, $\mathbf{n}_i(t)$, as the sum of several
-components,
-%
-\begin{equation}
-  \mathbf{n}_i(t) = \mathbf{n}^\mathrm{w}_i(t) + \mathbf{n}^\mathrm{f}_i(t) +
-  \mathbf{n}^\mathrm{c}(t),
-\label{eq:noise}
-\end{equation}
-%
-where $\mathbf{n}^\mathrm{w}_i(t)$ is uncorrelated white (thermal)
-noise, $\mathbf{n}^\mathrm{f}_i(t)$ is low-frequency noise that is
-\emph{not} correlated from bolometer-to-bolometer, and
-$\mathbf{n}^\mathrm{c}(t)$ is a (predominantly low-frequency)
-correlated or \emph{common-mode} component. The primary goal of
-map-making is to model and remove $\mathbf{n}^\mathrm{f}_i$ and
-$\mathbf{n}^\mathrm{c}$ from the time-series, and then by re-gridding
-the remaining data we can hope to approach the theoretical noise limit
-in the map. This procedure is summarized in Fig.~\ref{fig:dimm}.
 
@@
 \end{figure}
+
+As described in the introduction, the approach taken by SMURF to
+reducing \scuba\ data is to model (and remove) predominantly
+low-frequency noise sources that are correlated amongst detectors, and
+to iterate this process along with estimates of the map. This sort of
+algorithm has been used successfully in the past with instruments
+including SCUBA \citep{1999ASPC..172..171J}, SHARC-2
+\citep{kovacs2008}, and BOLOCAM \citep{aguirre2010}, and is capable of
+recovering modest angular scales up to of order the footprint of the
+array (depending, of course, on the noise in the data). The main
+benefit of this approach over CMB-style maximum-likelihood methods
+(which are generally superior at recovering structure on all scales
+available in the data) is that the computing resources are
+significantly smaller, both in terms of memory usage and execution
+time. Given the massive volume of \scuba\ data (produced at a rate of
+approximately 10 GB\,hr$^{-1}$ in each focal plane), iterative
+techniques were the only viable option for providing real-time
+feedback at the telescope, and for automatically reducing near-science
+grade maps on a timescale comensurate with the observation times,
+using single high-end desktop computers.  We do note, however, that
+recent advances in direct inversion methods such as SANEPIC
+\citep{patanchon2008} may in fact be tenable for \scuba\ data. A
+future goal of the software team is to investigate meshing SMURF
+(perhaps as a cleaning step and initial condition) with a SANEPIC-like
+algorithm.
+
+Significant experimentation with different models for noise sources
+during \scuba\ commissioning required a highly flexible software
+framework, and a configurable user interface. To achieve this goal,
+while minimizing development time, it was decided to build SMURF as a
+Starlink package \citep{2009ASPC..411..418J}, which provides access to
+a large suite of libraries (including a commanding and messaging
+interface, astrometric coordinate conversions and generation of
+standard WCS information, file formats etc.). Furthermore, Starlink is
+already used extensively at the Joint Astronomy Centre (host of the
+JCMT) for many other systems, which helps with
+interoperability. Though originally written primarily in FORTRAN, many
+of the core Starlink libraries are now ported into native C, or else
+have a C interface. It was therefore decided to develop SMURF in C
+(rather than C++ which would have required adding further dependencies
+to Starlink), although we have taken an object-oriented approach. For
+example, all of the data for a given noise model is stored in a C
+structure, and there is a standard interface for all functions that
+handle models (member data and functions for the class,
+respectively). In this way it is easy to extend SMURF with new
+models. Parallelisation is incorporated in the most time-consuming
+low-level routines using threads (e.g. performing Fast Fourier
+Transforms, re-gridding the data), usually handling either independent
+blocks of bolometers, or blocks of time, in each thread, depending on
+the nature of the calculation. \textbf{... something about scalability
+...}
+
+\begin{itemize}
+
+\item Discuss SMURF configurability
+
+\item Some sort of proof/demonstration showing why the algorithm works
+
+\item Discuss performance (execution time, memory, disk usage etc.)
+
+\item Where do you get SMURF/starlink, usage of NDF
+
+\item Talk about pipeline?
+
+\end{itemize}
+
+As we have seen in Sections~\ref{sec:bolosignal}--\ref{sec:pca},
+\scuba\ data are dominated at low frequencies ($\lsim$2\,Hz) by highly
+correlated signals. In addition to this $1/f$ noise, there is also
+uncorrelated white noise, and finally residual correlated (but
+non-astronomical) signals that appear at all frequencies with more
+complicated relationships amongst groups of bolometers
+(e.g. Fig.~\ref{fig:pca}). While the low-frequency noise can be
+significantly reduced using simple common-mode removal (subtracting
+the average signal from all bolometers at each time step), these
+remaining correlated noise patterns are much more difficult to
+model. Ultimately, a very simple approach has been taken to handling
+these two contaminating sources, for which the main steps in a typical
+reduction are shown in Fig.~\ref{fig:dimm}.
+
+First, the raw data are passed through a pre-processing stage which
+corrects some of the more significant glitches, applies flat-field
+corrections etc. Next, the iterative process begins (dashed box). Most
+of the low-frequency noise is removed using common-mode
+subtraction. Next, the extinction correction is applied (it is done at
+this stage rather than during pre-processing to avoid multiplying
+large atmospheric signals by any small errors in the flatfield
+correction). At this stage the data resemble the grey bolometer traces
+in Fig.~\ref{fig:bolos_mix}, with some residual baseline drifts still
+visible. These drifts are removed using a high-pass filter implemented
+with Fast Fourier Transforms (FFTs). The residual signal is then
+re-gridded to estimate the map. Finally, the map is used to estimate
+and remove the astronomical signal contamination in the bolometer
+data, leaving a relatively clean data set in which to measure the
+white noise level of each bolometer. Since the common-mode and
+filtering stages will have introduced ringing in the map in the
+vicinity of bright sources, the entire process is iterated. Each
+component in the dashed box of Fig.~\ref{fig:dimm} is re-calculated in
+sequence. In this way, the second time the common-mode is calculated,
+for example, most of the bright astronomical sources in the data have
+already been removed in the previous iteration when the map was
+estimated and subtracted, reducing the amount of ringing in the map.
+
+In Sections~\ref{sec:dataprep} and \ref{sec:components} we describe
+the data pre-processing steps and iterative algorithm in
+detail. Section~\ref{sec:converge} explores convergence tests and
+degeneracies between model components.
+
+We express the signal observed by the $i$th bolometer as a function of
+time,
+%
+\begin{equation}
+\mathbf{b}_i(t) = g_i[\mathbf{e}_i(t) \mathbf{a}_i(t) + \mathbf{n}_i(t)],
+\end{equation}
+%
+where $\mathbf{a}(t)$ is the time-varying signal produced by scanning
+the telescope across astronomical sources, $\mathbf{e}(t)$ is the
+time-varying extinction, which is a function of the telescope
+elevation and atmospheric conditions, and $\mathbf{n}_i(t)$ represents
+sources of noise. The two terms in square brackets, as written, have
+units of power delivered to the detectors (pW). The scale factor $g_i$
+converts this effective power to the digitized units recorded on disk
+(DAC) -- the \emph{flatfield} multiplied by a digitization constant --
+and in this formulation it is assumed to be constant in time.
+
+We then express the noise, $\mathbf{n}_i(t)$, as the sum of several
+components,
+%
+\begin{equation}
+  \mathbf{n}_i(t) = \mathbf{n}^\mathrm{w}_i(t) + \mathbf{n}^\mathrm{f}_i(t) +
+  \mathbf{n}^\mathrm{c}(t),
+\label{eq:noise}
+\end{equation}
+%
+where $\mathbf{n}^\mathrm{w}_i(t)$ is uncorrelated white (thermal)
+noise, $\mathbf{n}^\mathrm{f}_i(t)$ is low-frequency noise that is
+\emph{not} correlated from bolometer-to-bolometer, and
+$\mathbf{n}^\mathrm{c}(t)$ is a (predominantly low-frequency)
+correlated or \emph{common-mode} component. The primary goal of
+map-making is to model and remove $\mathbf{n}^\mathrm{f}_i$ and
+$\mathbf{n}^\mathrm{c}$ from the time-series, and then by re-gridding
+the remaining data we can hope to approach the theoretical noise limit
+in the map. This procedure is summarized in Fig.~\ref{fig:dimm}.
 
@@
 %-------------------------------------------------
-\subsection{Model components}
+\subsection{Iterative model calculation}
 \label{sec:components}
@@
 %-------------------------------------------------
-\subsection{Data preparation}
+\subsection{Data pre-processing}
 \label{sec:dataprep}

### 2012-09-05 smurf_paper.tex
move pre-processing section before model components
section 3 on map-maker workings mostly done

@@
 JCMT) for many other systems, which helps with
-interoperability. Though originally written primarily in FORTRAN, many
-of the core Starlink libraries are now ported into native C, or else
-have a C interface. It was therefore decided to develop SMURF in C
-(rather than C++ which would have required adding further dependencies
-to Starlink), although we have taken an object-oriented approach. For
-example, all of the data for a given noise model is stored in a C
-structure, and there is a standard interface for all functions that
-handle models (member data and functions for the class,
-respectively). In this way it is easy to extend SMURF with new
-models. Parallelisation is incorporated in the most time-consuming
-low-level routines using threads (e.g. performing Fast Fourier
-Transforms, re-gridding the data), usually handling either independent
-blocks of bolometers, or blocks of time, in each thread, depending on
-the nature of the calculation. \textbf{... something about scalability
-...}
+interoperability. Though originally written in FORTRAN, many of the
+core Starlink libraries are now ported to native C, or at least have a
+C interface. It was therefore decided to develop SMURF in C as well
+(rather than C++, for example, which would have required adding
+further dependencies to Starlink), although we have taken an
+object-oriented approach. For example, all of the data for a given
+noise model is encapsulated in a C structure, and there is a standard
+interface for all functions that handle models (member data and
+functions for the class, respectively). In this way it is easy to
+extend SMURF with new models. Parallelisation is incorporated in the
+most time-consuming low-level routines using threads (e.g. performing
+Fast Fourier Transforms, re-gridding the data), usually handling
+either independent blocks of bolometers, or blocks of time, in each
+thread, depending on the nature of the
+calculation. \textbf{... something about scalability ...}
 
@@
 
+%-------------------------------------------------
+\subsection{Data pre-processing}
+\label{sec:dataprep}
+%-------------------------------------------------
+
+Prior to executing the iterative part of the SMURF algorithm, the data
+must undergo several pre-processing steps. First, the data files are
+read into memory and concatenated into continuous arrays (\scuba\ data
+are broken up and written to disk every $\sim$30\,s). As the data are
+loaded, the flatfield correction is also applied \citep[see
+Section~2.1 in][]{dempsey2012}. Next, a series of data cleaning and
+filtering procedures are applied; these include the removal of large
+glitches that may hinder the iterative solution from converging, or
+simply tasks that do not need to be iterated.
+
+\subsubsection{Time-series down-sampling and map pixel size}
+\label{sec:downsamp}
+
+The highest useful frequency in the nominally 200\,Hz-sampled \scuba\
+data is that which corresponds to the smallest angular scale that the
+instrument is sensitive to. As mentioned earlier, the usual
+rule-of-thumb for a Gaussian beam is to provide at least 3 samples for
+each FWHM, or roughly 2.5\,arcsec for the 7.5\,arcsec 450\,\micron\
+channel, and 5\,arcsec for the 14.5\,arcsec 850\,\micron\ channel. For
+a typical scan speed of 300\,arcsec\,s$^{-1}$, the maximum useful
+sample rate is therefore about 120\,Hz at 450\,\micron, and 40\,Hz at
+850\,\micron. In order to save execution time and memory usage (both
+of which scale linearly with data volume), it is clearly advantageous
+to re-sample the data to these lower rates. In practice, the default
+map pixel sizes are set to 2 and 4\,arcsec at 450 and 850\,\micron,
+respectively (slightly over-sampled), and down-sampling occurs as the
+data are loaded to match this spatial sampling given the slew speed.
+
+The method used by SMURF to down-sample is to average together
+multiple samples from the original time-series, $x_i$ to estimate the
+lower-frequency output time-series, $y_j$. In general there will not
+be an integer number of samples from $x_i$ in each of the $y_j$, so
+fractional samples from $x_i$ are used at the $y_j$ time
+boundaries. The algorithm is fast, since each sample in $x_i$ need
+only be visited once. From a spectral point of view, this boxcar
+average is equivalent to applying a sinc-function low-pass filter in
+the frequency domain. Such behaviour is desirable since it serves as
+an anti-aliasing filter; most non-white features above the Nyquist
+frequency in $y_j$ that could contaminate the output data are largely
+removed.
+
+As an alternative, we also investigated an algorithm in which the FFT
+of $x_i$ is simply truncated to the target sample rate before
+transforming back to the time domain (i.e., applying a hard-edged
+low-pass filter). In practice the noise performance was
+indistinguishable, and required slightly more execution time using
+this method. It is therefore not currently available to users.
+
+\subsubsection{Bolometer filtering}
+
+Despite the ability of the map-maker to iteratively remove many noise
+components, under some circumstances it may be desirable to filter the
+data \emph{once} during the pre-processing step. Three main filtering
+options are available:
+
+\begin{itemize}
+
+\item The most commonly-used pre-processing filter is polynomial
+subtraction. A polynomial of the requested order is fit and removed
+from each bolometer time-series. At a bare minimum, the mean is
+removed from all of the bolometers (order 0) in all reductions
+described in this paper.
+
+\item All of the filters available as part of the iterative Fourier
+Transform Filter (Section~\ref{sec:flt}) can also be applied once
+during pre-processing.
+
+\item As an alternative to the iterative map-making procedure,
+cleaning by principal component analysis (PCA) is available as an
+experimental pre-processing option. As described in
+Section~\ref{sec:pca}, a new set of basic vectors for the bolometer
+data are found which diagonalize their covariance matrix. The
+brightest components can then be subtracted from the data. A single
+parameter specifies the threshold on the amplitude of the eigenvalues
+to be removed, as a number of standard deviations away from the mean
+value. Subarrays are cleaned independently once for the full length of
+each continuousy chunk of data. Given the computational expense of
+this method, and initial tests which showed little improvement over
+simple high-pass filtering, this method has not yet been explored in
+detail with \scuba\ data. However, since systematic effects seem to
+come and go, it is possible that PCA could be useful for particular
+data sets.
+
+\end{itemize}
+
+\subsubsection{Time-domain de-spiking}
+\label{sec:timedespike}
+
+Spikes of short duration and high amplitude are often seen in the time
+series data. If not removed, they can cause ringing when filtering the
+data. Two alternative approaches may be used to remove these
+spikes. This section describes the detection and removal of spikes
+within the time series of each bolometer, and Section~\ref{sec:ast}
+describes the iterative detection and removal of spikes as part of map
+estimation. In practice, map-based despiking usually gives superior
+results, and so time-domain despiking is switched off by default.
+
+Each one-dimensional bolometer time series is processed
+independently. At each time slice, the median value of the current
+bolometer is found in a box centred on the time slice, and containing
+a specified number of time slices (typically 50). If the residual
+between the time slice value and the median value is greater than some
+specified multiple (typically 10) of the local noise level, the time
+slice is flagged as a spike.
+
+If the local noise level were estimated within the same box used to
+determine the median value, a spike in the box would cause the local
+noise level to be over-estimated severely. For this reason, the local
+noise level is taken as the standard deviation of the values within a
+neighbouring box on the ``down-stream'' side of the median box (that
+is, the side that has already been checked for spikes). In other
+words, the high end of the noise box is just below the low end of the
+median filter box. This introduces a slight asymmetry in the noise,
+but this should not matter unless the noise varies significantly on a
+time scale shorter than the box size.
+
+This simple algorithm is not very good at distinguishing between spikes
+and bright point sources, and so the threshold for spike detection is
+usually raised when making maps of bright point sources.
+
+\subsubsection{Step correction}
+\label{sec:steps}
+
+\begin{figure*}
+\centering
+\includegraphics[width=\linewidth]{steps1.pdf}
+\caption{Examples of steps in time series data. Steps occur with a wide
+range of heights from the large steps shown on the left to the small
+steps shown on the right. Large steps are often followed by a brief
+``over-shoot'' as shown on the left. In both plots, the black curve is
+the uncorrected time series, and the red curve is the corrected time
+series. Samples close to a step are omitted in the corrected time series.
+In the right hand plot, the blue curve is the uncorrected time series for
+a nearby bolometer. The similarity between the red and blue curves shows
+that the step correction is performing well.
+}
+\label{fig:steps1}
+\end{figure*}
+
+Sudden steps can occur in the time stream data from each bolometer,
+with the most likely cause being cosmic ray events \citep[see
+Section~3.5.3 in][]{holland2012}. The black curves in
+Fig.~\ref{fig:steps1} show examples of such steps in the time streams
+for two bolometers. If not removed, these steps can cause severe
+ringing in the FLT model (Section~\ref{sec:flt}) and visible streaks
+in the final map, corresponding to the paths of individual bolometers
+over the sky.
+
+Steps occur with a wide range of heights and shapes. The ratio of step
+height to noise can vary from less than 10 to several hundred. Some
+steps occur over a single sample, such as the step close to sample
+5000 in the right-hand plot of Fig.~\ref{fig:steps1}, but others
+happen more gradually, such as the step close to sample 5300. In
+addition, a step can be preceded or followed by a short period of
+instability, as is visible at the bottom of the step in the left hand
+plot of Fig.~\ref{fig:steps1}. Further problems are caused by steps
+that occur close together in time, such as the large downward step
+followed by a smaller upward step close to sample 5000 in the right
+hand plot of Fig.~\ref{fig:steps1}.
+
+Detecting and correcting such a wide variety of steps reliably has proved
+to be a challenge. In outline, the following stages
+are involved in detecting steps in a single bolometer time stream:
+
+\begin{enumerate}
+
+\item median smooth the whole time stream;
+
+\item find the gradient of the median smoothed time series at each
+sample;
+
+\item smooth the gradient values to determine the local mean gradient
+and subtract this local mean from the total gradient to get residual
+gradient;
+
+\item find residual gradient values that exceed 25 times the local RMS
+of the residual gradients;
+
+\item Group these high residual gradients into contiguous blocks of
+samples;
+
+\item merge blocks that are separated by less than 100 samples.
+\end{enumerate}
+
+The above process produces a list of candidate steps in each bolometer
+time series. Each candidate step is then verified, measured and corrected
+using the following procedure:
+
+\begin{enumerate}
+
+\item the above process can mis-interpret the edges of a bright source
+as a step, so we ignore blocks that occur close to bright sources;
+
+\item if the block passes the above test, a least squares linear fit
+is performed to the median-filtered bolometer data just before the
+block, and this fit is extrapolated to predict the data value expected
+at the centre of the block on the basis of the preceding data;
+
+\item a least squares linear fit is performed to the median-filtered
+bolometer data just after the block, and this fit is extrapolated to
+predict the data value expected at the centre of the block on the
+basis of the following data;
+
+\item the difference between these two expected data values is taken
+as the step height;
+
+\item the preceding three steps are repeated several times, each time
+including a different selection of samples in the two least squares
+fits, with the mean and standard deviation of the corresponding set of
+step heights found;
+
+\item if the mean step height is small compared to the standard deviation
+of the step heights, or compared to the noise in the bolometer data, then
+the step is ignored;
+
+\item if the above checks are passed, all subsequent bolometer samples
+are corrected by the mean step height;
+
+\item bolometer samples within the duration of the step, and a few
+samples on either side, are flagged as unusable.
+
+\end{enumerate}
+
+Once all steps have been corrected within a bolometer time stream, a
+constant value is added to all samples in the time stream to restore its
+original mean value.
+
+The results of step correction are shown by the red curves in
+Fig.~\ref{fig:steps1}. For comparison, the blue curve shows the
+uncorrected time stream from a nearby bolometer that does not suffer from
+steps. The agreement between the red and blue curves confirms that
+the step correction algorithm is working satisfactorily.
+
+\subsubsection{Gap filling / apodization}
+\label{sec:gaps}
+
+SMURF uses FFTs of the bolometer data extensively for filtering.  Data
+that have been flagged as bad for any reason (for instance, due to the
+presence of spikes, steps, or unusual common-mode signal) need to be
+excluded from this FFT. For this reason, each contiguous block of bad
+data samples is filled with artificial data before taking the FFT. A
+least squares linear fit is performed to the 50 samples preceding the
+block, and a similar fit is performed to the 50 samples following the
+block. These are used to estimate the expected values at the start and
+end of the block of bad values. The bad values are then replaced by
+linear interpolation between the expected start and end
+values. Gaussian noise is added with a standard deviation equal to the
+mean of the RMS residuals in the two fits.
+
+In addition to replacing bad samples before the FFT, it is also
+necessary to ensure that the data values at the start and end of the
+time series are similar. Since an FFT treats the data as a single
+cycle in an infinitely repeating waveform, any large difference
+between starting and ending value will effectively introduce sudden
+steps at the start and end of each cycle, causing unwanted
+oscillations (ringing) in the transform. Another consequence of the
+cyclic nature of the FFT is that features at one end of the time
+series can affect the filtered values at the other end of the time
+series. Two methods are available to avoid these two problems:
+
+\begin{itemize}
+\item Apodisation: a number of samples at the start and end of each
+bolometer time series are multiplied by a cosine function in order to
+roll the data values off smoothly to zero. The default number of
+samples modified at each end of the time series is given by half the
+ratio of the sampling frequency to the lowest retained frequency. In
+addition, each end of the time stream is padded with double this
+number of zeros. This method is illustrated in
+Fig.~\ref{fig:pad2}. This method is not used by default as it reduces
+the amount of data available for the map, and can significantly hinder
+very short observations (e.g. of calibrators when focussing the
+telescope).
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{pad2.pdf}
+\caption{The black curve shows a bolometer time stream, padded
+with zeros. The red curve shows the time stream after apodisation.}
+\label{fig:pad2}
+\end{figure}
+
+\item Padding with artificial data: instead of padding with zeros,
+each time stream is padded with artificial data that connects the two
+ends of the data stream smoothly and includes Gaussian noise. No
+apodisation is performed. The number of samples of padding at each end
+is again equal to the ratio of the sampling frequency to the lowest
+retained frequency.  This is illustrated in Fig.~\ref{fig:pad1}. See
+\citet{stompor2002} for a thorough discussion of this procedure within
+the context of CMB map-making. This method is used by default.
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{pad1.pdf}
+\caption{The black curve shows the same bolometer time stream as in
+Fig.~\ref{fig:pad2}, againpadded with zeros. The red curve shows the
+time stream after padding with artificial data.}
+\label{fig:pad1}
+\end{figure}
+
+\end{itemize}
+
+\subsubsection{Additional bad bolometer rejection}
+
+Despite all of the cleaning operations described in the previous
+sections, the data from a given bolometer may be unusable do to it
+being poorly biased, having an incorrect flatfield correction applied,
+or having some other particularly pathological noise
+contamination. Most of these bolometers can be flagged simply by
+identifying outliers in the bolometer noise distribution. By default,
+SMURF measures the PSD of all bolometers between 2 and 10\,Hz, after
+all other pre-processing steps have been run (identical to the
+measurement in Section~\ref{sec:noi}). Despite having significant
+low-frequency noise (and possibly bright astronomical source)
+contamination, this higher-frequency portion of the PSD is generally
+quite clean. Furthermore, even if there is contamination, the purpose
+of this measurement is to identify outliers, rather than provide a
+meaningful absolute measurement of a given bolometer's white noise
+level. Both high and low (e.g., due to an incorrect and very small
+flatfield correction being applied) outliers from the centre of the
+distribution are flagged. The measurement is based on the logarithm of
+the noise values, and iterative clipping of the mean and standard
+deviation are used, to reduce the impact of large outliers.
+
+%-------------------------------------------------
+\subsection{Iterative model calculation}
+\label{sec:components}
+%-------------------------------------------------
+
 We express the signal observed by the $i$th bolometer as a function of
@@
 
-First, the data files are read from disk, concatenated into contiguous
-arrays in memory, and run through a pre-processing stage that performs
+ and run through a pre-processing stage that performs
 several operations including spike and sudden level step removal to
@@
 
-%-------------------------------------------------
-\subsection{Iterative model calculation}
-\label{sec:components}
-%-------------------------------------------------
-
 In addition to specifying when each model component is computed
@@
 Map estimation is accomplished using a nearest-neighbour resampling of
-the data onto a pre-defined map grid. For the $i$th map pixel
+the data onto a pre-defined map grid with. For the $i$th map pixel
 $\mathbf{m}(x_i,y_i)$, the brightness is estimated as the weighted
@@
 In addition to map estimation, the \model{AST} model can also be used
-to perform map-based despiking of the time-series
-(Section~\ref{sec:mapdespike}), and to apply constraints to the map to
-improve convergence (Section~\ref{sec:converge}).
+to perform map-based despiking of the time-series. Unlike the
+time-domain despiker (Section~\ref{sec:timedespike}), this calculation
+uses the scatter in the population of samples used to estimate the map
+pixel values (from different times and bolometers) to reject
+outliers. This method is more robust against false-positive detections
+of bright/compact astronomical sources, since transient features in
+the time-series are unlikely to occur by chance whenever a bolometer
+crosses a specific location on the sky, whereas real astronomical
+sources are at fixed spatial locations.
+
+The estimated variance, $\mathbf{v_p}(x_i,y_i)$, of the normalized
+weighted samples that land in the $i$th map pixel is simply the biased
+weighted sample variance (i.e., the variance map value multiplied by
+the number of samples)
+%
+\begin{equation}
+  \mathbf{v_p}(x_i,y_i) = N_i \mathbf{v}(x_i,y_i).
+\end{equation}
+%
+In order to compare the weighted \emph{differences} between the
+samples and the map values, [$\mathbf{b}_j - \mathbf{m}(x_i,y_i)$] to
+$\mathbf{v_p}$, they must be scaled appropriately. We define a
+normalized difference, $\mathbf{d}_j$, in such a way that the variance
+of this new variable gives the weighted sample variance of the
+underlying data points:
+%
+\begin{eqnarray}
+  \frac{\sum_j \mathbf{d}_j^2}{N} &=&
+  \frac{\sum_j \mathbf{w}_j [\mathbf{b}_j - \mathbf{m}(x_i,y_i)]^2}
+       {\sum_k \mathbf{w}_k} \\
+   \Rightarrow \mathbf{d}_j^2 &=& \frac{N \mathbf{w}_j [\mathbf{b}_j -
+       \mathbf{m}(x_i,y_i)]^2}{ \sum_k \mathbf{w}_k} .
+\end{eqnarray}
+%
+The map-based despiker flags those $\mathbf{d}_j$ that are further
+than some threshold number of standard deviations,
+$\sqrt{\mathbf{v_p}}$, away from zero so that they are not used in
+subsequent iterations.
+
+A final option available as part of the \model{AST} model is to apply
+constraints to the map to improve convergence. See the discussion in
+Section~\ref{sec:converge}.
 
@@
 
+%-------------------------------------------------
 \subsection{Convergence tests and model degeneracies}
 \label{sec:converge}
+%-------------------------------------------------
 
@@
 even at low-frequencies, do not leak into \model{FLT}.
-
-%-------------------------------------------------
-\subsection{Data pre-processing}
-\label{sec:dataprep}
-%-------------------------------------------------
-
-In this section we describe a number of data preparation steps that
-preceed the calculation of model components as described in the
-previous section.
-
-\subsubsection{Time-series down-sampling and map pixel size}
-\label{sec:downsamp}
-
-The highest useful frequency in the nominally 200\,Hz-sampled
-\scuba\ data is that which corresponds to the smallest angular scale
-that the instrument is sensitive to. As mentioned earlier, the usual
-rule-of-thumb for a Gaussian beam is to provide at least 3 samples for
-each FWHM, or roughly 2.5\,arcsec for the 7.5\,arcsec
-450\,\micron\ channel, and 5\,arcsec for the 14.5\,arcsec
-850\,\micron\ channel. For a typical scan speed of
-300\,arcsec\,s$^{-1}$, the maximum useful sample rate is therefore
-about 120\,Hz at 450\,\micron, and 40\,Hz at 850\,\micron. In order to
-save execution time and memory usage (both of which scale linearly
-with data volume), it is clearly advantageous to re-sample the data to
-these lower rates.
-
-Similarly, from the point of view of retaining all of the useful
-information, the size of the output map pixels need not be any smaller
-than the FWHM/3 rule. Larger pixels may be selected, but at the
-expense of retaining information at the diffraction limit of the
-telescope.  \textbf{ Per: loss of information?}.
-
-The default behaviour of SMURF is to select output map pixel sizes of
-2 and 4\,arcsec at 450 and 850\,\micron, respectively (slightly
-over-sampled). The slew speed of the telescope (approximately constant
-for all standard scan modes) is then used to calculate the highest
-useful frequency for the time-series given the map pixel size,
-$f_\mathrm{max}$. The data are then re-sampled to this new rate prior
-to map-making (as well as the data cleaning steps).
-
-The method used by SMURF to down-sample simply averages together
-multiple samples from the original time-series, $x_i$ to estimate the
-lower-frequency output time-series, $y_j$. In general there will not
-be an integer number of samples from $x_i$ in each of the $y_j$, so
-fractional samples from $x_i$ are used at the $y_j$ time
-boundaries. The algorithm is fast, since each sample in $x_i$ need
-only be visited once. From a spectral point of view, this boxcar
-average is equivalent to applying a sinc-function low-pass filter in
-frequency space. Such behaviour is desirable, since it serves as an
-anti-aliasing filter; most non-white features above the Nyquist
-frequency in $y_j$ that could contaminate the output data are largely
-removed.
-
-As an alternative, we also investigated an algorithm in which the FFT
-of $x_i$ is simply truncated to the target sample rate before
-transforming back to the time domain (i.e., applying a hard-edged
-low-pass filter). In practice the noise performance was
-indistinguishable, and required slightly more execution time, using
-this method compared to the boxcar method described above. It is
-therefore not currently available to users.
-
-\subsubsection{Bolometer filtering}
-
-Despite the ability of the map-maker to iteratively remove many noise
-components, under some circumstances it may be desirable to filter the
-data \emph{once} during the pre-processing step. Three main filtering
-options are available:
-
-\begin{itemize}
-
-\item The most commonly-used pre-processing filter is polynomial
-subtraction. A polynomial of the requested order is fit and removed
-from each bolometer time-series. At a bare minimum, the mean is
-removed from all of the bolometers (order 0) in all reductions
-described in this paper.
-
-\item All of the filters available as part of the iterative Fourier
-Transform Filter (Section~\ref{sec:flt}) can also be applied once
-during pre-processing.
-
-\item As a complete alternative to the iterative map-making procedure,
-cleaning by principal component analysis (PCA) is available as an
-experimental pre-processing option. As described in
-Section~\ref{sec:pca}, a new set of basic vectors for the bolometer
-data are found which diagonalize their covariance matrix. The
-brightest components can then be subtracted from the data. A single
-parameter specifies the threshold on the amplitude of the eigenvalues
-to be removed, as a number of standard deviations away from the mean
-value. Subarrays are cleaned independently once for the full length of
-each continuousy chunk of data. Given the computational expense of
-this method, and initial tests which showed little improvement over
-simple high-pass filtering, this method has not yet been explored in
-detail with \scuba\ data. However, since systematic effects seem to
-come and go, it is possible that PCA could be useful for particular
-data sets.
-
-\end{itemize}
-
-\subsubsection{Time-domain de-spiking}
-\label{sec:timedespike}
-
-Spikes of short duration and high amplitude are often seen in the time
-series data. If not removed, they can cause ringing in the FLT model
-(Section~\ref{sec:flt}). Two alternative approaches may be used to
-remove these spikes. This section describes the detection and removal
-of spikes within the time series of each bolometer, and
-section~\ref{sec:mapdespike} describes the detection and removal of
-spikes within the final map. In practice, map-based despiking
-(Section~\ref{sec:mapdespike}) usually gives superior results, and so
-time-domain despiking is usually switched off.
-
-Each one-dimensional bolometer time series is processed
-independently. At each time slice, the median value of the current
-bolometer is found in a box centred on the time slice, and containing
-a specified number of time slices (typically 50). If the residual
-between the time slice value and the median value is greater than some
-specified multiple (typically 10) of the local noise level, the time
-slice is flagged as a spike.
-
-If the local noise level were estimated within the same box used to
-determine the median value, a spike in the box would cause the local
-noise level to be over-estimated severely. For this reason, the local
-noise level is taken as the standard deviation of the values within a
-neighbouring box on the ``down-stream'' side of the median box (that
-is, the side that has already been checked for spikes). In other
-words, the high end of the noise box is just below the low end of the
-median filter box. This introduces a slight asymmetry in the noise,
-but this should not matter unless the noise varies significantly on a
-time scale shorter than the box size.
-
-This simple algorithm is not very good at distinguishing between spikes
-and bright point sources, and so the threshold for spike detection is
-usually raised when making maps of bright point sources.
-
-\subsubsection{Step correction}
-\label{sec:steps}
-
-\begin{figure*}
-\centering
-\includegraphics[width=\linewidth]{steps1.pdf}
-\caption{Examples of steps in time series data. Steps occur with a wide
-range of heights from the large steps shown on the left to the small
-steps shown on the right. Large steps are often followed by a brief
-``over-shoot'' as shown on the left. In both plots, the black curve is
-the uncorrected time series, and the red curve is the corrected time
-series. Samples close to a step are omitted in the corrected time series.
-In the right hand plot, the blue curve is the uncorrected time series for
-a nearby bolometer. The similarity between the red and blue curves shows
-that the step correction is performing well.
-}
-\label{fig:steps1}
-\end{figure*}
-
-Sudden steps can occur in the time stream data from each bolometer,
-with the most likely cause being wraparound in the first stage squid
-amplification, the details of which are given in \citet{holland2012}
-\textbf{Per gave a technical explanation, but this is probably best in
-another paper?}.
-
-The black curves in Fig.~\ref{fig:steps1} show examples of such steps in
-the time streams for two bolometers. If not removed, these steps can
-cause severe ringing in the FLT model (Section~\ref{sec:flt}) and visible
-streaks in the final map, corresponding to the paths of individual
-bolometers over the sky.
-
-Steps occur with a wide range of heights and shapes. The ratio of step
-height to noise can vary from less than 10 to several hundred. Some
-steps occur over a single sample, such as the step close to sample
-5000 in the right-hand plot of Fig.~\ref{fig:steps1}, but others
-happen more gradually, such as the step close to sample 5300. In
-addition, a step can be preceded or followed by a short period of
-instability, as is visible at the bottom of the step in the left hand
-plot of Fig.~\ref{fig:steps1}. Further problems are caused by steps
-that occur close together in time, such as the large downward step
-followed by a smaller upward step close to sample 5000 in the right
-hand plot of Fig.~\ref{fig:steps1}.
-
-Detecting and correcting such a wide variety of steps reliably has proved
-to be a challenge. In outline, the following stages
-are involved in detecting steps in a single bolometer time stream:
-
-\begin{enumerate}
-
-\item median smooth the whole time stream;
-
-\item find the gradient of the median smoothed time series at each
-sample;
-
-\item smooth the gradient values to determine the local mean gradient
-and subtract this local mean from the total gradient to get residual
-gradient;
-
-\item find residual gradient values that exceed 25 times the local RMS
-of the residual gradients;
-
-\item Group these high residual gradients into contiguous blocks of
-samples;
-
-\item merge blocks that are separated by less than 100 samples.
-\end{enumerate}
-
-The above process produces a list of candidate steps in each bolometer
-time series. Each candidate step is then verified, measured and corrected
-using the following procedure:
-
-\begin{enumerate}
-
-\item the above process can mis-interpret the edges of a bright source
-as a step, so we ignore blocks that occur close to bright sources;
-
-\item if the block passes the above test, a least squares linear fit
-is performed to the median-filtered bolometer data just before the
-block, and this fit is extrapolated to predict the data value expected
-at the centre of the block on the basis of the preceding data;
-
-\item a least squares linear fit is performed to the median-filtered
-bolometer data just after the block, and this fit is extrapolated to
-predict the data value expected at the centre of the block on the
-basis of the following data;
-
-\item the difference between these two expected data values is taken
-as the step height;
-
-\item the preceding three steps are repeated several times, each time
-including a different selection of samples in the two least squares
-fits, with the mean and standard deviation of the corresponding set of
-step heights found;
-
-\item if the mean step height is small compared to the standard deviation
-of the step heights, or compared to the noise in the bolometer data, then
-the step is ignored;
-
-\item if the above checks are passed, all subsequent bolometer samples
-are corrected by the mean step height;
-
-\item bolometer samples within the duration of the step, and a few
-samples on either side, are flagged as unusable.
-
-\end{enumerate}
-
-Once all steps have been corrected within a bolometer time stream, a
-constant value is added to all samples in the time stream to restore its
-original mean value.
-
-The results of step correction are shown by the red curves in
-Fig.~\ref{fig:steps1}. For comparison, the blue curve shows the
-uncorrected time stream from a nearby bolometer that does not suffer from
-steps. The agreement between the red and blue curves confirms that
-the step correction algorithm is working satisfactorily.
-
-\subsubsection{Gap filling / apodization}
-\label{sec:gaps}
-
-The evaluation of the \model{FLT} model (Section~\ref{sec:flt}) involves
-taking the Fast Fourier Transform (FFT) of each bolometer time series.
-Data that have been flagged as bad for any reason (for instance, due to
-the presence of spikes, steps, or unusual common-mode signal) need to be
-excluded from this FFT. For this reason, each contiguous block of bad
-data samples is filled with artificial data before taking the FFT. A
-least squares linear fit is performed to the 50 samples preceding the
-block, and a similar fit is performed to the 50 samples following the
-block. These are used to estimate the expected values at the start and
-end of the block of bad values. The bad values are then replaced by
-linear interpolation between the expected start and end values. Gaussian
-noise is added with a standard deviation equal to the mean of the RMS
-residuals in the two fits.
-
-In addition to replacing bad samples before the FFT, it is also
-necessary to ensure that the data values at the start and end of the
-time series are similar. Since an FFT treats the data as a single
-cycle in an infinitely repeating waveform, any large difference
-between starting and ending value will effectively introduce sudden
-steps at the start and end of each cycle, causing unwanted
-oscillations (ringing) in the transform. Another consequence of the
-cyclic nature of the FFT is that features at one end of the time
-series can affect the filtered values at the other end of the time
-series. Two methods are available to avoid these two problems:
-
-\begin{itemize}
-\item Apodisation: a number of samples at the start and end of each
-bolometer time series are multiplied by a cosine function in order to
-roll the data values off smoothly to zero. The default number of samples
-modified at each end of the time series is given by half the ratio of the
-sampling frequency to the lowest retained frequency. In addition, each
-end of the time stream is padded with double this number of zeros. This
-method is illustrated in Fig.~\ref{fig:pad2}.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{pad2.pdf}
-\caption{The black curve shows a bolometer time stream, padded
-with zeros. The red curve shows the time stream after apodisation.}
-\label{fig:pad2}
-\end{figure}
-
-\item Padding with artificial data: instead of padding with zeros,
-each time stream is padded with artificial data that connects the two
-ends of the data stream smoothly and includes Gaussian noise. No
-apodisation is performed. The number of samples of padding at each end
-is again equal to the ratio of the sampling frequency to the lowest
-retained frequency.  This is illustrated in Fig.~\ref{fig:pad1}. See
-\citet{stompor2002} for a thorough discussion of this procedure within
-the context of CMB map-making.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{pad1.pdf}
-\caption{The black curve shows the same bolometer time stream as in
-Fig.~\ref{fig:pad2}, againpadded with zeros. The red curve shows the
-time stream after padding with artificial data.}
-\label{fig:pad1}
-\end{figure}
-
-\end{itemize}
-
-\subsubsection{Flatfields}
-\label{sec:flatfields}
-
-TJ
-
-Discuss heater ramps, normalizing off the common-mode signal
-amplitude, astronomical sources, and the internal flatfield source.
-
-\textit{Surely flatfields are discussed in the main \scuba\ paper so
-don't really have any useful purpose here since they aren't
-technically part of the map-maker.}
-
-\subsubsection{Additional bad bolometer rejection}
-
-TJ
-
-Measuring the noise power spectra and throwing away outliers. Data
-flagged bad by the DA system.
-
-\subsubsection{Map-based despiking}
-\label{sec:mapdespike}
-
-It is possible to use the map estimate (Section~\ref{sec:ast}) to
-provide additional despiking of the time-series. Unlike the
-time-domain despiker (Section~\ref{sec:timedespike}), this calculation
-uses the scatter in the population of samples used to estimate the map
-pixel values (from different times and bolometers) to reject
-outliers. This method is more robust against false-positive detections
-of bright/compact astronomical sources, since transient features in the
-time-series are unlikely to occur by chance whenever a bolometer
-crosses a specific location on the sky, whereas real astronomical
-sources are at fixed spatial locations.
-
-The estimated variance, $\mathbf{v_p}(x_i,y_i)$, of the normalized
-weighted samples that land in the $i$th map pixel is simply the biased
-weighted sample variance (i.e., the variance map value multiplied by
-the number of samples)
-%
-\begin{equation}
-  \mathbf{v_p}(x_i,y_i) = N_i \mathbf{v}(x_i,y_i).
-\end{equation}
-
-In order to compare the weighted \emph{differences} between the
-samples and the map values, [$\mathbf{b}_j - \mathbf{m}(x_i,y_i)$] to
-$\mathbf{v_p}$, they must be scaled appropriately. We define a
-normalized difference, $\mathbf{d}_j$, in such a way that the variance
-of this new variable gives the weighted sample variance of the
-underlying data points:
-%
-\begin{eqnarray}
-  \frac{\sum_j \mathbf{d}_j^2}{N} &=&
-  \frac{\sum_j \mathbf{w}_j [\mathbf{b}_j - \mathbf{m}(x_i,y_i)]^2}
-       {\sum_k \mathbf{w}_k} \\
-   \Rightarrow \mathbf{d}_j^2 &=& \frac{N \mathbf{w}_j [\mathbf{b}_j -
-       \mathbf{m}(x_i,y_i)]^2}{ \sum_k \mathbf{w}_k} .
-\end{eqnarray}
-%
-The map-based despiker flags those $\mathbf{d}_j$ that are further
-than some threshold number of standard deviations,
-$\sqrt{\mathbf{v_p}}$, away from zero so that they are not used in
-subsequent iterations.
 

### 2012-09-06 smurf_paper.tex
add Remo as author, went over section 3 on iterative model calc

@@
     Frossie~Economou$^2$\thanks{Present address: National Optical
-      Astronomy Observatory, 950 N.\ Cherry Avenue, Tucson, AZ 85719, USA}
+      Astronomy Observatory, 950 N.\ Cherry Avenue, Tucson, AZ 85719, USA},
+    Remo~P.~J.~Tilanus$^{2,3}$
   }
@@
   $^{2}$JointAstronomy Centre, 660 N. A`oh\={o}k\={u} Place, University
-  Park, Hilo, Hawaii 96720, USA}
-
+  Park, Hilo, Hawaii 96720, USA\\
+  $^{3}$Netherlands Organisation for Scientific Research,
+  Laan van Nieuw Oost-Indie 300, NL-2509 AC The Hague, The Netherlands}
 \begin{document}
@@
   imaging bands centered over 450 and 850\,\micron. The camera is
-  scanned across the sky and records data at a rate of 200\,Hz. This
-  data volume is much greater than that of previously existing
+  scanned across the sky and records data at a rate of 200\,Hz. The
+  data rate is much greater than that of previously existing
   submillimetre cameras, and represents a significant data anlaysis
-  challenge. We describe the production of SCUBA-2 maps using the
+  challenge. We describe the production of \scuba\ maps using the
   Sub-Millimetre User Reduction Facility (SMURF) in which we have
@@
 itself. The most promising maximum-likelihood method that may one day
-be applied to SCUBA-2 data is ``SANEPIC'', which successfully reduced
+be applied to \scuba\ data is ``SANEPIC'', which successfully reduced
 maps from the Balloon-borne Large-Aperture Submillimeter Telescope,
@@
 of the data analyzed in this paper are publicly available through the
-CADC SCUBA-2 raw-data queries page for the dates and observation
+CADC \scuba\ raw-data queries page for the dates and observation
 numbers given in the
@@
 is only seen at the input to the second stage, the pickup can appear
-with random signs for bolometers along a a column, giving it a
-distinct signature from other common signals that appear with the same
-sign.
+with random signs for bolometers along a column, giving it a distinct
+signature from other common signals that appear with the same sign.
 
@@
 standard WCS information, file formats etc.). Furthermore, Starlink is
-already used extensively at the Joint Astronomy Centre (host of the
-JCMT) for many other systems, which helps with
+open source, and already used extensively at the Joint Astronomy
+Centre (host of the JCMT) for many other systems, which helps with
 interoperability. Though originally written in FORTRAN, many of the
@@
 already been removed in the previous iteration when the map was
-estimated and subtracted, reducing the amount of ringing in the map.
-
-In Sections~\ref{sec:dataprep} and \ref{sec:components} we describe
-the data pre-processing steps and iterative algorithm in
-detail. Section~\ref{sec:converge} explores convergence tests and
-degeneracies between model components.
+estimated and subtracted, reducing the amount of ringing in the map
+once it is re-estimated. Note that the extinction and gain denoted in
+a second dashed box are \emph{multiplicative} rather than additive
+models, and so they must be inverted prior to the start of each
+iteration to preserve the units (see Sections~\ref{sec:comgai} and
+\ref{sec:ext}).
+
+In general, SMURF is highly configurable, including many options for
+both the pre-processing stage and the iterative model components
+(which models are used, what order they are calculated in, how they
+are calculated).  In Sections~\ref{sec:dataprep} and
+\ref{sec:components} we describe the data pre-processing steps and
+iterative algorithm in detail. Section~\ref{sec:converge} explores
+convergence tests and degeneracies between model components.
 
@@
 
-Prior to executing the iterative part of the SMURF algorithm, the data
-must undergo several pre-processing steps. First, the data files are
-read into memory and concatenated into continuous arrays (\scuba\ data
-are broken up and written to disk every $\sim$30\,s). As the data are
-loaded, the flatfield correction is also applied \citep[see
-Section~2.1 in][]{dempsey2012}. Next, a series of data cleaning and
-filtering procedures are applied; these include the removal of large
-glitches that may hinder the iterative solution from converging, or
-simply tasks that do not need to be iterated.
+Prior to executing the iterative part of the algorithm, the data must
+undergo several pre-processing steps. First, the data files are read
+into memory and concatenated into continuous arrays (\scuba\ data are
+broken up and written to disk every $\sim$30\,s during data
+acquisition). As the data are loaded, they are multiplied by the
+flatfield correction \citep[see Section~2.1 in][]{dempsey2012}. Next,
+a series of data cleaning and filtering procedures are applied; these
+include the removal of large glitches that may hinder the iterative
+solution from converging, or simply tasks that do not need to be
+iterated.
 
@@
 only be visited once. From a spectral point of view, this boxcar
-average is equivalent to applying a sinc-function low-pass filter in
-the frequency domain. Such behaviour is desirable since it serves as
-an anti-aliasing filter; most non-white features above the Nyquist
+average is equivalent to applying a sinc-function low-pass
+filter. Such behaviour is desirable since it serves as an
+anti-aliasing filter; most non-white features above the Nyquist
 frequency in $y_j$ that could contaminate the output data are largely
@@
 cleaning by principal component analysis (PCA) is available as an
-experimental pre-processing option. As described in
-Section~\ref{sec:pca}, a new set of basic vectors for the bolometer
-data are found which diagonalize their covariance matrix. The
-brightest components can then be subtracted from the data. A single
-parameter specifies the threshold on the amplitude of the eigenvalues
-to be removed, as a number of standard deviations away from the mean
-value. Subarrays are cleaned independently once for the full length of
-each continuousy chunk of data. Given the computational expense of
-this method, and initial tests which showed little improvement over
-simple high-pass filtering, this method has not yet been explored in
-detail with \scuba\ data. However, since systematic effects seem to
-come and go, it is possible that PCA could be useful for particular
-data sets.
+experimental pre-processing option (Section~\ref{sec:pca}). The most
+significant components in the analysis are identified, and the
+projection of each bolometer time series along their eigenvectors are
+removed. A single parameter specifies the threshold on the amplitude
+of the eigenvalues to be removed, as a number of standard deviations
+away from the mean value. Subarrays are cleaned independently, once,
+for the full length of each continuousy chunk of data. Given the
+computational expense of this method, and initial tests which showed
+little improvement over simple high-pass filtering, this method has
+not yet been explored in detail with \scuba\ data. However, since
+systematic effects seem to come and go, it is possible that PCA could
+be useful for particular data sets.
 
@@
 for two bolometers. If not removed, these steps can cause severe
-ringing in the FLT model (Section~\ref{sec:flt}) and visible streaks
-in the final map, corresponding to the paths of individual bolometers
-over the sky.
+ringing when filtering, and visible streaks in the final map,
+corresponding to the paths of individual bolometers over the sky.
 
@@
 instability, as is visible at the bottom of the step in the left hand
-plot of Fig.~\ref{fig:steps1}. Further problems are caused by steps
-that occur close together in time, such as the large downward step
-followed by a smaller upward step close to sample 5000 in the right
-hand plot of Fig.~\ref{fig:steps1}.
+plot of Fig.~\ref{fig:steps1} (this is probably due to the response of
+the \scuba\ anti-aliasing filter; the sudden large step occurs prior
+to the 200\,Hz re-sampling). Further problems are caused by steps that
+occur close together in time, such as the large downward step followed
+by a smaller upward step close to sample 5000 in the right hand plot
+of Fig.~\ref{fig:steps1}.
 
@@
 
-We express the signal observed by the $i$th bolometer as a function of
-time,
+Once the data have been cleaned, and the worst data flagged to be
+ignored, the iterative solution for the map and contaminating noise
+signals begins.
+
+First, we describe the generic model for \scuba\ data. We express the
+signal observed by the $i$th bolometer as a function of time,
 %
 \begin{equation}
-\mathbf{b}_i(t) = g_i[\mathbf{e}_i(t) \mathbf{a}_i(t) + \mathbf{n}_i(t)],
+\mathbf{b}_i(t) = f_i[\mathbf{e}_i(t) \mathbf{a}_i(t) + \mathbf{n}_i(t)],
+\label{eq:model}
 \end{equation}
@@
 sources of noise. The two terms in square brackets, as written, have
-units of power delivered to the detectors (pW). The scale factor $g_i$
+units of power delivered to the detectors (pW). The scale factor $f_i$
 converts this effective power to the digitized units recorded on disk
-(DAC) -- the \emph{flatfield} multiplied by a digitization constant --
-and in this formulation it is assumed to be constant in time.
+(DAC) -- the flatfield multiplied by a digitization constant (applied
+once during pre-processing) -- which in this formulation is assumed to
+be constant in time.
 
@@
 \begin{equation}
-  \mathbf{n}_i(t) = \mathbf{n}^\mathrm{w}_i(t) + \mathbf{n}^\mathrm{f}_i(t) +
-  \mathbf{n}^\mathrm{c}(t),
+  \mathbf{n}_i(t) = \mathbf{n}^\mathrm{w}_i(t) +
+  g_i\mathbf{n}^\mathrm{c}(t) + \mathbf{n}^\mathrm{f}_i(t),
 \label{eq:noise}
@@
 %
-where $\mathbf{n}^\mathrm{w}_i(t)$ is uncorrelated white (thermal)
-noise, $\mathbf{n}^\mathrm{f}_i(t)$ is low-frequency noise that is
-\emph{not} correlated from bolometer-to-bolometer, and
-$\mathbf{n}^\mathrm{c}(t)$ is a (predominantly low-frequency)
-correlated or \emph{common-mode} component. The primary goal of
-map-making is to model and remove $\mathbf{n}^\mathrm{f}_i$ and
-$\mathbf{n}^\mathrm{c}$ from the time-series, and then by re-gridding
-the remaining data we can hope to approach the theoretical noise limit
-in the map. This procedure is summarized in Fig.~\ref{fig:dimm}.
-
- and run through a pre-processing stage that performs
-several operations including spike and sudden level step removal to
-enforce continuity in the time-series. Then, the iterative process
-within the dashed box begins. It is generally assumed that the
-dominant noise component is the common-mode, which is presumably
-dominated by large variations in the fridge temperature and sky noise
-(as argued in Section~\ref{sec:pca}). This signal is estimated as the
-average signal observed by all of the bolometers at each time step,
-and then removed from all of the bolometers. Once this large signal
-has been removed, we apply the extinction correction, which is
-determined completely from external measurements
-\citep[see][]{dempsey2012}. Next, residual noise that differs from
-bolometer-to-bolometer (primarily low-frequency) is removed, usually
-using a high-pass filter. The resulting ``clean'' time-series are then
-re-gridded to produce a map estimate using nearest-neighbour
-sampling. Since each map pixel is calculated as the average of many
-bolometer samples, it is fundamentally much less noisy than the raw
-bolometer time-series. We can therefore project the map back into the
-time domain (i.e. scanning the array across the current estimate of
-the map), and create approximately noiseless estimates of what each
-bolometer would have seen in the absence of noise sources, and remove
-that from the time-series. This final step in the iteration gives a
-residual signal (with low-frequency and astronomical noise components
-removed) that may be used to measured the white noise level in each
-detector, and also to estimate an approximate $\chi^2$ for the model
-to track convergence. As we will discuss in the next sections, each
-signal component estimate is biased slightly due to the presence of
-signal from other components, so the entire process is
-iterated. However, before the next iteration begins, we first undo the
-multiplicative factors introduced by the extinction correction, and
-also relative gains that may optionally be determined in the
-common-mode estimation phase.
-
-In practice, the map-maker is highly configurable, both in terms of
-the order in which model components are fit (and whether they are fit
-at all), and the details of \emph{how} they are fit. A summary of the
-model components is given in the next section.
-
-In addition to specifying when each model component is computed
-(normally an attempt is made to order them by decreasing amplitude in
-the time-series), there are a number of parameters that control their
-operation. Table~\ref{tab:components} lists the model components and
-the typical order in which they are calculated. The following sections
-describe their implementation in detail.
+where $\mathbf{n}^\mathrm{w}_i(t)$ is uncorrelated white noise,
+$\mathbf{n}^\mathrm{c}(t)$ is a correlated or \emph{common-mode}
+signal (with an optional scale factor $g_i$ for each bolometer), and
+$\mathbf{n}^\mathrm{f}_i(t)$ is (predominantly low-frequency) noise in
+excess of the white noise level that is either uncorrelated from
+bolometer-to-bolometer, or at least does not have a simple correlation
+relationship that would lead to it being included in
+$\mathbf{n}^\mathrm{c}(t)$. The primary goal of map-making is to model
+and remove $\mathbf{n}^\mathrm{c}(t)$ and $\mathbf{n}^\mathrm{f}_i(t)$
+from the time-series, and producing a map from what is left.  This
+procedure is summarized in Fig.~\ref{fig:dimm}.
+
+Each model component is estimated from the data in a user-defined
+sequence, where the typically strongest components are estimated
+first, and then subtracted, leaving a cleaner signal for subsequent
+model estimates. One exception to this rule is the extinction
+correction, $\mathbf{e}_i(t)$ in Eq.~\ref{eq:model}, which is simply a
+multiplicative factor that is known prior to map-making. Note that the
+order in which this correction occurs will modify the units of
+subsequently calculated additive model components (e.g. additional
+brackets may be needed in Eq.~\ref{eq:model} after
+$\mathbf{e}_i(t)$). Another exception is the calculation of
+$\mathbf{n}^\mathrm{c}(t)$ and $\mathbf{n}^\mathrm{f}_i(t)$ in which
+the previous iterations are usually added back into the time-series
+simultaneously to avoid degeneracies (see Section~\ref{sec:converge}).
+
+Usually the last two components estimated are $\mathbf{a}_i(t)$ and
+$\mathbf{w}_i(t)$ once the other contaminating noise sources, and
+extinction, have been corrected for. Provided that these two signals
+are the only components of the residual bolometer signal, regridding
+into a map provides the highest \snr\ estimate of the astronomical
+sky. The signal produced by this map is then projected back into the
+time-domain to estimate and remove $\mathbf{a}_i(t)$, leaving a
+residual signal that should be dominated by white noise,
+$\mathbf{n}_i^\mathrm{w}(t)$.
+
+After the first iteration, there will almost certainly be correlated
+errors between the model estimates. For example, the presence of a
+bright astronomical source will contaminate $\mathbf{n}^\mathrm{c}(t)$
+(which is a simple average of all of the bolometer time-series at each
+instant in time), leading to an \emph{over-subtraction}, and in turn,
+negative bowls in the map.
+
+Subsequent iterations, however, diminish such problems. Each additive
+model component is re-estimated in the same sequence, after first
+adding the previous estimate back into the time-series (but
+importantly, not the other components). In this case, much of the
+astronomical signal will have been identified and removed in the
+previous calculation of $\mathbf{a}_i(t)$, and therefore there will be
+less contamination in $\mathbf{n}^\mathrm{c}(t)$, and less negative
+bowling in the map. Any multiplicative models (e.g.,
+$\mathbf{e}_i(t)$) are inverted immediately prior to the start of a
+new iteration to preserve the units of the data. The iterative
+solution will thus converge, barring degeneracies between model
+components.
+
+The full set of model components available, and the parameters that
+control them, are described in the following
+sections. Table~\ref{tab:components} shows the typical order in which
+they are calculated. For a discussion on convergence tests and
+degenerices, see Section~\ref{sec:converge}.
 
@@
 
-The \model{COM} model is the common-mode signal itself. It is a single
-time series that is estimated by finding the mean of all bolometers at
-each time step. Bolometer values that have been flagged as unusable
-are excluded from the mean. {\bf Per: has a comment in
-here about stability of the gain pattern... it should be constant
-right?}.
+The \model{COM} model is the common-mode signal itself
+($\mathbf{n}^{c}(t)$ in Eq.~\ref{eq:noise}). It is a single time
+series that is estimated by finding the mean of all bolometers at each
+time step. Bolometer values that have been flagged as unusable are
+excluded from the mean.
 
@@
 stream with the common-mode signal allows an estimate of the relative
-bolometer sensitivity to be obtained. In practice, a least squares linear
-fit is performed between the bolometer time series and the common-mode to
-determine a gain and offset for each bolometer.  The gain and offset for
-each bolometer is known as the \model{GAI} model. Each bolometer value
-can then optionally be scaled and shifted using these values so that all
-bolometers share a common (but as yet unknown) calibration. This provides
-an alternative (or additional) flat-fielding strategy to that described
-in section~\ref{sec:bolos}.
-
-An option exists to cater for time-varying sensitivities. If used, the
-least squares fits described above are performed on short blocks of
-contiguous time slices, providing multiple gain and offset values for
-each bolometer (one pair for each block of time slices). The gain and
-offset at any required time slice can then be found by interpolation
-between these values.
+bolometer sensitivity to be obtained. In practice, a least squares
+linear fit is performed between the bolometer time series and the
+common-mode to determine a gain ($g_i$ in Eq.~\ref{eq:noise}) and
+offset for each bolometer.  The gain and offset for each bolometer is
+known as the \model{GAI} model. Each bolometer value can then
+optionally be scaled and shifted using these values so that all
+bolometers share a common (but as yet unknown) calibration. This
+provides an alternative (or additional) flat-fielding strategy to that
+described in section~\ref{sec:bolos}.
+
+An option exists to cater for time-varying sensitivities. In
+principle, the gain of the bolometers should be constant in
+time. However, there is evidence that there are slight variations, and
+this option does tend to slightly improve the noise in maps.  If used,
+the least squares fits described above are performed on short blocks
+of contiguous time slices, providing multiple gain and offset values
+for each bolometer (one pair for each block of time slices). The gain
+and offset at any required time slice can then be found by
+interpolation between these values.
 
@@
 common mode. Bolometers for which the correlation coefficient is below
-a specified limit, or which have unusually high or low gain compared to
-the other bolometers, are flagged as bad in order to omit them from the
-final map. If the option described
-above for handling time-varying sensitivities is used, then a
-correlation coefficient can be determined for each individual block of
-time slices.  This allows individual bad blocks to be rejected from a
-bolometer time stream, rather than rejecting the whole bolometer. This
-is beneficial for bolometers that follow the common-mode for most of
-the time but suffer from transient problems at specific times. {\bf
-Per: If magnetic pickup dominates the common mode a negative gain is
-normal and should not be removed. This might be a risk at high
-elevation in very good stable weather!}
-
-The common-mode value at each time slice is calculated as the unweighted
-mean of the values of all bolometers that have not previously been
-flagged as bad for some other reason. Bolometer values that were flagged
-as bad simply because they were poorly correlated with the common-mode on
-the previous iteration are included in the new common-mode estimate. If
-such samples are excluded, there is a strong possibility of
-discontinuities appearing in the COM model at block boundaries.
-These in turn can lead to ringing in the low pass filter, and
-instabilities in the convergence process.
+a specified limit, or which have unusually high or low gain compared
+to the other bolometers, are flagged as bad in order to omit them from
+the final map. If the option described above for handling time-varying
+sensitivities is used, then a correlation coefficient can be
+determined for each individual block of time slices.  This allows
+individual bad blocks to be rejected from a bolometer time stream,
+rather than rejecting the whole bolometer. This is beneficial for
+bolometers that follow the common-mode for most of the time but suffer
+from transient problems at specific times. As a warning, rejecting
+data based on outlier gain values can be misleading in cases where the
+data are dominated by magnetic field pickup. For example, the
+common-mode signal for the data in Fig.~\ref{fig:magpickup} would
+resemble the azimuthal scan pattern, but clearly the gains will have
+opposite signs (such that one or the other bolometer would be
+erroneously flagged).
+
+The common-mode value at each time slice is calculated as the
+unweighted mean of the values of all bolometers that have not
+previously been flagged as bad for some other reason. Bolometer values
+that were flagged as bad simply because they were poorly correlated
+with the common-mode on the previous iteration are included in the new
+common-mode estimate. If such samples are excluded, there is a strong
+possibility of discontinuities appearing in the \model{COM} model at
+block boundaries.  These in turn can lead to ringing when filtering,
+and instabilities in the convergence process.
 
@@
 Any extended astronomical emission on a scale comparable to or larger
-than the spatial extent of the area used to estimate the common-mode will
-contribute a similar signal to all bolometers. Therefore such extended
-emission is indistinguishable from the other sources of common-mode
-signal (fridge and atmosphere variations) and will be removed by the
-model. This places a limit on the spatial extent of astronomical
+than the spatial extent of the area used to estimate the common-mode
+will contribute a similar signal to all bolometers. Therefore such
+extended emission is indistinguishable from the other sources of
+common-mode signal (e.g. atmosphere variations) and will be removed by
+the model. This places a limit on the spatial extent of astronomical
 structure that can be recovered.
@@
 derived using the JCMT water vapour radiometer (WVM), and is not
-considered to be a free parameter in the solution. However, it is
-applied as part of the iterative solution, rather than a
-pre-processing step, since any small errors will be amplified by the
-large low-frequency drifts in the raw bolometer time-series. For
-example, if the bolometer drift is 1000 times greater than an
-astronomical source of interest, a 1 per cent error in the flatfield
-will produce stripes of order 10 times the astronomical signal
-amplitude in the final map! If \model{EXT} is applied after the bulk
-of the low-frequency noise has been removed (e.g., \model{COM,GAI}),
-then there is little potential for such small errors to affect the
-final map.  For details on how it is calculated see
-\citet{dempsey2012}. Note that its numerical value is calculated only
-once, and simply applied as a scale factor in the iterative solution.
+considered to be a free parameter in the solution ($\mathbf{e}_i(t)$
+in Eq.~\ref{eq:model}). However, it is applied as part of the
+iterative solution, rather than a pre-processing step, since any small
+errors will be amplified by the large low-frequency drifts in the raw
+bolometer time-series. For example, if the bolometer drift is 1000
+times greater than an astronomical source of interest, a 1 per cent
+error in the flatfield will produce stripes of order 10 times the
+astronomical signal amplitude in the final map! If \model{EXT} is
+applied after the bulk of the low-frequency noise has been removed
+(e.g., \model{COM,GAI}), then there is little potential for such small
+errors to affect the final map.  For details on how it is calculated
+see \citet{dempsey2012}. Note that its numerical value is calculated
+only once, and simply applied as a scale factor in the iterative
+solution. Unlike the additive model components, it is inverted at the
+start of each iteration to preserve the amplitude of the data before
+the re-calculation of other model components.
 
@@
 are stored in the \model{FLT} container array. Typically this model is
-used purely as a high-pass filter to remove most of the residual $1/f$
-noise following common-mode removal. Low-pass filtering is redundant
-for two reasons: (i) SMURF already low-pass filters and re-samples to
-a lower sample rate, as described in Section~\ref{sec:downsamp}; and
-(ii) the act of re-gridding the data to produce map estimates
-effectively low-pass filters the data at a frequency that corresponds
-to the inverse of the crossing time of a single map pixel. Notch
-filters have not been proven to be useful with \scuba\ data,
-particularly since line features tend to move around (i.e. a dynamic
-line-detection system would have to be developed, and care would have
-to be taken that astronomical sources are not suppressed).
+used purely as a high-pass filter to remove most of the residual
+uncorrelated $1/f$ noise following common-mode removal
+($\mathbf{n}^\mathrm{f}_i(t)$ in Eq.~\ref{eq:noise}). Low-pass
+filtering is redundant for two reasons: (i) SMURF already low-pass
+filters and re-samples to a lower sample rate, as described in
+Section~\ref{sec:downsamp}; and (ii) the act of re-gridding the data
+to produce map estimates effectively low-pass filters the data at a
+frequency that corresponds to the inverse of the crossing time of a
+single map pixel. Notch filters have not been proven to be useful with
+\scuba\ data, particularly since line features tend to move around
+(i.e. a dynamic line-detection system would have to be developed, and
+care would have to be taken that astronomical sources are not
+suppressed).
 
@@
 into the time domain (the signal that would be produced in each
-bolometer by the signal represented by the map) and removed.
+bolometer by the signal represented by the map, $\mathbf{a}_i(t)$ in
+Eq.~\ref{eq:model}) and removed.
 
@@
 Research, and the National Research Council of Canada. Additional
-funds for the contruction of SCUBA-2 were provided by the Canada
+funds for the contruction of \scuba\ were provided by the Canada
 Foundation for Innovation. EC thanks CANARIE/CANFAR for additional

### 2012-09-06 smurf_paper.tex
numerous fiddles throughout

@@
 
-Submillimetre imaging cameras generally use bolometers to maximise
-sensitivity, in which an absorber collects photons that pass through
-broad-band filters, and the signal is read out using a thermometer
-\citep[e.g.][]{griffin2002}. The sensitivity of bolometers in a camera
-like \scuba\ are limited by white photon and phonon noise from the
+Submillimetre imaging cameras generally use bolometers, rather than
+coherent detectors, to maximise sensitivity. The sensitivity of
+bolometers is limited by white photon and phonon noise from the
 instrument and ambient backgrounds. The low-frequency noise, however,
-is dominated by sources which produce slow variations in mean backgrounds
-(e.g., thermal variations within the cryostat, and the atmosphere for
-ground-based cameras), and drifts in the readout electronics. Such
-noise has a power spectrum $\propto 1/f^\alpha$ ($\alpha>0$), and the
-frequency at which it is comparable to the white noise level is called
-the ``$1/f$ knee''. Since the low-frequency noise is largely
-correlated between all, or subsets, of the bolometers in time, it can
-be suppressed during map-making, since astronomical signals have the
-distinct property that they are fixed in a sky reference frame
-(assuming they are not time-varying). If successful, the noise in the
-resulting map is said to be ``white noise limited'', meaning that the
-map noise is uncorrelated spatially, and has an amplitude that scales
-as the $\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the
-noise-equivalent flux-density (the white noise level of a bolometer in
-1\,s of integration), and $t$ is the amount of integration time in a
-map pixel.
+is typically dominated by sources which produce slow variations the
+background (e.g., thermal variations within the cryostat, and the
+atmosphere for ground-based cameras), and drifts in the readout
+electronics. Such noise has a power spectrum $\propto 1/f^\alpha$
+($\alpha>0$), and the frequency at which it is comparable to the white
+noise level is called the ``$1/f$ knee''. Since the low-frequency
+noise is largely correlated between all, or subsets, of the bolometers
+in time, it can be suppressed during map-making, since astronomical
+signals have the distinct property that they are fixed in a sky
+reference frame (assuming they are not time-varying). If successful,
+the noise in the resulting map is said to be ``white noise limited'',
+meaning that the map noise is uncorrelated spatially, and has an
+amplitude that scales as the $\mathrm{NEFD}/\sqrt{t}$, where the NEFD
+is the noise-equivalent flux-density (the white noise level of a
+bolometer in 1\,s of integration), and $t$ is the amount of
+integration time in a map pixel.
 
@@
 ground-based submillimetre cameras
-\citep[e.g.][]{glenn1998,weferling2002,wilson2008,kovacs2008b}., in
+\citep[e.g.,][]{glenn1998,weferling2002,wilson2008,kovacs2008b}., in
 favour of ``chopping'' methods (where the secondary is quickly moved
@@
 \citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO
-\citep[e.g.][]{omont2001}, another camera from the same
+\citep[e.g.,][]{omont2001}, another camera from the same
 generation. Generally speaking, such methods are fast, although
@@
 reducing data in fields of faint point sources with BOLOCAM
-\citep[e.g.][]{laurent2005}, and its younger sibling AzTEC
-\citep[e.g.][]{scott2008}, is ``principal component analysis'' (PCA)
+\citep[e.g.,][]{laurent2005}, and its younger sibling AzTEC
+\citep[e.g.,][]{scott2008}, is ``principal component analysis'' (PCA)
 cleaning. A statistical black-box removes the most correlated
@@
 fast iterative methods for the inversion do exist with minimal memory
-usage \citep[e.g.][]{wright1996}, for the more general map-making
-problem involving many detectors, the problem is significantly
-complicated both by the need to measure the cross power-spectra for
-all unique pairs of bolometers, and then to perform the inversion
-itself. The most promising maximum-likelihood method that may one day
-be applied to \scuba\ data is ``SANEPIC'', which successfully reduced
-maps from the Balloon-borne Large-Aperture Submillimeter Telescope,
-involving hundreds of bolometers, and correctly incorporating
-inter-bolometer noise correlations \citep[][]{patanchon2008}.
+usage \citep[e.g.,][]{wright1996}, for the more general map-making
+problem involving many detectors, things are significantly complicated
+both by the need to measure the cross power-spectra for all unique
+pairs of bolometers, and then to perform the inversion itself. The
+most promising maximum-likelihood method that may one day be applied
+to \scuba\ data is ``SANEPIC'', which successfully reduced maps from
+the Balloon-borne Large-Aperture Submillimeter Telescope, involving
+hundreds of bolometers, and correctly incorporating inter-bolometer
+noise correlations \citep[][]{patanchon2008}.
 
@@
 %\citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA analysis
-%\citep[e.g.][]{laurent2005,scott2008,aguirre2010}.
+%\citep[e.g.,][]{laurent2005,scott2008,aguirre2010}.
 
@@
 
-This paper is divided in the following way. We first describe the
-properties of \scuba\ data and scan patterns in Section~\ref{sec:data}
-to motivate the data model that we have adopted. Note that most of the
-analysis described here applies equally to all of the data taken
-to-date with \scuba, including during the Shared-Risks Observing
-period (S2SRO) from 2010 February 11 to 2010 March 25 (during which
-only single subarrays were available at each of the two observing
-wavelengths). A comparison between the S2SRO and fully-commissioned
-instrument is given in Section~\ref{sec:s2sro}. Next, the details of
-the SMURF algorithm are given in Section~\ref{sec:algorithm}. The
-paper is concluded in Section~\ref{sec:examples} with three detailed
-test cases that span the majority of observation types likely to be
-undertaken with \scuba, with an emphasis on the mitigation of
-divergence problems, and characterizing the output maps: (i) Uranus, a
-bright, compact source (Section~\ref{sec:point}); (ii) the Lockman
-Hole, a blind survey of faint point-like source
-(Section~\ref{sec:cosmo}); and (iii) the star-forming region M17,
-including bright, extended emission (Section~\ref{sec:extended}). All
-of the data analyzed in this paper are publicly available through the
-CADC \scuba\ raw-data queries page for the dates and observation
-numbers given in the
+This paper is organized in the following way. We first describe the
+properties of \scuba\ data in Section~\ref{sec:data} to motivate the
+data model that we have adopted. Next, the details of the SMURF
+algorithm are given in Section~\ref{sec:algorithm}. The paper is
+concluded in Section~\ref{sec:examples} with three detailed test cases
+that span the majority of observation types likely to be undertaken
+with \scuba, with an emphasis on the mitigation of divergence
+problems, and characterizing the output maps: (i) Uranus, a bright,
+compact source (Section~\ref{sec:point}); (ii) the Lockman Hole, a
+blind survey of faint point-like source (Section~\ref{sec:cosmo}); and
+(iii) the star-forming region M17, including bright, extended emission
+(Section~\ref{sec:extended}). All of the data analyzed in this paper
+are publicly available through the CADC \scuba\ raw-data queries page
+for the dates and observation numbers given in the
 text\footnote{\texttt{http://www3.cadc-ccda.hia-iha.nrc-cnrc.gc.ca/
@@
 
-In this section we summarize how \scuba\ bolometers detect temperature
-variations (Section~\ref{sec:bolos}), give exampels of typical
-bolometers signals (Section~\ref{sec:bolosignal}), and the impact of
-magnetic field pickup (Section~\ref{sec:magpickup}), and finally
-illustrate the use of principal component analysis to explore the
-correlated noise properties of \scuba\ data (Section~\ref{sec:pca}).
+In this section we summarize how \scuba\ works,
+(Section~\ref{sec:bolos}), give exampels of typical bolometers signals
+(Section~\ref{sec:bolosignal}), and the impact of magnetic field
+pickup (Section~\ref{sec:magpickup}), and finally illustrate the use
+of principal component analysis to explore the correlated noise
+properties of \scuba\ data (Section~\ref{sec:pca}).
 
@@
 
-Incoming light passes first through a beam-splitter, and then bandpass
+Incoming light passes through a beam-splitter, and then bandpass
 filters, providing simultaneous illumination and wavelength definition
-of both the 450 and 850\,\micron\ focal planes. The bolometers in each
-focal plane consist of absorbers coupled to superconducting
-transition-edge sensors (TESs) for thermometry. The focal planes are
-held at a temperature somewhat below the critical temperatures of the
-TESs by the fridge. The devices are then biased into the
-superconducting transition using a combination of individual pixel
-heaters (to compensate for most of the bulk variations in loading from
-the atmosphere and other backgrounds), and an electrical voltage
-bias. Throughout an observation the heater bias is fixed, but
-smaller-scale load variations both from the backgrounds and, to a much
-lesser extent, astronomical sources, significantly alter the TES
-impedences as their temperatures vary. Thermoelectrical feedback then
-keeps the TESs ``locked'' in the superconducting transition; an
-increase in temperature increases the impedence, and lowers the
-current, and therefore the heating resulting from the voltage
-bias. These current changes produce varying magnetic fields that are
-amplified using chains of superconducting quantum interference devices
-(SQUIDs), and the larger output currents are digitized. Each detector
-has its own SQUID for the first-stage of the amplification, but the
-remaining stages occur within a common chain of SQUIDs for each of the
-32 columns. All 40 rows are read-out in sequence at a row-visit
-frequency of about 12\,kHz. Such a high sample rate is unnecessary to
-produce maps, so the data are re-sampled to approximately 200\,Hz
-before writing to disk.  It should also be noted that there is an
-additional 41st row of SQUID readouts that are not connected to
-TESs. These ``dark SQUIDs'' track non-thermal noise sources that are
-common to each column's amplifier chain. The relationship between the
-output digitized current, $I$, and the input power, $P$, or $dI/dP$ is
-established using \emph{flatfield} observations immediately prior to
-science observations, in which the output signal is measured
-throughout a ramp of the pixel heaters (which provide a known input
-power); see Section~2.1 in \citet{dempsey2012} and Section~5.3 in
-\citet{holland2012}. Finally, the conversion to astronomical flux
-units from power involves a correction for atmospheric extinction
-\citep[primarily using the JCMT Water Vapour Monitor to track
-line-of-sight opacity variations, see Section~3 in][]{dempsey2012},
-and the application of a flux conversion factor (FCF) which is
-established from regular measurements of calibrators such as Uranus
-\citep[Section~5 in][]{dempsey2012}.
+of both the 450 and 850\,\micron\ focal planes. Each focal plan is
+populated by four \emph{subarrays} (labeled s4a--s4d at 450\,\micron,
+and s8a--s8d at 850\,\micron), each consisting of 32 columns and 40
+rows of bolometers. The bolometers are thermal absorbers coupled to
+superconducting transition-edge sensors (TESs) for thermometry;
+temperature variations in the TESs produce changing currents, and
+therefore varying magnetic fields.  The changing magnetic fields are
+detected and amplified using chains of superconducting quantum
+interference devices (SQUIDs), and the larger output currents are
+digitized. Each detector has its own SQUID for the first-stage of the
+amplification, but the remaining stages occur within a common chain of
+SQUIDs for each of the 32 columns. All 40 rows are read-out in
+sequence at a row-visit frequency of about 12\,kHz. Such a high sample
+rate is unnecessary to produce maps, so the data are re-sampled to
+approximately 200\,Hz before writing to disk.  It should also be noted
+that there is an additional 41st row of SQUID readouts that are not
+connected to TESs. These \emph{dark SQUIDs} track non-thermal noise
+sources that are common to each column's amplifier chain. The
+relationship between the output digitized current, $I$, and the input
+power, $P$, or $dI/dP$ is established using \emph{flatfield}
+observations immediately prior to science observations, in which the
+output signal is measured throughout a ramp of the pixel heaters
+(which provide a known input power); see Section~2.1 in
+\citet{dempsey2012} and Section~5.3 in \citet{holland2012}. Finally,
+the conversion to astronomical flux units from power involves a
+correction for atmospheric extinction \citep[primarily using the JCMT
+Water Vapour Monitor to track line-of-sight opacity variations, see
+Section~3 in][]{dempsey2012}, and the application of a flux conversion
+factor (FCF) which is established from regular measurements of
+calibrators such as Uranus \citep[Section~5 in][]{dempsey2012}.
 
@@
 bolometer); and (ii) longer timescale drifts resulting from slow
-variations in the mean load (e.g. fridge, telescope, or atmospheric
+variations in the mean load (e.g., fridge, telescope, or atmospheric
 changes), and also the readout electronics (often correlated amongst
@@
 simplest scan one might use is a raster ``boustrophedon'' pattern,
-i.e. a series of parallel passes across the field. However, such a
+i.e., a series of parallel passes across the field. However, such a
 scan yields extremely poor performace for structures that are
@@
 \includegraphics[width=0.49\linewidth]{bolos_point_mix.pdf}
-\caption{A comparison between single bolometer time series in each of
+\caption{A comparison between single bolometer time-series in each of
   the 450 and 850\,\micron\ bands with the mixing chamber temperature
@@
 
-In Fig.~\ref{fig:bolos_mix} we show sample time series from single
+In Fig.~\ref{fig:bolos_mix} we show sample time-series from single
 bolometers in each of the 450 and 850\,\micron\ focal planes, as well
@@
 have calculated a common-mode signal, $\mathbf{c}(t)$, as the average
-time series of all the working bolometers. We then fit the amplitude
+time-series of all the working bolometers. We then fit the amplitude
 of $\mathbf{c}(t)$ at each wavelength to the signals shown in
@@
 gives the same variance as the time-series variance across the full
-time series. In other words, given a bolometer signal $\mathbf{b}(t)$,
+time-series. In other words, given a bolometer signal $\mathbf{b}(t)$,
 %
@@
 the amplification of magnetic fields, any additional changing fields
-within the cryostat will also be detected.
+within the instrument will also be detected.
 
@@
   of Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec}. For each component
-  the top plot shows the time series of the normalized eigenvectors,
+  the top plot shows the time-series of the normalized eigenvectors,
   the middle plot its PSD, and the bottom coloured panels the
   eigenvalues for the bolometers at each wavelength (the amplitudes of
-  the eigenvectors in the time series). For reference, both the mean,
+  the eigenvectors in the time-series). For reference, both the mean,
   $\bar{\lambda}$, and \rms, $\lambda_\mathrm{rms}$, eigenvalues for
@@
 The basic method is as follows: (i) a covariance matrix is built up
-for all pairs $(i,j)$ of the $N$ bolometer time series,
+for all pairs $(i,j)$ of the $N$ bolometer time-series,
 $\langle\mathbf{b}_i(t),\mathbf{b}_j(t)\rangle$; and (ii) a singular
 value decomposition identifies a new set of statistically uncorrelated
-eigenvectors, $\mathbf{\xi}_i(t)$ (i.e. whose covariance matrix is
-diagonal), such that each of the bolometer time series is a linear
+eigenvectors, $\mathbf{\xi}_i(t)$ (i.e., whose covariance matrix is
+diagonal), such that each of the bolometer time-series is a linear
 combination of the eigenvectors, or \emph{components}
-i.e. $\mathbf{b}_i(t) = \bar{\mathbf{\xi}}
+i.e., $\mathbf{b}_i(t) = \bar{\mathbf{\xi}}
 \mathbf{\lambda}_i^\mathrm{T}$, where each row of the matrix
@@
 Fig.~\ref{fig:pca} is clearly produced by Components~1 and 2. In both
-cases, the eigenvector time series exhibit a roughly periodic signal
+cases, the eigenvector time-series exhibit a roughly periodic signal
 that resembles the scan pattern in the right-hand panel of
@@
 noise, in part due to the clear correlation patterns that appear along
-columns in the eigenvalue maps at both wavelengths (e.g. columns 26
+columns in the eigenvalue maps at both wavelengths (e.g., columns 26
 and 29 at 450\,\micron in Component~6). Finally, these components also
@@
 offsets that are likely residuals due to imperfect step correction
-(e.g. $\sim$50\,s in the eigenvector time-series for Component~6; also
+(e.g., $\sim$50\,s in the eigenvector time-series for Component~6; also
 see Section~\ref{sec:steps}).
@@
 Finally, Component~12 in Fig.~\ref{fig:pca} is again significant in
-only a handful of detectors in the 850\,\micron\ subarray (e.g. column
+only a handful of detectors in the 850\,\micron\ subarray (e.g., column
 21 row 25). The time-series plot for the eigenvector exhibits a large
@@
 due to the fact that many of the noise sources are not stationary in
-time (e.g. scan-synchronous noise which obviously depends on the scan
+time (e.g., scan-synchronous noise which obviously depends on the scan
 pattern). It should also be clear from this example that while PCA
@@
 \caption{Typical map-making algorithm. Raw data (stored in multiple
-  files) are read and concatenated into continuous time series. A
+  files) are read and concatenated into continuous time-series. A
   pre-processing stage repairs DC steps, applies a flatfield
@@
   estimating the map by re-gridding the data; and finally measuring
-  the noise in the residual time series. If the \rms\ has converged,
-  the final map is written to disk. Otherwise any multiplicative
+  the noise in the residual time-series. If the solution has
+  converged, the map is written to disk. Otherwise any multiplicative
   factors that may have been applied to the data are inverted (e.g.,
-  extinction correction and relative gains measured during the
-  common-mode calculation), and each additive component is added back
-  into the time series and re-estimated in turn.  {\bf Per: The data
-  flow is not clear to me. One way would be to take the ast map and
-  remove extinction and gain correction and subtract it from the raw
-  data and then simply start the iteration again. That would refine
-  the com model from ast model pollution. But the text suggest that
-  the other models also are removed from the data. Indeed this would
-  be way different models can interact diverge. This process how and
-  when this is done is not clear and there must be a number of
-  subtleties affecting the subtraction of the models.}}
+  extinction correction), and then each additive component is
+  re-estimated in turn, with the previous estimate first being added
+  back into the time-series (and importantly, \emph{not} the other
+  components).}
 \label{fig:dimm}
@@
 extend SMURF with new models. Parallelisation is incorporated in the
-most time-consuming low-level routines using threads (e.g. performing
+most time-consuming low-level routines using threads (e.g., performing
 Fast Fourier Transforms, re-gridding the data), usually handling
@@
 complicated relationships amongst groups of bolometers
-(e.g. Fig.~\ref{fig:pca}). While the low-frequency noise can be
+(e.g., Fig.~\ref{fig:pca}). While the low-frequency noise can be
 significantly reduced using simple common-mode removal (subtracting
@@
 significant components in the analysis are identified, and the
-projection of each bolometer time series along their eigenvectors are
+projection of each bolometer time-series along their eigenvectors are
 removed. A single parameter specifies the threshold on the amplitude
@@
 spikes. This section describes the detection and removal of spikes
-within the time series of each bolometer, and Section~\ref{sec:ast}
+within the time-series of each bolometer, and Section~\ref{sec:ast}
 describes the iterative detection and removal of spikes as part of map
@@
 
-Each one-dimensional bolometer time series is processed
+Each one-dimensional bolometer time-series is processed
 independently. At each time slice, the median value of the current
@@
 \includegraphics[width=\linewidth]{steps1.pdf}
-\caption{Examples of steps in time series data. Steps occur with a wide
+\caption{Examples of steps in time-series data. Steps occur with a wide
 range of heights from the large steps shown on the left to the small
@@
 ``over-shoot'' as shown on the left. In both plots, the black curve is
-the uncorrected time series, and the red curve is the corrected time
-series. Samples close to a step are omitted in the corrected time series.
-In the right hand plot, the blue curve is the uncorrected time series for
+the uncorrected time-series, and the red curve is the corrected time
+series. Samples close to a step are omitted in the corrected time-series.
+In the right hand plot, the blue curve is the uncorrected time-series for
 a nearby bolometer. The similarity between the red and blue curves shows
@@
 
-Sudden steps can occur in the time stream data from each bolometer,
+Sudden steps can occur in the time-series data from each bolometer,
 with the most likely cause being cosmic ray events \citep[see
 Section~3.5.3 in][]{holland2012}. The black curves in
-Fig.~\ref{fig:steps1} show examples of such steps in the time streams
+Fig.~\ref{fig:steps1} show examples of such steps in the time-series
 for two bolometers. If not removed, these steps can cause severe
@@
 
-Detecting and correcting such a wide variety of steps reliably has proved
-to be a challenge. In outline, the following stages
-are involved in detecting steps in a single bolometer time stream:
+Detecting and correcting such a wide variety of steps reliably has
+proved to be a challenge. In outline, the following stages are
+involved in detecting steps in a single bolometer time-series:
 
@@
 
-\item median smooth the whole time stream;
-
-\item find the gradient of the median smoothed time series at each
+\item median smooth the whole time-series;
+
+\item find the gradient of the median smoothed time-series at each
 sample;
@@
 The above process produces a list of candidate steps in each bolometer
-time series. Each candidate step is then verified, measured and corrected
+time-series. Each candidate step is then verified, measured and corrected
 using the following procedure:
@@
 
-Once all steps have been corrected within a bolometer time stream, a
-constant value is added to all samples in the time stream to restore its
+Once all steps have been corrected within a bolometer time-series, a
+constant value is added to all samples in the time-series to restore its
 original mean value.
@@
 Fig.~\ref{fig:steps1}. For comparison, the blue curve shows the
-uncorrected time stream from a nearby bolometer that does not suffer from
+uncorrected time-series from a nearby bolometer that does not suffer from
 steps. The agreement between the red and blue curves confirms that
@@
 necessary to ensure that the data values at the start and end of the
-time series are similar. Since an FFT treats the data as a single
+time-series are similar. Since an FFT treats the data as a single
 cycle in an infinitely repeating waveform, any large difference
@@
 \item Apodisation: a number of samples at the start and end of each
-bolometer time series are multiplied by a cosine function in order to
+bolometer time-series are multiplied by a cosine function in order to
 roll the data values off smoothly to zero. The default number of
-samples modified at each end of the time series is given by half the
+samples modified at each end of the time-series is given by half the
 ratio of the sampling frequency to the lowest retained frequency. In
-addition, each end of the time stream is padded with double this
+addition, each end of the time-series is padded with double this
 number of zeros. This method is illustrated in
@@
 the amount of data available for the map, and can significantly hinder
-very short observations (e.g. of calibrators when focussing the
+very short observations (e.g., of calibrators when focussing the
 telescope).
@@
 \includegraphics[width=\linewidth]{pad2.pdf}
-\caption{The black curve shows a bolometer time stream, padded
-with zeros. The red curve shows the time stream after apodisation.}
+\caption{The black curve shows a bolometer time-series, padded
+with zeros. The red curve shows the time-series after apodisation.}
 \label{fig:pad2}
@@
 \item Padding with artificial data: instead of padding with zeros,
-each time stream is padded with artificial data that connects the two
+each time-series is padded with artificial data that connects the two
 ends of the data stream smoothly and includes Gaussian noise. No
@@
 \includegraphics[width=\linewidth]{pad1.pdf}
-\caption{The black curve shows the same bolometer time stream as in
+\caption{The black curve shows the same bolometer time-series as in
 Fig.~\ref{fig:pad2}, againpadded with zeros. The red curve shows the
-time stream after padding with artificial data.}
+time-series after padding with artificial data.}
 \label{fig:pad1}
@@
 order in which this correction occurs will modify the units of
-subsequently calculated additive model components (e.g. additional
+subsequently calculated additive model components (e.g., additional
 brackets may be needed in Eq.~\ref{eq:model} after
@@
 
-﻿Fig.~\ref{fig:com} shows the time series from a selection of typical
+﻿Fig.~\ref{fig:com} shows the time-series from a selection of typical
 bolometers.\footnote{The data have been flat-fielded and each time
@@
 \includegraphics[width=\linewidth]{com.pdf}
-\caption{A selection of typical bolometer time series after
+\caption{A selection of typical bolometer time-series after
 flatfielding and removal of a constant baseline. It can be seen that
@@
 bolometer sensitivity to be obtained. In practice, a least squares
-linear fit is performed between the bolometer time series and the
+linear fit is performed between the bolometer time-series and the
 common-mode to determine a gain ($g_i$ in Eq.~\ref{eq:noise}) and
@@
 determined for each individual block of time slices.  This allows
-individual bad blocks to be rejected from a bolometer time stream,
+individual bad blocks to be rejected from a bolometer time-series,
 rather than rejecting the whole bolometer. This is beneficial for
@@
 Fig.~\ref{fig:dimm}, a large fraction of the remaining astronomical
-signal is extracted from the bolometer time series and transferred to the
+signal is extracted from the bolometer time-series and transferred to the
 output map, resulting in subsequent estimates of the common-mode being
@@
 extended emission is indistinguishable from the other sources of
-common-mode signal (e.g. atmosphere variations) and will be removed by
+common-mode signal (e.g., atmosphere variations) and will be removed by
 the model. This places a limit on the spatial extent of astronomical
@@
 
-\subsubsection{\model{FLT}: fourier transform filter}
+\subsubsection{\model{FLT}: Fourier transform filter}
 \label{sec:flt}
@@
 \scuba\ data, particularly since line features tend to move around
-(i.e. a dynamic line-detection system would have to be developed, and
+(i.e., a dynamic line-detection system would have to be developed, and
 care would have to be taken that astronomical sources are not
@@
 Eq.~\ref{eq:psd}. If the bolometer noise were produced purely by
-uncorrelated thermal sources (i.e. no other long-scale drifts), with
+uncorrelated thermal sources (i.e., no other long-scale drifts), with
 no high-frequency line features, and without the attenuation at even
@@
 source for \scuba\ that is commonly used as a primary flux
-calibrator. The now standard constant-velocity daisy scan pattern was
-used (Section~\ref{sec:scan}), with a scan speed of 155
-arcsec\,sec$^{-1}$. We perform several different reductions of the
+calibrator. The CV daisy scan pattern was used, with a scan speed of
+155 arcsec\,sec$^{-1}$. We perform several different reductions of the
 data to illustrate the purpose of various model components and the
@@
 \scuba\ surveys designed to detect extremely faint point-sources
-(e.g. high-redshift star-forming galaxies, and debris disks) are
+(e.g., high-redshift star-forming galaxies, and debris disks) are
 ideally limited by the white-noise performance of the instrument. The
@@
 community using previous instruments
-\citep[e.g.][]{scott2002,borys2003,laurent2005,coppin2006,scott2008,perera2008,devlin2009}.
+\citep[e.g.,][]{scott2002,borys2003,laurent2005,coppin2006,scott2008,perera2008,devlin2009}.
 In this section we will reduce scans of the Lockman Hole taken during
@@
 brightnesses verify that the map-maker response to them is
-\emph{linear} (i.e. the relative shape and amplitude compared to the
+\emph{linear} (i.e., the relative shape and amplitude compared to the
 input source is independent of brightness). The response to a
@@
 correlation between the radio and submillimetre emission mechanisms
-\citep[e.g.][]{smail2000,pope2006,ivison2007,chapin2009b}. Taking a
+\citep[e.g.,][]{smail2000,pope2006,ivison2007,chapin2009b}. Taking a
 representative search radius of 8\,arcsec from these studies with
@@
 maps may be to detect \emph{previously unknown} cool, dense regions of
-the ISM that may not have appeared at other wavelengths (e.g. the
+the ISM that may not have appeared at other wavelengths (e.g., the
 first optically-thick cloud-collapse stages of star-formation). While
@@
 which is appropriate for Galactic cirrus clouds
-\citep[e.g.][]{gautier1992}. It is then filtered again with a
+\citep[e.g.,][]{gautier1992}. It is then filtered again with a
 14.5\,arcsec Gaussian to model the effect of the \scuba\ optical
@@
 as any point in the curve where $P_\mathrm{M}(k)$ exceeds
-$P_\mathrm{S}(k)$ (i.e. $G(K)$ is assumed to be $\le$1). As expected,
+$P_\mathrm{S}(k)$ (i.e., $G(K)$ is assumed to be $\le$1). As expected,
 the larger the scale of the filter, the lower the spatial frequency at
@@
 shows us that, even though the raw noise in the left panel is lower at
-small scales when a smaller-scale filter is used (e.g. the red dashed
+small scales when a smaller-scale filter is used (e.g., the red dashed
 line), once we correct for the transfer function, we actually win in a

### 2012-09-06 smurf_paper.tex
comments from Gary

@@
   data rate is much greater than that of previously existing
-  submillimetre cameras, and represents a significant data anlaysis
+  submillimetre cameras, and represents a significant data analysis
   challenge. We describe the production of \scuba\ maps using the
@@
   feedback to observers, as well as for the production of generic
-  science products at the Canadian Astronomical Data Centre. A suite
-  of tools are also provided for interactive data analysis.  Three
+  science products at the Canadian Astronomy Data Centre. A suite of
+  tools is also provided for interactive data analysis.  Three
   detailed case studies are used to: (i) explore convergence
@@
 
-%\textbf{Guideline for figures:}
-
-%\begin{itemize}
-%\item minimize excess whitespace around the figure to optimize usage of space
-%\item ensure a decent line thickness
-%\item use a Times font for labels with a thickness that resembles the
-%  text in the body. The font size should be similar to the main body
-%  text or larger
-%\item avoid using colour unless it is necessary for emphasis
-%\end{itemize}
-
-%\textbf{start intro here...}
-
 The Submillimetre Common User Bolometer Array 2
-\citep[\scuba,][]{2006SPIE.6275E..45H,holland2012} is a new instrument
-for the 15-m James Clerk Maxwell Telescope (JCMT) on Mauna Kea,
-Hawai'i. The camera simultaneously images the sky in two broad bands
-centered over 450 and 850\,\micron, with approximately 7.5 and
-14.5\,arcsec full-width at half-maximum (FWHM) point spread functions
-(PSFs). The focal planes at each wavelength are populated with 4
-rectangular subarrays, consisting of $40 \times 32$ bolometers each,
-and together subtend a nearly 7\,arcmin $\times$ 7\,arcmin footprint
-on the sky (excluding gaps, the continuous solid angle is about
-43\,arcmin$^{2}$ per focal plane).  The instrument is also equipped
-with a Fourier transform spectrometer \citep[FTS-2,][]{gom2010}, and a
-polarimeter \cite[POL-2,][]{2011ASPC..449...68B}. This paper describes
-the properties of \scuba\ data that are relevant to producing maps of
-the imaging data, and the Submillimetre User Reduction Facility,
-SMURF, a software package for perfoming the reduction written using
-the Starlink Software Environment \citep{2009ASPC..411..418J}. The
-analysis of FTS-2 and POL-2 data will be described at a later date
-once these additional instruments have been commissioned and
-characterized. The details of the instrument design, performance, and
-its calibration are given in two companion papers: \citet{holland2012}
-and \citet{dempsey2012}.
+\citep[\scuba,][]{holland2012} is a new instrument for the 15-m James
+Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
+simultaneously images the sky in two broad bands centered over 450 and
+850\,\micron, with approximately 7.5 and 14.5\,arcsec full-width at
+half-maximum (FWHM) point spread functions (PSFs). The focal planes at
+each wavelength are populated with 4 rectangular subarrays, consisting
+of $40 \times 32$ bolometers each, and together subtend a nearly
+7\,arcmin $\times$ 7\,arcmin footprint on the sky (excluding gaps, the
+continuous solid angle is about 43\,arcmin$^{2}$ per focal plane).
+This paper describes the properties of \scuba\ data that are relevant
+to producing maps of the imaging data, and the Submillimetre User
+Reduction Facility, SMURF, a software package for performing the
+reduction written using the Starlink Software Environment
+\citep{2009ASPC..411..418J}. The details of the instrument design,
+performance, and calibration are given in two companion papers:
+\citet{holland2012} and \citet{dempsey2012}.
 
@@
 several important areas of astrophysics, including: the
-characterisation of the early stages of star-formation by identifying the
-dense, cold regions in molecular clouds where stars may eventually
+characterisation of the early stages of star-formation by identifying
+the dense, cold regions in molecular clouds where stars may eventually
 form; discovering through blind surveys a class of massive dusty
@@
 as submillimetre galaxies, or SMGs; and identifying debris disks
-around nearby stars, helping us understand the early stages of planet formation.  With
-$\sim10000$ total detectors, \scuba\ is presently the largest
-submillimetre camera in the world, and the fastest wide-area
+around nearby stars, helping us understand the early stages of planet
+formation.  With $\sim10000$ total detectors, \scuba\ is presently the
+largest submillimetre camera in the world, and the fastest wide-area
 ground-based submillimetre imager.
@@
 instruments that had poorer low-frequency noise performance, and were
-only sensivitve to modest angular scales
+only sensitive to modest angular scales
 
 There are three general styles of map-making that are relevant to
-reducing bolomter data in the literature. The simplest ``direct
+reducing bolometer data in the literature. The simplest ``direct
 methods'' involve some basic processing of the data to remove as much
@@
 generation. Generally speaking, such methods are fast, although
-dependending on the science goals and noise properties of the data,
+depending on the science goals and noise properties of the data,
 they may not achieve the best noise performance on the angular scales
@@
 The best existing map-making strategies for recovering information on
-all angular scales are \emph{maximum likelihood} techniques, in which
-the time-series data are expressed as a sampling of the ``true'' map
-of the sky plus noise, and then the equation is inverted to estimate
-the map as some weighted linear combination of the data that minimizes
-the variance.  The first good description of this method appears in
+all angular scales are maximum likelihood techniques, in which the
+time-series data are expressed as a sampling of the ``true'' map of
+the sky plus noise, and then the equation is inverted to estimate the
+map as some weighted linear combination of the data that minimizes the
+variance.  The first good description of this method appears in
 \citet{janssen1992} for the production of maps from the COsmic
@@
 portion of the (predominantly low-frequency) non-white noise sources
-can be modeled, an iterative solution is obtained for both the
+can be modelled, an iterative solution is obtained for both the
 astronomical image and the parameters of the noise model. Since the
@@
 estimation and removal of this signal significantly lowers the $1/f$
-knee, without compromosing structures on angular scales smaller than
+knee, without compromising structures on angular scales smaller than
 the array footprint. Residual independent drifts at lower frequencies
@@
 products using the JCMT Science Archive \citep{2011ASPC..442..203E}
-hosted by the Canadian Astronomical Data Centre (CADC). Details of the
+hosted by the Canadian Astronomy Data Centre (CADC). Details of the
 pipeline design are given in \citet{gibb2005} and
@@
 (iii) the star-forming region M17, including bright, extended emission
-(Section~\ref{sec:extended}). All of the data analyzed in this paper
+(Section~\ref{sec:extended}). All of the data analysed in this paper
 are publicly available through the CADC \scuba\ raw-data queries page
@@
 In this section we summarize how \scuba\ works,
-(Section~\ref{sec:bolos}), give exampels of typical bolometers signals
+(Section~\ref{sec:bolos}), give examples of typical bolometers signals
 (Section~\ref{sec:bolosignal}), and the impact of magnetic field
@@
 of both the 450 and 850\,\micron\ focal planes. Each focal plan is
-populated by four \emph{subarrays} (labeled s4a--s4d at 450\,\micron,
-and s8a--s8d at 850\,\micron), each consisting of 32 columns and 40
-rows of bolometers. The bolometers are thermal absorbers coupled to
+populated by four ``subarrays'' (labeled s4a--s4d at 450\,\micron, and
+s8a--s8d at 850\,\micron), each consisting of 32 columns and 40 rows
+of bolometers. The bolometers are thermal absorbers coupled to
 superconducting transition-edge sensors (TESs) for thermometry;
@@
 that there is an additional 41st row of SQUID readouts that are not
-connected to TESs. These \emph{dark SQUIDs} track non-thermal noise
+connected to TESs. These ``dark SQUIDs'' track non-thermal noise
 sources that are common to each column's amplifier chain. The
 relationship between the output digitized current, $I$, and the input
-power, $P$, or $dI/dP$ is established using \emph{flatfield}
-observations immediately prior to science observations, in which the
-output signal is measured throughout a ramp of the pixel heaters
-(which provide a known input power); see Section~2.1 in
-\citet{dempsey2012} and Section~5.3 in \citet{holland2012}. Finally,
-the conversion to astronomical flux units from power involves a
-correction for atmospheric extinction \citep[primarily using the JCMT
-Water Vapour Monitor to track line-of-sight opacity variations, see
-Section~3 in][]{dempsey2012}, and the application of a flux conversion
-factor (FCF) which is established from regular measurements of
-calibrators such as Uranus \citep[Section~5 in][]{dempsey2012}.
+power, $P$, or $dI/dP$ is established using flatfield observations
+immediately prior to science observations, in which the output signal
+is measured throughout a ramp of the pixel heaters (which provide a
+known input power); see Section~2.1 in \citet{dempsey2012} and
+Section~5.3 in \citet{holland2012}. Finally, the conversion to
+astronomical flux units from power involves a correction for
+atmospheric extinction \citep[primarily using the JCMT Water Vapour
+Monitor to track line-of-sight opacity variations, see Section~3
+in][]{dempsey2012}, and the application of a flux conversion factor
+(FCF) which is established from regular measurements of calibrators
+such as Uranus \citep[Section~5 in][]{dempsey2012}.
 
@@
 
-We note that the noise performace of \scuba\ has evolved over time. An
+We note that the noise performance of \scuba\ has evolved over time. An
 initial ``\scuba\ shared-risk observing'' period (S2SRO) took place
@@
 upgrading and commissioning the instrument, a servo using newly added
-NTD thermometers in the focal planes has effectively mitigated this
+thermometers in the focal planes has effectively mitigated this
 problem \citep[Section~2.5 in][]{holland2012}. In addition, there were
@@
 (FWHM) -- a typical rule-of-thumb for adequately sampling a Gaussian
-point spread function (PSF). The scan \emph{patterns} are designed,
+point spread function (PSF). The scan patterns are designed,
 primarily, to mitigate the effects of low-frequency $1/f$ drift. The
@@
 i.e., a series of parallel passes across the field. However, such a
-scan yields extremely poor performace for structures that are
+scan yields extremely poor performance for structures that are
 transverse to the scan direction, since a long time will have elapsed
@@
 may conspire with such a regularly-spaced pattern to produce periodic
-structures in the map, or \emph{scan-synchronous} noise.  It has
-therefore been recognized that scan strategies that visit each region
-of the map on as many different time scales, and scan angles as
-possible, offer the best protection against large-scale noise
-\citep[see discussion in][and references therein]{kovacs2008b}; such
-strategies are said to offer good \emph{cross-linking}. As a tradeoff
-between truly random scan patterns and feasibility of implementation
-taking into consideration both mechanical and scheduling constraints,
-\scuba\ presently has two primary scan modes: PONG, which covers
-rectangular areas larger than the array footprint in a closed pattern
-that ``bounces'' off the edges at 45\,degree angles (and also usually
-combined with a rotation through a number of fixed position angles to
-create a ``rotating PONG'' with even better cross-linking) which was
-based on a similar scan mode developed at the Caltech Submillimeter
-Observatory; and a ``constant velocity daisy'' (CV Daisy) for small
-areas (of order the array footprint), in which the telescope moves in
-a circle, whose centre also slowly traces out a circle. For a more
-complete description of the \scuba\ observing modes, see Section~5 in
-\citet{holland2012}.
+structures in the map, or ``scan-synchronous'' noise. Scan strategies
+that visit each region of the map on as many different time scales,
+and scan angles as possible, offer the best protection against
+large-scale noise; such strategies are said to offer good
+cross-linking. As a trade-off between truly random scan patterns and
+feasibility of implementation taking into consideration both
+mechanical and scheduling constraints, \scuba\ presently has two
+primary scan modes: PONG, which covers rectangular areas larger than
+the array footprint in a closed pattern that ``bounces'' off the edges
+at 45\,degree angles (and also usually combined with a rotation
+through a number of fixed position angles to create a ``rotating
+PONG'' with even better cross-linking); and a ``constant velocity
+daisy'' (CV Daisy) for small areas (of order the array footprint), in
+which the telescope moves in a circle, whose centre also slowly traces
+out a circle. For a more complete description of the \scuba\ observing
+modes, see Section~5 in \citet{holland2012}.
 
@@
   and azimuth/elevation pointing offsets, before and after upgrading
-  the instrument. The grey signals overplotted in the top two panels
+  the instrument. The grey signals over-plotted in the top two panels
   show the residual time-series after removing the common-mode
@@
 is likely that these are not observed at the focal plane itself, where
-the NTD thermometry used in the servo loop is performed). There is
-minimal correlated low-frequency noise in these 450\,\micron\ data
-(note that there is almost no difference once the common-mode has been
-subtarcted). The 850\,\micron\ channel, however, exhibits a more
+the thermometry used in the servo loop is performed). There is minimal
+correlated low-frequency noise in these 450\,\micron\ data (note that
+there is almost no difference once the common-mode has been
+subtracted). The 850\,\micron\ channel, however, exhibits a more
 significant signal that is obviously correlated with the telescope
@@
 individual bolometer PSDs at high-frequency, as expected if the
-bolometers are dominated by un-correlated white noise in that part of
+bolometers are dominated by uncorrelated white noise in that part of
 the spectrum. The common-mode signals are fit to each bolometer time
@@
 An additional noise source that is significant in only a small subset
-of the data (and moreso during the S2SRO period) is magnetic field
+of the data (and more so during the S2SRO period) is magnetic field
 pickup. Since the bolometer signals are ultimately detected through
@@
 the bolometer and dark squid signals are detecting a noise source that
-is correlated only with the azimuthal motion and \emph{not} the
+is correlated only with the azimuthal motion and not the
 elevation. This behaviour would be expected if if there were a strong
@@
 diagonal), such that each of the bolometer time-series is a linear
-combination of the eigenvectors, or \emph{components}
-i.e., $\mathbf{b}_i(t) = \bar{\mathbf{\xi}}
-\mathbf{\lambda}_i^\mathrm{T}$, where each row of the matrix
-$\bar{\mathbf{\xi}}$ is an eigenvector, and
+combination of the eigenvectors, or components i.e., $\mathbf{b}_i(t)
+= \bar{\mathbf{\xi}} \mathbf{\lambda}_i^\mathrm{T}$, where each row of
+the matrix $\bar{\mathbf{\xi}}$ is an eigenvector, and
 $\mathbf{\lambda}_i^\mathrm{T}$ is a column vector containing the
@@
 evident at 850\,\micron) suggests that the scan-synchronous noise
-(presumably magentic field pickup) has a different response across the
+(presumably magnetic field pickup) has a different response across the
 focal plane; in this case the PCA may have identified two orthogonal
@@
 the details can vary significantly from data set to data set, and the
-lengths of the time-series analyzed. The reason for this is probably
+lengths of the time-series analysed. The reason for this is probably
 due to the fact that many of the noise sources are not stationary in
@@
 not necessarily identify clean patterns that can easily be
-modeled. For example, while there are clearly correlation patterns
+modelled. For example, while there are clearly correlation patterns
 along columns (Components~6 and 8--10, particularly at 450\,\micron),
@@
   re-estimated in turn, with the previous estimate first being added
-  back into the time-series (and importantly, \emph{not} the other
+  back into the time-series (and importantly, not the other
   components).}
@@
 low-frequency noise sources that are correlated amongst detectors, and
-to iterate this process along with estimates of the map. This sort of
-algorithm has been used successfully in the past with instruments
-including SCUBA \citep{1999ASPC..172..171J}, SHARC-2
-\citep{kovacs2008}, and BOLOCAM \citep{aguirre2010}, and is capable of
-recovering modest angular scales up to of order the footprint of the
-array (depending, of course, on the noise in the data). The main
-benefit of this approach over CMB-style maximum-likelihood methods
-(which are generally superior at recovering structure on all scales
-available in the data) is that the computing resources are
-significantly smaller, both in terms of memory usage and execution
-time. Given the massive volume of \scuba\ data (produced at a rate of
-approximately 10 GB\,hr$^{-1}$ in each focal plane), iterative
-techniques were the only viable option for providing real-time
-feedback at the telescope, and for automatically reducing near-science
-grade maps on a timescale comensurate with the observation times,
-using single high-end desktop computers.  We do note, however, that
-recent advances in direct inversion methods such as SANEPIC
-\citep{patanchon2008} may in fact be tenable for \scuba\ data. A
-future goal of the software team is to investigate meshing SMURF
-(perhaps as a cleaning step and initial condition) with a SANEPIC-like
-algorithm.
-
-Significant experimentation with different models for noise sources
-during \scuba\ commissioning required a highly flexible software
-framework, and a configurable user interface. To achieve this goal,
-while minimizing development time, it was decided to build SMURF as a
-Starlink package \citep{2009ASPC..411..418J}, which provides access to
-a large suite of libraries (including a commanding and messaging
-interface, astrometric coordinate conversions and generation of
-standard WCS information, file formats etc.). Furthermore, Starlink is
-open source, and already used extensively at the Joint Astronomy
-Centre (host of the JCMT) for many other systems, which helps with
-interoperability. Though originally written in FORTRAN, many of the
-core Starlink libraries are now ported to native C, or at least have a
-C interface. It was therefore decided to develop SMURF in C as well
-(rather than C++, for example, which would have required adding
-further dependencies to Starlink), although we have taken an
-object-oriented approach. For example, all of the data for a given
-noise model is encapsulated in a C structure, and there is a standard
-interface for all functions that handle models (member data and
-functions for the class, respectively). In this way it is easy to
-extend SMURF with new models. Parallelisation is incorporated in the
-most time-consuming low-level routines using threads (e.g., performing
-Fast Fourier Transforms, re-gridding the data), usually handling
-either independent blocks of bolometers, or blocks of time, in each
-thread, depending on the nature of the
-calculation. \textbf{... something about scalability ...}
+to iterate this process along with estimates of the map. Significant
+experimentation with different models for noise sources during \scuba\
+commissioning required a highly flexible software framework, and a
+configurable user interface. To achieve this goal, while minimizing
+development time, it was decided to build SMURF as a Starlink package
+\citep{2009ASPC..411..418J}, which provides access to a large suite of
+libraries (including a commanding and messaging interface, astrometric
+coordinate conversions and generation of standard WCS information,
+file formats etc.). Furthermore, Starlink is open source, and already
+used extensively at the Joint Astronomy Centre (host of the JCMT) for
+many other systems, which helps with interoperability. Though
+originally written in FORTRAN, many of the core Starlink libraries are
+now ported to native C, or at least have a C interface. It was
+therefore decided to develop SMURF in C as well (rather than C++, for
+example, which would have required adding further dependencies to
+Starlink), although we have taken an object-oriented approach. For
+example, all of the data for a given noise model is encapsulated in a
+C structure, and there is a standard interface for all functions that
+handle models (member data and functions for the class,
+respectively). In this way it is easy to extend SMURF with new
+models. Parallelisation is incorporated in the most time-consuming
+low-level routines using threads (e.g., performing Fast Fourier
+Transforms, re-gridding the data), usually handling either independent
+blocks of bolometers, or blocks of time, in each thread, depending on
+the nature of the calculation. \textbf{... something about scalability
+...}
 
@@
 once it is re-estimated. Note that the extinction and gain denoted in
-a second dashed box are \emph{multiplicative} rather than additive
-models, and so they must be inverted prior to the start of each
-iteration to preserve the units (see Sections~\ref{sec:comgai} and
-\ref{sec:ext}).
+a second dashed box are multiplicative rather than additive models,
+and so they must be inverted prior to the start of each iteration to
+preserve the units (see Sections~\ref{sec:comgai} and \ref{sec:ext}).
 
@@
 components, under some circumstances it may be desirable to filter the
-data \emph{once} during the pre-processing step. Three main filtering
-options are available:
+data once during the pre-processing step. Three main filtering options
+are available:
 
@@
 away from the mean value. Subarrays are cleaned independently, once,
-for the full length of each continuousy chunk of data. Given the
+for the full length of each continuous chunk of data. Given the
 computational expense of this method, and initial tests which showed
@@
 describes the iterative detection and removal of spikes as part of map
-estimation. In practice, map-based despiking usually gives superior
-results, and so time-domain despiking is switched off by default.
+estimation. In practice, map-based de-spiking usually gives superior
+results, and so time-domain de-spiking is switched off by default.
 
@@
 
-\item the above process can mis-interpret the edges of a bright source
+\item the above process can misinterpret the edges of a bright source
 as a step, so we ignore blocks that occur close to bright sources;
@@
 
-\subsubsection{Gap filling / apodization}
+\subsubsection{Gap filling / apodisation}
 \label{sec:gaps}
@@
 \caption{The black curve shows the same bolometer time-series as in
-Fig.~\ref{fig:pad2}, againpadded with zeros. The red curve shows the
+Fig.~\ref{fig:pad2}, again padded with zeros. The red curve shows the
 time-series after padding with artificial data.}
@@
 where $\mathbf{n}^\mathrm{w}_i(t)$ is uncorrelated white noise,
-$\mathbf{n}^\mathrm{c}(t)$ is a correlated or \emph{common-mode}
-signal (with an optional scale factor $g_i$ for each bolometer), and
+$\mathbf{n}^\mathrm{c}(t)$ is a correlated or common-mode signal (with
+an optional scale factor $g_i$ for each bolometer), and
 $\mathbf{n}^\mathrm{f}_i(t)$ is (predominantly low-frequency) noise in
@@
 extinction, have been corrected for. Provided that these two signals
-are the only components of the residual bolometer signal, regridding
+are the only components of the residual bolometer signal, re-gridding
 into a map provides the highest \snr\ estimate of the astronomical
@@
 (which is a simple average of all of the bolometer time-series at each
-instant in time), leading to an \emph{over-subtraction}, and in turn,
+instant in time), leading to an over-subtraction, and in turn,
 negative bowls in the map.
@@
 they are calculated. For a discussion on convergence tests and
-degenerices, see Section~\ref{sec:converge}.
+degeneracies, see Section~\ref{sec:converge}.
 
@@
 The extinction correction is a multiplicative factor that is normally
-derived using the JCMT water vapour radiometer (WVM), and is not
+derived using the JCMT water vapour monitor (WVM), and is not
 considered to be a free parameter in the solution ($\mathbf{e}_i(t)$
@@
 hard frequency edges specified by the user. Alternatively, the
-frequency edges of the filters may be defined in terms of an
-\emph{angular scale}, but converted into a frequency through knowledge
-of the mean telescope slew speed. The time-series may optionally have
-apodization applied before the transform to avoid ringing (primarily
-caused by wrap-around discontinuities at the ends of the
-time-series). Finally, a whitening filter may also be applied in which
-a simple $1/f^\alpha + \mathrm{constant}$ is fit to the power spectrum
-of each bolometer, and then the bolometer FFT is multiplied by its
-inverse (though normalized to preserve the white noise level).  The
-signals that are \emph{removed} from the time-series by this process
-are stored in the \model{FLT} container array. Typically this model is
-used purely as a high-pass filter to remove most of the residual
-uncorrelated $1/f$ noise following common-mode removal
-($\mathbf{n}^\mathrm{f}_i(t)$ in Eq.~\ref{eq:noise}). Low-pass
-filtering is redundant for two reasons: (i) SMURF already low-pass
-filters and re-samples to a lower sample rate, as described in
-Section~\ref{sec:downsamp}; and (ii) the act of re-gridding the data
-to produce map estimates effectively low-pass filters the data at a
-frequency that corresponds to the inverse of the crossing time of a
-single map pixel. Notch filters have not been proven to be useful with
-\scuba\ data, particularly since line features tend to move around
-(i.e., a dynamic line-detection system would have to be developed, and
-care would have to be taken that astronomical sources are not
-suppressed).
+frequency edges of the filters may be defined in terms of an angular
+scale, but converted into a frequency through knowledge of the mean
+telescope slew speed. The time-series may optionally have apodisation
+applied before the transform to avoid ringing (primarily caused by
+wrap-around discontinuities at the ends of the time-series). Finally,
+a whitening filter may also be applied in which a simple $1/f^\alpha +
+\mathrm{constant}$ is fit to the power spectrum of each bolometer, and
+then the bolometer FFT is multiplied by its inverse (though normalized
+to preserve the white noise level).  The signals that are removed from
+the time-series by this process are stored in the \model{FLT}
+container array. Typically this model is used purely as a high-pass
+filter to remove most of the residual uncorrelated $1/f$ noise
+following common-mode removal ($\mathbf{n}^\mathrm{f}_i(t)$ in
+Eq.~\ref{eq:noise}). Low-pass filtering is redundant for two reasons:
+(i) SMURF already low-pass filters and re-samples to a lower sample
+rate, as described in Section~\ref{sec:downsamp}; and (ii) the act of
+re-gridding the data to produce map estimates effectively low-pass
+filters the data at a frequency that corresponds to the inverse of the
+crossing time of a single map pixel. Notch filters have not been
+proven to be useful with \scuba\ data, particularly since line
+features tend to move around (i.e., a dynamic line-detection system
+would have to be developed, and care would have to be taken that
+astronomical sources are not suppressed).
 
@@
 the mean given the scatter in the weighted samples. This is
-accomplished by dividing the \emph{biased weighted sample variance} by
-the number of samples that went into the average (akin to the formula
-for standard error on the mean, but accounting for weights),
+accomplished by dividing the biased weighted sample variance by the
+number of samples that went into the average (akin to the formula for
+standard error on the mean, but accounting for weights),
 %
@@
 
- We decided not to use an \emph{unbiased} estimator (e.g., the
-extension of the common $(1/N-1)\sum_j (\mathbf{b}_j -
-\bar{\mathbf{b}})^2$ estimator using weights), since in practice it
-would require accumulating an additional array of values at every map
-pixel, and only results in a small difference where there are less
-than $\sim10$ samples per pixel (a situation that is almost never
-encountered in a \scuba\ map except in the edge pixels).
+We decided not to use an unbiased estimator (e.g., the extension of
+the common $(1/N-1)\sum_j (\mathbf{b}_j - \bar{\mathbf{b}})^2$
+estimator using weights), since in practice it would require
+accumulating an additional array of values at every map pixel, and
+only results in a small difference where there are less than $\sim10$
+samples per pixel (a situation that is almost never encountered in a
+\scuba\ map except in the edge pixels).
 
@@
 %
-In order to compare the weighted \emph{differences} between the
-samples and the map values, [$\mathbf{b}_j - \mathbf{m}(x_i,y_i)$] to
+In order to compare the weighted differences between the samples and
+the map values, [$\mathbf{b}_j - \mathbf{m}(x_i,y_i)$] to
 $\mathbf{v_p}$, they must be scaled appropriately. We define a
@@
 Finally, it should be noted that the default behaviour is to calculate
-the white noise levels \emph{once} within \model{NOI}, after the second
+the white noise levels once within \model{NOI}, after the second
 iteration. The reason for fixing these values is to prevent any
@@
 the measured time-series bolometer variances and their white-noise
-levels measured between 2 and 10\,Hz. Note that this expression
-\emph{does not account for the degrees of freedom}. Clearly there are
-a large number of parameters in the model, which will ``fit-out'' some
-of the uncorrelated white noise, and potentially bias this estimate of
+levels measured between 2 and 10\,Hz. Note that this expression does
+not account for the degrees of freedom. Clearly there are a large
+number of parameters in the model, which will ``fit-out'' some of the
+uncorrelated white noise, and potentially bias this estimate of
 $\chi^2_\mathrm{r}$ low. One approach would be to keep track of the
@@
 bolometer after 2 (black) and 100 (grey) iterations which are nearly
-identical, yet the \emph{change} in the estimated \model{COM} (green)
-and \model{AST} (red) signals between 2 and 100 iterations are
+identical, yet the change in the estimated \model{COM} (green) and
+\model{AST} (red) signals between 2 and 100 iterations are
 large. However, it is also clear that the estimated \model{COM} and
-\model{AST} signals are \emph{complimentary}. In other words, the
-large change in the \model{AST} signal is cancelled by freedom in the
+\model{AST} signals are complimentary. In other words, the large
+change in the \model{AST} signal is cancelled by freedom in the
 \model{COM} signal to grow with opposite sign. For comparison, the
@@
   identical; the black line lies beneath the grey line). The green and
-  red lines show the \emph{difference} between the \model{COM} and
+  red lines show the difference between the \model{COM} and
   \model{AST} (the signal produced by the current map estimate for a
@@
 low-frequency noise sources with approximately linear response,
-without prior knowledge of the location of sources; (ii)
-\emph{whitening} (essentially flattening) the map to make the noise
-properties of the map easier to deal with; and (iii) detecting point
-sources using a \emph{matched-filter}. Note that variations on this
-general procedure have been used extensively in the submm cosmology
-community using previous instruments
+without prior knowledge of the location of sources; (ii) whitening
+(essentially flattening) the map to make the noise properties of the
+map easier to deal with; and (iii) detecting point sources using a
+``matched filter''. Note that variations on this general procedure
+have been used extensively in the submm cosmology community using
+previous instruments
 \citep[e.g.,][]{scott2002,borys2003,laurent2005,coppin2006,scott2008,perera2008,devlin2009}.
@@
   maps, using the even and odd scan numbers, respectively. Provided
-  that all noise sources are statistically un-correlated between the
+  that all noise sources are statistically uncorrelated between the
   two halves of the data, the map is a plausible realization of the
@@
 mitigated, and the method used in this example (the default processing
-in SMURF) is to apply a high-pass filter to the data \emph{once} as a
+in SMURF) is to apply a high-pass filter to the data once as a
 pre-processing step. The iterative solution is then run using only
@@
 structure; and (ii) adding synthetic sources to the data at a range of
-brightnesses verify that the map-maker response to them is
-\emph{linear} (i.e., the relative shape and amplitude compared to the
-input source is independent of brightness). The response to a
-synthetic point source (solid line) after map-making (dotted line) is
-shown in Fig.~\ref{fig:lockman_psf}. Clearly the use of a high-pass
-filter as a pre-processing step, and having no other map-constraints,
-has the down-side of introducing sidelobes around the main
-peak. However, the way this filter affects point-sources is both
-well-known, and linear.
+brightnesses verify that the map-maker response to them is linear
+(i.e., the relative shape and amplitude compared to the input source
+is independent of brightness). The response to a synthetic point
+source (solid line) after map-making (dotted line) is shown in
+Fig.~\ref{fig:lockman_psf}. Clearly the use of a high-pass filter as a
+pre-processing step, and having no other map-constraints, has the
+down-side of introducing sidelobes around the main peak. However, the
+way this filter affects point-sources is both well-known, and linear.
 
@@
 First, the angular power spectrum of noise is estimated from a
-\emph{jackknife map}: maps are produced from two independent halves of
+``jackknife map'': maps are produced from two independent halves of
 the total data set, and the jackknife signal in a map pixel,
@@
 
-To suppress noise in the map, we construct a \emph{whitening filter}
-from the reciprocal of the jackknife angular power spectrum (orange
-line in Fig.~\ref{fig:lockman_pspec}), normalized by the white-noise
-level estimated from the RMS power at angular frequencies $>
+To suppress noise in the map, we construct a whitening filter from the
+reciprocal of the jackknife angular power spectrum (orange line in
+Fig.~\ref{fig:lockman_pspec}), normalized by the white-noise level
+estimated from the RMS power at angular frequencies $>
 0.1$\,arcsec$^{-1}$. The filter is applied by scaling the power
@@
 ($\lsim 0.5$\,arcsec$^{-1}$) there is significantly less power than
-the raw map. However, it also clearly has \emph{some} power in excess
-of the white noise level. In theory, this residual signal is produced
-by astronomical sources, although its origin cannot be determined from
+the raw map. However, it also clearly has some power in excess of the
+white noise level. In theory, this residual signal is produced by
+astronomical sources, although its origin cannot be determined from
 this plot alone; nor are astronomical sources readily visible in the
 map (Fig.~\ref{fig:lockman_maps}b). To identify sources, we next apply
-a \emph{matched filter} to the whitened maps.
+a matched filter to the whitened maps.
 
@@
 photometric measurement strategy. In terms of the angular power
-spectrum, a matched-filter may also be thought of as an optimal
-\emph{low-pass} filter, that suppresses noise on scales smaller (and
+spectrum, a matched filter may also be thought of as an optimal
+low-pass filter, that suppresses noise on scales smaller (and
 therefore higher angular frequencies) than the beam. For the case at
@@
 
-Have real astronomical sources been detected using the matched-filter?
+Have real astronomical sources been detected using the matched filter?
 For both the smoothed signal and jackknife maps, orange circles denote
@@
 now deviates significantly from a Gaussian distribution, with a clear
-\emph{positive} tail (as one would expect for emitting sources). In
-fact, integrating the positive tails beyond our 3.8-$\sigma$
+positive tail (as one would expect for emitting sources). In fact,
+integrating the positive tails beyond our 3.8-$\sigma$
 source-detection threshold (vertical solid line) yields 229 map pixels
@@
   good agreement indicates that these maps are indeed dominated by
-  white noise. The lower panel shows the results for matched-filtered
+  white noise. The lower panel shows the results for matched filtered
   signal and jackknife maps (Figs.~\ref{fig:lockman_maps}c,f). The
   filtered jackknife map distribution is still close to the
-  expectation (Gaussian), but now the matched-filter has picked out
+  expectation (Gaussian), but now the matched filter has picked out
   significant signal, leading to the large positive tail. The vertical
@@
 PSF-smoothed map is dominated by point-source confusion, rather than
-instrumental noise, a modified matched-filter will offer improved
+instrumental noise, a modified matched filter will offer improved
 results. See Appendix~A in \citet{chapin2011}, which shows how to
@@
 clearly contains extended structure. Furthermore, the goal of such
-maps may be to detect \emph{previously unknown} cool, dense regions of
-the ISM that may not have appeared at other wavelengths (e.g., the
-first optically-thick cloud-collapse stages of star-formation). While
-the option does exist for the user to supply their own mask, a
-facility has been added to SMURF to generate one automatically by
-flagging pixels above some \snr\ threshold after each iteration as part
-of the ``bright extended'' configuration.
+maps may be to detect previously unknown cool, dense regions of the
+ISM that may not have appeared at other wavelengths (e.g., the first
+optically-thick cloud-collapse stages of star-formation). While the
+option does exist for the user to supply their own mask, a facility
+has been added to SMURF to generate one automatically by flagging
+pixels above some \snr\ threshold after each iteration as part of the
+``bright extended'' configuration.
 
@@
 edge scale was adopted (indicated in the top panels). First (top) row:
-average of two halves of the data analyzed independently, using the
+average of two halves of the data analysed independently, using the
 default configuration. The data have had added to them synthetic
@@
 transfer function as the ratio between the portion of the signal PSDs
-\emph{not} produced by noise, to the input PSD, or
+not produced by noise, to the input PSD, or
 %
@@
 
-One case in which the \snr\ is \emph{worse} using the bright extended
+One case in which the \snr\ is worse using the bright extended
 reduction is when using a 150\,arcsec filter. In this case the noise
@@
 filtering can provide misleading results as in this example. Maps of
-\emph{faint} extended emission will also suffer considerably, as the
+faint extended emission will also suffer considerably, as the
 structures of interest may not clear the \snr\ threshold for the mask.
@@
 during observations, and it is also regularly produces nearly
-science-grade products at the Canadian Astronomical Data Centre.
+science-grade products at the Canadian Astronomy Data Centre.
 

### 2012-09-06 smurf_paper.tex
typo

@@
   particularly at 450\,\micron). Finally, this example shows that the
-  white noise performance is in factor similar for the two subarrays
+  white noise performance is in fact similar for the two subarrays
   (s4a and s8b) that were used both before and after the upgrades.}

### 2012-09-06 smurf_paper.tex
Explicitly reference the pong paper
Frossie insists on being last author

@@
     Douglas~Scott$^{1}$,
+    Remo~P.~J.~Tilanus$^{2,3}$,
     Frossie~Economou$^2$\thanks{Present address: National Optical
       Astronomy Observatory, 950 N.\ Cherry Avenue, Tucson, AZ 85719, USA},
-    Remo~P.~J.~Tilanus$^{2,3}$
   }
@@
 out a circle. For a more complete description of the \scuba\ observing
-modes, see Section~5 in \citet{holland2012}.
+modes, see Section~5 in \citet{holland2012} and also \citet{2010SPIE.7740E..66K}.
 

### 2012-09-06 smurf_paper.tex
Add a bit of gravitas to the Starlink citation
explicitly mention kapuahi release
No reference to it as such since that would only be
(Berry et al, 2013, in preparation)
Use hyperref and add some URLs. Also add a final reference

@@
 \documentclass[useAMS,usenatbib,nofootinbib]{mn2e}
+
+%% Recommended by MN for handling hyperlinks
+\usepackage[pdftex,pdfpagemode={UseOutlines},bookmarks,bookmarksopen,colorlinks,linkcolor={blue},citecolor={green},urlcolor={red}]{hyperref}
 
@@
 reduction written using the Starlink Software Environment
-\citep{2009ASPC..411..418J}. The details of the instrument design,
+\citep{1993ASPC...52..229W,2009ASPC..411..418J}. The details of the instrument design,
 performance, and calibration are given in two companion papers:
@@
 for the dates and observation numbers given in the
-text\footnote{\texttt{http://www3.cadc-ccda.hia-iha.nrc-cnrc.gc.ca/
-jcmt/search/scuba2}}. All of the analysis was performed using
-\texttt{Starlink} version ... .
+text\footnote{\url{http://www.cadc-ccda.hia-iha.nrc-cnrc.gc.ca/jcmt/search/scuba2}}.
+All of the analysis was performed using the Starlink \textsc{kapuahi}
+release from 2012.
 
@@
 coordinate conversions and generation of standard WCS information,
-file formats etc.). Furthermore, Starlink is open source, and already
-used extensively at the Joint Astronomy Centre (host of the JCMT) for
-many other systems, which helps with interoperability. Though
+file formats etc.). Furthermore, Starlink\footnote{\url{http://www.starlink.ac.uk}} is
+open-source\footnote{\url{https://github.com/Starlink}},
+and already used extensively at the Joint Astronomy Centre (host of the JCMT) for
+many other systems \citep{jenness2011}, which helps with interoperability. Though
 originally written in FORTRAN, many of the core Starlink libraries are

### 2012-09-07 smurf_paper.tex
more updates from Douglas up to PCA section + acknowledgements

@@
   consisting of about 5000 bolometers in each of two simultaneous
-  imaging bands centered over 450 and 850\,\micron. The camera is
-  scanned across the sky and records data at a rate of 200\,Hz. The
-  data rate is much greater than that of previously existing
-  submillimetre cameras, and represents a significant data analysis
-  challenge. We describe the production of \scuba\ maps using the
-  Sub-Millimetre User Reduction Facility (SMURF) in which we have
+  imaging bands centred over 450 and 850\,\micron. The camera is
+  operated by scanning across the sky and records data at a rate of
+  200\,Hz. The data rate is much greater than that of previously
+  existing submillimetre cameras, and represents a significant
+  analysis challenge. We describe the production of \scuba\ maps using
+  the Sub-Millimetre User Reduction Facility (SMURF) in which we have
   adopted a fast, iterative approach to map-making that enables data
-  reduction on single, modern, high-end desktop computers with
+  reduction on single, modern, high-end desktop computers, with
   execution times that are shorter than the observing times.  SMURF is
-  used in an automated setting both at the telescope for real-time
+  used in an automated setting, both at the telescope for real-time
   feedback to observers, as well as for the production of generic
@@
 \begin{keywords}
-chicken, chicken: chicken, chicken: chicken -- chicken
+methods: data analysis, techniques: image processing, submillimetre:
+general, methods: observational
 \end{keywords}
@@
 Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
-simultaneously images the sky in two broad bands centered over 450 and
+simultaneously images the sky in two broad bands centred over 450 and
 850\,\micron, with approximately 7.5 and 14.5\,arcsec full-width at
@@
 This paper describes the properties of \scuba\ data that are relevant
-to producing maps of the imaging data, and the Submillimetre User
+for producing maps of the imaging data, and the Submillimetre User
 Reduction Facility, SMURF, a software package for performing the
@@
 around nearby stars, helping us understand the early stages of planet
-formation.  With $\sim10000$ total detectors, \scuba\ is presently the
+formation.  With $\sim10,000$ total detectors, \scuba\ is presently the
 largest submillimetre camera in the world, and the fastest wide-area
@@
 instrument and ambient backgrounds. The low-frequency noise, however,
-is typically dominated by sources which produce slow variations the
+is typically dominated by sources which produce slow variations in the
 background (e.g., thermal variations within the cryostat, and the
@@
 ground-based submillimetre cameras
-\citep[e.g.,][]{glenn1998,weferling2002,wilson2008,kovacs2008b}., in
-favour of ``chopping'' methods (where the secondary is quickly moved
-to modulate the signal) that were more appropriate for older
+\citep[e.g.,][]{glenn1998,weferling2002,wilson2008,kovacs2008b}, in
+preference to ``chopping'' methods (where the secondary is quickly
+moved to modulate the signal) that were more appropriate for older
 instruments that had poorer low-frequency noise performance, and were
@@
 generation. Generally speaking, such methods are fast, although
-depending on the science goals and noise properties of the data,
-they may not achieve the best noise performance on the angular scales
-of interest. In more recent years a method that has been popular for
+depending on the science goals and noise properties of the data, they
+may not achieve the best noise performance on the angular scales of
+interest. In more recent years a method that has been popular for
 reducing data in fields of faint point sources with BOLOCAM
@@
 \citep[e.g.,][]{scott2008}, is ``principal component analysis'' (PCA)
-cleaning. A statistical black-box removes the most correlated
+cleaning. A statistical ``black-box'' removes the most correlated
 components of the bolometer signals, enabling the detection of
 point-sources very close to the theoretical white-noise limits of the
-detectors with reasonable computation times when small numbers of
+detectors, with reasonable computation times when small numbers of
 bolometers are involved (hundreds rather than thousands of
@@
 the sky plus noise, and then the equation is inverted to estimate the
-map as some weighted linear combination of the data that minimizes the
+map as some weighted linear combination of the data that minimises the
 variance.  The first good description of this method appears in
@@
 both by the need to measure the cross power-spectra for all unique
-pairs of bolometers, and then to perform the inversion itself. The
+pairs of bolometers, as well as performing the inversion itself. The
 most promising maximum-likelihood method that may one day be applied
@@
 the Balloon-borne Large-Aperture Submillimeter Telescope, involving
-hundreds of bolometers, and correctly incorporating inter-bolometer
+hundreds of bolometers, while correctly incorporating inter-bolometer
 noise correlations \citep[][]{patanchon2008}.
@@
 a given bolometer (since the noise at any instant in time is
-uncorrelated with others, nor with the data from other bolometers),
+uncorrelated with others, and with the data from other bolometers),
 greatly simplifying the measurement of noise properties and the
@@
 methods. In principle, it should also be possible to recover a wider
-dynamic range in angular scales than the simplest ``direct'' methods,
-with an increase in calculation time that is merely linear in the
-number of iterations needed, while using a comparable amount of
+dynamic range in angular scales than possible for the simplest direct
+methods, with an increase in calculation time that is merely linear in
+the number of iterations needed, while using a comparable amount of
 memory. A fast iterative method for handling chopped data from CMB
@@
 
-This paper is organized in the following way. We first describe the
-properties of \scuba\ data in Section~\ref{sec:data} to motivate the
-data model that we have adopted. Next, the details of the SMURF
-algorithm are given in Section~\ref{sec:algorithm}. The paper is
-concluded in Section~\ref{sec:examples} with three detailed test cases
-that span the majority of observation types likely to be undertaken
-with \scuba, with an emphasis on the mitigation of divergence
-problems, and characterizing the output maps: (i) Uranus, a bright,
-compact source (Section~\ref{sec:point}); (ii) the Lockman Hole, a
-blind survey of faint point-like source (Section~\ref{sec:cosmo}); and
-(iii) the star-forming region M17, including bright, extended emission
+This paper is organised as follows. We first describe the properties
+of \scuba\ data in Section~\ref{sec:data} to motivate the data model
+that we have adopted. Next, the details of the SMURF algorithm are
+given in Section~\ref{sec:algorithm}. The paper is concluded in
+Section~\ref{sec:examples} with three detailed test cases that span
+the majority of observation types likely to be undertaken with \scuba,
+with an emphasis on the mitigation of divergence problems, and
+characterising the output maps: (i) Uranus, a bright, compact source
+(Section~\ref{sec:point}); (ii) the Lockman Hole, a blind survey of
+faint point-like sources (Section~\ref{sec:cosmo}); and (iii) the
+star-forming region M17, including bright, extended emission
 (Section~\ref{sec:extended}). All of the data analysed in this paper
@@
 
-In this section we summarize how \scuba\ works,
+In this section we summarise how \scuba\ works,
 (Section~\ref{sec:bolos}), give examples of typical bolometers signals
@@
 \citet{holland2012}, and its calibration in \citet{dempsey2012}, we
-summarize the basic concepts relevant to map-making here.
+summarise the basic concepts relevant to map-making here.
 
@@
 interference devices (SQUIDs), and the larger output currents are
-digitized. Each detector has its own SQUID for the first-stage of the
+digitised. Each detector has its own SQUID for the first-stage of the
 amplification, but the remaining stages occur within a common chain of
-SQUIDs for each of the 32 columns. All 40 rows are read-out in
+SQUIDs for each of the 32 columns. All 40 rows are read out in
 sequence at a row-visit frequency of about 12\,kHz. Such a high sample
@@
 known input power); see Section~2.1 in \citet{dempsey2012} and
-Section~5.3 in \citet{holland2012}. Finally, the conversion to
-astronomical flux units from power involves a correction for
-atmospheric extinction \citep[primarily using the JCMT Water Vapour
-Monitor to track line-of-sight opacity variations, see Section~3
-in][]{dempsey2012}, and the application of a flux conversion factor
-(FCF) which is established from regular measurements of calibrators
-such as Uranus \citep[Section~5 in][]{dempsey2012}.
+Section~5.3 in \citet{holland2012} for more details. Finally, the
+conversion to astronomical flux units from power involves a correction
+for atmospheric extinction \citep[primarily using the JCMT Water
+Vapour Monitor to track line-of-sight opacity variations, see
+Section~3 in][]{dempsey2012}, and the application of a flux conversion
+factor (FCF) which is established from regular measurements of
+calibrators such as Uranus \citep[Section~5 in][]{dempsey2012}.
 
@@
 relatively flat at higher frequencies. The spectra also include line
-features which are predominantly produced by aliased high-frequency
-noise sources detected during the multiplexed readout stage as
-described above.  Finally, an anti-aliasing filter produces roll-off
-in the spectrum at frequencies approaching Nyquist prior to the
-200\,Hz re-sampling.
-
-We note that the noise performance of \scuba\ has evolved over time. An
-initial ``\scuba\ shared-risk observing'' period (S2SRO) took place
-during February and March 2010, during which each of the 450 and
+features which are mainly produced by aliased high-frequency noise
+sources detected during the multiplexed readout stage, as described
+above.  Finally, an anti-aliasing filter produces roll-off in the
+spectrum at frequencies approaching Nyquist prior to the 200\,Hz
+re-sampling.
+
+We note that the noise performance of \scuba\ has evolved over
+time. An initial ``\scuba\ shared-risk observing'' period (S2SRO) took
+place during February and March 2010, during which each of the 450 and
 850\,\micron\ focal planes were populated with single subarrays, s4a
-and s8d respectively. In addition to having significantly fewer
+and s8d, respectively. In addition to having significantly fewer
 available bolometers than the current fully-commissioned instrument
 (and therefore a mapping speed reduced by a factor of $\sim$4),
-instabilities in the fridge lead to a large periodic signal with a
+instabilities in the fridge led to a large periodic signal with a
 period of about 25\,s that was correlated amongst the detectors. After
@@
 problem \citep[Section~2.5 in][]{holland2012}. In addition, there were
-improvements to the magnetic shielding (reducing magnetic field pickup
-as described in Section~\ref{sec:magpickup}), as well as more
+improvements to the magnetic shielding (reducing magnetic field
+pickup, as described in Section~\ref{sec:magpickup}), as well as more
 effective removal of aliased noise sources (which has reduced both the
@@
 slew rates, which are practically limited to about
-600\,arcsec\,s$^{-2}$ for the JCMT in order to have reasonable
+600\,arcsec\,s$^{-1}$ for the JCMT in order to have reasonable
 turn-around overheads (the amount of time spent changing direction and
@@
 structures in the map, or ``scan-synchronous'' noise. Scan strategies
-that visit each region of the map on as many different time scales,
-and scan angles as possible, offer the best protection against
-large-scale noise; such strategies are said to offer good
-cross-linking. As a trade-off between truly random scan patterns and
-feasibility of implementation taking into consideration both
-mechanical and scheduling constraints, \scuba\ presently has two
-primary scan modes: PONG, which covers rectangular areas larger than
-the array footprint in a closed pattern that ``bounces'' off the edges
-at 45\,degree angles (and also usually combined with a rotation
-through a number of fixed position angles to create a ``rotating
-PONG'' with even better cross-linking); and a ``constant velocity
-daisy'' (CV Daisy) for small areas (of order the array footprint), in
-which the telescope moves in a circle, whose centre also slowly traces
-out a circle. For a more complete description of the \scuba\ observing
-modes, see Section~5 in \citet{holland2012} and also \citet{2010SPIE.7740E..66K}.
+that visit each region of the map on as many different time scales and
+scan angles as possible, offer the best protection against large-scale
+noise; such strategies are said to offer good cross-linking. As a
+trade-off between truly random scan patterns and feasibility of
+implementation, taking into consideration both mechanical and
+scheduling constraints, \scuba\ presently has two primary scan modes:
+PONG, which covers rectangular areas larger than the array footprint
+in a closed pattern that ``bounces'' off the edges at 45\,degree
+angles (and also usually combined with a rotation through a number of
+fixed position angles to create a ``rotating PONG'' with even better
+cross-linking); and a ``constant velocity daisy'' (CV Daisy) for small
+areas (of order the array footprint), in which the telescope moves in
+a circle, whose centre also slowly traces out a circle. For a more
+complete description of the \scuba\ observing modes, see Section~5 in
+\citet{holland2012} and also \citet{2010SPIE.7740E..66K}.
 
@@
 problem, these data no longer exhibit a correlation with the mixing
-chamber (note that variations are still seen in the mixing chamber; it
-is likely that these are not observed at the focal plane itself, where
-the thermometry used in the servo loop is performed). There is minimal
-correlated low-frequency noise in these 450\,\micron\ data (note that
-there is almost no difference once the common-mode has been
-subtracted). The 850\,\micron\ channel, however, exhibits a more
-significant signal that is obviously correlated with the telescope
-motion. It is also correlated amongst many of the detectors, and
-common-mode removal corrects it to some extent. Part of the reason
-that this is seen here, and not in the S2SRO data set shown, is that
-the amplitude of the scan pattern is larger. Most of this
-``scan-synchronous'' noise is attributed to magnetic field pickup, as
-described in Section~\ref{sec:magpickup}.
-
-Next, we produce power spectral density (PSD) plots for four of the
-most sensitive bolometers from both focal planes, using the same two
-data sets, in Fig.~\ref{fig:pspec}. To produce this figure, we follow
-the convention that the PSD as a function of frequency,
-$\mathbf{P}(f)$, is normalized such that the integral over frequency
+chamber (note that variations are still seen in the mixing chamber;
+however, such variations do not necessarily reflect changes in the
+focal plane temperature). There is minimal correlated low-frequency
+noise in these 450\,\micron\ data (note that there is almost no
+difference once the common-mode has been subtracted). The
+850\,\micron\ channel, however, exhibits a more significant signal
+that is obviously correlated with the telescope motion. It is also
+correlated amongst many of the detectors, and common-mode removal
+corrects it to some extent. Part of the reason that this is seen here,
+and not in the S2SRO data set shown, is that the amplitude of the scan
+pattern is larger. Most of this ``scan-synchronous'' noise is
+attributed to magnetic field pickup, as described in
+Section~\ref{sec:magpickup}.
+
+Next, in Fig.~\ref{fig:pspec} we produce power spectral density (PSD)
+plots for four of the most sensitive bolometers from both focal
+planes, using the same two data sets. To produce this figure, we
+follow the convention that the PSD as a function of frequency,
+$\mathbf{P}(f)$, is normalised such that the integral over frequency
 gives the same variance as the time-series variance across the full
@@
 
-The dotted coloured lines in Fig~\ref{fig:pspec} show the PSDs for
+The dotted coloured lines in Fig.~\ref{fig:pspec} show the PSDs for
 raw, though flat-fielded and step-corrected (Section~\ref{sec:steps})
@@
 spectra of unresolved point-sources are also shown as dashed lines in
-each band (arbitrarily normalized), showing that the smallest features
+each band (arbitrarily normalised), showing that the smallest features
 resolvable by the telescope are only minimally affected by the excess
@@
 low-frequency noise is very correlated between the detectors (note the
-tight scatter in the dotted colour lines, and their resemblance to the
+tight scatter in the dotted coloured lines, and their resemblance to the
 common-mode), due to it being dominated by the fridge
@@
 considerably less correlated amongst detectors (larger spread in the
-dotted colour lines, particularly noticeable in these 450\,\micron\
+dotted coloured lines, particularly noticeable in these 450\,\micron\
 data), leading to a less drastic improvement upon common-mode
@@
 that the white noise performance (NEFD) is in fact similar before and
-the upgrades for these two subarrays (s4a at 450\,micron, and s8b at
-850\,\micron). The main improvements are reduction in correlated and
-line noise features mentioned above, and the larger number of working
-bolometers (approximately a factor of 4).
+after the upgrades for these two subarrays (s4a at 450\,micron, and
+s8b at 850\,\micron). The main improvements are reduction in
+correlated and line noise features mentioned above, and the larger
+number of working bolometers (approximately a factor of 4).
 
@@
 with random signs for bolometers along a column, giving it a distinct
-signature from other common signals that appear with the same sign.
+signature from other common signals that always appear with the same
+sign.
 
@@
   of Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec}. For each component
-  the top plot shows the time-series of the normalized eigenvectors,
+  the top plot shows the time-series of the normalised eigenvectors,
   the middle plot its PSD, and the bottom coloured panels the
@@
 map-making procedure for BOLOCAM and AzTEC is Principal Component
-Analysis \citep[PCA,][]{laurent2005,scott2008}. Here we use PCA to
-explore correlated signals in \scuba\ data.
+Analysis \citep[PCA,][]{laurent2005,scott2008,perera2008}. Here we use
+PCA to explore correlated signals in \scuba\ data.
 
@@
 $\mathbf{\lambda}_i^\mathrm{T}$ is a column vector containing the
-corresponding eigenvalues. The $\mathbf{\xi}_i(t)$ are normalized by
-$[\sum_t \mathbf{\xi}_i^2(t)]^{1/2}$ such that the root mean square
+corresponding eigenvalues. The $\mathbf{\xi}_i(t)$ are normalised by
+$[\sum_t \mathbf{\xi}_i^2(t)]^{1/2}$, such that the root mean square
 (\rms) amplitude of each component in a given bolometer signal is
@@
 bands). The top and middle panels for each component show the
-time-series and PSDs of the normalized eigenvectors. The bottom panels
+time-series and PSDs of the normalised eigenvectors. The bottom panels
 are maps indicating the eigenvalues (amplitudes) of the component
@@
 would expect the atmosphere to be more significant at 450 than at
-850\,\micron, although it is not obviously visible due to the strong
+850\,\micron, although it is not obviously visible, due to the strong
 presence of the scan-synchronous noise.
@@
 \mathrm{constant}$ is fit to the power spectrum of each bolometer, and
-then the bolometer FFT is multiplied by its inverse (though normalized
+then the bolometer FFT is multiplied by its inverse (though normalised
 to preserve the white noise level).  The signals that are removed from
@@
 
-The estimated variance, $\mathbf{v_p}(x_i,y_i)$, of the normalized
+The estimated variance, $\mathbf{v_p}(x_i,y_i)$, of the normalised
 weighted samples that land in the $i$th map pixel is simply the biased
@@
 $\mathbf{v_p}$, they must be scaled appropriately. We define a
-normalized difference, $\mathbf{d}_j$, in such a way that the variance
+normalised difference, $\mathbf{d}_j$, in such a way that the variance
 of this new variable gives the weighted sample variance of the
@@
 absolute change in the value of map pixels between subsequent
-iterations, normalized by the map pixel uncertainties (square root of
+iterations, normalised by the map pixel uncertainties (square root of
 Eq.~\ref{eq:varmap}), or
@@
 reciprocal of the jackknife angular power spectrum (orange line in
-Fig.~\ref{fig:lockman_pspec}), normalized by the white-noise level
+Fig.~\ref{fig:lockman_pspec}), normalised by the white-noise level
 estimated from the RMS power at angular frequencies $>
@@
 match-filtered) signal (blue) and jackknife (histograms), along with a
-Gaussian (mean 0, standard deviation 1, and area normalized to the
+Gaussian (mean 0, standard deviation 1, and area normalised to the
 number of map pixels) as a dashed line. In this case, it is clear that
@@
   distribution with mean zero, standard deviation one, and an area
-  normalized to the total number of pixels -- the expected
+  normalised to the total number of pixels -- the expected
   distribution for a map of spatially-uncorrelated Gaussian noise. The
@@
 14.5\,arcsec Gaussian to model the effect of the \scuba\ optical
-response. The RMS of these fluctuations is then normalized to 0.002 pW
+response. The RMS of these fluctuations is then normalised to 0.002 pW
 so that they are comparable to the dynamic range of M17 itself. The
@@
 Foundation for Innovation. EC thanks CANARIE/CANFAR for additional
-funding. Also mention NSERC.
+funding. This research was supported in part by the Natural Sciences
+and Engineering Research Council of Canada. ... general thanks to
+people at JAC ... also explicitly mention Dennis, Alex and Jen ...
 

### 2012-09-07 smurf_paper.tex
small typo
Remove the notes on additional background on performance etc
The paper is long and doesn't really need proof that it scales
with cores or details on how much memory is required. The Starlink
URLs should be enough to get the data and NDF is just a distraction.

@@
 blocks of bolometers, or blocks of time, in each thread, depending on
-the nature of the calculation. \textbf{... something about scalability
-...}
-
-\begin{itemize}
-
-\item Discuss SMURF configurability
-
-\item Some sort of proof/demonstration showing why the algorithm works
-
-\item Discuss performance (execution time, memory, disk usage etc.)
-
-\item Where do you get SMURF/starlink, usage of NDF
-
-\item Talk about pipeline?
-
-\end{itemize}
+the nature of the calculation.
 
@@
 It was immediately found in the analysis of \scuba\ data that the
-$\chi^2_\mathrm{r}$ test described above i not a sufficient stopping
+$\chi^2_\mathrm{r}$ test described above is not a sufficient stopping
 criterion. The reason for this is that various components of the model

### 2012-09-07 smurf_paper.tex
A further stab at the acknowledgements

@@
 funding. This research was supported in part by the Natural Sciences
-and Engineering Research Council of Canada. ... general thanks to
-people at JAC ... also explicitly mention Dennis, Alex and Jen ...
+and Engineering Research Council of Canada. The authors thank the
+members of the SCUBA-2 commissioning team for testing the
+map-maker and reporting anomalies; in particular Antonio Chrysostomou
+and Jessica Dempsey. We also thank Dennis Kelly, Alex van Engelen and
+Jennifer Balfour for work on the SCUBA-2 observation simulator;
+and Mark Thompson, Craig Walther and S\'{e}verin Gaudet for being
+on the Critical Design Review panel.
 

### 2012-09-09 smurf_paper.tex
update dimm flowchart... using open office now instead of xfig

@@
 \centering
-\includegraphics[width=0.75\linewidth]{dimm.pdf}
+\includegraphics[width=\linewidth]{dimm.pdf}
 \caption{Typical map-making algorithm. Raw data (stored in multiple
-  files) are read and concatenated into continuous time-series. A
-  pre-processing stage repairs DC steps, applies a flatfield
-  correction, and identifies and removes the noisiest bolometers. The
-  iterative portion then begins: estimating and removing the
-  common-mode signal; applying the extinction correction; high-pass
-  filtering to remove residual independent low-frequency noise;
-  estimating the map by re-gridding the data; and finally measuring
-  the noise in the residual time-series. If the solution has
+  files) are read and concatenated into continuous time-series, and
+  flatfieded. Based on the scan speed of the telescope, the data are
+  down-sampled to match the selected output map pixel size (usually 2
+  and 4\,arcsec at 450 and 850\,\micron, respectively, to ensure
+  adequate sampling of the \scuba\ PSFs). A cleaning stage repairs DC
+  steps and spikes, and fills gaps to ensure continuity. Finally, the
+  mean of each bolometer time-series is removed (higher-order
+  polynomials may also be used). The iterative portion then begins
+  (dashed box): estimating and removing the common-mode signal (the
+  combined \model{COM} and \model{GAI} models); applying the
+  extinction correction (\model{EXT}); high-pass filtering to remove
+  residual independent low-frequency noise (\model{FLT}); estimating
+  the map by re-gridding the data, and then removing its projection
+  from the time-series (\model{AST}); and finally measuring the noise
+  in the residual time-series (\model{NOI}). If the solution has
   converged, the map is written to disk. Otherwise any multiplicative
-  factors that may have been applied to the data are inverted (e.g.,
-  extinction correction), and then each additive component is
-  re-estimated in turn, with the previous estimate first being added
-  back into the time-series (and importantly, not the other
-  components).}
+  factors that may have been applied to the data are inverted (i.e.,
+  the extinction correction, \model{EXT}), and then the models for the
+  low-frequency noise; the common-mode (\model{COM,GAI}) and high-pass
+  filter (\model{FLT}). Each model is then re-estimated in turn until
+  \model{AST}, at which point the previous estimate of the
+  astronomical signal is added back into the data prior to its
+  re-calculation.}
 \label{fig:dimm}

### 2012-09-10 smurf_paper.tex
numerous minor edits to entire paper except for the conclusions

@@
   operated by scanning across the sky and records data at a rate of
-  200\,Hz. The data rate is much greater than that of previously
-  existing submillimetre cameras, and represents a significant
-  analysis challenge. We describe the production of \scuba\ maps using
-  the Sub-Millimetre User Reduction Facility (SMURF) in which we have
+  200\,Hz. This rate is much greater than that of previously existing
+  submillimetre cameras, and represents a significant analysis
+  challenge. We describe the production of \scuba\ maps using the
+  Sub-Millimetre User Reduction Facility (SMURF) in which we have
   adopted a fast, iterative approach to map-making that enables data
@@
   feedback to observers, as well as for the production of generic
-  science products at the Canadian Astronomy Data Centre. A suite of
-  tools is also provided for interactive data analysis.  Three
+  science products at the Canadian Astronomy Data Centre. Three
   detailed case studies are used to: (i) explore convergence
   properties of the map-maker using simple prior constraints (Uranus
-  -- a point source); (ii) achieve the white-noise limit of the
-  instrument for faint point-source studies (blank-field survey of the
+  -- a point source); (ii) achieve the white-noise limit for faint
+  point-source studies (extragalactic blank-field survey of the
   Lockman Hole); and (iii) demonstrate that our strategy is capable of
-  recovering angular scales up to the size of the array footprint
-  (approximately 5\,arcmin) for bright extended sources (star-forming
-  region M17).
+  recovering angular scales comparable to the size of the array
+  footprint (approximately 5\,arcmin) for bright extended sources
+  (star-forming region M17).
 \end{abstract}
@@
 (defined here to be 200--1200\,\micron) have helped revolutionise
-several important areas of astrophysics, including: the
-characterisation of the early stages of star-formation by identifying
-the dense, cold regions in molecular clouds where stars may eventually
-form; discovering through blind surveys a class of massive dusty
-star-forming galaxies in the early ($z>2$) Universe, now referred to
-as submillimetre galaxies, or SMGs; and identifying debris disks
+several important areas of astrophysics, including: discovering
+through blind surveys a class of massive dusty star-forming galaxies
+in the early ($z>2$) Universe, now referred to as submillimetre
+galaxies, or SMGs; the characterisation of the early stages of
+star-formation by identifying the dense, cold regions in molecular
+clouds where stars may eventually form; and identifying debris disks
 around nearby stars, helping us understand the early stages of planet
-formation.  With $\sim10,000$ total detectors, \scuba\ is presently the
-largest submillimetre camera in the world, and the fastest wide-area
-ground-based submillimetre imager.
+formation.  With $\sim10,000$ total detectors, \scuba\ is presently
+the largest submillimetre camera in the world, and the fastest
+wide-area ground-based submillimetre imager.
 
@@
 the noise in the resulting map is said to be ``white noise limited'',
-meaning that the map noise is uncorrelated spatially, and has an
-amplitude that scales as the $\mathrm{NEFD}/\sqrt{t}$, where the NEFD
-is the noise-equivalent flux-density (the white noise level of a
-bolometer in 1\,s of integration), and $t$ is the amount of
-integration time in a map pixel.
+meaning that it is uncorrelated spatially, and has an amplitude that
+scales as the $\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the
+noise-equivalent flux-density (the white noise level of a bolometer in
+1\,s of integration), and $t$ is the amount of integration time in a
+map pixel.
 
@@
 interest. In more recent years a method that has been popular for
-reducing data in fields of faint point sources with BOLOCAM
-\citep[e.g.,][]{laurent2005}, and its younger sibling AzTEC
-\citep[e.g.,][]{scott2008}, is ``principal component analysis'' (PCA)
-cleaning. A statistical ``black-box'' removes the most correlated
-components of the bolometer signals, enabling the detection of
-point-sources very close to the theoretical white-noise limits of the
-detectors, with reasonable computation times when small numbers of
-bolometers are involved (hundreds rather than thousands of
-detectors). However, PCA cleaning is not a good generic solution for
-producing maps of extended structures, since such sources produce
-correlated signals amongst many detectors, and are removed by this
-procedure.
+reducing data in fields of faint point sources with Bolocam
+\citep[e.g.,][]{laurent2005}, and its younger sibling the Aztronomical
+Thermal Emission Camera \citep[AzTEC, e.g.,][]{scott2008}, is
+``principal component analysis'' (PCA) cleaning. A statistical
+``black-box'' removes the most correlated components of the bolometer
+signals, enabling the detection of point-sources very close to the
+theoretical white-noise limits of the detectors, with reasonable
+computation times when small numbers of bolometers are involved
+(hundreds rather than thousands of detectors). However, PCA cleaning
+is not a good generic solution for producing maps of extended
+structures, since such sources produce correlated signals amongst many
+detectors, and are removed by this procedure. Furthermore, in the case
+of \scuba\, performing PCA on even a single subarray (1280 detectors,
+nominally) can be prohibitively slow.
 
@@
 fact that it used a differential radiometer instead of
-bolometers). Other descriptions in the CMB literature include
-\citet{tegmark1997} and \citet{stompor2002}, and an application to
-data from SCUBA is described in \citet{borys2004}.  The downside to
-such methods is that they can be both computationally expensive and
-have large memory requirements. While for some experimental designs
-fast iterative methods for the inversion do exist with minimal memory
-usage \citep[e.g.,][]{wright1996}, for the more general map-making
-problem involving many detectors, things are significantly complicated
-both by the need to measure the cross power-spectra for all unique
-pairs of bolometers, as well as performing the inversion itself. The
-most promising maximum-likelihood method that may one day be applied
-to \scuba\ data is ``SANEPIC'', which successfully reduced maps from
-the Balloon-borne Large-Aperture Submillimeter Telescope, involving
-hundreds of bolometers, while correctly incorporating inter-bolometer
-noise correlations \citep[][]{patanchon2008}.
-
-%  Styles of map-making: least
-%squares \citep{janssen1992} maximum likelihood
-%\citep[e.g.,][]{patanchon2008}; iterative -- see \citet{johnstone2000}
-%implementation of \citet{wright1996} to remove chop,
-%\citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA analysis
-%\citep[e.g.,][]{laurent2005,scott2008,aguirre2010}.
+bolometers). Other descriptions in the experimental cosmic microwave
+background (CMB) literature include \citet{tegmark1997} and
+\citet{stompor2002}, and an application to data from SCUBA is
+described in \citet{borys2004}.  The downside to such methods is that
+they can be both computationally expensive and have large memory
+requirements. While for some experimental designs fast iterative
+methods for the inversion do exist without requiring excessive amounts
+of memory \citep[e.g.,][]{wright1996}, for the more general map-making
+problem, involving many detectors, things are significantly
+complicated both by the need to measure the cross power spectra for
+all unique pairs of bolometers, as well as performing the inversion
+itself. The most promising maximum-likelihood method that may one day
+be applied to \scuba\ data is ``SANEPIC'', which successfully reduced
+maps from the Balloon-borne Large-Aperture Submillimeter Telescope,
+involving hundreds of bolometers, while correctly incorporating
+inter-bolometer noise correlations \citep[][]{patanchon2008}.
 
@@
 remaining (non-modelled) noise sources are assumed to be white, a
-single scalar noise may be calculated for all of the data points from
-a given bolometer (since the noise at any instant in time is
-uncorrelated with others, and with the data from other bolometers),
-greatly simplifying the measurement of noise properties and the
-inversion step that is so complicated in the maximum-likelihood
-methods. In principle, it should also be possible to recover a wider
-dynamic range in angular scales than possible for the simplest direct
-methods, with an increase in calculation time that is merely linear in
-the number of iterations needed, while using a comparable amount of
-memory. A fast iterative method for handling chopped data from CMB
-experiments was described in \citet{wright1996}, and later applied to
-chopped SCUBA data by \citet{johnstone2000}, although in each case the
-chopping renders the data essentially white without the need to model
-additional noise components. More similar to the algorithm described
-here was the iterative pipeline recipe for fitting and removing
-baseline drifts in SCUBA scan-map data as the astronomical image
-estimate improved \citep{1999ASPC..172..171J}. The two closest modern
-relatives of SMURF are the Comprehensive Reduction Utility for SHARC-2
+single scalar \rms\ may be calculated for all of the data points from
+a given bolometer to characterise its noise distribution (since the
+noise at any instant in time is uncorrelated with others, and with the
+data from other bolometers), greatly simplifying the measurement of
+noise properties and the inversion step that is so complicated in the
+maximum-likelihood methods. In principle, it should also be possible
+to recover a wider dynamic range in angular scales than possible for
+the simplest direct methods, with an increase in calculation time that
+is merely linear in the number of iterations needed, while using a
+comparable amount of memory. The iterative method of
+\citet{wright1996} specific to chopped (CMB) data was applied to SCUBA
+observations by \citet{johnstone2000}, although in this case the
+chopping removes most of the low-frequency noise, and greatly
+simplifies the problem (at the expensive of large-scale structure in
+the data). More similar to the algorithm described here was the
+iterative pipeline recipe for fitting and removing baseline drifts in
+SCUBA scan-map data as the astronomical image estimate improved
+\citep{1999ASPC..172..171J}. The two closest modern relatives of SMURF
+are the Comprehensive Reduction Utility for SHARC-2
 \citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
-Bolocam Galactic Plane Survey \citep{aguirre2010}.
+Bolocam Galactic Plane Survey \citep[][which uses iterative PCA
+cleaning to avoid the suppression of large-scale
+structure]{aguirre2010}.
 
@@
 This paper is organised as follows. We first describe the properties
-of \scuba\ data in Section~\ref{sec:data} to motivate the data model
-that we have adopted. Next, the details of the SMURF algorithm are
-given in Section~\ref{sec:algorithm}. The paper is concluded in
-Section~\ref{sec:examples} with three detailed test cases that span
-the majority of observation types likely to be undertaken with \scuba,
-with an emphasis on the mitigation of divergence problems, and
-characterising the output maps: (i) Uranus, a bright, compact source
-(Section~\ref{sec:point}); (ii) the Lockman Hole, a blind survey of
-faint point-like sources (Section~\ref{sec:cosmo}); and (iii) the
-star-forming region M17, including bright, extended emission
-(Section~\ref{sec:extended}). All of the data analysed in this paper
-are publicly available through the CADC \scuba\ raw-data queries page
-for the dates and observation numbers given in the
+of \scuba\ data, including a principal component analysis to reveal
+correlated noise features, in Section~\ref{sec:data}. Next, the
+details of the SMURF algorithm (pre-processing steps and the iterative
+solution) are given in Section~\ref{sec:algorithm}. The paper is
+concluded in Section~\ref{sec:examples} with three detailed test cases
+that span the majority of observation types likely to be undertaken
+with \scuba, with an emphasis on the mitigation of divergence
+problems, and characterising the output maps: (i) Uranus, a bright,
+compact source (Section~\ref{sec:point}); (ii) the Lockman Hole, a
+blind survey of faint point-like sources (Section~\ref{sec:cosmo});
+and (iii) the star-forming region M17, including bright, extended
+emission (Section~\ref{sec:extended}). All of the data analysed in
+this paper are publicly available through the CADC \scuba\ raw-data
+queries page for the dates and observation numbers given in the
 text\footnote{\url{http://www.cadc-ccda.hia-iha.nrc-cnrc.gc.ca/jcmt/search/scuba2}}.
@@
 release from 2012.
-
-%\scuba\ is presently unique in terms of the shear volume of data that
-%it produces on a nightly basis, and the fact that it is observing a
-%wide variety of science targets for multiple users (unlike dedicated
-%CMB experiments, for example). For this reason, SMURF has been
-%designed to use a faster, and less memory-intensive algorithm than
-%CMB-style maximum likelihood map-makers, although it is less accurate
-%at reproducing large-scale structures. The basic procedure is to
-%iteratively model and remove most of the bolometer-bolometer
-%correlated signal components and independent low-frequency noise, and
-%then regrid the residual signals.
-
-% Historically, both radio
-%and submillimetre telescope systems have used ``chopping'' to perform
-%rapid differences between target sky locations and some reference
-%(e.g., another point on the sky, internal calibrators, and different
-%frequencies in the case of heterodyne systems). In fact, spatial
-%chopping was used by SCUBA, the predecessor to \scuba, to remove
-%low-frequency drifts \citep{holland1999}. However, chopping (in
-%particular spatial chopping) has the unfortunate side-effect of
-%filtering out information on scales larger than the chop-throw
-%\textbf{(references)}.For this reason, modern cameras like \scuba\ aim
-%to provide a high-degree of inherent detector stability, and rapidly
-%slew the telescope so that astronomical signals up to large scales are
-%captured in a signal band at considerably higher frequency than the
-%residual low-frequency drift.
-
-%In the case of chopped data, ideally the resulting data are completely
-%free of low-frequency noise, and therefore the data samples are
-%statistically uncorrelated.
-
-%often used various
-%``chopping'' methods to remove low-frequency drifts -- rapid
-%differences are made between the sky signal at a given position and an
-%internal reference signal; the sky signal in two different bands for
-%heterodyne systems (frequency chopping); and a different position in
-%the sky \citep[a method employed by SCUBA, the predecessor of
-%\scuba][]{holland1999}.
 
@@
 of bolometers. The bolometers are thermal absorbers coupled to
-superconducting transition-edge sensors (TESs) for thermometry;
-temperature variations in the TESs produce changing currents, and
-therefore varying magnetic fields.  The changing magnetic fields are
-detected and amplified using chains of superconducting quantum
-interference devices (SQUIDs), and the larger output currents are
-digitised. Each detector has its own SQUID for the first-stage of the
-amplification, but the remaining stages occur within a common chain of
-SQUIDs for each of the 32 columns. All 40 rows are read out in
-sequence at a row-visit frequency of about 12\,kHz. Such a high sample
-rate is unnecessary to produce maps, so the data are re-sampled to
-approximately 200\,Hz before writing to disk.  It should also be noted
-that there is an additional 41st row of SQUID readouts that are not
-connected to TESs. These ``dark SQUIDs'' track non-thermal noise
-sources that are common to each column's amplifier chain. The
-relationship between the output digitized current, $I$, and the input
-power, $P$, or $dI/dP$ is established using flatfield observations
-immediately prior to science observations, in which the output signal
-is measured throughout a ramp of the pixel heaters (which provide a
-known input power); see Section~2.1 in \citet{dempsey2012} and
-Section~5.3 in \citet{holland2012} for more details. Finally, the
-conversion to astronomical flux units from power involves a correction
-for atmospheric extinction \citep[primarily using the JCMT Water
-Vapour Monitor to track line-of-sight opacity variations, see
-Section~3 in][]{dempsey2012}, and the application of a flux conversion
-factor (FCF) which is established from regular measurements of
-calibrators such as Uranus \citep[Section~5 in][]{dempsey2012}.
-
-The predominant noise features in \scuba\ bolometers are: (i) the
-effectively white thermal noise produced by the phonon and optical
-loading on the absorbers (largely independent from bolometer to
-bolometer); and (ii) longer timescale drifts resulting from slow
-variations in the mean load (e.g., fridge, telescope, or atmospheric
-changes), and also the readout electronics (often correlated amongst
-many or all of the bolometers). As such, the power spectra of
-bolometer time-series exhibit a ``$1/f^\alpha$'' ($\alpha>0$) shape
-below some knee frequency, $f_\mathrm{knee}$, and they are then
-relatively flat at higher frequencies. The spectra also include line
-features which are mainly produced by aliased high-frequency noise
-sources detected during the multiplexed readout stage, as described
-above.  Finally, an anti-aliasing filter produces roll-off in the
-spectrum at frequencies approaching Nyquist prior to the 200\,Hz
-re-sampling.
+superconducting transition-edge sensors (TESs) for thermometry.
+Temperature variations in the TESs produce changing currents, and
+therefore varying magnetic fields, which are detected and amplified
+using chains of superconducting quantum interference devices (SQUIDs),
+before the larger output currents are digitised. Each detector has its
+own SQUID for the first-stage of the amplification, but the remaining
+stages occur within a common chain of SQUIDs for each of the 32
+columns. All 40 rows are read out in sequence at a row-visit frequency
+of about 12\,kHz. Such a high sample rate is unnecessary to produce
+maps, so the data are re-sampled to approximately 200\,Hz before
+writing to disk. This rate provides a sample every 3.0\,arcsec, or
+approximately one third of the 450\,\micron\ diffraction-limited
+full-width at half-maximum (FWHM) -- a typical rule-of-thumb for
+adequately sampling a Gaussian point spread function (PSF). There is
+an additional 41st row of SQUID readouts that are not connected to
+TESs. These ``dark SQUIDs'' track non-thermal noise sources that are
+common to each column's amplifier chain. The relationship between the
+output digitized current, $I$, and the input power, $P$, or $dI/dP$ is
+established using flatfield observations immediately prior to science
+observations, in which the output signal is measured throughout a ramp
+of the pixel heaters (which provide a known input power); see
+Section~2.1 in \citet{dempsey2012} and Section~5.3 in
+\citet{holland2012} for more details. Finally, the conversion to
+astronomical flux units from power involves a correction for
+atmospheric extinction \citep[primarily using the JCMT Water Vapour
+Monitor to track line-of-sight opacity variations, see Section~3
+in][]{dempsey2012}, and the application of a flux conversion factor
+(FCF) which is established from regular measurements of calibrators
+such as Uranus \citep[Section~5 in][]{dempsey2012}.
+
+In addition to uncorrelated white noise, and low-frequency (and often
+correlated from bolometer to bolometer) $1/f$ drifts, bolometer power
+spectra include a roll-off in their spectra at frequencies approaching
+Nyquist, which is due to the anti-aliasing filter that forms part of
+the 200\,Hz re-sampling stage. There are also line features which are
+thought to be produced at frequencies far above the final 200\,Hz
+sample rate, and are aliased to lower frequencies during the
+multiplexed readout stage before they can be removed by the
+anti-aliasing filter.
 
@@
 
-In order to reduce the impact of noise on the final maps, it is
-essential that an appropriate scan strategy be used. First, it is
-clearly advantageous that the astronomical signals appear in the
-lowest-noise region of the power spectrum, namely frequencies above
-$f_\mathrm{knee}$. This is achieved by scanning the telescope at large
-slew rates, which are practically limited to about
-600\,arcsec\,s$^{-1}$ for the JCMT in order to have reasonable
-turn-around overheads (the amount of time spent changing direction and
-speed, over the largest elevation range possible). At this maximum
-speed there is a sample every 3.0\,arcsec, or approximately one third
-of the 450\,\micron\ diffraction-limited full-width at half-maximum
-(FWHM) -- a typical rule-of-thumb for adequately sampling a Gaussian
-point spread function (PSF). The scan patterns are designed,
-primarily, to mitigate the effects of low-frequency $1/f$ drift. The
-simplest scan one might use is a raster ``boustrophedon'' pattern,
-i.e., a series of parallel passes across the field. However, such a
-scan yields extremely poor performance for structures that are
-transverse to the scan direction, since a long time will have elapsed
-between each parallel step. In addition, other periodic noise features
-may conspire with such a regularly-spaced pattern to produce periodic
-structures in the map, or ``scan-synchronous'' noise. Scan strategies
-that visit each region of the map on as many different time scales and
-scan angles as possible, offer the best protection against large-scale
-noise; such strategies are said to offer good cross-linking. As a
-trade-off between truly random scan patterns and feasibility of
-implementation, taking into consideration both mechanical and
-scheduling constraints, \scuba\ presently has two primary scan modes:
-PONG, which covers rectangular areas larger than the array footprint
-in a closed pattern that ``bounces'' off the edges at 45\,degree
-angles (and also usually combined with a rotation through a number of
-fixed position angles to create a ``rotating PONG'' with even better
-cross-linking); and a ``constant velocity daisy'' (CV Daisy) for small
-areas (of order the array footprint), in which the telescope moves in
-a circle, whose centre also slowly traces out a circle. For a more
-complete description of the \scuba\ observing modes, see Section~5 in
-\citet{holland2012} and also \citet{2010SPIE.7740E..66K}.
+In order to reduce the impact of low-frequency noise on the final
+maps, \scuba\ scan strategies have been designed to provide: (i) good
+cross-linking (visiting every point of the mapped area on several
+different time scales, and at different scan angles); and (ii) minimal
+accelerations to reduce turn-around overheads (which could be quite
+large given the 600\,arcsec\,s$^{-1}$ maximum scan speed). For areas
+larger than the array footprint a closed ``PONG'' pattern is used, in
+which the boresight travels in approximately straight lines and
+``bounces'' off the edges at 45\,degree angles until the area is
+uniformly filled in. It is also usually combined with a rotation
+through a number of fixed position angles to create a ``rotating
+PONG'' with even better cross-linking. For smaller areas (of order the
+array footprint, or pont sources), in which the PONG turn-around
+overheads would be large, a ``constant velocity daisy'' (CV Daisy) is
+used.  Here the telescope moves in a circle, whose centre also slowly
+traces out a small circle. For a more complete description of the
+\scuba\ observing modes, see Section~5 in \citet{holland2012} and also
+\citet{2010SPIE.7740E..66K}.
 
@@
 bolometers in each of the 450 and 850\,\micron\ focal planes, as well
-as variations in the mixing chamber temperature (as a proxy for the
-focal-plane temperature), and the telescope pointing, for two data
-sets, before and after the upgrades that followed S2SRO.
+as variations in the mixing chamber temperature (though not located in
+the focal plane itself, it is certainly correlated with the
+temperature of the detectors), and the telescope pointing, for two
+data sets, before and after the upgrades that followed S2SRO.
 
@@
 problem, these data no longer exhibit a correlation with the mixing
-chamber (note that variations are still seen in the mixing chamber;
-however, such variations do not necessarily reflect changes in the
-focal plane temperature). There is minimal correlated low-frequency
-noise in these 450\,\micron\ data (note that there is almost no
-difference once the common-mode has been subtracted). The
+chamber (however, note that variations are still seen in the mixing
+chamber signal; such variations do not necessarily reflect changes in
+the focal plane temperature). There is minimal correlated
+low-frequency noise in these 450\,\micron\ data (note that there is
+almost no difference once the common-mode has been subtracted). The
 850\,\micron\ channel, however, exhibits a more significant signal
@@
 frequencies, and finally a roll-off caused by the anti-aliasing filter
-above $\gsim 70$\,Hz. As indicated in the previous section, the
-correlation between the low-frequency components of the different
-bolometer signals is large. The solid black lines in
-Fig.~\ref{fig:pspec} indicate the PSDs of the common-modes
-$\mathbf{c}(t)$ at each wavelength, which reproduce much of the
-low-frequency structure, as well as some of the higher-frequency line
-features. The $\mathbf{c}(t)$ otherwise drop substantially below the
-individual bolometer PSDs at high-frequency, as expected if the
-bolometers are dominated by uncorrelated white noise in that part of
-the spectrum. The common-mode signals are fit to each bolometer time
-series and removed as in Fig.~\ref{fig:bolos_mix}, and the resulting
-PSDs are shown as solid coloured lines.  For reference, the top
-horizontal axes have been converted to angular scale assuming the scan
-speeds of each observation (120\,arcsec\,s$^{-1}$ in the left-hand
-plot, and 190\,arcsec\,s$^{-1}$ in the right-hand plot). The power
-spectra of unresolved point-sources are also shown as dashed lines in
-each band (arbitrarily normalised), showing that the smallest features
+above $\gsim 70$\,Hz. The correlation between the low-frequency
+components of the different bolometer signals is large. The solid
+black lines in Fig.~\ref{fig:pspec} indicate the PSDs of the
+common-modes $\mathbf{c}(t)$ at each wavelength, which reproduce much
+of the low-frequency structure, as well as some of the
+higher-frequency line features. The $\mathbf{c}(t)$ otherwise drop
+substantially below the individual bolometer PSDs at high-frequency,
+as expected if the bolometers are dominated by uncorrelated white
+noise in that part of the spectrum. The common-mode signals are fit to
+each bolometer time series and removed as in Fig.~\ref{fig:bolos_mix},
+and the resulting PSDs are shown as solid coloured lines.  For
+reference, the top horizontal axes have been converted to angular
+scale assuming the scan speeds of each observation
+(120\,arcsec\,s$^{-1}$ in the left-hand plot, and
+190\,arcsec\,s$^{-1}$ in the right-hand plot). The power spectra of
+unresolved point-sources are also shown as dashed lines in each band
+(arbitrarily normalised), showing that the smallest features
 resolvable by the telescope are only minimally affected by the excess
@@
 Fig.~\ref{fig:pspec}) generally have less significant $1/f$ noise as
-compared with the S2SRO data. The residual low-frequency noise is
-considerably less correlated amongst detectors (larger spread in the
-dotted coloured lines, particularly noticeable in these 450\,\micron\
-data), leading to a less drastic improvement upon common-mode
-removal. However, since the data are less dominated by low-frequency
-noise to begin with, the signals generally require less low-frequency
-filtering to produce maps, and therefore retain larger-scale
-structures than with the S2SRO data. Furthermore, there are generally
-fewer line features in the PSDs of bolometers in the
-fully-commissioned instrument. Finally, these two data sets illustrate
-that the white noise performance (NEFD) is in fact similar before and
-after the upgrades for these two subarrays (s4a at 450\,micron, and
-s8b at 850\,\micron). The main improvements are reduction in
-correlated and line noise features mentioned above, and the larger
-number of working bolometers (approximately a factor of 4).
+compared with the S2SRO data, and it is considerably less correlated
+amongst detectors (larger spread in the dotted coloured lines,
+particularly noticeable in these 450\,\micron\ data), leading to a
+less drastic improvement upon common-mode removal. However, since the
+data are less dominated by low-frequency noise to begin with, the
+signals generally require less agressive high-pass filtering to
+produce maps, and therefore retain larger-scale structures than with
+the S2SRO data. Furthermore, there are generally fewer line features
+in the PSDs of bolometers in the fully-commissioned
+instrument. Finally, these two data sets illustrate that the white
+noise performance (NEFD) is in fact similar before and after the
+upgrades for these two subarrays (s4a at 450\,micron, and s8b at
+850\,\micron). The main improvements are a reduction in correlated and
+line noise features mentioned above, the larger number of working
+bolometers and field-of-view (approximately a factor of 4).
 
@@
 %-------------------------------------------------
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{magpickup.pdf}
+\caption{Evidence for significant magnetic field pickup for
+  observation 16 on 20100228 at 450\,\micron\ (s4a subarray).  The top
+  panel shows two un-flatfielded (but mean-subtracted and step
+  corrected) bolometer time-series from the same column, with a 200
+  sample boxcar smooth (approximately 1\,s), illustrating that they
+  are dominated by a similar signal with opposite signs. The second
+  panel shows the dark squid signal for the column, also
+  mean-subtracted and with the same boxcar smooth. The bottom panel
+  shows the azimuthal and elevation offsets from the map centre (mean
+  azimuth and elevation 171.9$^\circ$ and 68.0$^\circ$,
+  respectively). Only the azimuthal signal is obviously correlated
+  with the dark squids and bolometer signals, which suggests a
+  magnetic field stationary with respect to the telescope dome as the
+  source, since its direction with respect to the cryostat only
+  changes with azimuthal motion.}
+\label{fig:magpickup}
+\end{figure}
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{magpickup.pdf}
-\caption{Evidence for significant magnetic field pickup for
-  observation 16 on 20100228 at 450\,\micron\ (s4a subarray).  The top
-  panel shows two un-flatfielded (but mean-subtracted and step
-  corrected) bolometer time-series from the same column, with a 200
-  sample boxcar smooth (approximately 1\,s), illustrating that they
-  are dominated by a similar signal with opposite signs. The second
-  panel shows the dark squid signal for the column, also
-  mean-subtracted and with the same boxcar smooth. The bottom panel
-  shows the azimuthal and elevation offsets from the map centre. For
-  reference, the mean azimuth was 171.9$^\circ$ and the mean elevation
-  68.0$^\circ$. Only the azimuthal signal appears to be correlated
-  with the dark squids and bolometer signals, which suggests a
-  magnetic field stationary with respect to the telescope dome as the
-  source, since its direction with respect to the cryostat only
-  changes with azimuthal motion.}
-\label{fig:magpickup}
-\end{figure}
-
 The telescope pointing offsets for this approximately 0.5\,degree
@@
 earth's magnetic field). Since \scuba\ is mounted on a Nasmyth
-platform, only azimuthal motion will change the effective direction of
-such a field with respect to the cryostat. Tests have shown that, as
-in this example, large scans in azimuth generically produce pickup. In
+platform, only azimuthal motion will change the direction of such a
+field with respect to the cryostat. Tests have shown that, as in this
+example, large scans in azimuth generically produce pickup. In
 contrast, and as expected, changes in elevation results in pickup that
@@
 %-------------------------------------------------
+
+A method that has been used to remove correlated noise as part of the
+map-making procedure for Bolocam and AzTEC is Principal Component
+Analysis \citep[PCA,][]{laurent2005,scott2008,perera2008}. Here we use
+PCA to explore correlated signals in \scuba\ data.
+
+The basic method is as follows: (i) a covariance matrix is built up
+for all pairs $(i,j)$ of the $N$ bolometer time-series,
+$\langle\mathbf{b}_i(t),\mathbf{b}_j(t)\rangle$; and (ii) a singular
+value decomposition identifies a new set of statistically uncorrelated
+eigenvectors, $\mathbf{\xi}_i(t)$ (i.e., whose covariance matrix is
+diagonal), such that each of the bolometer time-series is a linear
+combination of the eigenvectors, or components i.e., $\mathbf{b}_i(t)
+= \bar{\mathbf{\xi}} \mathbf{\lambda}_i^\mathrm{T}$, where each row of
+the matrix $\bar{\mathbf{\xi}}$ is an eigenvector, and
+$\mathbf{\lambda}_i^\mathrm{T}$ is a column vector containing the
+corresponding eigenvalues. The $\mathbf{\xi}_i(t)$ are normalised by
+$[\sum_t \mathbf{\xi}_i^2(t)]^{1/2}$, such that the root mean square
+(\rms) amplitude of each component in a given bolometer signal is
+stored in the eigenvalues. In the earlier analyses mentioned, the
+low-frequency noise is assumed to be encapsulated in those components
+with the largest eigenvalues. Removing the projection of the time
+series along those components then significantly reduces $1/f$ noise
+while retaining most of the (higher-frequency) signal in point-like
+sources
+
+A novel feature of our \scuba\ analysis is that we can perform PCA
+with 450 and 850\,\micron\ data simultaneously, potentially helping us
+to differentiate thermal and optical noise signals (e.g., from the
+atmosphere) that might appear in both wavelengths, from other noise
+sources that are restricted to single subarrays (such as readout
+noise). In Fig.~\ref{fig:pca} we show the results for a combined
+analysis of the s4a and s8b subarrays for the first $\sim$100\,s of
+the same observation that was used in the right-hand
+(fully-commissioned) panels of Figs.~\ref{fig:bolos_mix} and
+\ref{fig:pspec}. The 12 most significant components are shown (ranked
+by the mean eigenvalues across all working bolometers in both
+bands). The top and middle panels for each component show the
+time-series and PSDs of the normalised eigenvectors. The bottom panels
+are maps indicating the eigenvalues (amplitudes) of the component
+across the focal plane (with the mean $\bar{\lambda}$ and and \rms,
+$\lambda_\mathrm{rms}$, of the eigenvalues for all of the bolometers
+calculated separately at each wavelength also shown).
+
+The majority of the correlated signal at both wavelengths in
+Fig.~\ref{fig:pca} is clearly produced by Components~1 and 2. In both
+cases, the eigenvector time-series exhibit a roughly periodic signal
+that resembles the scan pattern in the right-hand panel of
+Fig.~\ref{fig:pca}. While these correlated signals also clearly appear
+at both wavelengths, referring to the maps of eigenvalues, they are
+considerably stronger at 850\,\micron\ (consistent with the visual
+appearance of the bolometer signals in Fig.~\ref{fig:pca}). Also note
+that the PSDs for these eigenvectors exhibit nearly a pure $1/f$
+signature with almost no line features, or a white-noise plateau
+(suggesting that these are high-\snr\ measurements of a purely
+low-frequency drift).
 
@@
 
-A method that has been used to remove correlated noise as part of the
-map-making procedure for BOLOCAM and AzTEC is Principal Component
-Analysis \citep[PCA,][]{laurent2005,scott2008,perera2008}. Here we use
-PCA to explore correlated signals in \scuba\ data.
-
-The basic method is as follows: (i) a covariance matrix is built up
-for all pairs $(i,j)$ of the $N$ bolometer time-series,
-$\langle\mathbf{b}_i(t),\mathbf{b}_j(t)\rangle$; and (ii) a singular
-value decomposition identifies a new set of statistically uncorrelated
-eigenvectors, $\mathbf{\xi}_i(t)$ (i.e., whose covariance matrix is
-diagonal), such that each of the bolometer time-series is a linear
-combination of the eigenvectors, or components i.e., $\mathbf{b}_i(t)
-= \bar{\mathbf{\xi}} \mathbf{\lambda}_i^\mathrm{T}$, where each row of
-the matrix $\bar{\mathbf{\xi}}$ is an eigenvector, and
-$\mathbf{\lambda}_i^\mathrm{T}$ is a column vector containing the
-corresponding eigenvalues. The $\mathbf{\xi}_i(t)$ are normalised by
-$[\sum_t \mathbf{\xi}_i^2(t)]^{1/2}$, such that the root mean square
-(\rms) amplitude of each component in a given bolometer signal is
-stored in the eigenvalues. In the earlier analyses mentioned, the
-low-frequency noise is assumed to be encapsulated in those components
-with the largest eigenvalues. Removing the projection of the time
-series along those components then significantly reduces $1/f$ noise
-while retaining most of the (higher-frequency) signal in point-like
-sources
-
-A novel feature of our \scuba\ analysis is that we can perform PCA
-with 450 and 850\,\micron\ data simultaneously, potentially helping us
-to differentiate thermal and optical noise signals (e.g., from the
-atmosphere) that might appear in both wavelengths, from other noise
-sources that are restricted to single subarrays (such as readout
-noise). In Fig.~\ref{fig:pca} we show the results for a combined
-analysis of the s4a and s8b subarrays for the first $\sim$100\,s of
-the same observation that was used in the right-hand
-(fully-commissioned) panels of Figs.~\ref{fig:bolos_mix} and
-\ref{fig:pspec}. The 12 most significant components are shown (ranked
-by the mean eigenvalues across all working bolometers in both
-bands). The top and middle panels for each component show the
-time-series and PSDs of the normalised eigenvectors. The bottom panels
-are maps indicating the eigenvalues (amplitudes) of the component
-across the focal plane (with the mean $\bar{\lambda}$ and and \rms,
-$\lambda_\mathrm{rms}$, of the eigenvalues for all of the bolometers
-calculated separately at each wavelength also shown).
-
-The majority of the correlated signal at both wavelengths in
-Fig.~\ref{fig:pca} is clearly produced by Components~1 and 2. In both
-cases, the eigenvector time-series exhibit a roughly periodic signal
-that resembles the scan pattern in the right-hand panel of
-Fig.~\ref{fig:pca}. While these correlated signals also clearly appear
-at both wavelengths, referring to the maps of eigenvalues, they are
-considerably stronger at 850\,\micron\ (consistent with the visual
-appearance of the bolometer signals in Fig.~\ref{fig:pca}). Also note
-that the PSDs for these eigenvectors exhibit nearly a pure $1/f$
-signature with almost no line features, or a white-noise plateau
-(suggesting that these are high-\snr\ measurements of a purely
-low-frequency drift).
-
-What is the physical interpretation of these two components? First,
-the smoothly-varying pattern in the eigenvalue maps (particularly
-evident at 850\,\micron) suggests that the scan-synchronous noise
-(presumably magnetic field pickup) has a different response across the
-focal plane; in this case the PCA may have identified two orthogonal
-shapes that, when mixed in different quantities, can reproduce most of
-the signal in each bolometer. It is also likely that a varying
-atmospheric contribution is also included in these eigenvectors. One
-would expect the atmosphere to be more significant at 450 than at
-850\,\micron, although it is not obviously visible, due to the strong
-presence of the scan-synchronous noise.
+Why are there two dominant components? First, the smoothly-varying
+pattern in the eigenvalue maps (particularly evident at 850\,\micron)
+suggests that the scan-synchronous noise (presumably magnetic field
+pickup) has a different response across the focal plane; in this case
+the PCA may have identified two orthogonal shapes that, when mixed in
+different quantities, can reproduce most of the signal in each
+bolometer. It is also likely that a varying atmospheric contribution
+is also included in these eigenvectors. One would expect the
+atmosphere to be more significant at 450 than at 850\,\micron,
+although it is not obviously visible due to the strong presence of the
+scan-synchronous noise.
 
@@
 such as the two black bolometers near column 8, row 21 at
-850\,\micron\ in Component~3.
+850\,\micron\ in Component~3. Such events are typically flagged and
+removed during iterative de-spiking (Section~\ref{sec:ast}).
 
@@
 columns in the eigenvalue maps at both wavelengths (e.g., columns 26
-and 29 at 450\,\micron in Component~6). Finally, these components also
-have potentially unrelated features mixed into them, like sudden
-offsets that are likely residuals due to imperfect step correction
-(e.g., $\sim$50\,s in the eigenvector time-series for Component~6; also
-see Section~\ref{sec:steps}).
-
-Finally, Component~12 in Fig.~\ref{fig:pca} is again significant in
-only a handful of detectors in the 850\,\micron\ subarray (e.g., column
-21 row 25). The time-series plot for the eigenvector exhibits a large
+and 29 at 450\,\micron in Component~6). Finally, these components have
+potentially unrelated features mixed into them, like sudden offsets
+that are likely residuals due to imperfect step correction (e.g., at
+$t\sim50$\,s in the eigenvector time-series for Component~6; also see
+Section~\ref{sec:steps}).
+
+Finally, Component~12 in Fig.~\ref{fig:pca} is significant in only a
+handful of detectors in the 850\,\micron\ subarray (e.g., column 21
+row 25). The time-series plot for the eigenvector exhibits a large
 positive tail of sudden excursions which gradually worsen over
-time. Bolometers that exhibit this signal are probably poorly biased.
+time. Bolometers that exhibit this signal are probably poorly biased,
+and are usually flagged due to elevated noise levels during
+pre-processing (Section~\ref{sec:flagbadbol}).
 
@@
 time (e.g., scan-synchronous noise which obviously depends on the scan
-pattern). It should also be clear from this example that while PCA
-offers some helpful insight into the various sources of noise, it does
-not necessarily identify clean patterns that can easily be
-modelled. For example, while there are clearly correlation patterns
-along columns (Components~6 and 8--10, particularly at 450\,\micron),
-as one might expect given the common amplification chain, the
-intensities of these high-frequency signals are random in the
-eigenvalue maps, and a common-mode signal estimated from the data for
-each column would not remove it.
+pattern, and the tuning of the subarrays). It should also be clear
+from this example that while PCA offers some helpful insight into the
+various sources of noise, it does not necessarily identify clean
+patterns that can easily be modelled. For example, while there are
+clearly correlation patterns along columns (Components~6 and 8--10,
+particularly at 450\,\micron), as one might expect given the common
+amplification chain, the intensities of these high-frequency signals
+are random in the eigenvalue maps, and a common-mode signal estimated
+from the data for each column would not remove it. Finally, we note
+that most of the high-frequency ($\gsim2$\,Hz) correlated features
+that have been identified tend to be restricted to a small number of
+detectors, and can often be identified and removed using other
+(faster) means. What remains is predominantly low-frequency
+noise. This conclusion is central to our data reduction strategy
+described in the next section.
 
@@
 %------------------------------------------------------------------------------
+
+The approach taken by SMURF to reducing \scuba\ data is to model (and
+remove) predominantly low-frequency noise sources that are correlated
+amongst detectors, and to iterate this process along with estimates of
+the map. Significant experimentation with different models for noise
+sources during \scuba\ commissioning required a highly flexible
+software framework, and a configurable user interface. To achieve this
+goal, while minimizing development time, it was decided to build SMURF
+as a Starlink package \citep{2009ASPC..411..418J}, which provides
+access to a large suite of libraries (including a commanding and
+messaging interface, astrometric coordinate conversions and generation
+of standard WCS information, file formats etc.). Furthermore,
+Starlink\footnote{\url{http://www.starlink.ac.uk}} is
+open-source\footnote{\url{https://github.com/Starlink}}, and already
+used extensively at the Joint Astronomy Centre (host of the JCMT) for
+many other systems \citep{jenness2011}, which helps with
+interoperability. Though originally written in FORTRAN, many of the
+core Starlink libraries are now ported to native C, or at least have a
+C interface. It was therefore decided to develop SMURF in C as well
+(rather than C++, for example, which would have required adding
+further dependencies to Starlink), although we have taken an
+object-oriented approach. For example, all of the data for a given
+noise model is encapsulated in a C structure, and there is a standard
+interface for all functions that handle models (member data and
+functions for the class, respectively). In this way it is easy to
+extend SMURF with new models. Parallelisation is incorporated in the
+most time-consuming low-level routines using threads (e.g., performing
+Fast Fourier Transforms, re-gridding the data), usually handling
+either independent blocks of bolometers, or blocks of time, in each
+thread, depending on the nature of the calculation. We have found that
+for typical data sets the processing time scales well with the number
+of central processing unit (CPU) cores, although beyond 8 the returns
+are diminished due to tasks that cannot be parallelised (e.g., reading
+the data from disk).
+
+As we have seen in Sections~\ref{sec:bolosignal}--\ref{sec:pca},
+\scuba\ data are dominated at low frequencies ($\lsim$2\,Hz) by highly
+correlated signals. While it can be significantly reduced using simple
+common-mode removal (subtracting the average signal from all
+bolometers at each time step), the more complicated correlated
+residuals at these low frequencies are difficult to model. Ultimately,
+a very simple approach has been taken to handle both components, for
+which the main steps in a typical reduction are shown in
+Fig.~\ref{fig:dimm}.
 
@@
 
-As described in the introduction, the approach taken by SMURF to
-reducing \scuba\ data is to model (and remove) predominantly
-low-frequency noise sources that are correlated amongst detectors, and
-to iterate this process along with estimates of the map. Significant
-experimentation with different models for noise sources during \scuba\
-commissioning required a highly flexible software framework, and a
-configurable user interface. To achieve this goal, while minimizing
-development time, it was decided to build SMURF as a Starlink package
-\citep{2009ASPC..411..418J}, which provides access to a large suite of
-libraries (including a commanding and messaging interface, astrometric
-coordinate conversions and generation of standard WCS information,
-file formats etc.). Furthermore, Starlink\footnote{\url{http://www.starlink.ac.uk}} is
-open-source\footnote{\url{https://github.com/Starlink}},
-and already used extensively at the Joint Astronomy Centre (host of the JCMT) for
-many other systems \citep{jenness2011}, which helps with interoperability. Though
-originally written in FORTRAN, many of the core Starlink libraries are
-now ported to native C, or at least have a C interface. It was
-therefore decided to develop SMURF in C as well (rather than C++, for
-example, which would have required adding further dependencies to
-Starlink), although we have taken an object-oriented approach. For
-example, all of the data for a given noise model is encapsulated in a
-C structure, and there is a standard interface for all functions that
-handle models (member data and functions for the class,
-respectively). In this way it is easy to extend SMURF with new
-models. Parallelisation is incorporated in the most time-consuming
-low-level routines using threads (e.g., performing Fast Fourier
-Transforms, re-gridding the data), usually handling either independent
-blocks of bolometers, or blocks of time, in each thread, depending on
-the nature of the calculation.
-
-As we have seen in Sections~\ref{sec:bolosignal}--\ref{sec:pca},
-\scuba\ data are dominated at low frequencies ($\lsim$2\,Hz) by highly
-correlated signals. In addition to this $1/f$ noise, there is also
-uncorrelated white noise, and finally residual correlated (but
-non-astronomical) signals that appear at all frequencies with more
-complicated relationships amongst groups of bolometers
-(e.g., Fig.~\ref{fig:pca}). While the low-frequency noise can be
-significantly reduced using simple common-mode removal (subtracting
-the average signal from all bolometers at each time step), these
-remaining correlated noise patterns are much more difficult to
-model. Ultimately, a very simple approach has been taken to handling
-these two contaminating sources, for which the main steps in a typical
-reduction are shown in Fig.~\ref{fig:dimm}.
-
 First, the raw data are passed through a pre-processing stage which
@@
 corrections etc. Next, the iterative process begins (dashed box). Most
-of the low-frequency noise is removed using common-mode
-subtraction. Next, the extinction correction is applied (it is done at
-this stage rather than during pre-processing to avoid multiplying
-large atmospheric signals by any small errors in the flatfield
-correction). At this stage the data resemble the grey bolometer traces
-in Fig.~\ref{fig:bolos_mix}, with some residual baseline drifts still
-visible. These drifts are removed using a high-pass filter implemented
-with Fast Fourier Transforms (FFTs). The residual signal is then
-re-gridded to estimate the map. Finally, the map is used to estimate
-and remove the astronomical signal contamination in the bolometer
-data, leaving a relatively clean data set in which to measure the
-white noise level of each bolometer. Since the common-mode and
-filtering stages will have introduced ringing in the map in the
-vicinity of bright sources, the entire process is iterated. Each
-component in the dashed box of Fig.~\ref{fig:dimm} is re-calculated in
-sequence. In this way, the second time the common-mode is calculated,
-for example, most of the bright astronomical sources in the data have
-already been removed in the previous iteration when the map was
-estimated and subtracted, reducing the amount of ringing in the map
-once it is re-estimated. Note that the extinction and gain denoted in
-a second dashed box are multiplicative rather than additive models,
-and so they must be inverted prior to the start of each iteration to
-preserve the units (see Sections~\ref{sec:comgai} and \ref{sec:ext}).
+of the low-frequency noise is removed using common-mode subtraction
+(\model{COM,GAI}, Section~\ref{sec:comgai}). Then, the extinction
+correction (\model{EXT}, Section~\ref{sec:ext}) is applied. At this
+stage the data resemble the grey bolometer traces in
+Fig.~\ref{fig:bolos_mix}, with some residual baseline drifts still
+visible. These drifts are removed using a high-pass filter
+(\model{FLT}, Section~\ref{sec:flt}) implemented with Fast Fourier
+Transforms (FFTs). The residual signal is then re-gridded to estimate
+the map. Finally, the map is used to estimate and remove the
+astronomical signal contamination in the bolometer data (\model{AST},
+Section~\ref{sec:ast}), leaving a relatively clean data set in which
+to measure the white noise level of each bolometer (\model{NOI},
+Section~\ref{sec:noi}). Since the common-mode and filtering stages
+will have introduced ringing in the map in the vicinity of bright
+sources, the entire process is iterated. Each component in the dashed
+box of Fig.~\ref{fig:dimm} is re-calculated in sequence. In this way,
+the second time the common-mode is calculated, for example, most of
+the bright astronomical sources in the data have already been removed
+in the previous iteration when the map was estimated, reducing the
+amount of ringing in the map once it is re-estimated. Note that the
+extinction correction is a multiplicative factor, and must be inverted
+prior to re-calculating any of the additive model components to
+preserve the data amplitude. Additive model components from the
+previous iteration are generically added back into the time-series
+immediately prior to their re-calculation. However, for the special
+case of the common-mode and high-pass filter stages, they are replaced
+simultaneously at the start of the iteration to assist with
+convergence (see Sections~\ref{sec:converge}).
 
@@
 into memory and concatenated into continuous arrays (\scuba\ data are
-broken up and written to disk every $\sim$30\,s during data
-acquisition). As the data are loaded, they are multiplied by the
-flatfield correction \citep[see Section~2.1 in][]{dempsey2012}. Next,
-a series of data cleaning and filtering procedures are applied; these
-include the removal of large glitches that may hinder the iterative
-solution from converging, or simply tasks that do not need to be
-iterated.
+broken up and written to disk every $\sim$30\,s regardless of the
+observation length during data acquisition). As the data are loaded,
+they are multiplied by the flatfield correction \citep[see Section~2.1
+in][]{dempsey2012}. Next, a series of configurable data cleaning and
+filtering procedures are applied; these include the removal of large
+glitches that may hinder the iterative solution from converging, or
+simply tasks that do not need to be iterated.
 
@@
 
-The method used by SMURF to down-sample is to average together
-multiple samples from the original time-series, $x_i$ to estimate the
-lower-frequency output time-series, $y_j$. In general there will not
-be an integer number of samples from $x_i$ in each of the $y_j$, so
-fractional samples from $x_i$ are used at the $y_j$ time
-boundaries. The algorithm is fast, since each sample in $x_i$ need
-only be visited once. From a spectral point of view, this boxcar
-average is equivalent to applying a sinc-function low-pass
-filter. Such behaviour is desirable since it serves as an
-anti-aliasing filter; most non-white features above the Nyquist
-frequency in $y_j$ that could contaminate the output data are largely
+The method used by SMURF is to average together multiple samples from
+the original time-series, $x_i$ to estimate the lower-frequency output
+time-series, $y_j$. In general there will not be an integer number of
+samples from $x_i$ in each of the $y_j$, so fractional samples from
+$x_i$ are used at the $y_j$ time boundaries. The algorithm is fast,
+since each sample in $x_i$ need only be visited once. From a spectral
+point of view, this boxcar average is equivalent to applying a
+sinc-function low-pass filter. Such behaviour is desirable since it
+serves as an anti-aliasing filter; most non-white features above the
+Nyquist frequency in $y_j$ that could contaminate the output data are
 removed.
@@
 this method. It is therefore not currently available to users.
-
-\subsubsection{Bolometer filtering}
-
-Despite the ability of the map-maker to iteratively remove many noise
-components, under some circumstances it may be desirable to filter the
-data once during the pre-processing step. Three main filtering options
-are available:
-
-\begin{itemize}
-
-\item The most commonly-used pre-processing filter is polynomial
-subtraction. A polynomial of the requested order is fit and removed
-from each bolometer time-series. At a bare minimum, the mean is
-removed from all of the bolometers (order 0) in all reductions
-described in this paper.
-
-\item All of the filters available as part of the iterative Fourier
-Transform Filter (Section~\ref{sec:flt}) can also be applied once
-during pre-processing.
-
-\item As an alternative to the iterative map-making procedure,
-cleaning by principal component analysis (PCA) is available as an
-experimental pre-processing option (Section~\ref{sec:pca}). The most
-significant components in the analysis are identified, and the
-projection of each bolometer time-series along their eigenvectors are
-removed. A single parameter specifies the threshold on the amplitude
-of the eigenvalues to be removed, as a number of standard deviations
-away from the mean value. Subarrays are cleaned independently, once,
-for the full length of each continuous chunk of data. Given the
-computational expense of this method, and initial tests which showed
-little improvement over simple high-pass filtering, this method has
-not yet been explored in detail with \scuba\ data. However, since
-systematic effects seem to come and go, it is possible that PCA could
-be useful for particular data sets.
-
-\end{itemize}
 
@@
 \item smooth the gradient values to determine the local mean gradient
-and subtract this local mean from the total gradient to get residual
-gradient;
+and subtract this local mean from the total gradient to get the
+residual gradient;
 
@@
 
-\item Group these high residual gradients into contiguous blocks of
+\item group these high residual gradients into contiguous blocks of
 samples;
@@
 
-\begin{itemize}
+\begin{enumerate}
+
 \item Apodisation: a number of samples at the start and end of each
@@
 
-\end{itemize}
+\end{enumerate}
+
+\subsubsection{Bolometer filtering}
+
+Despite the ability of the map-maker to iteratively remove many noise
+components, under some circumstances it may be desirable to filter the
+data once during the pre-processing step. Three main filtering options
+are available:
+
+\begin{enumerate}
+
+\item The most commonly-used pre-processing filter is polynomial
+subtraction. A polynomial of the requested order is fit and removed
+from each bolometer time-series. At a bare minimum, the mean is
+removed from all of the bolometers (order 0) in all reductions
+described in this paper.
+
+\item All of the filters available as part of the iterative Fourier
+Transform Filter (Section~\ref{sec:flt}) can also be applied once
+during pre-processing.
+
+\item As an alternative to the iterative map-making procedure,
+cleaning by principal component analysis (PCA) is available as an
+experimental pre-processing option (Section~\ref{sec:pca}). The most
+significant components in the analysis are identified, and the
+projection of each bolometer time-series along their eigenvectors are
+removed. A single parameter specifies the threshold on the amplitude
+of the eigenvalues to be removed, as a number of standard deviations
+away from the mean value. Subarrays are cleaned independently, once,
+for the full length of each continuous chunk of data. Given the
+computational expense of this method, and initial tests which showed
+little improvement over simple high-pass filtering, this method has
+not yet been explored in detail with \scuba\ data. However, since
+systematic effects seem to come and go, it is possible that PCA could
+be useful for particular data sets.
+
+\end{enumerate}
 
 \subsubsection{Additional bad bolometer rejection}
-
-Despite all of the cleaning operations described in the previous
-sections, the data from a given bolometer may be unusable do to it
-being poorly biased, having an incorrect flatfield correction applied,
-or having some other particularly pathological noise
-contamination. Most of these bolometers can be flagged simply by
-identifying outliers in the bolometer noise distribution. By default,
-SMURF measures the PSD of all bolometers between 2 and 10\,Hz, after
-all other pre-processing steps have been run (identical to the
-measurement in Section~\ref{sec:noi}). Despite having significant
-low-frequency noise (and possibly bright astronomical source)
-contamination, this higher-frequency portion of the PSD is generally
-quite clean. Furthermore, even if there is contamination, the purpose
-of this measurement is to identify outliers, rather than provide a
+\label{sec:flagbadbol}
+
+Despite the cleaning operations described in the previous sections,
+the data from a given bolometer may be unusable do to it being poorly
+biased, having an incorrect flatfield correction applied, or having
+some other particularly pathological noise contamination. Most of
+these bolometers can be flagged simply by identifying outliers in the
+distribution of bolometer white noise levels. By default, SMURF
+measures the PSD of all bolometers between 2 and 10\,Hz, after all
+other pre-processing steps have been run (identical to the measurement
+in Section~\ref{sec:noi}). Despite having significant low-frequency
+noise (and possibly bright astronomical source) contamination, this
+higher-frequency portion of the PSD is generally quite
+clean. Furthermore, even if there is contamination, the purpose of
+this measurement is to identify outliers, rather than provide a
 meaningful absolute measurement of a given bolometer's white noise
@@
 flatfield correction being applied) outliers from the centre of the
-distribution are flagged. The measurement is based on the logarithm of
-the noise values, and iterative clipping of the mean and standard
-deviation are used, to reduce the impact of large outliers.
+logarithm (to reduce the impact of outliers) of the bolometer noise
+distribution are flagged.
 
@@
 
-Once the data have been cleaned, and the worst data flagged to be
-ignored, the iterative solution for the map and contaminating noise
-signals begins.
+Once the data have been cleaned, and the worst data flagged, the
+iterative solution for the map and contaminating noise signals begins.
 
 First, we describe the generic model for \scuba\ data. We express the
-signal observed by the $i$th bolometer as a function of time,
+digitised signal observed by the $i$th bolometer as a function of
+time,
 %
@@
 units of power delivered to the detectors (pW). The scale factor $f_i$
-converts this effective power to the digitized units recorded on disk
-(DAC) -- the flatfield multiplied by a digitization constant (applied
+converts this effective power to the digitised units recorded on disk
+(DAC) -- the flatfield multiplied by a digitisation constant (applied
 once during pre-processing) -- which in this formulation is assumed to
@@
 relationship that would lead to it being included in
-$\mathbf{n}^\mathrm{c}(t)$. The primary goal of map-making is to model
-and remove $\mathbf{n}^\mathrm{c}(t)$ and $\mathbf{n}^\mathrm{f}_i(t)$
-from the time-series, and producing a map from what is left.  This
-procedure is summarized in Fig.~\ref{fig:dimm}.
-
-Each model component is estimated from the data in a user-defined
-sequence, where the typically strongest components are estimated
-first, and then subtracted, leaving a cleaner signal for subsequent
-model estimates. One exception to this rule is the extinction
-correction, $\mathbf{e}_i(t)$ in Eq.~\ref{eq:model}, which is simply a
-multiplicative factor that is known prior to map-making. Note that the
-order in which this correction occurs will modify the units of
-subsequently calculated additive model components (e.g., additional
-brackets may be needed in Eq.~\ref{eq:model} after
-$\mathbf{e}_i(t)$). Another exception is the calculation of
-$\mathbf{n}^\mathrm{c}(t)$ and $\mathbf{n}^\mathrm{f}_i(t)$ in which
-the previous iterations are usually added back into the time-series
-simultaneously to avoid degeneracies (see Section~\ref{sec:converge}).
-
-Usually the last two components estimated are $\mathbf{a}_i(t)$ and
-$\mathbf{w}_i(t)$ once the other contaminating noise sources, and
-extinction, have been corrected for. Provided that these two signals
-are the only components of the residual bolometer signal, re-gridding
-into a map provides the highest \snr\ estimate of the astronomical
-sky. The signal produced by this map is then projected back into the
-time-domain to estimate and remove $\mathbf{a}_i(t)$, leaving a
-residual signal that should be dominated by white noise,
-$\mathbf{n}_i^\mathrm{w}(t)$.
+$\mathbf{n}^\mathrm{c}(t)$.
+
+During pre-processing, the map-maker divides by $f_i$ once. Then, in
+each iteration, the map-maker divides by $f_i$, models and removes
+$g_i\mathbf{n}^\mathrm{c}(t)$, divides by $\mathbf{e}_i(t)$, and
+applies a high-pass filter to remove $\mathbf{n}^\mathrm{f}_i(t)$ from
+the bolometer time-series. This procedure isolates the astronomical
+signal and white noise, $\mathbf{a}_i(t) +
+\mathbf{n}^\mathrm{w}_i(t)$, for estimating the map. Finally, once the
+map is estimated it is projected into the time-domain and removed from
+the bolometer time-series leaving $\mathbf{n}^\mathrm{w}_i(t)$, in
+which the bolometer noise can be measured. This is the most typical
+sequence; in practice the user can select an arbitrary order, although
+the solution converges much faster when the strongest components are
+estimated first, and then subtracted, leaving a cleaner signal for
+subsequent model estimates.
 
@@
 less contamination in $\mathbf{n}^\mathrm{c}(t)$, and less negative
-bowling in the map. Any multiplicative models (e.g.,
-$\mathbf{e}_i(t)$) are inverted immediately prior to the start of a
+bowling in the map. Any multiplicative models [e.g.,
+$\mathbf{e}_i(t)$] are inverted immediately prior to the start of a
 new iteration to preserve the units of the data. The iterative
@@
 sensitivities and so the amplitude of the common-mode variations will
-also vary from bolometer to bolometer. Comparing each bolometer time
-stream with the common-mode signal allows an estimate of the relative
-bolometer sensitivity to be obtained. In practice, a least squares
-linear fit is performed between the bolometer time-series and the
-common-mode to determine a gain ($g_i$ in Eq.~\ref{eq:noise}) and
+also vary from bolometer to bolometer. Comparing each bolometer
+time-series with the common-mode signal allows an estimate of the
+relative bolometer sensitivity to be obtained. In practice, a least
+squares linear fit is performed between the bolometer time-series and
+the common-mode to determine a gain ($g_i$ in Eq.~\ref{eq:noise}) and
 offset for each bolometer.  The gain and offset for each bolometer is
@@
 provides an alternative (or additional) flat-fielding strategy to that
-described in section~\ref{sec:bolos}.
+described in Section~\ref{sec:bolos}.
 
@@
 It can be seen from Fig.~\ref{fig:com} that some bolometers depart
-radically from the common-mode, indicating some problem within the
+radically from the common-mode, indicating some problem with the
 bolometer. Such bolometers are identified by calculating the Pearson
-coefficient of correlation between each bolometer time-stream and the
+coefficient of correlation between each bolometer time-series and the
 common mode. Bolometers for which the correlation coefficient is below
-a specified limit, or which have unusually high or low gain compared
+a specified limit, or which have unusually high or low gains compared
 to the other bolometers, are flagged as bad in order to omit them from
@@
 individual bad blocks to be rejected from a bolometer time-series,
-rather than rejecting the whole bolometer. This is beneficial for
-bolometers that follow the common-mode for most of the time but suffer
-from transient problems at specific times. As a warning, rejecting
+rather than rejecting the whole bolometer. As a warning, rejecting
 data based on outlier gain values can be misleading in cases where the
@@
 that were flagged as bad simply because they were poorly correlated
-with the common-mode on the previous iteration are included in the new
-common-mode estimate. If such samples are excluded, there is a strong
-possibility of discontinuities appearing in the \model{COM} model at
-block boundaries.  These in turn can lead to ringing when filtering,
-and instabilities in the convergence process.
+with the common-mode on the previous iteration are, however, included
+in the new common-mode estimate. If such samples are excluded, there
+is a strong possibility of discontinuities appearing in the
+\model{COM} model at block boundaries.  These in turn can lead to
+ringing when filtering, and instabilities in the convergence process.
 
@@
 derived using the JCMT water vapour monitor (WVM), and is not
-considered to be a free parameter in the solution ($\mathbf{e}_i(t)$
-in Eq.~\ref{eq:model}). However, it is applied as part of the
+considered to be a free parameter in the solution [$\mathbf{e}_i(t)$
+in Eq.~\ref{eq:model}]. However, it is applied as part of the
 iterative solution, rather than a pre-processing step, since any small
@@
 
-This model takes the FFT of the bolometer time-stream data, and can
+This model takes the FFT of the bolometer time-series data, and can
 apply both high- and low-pass filters, as well as notch filteres, at
@@
 scale, but converted into a frequency through knowledge of the mean
-telescope slew speed. The time-series may optionally have apodisation
-applied before the transform to avoid ringing (primarily caused by
-wrap-around discontinuities at the ends of the time-series). Finally,
-a whitening filter may also be applied in which a simple $1/f^\alpha +
-\mathrm{constant}$ is fit to the power spectrum of each bolometer, and
-then the bolometer FFT is multiplied by its inverse (though normalised
-to preserve the white noise level).  The signals that are removed from
-the time-series by this process are stored in the \model{FLT}
-container array. Typically this model is used purely as a high-pass
-filter to remove most of the residual uncorrelated $1/f$ noise
-following common-mode removal ($\mathbf{n}^\mathrm{f}_i(t)$ in
-Eq.~\ref{eq:noise}). Low-pass filtering is redundant for two reasons:
-(i) SMURF already low-pass filters and re-samples to a lower sample
-rate, as described in Section~\ref{sec:downsamp}; and (ii) the act of
-re-gridding the data to produce map estimates effectively low-pass
-filters the data at a frequency that corresponds to the inverse of the
-crossing time of a single map pixel. Notch filters have not been
-proven to be useful with \scuba\ data, particularly since line
-features tend to move around (i.e., a dynamic line-detection system
-would have to be developed, and care would have to be taken that
-astronomical sources are not suppressed).
+telescope slew speed. The time-series are generally gap-filled
+(Section~\ref{sec:gaps}) before the transform to avoid ringing
+(primarily caused by wrap-around discontinuities at the ends of the
+time-series). Finally, a whitening filter may also be applied in which
+a simple $1/f^\alpha + \mathrm{constant}$ is fit to the power spectrum
+of each bolometer, and then the bolometer FFT is multiplied by the
+square root of its inverse (though normalised to preserve the white
+noise level).  The signals that are removed from the time-series by
+this process are stored in the \model{FLT} container array. Typically
+this model is used purely as a high-pass filter to remove most of the
+residual $1/f$ noise following common-mode removal
+[$\mathbf{n}^\mathrm{f}_i(t)$ in Eq.~\ref{eq:noise}]. Low-pass
+filtering is redundant for two reasons: (i) SMURF already low-pass
+filters and re-samples to a lower sample rate, as described in
+Section~\ref{sec:downsamp}; and (ii) the act of re-gridding the data
+to produce map estimates effectively low-pass filters the data below a
+frequency that corresponds to the inverse of the crossing time of a
+single map pixel. Notch filters have not been proven to be useful with
+\scuba\ data, particularly since line features tend to move around
+(i.e., a dynamic line-detection system would have to be developed, and
+care would have to be taken that astronomical sources are not
+suppressed).
 
@@
 Map estimation is accomplished using a nearest-neighbour resampling of
-the data onto a pre-defined map grid with. For the $i$th map pixel
+the data onto a pre-defined map grid. For the $i$th map pixel,
 $\mathbf{m}(x_i,y_i)$, the brightness is estimated as the weighted
@@
 in Section~\ref{sec:noi}. This weighting scheme is sensible provided
-that the bolometer data have no correlated (i.e., low-frequency)
+that the bolometer data have no correlated (e.g., low-frequency)
 noise.
@@
 In addition to the signal map, a variance map $\mathbf{v}(x_j,y_j)$ is
-estimated. The default behaviour is to estimate this weighted error on
-the mean given the scatter in the weighted samples. This is
-accomplished by dividing the biased weighted sample variance by the
-number of samples that went into the average (akin to the formula for
-standard error on the mean, but accounting for weights),
+estimated. The default behaviour is to estimate this quantity given
+the scatter in the weighted samples. This is accomplished by dividing
+the biased weighted sample variance by the number of samples that went
+into the average (akin to the formula for standard error on the mean,
+but accounting for weights),
 %
@@
 The map-based despiker flags those $\mathbf{d}_j$ that are further
-than some threshold number of standard deviations,
-$\sqrt{\mathbf{v_p}}$, away from zero so that they are not used in
+than some threshold number of standard deviations
+$\sqrt{\mathbf{v_p}}$ away from zero, so that they are not used in
 subsequent iterations.
@@
 A final option available as part of the \model{AST} model is to apply
-constraints to the map to improve convergence. See the discussion in
-Section~\ref{sec:converge}.
+constraints to the map to improve convergence, which presently include
+setting user-specified or high-\snr\ regions to a value of zero for
+all but the final iteration. See the discussion in
+Section~\ref{sec:converge} and examples in
+Sections~\ref{sec:examples}.
 
@@
 The primary purpose of the noise component, \model{NOI}, is to measure
-the white noise of each bolometer ($\mathbf{n}^\mathrm{w}_i(t)$ in
-Eq.~\ref{eq:noise}). The measurement generally occurs once all of the
-other models have been fit and removed. This noise may then be used to
-estimate weights for the data in subsequent iterations.
+the white noise levels of each bolometer ($\mathbf{n}^\mathrm{w}_i(t)$
+in Eq.~\ref{eq:noise}). The measurement generally occurs once all of
+the other models have been fit and removed. This noise may then be
+used to estimate weights for the data in subsequent iterations.
 
@@
 Eq.~\ref{eq:psd}. If the bolometer noise were produced purely by
-uncorrelated thermal sources (i.e., no other long-scale drifts), with
-no high-frequency line features, and without the attenuation at even
+uncorrelated sources (i.e., no other long time-scale drifts), with no
+high-frequency line features, and without the attenuation at even
 higher frequencies by the anti-aliasing filter, this is the
@@
 (Section~\ref{sec:timedespike}), which may work better with these
-cleaner time-streams.
+cleaner time-series.
 
@@
 noise in the map is measured empirically from the scatter of the data
-points that land in each pixel (Section~\ref{sec:ast}), and
-large-scale noise sources can be checked by analyzing the angular
-power spectra of jackknife maps (e.g., Section~\ref{sec:cosmo}).
+points that land in each pixel (Section~\ref{sec:ast}).
 
@@
 primary reason for developing this model was to make it more robust
-against ringing near the ends of the time-series (since the FFT
-assumes the signal is periodic, and there may be a large step between
-the start and finish), or residual spikes (for which the median
-filtering is particularly useful). However, the gap-filling and
-continuity algorithms that we have employed (Section~\ref{sec:gaps})
-successfully mitigate these problems, and the \model{FLT} model is
-substantially faster,
+against ringing near the ends of the time-series, or residual spikes
+(for which the median filtering is particularly useful). However, the
+de-spiking and gap-filling algorithms that we have employed
+(Section~\ref{sec:gaps}) successfully mitigate these problems, and the
+\model{FLT} model is substantially faster.
 
@@
 higher-order polynomials (see their Fig.~10, as well as the discussion
-and references to earlier work in their Section~4.3).
+and references to earlier work in their Section~4.3). In
+\citet{aguirre2010} iterative PCA cleaning was used to remove the
+atmosphere since the simpler methods mentioned left substantial
+residuals in their data (similar to the \scuba\ data described here,
+although PCA cleaning is prohibitively slow in this case).
 
@@
 solution has converged, this signal should look approximately
-white. Once the white nosie, $\mathbf{n}^\mathrm{w}_i(t)$, has been
+white. Once the white noise, $\mathbf{n}^\mathrm{w}_i(t)$, has been
 measured from the bolometer PSDs in \model{NOI}
@@
 \citet{kovacs2008}. However, note that we are only using this quantity
-as a convergence test, and its value is not used in an absolute
+as a convergence test, and its value is not important in an absolute
 sense. Furthermore, the reference white noise values are calculated
@@
 iterations. The convergence criterion is met once this average change
-is smaller than the parameter \texttt{maptol}. Experimentally we found
-that a threshold of 0.05 gives good results (on average, map pixels
-change by 5\% of the estimated map RMS in subsequent iterations).
-This provides good correspondance with what we would choose ``by
-eye'', and furthermore, letting the solver run for many more
-iterations in several test cases yields insignificant differences.
-
-Another major source of divergence is correlation between \model{COM}
-and \model{FLT}. Since \model{FLT} usually consists of a high-pass
-filter following the application of \model{COM}, \model{COM} is
-completely free to grow any large-scale structure at frequencies below
-the chosen filter edge. While such structure does not appear in the
-map (as it is removed by \model{FLT}), we found that the solution
-could be made to converge significantly faster by ``re-mixing''
-\model{COM} and \model{FLT} at the start of each iteration. In
-practice, at the start of each iteration, the previous iterations of
-each model are both added back into the residual immediately prior to
-re-calculating \model{COM}. In this way, truly common-mode signals,
-even at low-frequencies, do not leak into \model{FLT}.
+is smaller than some threshold. Experimentally we found that a value
+of 0.05 gives good results (on average, map pixels change by 5\% of
+the estimated map RMS in subsequent iterations).  This provides good
+correspondance with what we would choose ``by eye'', and letting the
+solver run for many more iterations in several test cases yields
+insignificant differences.
+
+A major source of divergence is correlation between \model{COM} and
+\model{FLT}. Since \model{FLT} usually consists of a high-pass filter
+following the application of \model{COM}, \model{COM} is completely
+free to grow any large-scale structure at frequencies below the chosen
+filter edge. While such structure does not appear in the map (as it is
+removed by \model{FLT}), we found that the solution could be made to
+converge significantly faster by ``re-mixing'' \model{COM} and
+\model{FLT} at the start of each iteration. In practice, at the start
+of each iteration, the previous iterations of each model are both
+added back into the residual immediately prior to re-calculating
+\model{COM}. In this way, truly common-mode signals, even at
+low-frequencies, do not leak into \model{FLT}.
+
+The most serious cause of divergence is the degeneracy between
+low-frequency signal that is removed, and large-scale structures in
+the map. For example, by subtracting the common-mode signal, the map
+is free to grow large-scale structures that exceed the angular scale
+of the array footprint. Our approach as been to constrain regions of
+the map devoid of sources to a value of zero for all but the final
+iteration in the \model{AST} model (Section~\ref{sec:ast}). Such
+regions are either user-defined in advance, or can be determined from
+the data using a cut on \snr. This technique is explored considerably
+in the examples from Sections~\ref{sec:point} and \ref{sec:extended}.
 
@@
 source for \scuba\ that is commonly used as a primary flux
-calibrator. The CV daisy scan pattern was used, with a scan speed of
-155 arcsec\,sec$^{-1}$. We perform several different reductions of the
+calibrator. The CV daisy pattern was used, with a scan speed of 155
+arcsec\,sec$^{-1}$. We perform several different reductions of the
 data to illustrate the purpose of various model components and the
@@
 Fig.~\ref{fig:degeneracy} shows the residual signals for a single
-bolometer after 2 (black) and 100 (grey) iterations which are nearly
+bolometer after 2 (black) and 100 (grey) iterations, which are nearly
 identical, yet the change in the estimated \model{COM} (green) and
@@
 \model{COM} signal to grow with opposite sign. For comparison, the
-bottom panel of Fig.~\ref{fig:degeneracy} shows the pointing solution
+bottom panel of Fig.~\ref{fig:degeneracy} shows the telescope pointing
 for this section of data, and the shapes of the \model{AST} and
@@
 point-like source at the centre, the result is an
-azimuthally-symmetric $\sinc$ function-like pattern in the map. Note
-that due to the iterative nature of the solution, much of the signal
-in the source is actually removed for subsequent iterations of
-\model{FLT}, but as in the previous case, degeneracies between the map
-and the \model{FLT} and \model{COM} signal components can still lead
-to complicated structures.
-
-The preceding examples illustrate the need for constraints in the map
-solution in many situations. For calibrators (and other previously
-known bright, compact sources), a good, simple prior is to constrain
-the map to a value of zero away from the known locations of
-emission. In Fig.~\ref{fig:pointmaps}d, a solution is produced in an
-identical manner to the previous example, but now setting the map
+azimuthally-symmetric $\sinc$ function-like pattern in the map.
+
+This example illustrate the need for constraints in the map solution
+in many situations. For calibrators (and other previously known
+bright, compact sources), a good, simple prior is to constrain the map
+to a value of zero away from the known locations of emission (inside
+the \model{AST} model, Section~\ref{sec:ast}). In
+Fig.~\ref{fig:pointmaps}d, a solution is produced in an identical
+manner to Fig.~\ref{fig:pointmaps}c, but now setting the map
 explicitly to zero beyond a radius of 60\,arcsec from the location of
@@
 \includegraphics[width=\linewidth]{pointmaps.pdf}
-\caption{Multiple reductions of a 450\,\micron\ constant-velocity
-  daisy scan of Uranus, all scaled between $-$0.002\,pW (white) and
-  +0.002\,pW (black). (a) Reduction in which only common-mode
-  subtraction is used to suppress low-frequency noise, and the
-  reduction is forced to use 2 iterations (after the first iteration
-  an estimate of the source flux is removed, and the noise is measured
-  in the residuals to obtain appropriate weighting for the second and
-  final iteration). Circular streaks are caused by independent
-  low-frequency noise that is not removed by the common-mode (Uranus
-  peak 0.15\,pW). (b) Same as (a) but using 50 iterations,
-  illustrating the degeneracy between large-scale structure and the
-  common-mode (the footprint of working bolometers is also shown for
-  reference, Uranus peak 0.15\,pW). (c) Reduction in which high-pass
-  filtering above 0.775\,Hz (corresponding to a spatial scale of
-  200\,arcsec as indicated) is applied after common-mode removal, but
-  before the map estimate. The map-based convergence test is activated
-  and the solution halts after 10 iterations, but leaving large-scale
-  ringing due to the filter (Uranus peak 0.27\,pW). (d) Same as (c),
-  but now the region of the map beyond the white dot-dashed circle is
+\caption{Multiple reductions of a 450\,\micron\ CV daisy scan of
+  Uranus, all scaled between $-$0.002\,pW (white) and +0.002\,pW
+  (black). (a) Reduction in which only common-mode subtraction is used
+  to suppress low-frequency noise, and the reduction is forced to use
+  2 iterations (after the first iteration an estimate of the source
+  flux is removed, and the noise is measured in the residuals to
+  obtain appropriate weighting for the second and final
+  iteration). Circular streaks are caused by independent low-frequency
+  noise that is not removed by the common-mode (Uranus peak
+  0.15\,pW). (b) Same as (a) but using 50 iterations, illustrating the
+  degeneracy between large-scale structure and the common-mode (the
+  footprint of working bolometers is also shown for reference, Uranus
+  peak 0.15\,pW). (c) Reduction in which high-pass filtering above
+  0.775\,Hz (corresponding to a spatial scale of 200\,arcsec as
+  indicated) is applied after common-mode removal, but before the map
+  estimate. The map-based convergence test is activated and the
+  solution halts after 10 iterations, but leaving large-scale ringing
+  due to the filter (Uranus peak 0.27\,pW). (d) Same as (c), but now
+  the region of the map beyond the white dot-dashed circle is
   constrained to a value of zero until all but the final iteration.
@@
 %-------------------------------------------------
-\subsection{Deep blind survey}
+\subsection{Deep point source survey}
 \label{sec:cosmo}
@@
 \citep[e.g.,][]{scott2002,borys2003,laurent2005,coppin2006,scott2008,perera2008,devlin2009}.
-In this section we will reduce scans of the Lockman Hole taken during
-S2SRO as a pilot project for the Cosmology Legacy Survey. It consists
-of $\sim8.5$\,hours of data taken in 36 separate scans (average
+In this section we reduce scans of the Lockman Hole taken during S2SRO
+as a pilot project for the Cosmology Legacy Survey. It consists of
+$\sim8.5$\,hours of data taken in 36 separate scans (average
 15\,min. each) spread over February and March 2010, covering an area
@@
 the cutoff is only removed through iterative common-mode
-subtraction. The map is shown in Fig.~\ref{fig:lockman_maps}a. The
-map-maker has been tested in two ways: (i) large numbers of iterations
-are used to verify that the maps converge without the growth of large
-structure; and (ii) adding synthetic sources to the data at a range of
-brightnesses verify that the map-maker response to them is linear
-(i.e., the relative shape and amplitude compared to the input source
-is independent of brightness). The response to a synthetic point
-source (solid line) after map-making (dotted line) is shown in
-Fig.~\ref{fig:lockman_psf}. Clearly the use of a high-pass filter as a
-pre-processing step, and having no other map-constraints, has the
-down-side of introducing sidelobes around the main peak. However, the
-way this filter affects point-sources is both well-known, and linear.
+subtraction. Since the data are high-pass filtered prior to the
+iterative solution, \model{GAI} (fitting an independent amplitude of
+\model{COM} to each bolometer) has been de-activated, since there is
+very little structure in the common-mode with which to fit an accurate
+gain. The map is shown in Fig.~\ref{fig:lockman_maps}a. The map-maker
+has been tested in two ways: (i) large numbers of iterations are used
+to verify that the maps converge without the growth of large
+structure; and (ii) adding synthetic sources to the time-series data
+at a range of brightnesses verify that the map-maker response to them
+is linear (i.e., the relative shape and amplitude compared to the
+input source is independent of brightness). The response to a
+synthetic point source (solid line) after map-making (dotted line) is
+shown in Fig.~\ref{fig:lockman_psf}. Clearly the use of a high-pass
+filter as a pre-processing step, and having no other map-constraints,
+has the down-side of introducing sidelobes around the main
+peak. However, the way this filter affects point-sources is both
+well-known, and linear.
 
@@
 astronomical sources, and what is probably residual low-frequency
-noise,q causing faint ripples visible to the naked eye. Since the
+noise, causing faint patchiness visible to the naked eye. Since the
 mixture of the two components is unknown, the first step is to
@@
 frequencies, both the real map and the jackknife power spectra grow
-significantly, as a result of the more obvious large-scale undulations
+significantly, as a result of the more obvious large-scale patchiness
 in Figs.~\ref{fig:lockman_maps}(a) and (d).
@@
 pixels in the entire region). Recall that given the small pixel size
-of the map, several pixels generally contribute to each peak. Note
-that this procedure gives only a flavour of the analysis that is
-usually required to produce a robust source lists. Additional tests,
-along with a careful consideration of completeness and false-positive
-rates usual require a series of Monte Carlo simulations that are
-beyond the scope of this work. We direct the interested reader to a
-selection of papers on the subject:
+of the map, several pixels generally contribute to each peak. This
+procedure gives only a flavour of the analysis that is usually
+required to produce a robust source lists. Additional tests, along
+with a careful consideration of completeness and false-positive rates
+usual require a series of Monte Carlo simulations that are beyond the
+scope of this work. We direct the interested reader to a selection of
+papers on the subject:
 \citet{scott2002,coppin2006,perera2008,2009ApJ...707.1201W} and
@@
 submillimetre galaxies due to their relatively low surface densities
-(compare with optical catalogues, for example), and a strong
+(compared with optical catalogues, for example), and a strong
 correlation between the radio and submillimetre emission mechanisms
@@
 default parameters (the solution halted after reaching the default
-map-based convergence criterion in 30 iterations). Panel (a) also
+map-based convergence criterion in 17 iterations). Panel (a) also
 depicts the array footprint (position angle indicative of the start of
@@
 In this final example, we analyze a map of M17 which contains bright,
-extended emission, in contrast with the earlier two examples. A
-promise of \scuba\ is to recover larger angular scales than what was
-possible with earlier-generation ground-based submillimetre
-cameras. The two main factors contributing to this expectation are:
-(i) a larger array footprint on the sky which should make the
-subtraction of common sky-emission signals more effective; and (ii)
-fast detectors (with a correspondingly fast sample rate) to enable
-higher slew-rates (up to 600\,arcsec\,s$^{-1}$) to capture as much
-large-scale structure as possible above the detector $1/f$ knees. We
-produce maps of observation 11 on 20110531 using the 850\,\micron\
-array. It is a rotating PONG covering a diameter of 0.375\,deg, with a
-scan speed of 300\,arcsec\,s$^{-1}$, and a transverse spacing of
-180\,arcsec, taking 37.5\,min to complete.
+extended emission. The data are from observation 11 on 20110531 using
+the 850\,\micron\ array. It is a rotating PONG covering a diameter of
+0.375\,deg, with a scan speed of 300\,arcsec\,s$^{-1}$, and a
+transverse spacing of 180\,arcsec, taking 37.5\,min to complete.
 
@@
 encompassed by the mask, negative bowling is significantly reduced,
-and the brightest regions are clearly more extended.
+and the brightest regions are more extended.
 
@@
 zero-masked regions for each half of the data (note for the
-300\,arcsec case that the region about M17 matches fairly closely the
+300\,arcsec case that the region about M17 fairly closely matches the
 mask in Fig.~\ref{fig:m17}d). Clearly as the filter scale is
@@
 \includegraphics[width=0.49\linewidth]{cor_pspec_m17_default.pdf}
-\caption{Angular power spectral densities (PSDs) for the region of the
-M17 map in Fig.~\ref{fig:m17_jk} containing synthetic data, using the
-default configuration. Left: raw PSDs for the input (noiseless)
-simulated data (thick black line), the signal (average of each half)
-PSDs (thin solid lines), and noise PSDs estimated from the jackknives
-(dashed lines). Vertical dotted lines indicate the high-pass filter
-scales: 150\,arcsec (red); 300\,arcsec (orange); 600\,arcsec (green);
-and 900\,arcsec (blue). PSDs are also colour-coded by filter
-scale. Right: since the input PSD is known, it is possible to measure
-the transfer function of the map-maker as the ratio between the output
-map signal PSDs and the input PSD (from the left panel), giving the
-thin coloured lines (linear vertical axis shown on right of plot). The
-remaining lines are as in the left panel, but now corrected by the
-transfer function. This plot shows that increasing the filter scale
-improves the \snr\ at intermediates scales ($\sim$200--600\,arcsec),
-although the \snr\ worsens at smaller scales ($\lsim$200\,arcsec).}
+\caption{Angular PSDs for the region of the M17 map in
+Fig.~\ref{fig:m17_jk} containing synthetic data, using the default
+configuration. Left: raw PSDs for the input (noiseless) simulated data
+(thick black line), the signal (average of each half) PSDs (thin solid
+lines), and noise PSDs estimated from the jackknives (dashed
+lines). Vertical dotted lines indicate the high-pass filter scales:
+150\,arcsec (red); 300\,arcsec (orange); 600\,arcsec (green); and
+900\,arcsec (blue). PSDs are also colour-coded by filter scale. Right:
+since the input PSD is known, it is possible to measure the transfer
+function of the map-maker as the ratio between the difference of the
+output map signal PSDs and jackknife PSDs, and the input PSDs, giving
+the thin coloured lines (linear vertical axis shown on right of
+plot). The remaining lines are as in the left panel, but now corrected
+by the transfer function. This plot shows that increasing the filter
+scale improves the \snr\ at intermediates scales
+($\sim$200--600\,arcsec), although the \snr\ worsens at smaller scales
+($\lsim$200\,arcsec).}
 \label{fig:m17_def_ps}
@@
 
-While the reduction in the bottom panels of Fig.~\ref{fig:m17} is
-clearly superior, in at least a cosmetic sense to those in the top
-panels, it is important to quantify both the noise properties of the
-maps and their response to real structures (the transfer function). We
-would also like to know how each are affected by our choice of filter
-scale.  Similar to the previous section, we will use a jackknife test
-to estimate the noise, as well as injecting known sources into the
-data to observe how they are attenuated.
+While the reduction in the bottom panels of Fig.~\ref{fig:m17} is, at
+least a cosmetic sense, superior, to those in the top panels, it is
+important to quantify both the noise properties of the maps, and the
+response to real structures (the transfer function). We would also
+like to know how each are affected by our choice of filter scale.
+Similar to the previous section, we will use a jackknife test to
+estimate the noise, as well as injecting known sources into the data
+to observe how they are attenuated.
 
@@
 
-Next, we analyse the angular power spectra densities (PSDs) of the
+Next, we analyse the angular power spectral densities (PSDs) of the
 maps to understand the signal and noise properties of the map-maker in
@@
 line), once we correct for the transfer function, we actually win in a
-\snr\ sense at large scales using the larger-scale filter, as would be
-expected.
+\snr\ sense at large scales using the larger-scale filter (the
+corrected noise is lower), as would be expected.
 
@@
 iterative map-makers. For example, \citet{kovacs2008} typically
-restricts the solution to a small fixed number of iterations (of order
-$\sim$10), although we note that there is no guarantee that maps will
-have reasonably converged with such a strategy. For BOLOCAM data of
-the Galactic Plane, \citet{aguirre2010} used a Maximum-Entropy
-filtering step to suppress large-scales. As in the case of the
-\snr-based masking described here, Maximum-Entropy filtering is
-non-linear, meaning that it too requires simulations to establish its
-transfer function and noise properties as a function of angular scale.
+restricts the solution to a small fixed number of iterations
+($\lsim$10), so that there is probably little chance that such
+structures have the chance to grow. Experimentation with the model
+order is also advocated to gain an impression of the convergence
+properties. This approach is perhaps more relevant to their SHARC-2
+data for which a more complicated model is developed; the degeneracies
+we have observed for \scuba\ maps using our more brute-force approach
+with a high-pass filter tend to look similar regardless of the order
+(only the convergence time is affected).  For Bolocam data of the
+Galactic Plane, \citet{aguirre2010} used a maximum-entropy filtering
+step to suppress large-scales. Regardless of the method used,
+simulations are always required to establish the transfer function and
+noise properties as a function of angular scale.
 
@@
 \end{enumerate}
-
-%\scuba\ is being re-commissioned. Maybe we will have some initial data
-%on fridge performance? If both it, and sky noise are relatively
-%well-behaved, might be able to restrict filtering to lower frequencies
-%and get larger-scale structures.
-
-%Larger array footprint also helps.
-
-%Talk about potential of Pascale et al. method to get closer to
-%max-likelihood map solution (handle maps at constant scan angles with
-%weights in Fourier space separately).
 

### 2012-09-10 smurf_paper.tex
conclusions and acknowledgements

@@
   feedback to observers, as well as for the production of generic
-  science products at the Canadian Astronomy Data Centre. Three
-  detailed case studies are used to: (i) explore convergence
-  properties of the map-maker using simple prior constraints (Uranus
-  -- a point source); (ii) achieve the white-noise limit for faint
-  point-source studies (extragalactic blank-field survey of the
-  Lockman Hole); and (iii) demonstrate that our strategy is capable of
-  recovering angular scales comparable to the size of the array
-  footprint (approximately 5\,arcmin) for bright extended sources
-  (star-forming region M17).
+  science products for the JCMT Science Archive at the Canadian
+  Astronomy Data Centre. Three detailed case studies are used to: (i)
+  explore convergence properties of the map-maker using simple prior
+  constraints (Uranus -- a point source); (ii) achieve the white-noise
+  limit for faint point-source studies (extragalactic blank-field
+  survey of the Lockman Hole); and (iii) demonstrate that our strategy
+  is capable of recovering angular scales comparable to the size of
+  the array footprint (approximately 5\,arcmin) for bright extended
+  sources (star-forming region M17).
 \end{abstract}
@@
 cleaning to avoid the suppression of large-scale
-structure]{aguirre2010}.
+structure]{aguirre2011}.
 
@@
 telescope. The pipeline is also used to generate near-science grade
-products using the JCMT Science Archive \citep{2011ASPC..442..203E}
+products for the JCMT Science Archive \citep{2011ASPC..442..203E}
 hosted by the Canadian Astronomy Data Centre (CADC). Details of the
@@
 and references to earlier work in their Section~4.3). In
-\citet{aguirre2010} iterative PCA cleaning was used to remove the
+\citet{aguirre2011} iterative PCA cleaning was used to remove the
 atmosphere since the simpler methods mentioned left substantial
@@
 (only the convergence time is affected).  For Bolocam data of the
-Galactic Plane, \citet{aguirre2010} used a maximum-entropy filtering
+Galactic Plane, \citet{aguirre2011} used a maximum-entropy filtering
 step to suppress large-scales. Regardless of the method used,
@@
 (SMURF), which was designed to produce maps from the rapidly sampled
-$\sim10^4$ bolometers of the \scuba\ instrument. We have presented
-sufficient detail in order to be useful not only to users of \scuba,
-but for similar detector arrays in the future. The basic elements are:
-
-\begin{enumerate}
-
-\item A simple model which removes most noise in a common-mode
-rejection step, along with high-pass filtering, results in maps that
-approach the white-noise performance of the detectors.
-
-\item The main problem to overcome with our strategy is the divergence
-of the map on large angular scales, due to the degeneracy between the
-map, common-mode, and other residual low-frequency noise
-contributions. A simple strategy of constraining empty regions of the
-map to zero (either a user-supplied mask for known sources, or an
-iterative determination of signal above some \snr\ threshold) provides
-good constraints for compact objects, and bright/extended structures.
-
-\item For maps of faint point sources, a single (non-iterative)
-high-pass filter at the start of the reduction produces maps that are
-nearly white-noise limited. We illustatre how to completely whiten the
-maps using jackknife measurements of the angular noise power spectrum
-in order to reliably extract sources down to the limits of the map.
-
-\item The map-maker has an automatic stopping criterion based on the
-change in the map and $\chi^2$ between subsequent iterations, allowing
-it to adapt to a wide range of data in a pipeline setting.
-
-\item Principal Component Analysis (PCA) has been used to identify
-correlated noise contributions to \scuba\ data, and it is available as
-a generic data cleaning option.
-
-\item The memory requirements and execution times are reasonable
-(possible to reduce data in less time than it takes to collect) on
-modern high-end desktop computers due to the simplicity of the SMURF
-algorithm. It is used in real-time at the telescope for feedback
-during observations, and it is also regularly produces nearly
-science-grade products at the Canadian Astronomy Data Centre.
-
-\item SMURF is highly configurable, making it easy to vary the order
-and types of models that are fit to the data. It is also easy to add
-new models to the existing infrastructure.
-
-\item It is believed that a maximum-likelihood map-maker, such as
-SANEPIC, will provide improved maps over the iterative algorithm
-described here. However, SMURF will most likely be used as an initial
-step in such processing since it can quickly clean the bolometer
-time-series, as well as perform map-based despiking (a necessarily
-iterative step).
-
-\end{enumerate}
+$\sim10^4$ bolometers of the \scuba\ instrument. While similar to
+other algorithms in the literature that iteratively model and remove
+correlated noise components from the data, successively improving the
+map \citep[e.g.,][]{kovacs2008,aguirre2011}, we have shown that a
+relatively simpler approach provides good results for \scuba\ data,
+with reasonable computational requirements. This conclusion is
+encouraging for the development of future large-format bolometer
+arrays.
+
+%We have presented
+%sufficient detail in order to be useful not only to users of \scuba,
+%but for similar detector arrays in the future. The basic elements are:
+
+The most significant correlated noise sources in \scuba\ data occur at
+low-frequencies ($\lsim$2\,Hz), and are probably a mixture of
+atmospheric signals and magnetic field pickup. Most of this signal can
+be removed using common-mode rejection, although a principal component
+analysis (PCA) identifies a number of other less-significant
+correlated noise patterns that also dominate at low
+frequencies. Unfortunately these patterns are difficult to model, and
+their character varies from observation to observation. One could use
+the PCA directly to remove them, but this technique is prohibitively
+slow for existing high-end desktop computers given the volume of
+\scuba\ data. However, we have demonstrated that high-pass filtering
+can remove the bulk of these residual correlated noise components
+using significantly more efficient Fast Fourier Transforms (FFTs). The
+basic algorithm therefore consists of iterative common-mode rejection
+and high-pass filtering along with estimates of the map, which allows
+us to approach the white-noise limit of the instrument.
+
+The main problem to overcome is divergence on large angular scales,
+due to the degeneracy between the map, and the low-frequency signal
+components that are removed (model degeneracies are a generic problem
+of iterative map-makers). A simple strategy of constraining empty
+regions of the map to zero (using either a user-supplied mask for
+known sources, or an iterative determination of signal above some
+\snr\ threshold) provides good constraints for compact objects, and
+bright/extended structures, allowing us to effectively recover angular
+scales up to the order of the array footprint (approximately
+5\,arcmin).
+
+For maps of faint point sources, a single (non-iterative) high-pass
+filter at the start of the reduction produces maps that are nearly
+white-noise limited. Residual large-scale noise can be removed with a
+whitening filter, and sources detected using a matched-filter
+(smoothing with the effective filtered point spread function).
+
+The iterative solution is stopped once convergence in the $\chi^2$
+statistic, and/or variations in the map itself is achieved. This
+enables SMURF to run in a pipeline setting without user interaction
+for a wide variety of observations. Furthermore, the execution times
+are shorter than the observation lengths, and memory requirements for
+even the longest \scuba\ observations are within the capabilities of
+single, high-end desktop computers. SMURF can therefore provide real-time
+feedback at the telescope to observers, and it also produces near
+science-grade data products for the JCMT Science Archive hosted by the
+Canadian Astronomy Data Centre.
+
+%SMURF is highly configurable, making it easy to vary the order
+%and types of models that are fit to the data. It is also easy to add
+%new models to the existing infrastructure.
+
+One regime in which SMURF does not presently perform well is in maps
+of faint extended structures, since the zero-masking technique we have
+adopted cannot be used. Since SMURF is both highly configurable and
+extensible, it may be possible to develop an improved data model
+and/or map constraint to assist in these situations, as more
+experience with the instrument is gained. However, the best long-term
+prospects are probably to turn to a maximum-likelihood algorithm, such
+as SANEPIC \citep{patanchon2008}.
 
@@
 funding. This research was supported in part by the Natural Sciences
-and Engineering Research Council of Canada. The authors thank the
-members of the SCUBA-2 commissioning team for testing the
-map-maker and reporting anomalies; in particular Antonio Chrysostomou
-and Jessica Dempsey. We also thank Dennis Kelly, Alex van Engelen and
-Jennifer Balfour for work on the SCUBA-2 observation simulator;
-and Mark Thompson, Craig Walther and S\'{e}verin Gaudet for being
-on the Critical Design Review panel.
+and Engineering Research Council of Canada. EC also thanks Matthew
+Hasselfield and Mark Halpern for helpful discussions about multiplexed
+TES arrays and their noise characteristics.  The authors thank the
+members of the SCUBA-2 commissioning team for testing the map-maker
+and reporting anomalies; in particular Antonio Chrysostomou and
+Jessica Dempsey. We also thank Dennis Kelly, Alex van Engelen and
+Jennifer Balfour for work on the SCUBA-2 observation simulator; and
+Mark Thompson, Craig Walther and S\'{e}verin Gaudet for being on the
+Critical Design Review panel. Finally, we thank Per Friberg and Gary
+Davis for their helpful comments on the manuscript.
 

### 2012-09-11 smurf_paper.tex
a few more minor edits

@@
 large given the 600\,arcsec\,s$^{-1}$ maximum scan speed). For areas
-larger than the array footprint a closed ``PONG'' pattern is used, in
-which the boresight travels in approximately straight lines and
-``bounces'' off the edges at 45\,degree angles until the area is
+larger than the array footprint a rectangular ``PONG'' pattern is
+used, in which the boresight travels in approximately straight lines
+and ``bounces'' off the edges at 45\,degree angles until the area is
 uniformly filled in. It is also usually combined with a rotation
@@
 The \model{COM} model is the common-mode signal itself
-($\mathbf{n}^{c}(t)$ in Eq.~\ref{eq:noise}). It is a single time
+[$\mathbf{n}^{c}(t)$ in Eq.~\ref{eq:noise}]. It is a single time
 series that is estimated by finding the mean of all bolometers at each
@@
 squares linear fit is performed between the bolometer time-series and
-the common-mode to determine a gain ($g_i$ in Eq.~\ref{eq:noise}) and
+the common-mode to determine a gain [$g_i$ in Eq.~\ref{eq:noise}] and
 offset for each bolometer.  The gain and offset for each bolometer is
@@
 The primary purpose of the noise component, \model{NOI}, is to measure
-the white noise levels of each bolometer ($\mathbf{n}^\mathrm{w}_i(t)$
-in Eq.~\ref{eq:noise}). The measurement generally occurs once all of
+the white noise levels of each bolometer [$\mathbf{n}^\mathrm{w}_i(t)$
+in Eq.~\ref{eq:noise}]. The measurement generally occurs once all of
 the other models have been fit and removed. This noise may then be
@@
 %------------------------------------------------------------------------------
-\section{Summary and future work}
-\label{sec:summary}
+\section{Conclusions}
+\label{sec:conclusions}
 %------------------------------------------------------------------------------

### 2012-09-11 smurf_paper.tex
Thank an extra person

@@
 members of the SCUBA-2 commissioning team for testing the map-maker
-and reporting anomalies; in particular Antonio Chrysostomou and
-Jessica Dempsey. We also thank Dennis Kelly, Alex van Engelen and
+and reporting anomalies; in particular Antonio Chrysostomou,
+Jessica Dempsey and Wayne Holland. We also thank Dennis Kelly, Alex van Engelen and
 Jennifer Balfour for work on the SCUBA-2 observation simulator; and

### 2012-09-12 smurf_paper.tex
more edits from Douglas

--- 2012-09-11 smurf_paper.tex
@@
 (i) is achieved through fast-scanning of the entire telescope (up to
-600\,arcsec\,sec$^{-1}$ slew speeds), such that significant drift in
-the bolometers due to low-frequency noise occurs more slowly than the
+600\,arcsec\,sec$^{-1}$), such that significant drift in the
+bolometers due to low-frequency noise occurs more slowly than the
 crossing times for the astronomical scales of interest; and (ii) by
@@
 \citep[e.g.,][]{glenn1998,weferling2002,wilson2008,kovacs2008b}, in
-preference to ``chopping'' methods (where the secondary is quickly
-moved to modulate the signal) that were more appropriate for older
+preference to ``chopping'' methods (where the secondary is moved
+quickly to modulate the signal) that were more appropriate for older
 instruments that had poorer low-frequency noise performance, and were
@@
 common to each column's amplifier chain. The relationship between the
-output digitized current, $I$, and the input power, $P$, or $dI/dP$ is
+output digitised current, $I$, and the input power, $P$, or $dI/dP$ is
 established using flatfield observations immediately prior to science
@@
 columns in the eigenvalue maps at both wavelengths (e.g., columns 26
-and 29 at 450\,\micron in Component~6). Finally, these components have
-potentially unrelated features mixed into them, like sudden offsets
-that are likely residuals due to imperfect step correction (e.g., at
-$t\sim50$\,s in the eigenvector time-series for Component~6; also see
-Section~\ref{sec:steps}).
+and 29 at 450\,\micron\ in Component~6). Finally, these components
+have potentially unrelated features mixed into them, like sudden
+offsets that are likely residuals due to imperfect step correction
+(e.g., around $t\sim50$\,s in the eigenvector time-series for
+Component~6; also see Section~\ref{sec:steps}).
 
@@
 row 25). The time-series plot for the eigenvector exhibits a large
-positive tail of sudden excursions which gradually worsen over
+positive tail of sudden excursions, which gradually worsen over
 time. Bolometers that exhibit this signal are probably poorly biased,
@@
 amplification chain, the intensities of these high-frequency signals
-are random in the eigenvalue maps, and a common-mode signal estimated
-from the data for each column would not remove it. Finally, we note
-that most of the high-frequency ($\gsim2$\,Hz) correlated features
-that have been identified tend to be restricted to a small number of
-detectors, and can often be identified and removed using other
-(faster) means. What remains is predominantly low-frequency
+appear random in the eigenvalue maps, and a common-mode signal
+estimated from the data for each column would not remove it. Finally,
+we note that most of the high-frequency ($\gsim2$\,Hz) correlated
+features that have been identified tend to be restricted to a small
+number of detectors, and can often be identified and removed using
+other (faster) means. What remains is predominantly low-frequency
 noise. This conclusion is central to our data reduction strategy
@@
 object-oriented approach. For example, all of the data for a given
-noise model is encapsulated in a C structure, and there is a standard
+noise model are encapsulated in a C structure, and there is a standard
 interface for all functions that handle models (member data and
@@
 correction (\model{EXT}, Section~\ref{sec:ext}) is applied. At this
-stage the data resemble the grey bolometer traces in
+point the data resemble the grey bolometer traces in
 Fig.~\ref{fig:bolos_mix}, with some residual baseline drifts still
@@
 point of view, this boxcar average is equivalent to applying a
-sinc-function low-pass filter. Such behaviour is desirable since it
+sinc-function low-pass filter. Such behaviour is desirable, since it
 serves as an anti-aliasing filter; most non-white features above the
@@
 cycle in an infinitely repeating waveform, any large difference
-between starting and ending value will effectively introduce sudden
+between starting and ending values will effectively introduce sudden
 steps at the start and end of each cycle, causing unwanted
@@
 addition, each end of the time-series is padded with double this
-number of zeros. This method is illustrated in
-Fig.~\ref{fig:pad2}. This method is not used by default as it reduces
-the amount of data available for the map, and can significantly hinder
-very short observations (e.g., of calibrators when focussing the
-telescope).
+number of zeros. This method is illustrated in Fig.~\ref{fig:pad2}. It
+is not used by default as it reduces the amount of data available for
+the map, and can significantly hinder very short observations (e.g.,
+of calibrators when focussing the telescope).
 
@@
 Despite the cleaning operations described in the previous sections,
-the data from a given bolometer may be unusable do to it being poorly
+the data from a given bolometer may be unusable due to it being poorly
 biased, having an incorrect flatfield correction applied, or having
@@
 $\mathbf{n}^\mathrm{f}_i(t)$ is (predominantly low-frequency) noise in
-excess of the white noise level that is either uncorrelated from
+excess of the white noise level, that is either uncorrelated from
 bolometer-to-bolometer, or at least does not have a simple correlation
@@
 \begin{table}
-  \caption{Summary of the model components that can be fit to
-    \scuba\ time-series data with SMURF. Only the first group of
-    models (in the order indicated) are typically fit to the data
-    (\model{COM}--\model{NOI}) in the indicated order. The remaining
-    models (\model{DKS}--\model{PLN}) have only been included for
-    completeness}
+  \caption{Summary of the model components that can be fit to \scuba\
+    time-series data with SMURF. Only the first group of models are
+    typically fit to the data (\model{COM}--\model{NOI}) in the
+    indicated order. The remaining models (\model{DKS}--\model{PLN})
+    are usually omitted, although they are available as options.}
   \vspace{0.2cm}
@@
 flatfielding and removal of a constant baseline. It can be seen that
-most bolometers exhibit a common time variation overlayed with other
-features. Empty squares indicate the locations of dead bolometers.}
+most bolometers exhibit a common time variation overlaid with other
+features. Empty squares indicate the locations of broken bolometers.}
 \label{fig:com}
@@
 The \model{COM} model is the common-mode signal itself
-[$\mathbf{n}^{c}(t)$ in Eq.~\ref{eq:noise}]. It is a single time
-series that is estimated by finding the mean of all bolometers at each
-time step. Bolometer values that have been flagged as unusable are
-excluded from the mean.
+[$\mathbf{n}^\mathrm{c}(t)$ in Eq.~\ref{eq:noise}]. It is a single
+time-series that is estimated by finding the mean of all bolometers at
+each time step. Bolometer values that have been flagged as unusable
+are excluded from the mean.
 
@@
 bolometer. Such bolometers are identified by calculating the Pearson
-coefficient of correlation between each bolometer time-series and the
+correlation coefficient between each bolometer time-series and the
 common mode. Bolometers for which the correlation coefficient is below
@@
 extended emission is indistinguishable from the other sources of
-common-mode signal (e.g., atmosphere variations) and will be removed by
-the model. This places a limit on the spatial extent of astronomical
-structure that can be recovered.
+common-mode signal (e.g., atmosphere variations) and will be removed
+by the \model{COM} model. This places a limit on the spatial extent of
+astronomical structure that can be recovered.
 
@@
 This model takes the FFT of the bolometer time-series data, and can
-apply both high- and low-pass filters, as well as notch filteres, at
+apply both high- and low-pass filters, as well as notch filters, at
 hard frequency edges specified by the user. Alternatively, the
@@
 time-series). Finally, a whitening filter may also be applied in which
-a simple $1/f^\alpha + \mathrm{constant}$ is fit to the power spectrum
-of each bolometer, and then the bolometer FFT is multiplied by the
-square root of its inverse (though normalised to preserve the white
-noise level).  The signals that are removed from the time-series by
-this process are stored in the \model{FLT} container array. Typically
-this model is used purely as a high-pass filter to remove most of the
-residual $1/f$ noise following common-mode removal
+a simple form, $1/f^\alpha + \mathrm{constant}$, is fit to the power
+spectrum of each bolometer, and then the bolometer FFT is multiplied
+by the square root of its inverse (though normalised to preserve the
+white noise level).  The signals that are removed from the time-series
+by this process are stored in the \model{FLT} container
+array. Typically this model is used purely as a high-pass filter to
+remove most of the residual $1/f$ noise following common-mode removal
 [$\mathbf{n}^\mathrm{f}_i(t)$ in Eq.~\ref{eq:noise}]. Low-pass
@@
 In addition to the signal map, a variance map $\mathbf{v}(x_j,y_j)$ is
-estimated. The default behaviour is to estimate this quantity given
+estimated. The default procedure is to estimate this quantity given
 the scatter in the weighted samples. This is accomplished by dividing
@@
 samples per pixel (a situation that is almost never encountered in a
-\scuba\ map except in the edge pixels).
+\scuba\ map, except in the edge pixels).
 
@@
 time-domain despiker (Section~\ref{sec:timedespike}), this calculation
-uses the scatter in the population of samples used to estimate the map
-pixel values (from different times and bolometers) to reject
+utilises the scatter in the population of samples used to estimate the
+map pixel values (from different times and bolometers) to reject
 outliers. This method is more robust against false-positive detections
@@
 weighted sample variance (i.e., the variance map value multiplied by
-the number of samples)
+the number of samples):
 %
@@
 
-Finally, it should be noted that the default behaviour is to calculate
+Finally, it should be noted that the default procedure is to calculate
 the white noise levels once within \model{NOI}, after the second
@@
 (Section~\ref{sec:converge}). Note that the absolute values of the
-noise, and therefore weights calculated by the model component are
-irrelevant (only their relative values). The reason is that the final
-noise in the map is measured empirically from the scatter of the data
-points that land in each pixel (Section~\ref{sec:ast}).
+noise, and therefore weights calculated by the model, are irrelevant
+(only their relative values matter). The reason is that the final
+noise in the map is measured empirically from the scatter of the
+weighted data points that land in each pixel (Section~\ref{sec:ast}).
 
@@
 signal distribution across the focal plane at each time slice in an
-attempt to remove possible coherent atmospheric sky-noise structures.
+attempt to remove possible coherent atmospheric sky-noise structure.
 
@@
 cases, while the models had some success at removing the magnetic
-field pickup, it did not reduce the $1/f$ knee substantially,
+field pickup, they did not reduce the $1/f$ knee substantially,
 necessitating the continued use of the \model{FLT} model to remove the
@@
 finding only marginal improvements by fitting a plane, or even using
-higher-order polynomials (see their Fig.~10, as well as the discussion
-and references to earlier work in their Section~4.3). In
+higher-order polynomials (see their figure~10, as well as the
+discussion and references to earlier work in their section~4.3). In
 \citet{aguirre2011} iterative PCA cleaning was used to remove the
@@
 convergence criterion is met if the change in subsequent values of
-$\chi^2_\mathrm{r}$ is smaller than the parameter \texttt{chitol},
-which is set to $10^{-3}$ by default.
+$\chi^2_\mathrm{r}$ is smaller than the requsted threshikd, which is
+set to $10^{-3}$ by default.
 
@@
 the estimated map RMS in subsequent iterations).  This provides good
-correspondance with what we would choose ``by eye'', and letting the
+correspondence with what we would choose ``by eye'', and letting the
 solver run for many more iterations in several test cases yields
@@
 and final iteration. The resulting map after these two iterations is
-shown in Fig.~\ref{fig:pointmaps}a. While the \snr\ of Uranus is clearly
-large ($\sim$250 as estimated by SMURF), enabling us to see the faint
-sidelobes (circle and cross pattern), the map also has obvious
-circular streaks and other large-scale ripples. The circular streaks
-are due to the fact that \model{COM} does not account for all of the
-low-frequency noise (see Fig.~\ref{fig:pspec}), and therefore each
-bolometer leaves a trace of the circular scan pattern in the map as
-each of their baselines slowly drift independently. A significant
+shown in Fig.~\ref{fig:pointmaps}a. While the \snr\ of Uranus is
+clearly large ($\sim$250 as estimated by SMURF), enabling us to see
+the faint sidelobes (circle and cross pattern), the map also has
+obvious circular streaks and other large-scale ripples. The circular
+streaks are due to the fact that \model{COM} does not account for all
+of the low-frequency noise (see Fig.~\ref{fig:pspec}), and therefore
+each bolometer leaves a trace of the circular scan pattern in the map,
+as each of their baselines slowly drift independently. A significant
 contribution to the larger-scale ripples in the map, however, can be
@@
 
-One improvement is to simply restrict the low-frequency signal that
-goes into the maps by applying a high-pass filter to the bolometer
-time-series. A third reduction of the data uses the default map-making
-parameters, as described in Section~\ref{sec:algorithm}, which adds
-the \model{FLT} model to accomplish this task immediately prior to map
-estimation. We set the filter edge based on an angular scale of
-200\,arcsec, which, given the scan speed of 155\,arcsec\,sec$^{-1}$,
-corresponds to a frequency of 0.78\,Hz. Now using the automatic
-map-based convergence test (Section~\ref{sec:converge}), the resulting
-map converges after 10 iterations (Fig.~\ref{fig:pointmaps}c). Both
-the circular streaks and the large-scale gradient are now removed, but
-they have been replaced by an obvious, circularly-symmetric ringing
-pattern about Uranus. The reason for this ringing is that the
-hard-edged high-pass filter in frequency space is equivalent to a
-$\sinc$ function-like response in map space. Since the scan pattern is
-fairly isotropic (scans at all position angles), and there is a bright
-point-like source at the centre, the result is an
-azimuthally-symmetric $\sinc$ function-like pattern in the map.
+One improvement to remove residual low-frequency noise after
+common-mode removal in this Uranus image, is to simply apply a
+high-pass filter to the data. A third reduction of the data uses the
+default map-making parameters, as described in
+Section~\ref{sec:algorithm}, which adds the \model{FLT} model to
+accomplish this task immediately prior to map estimation. We set the
+filter edge based on an angular scale of 200\,arcsec, which, given the
+scan speed of 155\,arcsec\,sec$^{-1}$, corresponds to a frequency of
+0.78\,Hz. Now using the automatic map-based convergence test
+(Section~\ref{sec:converge}), the resulting map converges after 10
+iterations (Fig.~\ref{fig:pointmaps}c). Both the circular streaks and
+the large-scale gradient are now removed, but they have been replaced
+by an obvious, circularly-symmetric ringing pattern about Uranus. The
+reason for this ringing is that the hard-edged high-pass filter in
+frequency space is equivalent to a $\sinc$ function-like response in
+map space. Since the scan pattern is fairly isotropic (scans at all
+position angles), and there is a bright point-like source at the
+centre, the result is an azimuthally-symmetric $\sinc$ function-like
+pattern in the map.
 
@@
   peak 0.15\,pW). (c) Reduction in which high-pass filtering above
-  0.775\,Hz (corresponding to a spatial scale of 200\,arcsec as
+  0.775\,Hz (corresponding to a spatial scale of 200\,arcsec, as
   indicated) is applied after common-mode removal, but before the map
@@
 \scuba\ surveys designed to detect extremely faint point-sources
-(e.g., high-redshift star-forming galaxies, and debris disks) are
-ideally limited by the white-noise performance of the instrument. The
-approach described here for maximizing the \snr\ of point-sources
-involves three major steps: (i) generating a map that removes most
-low-frequency noise sources with approximately linear response,
-without prior knowledge of the location of sources; (ii) whitening
-(essentially flattening) the map to make the noise properties of the
-map easier to deal with; and (iii) detecting point sources using a
-``matched filter''. Note that variations on this general procedure
-have been used extensively in the submm cosmology community using
-previous instruments
+(e.g., high-redshift star-forming galaxies, and features in debris
+disks) are ideally limited by the white-noise performance of the
+instrument. The approach described here for maximising the \snr\ of
+point-sources involves three major steps: (i) generating a map that
+removes most low-frequency noise sources with approximately linear
+response, without prior knowledge of the location of sources; (ii)
+whitening (essentially flattening) the map to make the noise
+properties of the map easier to deal with; and (iii) detecting point
+sources using a ``matched filter''. Note that variations on this
+general procedure have been used extensively in the submm cosmology
+community using previous instruments
 \citep[e.g.,][]{scott2002,borys2003,laurent2005,coppin2006,scott2008,perera2008,devlin2009}.
@@
   second contour of panel (a)]. The image shows the \snr\, with
-  3.8-$\sigma$ peaks indicated with blue circles (radius
-  8\,arcsec). The orange ``$+$'' signs show the locations of 1.4\,GHz
-  radio sources from \citet{owen2008} with $S\gsim15$\,$\mu$Jy. Of the
-  10 submm peaks, 9 are within this search radius of at least one
-  radio source. (d) Jackknife map produced from the difference of two
-  maps, using the even and odd scan numbers, respectively. Provided
-  that all noise sources are statistically uncorrelated between the
-  two halves of the data, the map is a plausible realization of the
-  noise without contamination from astronomical sources. (e) The
-  jackknife map whitened using the same filter as that in panel
-  (b). The whitened jackknife map cross-correlated with the same
-  whitened PSF as in panel (c). Again, orange ``$+$'' and blue circles
-  indicate radio sources and 3.8-$\sigma$ peaks, respectively. Unlike
-  panel (c), there is only a single (apparently) significant peak, and
-  it is not in the vicinity of a radio source.}
+  3.8-$\sigma$ peaks indicated by blue circles (radius 8\,arcsec). The
+  orange ``$+$'' signs show the locations of 1.4\,GHz radio sources
+  from \citet{owen2008} with $S\gsim15$\,$\mu$Jy. Of the 10 submm
+  peaks, 9 are within this search radius of at least one radio
+  source. (d) Jackknife map produced from the difference of two maps,
+  using the even and odd scan numbers, respectively. Provided that all
+  noise sources are statistically uncorrelated between the two halves
+  of the data, the map is a plausible realisation of the noise without
+  contamination from astronomical sources. (e) The jackknife map
+  whitened using the same filter as that in panel (b). The whitened
+  jackknife map cross-correlated with the same whitened PSF as in
+  panel (c). Again, orange ``$+$'' and blue circles indicate radio
+  sources and 3.8-$\sigma$ peaks, respectively. Unlike panel (c),
+  there is only a single (apparently) significant peak, and it is not
+  in the vicinity of a radio source.}
 \label{fig:lockman_maps}
@@
 mitigated, and the method used in this example (the default processing
-in SMURF) is to apply a high-pass filter to the data once as a
+in SMURF) is to apply a high-pass filter to the data once, as a
 pre-processing step. The iterative solution is then run using only
@@
   Fig.~\ref{fig:lockman_pspec}) to the dotted line gives the whitened
-  line, which is the expected shape of a source in panel (b) from
-  Fig.~\ref{fig:lockman_maps}. Finally, cross-correlating the whitened
-  map with this whitened PSF is an effective ``matched filter'' for
-  identifying point-like sources, and this smoothed map is shown in
-  panel (c) of Fig.~\ref{fig:lockman_maps}.}
+  line (dashed) , which is the expected shape of a source in panel (b)
+  from Fig.~\ref{fig:lockman_maps}. Finally, cross-correlating the
+  whitened map with this whitened PSF is an effective ``matched
+  filter'' for identifying point-like sources, and this smoothed map
+  is shown in panel (c) of Fig.~\ref{fig:lockman_maps}.}
 \label{fig:lockman_psf}
@@
 to produce the two maps using odd and even scan numbers (i.e., each
-map will contain a nearly evenly-spaced sampling of data across the
+map will contain a nearly uniformly-spaced sampling of data across the
 full data set).
@@
 approximately flat at spatial frequencies $\gsim 0.06$\,arcsec$^{-1}$
-(scales $\lsim16$\,arcsec), with the exception of a line at
+(scales $\lsim16$\,arcsec), with the exception of a spike at
 $\sim0.175$\,arcsec$^{-1}$ (a scale of $\sim5.7$\,arcsec). One
@@
   (panels (a), (d), and (b) from Fig.~\ref{fig:lockman_maps},
-  respectively, considering only the square region indicated in
+  respectively), considering only the square region indicated in
   Fig.~\ref{fig:lockman_maps}d. Since the raw map contains spatially
@@
   (difference of two approximately equal-length subsets of the data)
-  is used to generate a plausible realization of pure noise. Assuming
+  is used to generate a plausible realisation of pure noise. Assuming
   that the noise properties are isotropic, a whitening filter is
@@
 ($\lsim 0.5$\,arcsec$^{-1}$) there is significantly less power than
-the raw map. However, it also clearly has some power in excess of the
-white noise level. In theory, this residual signal is produced by
+for the raw map. However, it also clearly has some power in excess of
+the white noise level. In theory, this residual signal is produced by
 astronomical sources, although its origin cannot be determined from
@@
 assumption, cross-correlation between the map and the known PSF yields
-the maximum-likelihood flux density of supposed point-sources centered
+the maximum-likelihood flux density of supposed point-sources centred
 over every location in the resulting map \citep[an extremely
@@
 identification in such smoothed maps have been used extensively in the
-submillimetre community as both an efficient source-detection and
+submillimetre community, as both an efficient source-detection and
 photometric measurement strategy. In terms of the angular power
@@
 low-pass filter, that suppresses noise on scales smaller (and
-therefore higher angular frequencies) than the beam. For the case at
-hand, the effective PSF for the whitened maps is given by the dotted
-curve in Fig.~\ref{fig:lockman_psf}. Maps smoothed by this shape are
-shown for the whitened signal, and jackknife maps in
+therefore at higher angular frequencies) than the beam. For the case
+at hand, the effective PSF for the whitened maps is given by the
+dotted curve in Fig.~\ref{fig:lockman_psf}. Maps smoothed by this
+shape are shown for the whitened signal, and jackknife maps in
 Figs.~\ref{fig:lockman_maps}c and f, respectively. Note that these
@@
   good agreement indicates that these maps are indeed dominated by
-  white noise. The lower panel shows the results for matched filtered
+  white noise. The lower panel shows the results for matched-filtered
   signal and jackknife maps (Figs.~\ref{fig:lockman_maps}c,f). The
@@
 extended emission. The data are from observation 11 on 20110531 using
-the 850\,\micron\ array. It is a rotating PONG covering a diameter of
-0.375\,deg, with a scan speed of 300\,arcsec\,s$^{-1}$, and a
-transverse spacing of 180\,arcsec, taking 37.5\,min to complete.
+the 850\,\micron\ array. It is a rotating PONG scan covering a
+diameter of 0.375\,deg, with a scan speed of 300\,arcsec\,s$^{-1}$,
+and a transverse spacing of 180\,arcsec, taking 37.5\,min to complete.
 
@@
 \includegraphics[width=\linewidth]{m17_jk.pdf}
-\caption{Maps of M17 used to characterize the noise properties and
+\caption{Maps of M17 used to characterise the noise properties and
 transfer function of SMURF (same intensity scale as
@@
 signal within a 600\,arcsec diameter circle (south of M17) created as
-a realization of noise from a $P(k) \propto k^{-3}$ angular power
+a realisation of noise from a $P(k) \propto k^{-3}$ angular power
 spectrum multiplied by the PSF, subtracting the minimum to make it
@@
 
-While the reduction in the bottom panels of Fig.~\ref{fig:m17} is, at
-least a cosmetic sense, superior, to those in the top panels, it is
+While the reduction in the bottom panels of Fig.~\ref{fig:m17} is (at
+least a cosmetic sense) superior to those in the top panels, it is
 important to quantify both the noise properties of the maps, and the
@@
 the noise properties may evolve with time (e.g., due to changing sky
-conditions) leading to a biased estimate of the parent noise
+conditions), leading to a biased estimate of the parent noise
 distribution in the complete map from the jackknife. Also, since the
@@
 power at a range of scales into a relatively empty region of the map.
-It is created by drawing a realization of noise within an 800\,arcsec
-on-a-side box, with an angular power spectrum $P(k) \propto k^{-3}$
+It is created by drawing a realisation of noise within an 800\,arcsec
+on-a-side box, with an angular power spectrum $P(k) \propto k^{-3}$,
 which is appropriate for Galactic cirrus clouds
@@
 columns show reductions using 150, 300, 600, and 900\,arcsec filter
-edges in the reductions. For reference, the largest scale that is
-completely inscribed by the array footprint is about 400\,arcsec, and
-the diagonal of the array is about 600\,arcsec (meaning that scales
-beyond 600\,arcsec are completely unconstrained when common-mode
-rejection is used).  The synthetic data are clearly seen as the
-circular region south of M17. As the filter scale is increased, the
-size of the ripples increases accordingly. While larger structures do
-seem to appear, negative bowls are a major problem without any other
-map constraints.The second row of Fig.~\ref{fig:m17_jk} shows the
+edges. For reference, the largest scale that is completely inscribed
+by the array footprint is about 400\,arcsec, and the diagonal of the
+array is about 600\,arcsec (meaning that scales beyond 600\,arcsec are
+completely unconstrained when common-mode rejection is used).  The
+synthetic data are clearly seen as the circular region south of
+M17. As the filter scale is increased, the size of the ripples
+increases accordingly. While larger structures do seem to appear,
+negative bowls are a major problem without any other map
+constraints.The second row of Fig.~\ref{fig:m17_jk} shows the
 jackknife maps. The astronomical emission is almost perfectly removed,
@@
 is probably due to some combination of detector gain and pointing
-drifts. Otherwise the jackknife appears to be a plausible realization
+drifts. Otherwise the jackknife appears to be a plausible realisation
 of noise from the same parent distribution as for the averaged maps.
@@
 the maps in Fig.~\ref{fig:m17_jk}). In fact, using the 900\,arcsec
-filter edge, the map response is still above 80 per centre right out to
+filter edge, the map response is still above 80 per cent right out to
 the largest scale accurately measured in the PSDs, 533\,arcsec, which
@@
 reduction is when using a 150\,arcsec filter. In this case the noise
-is considerably larger in the bright extended reduction as evidenced
+is considerably larger in the bright extended reduction, as evidenced
 by the ``kink'' near 150\,arcsec. Referring to the mask contours in
@@
 the map-maker response is non-linear when using \snr\ masking. Harsh
-filtering can provide misleading results as in this example. Maps of
+filtering can provide misleading results, as in this example. Maps of
 faint extended emission will also suffer considerably, as the
@@
 encouraging for the development of future large-format bolometer
-arrays.
-
-%We have presented
-%sufficient detail in order to be useful not only to users of \scuba,
-%but for similar detector arrays in the future. The basic elements are:
-
-The most significant correlated noise sources in \scuba\ data occur at
-low-frequencies ($\lsim$2\,Hz), and are probably a mixture of
-atmospheric signals and magnetic field pickup. Most of this signal can
-be removed using common-mode rejection, although a principal component
-analysis (PCA) identifies a number of other less-significant
-correlated noise patterns that also dominate at low
+arrays, for which our specific approach should be useful when even
+larger volumes of data are involved.
+
+A major obstacle to making maps of \scuba\ data is low-frequency
+correlated noise (probably a mixture of atmospheric signals and
+magnetic field pickup), which occurs at predominantly $\lsim$2\,Hz.
+Much of this signal can be removed using common-mode rejection,
+although principal component analysis (PCA) identifies a number of
+other less-significant correlated noise patterns at low
 frequencies. Unfortunately these patterns are difficult to model, and
@@
 the PCA directly to remove them, but this technique is prohibitively
-slow for existing high-end desktop computers given the volume of
+slow for existing high-end desktop computers, given the volume of
 \scuba\ data. However, we have demonstrated that high-pass filtering
@@
 
-The main problem to overcome is divergence on large angular scales,
-due to the degeneracy between the map, and the low-frequency signal
-components that are removed (model degeneracies are a generic problem
-of iterative map-makers). A simple strategy of constraining empty
-regions of the map to zero (using either a user-supplied mask for
-known sources, or an iterative determination of signal above some
-\snr\ threshold) provides good constraints for compact objects, and
-bright/extended structures, allowing us to effectively recover angular
-scales up to the order of the array footprint (approximately
-5\,arcmin).
+The main problem with the iterative solution is divergence on large
+angular scales, due to the degeneracy between the map, and the
+low-frequency signal components that are removed. A simple strategy of
+constraining empty regions of the map to zero (using either a
+user-supplied mask for known sources, or an iterative determination of
+signal above some \snr\ threshold) provides good constraints for
+compact objects, and bright/extended structures, allowing us to
+effectively recover angular scales up to the order of the array
+footprint (approximately 5\,arcmin).
 
@@
 single, high-end desktop computers. SMURF can therefore provide real-time
-feedback at the telescope to observers, and it also produces near
+feedback at the telescope to observers, and it also produces nearly
 science-grade data products for the JCMT Science Archive hosted by the
@@
 and/or map constraint to assist in these situations, as more
-experience with the instrument is gained. However, the best long-term
-prospects are probably to turn to a maximum-likelihood algorithm, such
-as SANEPIC \citep{patanchon2008}.
+experience with the instrument is gained. However, provided sufficient
+computing power is available, the best solution in the long-term will
+likely be a maximum-likelihood algorithm, such as SANEPIC
+\citep{patanchon2008}. Even in this case, the existing iterative
+solution from SMURF will probably be used as an initial step, since it
+can quickly clean the bolometer time-series, as well as perform
+map-based despiking (a necessarily iterative procedure).
 
@@
 funding. This research was supported in part by the Natural Sciences
-and Engineering Research Council of Canada. EC also thanks Matthew
-Hasselfield and Mark Halpern for helpful discussions about multiplexed
-TES arrays and their noise characteristics.  The authors thank the
-members of the SCUBA-2 commissioning team for testing the map-maker
-and reporting anomalies; in particular Antonio Chrysostomou and
-Jessica Dempsey. We also thank Dennis Kelly, Alex van Engelen and
-Jennifer Balfour for work on the SCUBA-2 observation simulator; and
-Mark Thompson, Craig Walther and S\'{e}verin Gaudet for being on the
-Critical Design Review panel. Finally, we thank Per Friberg and Gary
-Davis for their helpful comments on the manuscript.
+and Engineering Research Council of Canada. We thank Mark Halpern,
+Matthew Hasselfield, and Gaelen Marsden for many useful discussions.
+The authors thank the members of the SCUBA-2 commissioning team for
+testing the map-maker and reporting anomalies; in particular Antonio
+Chrysostomou, Jessica Dempsey, and Wayne Holland. We acknowledge the
+contributions of Dennis Kelly, Alex van Engelen and Jennifer Balfour
+for early investigations related to SMURF; and Mark Thompson, Craig
+Walther and S\'{e}verin Gaudet for being on the Critical Design Review
+panel. Finally, we thank Per Friberg and Gary Davis for their helpful
+comments on the manuscript.
 

### 2012-09-12 smurf_paper.tex
add ref for LABOCA map-maker, more acknowledgements

@@
 SCUBA scan-map data as the astronomical image estimate improved
-\citep{1999ASPC..172..171J}. The two closest modern relatives of SMURF
-are the Comprehensive Reduction Utility for SHARC-2
-\citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
-Bolocam Galactic Plane Survey \citep[][which uses iterative PCA
-cleaning to avoid the suppression of large-scale
+\citep{1999ASPC..172..171J}. The closest modern relatives of SMURF are
+the Comprehensive Reduction Utility \citep[CRUSH,][]{kovacs2008} for
+the Submillimeter High-Angular Resolution Camera 2 (SHARC-2), the
+Bolometer array data Analysis software (BoA) for the Large APEX
+BOlometer CAmera \citep[LABOCA,][]{siringo2009}, and the pipeline
+developed for the Bolocam Galactic Plane Survey \citep[][which uses
+iterative PCA cleaning to avoid the suppression of large-scale
 structure]{aguirre2011}.
@@
 residuals in their data (similar to the \scuba\ data described here,
-although PCA cleaning is prohibitively slow in this case).
+although PCA cleaning is prohibitively slow in our case).
 
@@
 correlated noise components from the data, successively improving the
-map \citep[e.g.,][]{kovacs2008,aguirre2011}, we have shown that a
-relatively simpler approach provides good results for \scuba\ data,
-with reasonable computational requirements. This conclusion is
+map \citep[e.g.,][]{kovacs2008,siringo2009,aguirre2011}, we have shown
+that a relatively simpler approach provides good results for \scuba\
+data, with reasonable computational requirements. This conclusion is
 encouraging for the development of future large-format bolometer
@@
 user-supplied mask for known sources, or an iterative determination of
-signal above some \snr\ threshold) provides good constraints for
+signal above some \snr\ threshold) provides good constraints for both
 compact objects, and bright/extended structures, allowing us to
@@
 funds for the contruction of \scuba\ were provided by the Canada
-Foundation for Innovation. EC thanks CANARIE/CANFAR for additional
-funding. This research was supported in part by the Natural Sciences
-and Engineering Research Council of Canada. We thank Mark Halpern,
-Matthew Hasselfield, and Gaelen Marsden for many useful discussions.
-The authors thank the members of the SCUBA-2 commissioning team for
-testing the map-maker and reporting anomalies; in particular Antonio
-Chrysostomou, Jessica Dempsey, and Wayne Holland. We acknowledge the
-contributions of Dennis Kelly, Alex van Engelen and Jennifer Balfour
-for early investigations related to SMURF; and Mark Thompson, Craig
-Walther and S\'{e}verin Gaudet for being on the Critical Design Review
-panel. Finally, we thank Per Friberg and Gary Davis for their helpful
-comments on the manuscript.
+Foundation for Innovation.  This research was supported in part by the
+Natural Sciences and Engineering Research Council of Canada.  EC
+thanks CANARIE/CANFAR for additional funding.  The authors thank the
+members of the SCUBA-2 commissioning team for testing the map-maker
+and reporting anomalies; in particular Antonio Chrysostomou, Jessica
+Dempsey, and Wayne Holland.  We also Mark Halpern, Matthew
+Hasselfield, and Gaelen Marsden for many useful discussions, and
+observers who provided useful feedback; especially David Nutter and
+Todd McKenzie.  We acknowledge the contributions of Dennis Kelly, Alex
+van Engelen and Jennifer Balfour for early investigations related to
+SMURF; and Mark Thompson, Craig Walther and S\'{e}verin Gaudet for
+being on the Critical Design Review panel. Finally, we thank Per
+Friberg and Gary Davis for their helpful comments on the manuscript.
 

### 2012-09-12 smurf_paper.tex
mention jackknives in conclusions

@@
 
-The main problem with the iterative solution is divergence on large
-angular scales, due to the degeneracy between the map, and the
+We have found that the iterative solution tends to diverge on large
+angular scales due to the degeneracy between the map, and the
 low-frequency signal components that are removed. A simple strategy of
@@
 signal above some \snr\ threshold) provides good constraints for both
-compact objects, and bright/extended structures, allowing us to
+compact objects, and bright/extended structures. Particularly in the
+latter case, using a combination of synthetic sources and an empirical
+measurement of the map noise from jackknife tests (differences of
+independent portions of the data), we have demonstrated that we can
 effectively recover angular scales up to the order of the array
@@
 white-noise limited. Residual large-scale noise can be removed with a
-whitening filter, and sources detected using a matched-filter
-(smoothing with the effective filtered point spread function).
+whitening filter (also established from jackknife estimates of the
+noise), and sources detected using a matched-filter (smoothing with
+the effective filtered point spread function).
 
@@
 computing power is available, the best solution in the long-term will
-likely be a maximum-likelihood algorithm, such as SANEPIC
+be a maximum-likelihood algorithm, such as SANEPIC
 \citep{patanchon2008}. Even in this case, the existing iterative

### 2012-09-12 smurf_paper.tex
fiddle in acknowldegements

@@
 and reporting anomalies; in particular Antonio Chrysostomou, Jessica
-Dempsey, and Wayne Holland.  We also Mark Halpern, Matthew
+Dempsey, and Wayne Holland.  We also thank Mark Halpern, Matthew
 Hasselfield, and Gaelen Marsden for many useful discussions, and
-observers who provided useful feedback; especially David Nutter and
+observers who provided helpful feedback; especially David Nutter and
 Todd McKenzie.  We acknowledge the contributions of Dennis Kelly, Alex

### 2012-09-12 smurf_paper.tex
Minor typos/corrections

@@
 full-width at half-maximum (FWHM) -- a typical rule-of-thumb for
-adequately sampling a Gaussian point spread function (PSF). There is
+adequately sampling a Gaussian point spread function (PSF) -- at the maximum scanning speed. There is
 an additional 41st row of SQUID readouts that are not connected to
@@
 Starlink\footnote{\url{http://www.starlink.ac.uk}} is
-open-source\footnote{\url{https://github.com/Starlink}}, and already
+open-source\footnote{\url{https://github.com/Starlink}} (distributed under the GNU General Public Licence v3), and already
 used extensively at the Joint Astronomy Centre (host of the JCMT) for
@@
 low-pass filter). In practice the noise performance was
-indistinguishable, and required slightly more execution time using
-this method. It is therefore not currently available to users.
+indistinguishable, and required slightly longer execution time than using
+the method above. Therefore this method is not used.
 
@@
 (absolute calibration), and hence noise performance of the instrument
-in astronomical units.  In this example we produce a 450\,\micron\ map
+in astronomically-useful units.  In this example we produce a 450\,\micron\ map
 of Uranus (observation 26 on 20111017), which is a nearly point-like
@@
 and for comparison, the jackknife map has also been whitened in
-Fig.~\ref{fig:lockman_maps}e. In both cases, the maps are visually
+Fig.~\ref{fig:lockman_maps}e. In both cases, the maps are visibly
 flatter than the non-whitened cases.
@@
 Have real astronomical sources been detected using the matched filter?
-For both the smoothed signal and jackknife maps, orange circles denote
+For both the smoothed signal and jackknife maps, blue circles denote
 3.8-$\sigma$ peaks. While not justified here, this threshold is fairly
@@
 as any point in the curve where $P_\mathrm{M}(k)$ exceeds
-$P_\mathrm{S}(k)$ (i.e., $G(K)$ is assumed to be $\le$1). As expected,
+$P_\mathrm{S}(k)$ (i.e., $G(k)$ is assumed to be $\le$1). As expected,
 the larger the scale of the filter, the lower the spatial frequency at

### 2012-09-12 smurf_paper.tex
After much discussion Wayne Holland is now an author

@@
     Frossie~Economou$^2$\thanks{Present address: National Optical
-      Astronomy Observatory, 950 N.\ Cherry Avenue, Tucson, AZ 85719, USA},
+      Astronomy Observatory, 950 N.\ Cherry Avenue, Tucson, AZ 85719,
+      USA},
+    Wayne ~S.~Holland$^{4,5}$
   }
@@
   $^{3}$Netherlands Organisation for Scientific Research,
-  Laan van Nieuw Oost-Indie 300, NL-2509 AC The Hague, The Netherlands}
+  Laan van Nieuw Oost-Indie 300, NL-2509 AC The Hague, The Netherlands\\
+  $^{4}$UK Astronomy Technology Centre, Royal Observatory, Blackford
+  Hill, Edinburgh EH9 3HJ\\
+  $^{5}$Institute for Astronomy, University of Edinburgh, Royal Observatory, Blackford Hill Edinburgh, EH9 3HJ}
 \begin{document}
@@
 members of the SCUBA-2 commissioning team for testing the map-maker
-and reporting anomalies; in particular Antonio Chrysostomou, Jessica
-Dempsey, and Wayne Holland.  We also thank Mark Halpern, Matthew
+and reporting anomalies; in particular Antonio Chrysostomou and Jessica
+Dempsey.  We also thank Mark Halpern, Matthew
 Hasselfield, and Gaelen Marsden for many useful discussions, and

### 2012-09-14 smurf_paper.tex
add support for latex+eps images instead of pdflatex+pdf images
First run convert_images to change all of the .pdf files into .eps.

Then uncomment the relevant block at the start of smurf_paper.tex. This
macro '\imagefile{image}' just tacks a .eps or .pdf on the end.
graphicx already knows how to deal with .eps vs .pdf suffix in image files
initial submission to MN
I put everything, including the information to fill out the on-line forms,
in the submission_mnras subdirectory. Currently smurf_paper.tex is set to
use postscript instead of PDF to follow their submission guidelines.

@@
 
+% --- Uncomment following block for use with pdflatex and pdf images ----------
 %% Recommended by MN for handling hyperlinks
-\usepackage[pdftex,pdfpagemode={UseOutlines},bookmarks,bookmarksopen,colorlinks,linkcolor={blue},citecolor={green},urlcolor={red}]{hyperref}
-
-% this is needed to get pdftex to work properly!
-\usepackage[pdftex]{graphicx}
+
+%\usepackage[pdftex,pdfpagemode={UseOutlines},bookmarks,bookmarksopen,colorlinks,linkcolor={blue},citecolor={green},urlcolor={red}]{hyperref}
+
+%\usepackage[pdftex]{graphicx}
+
+% -----------------------------------------------------------------------------
+
+% --- Uncomment following block for latex and postscript images ---------------
+%% Recommended by MN for handling hyperlinks - PS version
+
+\usepackage[bookmarks,bookmarksopen,colorlinks,linkcolor={blue},citecolor={green},urlcolor={red}]{hyperref}
+
+% -----------------------------------------------------------------------------
 
@@
 PONG'' with even better cross-linking. For smaller areas (of order the
-array footprint, or pont sources), in which the PONG turn-around
+array footprint, or point sources), in which the PONG turn-around
 overheads would be large, a ``constant velocity daisy'' (CV Daisy) is
@@
 \begin{figure*}
-\includegraphics[width=0.49\linewidth]{bolos_point_mix_s2sro.pdf}
-\includegraphics[width=0.49\linewidth]{bolos_point_mix.pdf}
+\includegraphics[width=0.49\linewidth]{bolos_point_mix_s2sro}
+\includegraphics[width=0.49\linewidth]{bolos_point_mix}
 \caption{A comparison between single bolometer time-series in each of
@@
 \centering
-\includegraphics[width=0.49\linewidth]{pspec_s2sro.pdf}
-\includegraphics[width=0.49\linewidth]{pspec.pdf}
+\includegraphics[width=0.49\linewidth]{pspec_s2sro}
+\includegraphics[width=0.49\linewidth]{pspec}
 \caption{Bolometer power spectral densities for the same two data sets
@@
 data are less dominated by low-frequency noise to begin with, the
-signals generally require less agressive high-pass filtering to
+signals generally require less aggressive high-pass filtering to
 produce maps, and therefore retain larger-scale structures than with
@@
 \centering
-\includegraphics[width=\linewidth]{magpickup.pdf}
+\includegraphics[width=\linewidth]{magpickup}
 \caption{Evidence for significant magnetic field pickup for
@@
 \centering
-\includegraphics[width=\linewidth]{pca.pdf}
+\includegraphics[width=\linewidth]{pca}
 \caption{The first twelve components from a principal component
@@
 \centering
-\includegraphics[width=\linewidth]{dimm.pdf}
+\includegraphics[width=\linewidth]{dimm}
 \caption{Typical map-making algorithm. Raw data (stored in multiple
@@
 \centering
-\includegraphics[width=\linewidth]{steps1.pdf}
+\includegraphics[width=\linewidth]{steps1}
 \caption{Examples of steps in time-series data. Steps occur with a wide
@@
 \centering
-\includegraphics[width=\linewidth]{pad2.pdf}
+\includegraphics[width=\linewidth]{pad2}
 \caption{The black curve shows a bolometer time-series, padded
@@
 \centering
-\includegraphics[width=\linewidth]{pad1.pdf}
+\includegraphics[width=\linewidth]{pad1}
 \caption{The black curve shows the same bolometer time-series as in
@@
 \centering
-\includegraphics[width=\linewidth]{com.pdf}
+\includegraphics[width=\linewidth]{com}
 \caption{A selection of typical bolometer time-series after
@@
 \centering
-\includegraphics[width=\linewidth]{pointmaps.pdf}
+\includegraphics[width=\linewidth]{pointmaps}
 \caption{Multiple reductions of a 450\,\micron\ CV daisy scan of
@@
 \centering
-\includegraphics[width=\linewidth]{degeneracy.pdf}
+\includegraphics[width=\linewidth]{degeneracy}
 \caption{Demonstration of the degeneracy between large-scale
@@
 \centering
-\includegraphics[width=\linewidth]{lockman_maps.pdf}
+\includegraphics[width=\linewidth]{lockman_maps}
 \caption{Reduction of a blank-field survey: the Lockman Hole. (a) Raw
@@
 \centering
-\includegraphics[width=\linewidth]{lockman_psf.pdf}
+\includegraphics[width=\linewidth]{lockman_psf}
 \caption{Slice through the angular response to an ideal Gaussian point
@@
 \centering
-\includegraphics[width=\linewidth]{lockman_pspec.pdf}
+\includegraphics[width=\linewidth]{lockman_pspec}
 \caption{Radial (azimuthally-averaged) angular power spectral
@@
 \centering
-\includegraphics[width=\linewidth]{lockman_hist.pdf}
+\includegraphics[width=\linewidth]{lockman_hist}
 \caption{Histograms of \snr\ using pixels from the central region of
@@
 \centering
-\includegraphics[width=\linewidth]{m17.pdf}
+\includegraphics[width=\linewidth]{m17}
 \caption{An 850\,\micron\ rotating PONG map of M17. Intensity is
@@
 \centering
-\includegraphics[width=\linewidth]{m17_jk.pdf}
+\includegraphics[width=\linewidth]{m17_jk}
 \caption{Maps of M17 used to characterise the noise properties and
@@
 \centering
-\includegraphics[width=0.49\linewidth]{pspec_m17_default.pdf}
-\includegraphics[width=0.49\linewidth]{cor_pspec_m17_default.pdf}
+\includegraphics[width=0.49\linewidth]{pspec_m17_default}
+\includegraphics[width=0.49\linewidth]{cor_pspec_m17_default}
 \caption{Angular PSDs for the region of the M17 map in
@@
 \centering
-\includegraphics[width=0.49\linewidth]{pspec_m17_bright_extended.pdf}
-\includegraphics[width=0.49\linewidth]{cor_pspec_m17_bright_extended.pdf}
+\includegraphics[width=0.49\linewidth]{pspec_m17_bright_extended}
+\includegraphics[width=0.49\linewidth]{cor_pspec_m17_bright_extended}
 \caption{Lines have same meaning as in Fig.~\ref{fig:m17_def_ps},
@@
 
-%SMURF is highly configurable, making it easy to vary the order
-%and types of models that are fit to the data. It is also easy to add
-%new models to the existing infrastructure.
-
 One regime in which SMURF does not presently perform well is in maps

### 2012-11-07 smurf_paper.tex
thank Mandana for help understanding the MCE xfer function

@@
 
-%\usepackage[pdftex,pdfpagemode={UseOutlines},bookmarks,bookmarksopen,colorlinks,linkcolor={blue},citecolor={green},urlcolor={red}]{hyperref}
-
-%\usepackage[pdftex]{graphicx}
+\usepackage[pdftex,pdfpagemode={UseOutlines},bookmarks,bookmarksopen,colorlinks,linkcolor={blue},citecolor={green},urlcolor={red}]{hyperref}
+
+\usepackage[pdftex]{graphicx}
 
@@
 
-\usepackage[bookmarks,bookmarksopen,colorlinks,linkcolor={blue},citecolor={green},urlcolor={red}]{hyperref}
+%\usepackage[bookmarks,bookmarksopen,colorlinks,linkcolor={blue},citecolor={green},urlcolor={red}]{hyperref}
 
@@
 members of the SCUBA-2 commissioning team for testing the map-maker
-and reporting anomalies; in particular Antonio Chrysostomou and Jessica
-Dempsey.  We also thank Mark Halpern, Matthew
-Hasselfield, and Gaelen Marsden for many useful discussions, and
-observers who provided helpful feedback; especially David Nutter and
-Todd McKenzie.  We acknowledge the contributions of Dennis Kelly, Alex
+and reporting anomalies; in particular Antonio Chrysostomou and
+Jessica Dempsey.  We also thank Mark Halpern, Matthew Hasselfield, and
+Gaelen Marsden for many useful discussions; observers who provided
+helpful feedback, especially David Nutter and Todd McKenzie; and
+Mandana Amiri for consultations regarding the SCUBA-2 Multi-Channel
+Electronics.  We acknowledge the contributions of Dennis Kelly, Alex
 van Engelen and Jennifer Balfour for early investigations related to

### 2012-11-09 smurf_paper.tex
addressing referee comments in intro
- added a bunch of references to SPT/ACT stuff
- fixed-up bolos_point_mix, pspec and pca plots now that
  flatfielding is turned on as it should
- some stuff about performance of algorithm

@@
   instrument operating on the 15-m James Clerk Maxwell Telescope,
-  consisting of about 5000 bolometers in each of two simultaneous
-  imaging bands centred over 450 and 850\,\micron. The camera is
-  operated by scanning across the sky and records data at a rate of
-  200\,Hz. This rate is much greater than that of previously existing
-  submillimetre cameras, and represents a significant analysis
-  challenge. We describe the production of \scuba\ maps using the
+  consisting nominally of 5120 bolometers (with a useful yield of
+  69\%) in each of two simultaneous imaging bands centred over 450 and
+  850\,\micron. The camera is operated by scanning across the sky and
+  recording data at a rate of 200\,Hz. As the largest of a new
+  generation of multiplexed kilopixel bolometer cameras operating in
+  the (sub)millimetre, \scuba\ data analysis represents a significant
+  challenge.  We describe the production of maps using the
   Sub-Millimetre User Reduction Facility (SMURF) in which we have
@@
   used in an automated setting, both at the telescope for real-time
-  feedback to observers, as well as for the production of generic
-  science products for the JCMT Science Archive at the Canadian
-  Astronomy Data Centre. Three detailed case studies are used to: (i)
-  explore convergence properties of the map-maker using simple prior
+  feedback to observers, as well as for the production of science
+  products for the JCMT Science Archive at the Canadian Astronomy Data
+  Centre. Three detailed case studies are used to: (i) explore
+  convergence properties of the map-maker using simple prior
   constraints (Uranus -- a point source); (ii) achieve the white-noise
@@
 around nearby stars, helping us understand the early stages of planet
-formation.  With $\sim10,000$ total detectors, \scuba\ is presently
-the largest submillimetre camera in the world, and the fastest
-wide-area ground-based submillimetre imager.
-
-Submillimetre imaging cameras generally use bolometers, rather than
-coherent detectors, to maximise sensitivity. The sensitivity of
-bolometers is limited by white photon and phonon noise from the
-instrument and ambient backgrounds. The low-frequency noise, however,
-is typically dominated by sources which produce slow variations in the
-background (e.g., thermal variations within the cryostat, and the
-atmosphere for ground-based cameras), and drifts in the readout
-electronics. Such noise has a power spectrum $\propto 1/f^\alpha$
-($\alpha>0$), and the frequency at which it is comparable to the white
-noise level is called the ``$1/f$ knee''. Since the low-frequency
-noise is largely correlated between all, or subsets, of the bolometers
-in time, it can be suppressed during map-making, since astronomical
-signals have the distinct property that they are fixed in a sky
-reference frame (assuming they are not time-varying). If successful,
-the noise in the resulting map is said to be ``white noise limited'',
-meaning that it is uncorrelated spatially, and has an amplitude that
-scales as the $\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the
-noise-equivalent flux-density (the white noise level of a bolometer in
-1\,s of integration), and $t$ is the amount of integration time in a
-map pixel.
+formation.  With $10,240$ nominal detectors \citep[of which $\sim$69\%
+work and are typically useful,][]{holland2012}, \scuba\ is presently
+the largest of a new generation of multiplexed kilopixel
+(sub)millimetre bolometer cameras, which also includes the cameras for
+the South Pole Telescope \citep[SPT,][]{carlstrom2011} and the Atacama
+Cosmology Telescope \citep[ACT,][]{swetz2011}, both dedicated
+experiments for studying anisotropies in the Cosmic Microwave
+Backround (CMB) using similar technology.
+
+Submillimetre imaging cameras generally maximise sensitivity using
+bolometers, rather than coherent detectors, which are limited by white
+photon and phonon noise from the instrument and ambient
+backgrounds. The low-frequency noise, however, is typically dominated
+by sources which produce slow variations in the background (e.g.,
+thermal variations within the cryostat, and the atmosphere for
+ground-based cameras), and drifts in the readout electronics. Such
+noise has a power spectrum $\propto 1/f^\alpha$ ($\alpha>0$), and the
+frequency at which it is comparable to the white noise level is called
+the ``$1/f$ knee''. Since the low-frequency noise is largely
+correlated between all, or subsets, of the bolometers in time, it can
+be suppressed during map-making, since astronomical signals have the
+distinct property that they are fixed in a sky reference frame
+(assuming they are not time-varying). If successful, the noise in the
+resulting map is said to be ``white noise limited'', meaning that it
+is uncorrelated spatially, and has an amplitude that scales as the
+$\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the noise-equivalent
+flux-density (the white noise level of a bolometer in 1\,s of
+integration), and $t$ is the amount of integration time in a map
+pixel.
 
@@
 spectrum, i.e., above the $1/f$ knee; and (ii) to provide good
-``cross-linking'', in which each portion of the map is sampled on a
-range of temporal scales, again, to help distinguish time-varying
+``cross-linking'', in which each portion of the map is scanned at a
+range of position angles, again, to help distinguish time-varying
 noise features from fixed astronomical sources. In the case of \scuba,
@@
 offering scan patterns that cross the sky at a wide range of position
-angles, ultimately observing each portion of the map at a number of
-different times. Such methods are now used by virtually all existing
-ground-based submillimetre cameras
+angles. Such methods are now used by virtually all existing
+ground-based submillimetre camerass
 \citep[e.g.,][]{glenn1998,weferling2002,wilson2008,kovacs2008b}, in
@@
 instruments that had poorer low-frequency noise performance, and were
-only sensitive to modest angular scales
+only sensitive to modest angular scales. We note that under some
+circumstances cross-linking may not be essential. For example, SPT,
+which was designed to measure CMB anisotropies, scans entirely in
+azimuth \citep{schaffer2011}. While this strategy limits its ability
+to recover large angular scales transverse to the scan direction, the
+anisotropic filtering in its maps can be accounted for during analysis
+to achieve SPTs focussed objectives \citep[in contrast, the similar
+ACT experiment uses cross-linking to improve its reponse to
+large-angular scales,][]{das2011}. This approach is less practical as
+a solution for \scuba, however, which must serve a broader range of
+scientific interests.
 
@@
 \citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO
-\citep[e.g.,][]{omont2001}, another camera from the same
-generation. Generally speaking, such methods are fast, although
-depending on the science goals and noise properties of the data, they
-may not achieve the best noise performance on the angular scales of
-interest. In more recent years a method that has been popular for
+\citep[e.g.,][]{omont2001}, another camera from the same generation. A
+more recent example is the analysis of SPT data
+\citep{schaffer2011}. Generally speaking, such methods are fast,
+although depending on the science goals and noise properties of the
+data, they may not achieve the best noise performance on the angular
+scales of interest. A related method that has been popular for
 reducing data in fields of faint point sources with Bolocam
@@
 Thermal Emission Camera \citep[AzTEC, e.g.,][]{scott2008}, is
-``principal component analysis'' (PCA) cleaning. A statistical
+principal component analysis (PCA) cleaning. A statistical
 ``black-box'' removes the most correlated components of the bolometer
-signals, enabling the detection of point-sources very close to the
+signals, enabling the detection of point-sources close to the
 theoretical white-noise limits of the detectors, with reasonable
@@
 (hundreds rather than thousands of detectors). However, PCA cleaning
-is not a good generic solution for producing maps of extended
-structures, since such sources produce correlated signals amongst many
-detectors, and are removed by this procedure. Furthermore, in the case
-of \scuba\, performing PCA on even a single subarray (1280 detectors,
-nominally) can be prohibitively slow.
+is not a good solution for producing maps of extended structures,
+since such sources produce correlated signals amongst many detectors,
+and are removed by this procedure \citep[an exception is the iterative
+PCA approach of][]{aguirre2011}. Furthermore, in the case of \scuba,
+performing PCA on even a single subarray (typically $\sim$900
+detectors) can be prohibitively slow.
 
@@
 fact that it used a differential radiometer instead of
-bolometers). Other descriptions in the experimental cosmic microwave
-background (CMB) literature include \citet{tegmark1997} and
-\citet{stompor2002}, and an application to data from SCUBA is
-described in \citet{borys2004}.  The downside to such methods is that
-they can be both computationally expensive and have large memory
-requirements. While for some experimental designs fast iterative
-methods for the inversion do exist without requiring excessive amounts
-of memory \citep[e.g.,][]{wright1996}, for the more general map-making
-problem, involving many detectors, things are significantly
-complicated both by the need to measure the cross power spectra for
-all unique pairs of bolometers, as well as performing the inversion
-itself. The most promising maximum-likelihood method that may one day
-be applied to \scuba\ data is ``SANEPIC'', which successfully reduced
-maps from the Balloon-borne Large-Aperture Submillimeter Telescope,
-involving hundreds of bolometers, while correctly incorporating
-inter-bolometer noise correlations \citep[][]{patanchon2008}.
+bolometers). Other descriptions in the experimental CMB literature
+include \citet{tegmark1997} and \citet{stompor2002}, and an
+application to data from SCUBA is described in \citet{borys2004}.  The
+downside to such methods is that they can be both computationally
+expensive and have large memory requirements. While for some
+experimental designs fast iterative methods for the inversion do exist
+without requiring excessive amounts of memory such as the method of
+\citet{wright1996} \citep[which was later implemented for SCUBA
+by][]{johnstone2000}, for the more general map-making problem,
+involving many detectors, matters are significantly complicated both
+by the need to measure the cross spectra for all unique pairs of
+bolometers, as well as performing the inversion itself. A
+maximum-likelihood method was successfully used to make maps from ACT
+data using approximately the same number of detectors as a single
+\scuba\ subarray at a single wavelength. However, the calculation is
+extremely resource intensive and requires a large cluster in order to
+run; a single map can take 10\,CPU\,years \citep{fowler2010}. Perhaps
+the most promising maximum-likelihood method that may one day be
+applied to \scuba\ data is ``SANEPIC'', which produced maps from
+Balloon-borne Large-Aperture Submillimeter Telescope data, involving
+hundreds of bolometers, while correctly incorporating inter-bolometer
+noise correlations \citep[][]{patanchon2008}.
 
@@
 noise at any instant in time is uncorrelated with others, and with the
-data from other bolometers), greatly simplifying the measurement of
-noise properties and the inversion step that is so complicated in the
-maximum-likelihood methods. In principle, it should also be possible
-to recover a wider dynamic range in angular scales than possible for
-the simplest direct methods, with an increase in calculation time that
-is merely linear in the number of iterations needed, while using a
-comparable amount of memory. The iterative method of
-\citet{wright1996} specific to chopped (CMB) data was applied to SCUBA
-observations by \citet{johnstone2000}, although in this case the
-chopping removes most of the low-frequency noise, and greatly
-simplifies the problem (at the expensive of large-scale structure in
-the data). More similar to the algorithm described here was the
-iterative pipeline recipe for fitting and removing baseline drifts in
-SCUBA scan-map data as the astronomical image estimate improved
-\citep{1999ASPC..172..171J}. The closest modern relatives of SMURF are
+data from other bolometers), greatly simplifying the inversion step
+that is so complicated in the maximum-likelihood methods. In the
+iterative approach memory usage usually scales linearly with $N$, the
+product of the number of bolometers, $N_\mathrm{b}$, with the number
+of samples in time, $N_\mathrm{t}$. The computation time is the
+product of the number of iterations, $N_\mathrm{iter}$, with the time
+per iteration, $t_\mathrm{iter}$. For the algorithm described in this
+paper, $N_\mathrm{iter}$ is typically in the range $\sim$5--100
+(depending on the \snr\ of large-scale structures, and the size of the
+map), and $t_\mathrm{iter}$ scales as $N_\mathrm{t} \log
+N_\mathrm{t}$, since Fast Fourier Transforms (FFTs) of the time-series
+are typically performed. By comparison, the fastest maximum-likelihood
+methods that account for inter-bolometer correlations have an
+execution time with an $N_\mathrm{b}^2$ dependence, although it is
+possible to limit memory usage to be linear in $N_\mathrm{t}$
+\citep[e.g.,][]{patanchon2008}.
+
+Iterative approaches to map-making have a long history in the submillimetre. For example, the iterative pipeline recipe for fitting and
+removing baseline drifts in SCUBA scan-map data as the astronomical
+image estimate improved \citep{1999ASPC..172..171J} is similar to the
+algorithm described in this paper. The closest modern relatives are
 the Comprehensive Reduction Utility \citep[CRUSH,][]{kovacs2008} for
@@
 BOlometer CAmera \citep[LABOCA,][]{siringo2009}, and the pipeline
-developed for the Bolocam Galactic Plane Survey \citep[][which uses
-iterative PCA cleaning to avoid the suppression of large-scale
-structure]{aguirre2011}.
+developed for the Bolocam Galactic Plane Survey \citep{aguirre2011}.
 
@@
 are removed with an iterative high-pass filter. This strategy enables
-SMURF to reduce data faster than they are taken, on single high-end
-desktop computers, and it has successfully been used as part of a
-real-time pipeline offering feedback to observers at the
-telescope. The pipeline is also used to generate near-science grade
-products for the JCMT Science Archive \citep{2011ASPC..442..203E}
-hosted by the Canadian Astronomy Data Centre (CADC). Details of the
-pipeline design are given in \citet{gibb2005} and
-\citet{2008AN....329..295C}.
+SMURF to reduce data on time scales comensurate with the observing
+times (or significantly faster), on single high-end desktop
+computers. For reference, all of the data analysis in this paper was
+performed on a machine with 64-bit 8-core central processing unit
+operating at 2.7\,GHz, and 48\,Gb of memory. SMURF has been
+successfully used as part of a real-time pipeline offering feedback to
+observers at the telescope. The pipeline is also used to generate
+near-science grade products for the JCMT Science Archive
+\citep{2011ASPC..442..203E} hosted by the Canadian Astronomy Data
+Centre (CADC). Details of the pipeline design are given in
+\citet{gibb2005} and \citet{2008AN....329..295C}.
 
@@
 full-width at half-maximum (FWHM) -- a typical rule-of-thumb for
-adequately sampling a Gaussian point spread function (PSF) -- at the maximum scanning speed. There is
-an additional 41st row of SQUID readouts that are not connected to
+adequately sampling a Gaussian point spread function (PSF) -- at the
+maximum scanning speed of 600\,arcsec\,s$^{-1}$. There is an
+additional 41st row of SQUID readouts that are not connected to
 TESs. These ``dark SQUIDs'' track non-thermal noise sources that are
@@
 \centering
-\includegraphics[width=\linewidth]{pca}
-\caption{The first twelve components from a principal component
-  analysis, ranked by the mean eigenvalues, of combined 450 (s4a
+\includegraphics[width=0.85\linewidth]{pca}
+\caption{The first nine components from a principal component
+  analysis, ranked by the mean eigenvalues of combined 450 (s4a
   subarray) and 850\,\micron\ (s8b subarray) time-series bolometer
@@
   of Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec}. For each component
-  the top plot shows the time-series of the normalised eigenvectors,
-  the middle plot its PSD, and the bottom coloured panels the
-  eigenvalues for the bolometers at each wavelength (the amplitudes of
-  the eigenvectors in the time-series). For reference, both the mean,
-  $\bar{\lambda}$, and \rms, $\lambda_\mathrm{rms}$, eigenvalues for
-  the bolometers in each subarray are also shown. Most of the
-  correlated $1/f$ signal is encapsulated in Components~1 and 2, and
+  the top plot shows the time-series of the eigenvectors normalised by
+  their \rms, the middle plot its PSD, and the bottom coloured panels
+  the eigenvalues for the bolometers at each wavelength (the
+  amplitudes of the eigenvectors in the time-series). For reference,
+  both the mean, $\bar{\lambda}$, and \rms, $\lambda_\mathrm{rms}$,
+  eigenvalues for the bolometers in each subarray are also shown. Most
+  of the correlated $1/f$ signal is encapsulated in Component~1, and
   appears to be dominated by scan-synchronous noise (compare with the
-  scan pattern in the right-hand panel of Fig.~\ref{fig:bolos_mix}).
-  Components~3, 4, 7 and 11 are transient glitches that only appear in
-  a small subset of the bolometers. Component~5 also appears in only a
-  few bolometers, and is dominated by a narrow line at 57\,Hz which
-  suggests that it is related to the 60\,Hz mains. Components~6 and
-  8--10 contain a mixture of broad lines above $\sim$30\,Hz (which are
-  probably aliased high-frequency noise), with some low-frequency
-  drifts. Finally, the erratic appearance of the eigenvector for
-  Component~12 is typical of a poorly-biased detector.}
+  scan pattern in the right-hand panel of
+  Fig.~\ref{fig:bolos_mix}). Some of this signal also appears in
+  Component~3. Component~2 is another $1/f$ signal that is dominant in
+  the 850\,\micron\ subarray. Most of the remaining components consist
+  of weaker $1/f$ noise and line features predominantly above
+  $\sim$30\,Hz. The eigenvalue maps show correlated patterns along
+  columns suggesting pickup in their common amplifier chains.}
 \label{fig:pca}
@@
 visible. These drifts are removed using a high-pass filter
-(\model{FLT}, Section~\ref{sec:flt}) implemented with Fast Fourier
-Transforms (FFTs). The residual signal is then re-gridded to estimate
-the map. Finally, the map is used to estimate and remove the
-astronomical signal contamination in the bolometer data (\model{AST},
+(\model{FLT}, Section~\ref{sec:flt}) implemented with FFTs. The
+residual signal is then re-gridded to estimate the map. Finally, the
+map is used to estimate and remove the astronomical signal
+contamination in the bolometer data (\model{AST},
 Section~\ref{sec:ast}), leaving a relatively clean data set in which

### 2012-11-13 smurf_paper.tex
Improve caption for Fig.9
addressing mostly smaller comments in 1st half of paper

@@
 \begin{figure*}
+\centering
 \includegraphics[width=0.49\linewidth]{bolos_point_mix_s2sro}
@@
 \caption{A comparison between single bolometer time-series in each of
-  the 450 and 850\,\micron\ bands with the mixing chamber temperature
-  and azimuth/elevation pointing offsets, before and after upgrading
-  the instrument. The grey signals over-plotted in the top two panels
-  show the residual time-series after removing the common-mode
-  signals.  Left: data taken during the S2SRO period, observation 29
-  on 20100313. There is a strong correlation between the bolometers
-  and the roughly $\sim$25\,s oscillation in the fridge, but only a
-  minor correlation with the telescope motion. The nearly flat
-  common-mode subtracted signals show: (i) that most of the
-  low-frequency signal is common to all of the bolometers; and (ii)
-  the non-correlated, and predominantly white noise at 450\,\micron\
-  is significantly larger than at 850\,\micron. Right: data taken with
-  the fully-commissioned instrument, observation 38 on
-  20111112. Unlike the S2SRO data, there is no strong signal produced
-  by variations in the fridge temperature. The 450\,\micron\ data have
-  minimal correlated low-frequency noise (as evidenced by the lack of
-  a difference between the raw and common-mode subtracted signals),
-  although the 850\,\micron\ data show a common signal correlated with
-  the telescope motion (note that this scan has a significantly larger
-  amplitude than the S2SRO data set), mostly likely caused by magnetic
-  field pickup (see Section~\ref{sec:magpickup} and
+  the 450 and 850\,\micron\ bands with the mixing chamber temperature,
+  the line-of-sight 450\,\micron\ opacity derived from the water
+  vapour monitor, and azimuth/elevation pointing offsets, before and
+  after upgrading the instrument. The grey signals over-plotted in the
+  top two panels show the residual time-series after removing the
+  common-mode signals.  Left: data taken during the S2SRO period,
+  observation 29 on 20100313. There is a strong correlation between
+  the bolometers and the roughly $\sim$25\,s oscillation in the
+  fridge, but only a minor correlation with the opacity and telescope
+  motion. The nearly flat common-mode subtracted signals show: (i)
+  that most of the low-frequency signal is common to all of the
+  bolometers; and (ii) the non-correlated, and predominantly white
+  noise at 450\,\micron\ is significantly larger than at
+  850\,\micron. Right: data taken with the fully-commissioned
+  instrument, observation 38 on 20111112. Unlike the S2SRO data, there
+  is no strong signal produced by variations in the fridge
+  temperature. The 450\,\micron\ data have minimal correlated
+  low-frequency noise (as evidenced by the lack of a difference
+  between the raw and common-mode subtracted signals), although the
+  850\,\micron\ data show a common signal correlated with the opacity
+  and telescope motion (note that this scan has a significantly larger
+  amplitude than the S2SRO data set), mostly likely caused by a
+  combination of changing airmass (anti-correlated with opacity), and
+  magnetic field pickup (see Section~\ref{sec:magpickup} and
   Fig.~\ref{fig:magpickup}).}
@@
 that resembles the scan pattern in the right-hand panel of
-Fig.~\ref{fig:pca}. While these correlated signals also clearly appear
-at both wavelengths, referring to the maps of eigenvalues, they are
-considerably stronger at 850\,\micron\ (consistent with the visual
+Fig.~\ref{fig:bolos_mix}. While these correlated signals also clearly
+appear at both wavelengths, referring to the maps of eigenvalues, they
+are considerably stronger at 850\,\micron\ (consistent with the visual
 appearance of the bolometer signals in Fig.~\ref{fig:pca}). Also note
@@
   and 4\,arcsec at 450 and 850\,\micron, respectively, to ensure
-  adequate sampling of the \scuba\ PSFs). A cleaning stage repairs DC
+  adequate sampling of the \scuba\ PSFs). A cleaning stage repairs
   steps and spikes, and fills gaps to ensure continuity. Finally, the
@@
 residual signal is then re-gridded to estimate the map. Finally, the
-map is used to estimate and remove the astronomical signal
-contamination in the bolometer data (\model{AST},
-Section~\ref{sec:ast}), leaving a relatively clean data set in which
-to measure the white noise level of each bolometer (\model{NOI},
-Section~\ref{sec:noi}). Since the common-mode and filtering stages
-will have introduced ringing in the map in the vicinity of bright
-sources, the entire process is iterated. Each component in the dashed
-box of Fig.~\ref{fig:dimm} is re-calculated in sequence. In this way,
-the second time the common-mode is calculated, for example, most of
-the bright astronomical sources in the data have already been removed
-in the previous iteration when the map was estimated, reducing the
-amount of ringing in the map once it is re-estimated. Note that the
-extinction correction is a multiplicative factor, and must be inverted
-prior to re-calculating any of the additive model components to
-preserve the data amplitude. Additive model components from the
-previous iteration are generically added back into the time-series
-immediately prior to their re-calculation. However, for the special
-case of the common-mode and high-pass filter stages, they are replaced
-simultaneously at the start of the iteration to assist with
-convergence (see Sections~\ref{sec:converge}).
+map is used to estimate and remove the astronomical signal from the
+bolometer data (\model{AST}, Section~\ref{sec:ast}), leaving a data
+set that is appropriate for measuring the white noise level of each
+bolometer (\model{NOI}, Section~\ref{sec:noi}). Since the common-mode
+and filtering stages will have introduced ringing in the map in the
+vicinity of bright sources, the entire process is iterated. Each
+component in the dashed box of Fig.~\ref{fig:dimm} is re-calculated in
+sequence. In this way, the second time the common-mode is calculated,
+for example, most of the bright astronomical sources in the data have
+already been removed in the previous iteration when the map was
+estimated, reducing the amount of ringing in the map once it is
+re-estimated. Note that the extinction correction is a multiplicative
+factor, and must be inverted prior to re-calculating any of the
+additive model components to preserve the data amplitude. Additive
+model components from the previous iteration are generically added
+back into the time-series immediately prior to their
+re-calculation. However, for the special case of the common-mode and
+high-pass filter stages, they are replaced simultaneously at the start
+of the iteration to assist with convergence (see
+Sections~\ref{sec:converge}).
 
@@
 samples from $x_i$ in each of the $y_j$, so fractional samples from
-$x_i$ are used at the $y_j$ time boundaries. The algorithm is fast,
-since each sample in $x_i$ need only be visited once. From a spectral
-point of view, this boxcar average is equivalent to applying a
-sinc-function low-pass filter. Such behaviour is desirable, since it
-serves as an anti-aliasing filter; most non-white features above the
-Nyquist frequency in $y_j$ that could contaminate the output data are
-removed.
+$x_i$ are used at the $y_j$ time boundaries: $y_j = (\sum w_i
+x_i)/(\sum w_i)$, where $i$ runs over all the samples from $x$ that
+land within, or partially overlap, $y_j$, and the weights,
+$0<w_i\le1$, indicate the fractions of each $x_i$ that land within
+$y_j$. The algorithm is fast since each sample in $x_i$ need only be
+visited once. From a spectral point of view, this boxcar average is
+equivalent to applying a sinc-function low-pass filter. Such behaviour
+is desirable, since it serves as an anti-aliasing filter; most
+non-white features above the Nyquist frequency in $y_j$ that could
+contaminate the output data are removed.
 
@@
 that have been flagged as bad for any reason (for instance, due to the
-presence of spikes, steps, or unusual common-mode signal) need to be
-excluded from this FFT. For this reason, each contiguous block of bad
-data samples is filled with artificial data before taking the FFT. A
-least squares linear fit is performed to the 50 samples preceding the
-block, and a similar fit is performed to the 50 samples following the
-block. These are used to estimate the expected values at the start and
-end of the block of bad values. The bad values are then replaced by
-linear interpolation between the expected start and end
+presence of spikes, steps, or unusual common-mode signal) when need to
+be excluded from this FFT. For this reason, each contiguous block of
+bad data samples is filled with artificial data before taking the
+FFT. A least squares linear fit is performed to the 50 samples
+preceding the block, and a similar fit is performed to the 50 samples
+following the block. These are used to estimate the expected values at
+the start and end of the block of bad values. The bad values are then
+replaced by linear interpolation between the expected start and end
 values. Gaussian noise is added with a standard deviation equal to the
-mean of the RMS residuals in the two fits.
+mean of the RMS residuals in the two fits. These flagged and filled
+portions of the data are then given a weight of zero when estimating
+the map.
 
@@
 
-First, we describe the generic model for \scuba\ data. We express the
+First, we describe the basic model for \scuba\ data. We express the
 digitised signal observed by the $i$th bolometer as a function of
@@
 During pre-processing, the map-maker divides by $f_i$ once. Then, in
-each iteration, the map-maker divides by $f_i$, models and removes
+each iteration, the map-maker models and removes
 $g_i\mathbf{n}^\mathrm{c}(t)$, divides by $\mathbf{e}_i(t)$, and
@@
 \caption{A selection of typical bolometer time-series after
-flatfielding and removal of a constant baseline. It can be seen that
-most bolometers exhibit a common time variation overlaid with other
+flatfielding and removal of a constant baseline. Each sub-plot shows 
+the power in pW from a single bolometer over a single 30 second period.  
+It can be seen that most bolometers exhibit a common time variation overlaid with other
 features. Empty squares indicate the locations of broken bolometers.}
@@
 opposite signs (such that one or the other bolometer would be
-erroneously flagged).
+erroneously flagged). Experimental models that were investigated for
+magnetic field pickup removal are described in
+Section~\ref{sec:experimental}.
 
@@
 \model{COM} model at block boundaries.  These in turn can lead to
-ringing when filtering, and instabilities in the convergence process.
+ringing when filtering, and instabilities in the convergence
+process.
 
@@
 Fig.~\ref{fig:dimm}, a large fraction of the remaining astronomical
-signal is extracted from the bolometer time-series and transferred to the
-output map, resulting in subsequent estimates of the common-mode being
-more accurate.
+signal is extracted from the bolometer time-series and transferred to
+the output map, resulting in subsequent estimates of the common-mode
+being more accurate. It is also for this reason that the previously
+described step of (re-)flagging outlier portions of the data is
+generally performed iteratively, rather than once as a pre-processing
+step, since bright/compact astronomical sources can be the cause of
+rejection in early iterations.
 
@@
 average of the bolometer data samples $\mathbf{b}_j$ that land within
-that pixel (from any bolometer or point in time),
+that pixel (from any bolometer or point in time, provided that they
+are not flagged as bad or gap-filled, Section~\ref{sec:gaps}),
 %
@@
 floating point precision errors can cause the difference to be
-significantly incorrect. In practice we use the more numerically
-stable ``weighted incremental algorithm'' that calculates incremental
-differences, as described in \citet{west1979}.
+significantly incorrect. In practice we use the superior ``weighted
+incremental algorithm'' that calculates incremental differences, as
+described in \citet{west1979}.
 
@@
 time-domain despiker (Section~\ref{sec:timedespike}), this calculation
-utilises the scatter in the population of samples used to estimate the
-map pixel values (from different times and bolometers) to reject
-outliers. This method is more robust against false-positive detections
-of bright/compact astronomical sources, since transient features in
-the time-series are unlikely to occur by chance whenever a bolometer
+utilises the scatter in the population of samples that land in a map
+pixel (from different times and bolometers) to reject outliers. This
+method is more robust against false-positive detections of
+bright/compact astronomical sources since transient features in the
+time-series are unlikely to occur by chance whenever a bolometer
 crosses a specific location on the sky, whereas real astronomical
@@
 \model{NOI} may optionally perform some cleaning options, such as the
-DC step fixer (Section~\ref{sec:steps}) and spike detection
+step fixer (Section~\ref{sec:steps}) and spike detection
 (Section~\ref{sec:timedespike}), which may work better with these
@@
 
-\subsubsection{Other experimental models}
+\subsubsection{Experimental models for the removal of magnetic field pickup and the atmosphere}
+\label{sec:experimental}
 
@@
 each column as a template that is simply scaled (gain and offset) to
-each bolometer time-series before removal. Since not all columns have
-working dark squids, another alternative model is \model{TMP} in which
-the azimuth of the telescope itself is used as the template. In both
-cases, while the models had some success at removing the magnetic
-field pickup, they did not reduce the $1/f$ knee substantially,
-necessitating the continued use of the \model{FLT} model to remove the
-remaining low-frequency noise. Little or no difference in the final
-maps could be seen with or without the inclusion of these models, and
-so they are not generally used.
+each bolometer time-series before removal. Unfortunately the dark
+squids do not work for every column (several in each subarray),
+meaning that those columns are usually discarded when producing
+maps. One possible remedy is to identify dead bolometers (with
+otherwise working TES readouts) or intentionally disconnect working
+bolometers, to create replacement dark readouts in those columns. This
+solution has not been pursued due to the limited success at removing
+pickup using the presently working dark squids. Another alternative
+model is \model{TMP} in which the azimuth of the telescope itself is
+used as the template. For both \model{DKS} and \model{TMP}, there is
+some success at visibly decreasing the low-frequency scan-synchronous
+noise in certain data sets, although often it does not (and there are
+also examples in which the noise is increased). In the case of
+\model{DKS}, there may simply be components of the signal seen by
+working detectors that are not detecting in the dark
+squids. Furthermore, in the case of \model{TMP}, the relationship
+between the projection of the Earth's magnetic field on the instrument
+and the magnitude of the pickup may not be linearly correlated as we
+have assumed (verified in some cases by a comparison of the azimuthal
+motion with working dark squid signals). In no case was the use of
+these models able to reduce the $1/f$ knee substantially,
+necessitating the continued use of \model{FLT} to remove the remaining
+low-frequency noise. Neither \model{DKS} nor \model{TMP} are typically
+used.
 
@@
 the noise in final maps). This result is basically consistent with
-those from earlier instruments, and suggests that the angular scale of
-the emitting regions in the atmosphere are un-resolved at the \scuba\
-focal plane. \citet{sayers2010}, in particular, examined several
-different ways of modelling and removing the correlated atmospheric
-noise from Bolocam data at 143 and 268\,GHz (also atop Mauna Kea),
-finding only marginal improvements by fitting a plane, or even using
-higher-order polynomials (see their figure~10, as well as the
-discussion and references to earlier work in their section~4.3). In
-\citet{aguirre2011} iterative PCA cleaning was used to remove the
-atmosphere since the simpler methods mentioned left substantial
-residuals in their data (similar to the \scuba\ data described here,
-although PCA cleaning is prohibitively slow in our case).
+those from earlier instruments \citep[e.g., the prediction for \scuba\
+based on SCUBA data described in][]{chapin2002}, and suggests that the
+angular scale of the emitting regions in the atmosphere are
+un-resolved at the \scuba\ focal plane. \citet{sayers2010}, in
+particular, examined several different ways of modelling and removing
+the correlated atmospheric noise from Bolocam data at 143 and 268\,GHz
+(also atop Mauna Kea), finding only marginal improvements by fitting a
+plane, or even higher-order polynomials (see their figure~10, as well
+as the discussion and references to earlier work in their
+section~4.3). In \citet{aguirre2011} iterative PCA cleaning was used
+to remove the atmosphere since the simpler methods mentioned left
+substantial residuals in their data (similar to the \scuba\ data
+described here, although PCA cleaning is prohibitively slow in our
+case).
 
@@
 convergence criterion is met if the change in subsequent values of
-$\chi^2_\mathrm{r}$ is smaller than the requsted threshikd, which is
+$\chi^2_\mathrm{r}$ is smaller than the requsted threshold, which is
 set to $10^{-3}$ by default.
@@
 
-This example illustrate the need for constraints in the map solution
+This example illustrates the need for constraints in the map solution
 in many situations. For calibrators (and other previously known
@@
   noise that is not removed by the common-mode (Uranus peak
-  0.15\,pW). (b) Same as (a) but using 50 iterations, illustrating the
+  0.15\,pW). (b) Same as (a) but using 100 iterations, illustrating the
   degeneracy between large-scale structure and the common-mode (the
@@
   from \citet{owen2008} with $S\gsim15$\,$\mu$Jy. Of the 10 submm
-  peaks, 9 are within this search radius of at least one radio
+  peaks, 9 are within a search radius 8\,arcsec of at least one radio
   source. (d) Jackknife map produced from the difference of two maps,
@@
 be un-resolved by the \scuba\ 7.5--14.5\,arcsec FWHM beams. Under this
-assumption, cross-correlation between the map and the known PSF yields
-the maximum-likelihood flux density of supposed point-sources centred
-over every location in the resulting map \citep[an extremely
-well-known result throughout astronomy, see][]{stetson1987}. Peak
-identification in such smoothed maps have been used extensively in the
-submillimetre community, as both an efficient source-detection and
-photometric measurement strategy. In terms of the angular power
-spectrum, a matched filter may also be thought of as an optimal
-low-pass filter, that suppresses noise on scales smaller (and
-therefore at higher angular frequencies) than the beam. For the case
-at hand, the effective PSF for the whitened maps is given by the
-dotted curve in Fig.~\ref{fig:lockman_psf}. Maps smoothed by this
-shape are shown for the whitened signal, and jackknife maps in
+assumption, and also assuming that the map noise is white,
+cross-correlation between the map and the known PSF yields the
+maximum-likelihood flux density of supposed point-sources centred over
+every location in the resulting map \citep[an extremely well-known
+result throughout astronomy, see][]{stetson1987}. Peak identification
+in such smoothed maps have been used extensively in the submillimetre
+community, as both an efficient source-detection and photometric
+measurement strategy. In terms of the angular power spectrum, a
+matched filter may also be thought of as an optimal low-pass filter,
+that suppresses noise on scales smaller (and therefore at higher
+angular frequencies) than the beam. For the case at hand, the
+effective PSF for the whitened maps is given by the dotted curve in
+Fig.~\ref{fig:lockman_psf}. Maps smoothed by this shape are shown for
+the whitened signal, and jackknife maps in
 Figs.~\ref{fig:lockman_maps}c and f, respectively. Note that these

### 2012-11-15 smurf_paper.tex
Add vertical limits to caption for Fig.9
didn't quite do the 3-way merge correctly...

@@
   Hill, Edinburgh EH9 3HJ\\
-  $^{5}$Institute for Astronomy, University of Edinburgh, Royal Observatory, Blackford Hill Edinburgh, EH9 3HJ}
+  $^{5}$Institute for Astronomy, University of Edinburgh, Royal
+  Observatory, Blackford Hill Edinburgh, EH9 3HJ}
 \begin{document}
@@
   instrument operating on the 15-m James Clerk Maxwell Telescope,
-  consisting nominally of 5120 bolometers (with a useful yield of
+  nominally consisting of 5120 bolometers (with a useful yield of
   69\%) in each of two simultaneous imaging bands centred over 450 and
@@
 reduction written using the Starlink Software Environment
-\citep{1993ASPC...52..229W,2009ASPC..411..418J}. The details of the instrument design,
-performance, and calibration are given in two companion papers:
-\citet{holland2012} and \citet{dempsey2012}.
+\citep{1993ASPC...52..229W,2009ASPC..411..418J}. The details of the
+instrument design, performance, and calibration are given in two
+companion papers: \citet{holland2012} and \citet{dempsey2012}.
 
@@
 experiments for studying anisotropies in the Cosmic Microwave
-Backround (CMB) using similar technology.
+Backround (CMB) using similar technology, the latter of which uses the
+same time-domain multiplexed readout electronics as \scuba\
+\citep{battistelli2008}.
 
@@
 large-angular scales,][]{das2011}. This approach is less practical as
-a solution for \scuba, however, which must serve a broader range of
-scientific interests.
+a general solution for \scuba, however, which must serve a broader
+range of scientific interests.
 
@@
 experimental designs fast iterative methods for the inversion do exist
-without requiring excessive amounts of memory such as the method of
+without requiring excessive amounts of memory, such as that of
 \citet{wright1996} \citep[which was later implemented for SCUBA
@@
 
-Iterative approaches to map-making have a long history in the submillimetre. For example, the iterative pipeline recipe for fitting and
-removing baseline drifts in SCUBA scan-map data as the astronomical
-image estimate improved \citep{1999ASPC..172..171J} is similar to the
-algorithm described in this paper. The closest modern relatives are
-the Comprehensive Reduction Utility \citep[CRUSH,][]{kovacs2008} for
-the Submillimeter High-Angular Resolution Camera 2 (SHARC-2), the
-Bolometer array data Analysis software (BoA) for the Large APEX
-BOlometer CAmera \citep[LABOCA,][]{siringo2009}, and the pipeline
-developed for the Bolocam Galactic Plane Survey \citep{aguirre2011}.
+Iterative approaches to map-making have a long history in the
+submillimetre, such as the pipeline recipe for fitting and removing
+baseline drifts in SCUBA scan-map data as the astronomical image
+estimate improved \citep{1999ASPC..172..171J}. The closest modern
+relatives are the Comprehensive Reduction Utility
+\citep[CRUSH,][]{kovacs2008} for the Submillimeter High-Angular
+Resolution Camera 2 (SHARC-2), the Bolometer array data Analysis
+software (BoA) for the Large APEX BOlometer CAmera
+\citep[LABOCA,][]{siringo2009}, and the pipeline developed for the
+Bolocam Galactic Plane Survey \citep{aguirre2011}.
 
@@
 computers. For reference, all of the data analysis in this paper was
-performed on a machine with 64-bit 8-core central processing unit
+performed on a machine with a 64-bit 8-core central processing unit
 operating at 2.7\,GHz, and 48\,Gb of memory. SMURF has been
@@
 s8a--s8d at 850\,\micron), each consisting of 32 columns and 40 rows
-of bolometers. The bolometers are thermal absorbers coupled to
-superconducting transition-edge sensors (TESs) for thermometry.
-Temperature variations in the TESs produce changing currents, and
-therefore varying magnetic fields, which are detected and amplified
-using chains of superconducting quantum interference devices (SQUIDs),
-before the larger output currents are digitised. Each detector has its
-own SQUID for the first-stage of the amplification, but the remaining
-stages occur within a common chain of SQUIDs for each of the 32
-columns. All 40 rows are read out in sequence at a row-visit frequency
-of about 12\,kHz. Such a high sample rate is unnecessary to produce
-maps, so the data are re-sampled to approximately 200\,Hz before
-writing to disk. This rate provides a sample every 3.0\,arcsec, or
-approximately one third of the 450\,\micron\ diffraction-limited
-full-width at half-maximum (FWHM) -- a typical rule-of-thumb for
-adequately sampling a Gaussian point spread function (PSF) -- at the
-maximum scanning speed of 600\,arcsec\,s$^{-1}$. There is an
-additional 41st row of SQUID readouts that are not connected to
-TESs. These ``dark SQUIDs'' track non-thermal noise sources that are
-common to each column's amplifier chain. The relationship between the
-output digitised current, $I$, and the input power, $P$, or $dI/dP$ is
-established using flatfield observations immediately prior to science
+of bolometers. The bolometers are thermal absorbers (with a response
+time constant of $\tau \sim 1$\,ms) coupled to superconducting
+transition-edge sensors (TESs) for thermometry. Temperature variations
+in the TESs produce changing currents, and therefore varying magnetic
+fields, which are detected and amplified using chains of
+superconducting quantum interference devices (SQUIDs), before the
+larger output currents are digitised. Each detector has its own SQUID
+for the first-stage of the amplification, but the remaining stages
+occur within a common chain of SQUIDs for each of the 32 columns. All
+40 rows are read out in sequence at a row-visit frequency of about
+12\,kHz. Such a high sample rate is unnecessary to produce maps, so
+the data are re-sampled to approximately 200\,Hz before writing to
+disk. This rate provides a sample every 3.0\,arcsec, or approximately
+one third of the 450\,\micron\ diffraction-limited full-width at
+half-maximum (FWHM) -- a typical rule-of-thumb for adequately sampling
+a Gaussian point spread function (PSF) -- at the maximum scanning
+speed of 600\,arcsec\,s$^{-1}$. There is an additional 41st row of
+SQUID readouts that are not connected to TESs. These ``dark SQUIDs''
+track non-thermal noise sources that are common to each column's
+amplifier chain. The relationship between the output digitised
+current, $I$, and the input power, $P$, or $dI/dP$ is established
+using flatfield observations immediately prior to science
 observations, in which the output signal is measured throughout a ramp
@@
 correlated from bolometer to bolometer) $1/f$ drifts, bolometer power
-spectra include a roll-off in their spectra at frequencies approaching
-Nyquist, which is due to the anti-aliasing filter that forms part of
-the 200\,Hz re-sampling stage. There are also line features which are
-thought to be produced at frequencies far above the final 200\,Hz
-sample rate, and are aliased to lower frequencies during the
-multiplexed readout stage before they can be removed by the
-anti-aliasing filter.
+spectra exhibit a roll-off at frequencies approaching Nyquist, which
+is due to the anti-aliasing filter that forms part of the 200\,Hz
+re-sampling stage. This anti-aliasing filter also introduces an
+effective lag of $\sim7$\,ms (which must be added to the 1\,ms thermal
+time constant of the bolometers mentioned above to obtain the total
+lag). There are also line features which are thought to be produced at
+frequencies far above the final 200\,Hz sample rate, and are aliased
+to lower frequencies during the multiplexed readout stage before they
+can be removed by the anti-aliasing filter.
 
@@
 maps, \scuba\ scan strategies have been designed to provide: (i) good
-cross-linking (visiting every point of the mapped area on several
-different time scales, and at different scan angles); and (ii) minimal
-accelerations to reduce turn-around overheads (which could be quite
-large given the 600\,arcsec\,s$^{-1}$ maximum scan speed). For areas
-larger than the array footprint a rectangular ``PONG'' pattern is
-used, in which the boresight travels in approximately straight lines
-and ``bounces'' off the edges at 45\,degree angles until the area is
-uniformly filled in. It is also usually combined with a rotation
-through a number of fixed position angles to create a ``rotating
-PONG'' with even better cross-linking. For smaller areas (of order the
-array footprint, or point sources), in which the PONG turn-around
-overheads would be large, a ``constant velocity daisy'' (CV Daisy) is
-used.  Here the telescope moves in a circle, whose centre also slowly
-traces out a small circle. For a more complete description of the
-\scuba\ observing modes, see Section~5 in \citet{holland2012} and also
+cross-linking (visiting every point of the mapped area at different
+scan angles); and (ii) minimal accelerations to reduce turn-around
+overheads (which could be quite large given the 600\,arcsec\,s$^{-1}$
+maximum scan speed). For areas larger than the array footprint a
+rectangular ``PONG'' pattern is used, in which the boresight travels
+in approximately straight lines and ``bounces'' off the edges at
+45\,degree angles until the area is uniformly filled in. It is also
+usually combined with a rotation through a number of fixed position
+angles to create a ``rotating PONG'' with even better
+cross-linking. For smaller areas (of order the array footprint, or
+point sources), in which the PONG turn-around overheads would be
+large, a ``constant velocity daisy'' (CV Daisy) is used.  Here the
+telescope moves in a circle, whose centre also slowly traces out a
+small circle. For a more complete description of the \scuba\ observing
+modes, see Section~5 in \citet{holland2012} and also
 \citet{2010SPIE.7740E..66K}.
@@
 Fig.~\ref{fig:bolos_mix} and remove it, yielding the grey residual
-signals. These residuals are quite flat, although still with
-noticeable variations. The white noise is also apparent, and larger at
+signals. These residuals are nearly white, although still with
+noticeable long-timescale variations. The white noise is greater at
 450\,\micron\ as one would expect from the larger backgrounds compared
@@
 correlated amongst many of the detectors, and common-mode removal
-corrects it to some extent. Part of the reason that this is seen here,
-and not in the S2SRO data set shown, is that the amplitude of the scan
-pattern is larger. Most of this ``scan-synchronous'' noise is
+corrects it to a large extent. Part of the reason that this is seen
+here, and not in the S2SRO data set shown, is that the amplitude of
+the scan pattern is larger. Most of this ``scan-synchronous'' noise is
 attributed to magnetic field pickup, as described in
@@
 Starlink\footnote{\url{http://www.starlink.ac.uk}} is
-open-source\footnote{\url{https://github.com/Starlink}} (distributed under the GNU General Public Licence v3), and already
-used extensively at the Joint Astronomy Centre (host of the JCMT) for
-many other systems \citep{jenness2011}, which helps with
-interoperability. Though originally written in FORTRAN, many of the
-core Starlink libraries are now ported to native C, or at least have a
-C interface. It was therefore decided to develop SMURF in C as well
-(rather than C++, for example, which would have required adding
-further dependencies to Starlink), although we have taken an
-object-oriented approach. For example, all of the data for a given
-noise model are encapsulated in a C structure, and there is a standard
-interface for all functions that handle models (member data and
-functions for the class, respectively). In this way it is easy to
-extend SMURF with new models. Parallelisation is incorporated in the
-most time-consuming low-level routines using threads (e.g., performing
-Fast Fourier Transforms, re-gridding the data), usually handling
-either independent blocks of bolometers, or blocks of time, in each
-thread, depending on the nature of the calculation. We have found that
-for typical data sets the processing time scales well with the number
-of central processing unit (CPU) cores, although beyond 8 the returns
-are diminished due to tasks that cannot be parallelised (e.g., reading
-the data from disk).
+open-source\footnote{\url{https://github.com/Starlink}} (distributed
+under the GNU General Public Licence v3), and already used extensively
+at the Joint Astronomy Centre (host of the JCMT) for many other
+systems \citep{jenness2011}, which helps with interoperability. Though
+originally written in FORTRAN, many of the core Starlink libraries are
+now ported to native C, or at least have a C interface. It was
+therefore decided to develop SMURF in C as well (rather than C++, for
+example, which would have required adding further dependencies to
+Starlink), although we have taken an object-oriented approach. For
+example, all of the data for a given noise model are encapsulated in a
+C structure, and there is a standard interface for all functions that
+handle models (member data and functions for the class,
+respectively). In this way it is easy to extend SMURF with new
+models. Parallelisation is incorporated in the most time-consuming
+low-level routines using threads (e.g., performing Fast Fourier
+Transforms, re-gridding the data), usually handling either independent
+blocks of bolometers, or blocks of time, in each thread, depending on
+the nature of the calculation. We have found that for typical data
+sets the processing time scales well with the number of central
+processing unit (CPU) cores, although beyond 8 the returns are
+diminished due to tasks that cannot be parallelised (e.g., reading the
+data from disk).
 
@@
 samples modified at each end of the time-series is given by half the
-ratio of the sampling frequency to the lowest retained frequency. In
+ratio of the sampling frequency to the lowest retained frequency, the
+edge frequency of the high-pass filter (Section~\ref{sec:flt}). In
 addition, each end of the time-series is padded with double this
@@
 $\mathbf{n}^\mathrm{c}(t)$.
+
+In principle there are two other effects that could be included in
+this model. First, signals detected by the bolometers (inside the
+brackets of Eq.~\ref{eq:model}) should be convolved with the thermal
+response function, $\exp(-t/\tau)$, where $\tau=1$\,ms. Second, the
+final signals should be convolved with the response of the
+anti-aliasing filter. In practice, the combined effect of these two
+filters is to introduce a lag of $\sim$7\,ms, and only minimal
+attentuation of a point source (at most 3\% for the smaller
+450\,\micron\ PSF at the maximum scan speed of 600\,arcsec). For all
+of the maps described in this paper, their impact is therefore
+negligible. The current version of SMURF (part of the Starlink
+\textsc{kapuahi} release) does not attempt to de-convolve these
+effects, although the next version will likely include an optional
+pre-processing step to perform this task.
 
@@
 flatfielding and removal of a constant baseline. Each sub-plot shows
-the power in pW from a single bolometer over a single 30 second period.  
-It can be seen that most bolometers exhibit a common time variation overlaid with other
-features. Empty squares indicate the locations of broken bolometers.}
+the power from a single bolometer over a 30 second period.
+The lower and upper vertical limits of each plot are -0.02 pW and +0.02
+pW respectively. It can be seen that most bolometers exhibit a common
+time variation overlaid with other features. Empty squares indicate the
+locations of broken bolometers.}
 \label{fig:com}
@@
 
-\subsubsection{Experimental models for the removal of magnetic field pickup and the atmosphere}
+\subsubsection{Experimental models for the removal of magnetic field
+  pickup and the atmosphere}
 \label{sec:experimental}
@@
 each bolometer time-series before removal. Unfortunately the dark
-squids do not work for every column (several in each subarray),
+squids do not work for every column (several cases in each subarray),
 meaning that those columns are usually discarded when producing
@@
 \model{DKS}, there may simply be components of the signal seen by
-working detectors that are not detecting in the dark
+working detectors that are not apparent in the dark
 squids. Furthermore, in the case of \model{TMP}, the relationship
@@
 (absolute calibration), and hence noise performance of the instrument
-in astronomically-useful units.  In this example we produce a 450\,\micron\ map
-of Uranus (observation 26 on 20111017), which is a nearly point-like
-source for \scuba\ that is commonly used as a primary flux
-calibrator. The CV daisy pattern was used, with a scan speed of 155
-arcsec\,sec$^{-1}$. We perform several different reductions of the
-data to illustrate the purpose of various model components and the
-convergence properties of the solution (Fig.~\ref{fig:pointmaps}). In
-all cases the maps are produced on a grid of azimuth (horizontal) and
-elevation (vertical) offsets from the position of Uranus (the origin),
-using 2\,arcsec pixels.
+in astronomically-useful units.  In this example we produce a
+450\,\micron\ map of Uranus (observation 26 on 20111017), which is a
+nearly point-like source for \scuba\ that is commonly used as a
+primary flux calibrator. The CV daisy pattern was used, with a scan
+speed of 155 arcsec\,sec$^{-1}$. We perform several different
+reductions of the data to illustrate the purpose of various model
+components and the convergence properties of the solution
+(Fig.~\ref{fig:pointmaps}). In all cases the maps are produced on a
+grid of azimuth (horizontal) and elevation (vertical) offsets from the
+position of Uranus (the origin), using 2\,arcsec pixels.
 
@@
 shown in Fig.~\ref{fig:pointmaps}a. While the \snr\ of Uranus is
-clearly large ($\sim$250 as estimated by SMURF), enabling us to see
-the faint sidelobes (circle and cross pattern), the map also has
-obvious circular streaks and other large-scale ripples. The circular
-streaks are due to the fact that \model{COM} does not account for all
-of the low-frequency noise (see Fig.~\ref{fig:pspec}), and therefore
+clearly large ($\sim$250), enabling us to see the faint sidelobes
+(circle and cross pattern), the map also has obvious circular streaks
+and other large-scale ripples. The circular streaks are due to the
+fact that \model{COM} does not account for all of the low-frequency
+(i.e., uncorrelated) noise (see Fig.~\ref{fig:pspec}), and therefore
 each bolometer leaves a trace of the circular scan pattern in the map,
-as each of their baselines slowly drift independently. A significant
+as their baselines slowly drift independently. A significant
 contribution to the larger-scale ripples in the map, however, can be
@@
 
-One improvement to remove residual low-frequency noise after
-common-mode removal in this Uranus image, is to simply apply a
-high-pass filter to the data. A third reduction of the data uses the
-default map-making parameters, as described in
+Attacking the problem of streaks, a simple method of removing residual
+(un-correlated) sources of low-frequency noise is to apply a high-pass
+filter after the common-mode removal. A third reduction of the data
+uses the default map-making parameters, as described in
 Section~\ref{sec:algorithm}, which adds the \model{FLT} model to
@@
 scan speed of 155\,arcsec\,sec$^{-1}$, corresponds to a frequency of
-0.78\,Hz. Now using the automatic map-based convergence test
-(Section~\ref{sec:converge}), the resulting map converges after 10
+0.78\,Hz (note in Fig.~\ref{fig:pspec} that this frequency is slightly
+above the $1/f$ knee at 450\,\micron\ after common-mode removal). Now
+using the automatic map-based convergence test
+(Section~\ref{sec:converge}), the solution converges after 10
 iterations (Fig.~\ref{fig:pointmaps}c). Both the circular streaks and
@@
 centre, the result is an azimuthally-symmetric $\sinc$ function-like
-pattern in the map.
+pattern in the map. The gradient resulting from the degeneracy between
+the common-mode and map has been removed because the filter scale of
+200\,arcsec (3.3\,arcmin) is smaller than the array field-of-view
+($\sim$5\,arcmin). If it were greater, degeneracies would again begin
+to appear on those larger scales.
 
@@
 bright, compact sources), a good, simple prior is to constrain the map
-to a value of zero away from the known locations of emission (inside
-the \model{AST} model, Section~\ref{sec:ast}). In
+to a value of zero away from the known locations of emission (part of
+the \model{AST} model calculation, Section~\ref{sec:ast}). In
 Fig.~\ref{fig:pointmaps}d, a solution is produced in an identical
@@
 final iteration. In this case, the map converges after 6 iterations,
-and all of the previous artifacts have been effectively removed. Since
-the map is now flat away from the source, and constrained to a value
-of zero, it is appropriate for performing aperture photometry
-directly, with no need for an additional reference annulus. This
-approach to constrained map-making is similar to that employed for
-poorly cross-linked scans of compact (though resolved) sources by
-\citet{wiebe2009} using BLAST data.
+and the ringing has been effectively removed. Since the map is now
+flat away from the source, and constrained to a value of zero, it is
+appropriate for performing aperture photometry directly, with no need
+for an additional reference annulus. This approach to constrained
+map-making is similar to that employed for poorly cross-linked scans
+of compact (though resolved) sources by \citet{wiebe2009} using BLAST
+data.
 
@@
 response, without prior knowledge of the location of sources; (ii)
-whitening (essentially flattening) the map to make the noise
-properties of the map easier to deal with; and (iii) detecting point
-sources using a ``matched filter''. Note that variations on this
-general procedure have been used extensively in the submm cosmology
-community using previous instruments
+apply a Fourier-space ``whitening filter'' to suppress residual
+large-scale noise; and (iii) detecting point sources using a ``matched
+filter''. Note that variations on this general procedure have been
+used extensively in the submm cosmology community using previous
+instruments
 \citep[e.g.,][]{scott2002,borys2003,laurent2005,coppin2006,scott2008,perera2008,devlin2009}.
@@
 $\sim8.5$\,hours of data taken in 36 separate scans (average
-15\,min. each) spread over February and March 2010, covering an area
-of about 50\,arcmin$^2$. The full list of dates and observation
-numbers includes: 20100218, 63, 64, 65, 70, 71, 72, 90, 91, 92, 97,
-98, 99; 20100220, 111, 112, 113, 118, 128, 129, 130; 20100303, 61, 62,
-64, 69, 70, 72, 73, 74, 75; 20100309, 87, 88, 90, 91; and 20100311,
-59, 64, 65, 72. These observations represent about 80 per cent of the
-total data taken as part of the project; the remaining observations
-were dropped due to problems keeping the arrays properly tuned during
-S2SRO, and were easily identified by their highly variable and erratic
-bolometer time-series (which resulted in maps full of large streaks).
+15\,min. each) spread over February and March 2010. Each scan is a
+360\,arcsec rotating PONG pattern, with a scan speed of
+240\,arcsec\,s$^{-1}$, covering an area of about 50\,arcmin$^2$. The
+full list of dates and observation numbers includes: 20100218, 63, 64,
+65, 70, 71, 72, 90, 91, 92, 97, 98, 99; 20100220, 111, 112, 113, 118,
+128, 129, 130; 20100303, 61, 62, 64, 69, 70, 72, 73, 74, 75; 20100309,
+87, 88, 90, 91; and 20100311, 59, 64, 65, 72. These observations
+represent about 80 per cent of the total data taken as part of the
+project; the remaining observations were dropped due to problems
+keeping the arrays properly tuned during S2SRO, and were easily
+identified by their highly variable and erratic bolometer time-series
+(which resulted in maps full of large streaks).
 
@@
   followed by 4 iterations using only the \model{COM} model to remove
-  residual low-frequency noise. White contours correspond to estimated
+  residual correlated noise. White contours correspond to estimated
   noise levels of 1.25, 2.5 and 5.0 times the minimum noise at the
@@
 the cutoff is only removed through iterative common-mode
-subtraction. Since the data are high-pass filtered prior to the
-iterative solution, \model{GAI} (fitting an independent amplitude of
-\model{COM} to each bolometer) has been de-activated, since there is
-very little structure in the common-mode with which to fit an accurate
-gain. The map is shown in Fig.~\ref{fig:lockman_maps}a. The map-maker
-has been tested in two ways: (i) large numbers of iterations are used
-to verify that the maps converge without the growth of large
-structure; and (ii) adding synthetic sources to the time-series data
-at a range of brightnesses verify that the map-maker response to them
-is linear (i.e., the relative shape and amplitude compared to the
-input source is independent of brightness). The response to a
-synthetic point source (solid line) after map-making (dotted line) is
-shown in Fig.~\ref{fig:lockman_psf}. Clearly the use of a high-pass
-filter as a pre-processing step, and having no other map-constraints,
-has the down-side of introducing sidelobes around the main
-peak. However, the way this filter affects point-sources is both
-well-known, and linear.
+subtraction. In this case, the filter edge has been chosen to remove
+scales larger than 200\,arcsec, or a high-pass filter above 1.2\,Hz
+given the scan speed. Since the data are high-pass filtered prior to
+the iterative solution, \model{GAI} (fitting an independent amplitude
+of \model{COM} to each bolometer) has been de-activated, since there
+is very little structure in the common-mode with which to fit an
+accurate gain. The map is shown in Fig.~\ref{fig:lockman_maps}a. The
+map-maker has been tested in two ways: (i) large numbers of iterations
+are used to verify that the maps converge without the growth of large
+structure; and (ii) adding synthetic sources to the real time-series
+data (a built-in feature of SMURF) at a range of brightnesses verify
+that the map-maker response to them is linear (i.e., the relative
+shape and amplitude compared to the input source is independent of
+brightness). The response to a synthetic point source (solid line)
+after map-making (dotted line) is shown in
+Fig.~\ref{fig:lockman_psf}. Clearly the use of a high-pass filter as a
+pre-processing step, and having no other map-constraints, has the
+down-side of introducing sidelobes around the main peak. Furthermore,
+the details of this shape depends on the high-pass filter edge that
+has been chosen (the higher the frequency, the greater the attenuation
+of the central peak, and the larger the negative side-lobes).
+However, the way this filter affects point-sources is measurable
+(using the synthetic source injection facility of SMURF), and linear.
 
@@
 corresponding to the default \model{FLT} high-pass filter. Similar to
-Fig.~\ref{fig:pointmaps}c, the lack of prior constraints on the map
-leads to degeneracies, and therefore ripples on scales larger than the
-array footprint (due to common-mode subtraction) and filter scale
-(especially evident around the brightest peak at the centre of the
-map). Panels (c) and (d) show the ``bright extended'' reduction, in
-which a zero-mask is created iteratively from all of the pixels that
-exceed an \snr\ of 5. While this region (red contour) only encompasses
-the brightest peaks early in the solution, in the final iteration it
-identifies most of the bright, extended emission, and significantly
-helps with negative ringing.}
+Fig.~\ref{fig:pointmaps}c, the high-pass filtering introduces ringing
+around bright sources. Panels (c) and (d) show the ``bright extended''
+reduction, in which a zero-mask is created iteratively from all of the
+pixels that lie below a \snr\ of 5. While this region (outside the red
+contour) only avoids the brightest peaks early in the solution, in the
+final iteration it skirts most of the bright, extended emission, and
+significantly helps with negative ringing.}
 \label{fig:m17}
@@
 the noise weights have been measured) and 17 iterations (when the map
-has converged). The first panel also depicts the array footprint, and
-the angular scale corresponding to the high-pass filter edge. Much
-like the reduction of a point source without any prior constraints
-(Fig.~\ref{fig:pointmaps}c), degeneracies on scales larger than the
-array footprint and filter produce ripples in the map.
+has converged), which uses iterative common-mode subtraction and
+high-pass filtering. The first panel also depicts the array footprint,
+and the angular scale (300\,arcsec) corresponding to the high-pass
+filter edge (0.6\,Hz). Much like the reduction of a point source
+without any prior constraints (Fig.~\ref{fig:pointmaps}c), there are
+ripples around bright sources due to the filtering.
 
@@
 has been added to SMURF to generate one automatically by flagging
-pixels above some \snr\ threshold after each iteration as part of the
-``bright extended'' configuration.
+pixels below some \snr\ threshold to be set to zero after each
+iteration as part of the ``bright extended'' configuration.
 
 The results of this automatic masking are shown in the bottom panels
-of Fig.~\ref{fig:m17}. After the second iteration, only narrow regions
-around the brightest peaks are identified. However, as the solution
-progresses, the negative bowls around the bright sources are slowly
-reduced and the mask ``grows'' out from the brightest areas. By the
-final iteration, most of the obvious structures in the data are
-encompassed by the mask, negative bowling is significantly reduced,
-and the brightest regions are more extended.
+of Fig.~\ref{fig:m17}. After the second iteration, everything but the
+brightest peaks are set to zero (outside the red contours). However,
+as the solution progresses, the negative bowls around the bright
+sources are slowly reduced and the mask ``grows'' out from the
+brightest areas. By the final iteration, most of the obvious
+structures in the data are excluded in the the mask, negative bowling
+is significantly reduced, and the brightest regions are more extended.
 
@@
 first row. Third row: average of the two halves of the data using the
-bright extended reduction. The blue and red contours indicate the
-zero-masked regions for each half of the data (note for the
-300\,arcsec case that the region about M17 fairly closely matches the
-mask in Fig.~\ref{fig:m17}d). Clearly as the filter scale is
-increased, due to the high \snr\ of the data, larger emission regions
-are (correctly) detected and reproduced in the map. Fourth (bottom)
-row: Jackknife maps for the bright extended reductions.}
+bright extended reduction. The regions outside the blue and red
+contours are constrained to zero for each half of the data (note for
+the 300\,arcsec case that these regions about M17 closely match the
+mask in Fig.~\ref{fig:m17}d using a full reduction). Clearly as the
+filter scale is increased, due to the high \snr\ of the data, larger
+emission regions are (correctly) detected and reproduced in the
+map. Fourth (bottom) row: Jackknife maps for the bright extended
+reductions.}
 \label{fig:m17_jk}
@@
 
-While the reduction in the bottom panels of Fig.~\ref{fig:m17} is (at
-least a cosmetic sense) superior to those in the top panels, it is
-important to quantify both the noise properties of the maps, and the
-response to real structures (the transfer function). We would also
-like to know how each are affected by our choice of filter scale.
-Similar to the previous section, we will use a jackknife test to
-estimate the noise, as well as injecting known sources into the data
-to observe how they are attenuated.
+While the reduction in the bottom panels of Fig.~\ref{fig:m17} is (in
+a cosmetic sense) superior to those in the top panels, it is important
+to quantify both the noise properties of the maps, and the response to
+real structures (the transfer function). We would also like to know
+how each are affected by our choice of filter scale.  Similar to the
+previous section, we will use a jackknife test to estimate the noise,
+as well as injecting known sources into the data to observe how they
+are attenuated.
 
@@
 columns show reductions using 150, 300, 600, and 900\,arcsec filter
-edges. For reference, the largest scale that is completely inscribed
+edges. The synthetic data are clearly seen as the circular region
+south of M17. As the filter scale is increased, the size of the
+ripples increases accordingly. While larger astronomical structures do
+seem to appear, negative bowls are a major problem without any other
+map constraints. Since the largest scale that is completely inscribed
 by the array footprint is about 400\,arcsec, and the diagonal of the
-array is about 600\,arcsec (meaning that scales beyond 600\,arcsec are
-completely unconstrained when common-mode rejection is used).  The
-synthetic data are clearly seen as the circular region south of
-M17. As the filter scale is increased, the size of the ripples
-increases accordingly. While larger structures do seem to appear,
-negative bowls are a major problem without any other map
-constraints.The second row of Fig.~\ref{fig:m17_jk} shows the
-jackknife maps. The astronomical emission is almost perfectly removed,
-except for a slight impression of M17 near the centre of the map, that
-is probably due to some combination of detector gain and pointing
-drifts. Otherwise the jackknife appears to be a plausible realisation
-of noise from the same parent distribution as for the averaged maps.
+array is about 600\,arcsec, scales ranging from $\sim$400--600\,arcsec
+up to the scale of the filter will be unconstrained due to the
+degeneracy between the common-mode and map (see discussion in
+Section~\ref{sec:point}). This degeneracy therefore causes some of the
+large-scale patchiness in the 600 and 900\,arcsec filtered maps that
+lie away from the central bright sources.  The second row of
+Fig.~\ref{fig:m17_jk} shows the jackknife maps. The astronomical
+emission is almost perfectly removed, except for a slight impression
+of M17 near the centre of the map, that is probably due to some
+combination of detector gain and pointing drifts. Otherwise the
+jackknife appears to be a plausible realisation of noise from the same
+parent distribution as for the averaged maps.
 
@@
 exercise using the bright extended configuration, in which the \snr\
-threshold of 5 is again used to constrain the map. As the filter scale
-is increased, more of the extended structure in M17 is reproduced in
-the map, as evidenced by the blue and red contours (masks generated
-from the first and second halves of the data, respectively). The
-masking does a generally good job of suppressing the largest-scale
-ripples that are produced by the default reduction. However, the noise
-away from regions of bright emission does increase noticeably (mottled
-appearance). With filter scales $\gsim600$\,arcsec, virtually the
-entire region of synthetic emission is correctly identified by the \snr\
-mask and allowed to vary freely in the solution.
+threshold of 5 is again used to identify low-significance pixels that
+are set to zero after each iteration. As the filter scale is
+increased, more of the extended structure in M17 is reproduced in the
+map, as evidenced by the blue and red contours (masks generated from
+the first and second halves of the data, respectively). The masking
+does a generally good job of suppressing the largest-scale ripples
+that are produced by the default reduction. However, the noise away
+from regions of bright emission does increase noticeably (mottled
+appearance) --- due to residual $1/f$ noise following common-mode
+removal that has a knee at a frequency above the filter edge. With
+filter scales $\gsim600$\,arcsec, virtually the entire region of
+synthetic emission is correctly identified by the \snr\ mask and
+allowed to vary freely in the solution.
 
@@
 One case in which the \snr\ is worse using the bright extended
-reduction is when using a 150\,arcsec filter. In this case the noise
-is considerably larger in the bright extended reduction, as evidenced
-by the ``kink'' near 150\,arcsec. Referring to the mask contours in
-the left panel of the third row, it is clear that the map-maker has
-failed to identify much of the bright, extended emission in the region
-of the synthetic source. Each area that is not within the contours is
-constrained to zero throughout the solution, therefore suppressing
-power (and lowering the transfer function), and subsequently reducing
-the \snr\ of the final result. This measurement serves as a warning:
-the map-maker response is non-linear when using \snr\ masking. Harsh
-filtering can provide misleading results, as in this example. Maps of
-faint extended emission will also suffer considerably, as the
-structures of interest may not clear the \snr\ threshold for the mask.
+reduction is when using a 150\,arcsec filter. Here the noise is
+considerably larger in the bright extended reduction, as evidenced by
+the ``kink'' near 150\,arcsec. Referring to the mask contours in the
+left panel of the third row in Fig,~\ref{fig:m17_jk}, it is clear
+that the map-maker has failed to identify much of the bright, extended
+emission in the region of the synthetic source. Each area that is not
+within the contours is constrained to zero throughout the solution,
+therefore suppressing power (and lowering the transfer function), and
+subsequently reducing the \snr\ of the final result. This measurement
+serves as a warning: the map-maker response is non-linear when using
+\snr\ masking. Harsh filtering can provide misleading results, as in
+this example. Maps of faint extended emission will also suffer
+considerably, as the structures of interest may not clear the \snr\
+threshold for the mask.
 
@@
 restricts the solution to a small fixed number of iterations
-($\lsim$10), so that there is probably little chance that such
-structures have the chance to grow. Experimentation with the model
-order is also advocated to gain an impression of the convergence
-properties. This approach is perhaps more relevant to their SHARC-2
-data for which a more complicated model is developed; the degeneracies
-we have observed for \scuba\ maps using our more brute-force approach
-with a high-pass filter tend to look similar regardless of the order
-(only the convergence time is affected).  For Bolocam data of the
-Galactic Plane, \citet{aguirre2011} used a maximum-entropy filtering
-step to suppress large-scales. Regardless of the method used,
-simulations are always required to establish the transfer function and
-noise properties as a function of angular scale.
+($\lsim$10), so that there is probably little opportunity for
+degeneracies between the map and common-mode to appear in the map.
+Experimentation with the model order is also advocated to gain an
+impression of the convergence properties. This approach is perhaps
+more relevant to their SHARC-2 data for which a more complicated model
+is developed; the degeneracies we have observed for \scuba\ maps using
+our more brute-force approach with a high-pass filter tend to look
+similar regardless of the order (only the convergence time is
+affected).  For Bolocam data of the Galactic Plane,
+\citet{aguirre2011} used a maximum-entropy filtering step to suppress
+large-scales. Regardless of the method used, simulations are always
+required to establish the transfer function and noise properties as a
+function of angular scale.
 
@@
 angular scales due to the degeneracy between the map, and the
-low-frequency signal components that are removed. A simple strategy of
+low-frequency signal components that are removed (namely the
+common-mode). In addition, the high-pass filtering produces
+significant ringing around bright sources. A simple strategy of
 constraining empty regions of the map to zero (using either a
 user-supplied mask for known sources, or an iterative determination of
-signal above some \snr\ threshold) provides good constraints for both
+signal below some \snr\ threshold) provides good constraints for both
 compact objects, and bright/extended structures. Particularly in the
@@
 filter at the start of the reduction produces maps that are nearly
-white-noise limited. Residual large-scale noise can be removed with a
-whitening filter (also established from jackknife estimates of the
-noise), and sources detected using a matched-filter (smoothing with
-the effective filtered point spread function).
+white-noise limited and linear (i.e., the response does not depend on
+\snr). Residual large-scale noise can be removed with a whitening
+filter (also established from jackknife estimates of the noise) based
+on the Fourier Transform of the maps, and sources detected using a
+matched-filter (smoothing with the effective filtered point spread
+function).
 
@@
 for a wide variety of observations. Furthermore, the execution times
-are shorter than the observation lengths, and memory requirements for
-even the longest \scuba\ observations are within the capabilities of
-single, high-end desktop computers. SMURF can therefore provide real-time
-feedback at the telescope to observers, and it also produces nearly
-science-grade data products for the JCMT Science Archive hosted by the
-Canadian Astronomy Data Centre.
+are typically shorter than the observation lengths, and memory
+requirements for even the longest \scuba\ observations are within the
+capabilities of single, high-end desktop computers. SMURF can
+therefore provide real-time feedback at the telescope to observers,
+and it also produces nearly science-grade data products for the JCMT
+Science Archive hosted by the Canadian Astronomy Data Centre.
 

### 2012-11-16 smurf_paper.tex
added some flagging statistics as requested by referee

@@
 glitches that may hinder the iterative solution from converging, or
-simply tasks that do not need to be iterated.
+simply tasks that do not need to be iterated. All of the data that are
+flagged as bad are ignored when estimating the map. As an example, for
+the 450\,\micron\ scan of Uranus that will be presented in
+Section~\ref{sec:point}: 31.00\% of the data (1587/5120 bolometers)
+are flagged due to data acquisition problems (dead or deactivated
+bolometers, or non-linear flatfields); 0.12\% of the data are flagged
+due to steps (including a small region about the steps); and 0.83\% if
+the data are flagged because the telescope was stationary.
 
@@
 
-\subsubsection{Additional bad bolometer rejection}
+\subsubsection{Additional data rejection}
 \label{sec:flagbadbol}
@@
 higher-frequency portion of the PSD is generally quite
-clean. Furthermore, even if there is contamination, the purpose of
-this measurement is to identify outliers, rather than provide a
-meaningful absolute measurement of a given bolometer's white noise
-level. Both high and low (e.g., due to an incorrect and very small
-flatfield correction being applied) outliers from the centre of the
-logarithm (to reduce the impact of outliers) of the bolometer noise
-distribution are flagged.
+flat. Furthermore, even if there is contamination, the purpose of this
+measurement is to identify outliers, rather than provide a meaningful
+absolute measurement of a given bolometer's white noise level. Both
+high and low (e.g., due to an incorrect and very small flatfield
+correction being applied) outliers from the centre of the logarithm
+(to reduce the impact of outliers) of the bolometer noise distribution
+are flagged.
+
+Finally, data that are taken while the telescope is either moving
+extremely slowly, or extremely fast, are flagged and ignored when
+estimating the map. When the telescope is moving slowly, astronomical
+signals appear at low frequencies in the bolometer data that are
+dominated by $1/f$ noise (in the extreme case of no motion, the
+bolometers have zero sensitivity). When the telescope is moving
+significantly faster than 600\,arcsec\,s$^{-1}$, both the thermal
+response of the bolometers, and the anti-aliasing filter in the
+readout electronics (Section~\ref{sec:bolos}) can significantly
+distort signals.  The default allowable range of speed is
+30--1000\,arcsec\,s$^{-1}$, which primarily rejects short periods when
+the telescope was stationary prior to commencing the scan pattern
+(generally speaking the speed is constant once the scan begins),
+although the scan speed can briefly exceed the upper limit, even for a
+requested speed of 600\,arcsec, if the pointing system is having
+trouble keeping-up for large-area scans at high elevation.
 
@@
 components.
+
+Additional flagging of data to be ignored in the map occurs during the
+iterations. For the example of the 450\,\micron\ Uranus scan in
+Section~\ref{sec:point}: an additional 55 spikes are detected using
+the map-based despiker ($1.2\times10^{-6}$ \% of the data); and 0.72\%
+of the data are flagged as outliers compared to the common-mode.
 

### 2012-11-21 smurf_paper.tex
more edits... figure panel labels, chi^2, acknowledge Dan...

@@
 \centering
-\includegraphics[width=0.49\linewidth]{bolos_point_mix_s2sro}
-\includegraphics[width=0.49\linewidth]{bolos_point_mix}
+
+\begin{minipage}[h]{0.495\linewidth}
+\textbf{(a) S2SRO} \\
+
+\includegraphics[width=\linewidth]{bolos_point_mix_s2sro}
+\end{minipage}
+\begin{minipage}[h]{0.495\linewidth}
+\textbf{(b) fully-commissioned} \\
+
+\includegraphics[width=\linewidth]{bolos_point_mix}
+\end{minipage}
+
 \caption{A comparison between single bolometer time-series in each of
@@
   top two panels show the residual time-series after removing the
-  common-mode signals.  Left: data taken during the S2SRO period,
+  common-mode signals.  (a) Data taken during the S2SRO period,
   observation 29 on 20100313. There is a strong correlation between
@@
   noise at 450\,\micron\ is significantly larger than at
-  850\,\micron. Right: data taken with the fully-commissioned
-  instrument, observation 38 on 20111112. Unlike the S2SRO data, there
-  is no strong signal produced by variations in the fridge
-  temperature. The 450\,\micron\ data have minimal correlated
-  low-frequency noise (as evidenced by the lack of a difference
-  between the raw and common-mode subtracted signals), although the
-  850\,\micron\ data show a common signal correlated with the opacity
-  and telescope motion (note that this scan has a significantly larger
-  amplitude than the S2SRO data set), mostly likely caused by a
-  combination of changing airmass (anti-correlated with opacity), and
-  magnetic field pickup (see Section~\ref{sec:magpickup} and
+  850\,\micron. (b) Data taken with the fully-commissioned instrument,
+  observation 38 on 20111112. Unlike the S2SRO data, there is no
+  strong signal produced by variations in the fridge temperature. The
+  450\,\micron\ data have minimal correlated low-frequency noise (as
+  evidenced by the lack of a difference between the raw and
+  common-mode subtracted signals), although the 850\,\micron\ data
+  show a common signal correlated with the opacity and telescope
+  motion (note that this scan has a significantly larger amplitude
+  than the S2SRO data set), mostly likely caused by a combination of
+  changing airmass (anti-correlated with opacity), and magnetic field
+  pickup (see Section~\ref{sec:magpickup} and
   Fig.~\ref{fig:magpickup}).}
@@
 \centering
-\includegraphics[width=0.49\linewidth]{pspec_s2sro}
-\includegraphics[width=0.49\linewidth]{pspec}
-\caption{Bolometer power spectral densities for the same two data sets
-  (before and after upgrades) used in Fig.~\ref{fig:bolos_mix}. The
-  PSDs have been boxcar smoothed with a width of 0.1\,Hz to reduce the
-  noise slightly and clarify some features. Four of the most sensitive
-  bolometers have been selected at each wavelength, and the
-  flat-fielded and step-corrected (but otherwise raw) PSDs are shown
-  as coloured dotted lines (the blue signals are for the same time
-  series as those shown in Fig.~\ref{fig:bolos_mix}).  The solid black
-  lines are the PSDs of the common-mode signals at each wavelength,
-  and the solid coloured lines show the PSDs of the bolometers once
-  the common-mode is removed.  Finally, the dashed black lines show
-  the spectral shape produced by a point source given the scan speeds
-  for the two observations (120\,arcsec\,s$^{-1}$ in the left-hand
-  plot, and 190\,arcsec\,s$^{-1}$ in the right-hand plot); for
-  reference, the top horizontal axes shows the conversion from
-  frequency to angular scale. Horizontal dotted lines at $2 \times
-  10^5$ and $1 \times 10^5$\,pW\,Hz$^{-1}$ at 450 and 850\,\micron,
-  respectively, are provided as a visual reference for the white-noise
-  levels. In addition to $1/f$ and white noise components, and line
-  features, all of the PSDs exhibit the gradual roll-off of the
-  anti-aliasing filter just below the Nyquist frequency.  Left: for
-  the S2SRO data there are broad line features in the PSDs at both
-  wavelengths above $\sim$35\,Hz.  At lower frequencies, the bolometer
-  signals exhibit clear $1/f$ knees at approximately 1 and 2\,Hz at
-  450 and 850\,\micron. Common-mode subtraction removes most of the
-  correlated fridge oscillation signal, lowering the $1/f$ knees to
-  approximately 0.2 and 0.7\,Hz at 450 and 850\,\micron,
-  respectively. Right: data from the fully-commissioned instrument
-  tend to have lower $1/f$ knees, and fewer line features. The
-  remaining low-frequency noise, now that the fridge oscillations have
-  been removed, is considerably more independent from bolometer to
-  bolometer than in the S2SRO data (note the larger spread in the
-  dotted lines at low frequencies compared to the left-hand plot,
-  particularly at 450\,\micron). Finally, this example shows that the
-  white noise performance is in fact similar for the two subarrays
-  (s4a and s8b) that were used both before and after the upgrades.}
+
+\begin{minipage}[h]{0.495\linewidth}
+\textbf{(a) S2SRO} \\
+
+\includegraphics[width=\linewidth]{pspec_s2sro}
+\end{minipage}
+\begin{minipage}[h]{0.495\linewidth}
+\textbf{(b) fully-commissioned} \\
+
+\includegraphics[width=\linewidth]{pspec}
+\end{minipage}
+
+\caption{Bolometer power spectral densities (PSDs) for the same two
+  data sets (before and after upgrades) used in
+  Fig.~\ref{fig:bolos_mix}. The PSDs have been boxcar smoothed with a
+  width of 0.1\,Hz to reduce the noise slightly and clarify some
+  features. Four of the most sensitive bolometers have been selected
+  at each wavelength, and the flat-fielded and step-corrected (but
+  otherwise raw) PSDs are shown as coloured dotted lines (the blue
+  signals are for the same time series as those shown in
+  Fig.~\ref{fig:bolos_mix}).  The solid black lines are the PSDs of
+  the common-mode signals at each wavelength, and the solid coloured
+  lines show the PSDs of the bolometers once the common-mode is
+  removed.  Finally, the dashed black lines show the spectral shape
+  produced by a point source given the scan speeds for the two
+  observations (120\,arcsec\,s$^{-1}$ in (a), and
+  190\,arcsec\,s$^{-1}$ in (b); for reference, the top horizontal axes
+  shows the conversion from frequency to angular scale. Horizontal
+  dotted lines at $10^{-7}$ and $10^{-8}$\,pW\,Hz$^{-1}$ at 450 and
+  850\,\micron, respectively, are provided as a visual reference for
+  the white-noise levels. In addition to $1/f$ and white noise
+  components, and line features, all of the PSDs exhibit the gradual
+  roll-off of the anti-aliasing filter just below the Nyquist
+  frequency.  (a) For the S2SRO data, particularly at 450\,\micron,
+  there are broad line features in the PSDs at both wavelengths above
+  $\sim$35\,Hz.  At lower frequencies, the bolometer signals exhibit
+  clear $1/f$ knees at approximately 2 and 3\,Hz at 450 and
+  850\,\micron. Common-mode subtraction removes most of the correlated
+  fridge oscillation signal, lowering the $1/f$ knees to approximately
+  0.6 and 1.0\,Hz at 450 and 850\,\micron, respectively. (b) Data from
+  the fully-commissioned instrument tend to have lower $1/f$ knees,
+  and fewer line features. For this example, the improvement is most
+  striking at 450\,micron\ where the PSD is approximately two
+  orders-of-magnitude lower in the fully-commissioned data at
+  0.1\,Hz. The white noise performance, however, is similar for the
+  two subarrays (s4a and s8b) that were used both before and after the
+  upgrades.}
 \label{fig:pspec}
@@
 
-In the S2SRO data (left panel), observation 29 on 20100313, both
-bolometers share significant long-timescale structure ($\gsim10$\,s)
-that appears to be related to variations in the fridge base
-temperature, although the similarity is clearly greater at
+In the S2SRO data (Fig.~\ref{fig:bolos_mix}a), observation 29 on
+20100313, both bolometers share significant long-timescale structure
+($\gsim10$\,s) that appears to be related to variations in the fridge
+base temperature, although the similarity is clearly greater at
 850\,\micron. In this particular case, the total power in the
@@
 Data from the fully commissioned instrument, observation 38 on
-20111112, are shown in the right panel of
-Fig.~\ref{fig:bolos_mix}. Having solved the fridge oscillation
-problem, these data no longer exhibit a correlation with the mixing
-chamber (however, note that variations are still seen in the mixing
-chamber signal; such variations do not necessarily reflect changes in
-the focal plane temperature). There is minimal correlated
-low-frequency noise in these 450\,\micron\ data (note that there is
-almost no difference once the common-mode has been subtracted). The
-850\,\micron\ channel, however, exhibits a more significant signal
-that is obviously correlated with the telescope motion. It is also
-correlated amongst many of the detectors, and common-mode removal
-corrects it to a large extent. Part of the reason that this is seen
-here, and not in the S2SRO data set shown, is that the amplitude of
-the scan pattern is larger. Most of this ``scan-synchronous'' noise is
-attributed to magnetic field pickup, as described in
-Section~\ref{sec:magpickup}.
+20111112, are shown in Fig.~\ref{fig:bolos_mix}b. Having solved the
+fridge oscillation problem, these data no longer exhibit a correlation
+with the mixing chamber (however, note that variations are still seen
+in the mixing chamber signal; such variations do not necessarily
+reflect changes in the focal plane temperature). There is minimal
+correlated low-frequency noise in these 450\,\micron\ data (note that
+there is almost no difference once the common-mode has been
+subtracted). The 850\,\micron\ channel, however, exhibits a more
+significant signal that is obviously correlated with the telescope
+motion. It is also correlated amongst many of the detectors, and
+common-mode removal corrects it to a large extent. Part of the reason
+that this is seen here, and not in the S2SRO data set shown, is that
+the amplitude of the scan pattern is larger. Most of this
+``scan-synchronous'' noise is attributed to magnetic field pickup, as
+described in Section~\ref{sec:magpickup}.
 
@@
 scale assuming the scan speeds of each observation
-(120\,arcsec\,s$^{-1}$ in the left-hand plot, and
-190\,arcsec\,s$^{-1}$ in the right-hand plot). The power spectra of
+(120\,arcsec\,s$^{-1}$ in Fig.~\ref{fig:pspec}a, and
+190\,arcsec\,s$^{-1}$ in Fig.~\ref{fig:pspec}b). The power spectra of
 unresolved point-sources are also shown as dashed lines in each band
@@
 
-In the S2SRO data (left-hand panel of Fig.~\ref{fig:pspec}), the
-low-frequency noise is very correlated between the detectors (note the
-tight scatter in the dotted coloured lines, and their resemblance to the
-common-mode), due to it being dominated by the fridge
-oscillations. Common-mode subtraction is very effective, reducing the
-$1/f$ knees in both wavelengths by about a factor of 5.
-
-Raw data from the fully-commissioned instrument (right-hand panel of
-Fig.~\ref{fig:pspec}) generally have less significant $1/f$ noise as
+In the S2SRO data (Fig.~\ref{fig:pspec}a), the low-frequency noise is
+very correlated between the detectors (note the tight scatter in the
+dotted coloured lines, and their resemblance to the common-mode), due
+to it being dominated by the fridge oscillations. Common-mode
+subtraction is very effective, reducing the $1/f$ knees in both
+wavelengths by about a factor of 5.
+
+Raw data from the fully-commissioned instrument
+(Fig.~\ref{fig:pspec}b) generally have less significant $1/f$ noise as
 compared with the S2SRO data, and it is considerably less correlated
@@
 analysis of the s4a and s8b subarrays for the first $\sim$100\,s of
-the same observation that was used in the right-hand
-(fully-commissioned) panels of Figs.~\ref{fig:bolos_mix} and
-\ref{fig:pspec}. The 12 most significant components are shown (ranked
+the same observation that was used in Figs.~\ref{fig:bolos_mix}b and
+\ref{fig:pspec}b. The 12 most significant components are shown (ranked
 by the mean eigenvalues across all working bolometers in both
@@
 cases, the eigenvector time-series exhibit a roughly periodic signal
-that resembles the scan pattern in the right-hand panel of
-Fig.~\ref{fig:bolos_mix}. While these correlated signals also clearly
-appear at both wavelengths, referring to the maps of eigenvalues, they
-are considerably stronger at 850\,\micron\ (consistent with the visual
-appearance of the bolometer signals in Fig.~\ref{fig:pca}). Also note
-that the PSDs for these eigenvectors exhibit nearly a pure $1/f$
-signature with almost no line features, or a white-noise plateau
-(suggesting that these are high-\snr\ measurements of a purely
-low-frequency drift).
+that resembles the scan pattern in Fig.~\ref{fig:bolos_mix}b. While
+these correlated signals also clearly appear at both wavelengths,
+referring to the maps of eigenvalues, they are considerably stronger
+at 850\,\micron\ (consistent with the visual appearance of the
+bolometer signals in Fig.~\ref{fig:pca}). Also note that the PSDs for
+these eigenvectors exhibit nearly a pure $1/f$ signature with almost
+no line features, or a white-noise plateau (suggesting that these are
+high-\snr\ measurements of a purely low-frequency drift).
 
@@
   subarray) and 850\,\micron\ (s8b subarray) time-series bolometer
-  data, for the same observation as that used in the right-hand panels
-  of Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec}. For each component
+  data, for the same observation as that used in
+  Figs.~\ref{fig:bolos_mix}b and \ref{fig:pspec}b. For each component
   the top plot shows the time-series of the eigenvectors normalised by
@@
   appears to be dominated by scan-synchronous noise (compare with the
-  scan pattern in the right-hand panel of
-  Fig.~\ref{fig:bolos_mix}). Some of this signal also appears in
-  Component~3. Component~2 is another $1/f$ signal that is dominant in
-  the 850\,\micron\ subarray. Most of the remaining components consist
-  of weaker $1/f$ noise and line features predominantly above
-  $\sim$30\,Hz. The eigenvalue maps show correlated patterns along
-  columns suggesting pickup in their common amplifier chains.}
+  scan pattern in Fig.~\ref{fig:bolos_mix}b). Some of this signal also
+  appears in Component~3. Component~2 is another $1/f$ signal that is
+  dominant in the 850\,\micron\ subarray. Most of the remaining
+  components consist of weaker $1/f$ noise and line features
+  predominantly above $\sim$30\,Hz. The eigenvalue maps show
+  correlated patterns along columns suggesting pickup in their common
+  amplifier chains.}
 \label{fig:pca}
@@
 Starlink), although we have taken an object-oriented approach. For
-example, all of the data for a given noise model are encapsulated in a
-C structure, and there is a standard interface for all functions that
-handle models (member data and functions for the class,
-respectively). In this way it is easy to extend SMURF with new
+example, all of the data for a given signal component model are
+encapsulated in a C structure, and there is a standard interface for
+all functions that handle models (member data and functions for the
+class, respectively). In this way it is easy to extend SMURF with new
 models. Parallelisation is incorporated in the most time-consuming
@@
 \begin{figure*}
-\centering
+\raggedright
+\textbf{(a) large step} \hspace{7.3cm} \textbf{(b) small steps}\\
+
+\vspace{0.2cm}
+
 \includegraphics[width=\linewidth]{steps1}
-\caption{Examples of steps in time-series data. Steps occur with a wide
-range of heights from the large steps shown on the left to the small
-steps shown on the right. Large steps are often followed by a brief
-``over-shoot'' as shown on the left. In both plots, the black curve is
-the uncorrected time-series, and the red curve is the corrected time
-series. Samples close to a step are omitted in the corrected time-series.
-In the right hand plot, the blue curve is the uncorrected time-series for
-a nearby bolometer. The similarity between the red and blue curves shows
-that the step correction is performing well.
-}
+\caption{Examples of steps in time-series data. Steps occur with a
+wide range of heights from the large steps shown in (a) to the small
+steps shown in (b). Large steps are often followed by a brief
+``over-shoot'' as shown in (a) which is ringing caused by the
+anti-aliasing filter. In both plots, the black curve is the
+uncorrected time-series, and the red curve is the corrected time
+series. Samples close to a step are omitted in the corrected
+time-series.  In (b) the blue curve is the uncorrected time-series for
+a nearby bolometer. The similarity between the red and blue curves
+shows that the step correction is performing well.  }
 \label{fig:steps1}
@@
 steps occur over a single sample, such as the step close to sample
-5000 in the right-hand plot of Fig.~\ref{fig:steps1}, but others
-happen more gradually, such as the step close to sample 5300. In
-addition, a step can be preceded or followed by a short period of
-instability, as is visible at the bottom of the step in the left hand
-plot of Fig.~\ref{fig:steps1} (this is probably due to the response of
-the \scuba\ anti-aliasing filter; the sudden large step occurs prior
-to the 200\,Hz re-sampling). Further problems are caused by steps that
-occur close together in time, such as the large downward step followed
-by a smaller upward step close to sample 5000 in the right hand plot
-of Fig.~\ref{fig:steps1}.
+5000 in Fig.~\ref{fig:steps1}b, but others happen more gradually, such
+as the step close to sample 5300. In addition, a step can be preceded
+or followed by a short period of instability, as is visible at the
+bottom of the step in Fig.~\ref{fig:steps1}a (this is probably due to
+the response of the \scuba\ anti-aliasing filter; the sudden large
+step occurs prior to the 200\,Hz re-sampling). Further problems are
+caused by steps that occur close together in time, such as the large
+downward step followed by a smaller upward step close to sample 5000
+in Fig.~\ref{fig:steps1}b.
 
@@
 The primary purpose of the noise component, \model{NOI}, is to measure
-the white noise levels of each bolometer [$\mathbf{n}^\mathrm{w}_i(t)$
-in Eq.~\ref{eq:noise}]. The measurement generally occurs once all of
-the other models have been fit and removed. This noise may then be
-used to estimate weights for the data in subsequent iterations.
+the white noise levels of each bolometer, which is approximated with a
+single (non-time varying) variance, $\sigma_i^{\mathbf{w}2}$. This
+value is then used to calculate weights for map estimation,
+$\mathbf{w}_i = 1/\sigma_i^{\mathbf{w}2}$. The measurement generally
+occurs once all of the other models have been fit and removed [i.e.,
+once $\mathbf{n}^\mathrm{w}_i(t)$ from Eq.~\ref{eq:noise} is
+isolated].
 
@@
 
-The map-maker will halt if the convergence criteria have been
-met. Presently there are two numerical tests that may be performed by
-SMURF, or alternatively, a fixed number of iterations may be
-specified. The first numerical test is an approximate check of the
-change in reduced chi-squared, $\chi^2_\mathrm{r}$. The standard
-deviations, $\sigma_i$, are measured for each of the residual
-bolometer signals once the modelled signal components have been
-removed. In the early iterations, the residual signals usually contain
-both white noise, and other long-timescale features. However, once the
-solution has converged, this signal should look approximately
-white. Once the white noise, $\mathbf{n}^\mathrm{w}_i(t)$, has been
-measured from the bolometer PSDs in \model{NOI}
-(Section~\ref{sec:noi}), $\chi^2_\mathrm{r}$ is calculated as $(1/N)
-\sum_i \sigma^2_i / \mathbf{n}^\mathrm{w}_i(t)$, where $i$ runs over
-the $N$ bolometers. In other words, it is the average ratio between
-the measured time-series bolometer variances and their white-noise
-levels measured between 2 and 10\,Hz. Note that this expression does
-not account for the degrees of freedom. Clearly there are a large
-number of parameters in the model, which will ``fit-out'' some of the
-uncorrelated white noise, and potentially bias this estimate of
-$\chi^2_\mathrm{r}$ low. One approach would be to keep track of the
-degrees of freedom, and therefore provide a correction, as in
-\citet{kovacs2008}. However, note that we are only using this quantity
-as a convergence test, and its value is not important in an absolute
-sense. Furthermore, the reference white noise values are calculated
-only once in SMURF (Section~\ref{sec:noi}), and the number of model
-parameters are fixed, so the value of $\chi^2_\mathrm{r}$ should
-decrease monotonically if the solution is behaving well. The
-convergence criterion is met if the change in subsequent values of
-$\chi^2_\mathrm{r}$ is smaller than the requsted threshold, which is
-set to $10^{-3}$ by default.
-
-It was immediately found in the analysis of \scuba\ data that the
-$\chi^2_\mathrm{r}$ test described above is not a sufficient stopping
-criterion. The reason for this is that various components of the model
-are, under normal circumstances, highly degenerate. One simple example
-is the degeneracy between large-scale astronomical structures (larger
-than the array footprint), and the common-mode rejection step
-(\model{COM}), which will be illustrated in
-Section~\ref{sec:point}. Since the flatness of the residual only
-reflects the ability of the model to fit the shape of the bolometer
-time-series, it does not necessarily correlate with convergence of the
-most important model component: the map. Indeed, examination of the
-map after large numbers of iterations (e.g., $>10$), shows the
-presence of large-scale structures that are anti-correlated with
-structures in other model components. For this reason, a second
-map-based convergence test was added.
-
-The map-based convergence statistic, $M_\mathrm{c}$, is the average
-absolute change in the value of map pixels between subsequent
-iterations, normalised by the map pixel uncertainties (square root of
-Eq.~\ref{eq:varmap}), or
+The map-maker will halt after a user-specified number of iterations,
+or once some convergence criterion has been achieved. Presently two
+numerical quantities are tracked after each iteration: the change in
+reduced chi-squared, $\chi^2_\mathrm{r}$; and the change in the
+map. While convergence in $\chi^2_\mathrm{r}$ is not, ultimately, a
+good indicator of when the solution should be stopped, we discuss its
+calculation and properties first as an introduction to the problem of
+model degeneracies.
+
+The bolometer residual variances, $\sigma^2_i$, are measured once the
+modelled signal components have been removed. In the early iterations,
+the residuals contain both white noise, and other long-timescale
+features. However, when the solution has converged, this signal should
+look white. Taking the component of bolometer variances due to white
+noise, $\sigma_i^{\mathbf{w}2}$ (measured from the bolometer PSDs in
+\model{NOI}, Section~\ref{sec:noi}), $\chi^2_\mathrm{r}$ is calculated
+as $(1/N) \sum_i \sigma^2_i / \sigma_i^{\mathbf{w}2}$, where $i$ runs
+over the $N$ bolometers. In other words, it is the average ratio
+between the measured time-series bolometer variances and their
+white-noise levels measured between 2 and 10\,Hz, and should tend to a
+value of $\sim$1 if the low-frequency noise (and any bright
+astronomical signals) have been successfully removed. However, this
+expression does not account for the degrees of freedom in the
+model. Clearly there are a large number of parameters that will
+``fit-out'' some of the uncorrelated white noise, and bias this
+estimate of $\chi^2_\mathrm{r}$ low. Indeed, for a converged map,
+values near $\sim$0.85 are common, depending primarily on the fraction
+of the total bandwidth that has been removed by high-pass filtering
+(Section~\ref{sec:flt}). Regardless of the normalisation, once the
+model parameterisation has been chosen, this quantity should converge
+to some fixed value.
+
+Practically speaking, $\chi^2_\mathrm{r}$ tends to converge before the
+map due to model degeneracies. One simple example is the degeneracy
+between large-scale astronomical structures (larger than the array
+footprint), and the common-mode rejection step (\model{COM}), which
+will be illustrated in Section~\ref{sec:point}. In this case, there is
+a significant space of maps with spurious large-scale structures that
+are anti-correlated with features in the common-mode, and the solution
+is free to wander this space while producing flat residuals.
+
+Instead, we typically use a map-based convergence statistic,
+$M_\mathrm{c}$, the average absolute change in the value of map pixels
+between subsequent iterations, normalised by the map pixel
+uncertainties (square root of Eq.~\ref{eq:varmap}), or
 %
@@
 where $i$ runs over the $N$ map pixels, and $j$ enumerates the
-iterations. The convergence criterion is met once this average change
-is smaller than some threshold. Experimentally we found that a value
-of 0.05 gives good results (on average, map pixels change by 5\% of
-the estimated map RMS in subsequent iterations).  This provides good
-correspondence with what we would choose ``by eye'', and letting the
+iterations. Experimentally we found that a change in this quantity
+$<$0.05 (on average, map pixels change by $<$5\% of the estimated map
+RMS in subsequent iterations) is a good point to stop, providing
+correspondence with what we would choose ``by eye''. Letting the
 solver run for many more iterations in several test cases yields
@@
 
-A major source of divergence is correlation between \model{COM} and
-\model{FLT}. Since \model{FLT} usually consists of a high-pass filter
-following the application of \model{COM}, \model{COM} is completely
-free to grow any large-scale structure at frequencies below the chosen
-filter edge. While such structure does not appear in the map (as it is
-removed by \model{FLT}), we found that the solution could be made to
-converge significantly faster by ``re-mixing'' \model{COM} and
-\model{FLT} at the start of each iteration. In practice, at the start
-of each iteration, the previous iterations of each model are both
-added back into the residual immediately prior to re-calculating
-\model{COM}. In this way, truly common-mode signals, even at
-low-frequencies, do not leak into \model{FLT}.
-
-The most serious cause of divergence is the degeneracy between
-low-frequency signal that is removed, and large-scale structures in
-the map. For example, by subtracting the common-mode signal, the map
-is free to grow large-scale structures that exceed the angular scale
-of the array footprint. Our approach as been to constrain regions of
-the map devoid of sources to a value of zero for all but the final
-iteration in the \model{AST} model (Section~\ref{sec:ast}). Such
-regions are either user-defined in advance, or can be determined from
-the data using a cut on \snr. This technique is explored considerably
-in the examples from Sections~\ref{sec:point} and \ref{sec:extended}.
+Another major source of divergence is correlation between \model{COM}
+and \model{FLT}. Since \model{FLT} usually consists of a high-pass
+filter following the application of \model{COM}, \model{COM} is
+completely free to grow any large-scale structure at frequencies below
+the chosen filter edge. While such structure does not appear in the
+map (as it is removed by \model{FLT}), we found that the solution
+could be made to converge significantly faster by ``re-mixing''
+\model{COM} and \model{FLT}. Immediately prior to the calculation of
+\model{COM}, the values of \model{COM} and \model{FLT} from the
+previous iteraion are added back into the residual simultaneously. In
+this way, truly common-mode signals, even at low-frequencies, do not
+leak into \model{FLT}.
+
+In order to control the degeneracy between low-frequency signal that
+is removed, and large-scale structures in the map, we have developed a
+simple system for constraining regions of the map devoid of sources to
+a value of zero for all but the final iteration in the \model{AST}
+model (Section~\ref{sec:ast}). Such regions are either user-defined in
+advance, or can be determined from the data using a cut on \snr. This
+technique is explored considerably in the examples in
+Sections~\ref{sec:point} and \ref{sec:extended}.
 
@@
 
-In this section three different common types of data are reduced: a
-bright point source, Uranus (Section~\ref{sec:point}); a blind survey
-of high-redshift galaxies in the Lockman Hole
-(Section~\ref{sec:cosmo}); and a map of a bright extended star-forming
-region in our Galaxy, M17 (Section~\ref{sec:extended}). In each case,
-variations on the default algorithm from Section~\ref{sec:algorithm}
-are described.
+While most of the steps described in the Section~\ref{sec:algorithm}
+for reducing data with SMURF are applicable to all data sets, there is
+no ``default'' reduction that provides good results for a wide range
+of source \snr\ and angular scales. However, SMURF does have a small
+number of configurations with variations on these baseline parameters
+that are applicable to most common types of data. In this section we
+illustrate the differences bewteen these configurations with the
+following examples: a bright point source, Uranus
+(Section~\ref{sec:point}); a blind survey of high-redshift galaxies in
+the Lockman Hole (Section~\ref{sec:cosmo}); and a map of a bright
+extended star-forming region in our Galaxy, M17
+(Section~\ref{sec:extended}).
 
@@
 filter after the common-mode removal. A third reduction of the data
-uses the default map-making parameters, as described in
+uses the baseline map-making parameters, as described in
 Section~\ref{sec:algorithm}, which adds the \model{FLT} model to
@@
 employed. Large-scale diverging structures in the map must be
-mitigated, and the method used in this example (the default processing
-in SMURF) is to apply a high-pass filter to the data once, as a
-pre-processing step. The iterative solution is then run using only
-\model{COM}, \model{EXT}, \model{AST}, and \model{NOI}. In other
+mitigated, and the method used in this example (the baseline
+processing in SMURF) is to apply a high-pass filter to the data once,
+as a pre-processing step. The iterative solution is then run using
+only \model{COM}, \model{EXT}, \model{AST}, and \model{NOI}. In other
 words, there is no information in the bolometer signals below some
@@
   source (solid line) along the R.A. axis following map-making using
-  the default blank-field processing with SMURF (dotted line), and
-  upon application of the whitening filter (dashed line). The ``map
-  filtered'' response is produced by adding the ideal Gaussian to the
-  real data near the centre of the Lockman Hole map, and re-reducing
-  the data. The resulting dotted line gives the expected shape of a
-  source in panel (a) from Fig.~\ref{fig:lockman_maps}. Applying the
-  whitening filter (reciprocal of the solid orange line in
+  the blank-field processing configuration with SMURF (dotted line),
+  and upon application of the whitening filter (dashed line). The
+  ``map filtered'' response is produced by adding the ideal Gaussian
+  to the real data near the centre of the Lockman Hole map, and
+  re-reducing the data. The resulting dotted line gives the expected
+  shape of a source in panel (a) from
+  Fig.~\ref{fig:lockman_maps}. Applying the whitening filter
+  (reciprocal of the solid orange line in
   Fig.~\ref{fig:lockman_pspec}) to the dotted line gives the whitened
@@
 panel. Panels (a) and (b) show the results for a reduction using the
-default parameters (the solution halted after reaching the default
-map-based convergence criterion in 17 iterations). Panel (a) also
-depicts the array footprint (position angle indicative of the start of
-the observation), and a 300\,arcsec line shows the spatial scale
-corresponding to the default \model{FLT} high-pass filter. Similar to
+baseline parameters (the solution halted after reaching the map-based
+convergence criterion in 17 iterations). Panel (a) also depicts the
+array footprint (position angle indicative of the start of the
+observation), and a 300\,arcsec line shows the spatial scale
+corresponding to the \model{FLT} high-pass filter. Similar to
 Fig.~\ref{fig:pointmaps}c, the high-pass filtering introduces ringing
@@
 
-The default reduction of these data is shown in the top panels of
+The baseline reduction of these data is shown in the top panels of
 Fig.~\ref{fig:m17}, after 2 iterations (the first map estimated after
@@
 average of two halves of the data analysed independently, using the
-default configuration. The data have had added to them synthetic
+baseline configuration. The data have had added to them synthetic
 signal within a 600\,arcsec diameter circle (south of M17) created as
@@
 \centering
-\includegraphics[width=0.49\linewidth]{pspec_m17_default}
-\includegraphics[width=0.49\linewidth]{cor_pspec_m17_default}
+\begin{minipage}[h]{0.495\linewidth}
+\textbf{(a) baseline reduction PSD} \\
+
+\includegraphics[width=\linewidth]{pspec_m17_default}
+\end{minipage}
+\begin{minipage}[h]{0.495\linewidth}
+\textbf{(b) baseline reduction transfer function} \\
+
+\includegraphics[width=\linewidth]{cor_pspec_m17_default}
+\end{minipage}
 \caption{Angular PSDs for the region of the M17 map in
-Fig.~\ref{fig:m17_jk} containing synthetic data, using the default
-configuration. Left: raw PSDs for the input (noiseless) simulated data
+Fig.~\ref{fig:m17_jk} containing synthetic data, using the baseline
+configuration. (a) Raw PSDs for the input (noiseless) simulated data
 (thick black line), the signal (average of each half) PSDs (thin solid
@@
 150\,arcsec (red); 300\,arcsec (orange); 600\,arcsec (green); and
-900\,arcsec (blue). PSDs are also colour-coded by filter scale. Right:
-since the input PSD is known, it is possible to measure the transfer
+900\,arcsec (blue). PSDs are also colour-coded by filter scale. (b)
+Since the input PSD is known, it is possible to measure the transfer
 function of the map-maker as the ratio between the difference of the
@@
 the thin coloured lines (linear vertical axis shown on right of
-plot). The remaining lines are as in the left panel, but now corrected
-by the transfer function. This plot shows that increasing the filter
-scale improves the \snr\ at intermediates scales
-($\sim$200--600\,arcsec), although the \snr\ worsens at smaller scales
-($\lsim$200\,arcsec).}
+plot). The remaining lines are as in (a), but now corrected by the
+transfer function. This plot shows that increasing the filter scale
+improves the \snr\ at intermediates scales ($\sim$200--600\,arcsec),
+although the \snr\ worsens at smaller scales ($\lsim$200\,arcsec).}
 \label{fig:m17_def_ps}
@@
 \centering
-\includegraphics[width=0.49\linewidth]{pspec_m17_bright_extended}
-\includegraphics[width=0.49\linewidth]{cor_pspec_m17_bright_extended}
+\begin{minipage}[h]{0.495\linewidth}
+\textbf{(a) bright extended reduction PSD} \\
+
+\includegraphics[width=\linewidth]{pspec_m17_bright_extended}
+\end{minipage}
+\begin{minipage}[h]{0.495\linewidth}
+\textbf{(b) bright extended reduction transfer function} \\
+
+\includegraphics[width=\linewidth]{cor_pspec_m17_bright_extended}
+\end{minipage}
 \caption{Lines have same meaning as in Fig.~\ref{fig:m17_def_ps},
@@
 angular scales. Also note that increasing the filter scale does not
-have as large an impact on the small-scale noise as in the default
+have as large an impact on the small-scale noise as in the baseline
 configuration, although it is still noticeably larger up to scales of
@@
 averaging the maps made of each independent half of the data, for the
-default configuration (inverse-variance weighting has been used). The
+baseline configuration (inverse-variance weighting has been used). The
 columns show reductions using 150, 300, 600, and 900\,arcsec filter
@@
 does a generally good job of suppressing the largest-scale ripples
-that are produced by the default reduction. However, the noise away
+that are produced by the baseline reduction. However, the noise away
 from regions of bright emission does increase noticeably (mottled
@@
 maps to understand the signal and noise properties of the map-maker in
-the region of synthetic sources, as a function of filter scale. The
-left panel of Fig.~\ref{fig:m17_def_ps} shows the raw PSDs for the
-input synthetic signal (thick black line), the output map signals
-(thin solid lines), and the jackknife maps (dashed lines). Colours
-encode the filter scales used: 150\,arcsec (red); 300\,arcsec
-(orange); 600\,arcsec (green); and 900\,arcsec (blue). Note that, with
-the exception of the synthetic data, we only plot the PSDs down to the
+the region of synthetic sources, as a function of filter scale. In
+Fig.~\ref{fig:m17_def_ps}a we show the raw PSDs for the input
+synthetic signal (thick black line), the output map signals (thin
+solid lines), and the jackknife maps (dashed lines). Colours encode
+the filter scales used: 150\,arcsec (red); 300\,arcsec (orange);
+600\,arcsec (green); and 900\,arcsec (blue). Note that, with the
+exception of the synthetic data, we only plot the PSDs down to the
 second-lowest spatial frequency bin of
@@
 jackknife PSDs by dividing by $G(k)$ to produce the thick solid, and
-dashed lines in the right panel of Fig.~\ref{fig:m17_def_ps}. This
-shows us that, even though the raw noise in the left panel is lower at
-small scales when a smaller-scale filter is used (e.g., the red dashed
+dashed lines in Fig.~\ref{fig:m17_def_ps}b. This shows us that, even
+though the raw noise in Fig.~\ref{fig:m17_def_ps}a is lower at small
+scales when a smaller-scale filter is used (e.g., the red dashed
 line), once we correct for the transfer function, we actually win in a
@@
 Fig.~\ref{fig:m17_be_ps}. The most obvious improvement with this
-reduction over the default reduction is that the transfer functions
+reduction over the baseline reduction is that the transfer functions
 fall more slowly at large angular scales, accompanied by a slower
@@
 in small-scale noise as the filter edge is increased is not as drastic
-as in the default reduction, and the \snr\ improves at scales
+as in the baseline reduction, and the \snr\ improves at scales
 $\gsim$60\,arcsec.
@@
 helpful feedback, especially David Nutter and Todd McKenzie; and
-Mandana Amiri for consultations regarding the SCUBA-2 Multi-Channel
-Electronics.  We acknowledge the contributions of Dennis Kelly, Alex
-van Engelen and Jennifer Balfour for early investigations related to
-SMURF; and Mark Thompson, Craig Walther and S\'{e}verin Gaudet for
-being on the Critical Design Review panel. Finally, we thank Per
-Friberg and Gary Davis for their helpful comments on the manuscript.
+Mandana Amiri and Dan Bintley for consultations regarding the SCUBA-2
+Multi-Channel Electronics.  We acknowledge the contributions of Dennis
+Kelly, Alex van Engelen and Jennifer Balfour for early investigations
+related to SMURF; and Mark Thompson, Craig Walther and S\'{e}verin
+Gaudet for being on the Critical Design Review panel. Finally, we
+thank Per Friberg and Gary Davis for their helpful comments on the
+manuscript.
 

### 2012-11-22 smurf_paper.tex
more explanation about why zeromasking works. fixed m17 figs

@@
 \includegraphics[width=\linewidth]{com}
-\caption{A selection of typical bolometer time-series after
-flatfielding and removal of a constant baseline. Each sub-plot shows
-the power from a single bolometer over a 30 second period.
-The lower and upper vertical limits of each plot are -0.02 pW and +0.02
-pW respectively. It can be seen that most bolometers exhibit a common
-time variation overlaid with other features. Empty squares indicate the
-locations of broken bolometers.}
+\caption{A selection of bolometer time-series from a representative
+region of the focal plane after flatfielding and removal of a constant
+baseline. Each sub-plot shows the power from a single bolometer over a
+30\,s interval.  The lower and upper vertical limits of each plot are
+$-0.02$\,pW and $+0.02$\,pW, respectively. It can be seen that most
+bolometers exhibit a common time variation overlaid with other
+features. Empty squares indicate the locations of broken bolometers.}
 \label{fig:com}
@@
 estimate of $\chi^2_\mathrm{r}$ low. Indeed, for a converged map,
-values near $\sim$0.85 are common, depending primarily on the fraction
-of the total bandwidth that has been removed by high-pass filtering
-(Section~\ref{sec:flt}). Regardless of the normalisation, once the
-model parameterisation has been chosen, this quantity should converge
-to some fixed value.
+values in the range $\sim$0.85--0.95 are common, depending on the
+precise configuration parameters. Regardless of the normalisation,
+once the model parameterisation has been chosen, this quantity should
+converge to some fixed value.
 
@@
 
-While most of the steps described in the Section~\ref{sec:algorithm}
-for reducing data with SMURF are applicable to all data sets, there is
-no ``default'' reduction that provides good results for a wide range
-of source \snr\ and angular scales. However, SMURF does have a small
+While most of the steps described in Section~\ref{sec:algorithm} for
+reducing data with SMURF are applicable to all data sets, there is no
+``default'' reduction that provides good results for a wide range of
+source \snr\ and angular scales. However, SMURF does have a small
 number of configurations with variations on these baseline parameters
@@
 appropriate for performing aperture photometry directly, with no need
-for an additional reference annulus. This approach to constrained
-map-making is similar to that employed for poorly cross-linked scans
-of compact (though resolved) sources by \citet{wiebe2009} using BLAST
-data.
+for an additional reference annulus.
+
+The reason this prior works is simple to understand if the bolometer
+data are thought of as providing useful differential measurements up
+to scales corresponding to the filter edge. In this example, the
+bolometers retain information on brightness differences up to
+200\,arcsec scales. Since the map is constrained to zero to within
+60\,arcsec of the peak of Uranus (well within 200\,arcsec), the
+solution is then able to accurately reconstruct the differential peak
+intensity of Uranus with respect to this constrained background.  This
+approach to map-making is similar to that employed for poorly
+cross-linked scans of compact (though resolved) sources by
+\citet{wiebe2009} using BLAST data.
 
@@
 transfer function. This plot shows that increasing the filter scale
-improves the \snr\ at intermediates scales ($\sim$200--600\,arcsec),
+improves the \snr\ at intermediate scales ($\sim$200--600\,arcsec),
 although the \snr\ worsens at smaller scales ($\lsim$200\,arcsec).}

### 2012-11-29 smurf_paper.tex
updates about lags in the bolo data, PCA section, and sky noise

@@
   reduction on single, modern, high-end desktop computers, with
-  execution times that are shorter than the observing times.  SMURF is
-  used in an automated setting, both at the telescope for real-time
-  feedback to observers, as well as for the production of science
-  products for the JCMT Science Archive at the Canadian Astronomy Data
-  Centre. Three detailed case studies are used to: (i) explore
-  convergence properties of the map-maker using simple prior
+  execution times that are typically shorter than the observing times.
+  SMURF is used in an automated setting, both at the telescope for
+  real-time feedback to observers, as well as for the production of
+  science products for the JCMT Science Archive at the Canadian
+  Astronomy Data Centre. Three detailed case studies are used to: (i)
+  explore convergence properties of the map-maker using simple prior
   constraints (Uranus -- a point source); (ii) achieve the white-noise
@@
 re-sampling stage. This anti-aliasing filter also introduces an
-effective lag of $\sim7$\,ms (which must be added to the 1\,ms thermal
-time constant of the bolometers mentioned above to obtain the total
-lag). There are also line features which are thought to be produced at
-frequencies far above the final 200\,Hz sample rate, and are aliased
-to lower frequencies during the multiplexed readout stage before they
-can be removed by the anti-aliasing filter.
+effective lag of $\sim$6\,ms. Adding this lag to the $\sim$1\,ms
+thermal time constant, as well as a negative shift of about $-$2.5\,ms
+caused by a half-sample offset due to the synchronization between the
+pointing and bolometer data, yields a net lag of approximately
+4.5\,ms. There are also line features in the spectra that are thought
+to be produced primarily at frequencies far above the final 200\,Hz
+sample rate, and are aliased to lower frequencies during the
+multiplexed readout stage before they can be removed by the
+anti-aliasing filter.
 
@@
   the 450 and 850\,\micron\ bands with the mixing chamber temperature,
-  the line-of-sight 450\,\micron\ opacity derived from the water
+  the line-of-sight 450\,\micron\ opacity derived from the JCMT water
   vapour monitor, and azimuth/elevation pointing offsets, before and
@@
 the focal plane itself, it is certainly correlated with the
-temperature of the detectors), and the telescope pointing, for two
-data sets, before and after the upgrades that followed S2SRO.
+temperature of the detectors), the line-of-sight atmoshperic opacity
+as derived from the JCMT water vapour monitor (WVM), and the telescope
+pointing, for two data sets, before and after the upgrades that
+followed S2SRO.
 
@@
 base temperature, although the similarity is clearly greater at
-850\,\micron. In this particular case, the total power in the
-fluctuations at 450\,\micron, are comparable to those at
-850\,\micron. Such behaviour might be expected if there is a
-comparable varying thermal load from the fridge at each wavelength
-that dominates. We also note that there is no obvious strong
-correlation between the low-frequency signal structure, at either
-wavelength, with the telescope motion.
+850\,\micron. Note that the total power in the fluctuations at
+450\,\micron\ are comparable to those at 850\,\micron\ as one might
+expect if there is a comparable varying thermal load from the fridge
+at each wavelength that dominates. We also note that there is no
+obvious strong correlation between the low-frequency signal structure,
+at either wavelength, with the telescope motion. However, there is a
+suggestion that the shorter-timescale behaviour of the opacity is
+anti-correlated with the elevation, as expected.
 
@@
 reflect changes in the focal plane temperature). There is minimal
-correlated low-frequency noise in these 450\,\micron\ data (note that
-there is almost no difference once the common-mode has been
-subtracted). The 850\,\micron\ channel, however, exhibits a more
-significant signal that is obviously correlated with the telescope
-motion. It is also correlated amongst many of the detectors, and
-common-mode removal corrects it to a large extent. Part of the reason
-that this is seen here, and not in the S2SRO data set shown, is that
-the amplitude of the scan pattern is larger. Most of this
-``scan-synchronous'' noise is attributed to magnetic field pickup, as
-described in Section~\ref{sec:magpickup}.
+correlated low-frequency noise in these 450\,\micron\ data, resulting
+in little difference between the raw and common-mode subtracted
+data. The 850\,\micron\ channel, however, exhibits a more significant
+signal that is obviously correlated with the telescope motion. It is
+also correlated amongst many of the detectors, and common-mode removal
+corrects it to a large extent. Part of the reason that this is seen
+here, and not in the S2SRO data set shown, is that the amplitude of
+the scan pattern is larger. Most of this ``scan-synchronous'' noise is
+attributed to magnetic field pickup, as described in
+Section~\ref{sec:magpickup}. Also similar to the S2SRO data, an
+anti-correlation between the opacity and elevation is apparent.
 
@@
 corresponding eigenvalues. The $\mathbf{\xi}_i(t)$ are normalised by
-$[\sum_t \mathbf{\xi}_i^2(t)]^{1/2}$, such that the root mean square
-(\rms) amplitude of each component in a given bolometer signal is
-stored in the eigenvalues. In the earlier analyses mentioned, the
-low-frequency noise is assumed to be encapsulated in those components
-with the largest eigenvalues. Removing the projection of the time
-series along those components then significantly reduces $1/f$ noise
-while retaining most of the (higher-frequency) signal in point-like
-sources
+$[\sum_t \mathbf{\xi}_i^2(t)]^{1/2}$, such that the amplitude of each
+component in a given bolometer signal is stored in the eigenvalues. In
+the earlier analyses mentioned, the low-frequency noise is assumed to
+be encapsulated in those components with the largest
+eigenvalues. Removing the projection of the time series along those
+components then significantly reduces $1/f$ noise while retaining most
+of the (higher-frequency) signal in point-like sources
 
@@
 the same observation that was used in Figs.~\ref{fig:bolos_mix}b and
-\ref{fig:pspec}b. The 12 most significant components are shown (ranked
+\ref{fig:pspec}b. The 9 most significant components are shown (ranked
 by the mean eigenvalues across all working bolometers in both
@@
 The majority of the correlated signal at both wavelengths in
-Fig.~\ref{fig:pca} is clearly produced by Components~1 and 2. In both
-cases, the eigenvector time-series exhibit a roughly periodic signal
-that resembles the scan pattern in Fig.~\ref{fig:bolos_mix}b. While
-these correlated signals also clearly appear at both wavelengths,
-referring to the maps of eigenvalues, they are considerably stronger
-at 850\,\micron\ (consistent with the visual appearance of the
-bolometer signals in Fig.~\ref{fig:pca}). Also note that the PSDs for
-these eigenvectors exhibit nearly a pure $1/f$ signature with almost
-no line features, or a white-noise plateau (suggesting that these are
-high-\snr\ measurements of a purely low-frequency drift).
+Fig.~\ref{fig:pca} is produced by Component~1, in which the
+eigenvector time-series exhibits a roughly periodic signal that
+resembles the scan pattern in Fig.~\ref{fig:bolos_mix}b. While this
+correlated signal appears at both wavelengths, referring to the maps
+of eigenvalues, they are considerably stronger at 850\,\micron\
+(consistent with the visual appearance of the bolometer signals in
+Fig.~\ref{fig:pca}). Also note that the PSD for theis eigenvector
+exhibit nearly a pure $1/f$ signature with minimal line features, or a
+white-noise plateau (suggesting that these are high-\snr\ measurements
+of a purely low-frequency drift). Component~3 is weaker, yet also
+shows evidence of the scan-synchronous noise, combined with
+high-frequency line features that appear primarily in columns 26, 29
+and 30 of the 450\,\micron\ subarray. Unlike Component~1, it is
+generally more dominant at 450\,\micron.
 
@@
 
-Why are there two dominant components? First, the smoothly-varying
-pattern in the eigenvalue maps (particularly evident at 850\,\micron)
-suggests that the scan-synchronous noise (presumably magnetic field
-pickup) has a different response across the focal plane; in this case
-the PCA may have identified two orthogonal shapes that, when mixed in
-different quantities, can reproduce most of the signal in each
-bolometer. It is also likely that a varying atmospheric contribution
-is also included in these eigenvectors. One would expect the
-atmosphere to be more significant at 450 than at 850\,\micron,
-although it is not obviously visible due to the strong presence of the
-scan-synchronous noise.
-
-Next, Components~3, 4, 7 and 11 in Fig.~\ref{fig:pca} have a common
-character. In all cases, the PCA appears to have identified strong,
-but transient glitches that appear in only one or several bolometers,
-such as the two black bolometers near column 8, row 21 at
-850\,\micron\ in Component~3. Such events are typically flagged and
-removed during iterative de-spiking (Section~\ref{sec:ast}).
-
-The most interesting feature of Component~5 in Fig.~\ref{fig:pca} is a
-strong narrow line in the PSD at a frequency of about 57\,Hz
-(suggesting the 60\,Hz mains as the culprit). The eigenvalue maps show
-that it is only significant for several bolometers in the
-850\,\micron\ subarray: at column 22 row 22; several bolometers at the
-bottom of column 28; and more near the top-right of the subarray.
-
-Components~6 and 8--10 in Fig.~\ref{fig:pca} have several generic
-features in common. First, the PSDs exhibit moderate low-frequency
-$1/f$ noise. Second, there are broad-line features at frequencies
-above $\sim$30\,Hz which we believe to be aliased high-frequency
-noise, in part due to the clear correlation patterns that appear along
-columns in the eigenvalue maps at both wavelengths (e.g., columns 26
-and 29 at 450\,\micron\ in Component~6). Finally, these components
-have potentially unrelated features mixed into them, like sudden
-offsets that are likely residuals due to imperfect step correction
-(e.g., around $t\sim50$\,s in the eigenvector time-series for
-Component~6; also see Section~\ref{sec:steps}).
-
-Finally, Component~12 in Fig.~\ref{fig:pca} is significant in only a
-handful of detectors in the 850\,\micron\ subarray (e.g., column 21
-row 25). The time-series plot for the eigenvector exhibits a large
-positive tail of sudden excursions, which gradually worsen over
-time. Bolometers that exhibit this signal are probably poorly biased,
-and are usually flagged due to elevated noise levels during
-pre-processing (Section~\ref{sec:flagbadbol}).
+Why are there two components that seem to be dominated by
+scan-synchronous noise (which is presumably a mixture of
+magnetic-field pickup and atmospheric noise)? First, the
+smoothly-varying gradients in the eigenvalue maps (primarily at
+850\,\micron) suggests that this noise has a different response across
+the focal plane; in this case the PCA has identified two orthogonal
+shapes that, when mixed in different quantities, can reproduce most of
+the signal in each bolometer. It is also likely that the atmospheric
+contribution is stronger at 450\,\micron\ (i.e., predominantly
+Component~3).
+
+Like Componen~1, Component~2 is a purely $1/f$ signal, with an
+amplitude comparable to Component~3, that appears almost exclusively
+at 850\,\micron. Since it is not obviously correlated with the
+telescope motion, yet has an amplitude that varies smoothly across the
+850\,\micron\ subarray, it may represent the PCAs attempt to reproduce
+some non-linear response to magnetic-field and/or atmospheric pickup
+
+The remaining, weaker, Components~4--9 share some common
+features. They all have some degree of $1/f$ noise, suggesting further
+residual noise related to the telescope motion and/or other slow
+baseline drifts. All of them also exhibit line features in their
+spectra, which in many cases appear at random amplitudes along columns
+(e.g. columns 15, and 26--30 at 450\,\micron\ in Component~4), or in a
+sampling of isolated bolometers (e.g. colummn 13, row 19 at
+850\,\micron\ in Component~6). The signals that appear along columns
+is probably related to magnetic-field pickup. Other sources of lines
+are thought to be aliased high-frequency noise, and also the 60\,Hz
+mains which appear in some bolometers.
 
@@
 the details can vary significantly from data set to data set, and the
-lengths of the time-series analysed. The reason for this is probably
-due to the fact that many of the noise sources are not stationary in
-time (e.g., scan-synchronous noise which obviously depends on the scan
-pattern, and the tuning of the subarrays). It should also be clear
-from this example that while PCA offers some helpful insight into the
-various sources of noise, it does not necessarily identify clean
-patterns that can easily be modelled. For example, while there are
-clearly correlation patterns along columns (Components~6 and 8--10,
-particularly at 450\,\micron), as one might expect given the common
+lengths of the time-series analysed. The reason for this is that many
+of the noise sources are not stationary in time (e.g.,
+scan-synchronous noise which obviously depends on the scan pattern,
+and the tuning of the subarrays). It should also be clear from this
+example that while PCA offers some helpful insight into the various
+sources of noise, it does not necessarily identify clean patterns that
+can easily be modelled. For example, while there are clearly
+correlation patterns along columns as expected from the common
 amplification chain, the intensities of these high-frequency signals
-appear random in the eigenvalue maps, and a common-mode signal
-estimated from the data for each column would not remove it. Finally,
-we note that most of the high-frequency ($\gsim2$\,Hz) correlated
-features that have been identified tend to be restricted to a small
-number of detectors, and can often be identified and removed using
-other (faster) means. What remains is predominantly low-frequency
-noise. This conclusion is central to our data reduction strategy
-described in the next section.
+appear random in the eigenvalue maps, and a simple common-mode signal
+estimated from the data for each column would not remove
+it. Regardless, the dominant noise, in all cases, is at
+low-frequencies. This conclusion is central to our data reduction
+strategy described in Section~\ref{sec:algorithm}.
+
+\subsection{Overall performance and mapping speed}
 
@@
 The extinction correction is a multiplicative factor that is normally
-derived using the JCMT water vapour monitor (WVM), and is not
-considered to be a free parameter in the solution [$\mathbf{e}_i(t)$
-in Eq.~\ref{eq:model}]. However, it is applied as part of the
-iterative solution, rather than a pre-processing step, since any small
-errors will be amplified by the large low-frequency drifts in the raw
-bolometer time-series. For example, if the bolometer drift is 1000
-times greater than an astronomical source of interest, a 1 per cent
-error in the flatfield will produce stripes of order 10 times the
-astronomical signal amplitude in the final map! If \model{EXT} is
-applied after the bulk of the low-frequency noise has been removed
-(e.g., \model{COM,GAI}), then there is little potential for such small
-errors to affect the final map.  For details on how it is calculated
-see \citet{dempsey2012}. Note that its numerical value is calculated
-only once, and simply applied as a scale factor in the iterative
+derived using the WVM, and is not considered to be a free parameter in
+the solution [$\mathbf{e}_i(t)$ in Eq.~\ref{eq:model}]. However, it is
+applied as part of the iterative solution, rather than a
+pre-processing step, since any small errors will be amplified by the
+large low-frequency drifts in the raw bolometer time-series. For
+example, if the bolometer drift is 1000 times greater than an
+astronomical source of interest, a 1 per cent error in the flatfield
+will produce stripes of order 10 times the astronomical signal
+amplitude in the final map! If \model{EXT} is applied after the bulk
+of the low-frequency noise has been removed (e.g., \model{COM,GAI}),
+then there is little potential for such small errors to affect the
+final map.  For details on how it is calculated see
+\citet{dempsey2012}. Note that its numerical value is calculated only
+once, and simply applied as a scale factor in the iterative
 solution. Unlike the additive model components, it is inverted at the

### 2012-11-29 smurf_paper.tex
add newer/better reference for BoA

@@
 \citep[CRUSH,][]{kovacs2008} for the Submillimeter High-Angular
-Resolution Camera 2 (SHARC-2), the Bolometer array data Analysis
-software (BoA) for the Large APEX BOlometer CAmera
-\citep[LABOCA,][]{siringo2009}, and the pipeline developed for the
-Bolocam Galactic Plane Survey \citep{aguirre2011}.
+Resolution Camera 2 (SHARC-2), the pipeline developed for the Bolocam
+Galactic Plane Survey \citep{aguirre2011}, and the Bolometer array
+data Analysis software (BoA) for the Large APEX BOlometer CAmera
+\citep[LABOCA,][]{schuller2012}.
 
@@
 correlated noise components from the data, successively improving the
-map \citep[e.g.,][]{kovacs2008,siringo2009,aguirre2011}, we have shown
+map \citep[e.g.,][]{kovacs2008,aguirre2011,schuller2012}, we have shown
 that a relatively simpler approach provides good results for \scuba\

### 2012-11-29 smurf_paper.tex
add section about lags to pre-processing

@@
 to be produced primarily at frequencies far above the final 200\,Hz
-sample rate, and are aliased to lower frequencies during the
+sample rate, which are aliased to lower frequencies during the
 multiplexed readout stage before they can be removed by the
@@
 
+\subsubsection{Impact of lags}
+\label{sec:lags}
+
+In principle, the effects of the bolometer thermal response,
+anti-aliasing filter, and data acquisition lags should be accounted
+for; together they produce a net lag in the bolometer signals of about
+4.5\,ms (Section~\ref{sec:bolos}). This can be modelled by convolving
+the ``pure'' bolometer signals with a single system response
+function. Similarly, this effect could be removed using de-convolution
+as a pre-processing step. The impact of this lag is primarily a
+reduction in point-source sensitivity since scans at different
+position angles will detect the source at slightly different positions
+(and therefore reduce the peak signal once averaged together). This
+attenuation is most significant at 450\,\micron\ given the smaller
+beam size, and at the greatest scan speeds. For reference, at a speed
+of 150\,arcsec\,s$^{-1}$ (typical for CV daisies of calibrators and
+deep fields) the attenuation is a negligible $<$1\% at 850\,\micron,
+and $\sim$1\% at 450\,\micron. However, at the maximum speed of
+600\,arcsec\,s$^{-1}$ (for the largest, and typically shallowest
+maps), this attenuation grows to about 5\% and 15\% at 850 and
+450\,\micron, respectively. It is unlikely that this response has had
+a significant impact on observations taken to date, since such scans
+have been used primarily for shallow cosmology fields for which only
+the 850\,\micron\ data are expected to be useful, or Galactic fields
+of extended structures (which are less sensitive to this effect). In
+the present Starlink \textsc{kapuahi} release there is no option to
+de-convolve the system response. However, it is presently being added
+and should be present in future releases.
+
 %-------------------------------------------------
@@
 $\mathbf{n}^\mathrm{c}(t)$.
-
-In principle there are two other effects that could be included in
-this model. First, signals detected by the bolometers (inside the
-brackets of Eq.~\ref{eq:model}) should be convolved with the thermal
-response function, $\exp(-t/\tau)$, where $\tau=1$\,ms. Second, the
-final signals should be convolved with the response of the
-anti-aliasing filter. In practice, the combined effect of these two
-filters is to introduce a lag of $\sim$7\,ms, and only minimal
-attentuation of a point source (at most 3\% for the smaller
-450\,\micron\ PSF at the maximum scan speed of 600\,arcsec). For all
-of the maps described in this paper, their impact is therefore
-negligible. The current version of SMURF (part of the Starlink
-\textsc{kapuahi} release) does not attempt to de-convolve these
-effects, although the next version will likely include an optional
-pre-processing step to perform this task.
 
@@
 
-For maps of faint point sources, a single (non-iterative) high-pass
+For maps of faint point-sources, a single (non-iterative) high-pass
 filter at the start of the reduction produces maps that are nearly

### 2012-11-30 smurf_paper.tex
update text and plots for point source with extinction correction fix

@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=\linewidth]{pointmaps}
+\caption{Multiple reductions of a 450\,\micron\ CV daisy scan of
+  Uranus, all scaled between $-$0.002\,pW (white) and +0.002\,pW
+  (black). In all cases extinction correction has been applied
+  immediately after common-mode removal. (a) Reduction in which only
+  common-mode subtraction is used to suppress low-frequency noise, and
+  the reduction is forced to use 2 iterations (after the first
+  iteration an estimate of the source flux is removed, and the noise
+  is measured in the residuals to obtain appropriate weighting for the
+  second and final iteration). Circular streaks are caused by
+  independent low-frequency noise that is not removed by the
+  common-mode (Uranus peak 0.358\,pW). (b) Same as (a) but using 100
+  iterations, illustrating the degeneracy between large-scale
+  structure and the common-mode (the footprint of working bolometers
+  is also shown for reference, Uranus peak 0.358\,pW). (c) Reduction
+  in which high-pass filtering above 0.775\,Hz (corresponding to a
+  spatial scale of 200\,arcsec, as indicated) is applied after
+  common-mode removal, but before the map estimate. The map-based
+  convergence test is activated and the solution halts after 10
+  iterations, but leaving large-scale ringing due to the filter
+  (Uranus peak 0.353\,pW). (d) Same as (c), but now the region of the
+  map beyond the white dot-dashed circle is constrained to a value of
+  zero until all but the final iteration.  The map is now extremely
+  flat, there is no attenuation of the source flux compared with the
+  first two reductions, and the diffraction pattern is clearly seen
+  (Uranus peak 0.358\,pW).  }
+\label{fig:pointmaps}
+\end{figure*}
+
 The accurate measurement of positions and brightnesses of known point
@@
 (absolute calibration), and hence noise performance of the instrument
-in astronomically-useful units.  In this example we produce a
+in astronomically-useful units.  In this example we reduce a
 450\,\micron\ map of Uranus (observation 26 on 20111017), which is a
@@
 and final iteration. The resulting map after these two iterations is
-shown in Fig.~\ref{fig:pointmaps}a. While the \snr\ of Uranus is
-clearly large ($\sim$250), enabling us to see the faint sidelobes
+shown in Fig.~\ref{fig:pointmaps}a. While the peak \snr\ of Uranus is
+clearly large ($\sim$160), enabling us to see the faint sidelobes
 (circle and cross pattern), the map also has obvious circular streaks
@@
 bolometer after 2 (black) and 100 (grey) iterations, which are nearly
-identical, yet the change in the estimated \model{COM} (green) and
-\model{AST} (red) signals between 2 and 100 iterations are
-large. However, it is also clear that the estimated \model{COM} and
-\model{AST} signals are complimentary. In other words, the large
-change in the \model{AST} signal is cancelled by freedom in the
-\model{COM} signal to grow with opposite sign. For comparison, the
+identical, yet the change in the estimated
+\model{COM}\footnote{\model{COM} has been multiplied by the
+time-varying extinction correction to enable direct comparison with
+\model{AST}.} (green) and \model{AST} (red) signals between 2 and 100
+iterations are large However, it is also clear that the estimated
+\model{COM} and \model{AST} signals are complimentary. In other words,
+the large change in the \model{AST} signal is cancelled by freedom in
+the \model{COM} signal to grow with opposite sign. For comparison, the
 bottom panel of Fig.~\ref{fig:degeneracy} shows the telescope pointing
@@
 centre, the result is an azimuthally-symmetric $\sinc$ function-like
-pattern in the map. The gradient resulting from the degeneracy between
-the common-mode and map has been removed because the filter scale of
+pattern in the map. The peak flux has also been attenuated to
+0.353\,pW in this reduction, compared with 0.358\,pW in the first two
+reductions. The gradient resulting from the degeneracy between the
+common-mode and map has been removed because the filter scale of
 200\,arcsec (3.3\,arcmin) is smaller than the array field-of-view
@@
 final iteration. In this case, the map converges after 6 iterations,
-and the ringing has been effectively removed. Since the map is now
-flat away from the source, and constrained to a value of zero, it is
-appropriate for performing aperture photometry directly, with no need
-for an additional reference annulus.
-
-The reason this prior works is simple to understand if the bolometer
-data are thought of as providing useful differential measurements up
-to scales corresponding to the filter edge. In this example, the
-bolometers retain information on brightness differences up to
-200\,arcsec scales. Since the map is constrained to zero to within
-60\,arcsec of the peak of Uranus (well within 200\,arcsec), the
-solution is then able to accurately reconstruct the differential peak
-intensity of Uranus with respect to this constrained background.  This
-approach to map-making is similar to that employed for poorly
-cross-linked scans of compact (though resolved) sources by
-\citet{wiebe2009} using BLAST data.
-
-\begin{figure*}
-\centering
-\includegraphics[width=\linewidth]{pointmaps}
-\caption{Multiple reductions of a 450\,\micron\ CV daisy scan of
-  Uranus, all scaled between $-$0.002\,pW (white) and +0.002\,pW
-  (black). (a) Reduction in which only common-mode subtraction is used
-  to suppress low-frequency noise, and the reduction is forced to use
-  2 iterations (after the first iteration an estimate of the source
-  flux is removed, and the noise is measured in the residuals to
-  obtain appropriate weighting for the second and final
-  iteration). Circular streaks are caused by independent low-frequency
-  noise that is not removed by the common-mode (Uranus peak
-  0.15\,pW). (b) Same as (a) but using 100 iterations, illustrating the
-  degeneracy between large-scale structure and the common-mode (the
-  footprint of working bolometers is also shown for reference, Uranus
-  peak 0.15\,pW). (c) Reduction in which high-pass filtering above
-  0.775\,Hz (corresponding to a spatial scale of 200\,arcsec, as
-  indicated) is applied after common-mode removal, but before the map
-  estimate. The map-based convergence test is activated and the
-  solution halts after 10 iterations, but leaving large-scale ringing
-  due to the filter (Uranus peak 0.27\,pW). (d) Same as (c), but now
-  the region of the map beyond the white dot-dashed circle is
-  constrained to a value of zero until all but the final iteration.
-  The map is now extremely flat, there is little attenuation of the
-  source flux, and the diffraction pattern is clearly seen (Uranus
-  peak 0.27\,pW).}
-\label{fig:pointmaps}
-\end{figure*}
+and the ringing has been effectively removed. The attenuation from the
+previous reduction is now removed, and Uranus again has a peak value
+of 0.358\,pW. For reference, the Uranus model distributed with
+Starlink predicts a peak brightness of 176\,Jy on this date, yielding
+an FCF of 492 Jy\,pW$^{-1}$, which is well within the typical range
+for \scuba\ of $(492 \pm 67)$\,Jy\,pW$^{-1}$ \citep[see][for
+details]{dempsey2012}. Since the map is now flat away from the source,
+and constrained to a value of zero, it is appropriate for performing
+aperture photometry directly, with no need for an additional reference
+annulus.
+
+The way this prior works can be understood from the point of view of
+differential measurements. Bolometer data contain information up to
+scales corresponding to the filter edge (or the scale of the array
+footprint, whichever is smallest). In this example, the relevant scale
+is 200\,arcsec. Since the map is constrained to zero within 60\,arcsec
+of the peak of Uranus (well within 200\,arcsec), the solution is able
+to accurately reconstruct the differential peak intensity of Uranus
+with respect to this constrained background. This approach to
+map-making is similar to that employed for poorly cross-linked scans
+of compact (though resolved) sources by \citet{wiebe2009} using BLAST
+data.
 

### 2012-12-01 smurf_paper.tex
sky noise and clarification of whitened PSF

@@
 here, and not in the S2SRO data set shown, is that the amplitude of
-the scan pattern is larger. Most of this ``scan-synchronous'' noise is
-attributed to magnetic field pickup, as described in
-Section~\ref{sec:magpickup}. Also similar to the S2SRO data, an
+the scan pattern is larger. This ``scan-synchronous'' noise is
+attributed to a combination of magnetic field pickup, as described in
+Section~\ref{sec:magpickup}, and sky brightness variations due to
+changes in elevation (not to be confused with underlying changes in
+the atmosphere; a homogeneous atmosphere will appear brighter with
+increasing airmass). Also similar to the S2SRO data, an
 anti-correlation between the opacity and elevation is apparent.
@@
 
-An additional noise source that is significant in only a small subset
-of the data (and more so during the S2SRO period) is magnetic field
+An additional noise source that is significant primarily for wide-area
+scans (and more so during the S2SRO period) is magnetic field
 pickup. Since the bolometer signals are ultimately detected through
@@
 %-------------------------------------------------
+\subsection{Sky noise}
+\label{sec:skynoise}
+%-------------------------------------------------
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{skynoise}
+\caption{Illustration of the dominance of atmospheric variations on
+  long-timescales using 35\,min. of data from the fully-commissioned
+  instrument, observation 18 on 20110324. Red and blue lines show the
+  850 and 450\,micron\ common-mode signals, corresponding to the left-
+  and right- vertical axes, respectively. The black line shows the
+  line-of-sight opacity from the WVM as a proxy for atmospheric
+  emission (arbitrary scale). Bolometer baseline drifts on timescales
+  $\gsim$100\,s are clearly correlated with the atmosphere in this
+  case.}
+\label{fig:skynoise}
+\end{figure}
+
+The data shown in Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec} are
+dominated at low-frequencies ($\lsim$2\,Hz) by scan-synchronous noise,
+rathern than underlying variations in the atmosphere, or ``sky
+noise''. However, only $\sim$100\,s of data were analysed. In
+Fig.~\ref{fig:skynoise}, an order-of-magnitude longer
+(fully-commissioned) data set is plotted from observation 18 on
+20110324. The common-mode signals from the 450 and 850\,\micron\ focal
+planes are compared with the line-of-sight opacity at 183\,GHz
+(1.6\,mm) from the WVM \citep[linear relationships between this
+opacity, and those at 450 and 850\,micron\ are given
+in][]{dempsey2012}. Unlike the earlier plots, it is clear that the
+bolometer baseline drifts on timescales $\gsim$100\,s are tightly
+correlated with the atmospheric opacity (especially at 850\,\micron),
+which serves as a proxy for atmospheric emission. The fact that the
+correlation between the WVM and the 450\,micron\ bolometers is weaker
+either reflects an underlying different in the atmosphere at higher
+frequencies, or the presence of some other significant low-frequency
+noise source.
+
+While the relative importance of sky noise at low-frequencies is
+highly time- and scan-pattern dependent, in general it does not have a
+major impact on map-making. Upon common-mode removal (which is an
+integral part of our map-making strategy) uncorrelated noise remains
+(e.g., solid lines in Fig~\ref{fig:pspec}). As we will see in
+Section~\ref{sec:pca} and \ref{sec:experimental}, this remaining noise
+does not have a smoothly-varying correlation pattern across the focal
+plane one might expect if it were due to resolved ``clouds'' of
+atmospheric emission. A high-pass filter above $\sim$0.1--1.0\,Hz is
+usually employed remove it (Section~\ref{sec:components}), which will
+also remove any sky noise that is not already subtracted with the
+common-mode.
+
+%-------------------------------------------------
 \subsection{Principal component analysis}
@@
 Analysis \citep[PCA,][]{laurent2005,scott2008,perera2008}. Here we use
-PCA to explore correlated signals in \scuba\ data.
+PCA to further explore correlated signals in \scuba\ data.
 
@@
 (consistent with the visual appearance of the bolometer signals in
-Fig.~\ref{fig:pca}). Also note that the PSD for theis eigenvector
-exhibit nearly a pure $1/f$ signature with minimal line features, or a
-white-noise plateau (suggesting that these are high-\snr\ measurements
-of a purely low-frequency drift). Component~3 is weaker, yet also
-shows evidence of the scan-synchronous noise, combined with
+Fig.~\ref{fig:pca}). Also note that the PSD for this eigenvector
+exhibits nearly a pure $1/f$ signature with minimal line features, and
+no white-noise plateau (suggesting that these are high-\snr\
+measurements of a purely low-frequency drift). Component~3 is weaker,
+yet also shows evidence of the scan-synchronous noise, combined with
 high-frequency line features that appear primarily in columns 26, 29
@@
 scan-synchronous noise (which is presumably a mixture of
-magnetic-field pickup and atmospheric noise)? First, the
-smoothly-varying gradients in the eigenvalue maps (primarily at
-850\,\micron) suggests that this noise has a different response across
-the focal plane; in this case the PCA has identified two orthogonal
-shapes that, when mixed in different quantities, can reproduce most of
-the signal in each bolometer. It is also likely that the atmospheric
-contribution is stronger at 450\,\micron\ (i.e., predominantly
-Component~3).
-
-Like Componen~1, Component~2 is a purely $1/f$ signal, with an
+magnetic-field pickup and brightness variations in the atmosphere)?
+First, the smoothly-varying gradients in the eigenvalue maps
+(primarily at 850\,\micron) suggests that this noise has a different
+response across the focal plane; in this case the PCA has identified
+two orthogonal shapes that, when mixed in different quantities, can
+reproduce most of the signal in each bolometer. It is also likely that
+the atmospheric contribution is stronger at 450\,\micron\ (i.e.,
+predominantly Component~3).
+
+Like Component~1, Component~2 is a purely $1/f$ signal, with an
 amplitude comparable to Component~3, that appears almost exclusively
@@
 850\,\micron\ in Component~6). The signals that appear along columns
-is probably related to magnetic-field pickup. Other sources of lines
+are probably related to magnetic-field pickup. Other sources of lines
 are thought to be aliased high-frequency noise, and also the 60\,Hz
@@
 estimated from the data for each column would not remove
-it. Regardless, the dominant noise, in all cases, is at
-low-frequencies. This conclusion is central to our data reduction
-strategy described in Section~\ref{sec:algorithm}.
-
-\subsection{Overall performance and mapping speed}
+it. Regardless, the dominant noise in excess of the fundamental white
+noise limit, in all cases, is at low-frequencies. This conclusion is
+central to our data reduction strategy described in
+Section~\ref{sec:algorithm}.
 
@@
   Fig.~\ref{fig:lockman_pspec}) to the dotted line gives the whitened
-  line (dashed) , which is the expected shape of a source in panel (b)
+  line (dashed), which is the expected shape of a source in panel (b)
   from Fig.~\ref{fig:lockman_maps}. Finally, cross-correlating the
@@
 map (Fig.~\ref{fig:lockman_maps}b). To identify sources, we next apply
-a matched filter to the whitened maps.
+a ``matched filter'' to the whitened maps.
 
@@
 assumption, and also assuming that the map noise is white,
-cross-correlation between the map and the known PSF yields the
-maximum-likelihood flux density of supposed point-sources centred over
-every location in the resulting map \citep[an extremely well-known
-result throughout astronomy, see][]{stetson1987}. Peak identification
-in such smoothed maps have been used extensively in the submillimetre
-community, as both an efficient source-detection and photometric
-measurement strategy. In terms of the angular power spectrum, a
-matched filter may also be thought of as an optimal low-pass filter,
-that suppresses noise on scales smaller (and therefore at higher
-angular frequencies) than the beam. For the case at hand, the
-effective PSF for the whitened maps is given by the dotted curve in
-Fig.~\ref{fig:lockman_psf}. Maps smoothed by this shape are shown for
-the whitened signal, and jackknife maps in
+cross-correlation between the map and the known PSF, or matched
+filtering, yields the maximum-likelihood flux density of supposed
+point-sources centred over every location in the resulting map
+\citep[an extremely well-known result throughout astronomy,
+see][]{stetson1987}. Peak identification in such smoothed maps have
+been used extensively in the submillimetre community, as both an
+efficient source-detection and photometric measurement strategy. In
+terms of the angular power spectrum, a matched filter may also be
+thought of as an optimal low-pass filter, that suppresses noise on
+scales smaller (and therefore at higher angular frequencies) than the
+beam. For the case at hand, we may use this cross-correlation
+technique since the map has been whitened. However, we must first
+establish the effective shape of point sources in this map due both to
+map-making itself, and the whitening filter. We determine the effect
+of map-making by adding a synthetic (and high-\snr) point source to
+the real data, and measure its resulting shape in the map (solid and
+dotted lines in Fig.~\ref{fig:lockman_psf}, respectively). Next, the
+Fourier transform of the map filtered PSF is divided by the jackknife
+noise power spectrum to calculate the final whitened PSF (dashed line
+in Fig.~\ref{fig:lockmand_psf}). Maps smoothed by this shape are shown
+for the whitened signal, and jackknife maps in
 Figs.~\ref{fig:lockman_maps}c and f, respectively. Note that these

### 2012-12-03 resubmit_mnras/response.txt
final changes, draft of response letter

Draft of my answers preceded by "-":

Scientific Editor's Comments:

Some of he figures (e.g., Fig. 4) are too small for comfortable reading, while others (e.g., Fig. 6) are unnecessarily large. In Fig. 9, the axes need to be explained and labelled (at least the time range); I'm not sure why you include broken bolometers as "typical" - the space could be better used! In figures with many panels, it helps to label the panels (a), (b), etc. and refer to the panels by label - see the instructions to authors.

- Fig.4 had been modified, using a smaller number of sub-windows and a
  larger font it is now more readable.

- Fig.6 has been made smaller.

- Fig.9 now has axis labels and an improved caption - however, we
  elected to leave in the "broken" bolometers as this is a fairly
  typical portion of the focal plane (remembering that there is a ~69%
  yield); this gives the reader a good visual sense.

Although the paper is clearly written, it is rather wordy, and I encourage you to review the text and see whether you can express it more concisely.

There are some awkward phrases, e.g.,
"about 5000 bolometers" (don't you know how many?),
"generic science products" (surely you don't mean "generic"?),
"relatively simpler" (isn't "simpler" sufficient?),
"relatively clean" (do you mean "cleaner"?),
"nearly science-grade" (this requires explanation, or just omit this sentence, which occurs both in Section 1 and the conclusions: this duplication is unnecessary).
"The reason for this is probably due to" (you mean "This is probably due to").

- done

Assistant Editor's Comments:

Holland et al and Dempsey et al are not yet 'in press' as they have not been accepted. Please cite these as 'submitted to MNRAS'.

When submitting a revised version of this paper, it would be helpful if you could include the latest versions of the two companion papers as supplementary material for the referee.

- done as requested

Reviewer's Comments:

- We would like to thank the thorough reviewer for identfying several
  areas that needed clarification, and perhaps more importantly, three
  careless errors (which led to incorrect units in Figs.1 & 2, and
  confusing results in Figs.11 & 20), and an ommission (impact /
  treatment of the response of the bolometers and data acquisition
  system). Detail responses follow.

Overall: The paper will be a useful reference for SMURF users and clearly represents a lot of work. Forgive me for only picking on the residual bits that I see need improving, but this will be the most efficient way to get to the end of this process.

Although I appreciate the division between this paper (data processing) and others to do with the instrument etc, it would be useful from a user perspective to have some sense of number of working channels, some parametrization of survey depth as a function of time&area -- using your "white noise limit", number of bad detectors, statistics on spikes etc. A lot of material is presented in section 2 to characterize the bolometer data. It would be very useful to bring it home with an focal-plane-wide characterization summary.

- A new, short, Section 3.4 has been added to address this.

In the examples it seemed like the "default" pipeline was never the correct one to use. Is there a case where the default is the one to use?

- Perhaps "default" is the wrong word - in reality, the description in
  Section 3 is of the "baseline" parameters for the reduction, which
  are then added-to for specific reduction, as in the examples from
  section 4. The word "default" has been replaced with "baseline"
  throughout, and clarification added to the intro of Section 4 on
  example reductions.

I really think it's useful and proper to cross-reference the millimeter kilopixel arrays with comparable data volumes and science overlap (SPT and ACT) and associated data processing literature out there. See specific comments below.

- Agreed. The artificial division between "submm" and "mm"
  is not useful in this paper, and SPT and ACT are certainly relevant
  comparisons.

+ Abstract: "… 200 Hz. This rate is much greater…"

This may be a true statement, but it neglects that very analogous arrays have been deployed in the millimeter. I'm thinking of SPT and ACT, which are arguably more similar to SCUBA2, by virtue of their kilopixel scale and multiplexed readout, than the previous generation submillimeter. SPT downsamples to 100 samp/sec. ACT writes to disk at 400 samp/sec and then downsamples to 200 samp/sec for analysis. SCUBA 2 has more pixels than either of these experiments (factor of 10 for SPT and 3 for ACT). Then again, as written the sentence is true, though maybe "data rate" would be better than "rate" which would make the claim stronger, even when including the millimeter kilopixel instruments.

- Agreed, and re-worded so that SCUBA-2 is now "the largest of a new
  generation of multiplexed kilopixel bolo cmaers"

+ Sec 1 Para 4: "``cross-linking'',in which each portion of the map is sampled on a range of temporal time scales"

This is an odd definition of cross-linking. The essence of cross-linking, as it's name implies, is that the sky in a given pixel is sampled in various directions, thus connecting pixels in the map isotropically and producing nice symmetric noise properties (and in the case of a mapper like SMURF also symmetrizing the transfer function). Coming back to timescales, in the SMURF approach any timescale corresponding to a revisit of a pixel (except at map boundaries) is going to be wiped out due to filtering/ common mode removal. In this sense the main (nice) benefit of the X-linking is isotropy. I realize this is a general discussion of mapping, but even then, it's really that you're connecting all the pixels and thus isotropizing the covariance in map space. In principal the different time scales could come into play in a full ML mapper, but the different time scale visitations could be achieved without cross-linking. For these reasons, I would emphasize position angle variation etc, as you do in most parts of the following text -- don't lead in with "different time-scales".

- Agreed, and description modified.

+ Sec 1 Para 5: "baseline removal and other simple filters"

This would be a nice place to cite SPT (Say, http://adsabs.harvard.edu/abs/2011ApJ...743...90S).

- Done.

+ Sec 1 Para 6: "maximum likelihood techniques"

This would be a nice place to cite ACT (Say, http://adsabs.harvard.edu/abs/2011ApJ...729...62D or http://adsabs.harvard.edu/abs/2012arXiv1208.0050D -- though this is not accepted).

- Done (although used earlier Fowler 2010 as a reference).

+ Sec 1 Para 7: "merely linear in the number of iterations needed, while using a comparable amount of memory"

I got tripped up here. Maybe you can clarify. Here are my thoughts:

The first part seems like a tautology: of course iterative methods scale as the number of iterations. Often times, the bottom line is that inversion is essentially impossible computationally (for a map with a few million pixels say) and the iterative methods let you go forward. The less trivial questions are: how long do the iterations take and how many iterations do you need? The iteration time is going to scale as the time ordered data length (or say NlogN for FFT based methods). You'll need more iterations to get to larger scales (e.g., if you're trying to go beyond your array footprint): for most techniques that I've encountered, small scales will be recovered in ~10 iterations while larger scales can take thousands of iterations, depending on the instrument etc.

Is the second bit true? Does the matrix inversion really take the same amount of memory as the iterative techniques for all map/time ordered data sizes?

- Agreed. Added some discussion of how algorithms scale.

+ Sec 1 Para 8: "single high-end desktop computers"

It would be useful for SMURF/SCUBA2 users to know an example of what would qualify as "high-end". (Number of processors and amount of memory would probably be enough to give the idea.)

- added a setence describing the machine used for analysis in the
  paper: 64-bit, 8 cores, 2.7GHz, 48 Gb of RAM.

+ Sec 2.2, 2.4, Figure 1, etc: e.g., "It is also likely that a varying atmospheric contribution is also included in these eigenvectors…"

I didn't get a good feeling from these sections as to how large a signal (say in pW over some time period) the atmospheric drift is in the data. There was a lot on scan synchronous noise (either before debugging the cryogenic problem or after where magnetic pickup seems to dominate). I was surprised to see that the change in elevation didn't give a signal due to changes in airmass. Could you give a better sense of how the atmosphere contributes? A possibly useful figure (though I know there are a lot of figures) is to plot time streams or spectra taken during different conditions (parametrized by PWV?). Even to just to give some quantitative sense of the magnitude in of the atmosphere in the data would be useful. The first components of the PCA were said to be scan synchronous noise dominated "presumably magnetic field pickup"… Is the atmosphere scan synchronous (e.g. b/c of airmass considerations?)… I was just left wondering here. From figure 1, right panel, it looks like after the magnetic field is removed, then the white noise totally dominates atmospheric drift. Figure 6 looks like a good example atmospheric drift maybe???

- After reading your comments we realised that the paper is lacking
  some detail on the atmosphere. While we felt that a really thorough
  analysis was beyond the scope of this paper (which is already very
  long!), we added a couple of things to help. First, Fig.1 has
  additional panels showing the line-of-sight opacity from the water
  vapour monitor. This *does* show the effect of elevation, so we've
  backed-off slightly on the claim that everything is magnetic field
  pickup (although it is all still scan-synchronous). Next, we've
  added a new Fig.4 and Section 2.4, in which we show some bolometer
  data over a much longer period (20 min.), in which the
  long-timescale behaviour is clearly dominated by underlying
  atmospheric variations. In the end, however, it is the
  higher-frequency scan-synchronous stuff that matters most for
  map-making. Finally, we re-ran the PCA code after first doing extra
  cleaning on the data (the first version of the plot had glitches in
  it which weren't particularly insightful). We were able to reduce
  the panels down to 9 from 12 to make them more readable, and the
  units are also now correct (see below). The discussion has been
  modified in Section 2.5 to reflect these changes, and the modified
  interpretation of sky noise.

+ Figure 2: "2x10^5 pW Hz^{-1}"

This number (and accompanying y-axis) seems reasonable given the plots in figure 1 (white noise is 2x10^4 pW Pk-Pk), but the white noise in Figure 6 looks much smaller (white noise is 0.01 pW Pk-Pk). Also glancing at Fig 9 in the companion paper (Holland et al) suggests NEPs of ~ 10^-16 W-rtHz or 10^-4 pW-rtHz. The latter numbers are more what I'm used to seeing in bolometer systems at these temps. Please clarify units. Is there a calibration missing…?

- Thank you for catching this! The original plots were not
  flatfielded, so they were in rather useless DAC units. Figs.1, 2,
  and the PCA plot have been re-made with the correct, sensible units
  of pW.

+ Sec 2.4, Para 4: "resembles the scan pattern in the right-hand panel of fig. 4"

I don't see a scan pattern in figure 4…

- Fixed.

+ Sec 3, Para 3: "the map is used to estimate and remove the astronomical signal contamination in the bolometer data…"

This seems like a oxymoron: astronomical signal contamination. I think it'll confuse people. Maybe "remove the astronomical signal from the bolometer data, leaving only white noise (NOI)". Or something like that. (It's also weird to talk about data which is only noise as "clean"!)

- Re-worded.

+ Figure 5. "DC steps"

There are enough acronyms running around this paper that it took me a while to realize what this meant. And that's only because I've worked in a lab where DC is a normal (and completely overloaded) term.  Maybe just drop "DC". You talk about "step correction" later. So probably "step" is less confusing that "DC step", which I don't see come up later.

- Agreed, term changed to "steps" throughout.

+ Sec 3.1

I didn't see any discussion of accounting for the time constants of the detectors or of the MCE anti-aliasing filter. Do you not deconvolve these as an initial step? Seems like this would give you a non-trivial transfer function at small scales if you did not…. (even if you have your nice cross-linked isotropic transfer function, you would have the unfortunate characteristic of wiping out small scale structure). Please clarify.

- This was a fairly major ommission in the first version. This is an
  effect that we considered long ago, and had been ignoring due to
  early estimates that showed they should be insignificant. In fact,
  there was no special code in SMURF to handle the de-convolution of
  these response functions. However, we have calculated and simulated
  these effects, finding that it can actually be quite significant at
  450um (15% attenuation of a point source at the maximum scan speed
  of 600 "/s). Fortunately this will not affect many users, but it is
  important to provide as a warning. The current development branch of
  SMURF now has code to handle this correction, but it is not in the
  release that is otherwise described in this paper. The details of
  the response are now added to Secion 2.1, and we've added Section
  3.1.7 to discuss the impact on maps.

+ Sec 3.1.1 Para 2: "fractional samples"

What is a fractional sample? Would be nice to clarify in text.

- The way input samples are divided in the lower sample rate output
  stream is now clarified in the text.

+ Sec 3.1.2, 3.1.3: de-spiking and step correction

For the SMURF user, it would be nice to know how often this occurs in a dataset. (Related to my general comment of giving an idea of the array/observation properties of the time stream data) It would be nice to have some statistics.

- some example values are now given in the new section 3.4.

+ Sec 3.1.4, Para 1:  "bad values are then replaced by linear interpolation between…"

Are these given zero weight in the map?

- correct. modified in the text.

+ Sec 3.1.4, Para 3: "The default number of samples modified at each end of the time-series is given by half the ratio of the sampling frequency to the lowest retained frequency."

Unclear at this point what "lowest retained frequency" means -- please clarify.

- This frequency corresponds to the edge of the high-pass filter that
  we used. Now clarified.

+ Sec 3.2, Para 4: "During pre-processing, the map-maker divides by f_i once. Then in each iteration, the map-maker divides by f_i…"

Seems like the data is repeatedly being divided by the flatfield and digitization constant… especially confusing b/c in the second paragraph we have "applied once during preprocessing"

- This was a mistake in the text. This division only happens once at
  the start. Now corrected.

+ Sec 3.2.1, Para 5: "cases where the data is dominated by magnetic field pickup"

Why not remove magnetic field pickup as a preprocessing step? It's generally on scales larger than you'll ever recover with SMURF and you can use the tricks (dark squids etc) that you have in the DKS module…

- It wouldn't work as a pre-processing step because the fit would be
  quite noisy due to the presence of other signals (e.g. bright
  astronomical sources).

+ Sec 3.2.1, Para 6: "Bolometer values that were flagged… are, however included in the new common-mode estimate"

Again, couldn't this exclusion be a pre-processing step (after the hypothetical DKS preprocessing) so you don't have to deal with these bad detectors?

- Sometimes yes, sometimes no. See comments below.

+ Sec 3.2.6, Para 2: "not all columns have working dark squids"

Glancing at Holland et al, it looks like the arrays have 60-80% yield, could one of the dead detectors be used as a dark channel (or are they dead because the mux is dead?).  In fact it's nice to have a "dark squid" on all rows and columns to get rid of noise (low and high frequency) related to the electronics.

- Yes, this was actually an idea at one point. However, since we had
  little success with the working dark squids, it was decided that it
  wasn't worth the effort.

+ Sec 3.2.6, Para 2: "necessitating the continued use of FLT… they [DKS/TMP] are not generally used"

Again, though, why not do this as a preprocessing step together with identifying all the bad non COM detectors? The dark detectors should not only carry information about the magnetic field pickup, but also contain high-frequency information related to row addressing and column readout that can be used to decontaminate the data. (Even if FLT does tend to kill everything on these scales, this could be a COM improving step where you can toss the bad detectors that don't follow common mode)

- In principle we agree with most of your comments on the dark squids;
  they should carry useful information that can beat-down the noise at
  all frequencies (not just the lowest frequencies). In practice the
  results were all over the place. Sometimes it helped a bit, but
  often the noise increased. It's rather frustrating that this did not
  work any better, but ultimately we had to resort to high-pass
  filtering to get good, repeatable results. We have tried to express
  this more clearly in this section.

+ Sec 3.3, Para 1: "Once the white noise, n_i^w(t) has been measured from the bolometer PSDs in NOI"

Is the noise time stream "model" n_i^w(t) really estimated from the PSDs or is this the weight w_i (or variance sigma_i^2…) that is estimated from the PSDs?

- We had some poor word choices in this section. We now clarify that
  "NOI" is just measuring the PSD to estimate a scalar variance, which
  later gets turned into the weight for map estimation, w_i = 1/sigma_i^2.

+ Sec 3.3, Para 1: "requsted threshikd"

typo

- fixed

+ Sec 3.3, Para 2: "does not necessarily correlate with convergence of the most important model component"

It was not clear to me when one would every use the "chi-sq" convergence criterion. The map-based one looks much more conventional to me (a la the conjugate gradient techniques). Please include a blurb on when the SMURF user would use the "chi-sq" convergence criterion in addition or as replacement for the map-based one.

- Correct. The chi^2 test is really more useful as an introduction to
  model degeneracies - it isn't useful as a stopping criterion as you
  say. This section has been re-written.

+ Sec 4.1, Para 4: "This example illustrate the need…"

typo

- fixed

+ Sec 4.1, Para 4: "a good, simple prior is to constrain the map to a value of zero away from the known locations of emission"

How does not masking a bright source prevent filter ringing? are you masking all the "sidelobes" of the filter? Please describe the mechanics of this "simple prior" in an additional one or more sentences.

- We have added some clarification in a new paragraph. In essence, the
  bolometer data are capable of constraining *differences* in the map
  up to the scale of the filter. Constraining (empty) portions of the
  map to zero provides references for those differences so that you
  get a reasonable answer out.

+ Figure 10:  "Same as (a) but using 50 iterations…"

I think you mean 100 iterations?

- Correct, fixed.

+ Figure 10: "Uranus peak 0.15 pW" -> "Uranus peak 0.27 pW"… "little attenuation of the source flux"

Can you explain why adding an extra filter to the process (just common mode -> filter + common mode) results in a increase (.15->.27 pW) in the peak brightness? Is there small scale noise (a stripe?) that is randomly knocking down the signal in the common mode removed case that gets nixed by the filter? Also can you quantify "little attenuation"? What do you expect for the uranus brightness in this measurement?

- Another careless mistake. The top two plots were made without
  extinction correction, while the bottom-two were. This figure has
  been re-made (note that the peak fluxes have changed since we were
  also forgetting to include some recent changes to the calibration),
  and things are more sensible: the extra filter gives you a *lower*
  flux now, as you correctly expected.

+ Sec 4.2, Para 1: "whitening (essentially flattening) the map"

flattening in Fourier space? As written a neophyte might think you mean making the map flat=constant! this is a potentially confusing parenthetical piece of jargon… maybe just drop since you describe it later?

- re-worded as a "Fourier-space filter to suppress residual large-scale noise".

+ Sec 4.2, Para 2: "However, the way this filter affects point-sources is both well-known, and linear."

Presumably this depends on the cut-off of the pre-processing filter. Does SMURF tell the user what this is going to be or does SMURF provide the tools to make the synthetic source so they can test it out? It would be useful to include such information.

- SMURF has a facility for adding in synthetic sources to real data
  on-the-fly. This is the mechanism we have used (and recommend to
  users), and we now state this in the text.

+ Figure 12: "this search radius"

8"? From the wording, it's unclear what is the "search radius". Please clarify.

- Correct, fixed.

+ Sec 4.2, Para 8: "Under this assumption, cross-correlation between between the map and the known PSF…"

This statement also assumes white noise.

- caveat added.

+ Sec 4.2, Para 8: "For the case at hand, the effective PSF for the whitened maps is given by the dotted curve in Fig. 13."

This is confusing language. Fig 13 list the dashed line as "whitened". It is true though that the dotted curve gives the correct PSF for your matched filter, since it's the one that corresponds to the pre-filtered beam. The matched filter in Fourier space is B(k)N^-1(k), where the N^-1(k) is your pre-whitening step (dividing by the power spectrum of the noise). The B(k) is the PSF before the whitening. It seems like you say this here, but it's a bit confusing with the figure labeling. Maybe it would be useful to include a simple equation like  \Phi(k) =  B(k)N^-1(k) or similar to describe your total operation?

- Yes, we are essentially breaking down the typical B(k)/N^-1(k)
  matched-filter expression into two steps. We've clarified the text
  here.

+ Sec 4.3, Para 2: "Much like the reduction of a point source without any prior constraints (Fig 10c), degeneracies on scales larger than the array footprint and filter produce ripples in the map."

This sentence confused me. In the source case you had two things: 1) degeneracies between your map and the common mode that introduced very large modes, which you eliminated by wiping out those scales with a filter before estimating the map. And then you found 2) that the filter was ringing and producing ripples in the map. I think you're saying you have these two problems here but the sentence, as written, made it seem like the two were somehow interacting or somehow the same thing. Maybe something more explicit like "Much like…, mapping extended emission would suffer from degeneracy between the common mode and map (if no filter were used) and, once the filter is introduced, ringing around sources.

- Yes, this is confusing because we're mixing two ideas. In fact, both
  the high-pass filtering and the common-mode removal can have similar
  problems with degeneracies, but it depends on the scale of the
  filter relative to the array footprint. Some new text in both
  Section 4.1 and 4.3 now explains this better.

+ Sec 4.3, Para 4: "By the final iteration, most of the obvious structures in the data are encompassed by the mask…"

This is opposite the bright source case (example 1) in which everything but the source was masked. Since you're drawing an analogy (see previous comment) with the point source reduction, would you add a bit more description to explain why you use the "opposite" mask here?

- Our description was very confusing here! Due to a poor word choice
  it sounds as though we were using the inverse of what we really
  do. The portion of the map contrained to zero is where the SNR is
  *below* some threshold (so similar to the case of Uranus). This has
  been clarified.

+ Fig 18

Why is the jackknife noise higher than the map noise at small (<25") angular scales?

- Another careless mistake. The average maps used variance weighting,
  whereas the jackknife (difference) maps did not, leading to greater
  noise. This has now been fixed and the plots look more sensible at
  small scales.

+ Sec 4.3, Para 5: "at least a cosmetic sense"

in a cosmetic sense?

- Fixed.

+ Sec 4.3, Para 9: "in which the SNR threshold of 5 is again used to constrain the map"

A bit confusing, esp. b/c I only found S/N=5  mentioned before in a caption. Maybe "S/N threshold is used to define the extended brightness mask and constrain the map" or something a little more explicit like this...?

- Extra text added to clarify that pixels below this threshold are set to zero.

+ Sec 4.3, Para 13 "In this case the noise… in the left panel of the third row"

Would really help to give figure numbers here.

- Added.

+ Sec 4.3, Final Paragraph: "such structures have a chance to grow"

too vague -- "such structures" could be interpreted as the large scale astro structure. Maybe something like "degeneracies with the common mode have the chance to show up in the map"

- Text changed to something resembling this suggestion

### 2012-12-03 smurf_paper.tex
final changes, draft of response letter

@@
 experiments for studying anisotropies in the Cosmic Microwave
-Backround (CMB) using similar technology, the latter of which uses the
+Background (CMB) using similar technology, the latter of which uses the
 same time-domain multiplexed readout electronics as \scuba\
@@
 angles. Such methods are now used by virtually all existing
-ground-based submillimetre camerass
+ground-based submillimetre cameras
 \citep[e.g.,][]{glenn1998,weferling2002,wilson2008,kovacs2008b}, in
@@
 to achieve SPTs focussed objectives \citep[in contrast, the similar
-ACT experiment uses cross-linking to improve its reponse to
+ACT experiment uses cross-linking to improve its response to
 large-angular scales,][]{das2011}. This approach is less practical as
@@
 are removed with an iterative high-pass filter. This strategy enables
-SMURF to reduce data on time scales comensurate with the observing
+SMURF to reduce data on time scales commensurate with the observing
 times (or significantly faster), on single high-end desktop
@@
 observers at the telescope. The pipeline is also used to generate
-near-science grade products for the JCMT Science Archive
-\citep{2011ASPC..442..203E} hosted by the Canadian Astronomy Data
-Centre (CADC). Details of the pipeline design are given in
-\citet{gibb2005} and \citet{2008AN....329..295C}.
+products for the JCMT Science Archive \citep{2011ASPC..442..203E}
+hosted by the Canadian Astronomy Data Centre (CADC). Details of the
+pipeline design are given in \citet{gibb2005} and
+\citet{2008AN....329..295C}.
 
@@
 (Section~\ref{sec:bolos}), give examples of typical bolometers signals
-(Section~\ref{sec:bolosignal}), and the impact of magnetic field
-pickup (Section~\ref{sec:magpickup}), and finally illustrate the use
-of principal component analysis to explore the correlated noise
-properties of \scuba\ data (Section~\ref{sec:pca}).
+(Sections~\ref{sec:bolosignal}), discuss the impacts of magnetic field
+pickup (Section~\ref{sec:magpickup}) and sky noise
+(Section~\ref{sec:skynoise}), and finally use of principal component
+analysis to explore the correlated noise properties of \scuba\ data
+(Section~\ref{sec:pca}).
 
@@
 astronomical flux units from power involves a correction for
-atmospheric extinction \citep[primarily using the JCMT Water Vapour
-Monitor to track line-of-sight opacity variations, see Section~3
-in][]{dempsey2012}, and the application of a flux conversion factor
-(FCF) which is established from regular measurements of calibrators
-such as Uranus \citep[Section~5 in][]{dempsey2012}.
+atmospheric extinction \citep[primarily using the 183\,GHz (1.6\,mm)
+JCMT Water Vapour Monitor to track line-of-sight opacity variations,
+see Section~3 in][]{dempsey2012}, and the application of a flux
+conversion factor (FCF) which is established from regular measurements
+of calibrators such as Uranus \citep[Section~5 in][]{dempsey2012}.
 
@@
 scans (and more so during the S2SRO period) is magnetic field
-pickup. Since the bolometer signals are ultimately detected through
-the amplification of magnetic fields, any additional changing fields
-within the instrument will also be detected.
+pickup. Since the bolometer signals are detected through the
+amplification of magnetic fields, any additional changing fields
+within the instrument will add to the noise.
 
@@
 dominated at low-frequencies ($\lsim$2\,Hz) by scan-synchronous noise,
-rathern than underlying variations in the atmosphere, or ``sky
+Rather than underlying variations in the atmosphere, or ``sky
 noise''. However, only $\sim$100\,s of data were analysed. In
@@
 20110324. The common-mode signals from the 450 and 850\,\micron\ focal
-planes are compared with the line-of-sight opacity at 183\,GHz
-(1.6\,mm) from the WVM \citep[linear relationships between this
-opacity, and those at 450 and 850\,micron\ are given
-in][]{dempsey2012}. Unlike the earlier plots, it is clear that the
-bolometer baseline drifts on timescales $\gsim$100\,s are tightly
-correlated with the atmospheric opacity (especially at 850\,\micron),
-which serves as a proxy for atmospheric emission. The fact that the
-correlation between the WVM and the 450\,micron\ bolometers is weaker
-either reflects an underlying different in the atmosphere at higher
-frequencies, or the presence of some other significant low-frequency
-noise source.
+planes are compared with the (arbitrarily scaled) line-of-sight
+opacity at 183\,GHz from the WVM. Unlike the earlier plots, it is
+clear that the bolometer baseline drifts on timescales $\gsim$100\,s
+are tightly correlated with the atmospheric opacity (especially at
+850\,\micron), which serves as a proxy for atmospheric emission. The
+fact that the correlation between the WVM and the 450\,micron\
+bolometers is weaker either reflects an underlying difference in the
+atmosphere at higher frequencies, or the presence of some other
+significant low-frequency noise source.
 
 While the relative importance of sky noise at low-frequencies is
-highly time- and scan-pattern dependent, in general it does not have a
+highly time- and scan pattern-dependent, in general it does not have a
 major impact on map-making. Upon common-mode removal (which is an
@@
 atmospheric emission. A high-pass filter above $\sim$0.1--1.0\,Hz is
-usually employed remove it (Section~\ref{sec:components}), which will
-also remove any sky noise that is not already subtracted with the
-common-mode.
+usually employed remove this residual noise
+(Section~\ref{sec:components}), which will also remove any sky noise
+that is not already subtracted by the common-mode.
 
@@
 (e.g. columns 15, and 26--30 at 450\,\micron\ in Component~4), or in a
-sampling of isolated bolometers (e.g. colummn 13, row 19 at
+sampling of isolated bolometers (e.g. column 13, row 19 at
 850\,\micron\ in Component~6). The signals that appear along columns
@@
 iterative algorithm in detail. Section~\ref{sec:converge} explores
-convergence tests and degeneracies between model components.
+convergence tests and degeneracies between model components. Finally,
+Section~\ref{sec:performance} summarizes the mapping performance of
+\scuba\ using SMURF.
 
@@
 simply tasks that do not need to be iterated. All of the data that are
-flagged as bad are ignored when estimating the map. As an example, for
-the 450\,\micron\ scan of Uranus that will be presented in
-Section~\ref{sec:point}: 31.00\% of the data (1587/5120 bolometers)
-are flagged due to data acquisition problems (dead or deactivated
-bolometers, or non-linear flatfields); 0.12\% of the data are flagged
-due to steps (including a small region about the steps); and 0.83\% if
-the data are flagged because the telescope was stationary.
+flagged as bad are ignored when estimating the map.
 
@@
 equivalent to applying a sinc-function low-pass filter. Such behaviour
-is desirable, since it serves as an anti-aliasing filter; most
+is desirable since it serves as an anti-aliasing filter; most
 non-white features above the Nyquist frequency in $y_j$ that could
@@
 de-convolve the system response. However, it is presently being added
-and should be present in future releases.
+and will be present in future releases.
 
@@
 components.
-
-Additional flagging of data to be ignored in the map occurs during the
-iterations. For the example of the 450\,\micron\ Uranus scan in
-Section~\ref{sec:point}: an additional 55 spikes are detected using
-the map-based despiker ($1.2\times10^{-6}$ \% of the data); and 0.72\%
-of the data are flagged as outliers compared to the common-mode.
 
@@
 contribute signal to some bolometers but not other bolometers, thus
-biassing the simple mean used to estimate the common mode. However, on
+biasing the simple mean used to estimate the common mode. However, on
 each iteration of the map-making algorithm illustrated in
@@
 First, the bolometer PSDs are calculated as in Eq.~\ref{eq:psd}. An
-average white noise level is then measured from 2 to 10\,Hz, a
-relatively clean region of the PSD that tends to lie above the $1/f$
-knee, but below the high-frequency line features for typical bolometer
-data (Fig.~\ref{fig:pspec}). Taking this constant level for
-$\mathbf{P}(f)$ we then calculate the expected variance of the
-time-series (in approximately 200\,Hz samples) using
-Eq.~\ref{eq:psd}. If the bolometer noise were produced purely by
-uncorrelated sources (i.e., no other long time-scale drifts), with no
-high-frequency line features, and without the attenuation at even
-higher frequencies by the anti-aliasing filter, this is the
-theoretical noise limit of the detectors. These measured variances are
-stored, and then used in subsequent iterations to weight the data
-points when calculating the map estimate (Section~\ref{sec:ast}). They
-are also used for convergence tests (Section~\ref{sec:converge}).
+average white noise level is then measured from 2 to 10\,Hz, a clean
+region of the PSD that tends to lie above the $1/f$ knee, but below
+the high-frequency line features for typical bolometer data
+(Fig.~\ref{fig:pspec}). Taking this constant level for $\mathbf{P}(f)$
+we then calculate the expected variance of the time-series (in
+approximately 200\,Hz samples) using Eq.~\ref{eq:psd}. If the
+bolometer noise were produced purely by uncorrelated sources (i.e., no
+other long time-scale drifts), with no high-frequency line features,
+and without the attenuation at even higher frequencies by the
+anti-aliasing filter, this is the theoretical noise limit of the
+detectors. These measured variances are stored, and then used in
+subsequent iterations to weight the data points when calculating the
+map estimate (Section~\ref{sec:ast}). They are also used for
+convergence tests (Section~\ref{sec:converge}).
 
@@
 
-Additional models exist as options, altough they are not generally
+Additional models exist as options, although they are not generally
 used: \model{DKS}, the use of dark squids as a template for removing
@@
 \model{COM}, the values of \model{COM} and \model{FLT} from the
-previous iteraion are added back into the residual simultaneously. In
+previous iteration are added back into the residual simultaneously. In
 this way, truly common-mode signals, even at low-frequencies, do not
@@
 
+%-------------------------------------------------
+\subsection{Instrument and map-making performance}
+\label{sec:performance}
+%-------------------------------------------------
+
+From a scientific perspective, the most important goal of map-making
+is to efficiently use all of the available data to achieve the
+greatest sensitivity that the instrument is capable of. Generally
+speaking, only a small percentage of the data are flagged as unusable
+during map-making beyond the $\sim$69\% detector yield.  As an
+example, for the 450\,\micron\ scan of Uranus that will be presented
+in Section~\ref{sec:point}, during pre-processing: 31.00\% of the data
+(1587/5120 bolometers) are flagged due to data acquisition problems
+(dead or deactivated bolometers, or non-linear flatfields); 0.12\% of
+the data are flagged due to steps (including a small region about the
+steps, Section~\ref{sec:steps}); and 0.83\% if the data are flagged
+because the telescope was stationary
+(Section~\ref{sec:flagbadbol}). Once the iterative solution begins: 55
+spikes are detected using the map-based despiker ($1.2\times10^{-6}$
+\% of the data, Section~\ref{sec:ast}); and 0.72\% of the data are
+flagged as outliers compared to the common-mode
+(Section~\ref{sec:comgai}).
+
+The white noise levels of \scuba\ bolometers are measured between 2
+and 10\,Hz in the PSDs (as described in Section~\ref{sec:noi}),
+leading to noise equivalent powers (NEPs, with units W\,s$^{1/2}$) --
+the \rms\ time-series noise in a 1\,s integration \citep[see section
+3.5.2 in][]{holland2012}. The NEPs are then converted into more useful
+noise equivalent flux densities (NEFDs, with units Jy\,s$^{1/2}$)
+through multiplication by the extinction correction and the FCF. For
+an ideal scan pattern in which every bolometer spends an equal amount
+of time observing each point on the sky, the expected noise in a map
+pixel is $\mathrm{NEFD_{eff}} t^{-1/2}$, where $t$ is the integration
+time, and the effective NEFD, $\mathrm{NEFD_{eff}} = (1/\sum
+\mathrm{NEFD}_i^{-2})^{1/2}$, is the sensitivity of the entire array
+including all bolometers $i$. Experimentally it has been found that
+the map noise based on the weighted scatter of the data
+(Eq.~\ref{eq:varmap}), and the scatter of map pixels in small
+apertures, both correlate well with this expected limiting noise
+performance for maps that are larger than the array footprint (for
+smaller maps the assumption of uniform coverage is incorrect). As an
+example of observing under favourable conditions with line-of-sight
+opacities of 0.87 and 0.34 at 450 and 850\,\micron\ \citep[zenith sky
+opacities at 225\,GHz of 0.04 and 0.065 at 450 and 850\,\micron,
+respectively, at an average airmass of 1.2, from table 3
+in][]{holland2012}, the mapping speed is approximately
+$6\times10^{-5}$ and $3\times10^{-3}$\,deg$^2$\,hr$^{-1}$\,mJy$^{-2}$,
+at 450 and 850\micron, for the default pixel sizes of 2 and 4\,arcsec,
+respectively. If maps are high-pass filtered and cross-correlated with
+the PSF for the optimal extraction of point sources (see
+Section~\ref{sec:cosmo}), the mapping speeds increase to approximately
+$5\times10^{-4}$ and
+$1.5\times10^{-2}$\,deg$^2$\,hr$^{-1}$\,mJy$^{-2}$, independently of
+the chosen map pixel sizes. See the on-line \scuba\ integration time
+calculator for mapping speeds under arbitrary observing conditions
+and scan
+modes\footnote{\url{http://www.jach.hawaii.edu/jac-bin/propscuba2itc.pl}}.
+
 %------------------------------------------------------------------------------
@@
 that are applicable to most common types of data. In this section we
-illustrate the differences bewteen these configurations with the
+illustrate the differences between these configurations with the
 following examples: a bright point source, Uranus
@@
 noise power spectrum to calculate the final whitened PSF (dashed line
-in Fig.~\ref{fig:lockmand_psf}). Maps smoothed by this shape are shown
+in Fig.~\ref{fig:lockman_psf}). Maps smoothed by this shape are shown
 for the whitened signal, and jackknife maps in
@@
 proven invaluable for the precise identifications of high-redshift
-submillimetre galaxies due to their relatively low surface densities
-(compared with optical catalogues, for example), and a strong
-correlation between the radio and submillimetre emission mechanisms
+submillimetre galaxies due to their low surface densities (compared
+with optical catalogues, for example), and a strong correlation
+between the radio and submillimetre emission mechanisms
 \citep[e.g.,][]{smail2000,pope2006,ivison2007,chapin2009b}. Taking a
@@
 \caption{An 850\,\micron\ rotating PONG map of M17. Intensity is
-logarithmically scaled between $-$0.0003\,pw (white) and +0.01\,pW
+logarithmically scaled between $-$0.0003\,pW (white) and +0.01\,pW
 (black). Iteration numbers are given in the corner of each
@@
 increase in noise; in other words, there is greatly improved \snr\ at
-large angular scales (an obvious conclusion given the appearence of
+large angular scales (an obvious conclusion given the appearance of
 the maps in Fig.~\ref{fig:m17_jk}). In fact, using the 900\,arcsec
@@
 correlated noise components from the data, successively improving the
-map \citep[e.g.,][]{kovacs2008,aguirre2011,schuller2012}, we have shown
-that a relatively simpler approach provides good results for \scuba\
+map \citep[e.g.,][]{kovacs2008,aguirre2011,schuller2012}, we have
+shown that a fairly simple approach provides good results for \scuba\
 data, with reasonable computational requirements. This conclusion is
@@
 capabilities of single, high-end desktop computers. SMURF can
-therefore provide real-time feedback at the telescope to observers,
-and it also produces nearly science-grade data products for the JCMT
-Science Archive hosted by the Canadian Astronomy Data Centre.
+therefore provide real-time feedback at the telescope to observers.
 
@@
 Research, and the National Research Council of Canada. Additional
-funds for the contruction of \scuba\ were provided by the Canada
+funds for the construction of \scuba\ were provided by the Canada
 Foundation for Innovation.  This research was supported in part by the
@@
 related to SMURF; and Mark Thompson, Craig Walther and S\'{e}verin
-Gaudet for being on the Critical Design Review panel. Finally, we
-thank Per Friberg and Gary Davis for their helpful comments on the
-manuscript.
+Gaudet for being on the Critical Design Review panel. We thank Per
+Friberg and Gary Davis for their helpful comments on the
+manuscript. Finally, we thank the anonymous referee for identifying
+several key areas that required clarification.
 

### 2012-12-04 resubmit_mnras/response.txt
minor changes based on comments from co-authors

@@
 
-- Fig.9 now has axis labels and an improved caption - however, we
-  elected to leave in the "broken" bolometers as this is a fairly
-  typical portion of the focal plane (remembering that there is a ~69%
-  yield); this gives the reader a good visual sense.
+- Fig.9 is now Fig. 10, and now has axis labels and an improved
+  caption - however, we elected to leave in the "broken" bolometers as
+  this is a fairly typical portion of the focal plane (remembering
+  that there is a ~70% yield); this gives the reader a good visual
+  sense of that.
 
@@
 
+- in the References the "in press" has been replaced with "submitted"
+  for these two papers
+
 When submitting a revised version of this paper, it would be helpful if you could include the latest versions of the two companion papers as supplementary material for the referee.
@@
   treatment of the response of the bolometers and data acquisition
-  system). Detail responses follow.
+  system). Detailed responses follow.
 

### 2012-12-04 smurf_paper.tex
minor changes based on comments from co-authors

@@
   instrument operating on the 15-m James Clerk Maxwell Telescope,
-  nominally consisting of 5120 bolometers (with a useful yield of
-  69\%) in each of two simultaneous imaging bands centred over 450 and
-  850\,\micron. The camera is operated by scanning across the sky and
-  recording data at a rate of 200\,Hz. As the largest of a new
-  generation of multiplexed kilopixel bolometer cameras operating in
-  the (sub)millimetre, \scuba\ data analysis represents a significant
-  challenge.  We describe the production of maps using the
-  Sub-Millimetre User Reduction Facility (SMURF) in which we have
-  adopted a fast, iterative approach to map-making that enables data
-  reduction on single, modern, high-end desktop computers, with
-  execution times that are typically shorter than the observing times.
-  SMURF is used in an automated setting, both at the telescope for
-  real-time feedback to observers, as well as for the production of
-  science products for the JCMT Science Archive at the Canadian
-  Astronomy Data Centre. Three detailed case studies are used to: (i)
-  explore convergence properties of the map-maker using simple prior
-  constraints (Uranus -- a point source); (ii) achieve the white-noise
-  limit for faint point-source studies (extragalactic blank-field
-  survey of the Lockman Hole); and (iii) demonstrate that our strategy
-  is capable of recovering angular scales comparable to the size of
-  the array footprint (approximately 5\,arcmin) for bright extended
-  sources (star-forming region M17).
+  nominally consisting of 5120 bolometers in each of two simultaneous
+  imaging bands centred over 450 and 850\,\micron. The camera is
+  operated by scanning across the sky and recording data at a rate of
+  200\,Hz. As the largest of a new generation of multiplexed kilopixel
+  bolometer cameras operating in the (sub)millimetre, \scuba\ data
+  analysis represents a significant challenge.  We describe the
+  production of maps using the Sub-Millimetre User Reduction Facility
+  (SMURF) in which we have adopted a fast, iterative approach to
+  map-making that enables data reduction on single, modern, high-end
+  desktop computers, with execution times that are typically shorter
+  than the observing times.  SMURF is used in an automated setting,
+  both at the telescope for real-time feedback to observers, as well
+  as for the production of science products for the JCMT Science
+  Archive at the Canadian Astronomy Data Centre. Three detailed case
+  studies are used to: (i) explore convergence properties of the
+  map-maker using simple prior constraints (Uranus -- a point source);
+  (ii) achieve the white-noise limit for faint point-source studies
+  (extragalactic blank-field survey of the Lockman Hole); and (iii)
+  demonstrate that our strategy is capable of recovering angular
+  scales comparable to the size of the array footprint (approximately
+  5\,arcmin) for bright extended sources (star-forming region M17).
 \end{abstract}
@@
 around nearby stars, helping us understand the early stages of planet
-formation.  With $10,240$ nominal detectors \citep[of which $\sim$69\%
+formation.  With $10,240$ nominal detectors \citep[of which $\sim$70\%
 work and are typically useful,][]{holland2012}, \scuba\ is presently
@@
 SMURF to reduce data on time scales commensurate with the observing
-times (or significantly faster), on single high-end desktop
-computers. For reference, all of the data analysis in this paper was
+times (or significantly faster), on single high-end desktop computers
+using the . For reference, all of the data analysis in this paper was
 performed on a machine with a 64-bit 8-core central processing unit
 operating at 2.7\,GHz, and 48\,Gb of memory. SMURF has been
-successfully used as part of a real-time pipeline offering feedback to
-observers at the telescope. The pipeline is also used to generate
-products for the JCMT Science Archive \citep{2011ASPC..442..203E}
-hosted by the Canadian Astronomy Data Centre (CADC). Details of the
-pipeline design are given in \citet{gibb2005} and
-\citet{2008AN....329..295C}.
+successfully used as part of a real-time pipeline \citep[based on the
+Observatory Reduction and Acquisition Control - Data Reduction system,
+or ORAC-DR,][]{gibb2005} offering feedback to observers at the
+telescope. The pipeline is also used to generate products for the JCMT
+Science Archive \citep{2011ASPC..442..203E} hosted by the Canadian
+Astronomy Data Centre (CADC). Details of the pipeline design are given
+in \citet{gibb2005} and \citet{2008AN....329..295C}.
 
@@
 the focal plane itself, it is certainly correlated with the
-temperature of the detectors), the line-of-sight atmoshperic opacity
+temperature of the detectors), the line-of-sight atmospheric opacity
 as derived from the JCMT water vapour monitor (WVM), and the telescope
@@
 \caption{Illustration of the dominance of atmospheric variations on
-  long-timescales using 35\,min. of data from the fully-commissioned
+  long-timescales using 35\,min. of data taken in opaque, and
+  highly-variable weather conditions (line-of-sight 450\,\micron\
+  opacity ranging from 12 to 22) from the fully-commissioned
   instrument, observation 18 on 20110324. Red and blue lines show the
-  850 and 450\,micron\ common-mode signals, corresponding to the left-
+  850 and 450\,\micron\ common-mode signals, corresponding to the left-
   and right- vertical axes, respectively. The black line shows the
@@
 
-The data shown in Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec} are
-dominated at low-frequencies ($\lsim$2\,Hz) by scan-synchronous noise,
-Rather than underlying variations in the atmosphere, or ``sky
-noise''. However, only $\sim$100\,s of data were analysed. In
-Fig.~\ref{fig:skynoise}, an order-of-magnitude longer
-(fully-commissioned) data set is plotted from observation 18 on
-20110324. The common-mode signals from the 450 and 850\,\micron\ focal
-planes are compared with the (arbitrarily scaled) line-of-sight
-opacity at 183\,GHz from the WVM. Unlike the earlier plots, it is
-clear that the bolometer baseline drifts on timescales $\gsim$100\,s
-are tightly correlated with the atmospheric opacity (especially at
-850\,\micron), which serves as a proxy for atmospheric emission. The
-fact that the correlation between the WVM and the 450\,micron\
-bolometers is weaker either reflects an underlying difference in the
-atmosphere at higher frequencies, or the presence of some other
-significant low-frequency noise source.
+The data shown in Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec} were
+taken during fairly stable weather conditions (although with different
+450\,\micron\ line-of-sight opacities $\sim$0.9 and 6.5 in the S2SRO
+and post-upgrade data sets, respectively), and are dominated at
+low-frequencies ($\lsim$2\,Hz) by scan-synchronous noise, rather than
+underlying variations in the atmosphere, or ``sky noise''. However,
+only $\sim$100\,s of data were analysed. In Fig.~\ref{fig:skynoise},
+an order-of-magnitude longer (fully-commissioned) data set is plotted
+from observation 18 on 20110324, taken during more opaque, and highly
+variable weather conditions (450\,\micron\ line-of-sight opacities
+ranging from 12 to 22, scaled from the WVM data which are shown as a
+black line). Unlike the earlier plots, it is clear that the bolometer
+baseline drifts on timescales $\gsim$100\,s are tightly correlated
+with the atmospheric opacity, which serves as a proxy for atmospheric
+emission. The weaker correlation between the WVM and the 450\,micron\
+bolometers is probably due to the atmosphere being more transmissive
+at 850\micron, therefore the emission detected at 450\,\micron\ comes
+from water vapour that is closer to the telescope.
 
@@
 plane one might expect if it were due to resolved ``clouds'' of
-atmospheric emission. A high-pass filter above $\sim$0.1--1.0\,Hz is
-usually employed remove this residual noise
-(Section~\ref{sec:components}), which will also remove any sky noise
-that is not already subtracted by the common-mode.
+atmospheric emission. A high-pass filter with an edge frequency in the
+range $\sim$0.1--1.0\,Hz is usually employed remove this residual
+noise (Section~\ref{sec:components}), which will also remove any sky
+noise that is not already subtracted by the common-mode.
 
@@
 convergence tests and degeneracies between model components. Finally,
-Section~\ref{sec:performance} summarizes the mapping performance of
+Section~\ref{sec:performance} summarises the mapping performance of
 \scuba\ using SMURF.
@@
 distort signals.  The default allowable range of speed is
-30--1000\,arcsec\,s$^{-1}$, which primarily rejects short periods when
-the telescope was stationary prior to commencing the scan pattern
-(generally speaking the speed is constant once the scan begins),
-although the scan speed can briefly exceed the upper limit, even for a
-requested speed of 600\,arcsec, if the pointing system is having
-trouble keeping-up for large-area scans at high elevation.
+30--1000\,arcsec\,s$^{-1}$. These thresholds primarily reject short
+periods when the telescope was stationary prior to commencing the scan
+pattern (generally speaking the speed is constant once the scan
+begins). The scan speed can also briefly exceed the upper limit, even
+for a requested speed of 600\,arcsec, if the pointing system is
+struggling for large-area scans at high elevation.
 
@@
 For this reason, the usual practice is to estimate a single
-\model{COM} model by examining data from all four sub-arrays in each
+\model{COM} model by examining data from all four subarrays in each
 waveband, since this allows spatial structure on the scale of the
@@
 speaking, only a small percentage of the data are flagged as unusable
-during map-making beyond the $\sim$69\% detector yield.  As an
+during map-making beyond the $\sim$70\% detector yield.  As an
 example, for the 450\,\micron\ scan of Uranus that will be presented
@@
 of time observing each point on the sky, the expected noise in a map
-pixel is $\mathrm{NEFD_{eff}} t^{-1/2}$, where $t$ is the integration
+pixel is $\mathrm{NEFD_{eff}}/t^{1/2}$, where $t$ is the integration
 time, and the effective NEFD, $\mathrm{NEFD_{eff}} = (1/\sum
@@
 at 450 and 850\micron, for the default pixel sizes of 2 and 4\,arcsec,
-respectively. If maps are high-pass filtered and cross-correlated with
-the PSF for the optimal extraction of point sources (see
+respectively. If maps are filtered and cross-correlated with the PSF
+for the optimal extraction of point sources (see
 Section~\ref{sec:cosmo}), the mapping speeds increase to approximately
 $5\times10^{-4}$ and
-$1.5\times10^{-2}$\,deg$^2$\,hr$^{-1}$\,mJy$^{-2}$, independently of
-the chosen map pixel sizes. See the on-line \scuba\ integration time
-calculator for mapping speeds under arbitrary observing conditions
-and scan
+$1.5\times10^{-2}$\,deg$^2$\,hr$^{-1}$\,mJy$^{-2}$ in this example,
+independently of the chosen map pixel sizes. See the on-line \scuba\
+integration time calculator for mapping speeds under arbitrary
+observing conditions and scan
 modes\footnote{\url{http://www.jach.hawaii.edu/jac-bin/propscuba2itc.pl}}.
@@
 Starlink predicts a peak brightness of 176\,Jy on this date, yielding
-an FCF of 492 Jy\,pW$^{-1}$, which is well within the typical range
+an FCF of 491 Jy\,pW$^{-1}$, which is well within the typical range
 for \scuba\ of $(492 \pm 67)$\,Jy\,pW$^{-1}$ \citep[see][for
@@
 assuming that errors due to calibration between the two halves are
-insignificant). The approach we have taken to minimize systematics is
+insignificant). The approach we have taken to minimise systematics is
 to produce the two maps using odd and even scan numbers (i.e., each
@@
 second-lowest spatial frequency bin of
-$1.875\times10^{-3}$\,arcsec$^{-1}$, or 533\,arcsec, due to the fact
-that the smallest is very poorly sampled, and therefore noisy. As the
-filter edge is increased, more power is measured in the map PSDs at
-larger scales. However, much of this power is clearly produced by
-noise which appears in the jackknife PSDs. We estimate the map-maker
-transfer function as the ratio between the portion of the signal PSDs
-not produced by noise, to the input PSD, or
+$1.875\times10^{-3}$\,arcsec$^{-1}$, or 533\,arcsec, since the lowest
+is very poorly sampled, and therefore noisy. At small scales
+($\lsim$20\,arcsec) the map and jackknife PSDs flatten. However,
+unlike Fig.~\ref{fig:lockman_pspec}, these white noise levels are
+slightly larger in the jackknife maps, suggesting that the differences
+have not completely removed the astronomical signals. A strong
+possibility is that the two half-maps that go into the differences
+have not converged equally, probably due to the lack of prior
+constraints.  As the filter edge is increased, more power is measured
+in the map PSDs at larger scales. However, much of this power is
+clearly produced by noise which appears in the jackknife PSDs. We
+estimate the map-maker transfer function as the ratio between the
+portion of the signal PSDs not produced by noise, to the input PSD, or
 %
@@
 as in the baseline reduction, and the \snr\ improves at scales
-$\gsim$60\,arcsec.
+$\gsim$60\,arcsec. Finally, note that both the map and jackknife white
+noise levels (at scales $\lsim$20\,arcsec) are in excellent agreement,
+unlike the previous example.
 
@@
 Friberg and Gary Davis for their helpful comments on the
-manuscript. Finally, we thank the anonymous referee for identifying
-several key areas that required clarification.
+manuscript. Finally, we thank the anonymous referee for a very helpful
+review that helped to clarify several key areas of the paper.
 

### 2012-12-04 smurf_paper.tex
some more minor edits

@@
 data, they may not achieve the best noise performance on the angular
-scales of interest. A related method that has been popular for
-reducing data in fields of faint point sources with Bolocam
-\citep[e.g.,][]{laurent2005}, and its younger sibling the Aztronomical
-Thermal Emission Camera \citep[AzTEC, e.g.,][]{scott2008}, is
-principal component analysis (PCA) cleaning. A statistical
-``black-box'' removes the most correlated components of the bolometer
-signals, enabling the detection of point-sources close to the
-theoretical white-noise limits of the detectors, with reasonable
-computation times when small numbers of bolometers are involved
-(hundreds rather than thousands of detectors). However, PCA cleaning
-is not a good solution for producing maps of extended structures,
-since such sources produce correlated signals amongst many detectors,
-and are removed by this procedure \citep[an exception is the iterative
-PCA approach of][]{aguirre2011}. Furthermore, in the case of \scuba,
-performing PCA on even a single subarray (typically $\sim$900
-detectors) can be prohibitively slow.
+scales of interest. A method for reducing data in fields of faint
+point sources with Bolocam \citep[e.g.,][]{laurent2005}, and its
+younger sibling the Aztronomical Thermal Emission Camera \citep[AzTEC,
+e.g.,][]{scott2008}, is principal component analysis (PCA) cleaning. A
+statistical ``black-box'' removes the most correlated components of
+the bolometer signals, enabling the detection of point-sources close
+to the theoretical white-noise limits of the detectors, with
+reasonable computation times when small numbers of bolometers are
+involved (hundreds rather than thousands of detectors). However, PCA
+cleaning is not a good solution for producing maps of extended
+structures, since such sources produce correlated signals amongst many
+detectors, and are removed by this procedure \citep[an exception is
+the iterative PCA approach of][]{aguirre2011}. Furthermore, in the
+case of \scuba, performing PCA on even a single subarray (typically
+$\sim$900 detectors) can be prohibitively slow.
 
@@
 that is so complicated in the maximum-likelihood methods. In the
-iterative approach memory usage usually scales linearly with $N$, the
-product of the number of bolometers, $N_\mathrm{b}$, with the number
-of samples in time, $N_\mathrm{t}$. The computation time is the
-product of the number of iterations, $N_\mathrm{iter}$, with the time
-per iteration, $t_\mathrm{iter}$. For the algorithm described in this
+iterative approach memory usage scales linearly with $N$, the product
+of the number of bolometers, $N_\mathrm{b}$, with the number of
+samples in time, $N_\mathrm{t}$. The computation time is the product
+of the number of iterations, $N_\mathrm{iter}$, with the time per
+iteration, $t_\mathrm{iter}$. For the algorithm described in this
 paper, $N_\mathrm{iter}$ is typically in the range $\sim$5--100
@@
 SMURF to reduce data on time scales commensurate with the observing
-times (or significantly faster), on single high-end desktop computers
-using the . For reference, all of the data analysis in this paper was
+times (or significantly faster), on single high-end desktop
+computers. For reference, all of the data analysis in this paper was
 performed on a machine with a 64-bit 8-core central processing unit
@@
 Observatory Reduction and Acquisition Control - Data Reduction system,
-or ORAC-DR,][]{gibb2005} offering feedback to observers at the
-telescope. The pipeline is also used to generate products for the JCMT
-Science Archive \citep{2011ASPC..442..203E} hosted by the Canadian
-Astronomy Data Centre (CADC). Details of the pipeline design are given
-in \citet{gibb2005} and \citet{2008AN....329..295C}.
+or ORAC-DR,][]{gibb2005,2008AN....329..295C} offering feedback to
+observers at the telescope. The pipeline is also used to generate
+products for the JCMT Science Archive \citep{2011ASPC..442..203E}
+hosted by the Canadian Astronomy Data Centre (CADC).
 
@@
 \caption{Illustration of the dominance of atmospheric variations on
-  long-timescales using 35\,min. of data taken in opaque, and
-  highly-variable weather conditions (line-of-sight 450\,\micron\
-  opacity ranging from 12 to 22) from the fully-commissioned
-  instrument, observation 18 on 20110324. Red and blue lines show the
-  850 and 450\,\micron\ common-mode signals, corresponding to the left-
-  and right- vertical axes, respectively. The black line shows the
-  line-of-sight opacity from the WVM as a proxy for atmospheric
-  emission (arbitrary scale). Bolometer baseline drifts on timescales
-  $\gsim$100\,s are clearly correlated with the atmosphere in this
-  case.}
+  long-timescales using 35\,min. of data taken in highly-variable
+  weather conditions (line-of-sight 450\,\micron\ opacity ranging from
+  3.6 to 7.7) from the fully-commissioned instrument, observation 18
+  on 20110324. Red and blue lines show the 850 and 450\,\micron\
+  common-mode signals, corresponding to the left- and right- vertical
+  axes, respectively. The black line shows the line-of-sight opacity
+  from the WVM as a proxy for atmospheric emission (arbitrary
+  scale). Bolometer baseline drifts on timescales $\gsim$100\,s are
+  clearly correlated with the atmosphere in this case.}
 \label{fig:skynoise}
@@
 an order-of-magnitude longer (fully-commissioned) data set is plotted
-from observation 18 on 20110324, taken during more opaque, and highly
-variable weather conditions (450\,\micron\ line-of-sight opacities
-ranging from 12 to 22, scaled from the WVM data which are shown as a
-black line). Unlike the earlier plots, it is clear that the bolometer
-baseline drifts on timescales $\gsim$100\,s are tightly correlated
-with the atmospheric opacity, which serves as a proxy for atmospheric
-emission. The weaker correlation between the WVM and the 450\,micron\
-bolometers is probably due to the atmosphere being more transmissive
-at 850\micron, therefore the emission detected at 450\,\micron\ comes
-from water vapour that is closer to the telescope.
+from observation 18 on 20110324, during highly variable weather
+conditions (450\,\micron\ line-of-sight opacities ranging from 3.6 to
+7.7, scaled from the WVM data which are shown as a black line). Unlike
+the earlier plots, it is clear that the bolometer baseline drifts on
+timescales $\gsim$100\,s are tightly correlated with the atmospheric
+opacity, which serves as a proxy for atmospheric emission. The weaker
+correlation between the WVM and the 450\,micron\ bolometers is
+probably due to the atmosphere being more transmissive at 850\micron,
+therefore the emission detected at 450\,\micron\ comes from water
+vapour that is closer to the telescope.
 
@@
 low-pass filter). In practice the noise performance was
-indistinguishable, and required slightly longer execution time than using
-the method above. Therefore this method is not used.
+indistinguishable, and required slightly longer execution time
+compared with the previous method, and is therefore not used.
 
@@
 that have been flagged as bad for any reason (for instance, due to the
-presence of spikes, steps, or unusual common-mode signal) when need to
-be excluded from this FFT. For this reason, each contiguous block of
-bad data samples is filled with artificial data before taking the
-FFT. A least squares linear fit is performed to the 50 samples
-preceding the block, and a similar fit is performed to the 50 samples
-following the block. These are used to estimate the expected values at
-the start and end of the block of bad values. The bad values are then
-replaced by linear interpolation between the expected start and end
-values. Gaussian noise is added with a standard deviation equal to the
-mean of the RMS residuals in the two fits. These flagged and filled
-portions of the data are then given a weight of zero when estimating
-the map.
+presence of spikes, steps, or unusual common-mode signal) need to be
+excluded. For this reason, each contiguous block of bad data samples
+is filled with artificial data before taking the FFT. A least squares
+linear fit is performed to the 50 samples preceding the block, and a
+similar fit is performed to the 50 samples following the block. These
+are used to estimate the expected values at the start and end of the
+block of bad values. The bad values are then replaced by linear
+interpolation between the expected start and end values. Gaussian
+noise is added with a standard deviation equal to the mean of the RMS
+residuals in the two fits. These flagged and filled portions of the
+data are then given a weight of zero when estimating the map.
 
@@
 pattern (generally speaking the speed is constant once the scan
-begins). The scan speed can also briefly exceed the upper limit, even
-for a requested speed of 600\,arcsec, if the pointing system is
-struggling for large-area scans at high elevation.
+begins). The scan speed can also briefly exceed the upper limit when
+approaching the maximum requested speed of 600\,arcsec as the pointing
+system struggles at high elevation.
 
@@
 for; together they produce a net lag in the bolometer signals of about
-4.5\,ms (Section~\ref{sec:bolos}). This can be modelled by convolving
-the ``pure'' bolometer signals with a single system response
+4.5\,ms with respect to the pointing data
+(Section~\ref{sec:bolos}). This can be modelled by convolving the
+``pure'' bolometer signals with a single system response
 function. Similarly, this effect could be removed using de-convolution
@@
 squares linear fit is performed between the bolometer time-series and
-the common-mode to determine a gain [$g_i$ in Eq.~\ref{eq:noise}] and
+the common-mode to determine a gain ($g_i$ in Eq.~\ref{eq:noise}) and
 offset for each bolometer.  The gain and offset for each bolometer is
@@
 For the initial iteration the weights $\mathbf{w}_j$ are set to 1, but
-subsequently they are set to $1/\sigma_j^2$, the estimated inverse
-variance expected from the bolometer white noise levels, as discussed
-in Section~\ref{sec:noi}. This weighting scheme is sensible provided
-that the bolometer data have no correlated (e.g., low-frequency)
-noise.
+subsequently they are set to $1/\sigma_j^{\mathbf{w}2}$, the estimated
+inverse variance expected from the bolometer white noise levels, as
+discussed in Section~\ref{sec:noi}. This weighting scheme is sensible
+provided that the bolometer data have no correlated (e.g.,
+low-frequency) noise.
 
@@
   from \citet{owen2008} with $S\gsim15$\,$\mu$Jy. Of the 10 submm
-  peaks, 9 are within a search radius 8\,arcsec of at least one radio
-  source. (d) Jackknife map produced from the difference of two maps,
-  using the even and odd scan numbers, respectively. Provided that all
-  noise sources are statistically uncorrelated between the two halves
-  of the data, the map is a plausible realisation of the noise without
-  contamination from astronomical sources. (e) The jackknife map
-  whitened using the same filter as that in panel (b). The whitened
-  jackknife map cross-correlated with the same whitened PSF as in
-  panel (c). Again, orange ``$+$'' and blue circles indicate radio
-  sources and 3.8-$\sigma$ peaks, respectively. Unlike panel (c),
-  there is only a single (apparently) significant peak, and it is not
-  in the vicinity of a radio source.}
+  peaks, 9 are within a search radius of 8\,arcsec of at least one
+  radio source. (d) Jackknife map produced from the difference of two
+  maps, using the even and odd scan numbers, respectively. Provided
+  that all noise sources are statistically uncorrelated between the
+  two halves of the data, the map is a plausible realisation of the
+  noise without contamination from astronomical sources. (e) The
+  jackknife map whitened using the same filter as that in panel
+  (b). The whitened jackknife map cross-correlated with the same
+  whitened PSF as in panel (c). Again, orange ``$+$'' and blue circles
+  indicate radio sources and 3.8-$\sigma$ peaks, respectively. Unlike
+  panel (c), there is only a single (apparently) significant peak, and
+  it is not in the vicinity of a radio source.}
 \label{fig:lockman_maps}
@@
 
-In this final example, we analyze a map of M17 which contains bright,
+In this final example, we analyse a map of M17 which contains bright,
 extended emission. The data are from observation 11 on 20110531 using
@@
 have as large an impact on the small-scale noise as in the baseline
-configuration, although it is still noticeably larger up to scales of
-about 60\,arcsec.}
+configuration.}
 \label{fig:m17_be_ps}
@@
 regions approximately $\sqrt{2}$ shallower in these maps than for the
-full data set. Finally, the cross-linking (positional angles sampled)
-is similar, though not identical in the two halves. Ideally one would
+full data set. Finally, the cross-linking (position angles sampled) is
+similar, though not identical in the two halves. Ideally one would
 have many full maps, as in the case of the Lockman Hole data in the
@@
 power at a range of scales into a relatively empty region of the map.
-It is created by drawing a realisation of noise within an 800\,arcsec
-on-a-side box, with an angular power spectrum $P(k) \propto k^{-3}$,
-which is appropriate for Galactic cirrus clouds
-\citep[e.g.,][]{gautier1992}. It is then filtered again with a
-14.5\,arcsec Gaussian to model the effect of the \scuba\ optical
-response. The RMS of these fluctuations is then normalised to 0.002 pW
-so that they are comparable to the dynamic range of M17 itself. The
-minimum is then subtracted so that the signal is positive. Finally,
+It is created by drawing a realisation of noise from an angular power
+spectrum $P(k) \propto k^{-3}$, which is appropriate for Galactic
+cirrus clouds \citep[e.g.,][]{gautier1992}, within an 800\,arcsec
+on-a-side box. It is then filtered again with a 14.5\,arcsec Gaussian
+to model the effect of the \scuba\ optical response. The RMS of these
+fluctuations is then normalised to 0.002 pW so that they are
+comparable to the dynamic range of M17 itself. The minimum is then
+subtracted so that the signal is positive. Finally, multiplication by
 half a period of a (1+cosine)/2 function is used to roll-off the
@@
 
-The transfer functions $G(k)$ are plotted as thin solid lines in the
-right panel of Fig.~\ref{fig:m17_def_ps}. This formula produces
-extremely noisy values at large frequencies (small scales), and we
-therefore set it to a value of 1 above 0.015\,arcsec$^{-1}$, as well
-as any point in the curve where $P_\mathrm{M}(k)$ exceeds
-$P_\mathrm{S}(k)$ (i.e., $G(k)$ is assumed to be $\le$1). As expected,
-the larger the scale of the filter, the lower the spatial frequency at
-which the map transfer function falls. We then correct the map and
-jackknife PSDs by dividing by $G(k)$ to produce the thick solid, and
-dashed lines in Fig.~\ref{fig:m17_def_ps}b. This shows us that, even
-though the raw noise in Fig.~\ref{fig:m17_def_ps}a is lower at small
-scales when a smaller-scale filter is used (e.g., the red dashed
-line), once we correct for the transfer function, we actually win in a
-\snr\ sense at large scales using the larger-scale filter (the
-corrected noise is lower), as would be expected.
+The transfer functions $G(k)$ are plotted as thin solid lines in
+Fig.~\ref{fig:m17_def_ps}b. This formula produces extremely noisy
+values at large frequencies (small scales), and we therefore set it to
+a value of 1 above 0.015\,arcsec$^{-1}$, as well as any point in the
+curve where $P_\mathrm{M}(k)$ exceeds $P_\mathrm{S}(k)$ [i.e., $G(k)$
+is assumed to be $\le$1]. As expected, the larger the scale of the
+filter, the lower the spatial frequency at which the map transfer
+function falls. We then correct the map and jackknife PSDs by dividing
+by $G(k)$ to produce the thick solid, and dashed lines in
+Fig.~\ref{fig:m17_def_ps}b. This shows us that, even though the raw
+noise in Fig.~\ref{fig:m17_def_ps}a is lower at small scales when a
+smaller-scale filter is used (e.g., the red dashed line), once we
+correct for the transfer function, we actually win in a \snr\ sense at
+large scales using the larger-scale filter (the corrected noise is
+lower), as would be expected.
 
@@
 in small-scale noise as the filter edge is increased is not as drastic
-as in the baseline reduction, and the \snr\ improves at scales
-$\gsim$60\,arcsec. Finally, note that both the map and jackknife white
-noise levels (at scales $\lsim$20\,arcsec) are in excellent agreement,
-unlike the previous example.
+as in the baseline reduction. Finally, note that both the map and
+jackknife white noise levels (at scales $\lsim$20\,arcsec) are in
+excellent agreement, unlike the previous example.
 
@@
 the ``kink'' near 150\,arcsec. Referring to the mask contours in the
-left panel of the third row in Fig,~\ref{fig:m17_jk}, it is clear
-that the map-maker has failed to identify much of the bright, extended
+left panel of the third row in Fig,~\ref{fig:m17_jk}, it is clear that
+the map-maker has failed to identify much of the bright, extended
 emission in the region of the synthetic source. Each area that is not
@@
 this example. Maps of faint extended emission will also suffer
-considerably, as the structures of interest may not clear the \snr\
+considerably, as the structures of interest may lie below the \snr\
 threshold for the mask.
@@
 
-The iterative solution is stopped once convergence in the $\chi^2$
-statistic, and/or variations in the map itself is achieved. This
-enables SMURF to run in a pipeline setting without user interaction
-for a wide variety of observations. Furthermore, the execution times
-are typically shorter than the observation lengths, and memory
-requirements for even the longest \scuba\ observations are within the
-capabilities of single, high-end desktop computers. SMURF can
-therefore provide real-time feedback at the telescope to observers.
+The iterative solution is stopped once convergence in the map itself
+is achieved. This enables SMURF to run in a pipeline setting without
+user interaction for a wide variety of observations. Furthermore, the
+execution times are typically shorter than the observation lengths,
+and memory requirements for even the longest \scuba\ observations are
+within the capabilities of single, high-end desktop computers. SMURF
+can therefore provide real-time feedback at the telescope to
+observers.
 

### 2012-12-06 smurf_paper.tex
minor edits from Douglas, including a change in date format: YYYY-MM-DD

@@
 the sky plus noise, and then the equation is inverted to estimate the
-map as some weighted linear combination of the data that minimises the
+map as some weighted linear combination of the data that minimizes the
 variance.  The first good description of this method appears in
@@
 
-This paper is organised as follows. We first describe the properties
+This paper is organized as follows. We first describe the properties
 of \scuba\ data, including a principal component analysis to reveal
@@
 of both the 450 and 850\,\micron\ focal planes. Each focal plan is
-populated by four ``subarrays'' (labeled s4a--s4d at 450\,\micron, and
+populated by four ``subarrays'' (labelled s4a--s4d at 450\,\micron, and
 s8a--s8d at 850\,\micron), each consisting of 32 columns and 40 rows
@@
   common-mode signals.  (a) Data taken during the S2SRO period,
-  observation 29 on 20100313. There is a strong correlation between
+  observation 29 on 2010-03-13. There is a strong correlation between
   the bolometers and the roughly $\sim$25\,s oscillation in the
@@
   850\,\micron. (b) Data taken with the fully-commissioned instrument,
-  observation 38 on 20111112. Unlike the S2SRO data, there is no
+  observation 38 on 2011-11-12. Unlike the S2SRO data, there is no
   strong signal produced by variations in the fridge temperature. The
@@
   and fewer line features. For this example, the improvement is most
-  striking at 450\,micron\ where the PSD is approximately two
+  striking at 450\,\micron\ where the PSD is approximately two
   orders-of-magnitude lower in the fully-commissioned data at
@@
 In the S2SRO data (Fig.~\ref{fig:bolos_mix}a), observation 29 on
-20100313, both bolometers share significant long-timescale structure
+2010-03-13, both bolometers share significant long-timescale structure
 ($\gsim10$\,s) that appears to be related to variations in the fridge
@@
 Data from the fully commissioned instrument, observation 38 on
-20111112, are shown in Fig.~\ref{fig:bolos_mix}b. Having solved the
+2011-11-12, are shown in Fig.~\ref{fig:bolos_mix}b. Having solved the
 fridge oscillation problem, these data no longer exhibit a correlation
@@
 higher-frequency line features. The $\mathbf{c}(t)$ otherwise drop
-substantially below the individual bolometer PSDs at high-frequency,
+substantially below the individual bolometer PSDs at high frequency,
 as expected if the bolometers are dominated by uncorrelated white
@@
 noise performance (NEFD) is in fact similar before and after the
-upgrades for these two subarrays (s4a at 450\,micron, and s8b at
+upgrades for these two subarrays (s4a at 450\,\micron, and s8b at
 850\,\micron). The main improvements are a reduction in correlated and
@@
 \caption{Evidence for significant magnetic field pickup for
-  observation 16 on 20100228 at 450\,\micron\ (s4a subarray).  The top
+  observation 16 on 2010-02-28 at 450\,\micron\ (s4a subarray).  The top
   panel shows two un-flatfielded (but mean-subtracted and step
@@
 Example data from the 450\,\micron\ subarray s4a where pickup appears
-to be significant (observation 16 on 20100228) are shown in
+to be significant (observation 16 on 2010-02-28) are shown in
 Fig.~\ref{fig:magpickup}. The time-series for two bolometers in the
@@
   3.6 to 7.7) from the fully-commissioned instrument, observation 18
-  on 20110324. Red and blue lines show the 850 and 450\,\micron\
+  on 2011-03-24. Red and blue lines show the 850 and 450\,\micron\
   common-mode signals, corresponding to the left- and right- vertical
@@
 and post-upgrade data sets, respectively), and are dominated at
-low-frequencies ($\lsim$2\,Hz) by scan-synchronous noise, rather than
+low frequencies ($\lsim$2\,Hz) by scan-synchronous noise, rather than
 underlying variations in the atmosphere, or ``sky noise''. However,
@@
 an order-of-magnitude longer (fully-commissioned) data set is plotted
-from observation 18 on 20110324, during highly variable weather
+from observation 18 on 2011-03-24, during highly variable weather
 conditions (450\,\micron\ line-of-sight opacities ranging from 3.6 to
@@
 opacity, which serves as a proxy for atmospheric emission. The weaker
-correlation between the WVM and the 450\,micron\ bolometers is
+correlation between the WVM and the 450\,\micron\ bolometers is
 probably due to the atmosphere being more transmissive at 850\micron,
@@
 
-While the relative importance of sky noise at low-frequencies is
+While the relative importance of sky noise at low frequencies is
 highly time- and scan pattern-dependent, in general it does not have a
@@
 integral part of our map-making strategy) uncorrelated noise remains
-(e.g., solid lines in Fig~\ref{fig:pspec}). As we will see in
+(e.g., solid lines in Fig.~\ref{fig:pspec}). As we will see in
 Section~\ref{sec:pca} and \ref{sec:experimental}, this remaining noise
 does not have a smoothly-varying correlation pattern across the focal
-plane one might expect if it were due to resolved ``clouds'' of
+plane as one might expect if it were due to resolved ``clouds'' of
 atmospheric emission. A high-pass filter with an edge frequency in the
-range $\sim$0.1--1.0\,Hz is usually employed remove this residual
-noise (Section~\ref{sec:components}), which will also remove any sky
+range $\sim$0.1--1.0\,Hz is usually employed to remove this residual
+noise (Section~\ref{sec:components}), and will also remove any sky
 noise that is not already subtracted by the common-mode.
@@
 the same observation that was used in Figs.~\ref{fig:bolos_mix}b and
-\ref{fig:pspec}b. The 9 most significant components are shown (ranked
-by the mean eigenvalues across all working bolometers in both
+\ref{fig:pspec}b. The nine most significant components are shown
+(ranked by the mean eigenvalues across all working bolometers in both
 bands). The top and middle panels for each component show the
@@
 scan-synchronous noise (which is presumably a mixture of
-magnetic-field pickup and brightness variations in the atmosphere)?
+magnetic field pickup and brightness variations in the atmosphere)?
 First, the smoothly-varying gradients in the eigenvalue maps
@@
 telescope motion, yet has an amplitude that varies smoothly across the
-850\,\micron\ subarray, it may represent the PCAs attempt to reproduce
-some non-linear response to magnetic-field and/or atmospheric pickup
+850\,\micron\ subarray, it may represent the PCA's attempt to reproduce
+some non-linear response to magnetic field and/or atmospheric pickup
 
@@
 850\,\micron\ in Component~6). The signals that appear along columns
-are probably related to magnetic-field pickup. Other sources of lines
+are probably related to magnetic field pickup. Other sources of lines
 are thought to be aliased high-frequency noise, and also the 60\,Hz
@@
 is to efficiently use all of the available data to achieve the
-greatest sensitivity that the instrument is capable of. Generally
-speaking, only a small percentage of the data are flagged as unusable
-during map-making beyond the $\sim$70\% detector yield.  As an
-example, for the 450\,\micron\ scan of Uranus that will be presented
-in Section~\ref{sec:point}, during pre-processing: 31.00\% of the data
-(1587/5120 bolometers) are flagged due to data acquisition problems
-(dead or deactivated bolometers, or non-linear flatfields); 0.12\% of
-the data are flagged due to steps (including a small region about the
-steps, Section~\ref{sec:steps}); and 0.83\% if the data are flagged
-because the telescope was stationary
-(Section~\ref{sec:flagbadbol}). Once the iterative solution begins: 55
-spikes are detected using the map-based despiker ($1.2\times10^{-6}$
-\% of the data, Section~\ref{sec:ast}); and 0.72\% of the data are
-flagged as outliers compared to the common-mode
+greatest sensitivity to astrononomical sources that the instrument is
+capable of. Generally speaking, only a small percentage of the data
+are flagged as unusable during map-making beyond the $\sim$70\%
+detector yield.  As an example, for the 450\,\micron\ scan of Uranus
+that will be presented in Section~\ref{sec:point}, during
+pre-processing: 31\% of the data (1587/5120 bolometers) are flagged
+due to data acquisition problems (dead or deactivated bolometers, or
+non-linear flatfields); 0.12\% of the data are flagged due to steps
+(including a small region about the steps, Section~\ref{sec:steps});
+and 0.83\% if the data are flagged because the telescope was
+stationary (Section~\ref{sec:flagbadbol}). Once the iterative solution
+begins: 55 spikes are detected using the map-based despiker
+($1.2\times10^{-6}$ \% of the data, Section~\ref{sec:ast}); and 0.72\%
+of the data are flagged as outliers compared to the common-mode
 (Section~\ref{sec:comgai}).
@@
 in astronomically-useful units.  In this example we reduce a
-450\,\micron\ map of Uranus (observation 26 on 20111017), which is a
+450\,\micron\ map of Uranus (observation 26 on 2011-10-17), which is a
 nearly point-like source for \scuba\ that is commonly used as a
@@
 240\,arcsec\,s$^{-1}$, covering an area of about 50\,arcmin$^2$. The
-full list of dates and observation numbers includes: 20100218, 63, 64,
-65, 70, 71, 72, 90, 91, 92, 97, 98, 99; 20100220, 111, 112, 113, 118,
-128, 129, 130; 20100303, 61, 62, 64, 69, 70, 72, 73, 74, 75; 20100309,
-87, 88, 90, 91; and 20100311, 59, 64, 65, 72. These observations
+full list of dates and observation numbers includes: 2010-02-18, 63, 64,
+65, 70, 71, 72, 90, 91, 92, 97, 98, 99; 2010-02-20, 111, 112, 113, 118,
+128, 129, 130; 2010-03-03, 61, 62, 64, 69, 70, 72, 73, 74, 75; 2010-03-09,
+87, 88, 90, 91; and 2010-03-11, 59, 64, 65, 72. These observations
 represent about 80 per cent of the total data taken as part of the
@@
 assuming that errors due to calibration between the two halves are
-insignificant). The approach we have taken to minimise systematics is
+insignificant). The approach we have taken to minimize systematics is
 to produce the two maps using odd and even scan numbers (i.e., each
@@
 In this final example, we analyse a map of M17 which contains bright,
-extended emission. The data are from observation 11 on 20110531 using
+extended emission. The data are from observation 11 on 2011-05-31 using
 the 850\,\micron\ array. It is a rotating PONG scan covering a
@@
 brightest areas. By the final iteration, most of the obvious
-structures in the data are excluded in the the mask, negative bowling
-is significantly reduced, and the brightest regions are more extended.
+structures in the data are excluded in the mask, negative bowling is
+significantly reduced, and the brightest regions are more extended.
 

### 2012-12-11 smurf_paper.tex
should have put actual files instead of links in 32b1a18cd

--- 2012-12-04 smurf_paper.tex
@@
   Hill, Edinburgh EH9 3HJ\\
-  $^{5}$Institute for Astronomy, University of Edinburgh, Royal
-  Observatory, Blackford Hill Edinburgh, EH9 3HJ}
+  $^{5}$Institute for Astronomy, University of Edinburgh, Royal Observatory, Blackford Hill Edinburgh, EH9 3HJ}
 \begin{document}
@@
   instrument operating on the 15-m James Clerk Maxwell Telescope,
-  nominally consisting of 5120 bolometers in each of two simultaneous
+  consisting of about 5000 bolometers in each of two simultaneous
   imaging bands centred over 450 and 850\,\micron. The camera is
-  operated by scanning across the sky and recording data at a rate of
-  200\,Hz. As the largest of a new generation of multiplexed kilopixel
-  bolometer cameras operating in the (sub)millimetre, \scuba\ data
-  analysis represents a significant challenge.  We describe the
-  production of maps using the Sub-Millimetre User Reduction Facility
-  (SMURF) in which we have adopted a fast, iterative approach to
-  map-making that enables data reduction on single, modern, high-end
-  desktop computers, with execution times that are typically shorter
-  than the observing times.  SMURF is used in an automated setting,
-  both at the telescope for real-time feedback to observers, as well
-  as for the production of science products for the JCMT Science
-  Archive at the Canadian Astronomy Data Centre. Three detailed case
-  studies are used to: (i) explore convergence properties of the
-  map-maker using simple prior constraints (Uranus -- a point source);
-  (ii) achieve the white-noise limit for faint point-source studies
-  (extragalactic blank-field survey of the Lockman Hole); and (iii)
-  demonstrate that our strategy is capable of recovering angular
-  scales comparable to the size of the array footprint (approximately
-  5\,arcmin) for bright extended sources (star-forming region M17).
+  operated by scanning across the sky and records data at a rate of
+  200\,Hz. This rate is much greater than that of previously existing
+  submillimetre cameras, and represents a significant analysis
+  challenge. We describe the production of \scuba\ maps using the
+  Sub-Millimetre User Reduction Facility (SMURF) in which we have
+  adopted a fast, iterative approach to map-making that enables data
+  reduction on single, modern, high-end desktop computers, with
+  execution times that are shorter than the observing times.  SMURF is
+  used in an automated setting, both at the telescope for real-time
+  feedback to observers, as well as for the production of generic
+  science products for the JCMT Science Archive at the Canadian
+  Astronomy Data Centre. Three detailed case studies are used to: (i)
+  explore convergence properties of the map-maker using simple prior
+  constraints (Uranus -- a point source); (ii) achieve the white-noise
+  limit for faint point-source studies (extragalactic blank-field
+  survey of the Lockman Hole); and (iii) demonstrate that our strategy
+  is capable of recovering angular scales comparable to the size of
+  the array footprint (approximately 5\,arcmin) for bright extended
+  sources (star-forming region M17).
 \end{abstract}
@@
 reduction written using the Starlink Software Environment
-\citep{1993ASPC...52..229W,2009ASPC..411..418J}. The details of the
-instrument design, performance, and calibration are given in two
-companion papers: \citet{holland2012} and \citet{dempsey2012}.
+\citep{1993ASPC...52..229W,2009ASPC..411..418J}. The details of the instrument design,
+performance, and calibration are given in two companion papers:
+\citet{holland2012} and \citet{dempsey2012}.
 
@@
 around nearby stars, helping us understand the early stages of planet
-formation.  With $10,240$ nominal detectors \citep[of which $\sim$70\%
-work and are typically useful,][]{holland2012}, \scuba\ is presently
-the largest of a new generation of multiplexed kilopixel
-(sub)millimetre bolometer cameras, which also includes the cameras for
-the South Pole Telescope \citep[SPT,][]{carlstrom2011} and the Atacama
-Cosmology Telescope \citep[ACT,][]{swetz2011}, both dedicated
-experiments for studying anisotropies in the Cosmic Microwave
-Background (CMB) using similar technology, the latter of which uses the
-same time-domain multiplexed readout electronics as \scuba\
-\citep{battistelli2008}.
-
-Submillimetre imaging cameras generally maximise sensitivity using
-bolometers, rather than coherent detectors, which are limited by white
-photon and phonon noise from the instrument and ambient
-backgrounds. The low-frequency noise, however, is typically dominated
-by sources which produce slow variations in the background (e.g.,
-thermal variations within the cryostat, and the atmosphere for
-ground-based cameras), and drifts in the readout electronics. Such
-noise has a power spectrum $\propto 1/f^\alpha$ ($\alpha>0$), and the
-frequency at which it is comparable to the white noise level is called
-the ``$1/f$ knee''. Since the low-frequency noise is largely
-correlated between all, or subsets, of the bolometers in time, it can
-be suppressed during map-making, since astronomical signals have the
-distinct property that they are fixed in a sky reference frame
-(assuming they are not time-varying). If successful, the noise in the
-resulting map is said to be ``white noise limited'', meaning that it
-is uncorrelated spatially, and has an amplitude that scales as the
-$\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the noise-equivalent
-flux-density (the white noise level of a bolometer in 1\,s of
-integration), and $t$ is the amount of integration time in a map
-pixel.
+formation.  With $\sim10,000$ total detectors, \scuba\ is presently
+the largest submillimetre camera in the world, and the fastest
+wide-area ground-based submillimetre imager.
+
+Submillimetre imaging cameras generally use bolometers, rather than
+coherent detectors, to maximise sensitivity. The sensitivity of
+bolometers is limited by white photon and phonon noise from the
+instrument and ambient backgrounds. The low-frequency noise, however,
+is typically dominated by sources which produce slow variations in the
+background (e.g., thermal variations within the cryostat, and the
+atmosphere for ground-based cameras), and drifts in the readout
+electronics. Such noise has a power spectrum $\propto 1/f^\alpha$
+($\alpha>0$), and the frequency at which it is comparable to the white
+noise level is called the ``$1/f$ knee''. Since the low-frequency
+noise is largely correlated between all, or subsets, of the bolometers
+in time, it can be suppressed during map-making, since astronomical
+signals have the distinct property that they are fixed in a sky
+reference frame (assuming they are not time-varying). If successful,
+the noise in the resulting map is said to be ``white noise limited'',
+meaning that it is uncorrelated spatially, and has an amplitude that
+scales as the $\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the
+noise-equivalent flux-density (the white noise level of a bolometer in
+1\,s of integration), and $t$ is the amount of integration time in a
+map pixel.
 
@@
 spectrum, i.e., above the $1/f$ knee; and (ii) to provide good
-``cross-linking'', in which each portion of the map is scanned at a
-range of position angles, again, to help distinguish time-varying
+``cross-linking'', in which each portion of the map is sampled on a
+range of temporal scales, again, to help distinguish time-varying
 noise features from fixed astronomical sources. In the case of \scuba,
@@
 offering scan patterns that cross the sky at a wide range of position
-angles. Such methods are now used by virtually all existing
+angles, ultimately observing each portion of the map at a number of
+different times. Such methods are now used by virtually all existing
 ground-based submillimetre cameras
@@
 instruments that had poorer low-frequency noise performance, and were
-only sensitive to modest angular scales. We note that under some
-circumstances cross-linking may not be essential. For example, SPT,
-which was designed to measure CMB anisotropies, scans entirely in
-azimuth \citep{schaffer2011}. While this strategy limits its ability
-to recover large angular scales transverse to the scan direction, the
-anisotropic filtering in its maps can be accounted for during analysis
-to achieve SPTs focussed objectives \citep[in contrast, the similar
-ACT experiment uses cross-linking to improve its response to
-large-angular scales,][]{das2011}. This approach is less practical as
-a general solution for \scuba, however, which must serve a broader
-range of scientific interests.
+only sensitive to modest angular scales
 
@@
 \citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO
-\citep[e.g.,][]{omont2001}, another camera from the same generation. A
-more recent example is the analysis of SPT data
-\citep{schaffer2011}. Generally speaking, such methods are fast,
-although depending on the science goals and noise properties of the
-data, they may not achieve the best noise performance on the angular
-scales of interest. A related method that has been popular for
+\citep[e.g.,][]{omont2001}, another camera from the same
+generation. Generally speaking, such methods are fast, although
+depending on the science goals and noise properties of the data, they
+may not achieve the best noise performance on the angular scales of
+interest. In more recent years a method that has been popular for
 reducing data in fields of faint point sources with Bolocam
@@
 Thermal Emission Camera \citep[AzTEC, e.g.,][]{scott2008}, is
-principal component analysis (PCA) cleaning. A statistical
+``principal component analysis'' (PCA) cleaning. A statistical
 ``black-box'' removes the most correlated components of the bolometer
-signals, enabling the detection of point-sources close to the
+signals, enabling the detection of point-sources very close to the
 theoretical white-noise limits of the detectors, with reasonable
@@
 (hundreds rather than thousands of detectors). However, PCA cleaning
-is not a good solution for producing maps of extended structures,
-since such sources produce correlated signals amongst many detectors,
-and are removed by this procedure \citep[an exception is the iterative
-PCA approach of][]{aguirre2011}. Furthermore, in the case of \scuba,
-performing PCA on even a single subarray (typically $\sim$900
-detectors) can be prohibitively slow.
+is not a good generic solution for producing maps of extended
+structures, since such sources produce correlated signals amongst many
+detectors, and are removed by this procedure. Furthermore, in the case
+of \scuba\, performing PCA on even a single subarray (1280 detectors,
+nominally) can be prohibitively slow.
 
@@
 fact that it used a differential radiometer instead of
-bolometers). Other descriptions in the experimental CMB literature
-include \citet{tegmark1997} and \citet{stompor2002}, and an
-application to data from SCUBA is described in \citet{borys2004}.  The
-downside to such methods is that they can be both computationally
-expensive and have large memory requirements. While for some
-experimental designs fast iterative methods for the inversion do exist
-without requiring excessive amounts of memory, such as that of
-\citet{wright1996} \citep[which was later implemented for SCUBA
-by][]{johnstone2000}, for the more general map-making problem,
-involving many detectors, matters are significantly complicated both
-by the need to measure the cross spectra for all unique pairs of
-bolometers, as well as performing the inversion itself. A
-maximum-likelihood method was successfully used to make maps from ACT
-data using approximately the same number of detectors as a single
-\scuba\ subarray at a single wavelength. However, the calculation is
-extremely resource intensive and requires a large cluster in order to
-run; a single map can take 10\,CPU\,years \citep{fowler2010}. Perhaps
-the most promising maximum-likelihood method that may one day be
-applied to \scuba\ data is ``SANEPIC'', which produced maps from
-Balloon-borne Large-Aperture Submillimeter Telescope data, involving
-hundreds of bolometers, while correctly incorporating inter-bolometer
-noise correlations \citep[][]{patanchon2008}.
+bolometers). Other descriptions in the experimental cosmic microwave
+background (CMB) literature include \citet{tegmark1997} and
+\citet{stompor2002}, and an application to data from SCUBA is
+described in \citet{borys2004}.  The downside to such methods is that
+they can be both computationally expensive and have large memory
+requirements. While for some experimental designs fast iterative
+methods for the inversion do exist without requiring excessive amounts
+of memory \citep[e.g.,][]{wright1996}, for the more general map-making
+problem, involving many detectors, things are significantly
+complicated both by the need to measure the cross power spectra for
+all unique pairs of bolometers, as well as performing the inversion
+itself. The most promising maximum-likelihood method that may one day
+be applied to \scuba\ data is ``SANEPIC'', which successfully reduced
+maps from the Balloon-borne Large-Aperture Submillimeter Telescope,
+involving hundreds of bolometers, while correctly incorporating
+inter-bolometer noise correlations \citep[][]{patanchon2008}.
 
@@
 noise at any instant in time is uncorrelated with others, and with the
-data from other bolometers), greatly simplifying the inversion step
-that is so complicated in the maximum-likelihood methods. In the
-iterative approach memory usage usually scales linearly with $N$, the
-product of the number of bolometers, $N_\mathrm{b}$, with the number
-of samples in time, $N_\mathrm{t}$. The computation time is the
-product of the number of iterations, $N_\mathrm{iter}$, with the time
-per iteration, $t_\mathrm{iter}$. For the algorithm described in this
-paper, $N_\mathrm{iter}$ is typically in the range $\sim$5--100
-(depending on the \snr\ of large-scale structures, and the size of the
-map), and $t_\mathrm{iter}$ scales as $N_\mathrm{t} \log
-N_\mathrm{t}$, since Fast Fourier Transforms (FFTs) of the time-series
-are typically performed. By comparison, the fastest maximum-likelihood
-methods that account for inter-bolometer correlations have an
-execution time with an $N_\mathrm{b}^2$ dependence, although it is
-possible to limit memory usage to be linear in $N_\mathrm{t}$
-\citep[e.g.,][]{patanchon2008}.
-
-Iterative approaches to map-making have a long history in the
-submillimetre, such as the pipeline recipe for fitting and removing
-baseline drifts in SCUBA scan-map data as the astronomical image
-estimate improved \citep{1999ASPC..172..171J}. The closest modern
-relatives are the Comprehensive Reduction Utility
-\citep[CRUSH,][]{kovacs2008} for the Submillimeter High-Angular
-Resolution Camera 2 (SHARC-2), the pipeline developed for the Bolocam
-Galactic Plane Survey \citep{aguirre2011}, and the Bolometer array
-data Analysis software (BoA) for the Large APEX BOlometer CAmera
-\citep[LABOCA,][]{schuller2012}.
+data from other bolometers), greatly simplifying the measurement of
+noise properties and the inversion step that is so complicated in the
+maximum-likelihood methods. In principle, it should also be possible
+to recover a wider dynamic range in angular scales than possible for
+the simplest direct methods, with an increase in calculation time that
+is merely linear in the number of iterations needed, while using a
+comparable amount of memory. The iterative method of
+\citet{wright1996} specific to chopped (CMB) data was applied to SCUBA
+observations by \citet{johnstone2000}, although in this case the
+chopping removes most of the low-frequency noise, and greatly
+simplifies the problem (at the expensive of large-scale structure in
+the data). More similar to the algorithm described here was the
+iterative pipeline recipe for fitting and removing baseline drifts in
+SCUBA scan-map data as the astronomical image estimate improved
+\citep{1999ASPC..172..171J}. The closest modern relatives of SMURF are
+the Comprehensive Reduction Utility \citep[CRUSH,][]{kovacs2008} for
+the Submillimeter High-Angular Resolution Camera 2 (SHARC-2), the
+Bolometer array data Analysis software (BoA) for the Large APEX
+BOlometer CAmera \citep[LABOCA,][]{siringo2009}, and the pipeline
+developed for the Bolocam Galactic Plane Survey \citep[][which uses
+iterative PCA cleaning to avoid the suppression of large-scale
+structure]{aguirre2011}.
 
@@
 are removed with an iterative high-pass filter. This strategy enables
-SMURF to reduce data on time scales commensurate with the observing
-times (or significantly faster), on single high-end desktop computers
-using the . For reference, all of the data analysis in this paper was
-performed on a machine with a 64-bit 8-core central processing unit
-operating at 2.7\,GHz, and 48\,Gb of memory. SMURF has been
-successfully used as part of a real-time pipeline \citep[based on the
-Observatory Reduction and Acquisition Control - Data Reduction system,
-or ORAC-DR,][]{gibb2005} offering feedback to observers at the
-telescope. The pipeline is also used to generate products for the JCMT
-Science Archive \citep{2011ASPC..442..203E} hosted by the Canadian
-Astronomy Data Centre (CADC). Details of the pipeline design are given
-in \citet{gibb2005} and \citet{2008AN....329..295C}.
+SMURF to reduce data faster than they are taken, on single high-end
+desktop computers, and it has successfully been used as part of a
+real-time pipeline offering feedback to observers at the
+telescope. The pipeline is also used to generate near-science grade
+products for the JCMT Science Archive \citep{2011ASPC..442..203E}
+hosted by the Canadian Astronomy Data Centre (CADC). Details of the
+pipeline design are given in \citet{gibb2005} and
+\citet{2008AN....329..295C}.
 
@@
 (Section~\ref{sec:bolos}), give examples of typical bolometers signals
-(Sections~\ref{sec:bolosignal}), discuss the impacts of magnetic field
-pickup (Section~\ref{sec:magpickup}) and sky noise
-(Section~\ref{sec:skynoise}), and finally use of principal component
-analysis to explore the correlated noise properties of \scuba\ data
-(Section~\ref{sec:pca}).
+(Section~\ref{sec:bolosignal}), and the impact of magnetic field
+pickup (Section~\ref{sec:magpickup}), and finally illustrate the use
+of principal component analysis to explore the correlated noise
+properties of \scuba\ data (Section~\ref{sec:pca}).
 
@@
 s8a--s8d at 850\,\micron), each consisting of 32 columns and 40 rows
-of bolometers. The bolometers are thermal absorbers (with a response
-time constant of $\tau \sim 1$\,ms) coupled to superconducting
-transition-edge sensors (TESs) for thermometry. Temperature variations
-in the TESs produce changing currents, and therefore varying magnetic
-fields, which are detected and amplified using chains of
-superconducting quantum interference devices (SQUIDs), before the
-larger output currents are digitised. Each detector has its own SQUID
-for the first-stage of the amplification, but the remaining stages
-occur within a common chain of SQUIDs for each of the 32 columns. All
-40 rows are read out in sequence at a row-visit frequency of about
-12\,kHz. Such a high sample rate is unnecessary to produce maps, so
-the data are re-sampled to approximately 200\,Hz before writing to
-disk. This rate provides a sample every 3.0\,arcsec, or approximately
-one third of the 450\,\micron\ diffraction-limited full-width at
-half-maximum (FWHM) -- a typical rule-of-thumb for adequately sampling
-a Gaussian point spread function (PSF) -- at the maximum scanning
-speed of 600\,arcsec\,s$^{-1}$. There is an additional 41st row of
-SQUID readouts that are not connected to TESs. These ``dark SQUIDs''
-track non-thermal noise sources that are common to each column's
-amplifier chain. The relationship between the output digitised
-current, $I$, and the input power, $P$, or $dI/dP$ is established
-using flatfield observations immediately prior to science
+of bolometers. The bolometers are thermal absorbers coupled to
+superconducting transition-edge sensors (TESs) for thermometry.
+Temperature variations in the TESs produce changing currents, and
+therefore varying magnetic fields, which are detected and amplified
+using chains of superconducting quantum interference devices (SQUIDs),
+before the larger output currents are digitised. Each detector has its
+own SQUID for the first-stage of the amplification, but the remaining
+stages occur within a common chain of SQUIDs for each of the 32
+columns. All 40 rows are read out in sequence at a row-visit frequency
+of about 12\,kHz. Such a high sample rate is unnecessary to produce
+maps, so the data are re-sampled to approximately 200\,Hz before
+writing to disk. This rate provides a sample every 3.0\,arcsec, or
+approximately one third of the 450\,\micron\ diffraction-limited
+full-width at half-maximum (FWHM) -- a typical rule-of-thumb for
+adequately sampling a Gaussian point spread function (PSF) -- at the maximum scanning speed. There is
+an additional 41st row of SQUID readouts that are not connected to
+TESs. These ``dark SQUIDs'' track non-thermal noise sources that are
+common to each column's amplifier chain. The relationship between the
+output digitised current, $I$, and the input power, $P$, or $dI/dP$ is
+established using flatfield observations immediately prior to science
 observations, in which the output signal is measured throughout a ramp
@@
 astronomical flux units from power involves a correction for
-atmospheric extinction \citep[primarily using the 183\,GHz (1.6\,mm)
-JCMT Water Vapour Monitor to track line-of-sight opacity variations,
-see Section~3 in][]{dempsey2012}, and the application of a flux
-conversion factor (FCF) which is established from regular measurements
-of calibrators such as Uranus \citep[Section~5 in][]{dempsey2012}.
+atmospheric extinction \citep[primarily using the JCMT Water Vapour
+Monitor to track line-of-sight opacity variations, see Section~3
+in][]{dempsey2012}, and the application of a flux conversion factor
+(FCF) which is established from regular measurements of calibrators
+such as Uranus \citep[Section~5 in][]{dempsey2012}.
 
@@
 correlated from bolometer to bolometer) $1/f$ drifts, bolometer power
-spectra exhibit a roll-off at frequencies approaching Nyquist, which
-is due to the anti-aliasing filter that forms part of the 200\,Hz
-re-sampling stage. This anti-aliasing filter also introduces an
-effective lag of $\sim$6\,ms. Adding this lag to the $\sim$1\,ms
-thermal time constant, as well as a negative shift of about $-$2.5\,ms
-caused by a half-sample offset due to the synchronization between the
-pointing and bolometer data, yields a net lag of approximately
-4.5\,ms. There are also line features in the spectra that are thought
-to be produced primarily at frequencies far above the final 200\,Hz
-sample rate, which are aliased to lower frequencies during the
+spectra include a roll-off in their spectra at frequencies approaching
+Nyquist, which is due to the anti-aliasing filter that forms part of
+the 200\,Hz re-sampling stage. There are also line features which are
+thought to be produced at frequencies far above the final 200\,Hz
+sample rate, and are aliased to lower frequencies during the
 multiplexed readout stage before they can be removed by the
@@
 maps, \scuba\ scan strategies have been designed to provide: (i) good
-cross-linking (visiting every point of the mapped area at different
-scan angles); and (ii) minimal accelerations to reduce turn-around
-overheads (which could be quite large given the 600\,arcsec\,s$^{-1}$
-maximum scan speed). For areas larger than the array footprint a
-rectangular ``PONG'' pattern is used, in which the boresight travels
-in approximately straight lines and ``bounces'' off the edges at
-45\,degree angles until the area is uniformly filled in. It is also
-usually combined with a rotation through a number of fixed position
-angles to create a ``rotating PONG'' with even better
-cross-linking. For smaller areas (of order the array footprint, or
-point sources), in which the PONG turn-around overheads would be
-large, a ``constant velocity daisy'' (CV Daisy) is used.  Here the
-telescope moves in a circle, whose centre also slowly traces out a
-small circle. For a more complete description of the \scuba\ observing
-modes, see Section~5 in \citet{holland2012} and also
+cross-linking (visiting every point of the mapped area on several
+different time scales, and at different scan angles); and (ii) minimal
+accelerations to reduce turn-around overheads (which could be quite
+large given the 600\,arcsec\,s$^{-1}$ maximum scan speed). For areas
+larger than the array footprint a rectangular ``PONG'' pattern is
+used, in which the boresight travels in approximately straight lines
+and ``bounces'' off the edges at 45\,degree angles until the area is
+uniformly filled in. It is also usually combined with a rotation
+through a number of fixed position angles to create a ``rotating
+PONG'' with even better cross-linking. For smaller areas (of order the
+array footprint, or point sources), in which the PONG turn-around
+overheads would be large, a ``constant velocity daisy'' (CV Daisy) is
+used.  Here the telescope moves in a circle, whose centre also slowly
+traces out a small circle. For a more complete description of the
+\scuba\ observing modes, see Section~5 in \citet{holland2012} and also
 \citet{2010SPIE.7740E..66K}.
@@
 \begin{figure*}
-\centering
-
-\begin{minipage}[h]{0.495\linewidth}
-\textbf{(a) S2SRO} \\
-
-\includegraphics[width=\linewidth]{bolos_point_mix_s2sro}
-\end{minipage}
-\begin{minipage}[h]{0.495\linewidth}
-\textbf{(b) fully-commissioned} \\
-
-\includegraphics[width=\linewidth]{bolos_point_mix}
-\end{minipage}
-
+\includegraphics[width=0.49\linewidth]{bolos_point_mix_s2sro}
+\includegraphics[width=0.49\linewidth]{bolos_point_mix}
 \caption{A comparison between single bolometer time-series in each of
-  the 450 and 850\,\micron\ bands with the mixing chamber temperature,
-  the line-of-sight 450\,\micron\ opacity derived from the JCMT water
-  vapour monitor, and azimuth/elevation pointing offsets, before and
-  after upgrading the instrument. The grey signals over-plotted in the
-  top two panels show the residual time-series after removing the
-  common-mode signals.  (a) Data taken during the S2SRO period,
-  observation 29 on 20100313. There is a strong correlation between
-  the bolometers and the roughly $\sim$25\,s oscillation in the
-  fridge, but only a minor correlation with the opacity and telescope
-  motion. The nearly flat common-mode subtracted signals show: (i)
-  that most of the low-frequency signal is common to all of the
-  bolometers; and (ii) the non-correlated, and predominantly white
-  noise at 450\,\micron\ is significantly larger than at
-  850\,\micron. (b) Data taken with the fully-commissioned instrument,
-  observation 38 on 20111112. Unlike the S2SRO data, there is no
-  strong signal produced by variations in the fridge temperature. The
-  450\,\micron\ data have minimal correlated low-frequency noise (as
-  evidenced by the lack of a difference between the raw and
-  common-mode subtracted signals), although the 850\,\micron\ data
-  show a common signal correlated with the opacity and telescope
-  motion (note that this scan has a significantly larger amplitude
-  than the S2SRO data set), mostly likely caused by a combination of
-  changing airmass (anti-correlated with opacity), and magnetic field
-  pickup (see Section~\ref{sec:magpickup} and
+  the 450 and 850\,\micron\ bands with the mixing chamber temperature
+  and azimuth/elevation pointing offsets, before and after upgrading
+  the instrument. The grey signals over-plotted in the top two panels
+  show the residual time-series after removing the common-mode
+  signals.  Left: data taken during the S2SRO period, observation 29
+  on 20100313. There is a strong correlation between the bolometers
+  and the roughly $\sim$25\,s oscillation in the fridge, but only a
+  minor correlation with the telescope motion. The nearly flat
+  common-mode subtracted signals show: (i) that most of the
+  low-frequency signal is common to all of the bolometers; and (ii)
+  the non-correlated, and predominantly white noise at 450\,\micron\
+  is significantly larger than at 850\,\micron. Right: data taken with
+  the fully-commissioned instrument, observation 38 on
+  20111112. Unlike the S2SRO data, there is no strong signal produced
+  by variations in the fridge temperature. The 450\,\micron\ data have
+  minimal correlated low-frequency noise (as evidenced by the lack of
+  a difference between the raw and common-mode subtracted signals),
+  although the 850\,\micron\ data show a common signal correlated with
+  the telescope motion (note that this scan has a significantly larger
+  amplitude than the S2SRO data set), mostly likely caused by magnetic
+  field pickup (see Section~\ref{sec:magpickup} and
   Fig.~\ref{fig:magpickup}).}
@@
 \centering
-
-\begin{minipage}[h]{0.495\linewidth}
-\textbf{(a) S2SRO} \\
-
-\includegraphics[width=\linewidth]{pspec_s2sro}
-\end{minipage}
-\begin{minipage}[h]{0.495\linewidth}
-\textbf{(b) fully-commissioned} \\
-
-\includegraphics[width=\linewidth]{pspec}
-\end{minipage}
-
-\caption{Bolometer power spectral densities (PSDs) for the same two
-  data sets (before and after upgrades) used in
-  Fig.~\ref{fig:bolos_mix}. The PSDs have been boxcar smoothed with a
-  width of 0.1\,Hz to reduce the noise slightly and clarify some
-  features. Four of the most sensitive bolometers have been selected
-  at each wavelength, and the flat-fielded and step-corrected (but
-  otherwise raw) PSDs are shown as coloured dotted lines (the blue
-  signals are for the same time series as those shown in
-  Fig.~\ref{fig:bolos_mix}).  The solid black lines are the PSDs of
-  the common-mode signals at each wavelength, and the solid coloured
-  lines show the PSDs of the bolometers once the common-mode is
-  removed.  Finally, the dashed black lines show the spectral shape
-  produced by a point source given the scan speeds for the two
-  observations (120\,arcsec\,s$^{-1}$ in (a), and
-  190\,arcsec\,s$^{-1}$ in (b); for reference, the top horizontal axes
-  shows the conversion from frequency to angular scale. Horizontal
-  dotted lines at $10^{-7}$ and $10^{-8}$\,pW\,Hz$^{-1}$ at 450 and
-  850\,\micron, respectively, are provided as a visual reference for
-  the white-noise levels. In addition to $1/f$ and white noise
-  components, and line features, all of the PSDs exhibit the gradual
-  roll-off of the anti-aliasing filter just below the Nyquist
-  frequency.  (a) For the S2SRO data, particularly at 450\,\micron,
-  there are broad line features in the PSDs at both wavelengths above
-  $\sim$35\,Hz.  At lower frequencies, the bolometer signals exhibit
-  clear $1/f$ knees at approximately 2 and 3\,Hz at 450 and
-  850\,\micron. Common-mode subtraction removes most of the correlated
-  fridge oscillation signal, lowering the $1/f$ knees to approximately
-  0.6 and 1.0\,Hz at 450 and 850\,\micron, respectively. (b) Data from
-  the fully-commissioned instrument tend to have lower $1/f$ knees,
-  and fewer line features. For this example, the improvement is most
-  striking at 450\,micron\ where the PSD is approximately two
-  orders-of-magnitude lower in the fully-commissioned data at
-  0.1\,Hz. The white noise performance, however, is similar for the
-  two subarrays (s4a and s8b) that were used both before and after the
-  upgrades.}
+\includegraphics[width=0.49\linewidth]{pspec_s2sro}
+\includegraphics[width=0.49\linewidth]{pspec}
+\caption{Bolometer power spectral densities for the same two data sets
+  (before and after upgrades) used in Fig.~\ref{fig:bolos_mix}. The
+  PSDs have been boxcar smoothed with a width of 0.1\,Hz to reduce the
+  noise slightly and clarify some features. Four of the most sensitive
+  bolometers have been selected at each wavelength, and the
+  flat-fielded and step-corrected (but otherwise raw) PSDs are shown
+  as coloured dotted lines (the blue signals are for the same time
+  series as those shown in Fig.~\ref{fig:bolos_mix}).  The solid black
+  lines are the PSDs of the common-mode signals at each wavelength,
+  and the solid coloured lines show the PSDs of the bolometers once
+  the common-mode is removed.  Finally, the dashed black lines show
+  the spectral shape produced by a point source given the scan speeds
+  for the two observations (120\,arcsec\,s$^{-1}$ in the left-hand
+  plot, and 190\,arcsec\,s$^{-1}$ in the right-hand plot); for
+  reference, the top horizontal axes shows the conversion from
+  frequency to angular scale. Horizontal dotted lines at $2 \times
+  10^5$ and $1 \times 10^5$\,pW\,Hz$^{-1}$ at 450 and 850\,\micron,
+  respectively, are provided as a visual reference for the white-noise
+  levels. In addition to $1/f$ and white noise components, and line
+  features, all of the PSDs exhibit the gradual roll-off of the
+  anti-aliasing filter just below the Nyquist frequency.  Left: for
+  the S2SRO data there are broad line features in the PSDs at both
+  wavelengths above $\sim$35\,Hz.  At lower frequencies, the bolometer
+  signals exhibit clear $1/f$ knees at approximately 1 and 2\,Hz at
+  450 and 850\,\micron. Common-mode subtraction removes most of the
+  correlated fridge oscillation signal, lowering the $1/f$ knees to
+  approximately 0.2 and 0.7\,Hz at 450 and 850\,\micron,
+  respectively. Right: data from the fully-commissioned instrument
+  tend to have lower $1/f$ knees, and fewer line features. The
+  remaining low-frequency noise, now that the fridge oscillations have
+  been removed, is considerably more independent from bolometer to
+  bolometer than in the S2SRO data (note the larger spread in the
+  dotted lines at low frequencies compared to the left-hand plot,
+  particularly at 450\,\micron). Finally, this example shows that the
+  white noise performance is in fact similar for the two subarrays
+  (s4a and s8b) that were used both before and after the upgrades.}
 \label{fig:pspec}
@@
 the focal plane itself, it is certainly correlated with the
-temperature of the detectors), the line-of-sight atmospheric opacity
-as derived from the JCMT water vapour monitor (WVM), and the telescope
-pointing, for two data sets, before and after the upgrades that
-followed S2SRO.
-
-In the S2SRO data (Fig.~\ref{fig:bolos_mix}a), observation 29 on
-20100313, both bolometers share significant long-timescale structure
-($\gsim10$\,s) that appears to be related to variations in the fridge
-base temperature, although the similarity is clearly greater at
-850\,\micron. Note that the total power in the fluctuations at
-450\,\micron\ are comparable to those at 850\,\micron\ as one might
-expect if there is a comparable varying thermal load from the fridge
-at each wavelength that dominates. We also note that there is no
-obvious strong correlation between the low-frequency signal structure,
-at either wavelength, with the telescope motion. However, there is a
-suggestion that the shorter-timescale behaviour of the opacity is
-anti-correlated with the elevation, as expected.
+temperature of the detectors), and the telescope pointing, for two
+data sets, before and after the upgrades that followed S2SRO.
+
+In the S2SRO data (left panel), observation 29 on 20100313, both
+bolometers share significant long-timescale structure ($\gsim10$\,s)
+that appears to be related to variations in the fridge base
+temperature, although the similarity is clearly greater at
+850\,\micron. In this particular case, the total power in the
+fluctuations at 450\,\micron, are comparable to those at
+850\,\micron. Such behaviour might be expected if there is a
+comparable varying thermal load from the fridge at each wavelength
+that dominates. We also note that there is no obvious strong
+correlation between the low-frequency signal structure, at either
+wavelength, with the telescope motion.
 
@@
 Fig.~\ref{fig:bolos_mix} and remove it, yielding the grey residual
-signals. These residuals are nearly white, although still with
-noticeable long-timescale variations. The white noise is greater at
+signals. These residuals are quite flat, although still with
+noticeable variations. The white noise is also apparent, and larger at
 450\,\micron\ as one would expect from the larger backgrounds compared
@@
 Data from the fully commissioned instrument, observation 38 on
-20111112, are shown in Fig.~\ref{fig:bolos_mix}b. Having solved the
-fridge oscillation problem, these data no longer exhibit a correlation
-with the mixing chamber (however, note that variations are still seen
-in the mixing chamber signal; such variations do not necessarily
-reflect changes in the focal plane temperature). There is minimal
-correlated low-frequency noise in these 450\,\micron\ data, resulting
-in little difference between the raw and common-mode subtracted
-data. The 850\,\micron\ channel, however, exhibits a more significant
-signal that is obviously correlated with the telescope motion. It is
-also correlated amongst many of the detectors, and common-mode removal
-corrects it to a large extent. Part of the reason that this is seen
-here, and not in the S2SRO data set shown, is that the amplitude of
-the scan pattern is larger. This ``scan-synchronous'' noise is
-attributed to a combination of magnetic field pickup, as described in
-Section~\ref{sec:magpickup}, and sky brightness variations due to
-changes in elevation (not to be confused with underlying changes in
-the atmosphere; a homogeneous atmosphere will appear brighter with
-increasing airmass). Also similar to the S2SRO data, an
-anti-correlation between the opacity and elevation is apparent.
+20111112, are shown in the right panel of
+Fig.~\ref{fig:bolos_mix}. Having solved the fridge oscillation
+problem, these data no longer exhibit a correlation with the mixing
+chamber (however, note that variations are still seen in the mixing
+chamber signal; such variations do not necessarily reflect changes in
+the focal plane temperature). There is minimal correlated
+low-frequency noise in these 450\,\micron\ data (note that there is
+almost no difference once the common-mode has been subtracted). The
+850\,\micron\ channel, however, exhibits a more significant signal
+that is obviously correlated with the telescope motion. It is also
+correlated amongst many of the detectors, and common-mode removal
+corrects it to some extent. Part of the reason that this is seen here,
+and not in the S2SRO data set shown, is that the amplitude of the scan
+pattern is larger. Most of this ``scan-synchronous'' noise is
+attributed to magnetic field pickup, as described in
+Section~\ref{sec:magpickup}.
 
@@
 scale assuming the scan speeds of each observation
-(120\,arcsec\,s$^{-1}$ in Fig.~\ref{fig:pspec}a, and
-190\,arcsec\,s$^{-1}$ in Fig.~\ref{fig:pspec}b). The power spectra of
+(120\,arcsec\,s$^{-1}$ in the left-hand plot, and
+190\,arcsec\,s$^{-1}$ in the right-hand plot). The power spectra of
 unresolved point-sources are also shown as dashed lines in each band
@@
 
-In the S2SRO data (Fig.~\ref{fig:pspec}a), the low-frequency noise is
-very correlated between the detectors (note the tight scatter in the
-dotted coloured lines, and their resemblance to the common-mode), due
-to it being dominated by the fridge oscillations. Common-mode
-subtraction is very effective, reducing the $1/f$ knees in both
-wavelengths by about a factor of 5.
-
-Raw data from the fully-commissioned instrument
-(Fig.~\ref{fig:pspec}b) generally have less significant $1/f$ noise as
+In the S2SRO data (left-hand panel of Fig.~\ref{fig:pspec}), the
+low-frequency noise is very correlated between the detectors (note the
+tight scatter in the dotted coloured lines, and their resemblance to the
+common-mode), due to it being dominated by the fridge
+oscillations. Common-mode subtraction is very effective, reducing the
+$1/f$ knees in both wavelengths by about a factor of 5.
+
+Raw data from the fully-commissioned instrument (right-hand panel of
+Fig.~\ref{fig:pspec}) generally have less significant $1/f$ noise as
 compared with the S2SRO data, and it is considerably less correlated
@@
 
-An additional noise source that is significant primarily for wide-area
-scans (and more so during the S2SRO period) is magnetic field
-pickup. Since the bolometer signals are detected through the
-amplification of magnetic fields, any additional changing fields
-within the instrument will add to the noise.
+An additional noise source that is significant in only a small subset
+of the data (and more so during the S2SRO period) is magnetic field
+pickup. Since the bolometer signals are ultimately detected through
+the amplification of magnetic fields, any additional changing fields
+within the instrument will also be detected.
 
@@
 %-------------------------------------------------
-\subsection{Sky noise}
-\label{sec:skynoise}
-%-------------------------------------------------
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{skynoise}
-\caption{Illustration of the dominance of atmospheric variations on
-  long-timescales using 35\,min. of data taken in opaque, and
-  highly-variable weather conditions (line-of-sight 450\,\micron\
-  opacity ranging from 12 to 22) from the fully-commissioned
-  instrument, observation 18 on 20110324. Red and blue lines show the
-  850 and 450\,\micron\ common-mode signals, corresponding to the left-
-  and right- vertical axes, respectively. The black line shows the
-  line-of-sight opacity from the WVM as a proxy for atmospheric
-  emission (arbitrary scale). Bolometer baseline drifts on timescales
-  $\gsim$100\,s are clearly correlated with the atmosphere in this
-  case.}
-\label{fig:skynoise}
-\end{figure}
-
-The data shown in Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec} were
-taken during fairly stable weather conditions (although with different
-450\,\micron\ line-of-sight opacities $\sim$0.9 and 6.5 in the S2SRO
-and post-upgrade data sets, respectively), and are dominated at
-low-frequencies ($\lsim$2\,Hz) by scan-synchronous noise, rather than
-underlying variations in the atmosphere, or ``sky noise''. However,
-only $\sim$100\,s of data were analysed. In Fig.~\ref{fig:skynoise},
-an order-of-magnitude longer (fully-commissioned) data set is plotted
-from observation 18 on 20110324, taken during more opaque, and highly
-variable weather conditions (450\,\micron\ line-of-sight opacities
-ranging from 12 to 22, scaled from the WVM data which are shown as a
-black line). Unlike the earlier plots, it is clear that the bolometer
-baseline drifts on timescales $\gsim$100\,s are tightly correlated
-with the atmospheric opacity, which serves as a proxy for atmospheric
-emission. The weaker correlation between the WVM and the 450\,micron\
-bolometers is probably due to the atmosphere being more transmissive
-at 850\micron, therefore the emission detected at 450\,\micron\ comes
-from water vapour that is closer to the telescope.
-
-While the relative importance of sky noise at low-frequencies is
-highly time- and scan pattern-dependent, in general it does not have a
-major impact on map-making. Upon common-mode removal (which is an
-integral part of our map-making strategy) uncorrelated noise remains
-(e.g., solid lines in Fig~\ref{fig:pspec}). As we will see in
-Section~\ref{sec:pca} and \ref{sec:experimental}, this remaining noise
-does not have a smoothly-varying correlation pattern across the focal
-plane one might expect if it were due to resolved ``clouds'' of
-atmospheric emission. A high-pass filter with an edge frequency in the
-range $\sim$0.1--1.0\,Hz is usually employed remove this residual
-noise (Section~\ref{sec:components}), which will also remove any sky
-noise that is not already subtracted by the common-mode.
-
-%-------------------------------------------------
 \subsection{Principal component analysis}
@@
 Analysis \citep[PCA,][]{laurent2005,scott2008,perera2008}. Here we use
-PCA to further explore correlated signals in \scuba\ data.
+PCA to explore correlated signals in \scuba\ data.
 
@@
 corresponding eigenvalues. The $\mathbf{\xi}_i(t)$ are normalised by
-$[\sum_t \mathbf{\xi}_i^2(t)]^{1/2}$, such that the amplitude of each
-component in a given bolometer signal is stored in the eigenvalues. In
-the earlier analyses mentioned, the low-frequency noise is assumed to
-be encapsulated in those components with the largest
-eigenvalues. Removing the projection of the time series along those
-components then significantly reduces $1/f$ noise while retaining most
-of the (higher-frequency) signal in point-like sources
+$[\sum_t \mathbf{\xi}_i^2(t)]^{1/2}$, such that the root mean square
+(\rms) amplitude of each component in a given bolometer signal is
+stored in the eigenvalues. In the earlier analyses mentioned, the
+low-frequency noise is assumed to be encapsulated in those components
+with the largest eigenvalues. Removing the projection of the time
+series along those components then significantly reduces $1/f$ noise
+while retaining most of the (higher-frequency) signal in point-like
+sources
 
@@
 analysis of the s4a and s8b subarrays for the first $\sim$100\,s of
-the same observation that was used in Figs.~\ref{fig:bolos_mix}b and
-\ref{fig:pspec}b. The 9 most significant components are shown (ranked
+the same observation that was used in the right-hand
+(fully-commissioned) panels of Figs.~\ref{fig:bolos_mix} and
+\ref{fig:pspec}. The 12 most significant components are shown (ranked
 by the mean eigenvalues across all working bolometers in both
@@
 The majority of the correlated signal at both wavelengths in
-Fig.~\ref{fig:pca} is produced by Component~1, in which the
-eigenvector time-series exhibits a roughly periodic signal that
-resembles the scan pattern in Fig.~\ref{fig:bolos_mix}b. While this
-correlated signal appears at both wavelengths, referring to the maps
-of eigenvalues, they are considerably stronger at 850\,\micron\
-(consistent with the visual appearance of the bolometer signals in
-Fig.~\ref{fig:pca}). Also note that the PSD for this eigenvector
-exhibits nearly a pure $1/f$ signature with minimal line features, and
-no white-noise plateau (suggesting that these are high-\snr\
-measurements of a purely low-frequency drift). Component~3 is weaker,
-yet also shows evidence of the scan-synchronous noise, combined with
-high-frequency line features that appear primarily in columns 26, 29
-and 30 of the 450\,\micron\ subarray. Unlike Component~1, it is
-generally more dominant at 450\,\micron.
+Fig.~\ref{fig:pca} is clearly produced by Components~1 and 2. In both
+cases, the eigenvector time-series exhibit a roughly periodic signal
+that resembles the scan pattern in the right-hand panel of
+Fig.~\ref{fig:pca}. While these correlated signals also clearly appear
+at both wavelengths, referring to the maps of eigenvalues, they are
+considerably stronger at 850\,\micron\ (consistent with the visual
+appearance of the bolometer signals in Fig.~\ref{fig:pca}). Also note
+that the PSDs for these eigenvectors exhibit nearly a pure $1/f$
+signature with almost no line features, or a white-noise plateau
+(suggesting that these are high-\snr\ measurements of a purely
+low-frequency drift).
 
@@
 \centering
-\includegraphics[width=0.85\linewidth]{pca}
-\caption{The first nine components from a principal component
-  analysis, ranked by the mean eigenvalues of combined 450 (s4a
+\includegraphics[width=\linewidth]{pca}
+\caption{The first twelve components from a principal component
+  analysis, ranked by the mean eigenvalues, of combined 450 (s4a
   subarray) and 850\,\micron\ (s8b subarray) time-series bolometer
-  data, for the same observation as that used in
-  Figs.~\ref{fig:bolos_mix}b and \ref{fig:pspec}b. For each component
-  the top plot shows the time-series of the eigenvectors normalised by
-  their \rms, the middle plot its PSD, and the bottom coloured panels
-  the eigenvalues for the bolometers at each wavelength (the
-  amplitudes of the eigenvectors in the time-series). For reference,
-  both the mean, $\bar{\lambda}$, and \rms, $\lambda_\mathrm{rms}$,
-  eigenvalues for the bolometers in each subarray are also shown. Most
-  of the correlated $1/f$ signal is encapsulated in Component~1, and
+  data, for the same observation as that used in the right-hand panels
+  of Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec}. For each component
+  the top plot shows the time-series of the normalised eigenvectors,
+  the middle plot its PSD, and the bottom coloured panels the
+  eigenvalues for the bolometers at each wavelength (the amplitudes of
+  the eigenvectors in the time-series). For reference, both the mean,
+  $\bar{\lambda}$, and \rms, $\lambda_\mathrm{rms}$, eigenvalues for
+  the bolometers in each subarray are also shown. Most of the
+  correlated $1/f$ signal is encapsulated in Components~1 and 2, and
   appears to be dominated by scan-synchronous noise (compare with the
-  scan pattern in Fig.~\ref{fig:bolos_mix}b). Some of this signal also
-  appears in Component~3. Component~2 is another $1/f$ signal that is
-  dominant in the 850\,\micron\ subarray. Most of the remaining
-  components consist of weaker $1/f$ noise and line features
-  predominantly above $\sim$30\,Hz. The eigenvalue maps show
-  correlated patterns along columns suggesting pickup in their common
-  amplifier chains.}
+  scan pattern in the right-hand panel of Fig.~\ref{fig:bolos_mix}).
+  Components~3, 4, 7 and 11 are transient glitches that only appear in
+  a small subset of the bolometers. Component~5 also appears in only a
+  few bolometers, and is dominated by a narrow line at 57\,Hz which
+  suggests that it is related to the 60\,Hz mains. Components~6 and
+  8--10 contain a mixture of broad lines above $\sim$30\,Hz (which are
+  probably aliased high-frequency noise), with some low-frequency
+  drifts. Finally, the erratic appearance of the eigenvector for
+  Component~12 is typical of a poorly-biased detector.}
 \label{fig:pca}
@@
 
-Why are there two components that seem to be dominated by
-scan-synchronous noise (which is presumably a mixture of
-magnetic-field pickup and brightness variations in the atmosphere)?
-First, the smoothly-varying gradients in the eigenvalue maps
-(primarily at 850\,\micron) suggests that this noise has a different
-response across the focal plane; in this case the PCA has identified
-two orthogonal shapes that, when mixed in different quantities, can
-reproduce most of the signal in each bolometer. It is also likely that
-the atmospheric contribution is stronger at 450\,\micron\ (i.e.,
-predominantly Component~3).
-
-Like Component~1, Component~2 is a purely $1/f$ signal, with an
-amplitude comparable to Component~3, that appears almost exclusively
-at 850\,\micron. Since it is not obviously correlated with the
-telescope motion, yet has an amplitude that varies smoothly across the
-850\,\micron\ subarray, it may represent the PCAs attempt to reproduce
-some non-linear response to magnetic-field and/or atmospheric pickup
-
-The remaining, weaker, Components~4--9 share some common
-features. They all have some degree of $1/f$ noise, suggesting further
-residual noise related to the telescope motion and/or other slow
-baseline drifts. All of them also exhibit line features in their
-spectra, which in many cases appear at random amplitudes along columns
-(e.g. columns 15, and 26--30 at 450\,\micron\ in Component~4), or in a
-sampling of isolated bolometers (e.g. column 13, row 19 at
-850\,\micron\ in Component~6). The signals that appear along columns
-are probably related to magnetic-field pickup. Other sources of lines
-are thought to be aliased high-frequency noise, and also the 60\,Hz
-mains which appear in some bolometers.
+Why are there two dominant components? First, the smoothly-varying
+pattern in the eigenvalue maps (particularly evident at 850\,\micron)
+suggests that the scan-synchronous noise (presumably magnetic field
+pickup) has a different response across the focal plane; in this case
+the PCA may have identified two orthogonal shapes that, when mixed in
+different quantities, can reproduce most of the signal in each
+bolometer. It is also likely that a varying atmospheric contribution
+is also included in these eigenvectors. One would expect the
+atmosphere to be more significant at 450 than at 850\,\micron,
+although it is not obviously visible due to the strong presence of the
+scan-synchronous noise.
+
+Next, Components~3, 4, 7 and 11 in Fig.~\ref{fig:pca} have a common
+character. In all cases, the PCA appears to have identified strong,
+but transient glitches that appear in only one or several bolometers,
+such as the two black bolometers near column 8, row 21 at
+850\,\micron\ in Component~3. Such events are typically flagged and
+removed during iterative de-spiking (Section~\ref{sec:ast}).
+
+The most interesting feature of Component~5 in Fig.~\ref{fig:pca} is a
+strong narrow line in the PSD at a frequency of about 57\,Hz
+(suggesting the 60\,Hz mains as the culprit). The eigenvalue maps show
+that it is only significant for several bolometers in the
+850\,\micron\ subarray: at column 22 row 22; several bolometers at the
+bottom of column 28; and more near the top-right of the subarray.
+
+Components~6 and 8--10 in Fig.~\ref{fig:pca} have several generic
+features in common. First, the PSDs exhibit moderate low-frequency
+$1/f$ noise. Second, there are broad-line features at frequencies
+above $\sim$30\,Hz which we believe to be aliased high-frequency
+noise, in part due to the clear correlation patterns that appear along
+columns in the eigenvalue maps at both wavelengths (e.g., columns 26
+and 29 at 450\,\micron\ in Component~6). Finally, these components
+have potentially unrelated features mixed into them, like sudden
+offsets that are likely residuals due to imperfect step correction
+(e.g., around $t\sim50$\,s in the eigenvector time-series for
+Component~6; also see Section~\ref{sec:steps}).
+
+Finally, Component~12 in Fig.~\ref{fig:pca} is significant in only a
+handful of detectors in the 850\,\micron\ subarray (e.g., column 21
+row 25). The time-series plot for the eigenvector exhibits a large
+positive tail of sudden excursions, which gradually worsen over
+time. Bolometers that exhibit this signal are probably poorly biased,
+and are usually flagged due to elevated noise levels during
+pre-processing (Section~\ref{sec:flagbadbol}).
 
@@
 the details can vary significantly from data set to data set, and the
-lengths of the time-series analysed. The reason for this is that many
-of the noise sources are not stationary in time (e.g.,
-scan-synchronous noise which obviously depends on the scan pattern,
-and the tuning of the subarrays). It should also be clear from this
-example that while PCA offers some helpful insight into the various
-sources of noise, it does not necessarily identify clean patterns that
-can easily be modelled. For example, while there are clearly
-correlation patterns along columns as expected from the common
+lengths of the time-series analysed. The reason for this is probably
+due to the fact that many of the noise sources are not stationary in
+time (e.g., scan-synchronous noise which obviously depends on the scan
+pattern, and the tuning of the subarrays). It should also be clear
+from this example that while PCA offers some helpful insight into the
+various sources of noise, it does not necessarily identify clean
+patterns that can easily be modelled. For example, while there are
+clearly correlation patterns along columns (Components~6 and 8--10,
+particularly at 450\,\micron), as one might expect given the common
 amplification chain, the intensities of these high-frequency signals
-appear random in the eigenvalue maps, and a simple common-mode signal
-estimated from the data for each column would not remove
-it. Regardless, the dominant noise in excess of the fundamental white
-noise limit, in all cases, is at low-frequencies. This conclusion is
-central to our data reduction strategy described in
-Section~\ref{sec:algorithm}.
+appear random in the eigenvalue maps, and a common-mode signal
+estimated from the data for each column would not remove it. Finally,
+we note that most of the high-frequency ($\gsim2$\,Hz) correlated
+features that have been identified tend to be restricted to a small
+number of detectors, and can often be identified and removed using
+other (faster) means. What remains is predominantly low-frequency
+noise. This conclusion is central to our data reduction strategy
+described in the next section.
 
@@
 Starlink\footnote{\url{http://www.starlink.ac.uk}} is
-open-source\footnote{\url{https://github.com/Starlink}} (distributed
-under the GNU General Public Licence v3), and already used extensively
-at the Joint Astronomy Centre (host of the JCMT) for many other
-systems \citep{jenness2011}, which helps with interoperability. Though
-originally written in FORTRAN, many of the core Starlink libraries are
-now ported to native C, or at least have a C interface. It was
-therefore decided to develop SMURF in C as well (rather than C++, for
-example, which would have required adding further dependencies to
-Starlink), although we have taken an object-oriented approach. For
-example, all of the data for a given signal component model are
-encapsulated in a C structure, and there is a standard interface for
-all functions that handle models (member data and functions for the
-class, respectively). In this way it is easy to extend SMURF with new
-models. Parallelisation is incorporated in the most time-consuming
-low-level routines using threads (e.g., performing Fast Fourier
-Transforms, re-gridding the data), usually handling either independent
-blocks of bolometers, or blocks of time, in each thread, depending on
-the nature of the calculation. We have found that for typical data
-sets the processing time scales well with the number of central
-processing unit (CPU) cores, although beyond 8 the returns are
-diminished due to tasks that cannot be parallelised (e.g., reading the
-data from disk).
+open-source\footnote{\url{https://github.com/Starlink}} (distributed under the GNU General Public Licence v3), and already
+used extensively at the Joint Astronomy Centre (host of the JCMT) for
+many other systems \citep{jenness2011}, which helps with
+interoperability. Though originally written in FORTRAN, many of the
+core Starlink libraries are now ported to native C, or at least have a
+C interface. It was therefore decided to develop SMURF in C as well
+(rather than C++, for example, which would have required adding
+further dependencies to Starlink), although we have taken an
+object-oriented approach. For example, all of the data for a given
+noise model are encapsulated in a C structure, and there is a standard
+interface for all functions that handle models (member data and
+functions for the class, respectively). In this way it is easy to
+extend SMURF with new models. Parallelisation is incorporated in the
+most time-consuming low-level routines using threads (e.g., performing
+Fast Fourier Transforms, re-gridding the data), usually handling
+either independent blocks of bolometers, or blocks of time, in each
+thread, depending on the nature of the calculation. We have found that
+for typical data sets the processing time scales well with the number
+of central processing unit (CPU) cores, although beyond 8 the returns
+are diminished due to tasks that cannot be parallelised (e.g., reading
+the data from disk).
 
@@
   and 4\,arcsec at 450 and 850\,\micron, respectively, to ensure
-  adequate sampling of the \scuba\ PSFs). A cleaning stage repairs
+  adequate sampling of the \scuba\ PSFs). A cleaning stage repairs DC
   steps and spikes, and fills gaps to ensure continuity. Finally, the
@@
 visible. These drifts are removed using a high-pass filter
-(\model{FLT}, Section~\ref{sec:flt}) implemented with FFTs. The
-residual signal is then re-gridded to estimate the map. Finally, the
-map is used to estimate and remove the astronomical signal from the
-bolometer data (\model{AST}, Section~\ref{sec:ast}), leaving a data
-set that is appropriate for measuring the white noise level of each
-bolometer (\model{NOI}, Section~\ref{sec:noi}). Since the common-mode
-and filtering stages will have introduced ringing in the map in the
-vicinity of bright sources, the entire process is iterated. Each
-component in the dashed box of Fig.~\ref{fig:dimm} is re-calculated in
-sequence. In this way, the second time the common-mode is calculated,
-for example, most of the bright astronomical sources in the data have
-already been removed in the previous iteration when the map was
-estimated, reducing the amount of ringing in the map once it is
-re-estimated. Note that the extinction correction is a multiplicative
-factor, and must be inverted prior to re-calculating any of the
-additive model components to preserve the data amplitude. Additive
-model components from the previous iteration are generically added
-back into the time-series immediately prior to their
-re-calculation. However, for the special case of the common-mode and
-high-pass filter stages, they are replaced simultaneously at the start
-of the iteration to assist with convergence (see
-Sections~\ref{sec:converge}).
+(\model{FLT}, Section~\ref{sec:flt}) implemented with Fast Fourier
+Transforms (FFTs). The residual signal is then re-gridded to estimate
+the map. Finally, the map is used to estimate and remove the
+astronomical signal contamination in the bolometer data (\model{AST},
+Section~\ref{sec:ast}), leaving a relatively clean data set in which
+to measure the white noise level of each bolometer (\model{NOI},
+Section~\ref{sec:noi}). Since the common-mode and filtering stages
+will have introduced ringing in the map in the vicinity of bright
+sources, the entire process is iterated. Each component in the dashed
+box of Fig.~\ref{fig:dimm} is re-calculated in sequence. In this way,
+the second time the common-mode is calculated, for example, most of
+the bright astronomical sources in the data have already been removed
+in the previous iteration when the map was estimated, reducing the
+amount of ringing in the map once it is re-estimated. Note that the
+extinction correction is a multiplicative factor, and must be inverted
+prior to re-calculating any of the additive model components to
+preserve the data amplitude. Additive model components from the
+previous iteration are generically added back into the time-series
+immediately prior to their re-calculation. However, for the special
+case of the common-mode and high-pass filter stages, they are replaced
+simultaneously at the start of the iteration to assist with
+convergence (see Sections~\ref{sec:converge}).
 
@@
 iterative algorithm in detail. Section~\ref{sec:converge} explores
-convergence tests and degeneracies between model components. Finally,
-Section~\ref{sec:performance} summarises the mapping performance of
-\scuba\ using SMURF.
+convergence tests and degeneracies between model components.
 
@@
 glitches that may hinder the iterative solution from converging, or
-simply tasks that do not need to be iterated. All of the data that are
-flagged as bad are ignored when estimating the map.
+simply tasks that do not need to be iterated.
 
@@
 samples from $x_i$ in each of the $y_j$, so fractional samples from
-$x_i$ are used at the $y_j$ time boundaries: $y_j = (\sum w_i
-x_i)/(\sum w_i)$, where $i$ runs over all the samples from $x$ that
-land within, or partially overlap, $y_j$, and the weights,
-$0<w_i\le1$, indicate the fractions of each $x_i$ that land within
-$y_j$. The algorithm is fast since each sample in $x_i$ need only be
-visited once. From a spectral point of view, this boxcar average is
-equivalent to applying a sinc-function low-pass filter. Such behaviour
-is desirable since it serves as an anti-aliasing filter; most
-non-white features above the Nyquist frequency in $y_j$ that could
-contaminate the output data are removed.
+$x_i$ are used at the $y_j$ time boundaries. The algorithm is fast,
+since each sample in $x_i$ need only be visited once. From a spectral
+point of view, this boxcar average is equivalent to applying a
+sinc-function low-pass filter. Such behaviour is desirable, since it
+serves as an anti-aliasing filter; most non-white features above the
+Nyquist frequency in $y_j$ that could contaminate the output data are
+removed.
 
@@
 \begin{figure*}
-\raggedright
-\textbf{(a) large step} \hspace{7.3cm} \textbf{(b) small steps}\\
-
-\vspace{0.2cm}
-
+\centering
 \includegraphics[width=\linewidth]{steps1}
-\caption{Examples of steps in time-series data. Steps occur with a
-wide range of heights from the large steps shown in (a) to the small
-steps shown in (b). Large steps are often followed by a brief
-``over-shoot'' as shown in (a) which is ringing caused by the
-anti-aliasing filter. In both plots, the black curve is the
-uncorrected time-series, and the red curve is the corrected time
-series. Samples close to a step are omitted in the corrected
-time-series.  In (b) the blue curve is the uncorrected time-series for
-a nearby bolometer. The similarity between the red and blue curves
-shows that the step correction is performing well.  }
+\caption{Examples of steps in time-series data. Steps occur with a wide
+range of heights from the large steps shown on the left to the small
+steps shown on the right. Large steps are often followed by a brief
+``over-shoot'' as shown on the left. In both plots, the black curve is
+the uncorrected time-series, and the red curve is the corrected time
+series. Samples close to a step are omitted in the corrected time-series.
+In the right hand plot, the blue curve is the uncorrected time-series for
+a nearby bolometer. The similarity between the red and blue curves shows
+that the step correction is performing well.
+}
 \label{fig:steps1}
@@
 steps occur over a single sample, such as the step close to sample
-5000 in Fig.~\ref{fig:steps1}b, but others happen more gradually, such
-as the step close to sample 5300. In addition, a step can be preceded
-or followed by a short period of instability, as is visible at the
-bottom of the step in Fig.~\ref{fig:steps1}a (this is probably due to
-the response of the \scuba\ anti-aliasing filter; the sudden large
-step occurs prior to the 200\,Hz re-sampling). Further problems are
-caused by steps that occur close together in time, such as the large
-downward step followed by a smaller upward step close to sample 5000
-in Fig.~\ref{fig:steps1}b.
+5000 in the right-hand plot of Fig.~\ref{fig:steps1}, but others
+happen more gradually, such as the step close to sample 5300. In
+addition, a step can be preceded or followed by a short period of
+instability, as is visible at the bottom of the step in the left hand
+plot of Fig.~\ref{fig:steps1} (this is probably due to the response of
+the \scuba\ anti-aliasing filter; the sudden large step occurs prior
+to the 200\,Hz re-sampling). Further problems are caused by steps that
+occur close together in time, such as the large downward step followed
+by a smaller upward step close to sample 5000 in the right hand plot
+of Fig.~\ref{fig:steps1}.
 
@@
 that have been flagged as bad for any reason (for instance, due to the
-presence of spikes, steps, or unusual common-mode signal) when need to
-be excluded from this FFT. For this reason, each contiguous block of
-bad data samples is filled with artificial data before taking the
-FFT. A least squares linear fit is performed to the 50 samples
-preceding the block, and a similar fit is performed to the 50 samples
-following the block. These are used to estimate the expected values at
-the start and end of the block of bad values. The bad values are then
-replaced by linear interpolation between the expected start and end
+presence of spikes, steps, or unusual common-mode signal) need to be
+excluded from this FFT. For this reason, each contiguous block of bad
+data samples is filled with artificial data before taking the FFT. A
+least squares linear fit is performed to the 50 samples preceding the
+block, and a similar fit is performed to the 50 samples following the
+block. These are used to estimate the expected values at the start and
+end of the block of bad values. The bad values are then replaced by
+linear interpolation between the expected start and end
 values. Gaussian noise is added with a standard deviation equal to the
-mean of the RMS residuals in the two fits. These flagged and filled
-portions of the data are then given a weight of zero when estimating
-the map.
+mean of the RMS residuals in the two fits.
 
@@
 samples modified at each end of the time-series is given by half the
-ratio of the sampling frequency to the lowest retained frequency, the
-edge frequency of the high-pass filter (Section~\ref{sec:flt}). In
+ratio of the sampling frequency to the lowest retained frequency. In
 addition, each end of the time-series is padded with double this
@@
 
-\subsubsection{Additional data rejection}
+\subsubsection{Additional bad bolometer rejection}
 \label{sec:flagbadbol}
@@
 higher-frequency portion of the PSD is generally quite
-flat. Furthermore, even if there is contamination, the purpose of this
-measurement is to identify outliers, rather than provide a meaningful
-absolute measurement of a given bolometer's white noise level. Both
-high and low (e.g., due to an incorrect and very small flatfield
-correction being applied) outliers from the centre of the logarithm
-(to reduce the impact of outliers) of the bolometer noise distribution
-are flagged.
-
-Finally, data that are taken while the telescope is either moving
-extremely slowly, or extremely fast, are flagged and ignored when
-estimating the map. When the telescope is moving slowly, astronomical
-signals appear at low frequencies in the bolometer data that are
-dominated by $1/f$ noise (in the extreme case of no motion, the
-bolometers have zero sensitivity). When the telescope is moving
-significantly faster than 600\,arcsec\,s$^{-1}$, both the thermal
-response of the bolometers, and the anti-aliasing filter in the
-readout electronics (Section~\ref{sec:bolos}) can significantly
-distort signals.  The default allowable range of speed is
-30--1000\,arcsec\,s$^{-1}$. These thresholds primarily reject short
-periods when the telescope was stationary prior to commencing the scan
-pattern (generally speaking the speed is constant once the scan
-begins). The scan speed can also briefly exceed the upper limit, even
-for a requested speed of 600\,arcsec, if the pointing system is
-struggling for large-area scans at high elevation.
-
-\subsubsection{Impact of lags}
-\label{sec:lags}
-
-In principle, the effects of the bolometer thermal response,
-anti-aliasing filter, and data acquisition lags should be accounted
-for; together they produce a net lag in the bolometer signals of about
-4.5\,ms (Section~\ref{sec:bolos}). This can be modelled by convolving
-the ``pure'' bolometer signals with a single system response
-function. Similarly, this effect could be removed using de-convolution
-as a pre-processing step. The impact of this lag is primarily a
-reduction in point-source sensitivity since scans at different
-position angles will detect the source at slightly different positions
-(and therefore reduce the peak signal once averaged together). This
-attenuation is most significant at 450\,\micron\ given the smaller
-beam size, and at the greatest scan speeds. For reference, at a speed
-of 150\,arcsec\,s$^{-1}$ (typical for CV daisies of calibrators and
-deep fields) the attenuation is a negligible $<$1\% at 850\,\micron,
-and $\sim$1\% at 450\,\micron. However, at the maximum speed of
-600\,arcsec\,s$^{-1}$ (for the largest, and typically shallowest
-maps), this attenuation grows to about 5\% and 15\% at 850 and
-450\,\micron, respectively. It is unlikely that this response has had
-a significant impact on observations taken to date, since such scans
-have been used primarily for shallow cosmology fields for which only
-the 850\,\micron\ data are expected to be useful, or Galactic fields
-of extended structures (which are less sensitive to this effect). In
-the present Starlink \textsc{kapuahi} release there is no option to
-de-convolve the system response. However, it is presently being added
-and will be present in future releases.
+clean. Furthermore, even if there is contamination, the purpose of
+this measurement is to identify outliers, rather than provide a
+meaningful absolute measurement of a given bolometer's white noise
+level. Both high and low (e.g., due to an incorrect and very small
+flatfield correction being applied) outliers from the centre of the
+logarithm (to reduce the impact of outliers) of the bolometer noise
+distribution are flagged.
 
@@
 
-First, we describe the basic model for \scuba\ data. We express the
+First, we describe the generic model for \scuba\ data. We express the
 digitised signal observed by the $i$th bolometer as a function of
@@
 During pre-processing, the map-maker divides by $f_i$ once. Then, in
-each iteration, the map-maker models and removes
+each iteration, the map-maker divides by $f_i$, models and removes
 $g_i\mathbf{n}^\mathrm{c}(t)$, divides by $\mathbf{e}_i(t)$, and
@@
 \includegraphics[width=\linewidth]{com}
-\caption{A selection of bolometer time-series from a representative
-region of the focal plane after flatfielding and removal of a constant
-baseline. Each sub-plot shows the power from a single bolometer over a
-30\,s interval.  The lower and upper vertical limits of each plot are
-$-0.02$\,pW and $+0.02$\,pW, respectively. It can be seen that most
-bolometers exhibit a common time variation overlaid with other
+\caption{A selection of typical bolometer time-series after
+flatfielding and removal of a constant baseline. It can be seen that
+most bolometers exhibit a common time variation overlaid with other
 features. Empty squares indicate the locations of broken bolometers.}
@@
 opposite signs (such that one or the other bolometer would be
-erroneously flagged). Experimental models that were investigated for
-magnetic field pickup removal are described in
-Section~\ref{sec:experimental}.
+erroneously flagged).
 
@@
 \model{COM} model at block boundaries.  These in turn can lead to
-ringing when filtering, and instabilities in the convergence
-process.
+ringing when filtering, and instabilities in the convergence process.
 
@@
 contribute signal to some bolometers but not other bolometers, thus
-biasing the simple mean used to estimate the common mode. However, on
+biassing the simple mean used to estimate the common mode. However, on
 each iteration of the map-making algorithm illustrated in
 Fig.~\ref{fig:dimm}, a large fraction of the remaining astronomical
-signal is extracted from the bolometer time-series and transferred to
-the output map, resulting in subsequent estimates of the common-mode
-being more accurate. It is also for this reason that the previously
-described step of (re-)flagging outlier portions of the data is
-generally performed iteratively, rather than once as a pre-processing
-step, since bright/compact astronomical sources can be the cause of
-rejection in early iterations.
+signal is extracted from the bolometer time-series and transferred to the
+output map, resulting in subsequent estimates of the common-mode being
+more accurate.
 
@@
 For this reason, the usual practice is to estimate a single
-\model{COM} model by examining data from all four subarrays in each
+\model{COM} model by examining data from all four sub-arrays in each
 waveband, since this allows spatial structure on the scale of the
@@
 The extinction correction is a multiplicative factor that is normally
-derived using the WVM, and is not considered to be a free parameter in
-the solution [$\mathbf{e}_i(t)$ in Eq.~\ref{eq:model}]. However, it is
-applied as part of the iterative solution, rather than a
-pre-processing step, since any small errors will be amplified by the
-large low-frequency drifts in the raw bolometer time-series. For
-example, if the bolometer drift is 1000 times greater than an
-astronomical source of interest, a 1 per cent error in the flatfield
-will produce stripes of order 10 times the astronomical signal
-amplitude in the final map! If \model{EXT} is applied after the bulk
-of the low-frequency noise has been removed (e.g., \model{COM,GAI}),
-then there is little potential for such small errors to affect the
-final map.  For details on how it is calculated see
-\citet{dempsey2012}. Note that its numerical value is calculated only
-once, and simply applied as a scale factor in the iterative
+derived using the JCMT water vapour monitor (WVM), and is not
+considered to be a free parameter in the solution [$\mathbf{e}_i(t)$
+in Eq.~\ref{eq:model}]. However, it is applied as part of the
+iterative solution, rather than a pre-processing step, since any small
+errors will be amplified by the large low-frequency drifts in the raw
+bolometer time-series. For example, if the bolometer drift is 1000
+times greater than an astronomical source of interest, a 1 per cent
+error in the flatfield will produce stripes of order 10 times the
+astronomical signal amplitude in the final map! If \model{EXT} is
+applied after the bulk of the low-frequency noise has been removed
+(e.g., \model{COM,GAI}), then there is little potential for such small
+errors to affect the final map.  For details on how it is calculated
+see \citet{dempsey2012}. Note that its numerical value is calculated
+only once, and simply applied as a scale factor in the iterative
 solution. Unlike the additive model components, it is inverted at the
@@
 average of the bolometer data samples $\mathbf{b}_j$ that land within
-that pixel (from any bolometer or point in time, provided that they
-are not flagged as bad or gap-filled, Section~\ref{sec:gaps}),
+that pixel (from any bolometer or point in time),
 %
@@
 floating point precision errors can cause the difference to be
-significantly incorrect. In practice we use the superior ``weighted
-incremental algorithm'' that calculates incremental differences, as
-described in \citet{west1979}.
+significantly incorrect. In practice we use the more numerically
+stable ``weighted incremental algorithm'' that calculates incremental
+differences, as described in \citet{west1979}.
 
@@
 time-domain despiker (Section~\ref{sec:timedespike}), this calculation
-utilises the scatter in the population of samples that land in a map
-pixel (from different times and bolometers) to reject outliers. This
-method is more robust against false-positive detections of
-bright/compact astronomical sources since transient features in the
-time-series are unlikely to occur by chance whenever a bolometer
+utilises the scatter in the population of samples used to estimate the
+map pixel values (from different times and bolometers) to reject
+outliers. This method is more robust against false-positive detections
+of bright/compact astronomical sources, since transient features in
+the time-series are unlikely to occur by chance whenever a bolometer
 crosses a specific location on the sky, whereas real astronomical
@@
 The primary purpose of the noise component, \model{NOI}, is to measure
-the white noise levels of each bolometer, which is approximated with a
-single (non-time varying) variance, $\sigma_i^{\mathbf{w}2}$. This
-value is then used to calculate weights for map estimation,
-$\mathbf{w}_i = 1/\sigma_i^{\mathbf{w}2}$. The measurement generally
-occurs once all of the other models have been fit and removed [i.e.,
-once $\mathbf{n}^\mathrm{w}_i(t)$ from Eq.~\ref{eq:noise} is
-isolated].
+the white noise levels of each bolometer [$\mathbf{n}^\mathrm{w}_i(t)$
+in Eq.~\ref{eq:noise}]. The measurement generally occurs once all of
+the other models have been fit and removed. This noise may then be
+used to estimate weights for the data in subsequent iterations.
 
 First, the bolometer PSDs are calculated as in Eq.~\ref{eq:psd}. An
-average white noise level is then measured from 2 to 10\,Hz, a clean
-region of the PSD that tends to lie above the $1/f$ knee, but below
-the high-frequency line features for typical bolometer data
-(Fig.~\ref{fig:pspec}). Taking this constant level for $\mathbf{P}(f)$
-we then calculate the expected variance of the time-series (in
-approximately 200\,Hz samples) using Eq.~\ref{eq:psd}. If the
-bolometer noise were produced purely by uncorrelated sources (i.e., no
-other long time-scale drifts), with no high-frequency line features,
-and without the attenuation at even higher frequencies by the
-anti-aliasing filter, this is the theoretical noise limit of the
-detectors. These measured variances are stored, and then used in
-subsequent iterations to weight the data points when calculating the
-map estimate (Section~\ref{sec:ast}). They are also used for
-convergence tests (Section~\ref{sec:converge}).
+average white noise level is then measured from 2 to 10\,Hz, a
+relatively clean region of the PSD that tends to lie above the $1/f$
+knee, but below the high-frequency line features for typical bolometer
+data (Fig.~\ref{fig:pspec}). Taking this constant level for
+$\mathbf{P}(f)$ we then calculate the expected variance of the
+time-series (in approximately 200\,Hz samples) using
+Eq.~\ref{eq:psd}. If the bolometer noise were produced purely by
+uncorrelated sources (i.e., no other long time-scale drifts), with no
+high-frequency line features, and without the attenuation at even
+higher frequencies by the anti-aliasing filter, this is the
+theoretical noise limit of the detectors. These measured variances are
+stored, and then used in subsequent iterations to weight the data
+points when calculating the map estimate (Section~\ref{sec:ast}). They
+are also used for convergence tests (Section~\ref{sec:converge}).
 
@@
 \model{NOI} may optionally perform some cleaning options, such as the
-step fixer (Section~\ref{sec:steps}) and spike detection
+DC step fixer (Section~\ref{sec:steps}) and spike detection
 (Section~\ref{sec:timedespike}), which may work better with these
@@
 
-\subsubsection{Experimental models for the removal of magnetic field
-  pickup and the atmosphere}
-\label{sec:experimental}
-
-Additional models exist as options, although they are not generally
+\subsubsection{Other experimental models}
+
+Additional models exist as options, altough they are not generally
 used: \model{DKS}, the use of dark squids as a template for removing
@@
 each column as a template that is simply scaled (gain and offset) to
-each bolometer time-series before removal. Unfortunately the dark
-squids do not work for every column (several cases in each subarray),
-meaning that those columns are usually discarded when producing
-maps. One possible remedy is to identify dead bolometers (with
-otherwise working TES readouts) or intentionally disconnect working
-bolometers, to create replacement dark readouts in those columns. This
-solution has not been pursued due to the limited success at removing
-pickup using the presently working dark squids. Another alternative
-model is \model{TMP} in which the azimuth of the telescope itself is
-used as the template. For both \model{DKS} and \model{TMP}, there is
-some success at visibly decreasing the low-frequency scan-synchronous
-noise in certain data sets, although often it does not (and there are
-also examples in which the noise is increased). In the case of
-\model{DKS}, there may simply be components of the signal seen by
-working detectors that are not apparent in the dark
-squids. Furthermore, in the case of \model{TMP}, the relationship
-between the projection of the Earth's magnetic field on the instrument
-and the magnitude of the pickup may not be linearly correlated as we
-have assumed (verified in some cases by a comparison of the azimuthal
-motion with working dark squid signals). In no case was the use of
-these models able to reduce the $1/f$ knee substantially,
-necessitating the continued use of \model{FLT} to remove the remaining
-low-frequency noise. Neither \model{DKS} nor \model{TMP} are typically
-used.
+each bolometer time-series before removal. Since not all columns have
+working dark squids, another alternative model is \model{TMP} in which
+the azimuth of the telescope itself is used as the template. In both
+cases, while the models had some success at removing the magnetic
+field pickup, they did not reduce the $1/f$ knee substantially,
+necessitating the continued use of the \model{FLT} model to remove the
+remaining low-frequency noise. Little or no difference in the final
+maps could be seen with or without the inclusion of these models, and
+so they are not generally used.
 
@@
 the noise in final maps). This result is basically consistent with
-those from earlier instruments \citep[e.g., the prediction for \scuba\
-based on SCUBA data described in][]{chapin2002}, and suggests that the
-angular scale of the emitting regions in the atmosphere are
-un-resolved at the \scuba\ focal plane. \citet{sayers2010}, in
-particular, examined several different ways of modelling and removing
-the correlated atmospheric noise from Bolocam data at 143 and 268\,GHz
-(also atop Mauna Kea), finding only marginal improvements by fitting a
-plane, or even higher-order polynomials (see their figure~10, as well
-as the discussion and references to earlier work in their
-section~4.3). In \citet{aguirre2011} iterative PCA cleaning was used
-to remove the atmosphere since the simpler methods mentioned left
-substantial residuals in their data (similar to the \scuba\ data
-described here, although PCA cleaning is prohibitively slow in our
-case).
+those from earlier instruments, and suggests that the angular scale of
+the emitting regions in the atmosphere are un-resolved at the \scuba\
+focal plane. \citet{sayers2010}, in particular, examined several
+different ways of modelling and removing the correlated atmospheric
+noise from Bolocam data at 143 and 268\,GHz (also atop Mauna Kea),
+finding only marginal improvements by fitting a plane, or even using
+higher-order polynomials (see their figure~10, as well as the
+discussion and references to earlier work in their section~4.3). In
+\citet{aguirre2011} iterative PCA cleaning was used to remove the
+atmosphere since the simpler methods mentioned left substantial
+residuals in their data (similar to the \scuba\ data described here,
+although PCA cleaning is prohibitively slow in our case).
 
@@
 
-The map-maker will halt after a user-specified number of iterations,
-or once some convergence criterion has been achieved. Presently two
-numerical quantities are tracked after each iteration: the change in
-reduced chi-squared, $\chi^2_\mathrm{r}$; and the change in the
-map. While convergence in $\chi^2_\mathrm{r}$ is not, ultimately, a
-good indicator of when the solution should be stopped, we discuss its
-calculation and properties first as an introduction to the problem of
-model degeneracies.
-
-The bolometer residual variances, $\sigma^2_i$, are measured once the
-modelled signal components have been removed. In the early iterations,
-the residuals contain both white noise, and other long-timescale
-features. However, when the solution has converged, this signal should
-look white. Taking the component of bolometer variances due to white
-noise, $\sigma_i^{\mathbf{w}2}$ (measured from the bolometer PSDs in
-\model{NOI}, Section~\ref{sec:noi}), $\chi^2_\mathrm{r}$ is calculated
-as $(1/N) \sum_i \sigma^2_i / \sigma_i^{\mathbf{w}2}$, where $i$ runs
-over the $N$ bolometers. In other words, it is the average ratio
-between the measured time-series bolometer variances and their
-white-noise levels measured between 2 and 10\,Hz, and should tend to a
-value of $\sim$1 if the low-frequency noise (and any bright
-astronomical signals) have been successfully removed. However, this
-expression does not account for the degrees of freedom in the
-model. Clearly there are a large number of parameters that will
-``fit-out'' some of the uncorrelated white noise, and bias this
-estimate of $\chi^2_\mathrm{r}$ low. Indeed, for a converged map,
-values in the range $\sim$0.85--0.95 are common, depending on the
-precise configuration parameters. Regardless of the normalisation,
-once the model parameterisation has been chosen, this quantity should
-converge to some fixed value.
-
-Practically speaking, $\chi^2_\mathrm{r}$ tends to converge before the
-map due to model degeneracies. One simple example is the degeneracy
-between large-scale astronomical structures (larger than the array
-footprint), and the common-mode rejection step (\model{COM}), which
-will be illustrated in Section~\ref{sec:point}. In this case, there is
-a significant space of maps with spurious large-scale structures that
-are anti-correlated with features in the common-mode, and the solution
-is free to wander this space while producing flat residuals.
-
-Instead, we typically use a map-based convergence statistic,
-$M_\mathrm{c}$, the average absolute change in the value of map pixels
-between subsequent iterations, normalised by the map pixel
-uncertainties (square root of Eq.~\ref{eq:varmap}), or
+The map-maker will halt if the convergence criteria have been
+met. Presently there are two numerical tests that may be performed by
+SMURF, or alternatively, a fixed number of iterations may be
+specified. The first numerical test is an approximate check of the
+change in reduced chi-squared, $\chi^2_\mathrm{r}$. The standard
+deviations, $\sigma_i$, are measured for each of the residual
+bolometer signals once the modelled signal components have been
+removed. In the early iterations, the residual signals usually contain
+both white noise, and other long-timescale features. However, once the
+solution has converged, this signal should look approximately
+white. Once the white noise, $\mathbf{n}^\mathrm{w}_i(t)$, has been
+measured from the bolometer PSDs in \model{NOI}
+(Section~\ref{sec:noi}), $\chi^2_\mathrm{r}$ is calculated as $(1/N)
+\sum_i \sigma^2_i / \mathbf{n}^\mathrm{w}_i(t)$, where $i$ runs over
+the $N$ bolometers. In other words, it is the average ratio between
+the measured time-series bolometer variances and their white-noise
+levels measured between 2 and 10\,Hz. Note that this expression does
+not account for the degrees of freedom. Clearly there are a large
+number of parameters in the model, which will ``fit-out'' some of the
+uncorrelated white noise, and potentially bias this estimate of
+$\chi^2_\mathrm{r}$ low. One approach would be to keep track of the
+degrees of freedom, and therefore provide a correction, as in
+\citet{kovacs2008}. However, note that we are only using this quantity
+as a convergence test, and its value is not important in an absolute
+sense. Furthermore, the reference white noise values are calculated
+only once in SMURF (Section~\ref{sec:noi}), and the number of model
+parameters are fixed, so the value of $\chi^2_\mathrm{r}$ should
+decrease monotonically if the solution is behaving well. The
+convergence criterion is met if the change in subsequent values of
+$\chi^2_\mathrm{r}$ is smaller than the requsted threshikd, which is
+set to $10^{-3}$ by default.
+
+It was immediately found in the analysis of \scuba\ data that the
+$\chi^2_\mathrm{r}$ test described above is not a sufficient stopping
+criterion. The reason for this is that various components of the model
+are, under normal circumstances, highly degenerate. One simple example
+is the degeneracy between large-scale astronomical structures (larger
+than the array footprint), and the common-mode rejection step
+(\model{COM}), which will be illustrated in
+Section~\ref{sec:point}. Since the flatness of the residual only
+reflects the ability of the model to fit the shape of the bolometer
+time-series, it does not necessarily correlate with convergence of the
+most important model component: the map. Indeed, examination of the
+map after large numbers of iterations (e.g., $>10$), shows the
+presence of large-scale structures that are anti-correlated with
+structures in other model components. For this reason, a second
+map-based convergence test was added.
+
+The map-based convergence statistic, $M_\mathrm{c}$, is the average
+absolute change in the value of map pixels between subsequent
+iterations, normalised by the map pixel uncertainties (square root of
+Eq.~\ref{eq:varmap}), or
 %
@@
 where $i$ runs over the $N$ map pixels, and $j$ enumerates the
-iterations. Experimentally we found that a change in this quantity
-$<$0.05 (on average, map pixels change by $<$5\% of the estimated map
-RMS in subsequent iterations) is a good point to stop, providing
-correspondence with what we would choose ``by eye''. Letting the
+iterations. The convergence criterion is met once this average change
+is smaller than some threshold. Experimentally we found that a value
+of 0.05 gives good results (on average, map pixels change by 5\% of
+the estimated map RMS in subsequent iterations).  This provides good
+correspondence with what we would choose ``by eye'', and letting the
 solver run for many more iterations in several test cases yields
@@
 
-Another major source of divergence is correlation between \model{COM}
-and \model{FLT}. Since \model{FLT} usually consists of a high-pass
-filter following the application of \model{COM}, \model{COM} is
-completely free to grow any large-scale structure at frequencies below
-the chosen filter edge. While such structure does not appear in the
-map (as it is removed by \model{FLT}), we found that the solution
-could be made to converge significantly faster by ``re-mixing''
-\model{COM} and \model{FLT}. Immediately prior to the calculation of
-\model{COM}, the values of \model{COM} and \model{FLT} from the
-previous iteration are added back into the residual simultaneously. In
-this way, truly common-mode signals, even at low-frequencies, do not
-leak into \model{FLT}.
-
-In order to control the degeneracy between low-frequency signal that
-is removed, and large-scale structures in the map, we have developed a
-simple system for constraining regions of the map devoid of sources to
-a value of zero for all but the final iteration in the \model{AST}
-model (Section~\ref{sec:ast}). Such regions are either user-defined in
-advance, or can be determined from the data using a cut on \snr. This
-technique is explored considerably in the examples in
-Sections~\ref{sec:point} and \ref{sec:extended}.
-
-%-------------------------------------------------
-\subsection{Instrument and map-making performance}
-\label{sec:performance}
-%-------------------------------------------------
-
-From a scientific perspective, the most important goal of map-making
-is to efficiently use all of the available data to achieve the
-greatest sensitivity that the instrument is capable of. Generally
-speaking, only a small percentage of the data are flagged as unusable
-during map-making beyond the $\sim$70\% detector yield.  As an
-example, for the 450\,\micron\ scan of Uranus that will be presented
-in Section~\ref{sec:point}, during pre-processing: 31.00\% of the data
-(1587/5120 bolometers) are flagged due to data acquisition problems
-(dead or deactivated bolometers, or non-linear flatfields); 0.12\% of
-the data are flagged due to steps (including a small region about the
-steps, Section~\ref{sec:steps}); and 0.83\% if the data are flagged
-because the telescope was stationary
-(Section~\ref{sec:flagbadbol}). Once the iterative solution begins: 55
-spikes are detected using the map-based despiker ($1.2\times10^{-6}$
-\% of the data, Section~\ref{sec:ast}); and 0.72\% of the data are
-flagged as outliers compared to the common-mode
-(Section~\ref{sec:comgai}).
-
-The white noise levels of \scuba\ bolometers are measured between 2
-and 10\,Hz in the PSDs (as described in Section~\ref{sec:noi}),
-leading to noise equivalent powers (NEPs, with units W\,s$^{1/2}$) --
-the \rms\ time-series noise in a 1\,s integration \citep[see section
-3.5.2 in][]{holland2012}. The NEPs are then converted into more useful
-noise equivalent flux densities (NEFDs, with units Jy\,s$^{1/2}$)
-through multiplication by the extinction correction and the FCF. For
-an ideal scan pattern in which every bolometer spends an equal amount
-of time observing each point on the sky, the expected noise in a map
-pixel is $\mathrm{NEFD_{eff}}/t^{1/2}$, where $t$ is the integration
-time, and the effective NEFD, $\mathrm{NEFD_{eff}} = (1/\sum
-\mathrm{NEFD}_i^{-2})^{1/2}$, is the sensitivity of the entire array
-including all bolometers $i$. Experimentally it has been found that
-the map noise based on the weighted scatter of the data
-(Eq.~\ref{eq:varmap}), and the scatter of map pixels in small
-apertures, both correlate well with this expected limiting noise
-performance for maps that are larger than the array footprint (for
-smaller maps the assumption of uniform coverage is incorrect). As an
-example of observing under favourable conditions with line-of-sight
-opacities of 0.87 and 0.34 at 450 and 850\,\micron\ \citep[zenith sky
-opacities at 225\,GHz of 0.04 and 0.065 at 450 and 850\,\micron,
-respectively, at an average airmass of 1.2, from table 3
-in][]{holland2012}, the mapping speed is approximately
-$6\times10^{-5}$ and $3\times10^{-3}$\,deg$^2$\,hr$^{-1}$\,mJy$^{-2}$,
-at 450 and 850\micron, for the default pixel sizes of 2 and 4\,arcsec,
-respectively. If maps are filtered and cross-correlated with the PSF
-for the optimal extraction of point sources (see
-Section~\ref{sec:cosmo}), the mapping speeds increase to approximately
-$5\times10^{-4}$ and
-$1.5\times10^{-2}$\,deg$^2$\,hr$^{-1}$\,mJy$^{-2}$ in this example,
-independently of the chosen map pixel sizes. See the on-line \scuba\
-integration time calculator for mapping speeds under arbitrary
-observing conditions and scan
-modes\footnote{\url{http://www.jach.hawaii.edu/jac-bin/propscuba2itc.pl}}.
+A major source of divergence is correlation between \model{COM} and
+\model{FLT}. Since \model{FLT} usually consists of a high-pass filter
+following the application of \model{COM}, \model{COM} is completely
+free to grow any large-scale structure at frequencies below the chosen
+filter edge. While such structure does not appear in the map (as it is
+removed by \model{FLT}), we found that the solution could be made to
+converge significantly faster by ``re-mixing'' \model{COM} and
+\model{FLT} at the start of each iteration. In practice, at the start
+of each iteration, the previous iterations of each model are both
+added back into the residual immediately prior to re-calculating
+\model{COM}. In this way, truly common-mode signals, even at
+low-frequencies, do not leak into \model{FLT}.
+
+The most serious cause of divergence is the degeneracy between
+low-frequency signal that is removed, and large-scale structures in
+the map. For example, by subtracting the common-mode signal, the map
+is free to grow large-scale structures that exceed the angular scale
+of the array footprint. Our approach as been to constrain regions of
+the map devoid of sources to a value of zero for all but the final
+iteration in the \model{AST} model (Section~\ref{sec:ast}). Such
+regions are either user-defined in advance, or can be determined from
+the data using a cut on \snr. This technique is explored considerably
+in the examples from Sections~\ref{sec:point} and \ref{sec:extended}.
 
@@
 
-While most of the steps described in Section~\ref{sec:algorithm} for
-reducing data with SMURF are applicable to all data sets, there is no
-``default'' reduction that provides good results for a wide range of
-source \snr\ and angular scales. However, SMURF does have a small
-number of configurations with variations on these baseline parameters
-that are applicable to most common types of data. In this section we
-illustrate the differences between these configurations with the
-following examples: a bright point source, Uranus
-(Section~\ref{sec:point}); a blind survey of high-redshift galaxies in
-the Lockman Hole (Section~\ref{sec:cosmo}); and a map of a bright
-extended star-forming region in our Galaxy, M17
-(Section~\ref{sec:extended}).
+In this section three different common types of data are reduced: a
+bright point source, Uranus (Section~\ref{sec:point}); a blind survey
+of high-redshift galaxies in the Lockman Hole
+(Section~\ref{sec:cosmo}); and a map of a bright extended star-forming
+region in our Galaxy, M17 (Section~\ref{sec:extended}). In each case,
+variations on the default algorithm from Section~\ref{sec:algorithm}
+are described.
 
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=\linewidth]{pointmaps}
-\caption{Multiple reductions of a 450\,\micron\ CV daisy scan of
-  Uranus, all scaled between $-$0.002\,pW (white) and +0.002\,pW
-  (black). In all cases extinction correction has been applied
-  immediately after common-mode removal. (a) Reduction in which only
-  common-mode subtraction is used to suppress low-frequency noise, and
-  the reduction is forced to use 2 iterations (after the first
-  iteration an estimate of the source flux is removed, and the noise
-  is measured in the residuals to obtain appropriate weighting for the
-  second and final iteration). Circular streaks are caused by
-  independent low-frequency noise that is not removed by the
-  common-mode (Uranus peak 0.358\,pW). (b) Same as (a) but using 100
-  iterations, illustrating the degeneracy between large-scale
-  structure and the common-mode (the footprint of working bolometers
-  is also shown for reference, Uranus peak 0.358\,pW). (c) Reduction
-  in which high-pass filtering above 0.775\,Hz (corresponding to a
-  spatial scale of 200\,arcsec, as indicated) is applied after
-  common-mode removal, but before the map estimate. The map-based
-  convergence test is activated and the solution halts after 10
-  iterations, but leaving large-scale ringing due to the filter
-  (Uranus peak 0.353\,pW). (d) Same as (c), but now the region of the
-  map beyond the white dot-dashed circle is constrained to a value of
-  zero until all but the final iteration.  The map is now extremely
-  flat, there is no attenuation of the source flux compared with the
-  first two reductions, and the diffraction pattern is clearly seen
-  (Uranus peak 0.358\,pW).  }
-\label{fig:pointmaps}
-\end{figure*}
-
 The accurate measurement of positions and brightnesses of known point
@@
 (absolute calibration), and hence noise performance of the instrument
-in astronomically-useful units.  In this example we reduce a
-450\,\micron\ map of Uranus (observation 26 on 20111017), which is a
-nearly point-like source for \scuba\ that is commonly used as a
-primary flux calibrator. The CV daisy pattern was used, with a scan
-speed of 155 arcsec\,sec$^{-1}$. We perform several different
-reductions of the data to illustrate the purpose of various model
-components and the convergence properties of the solution
-(Fig.~\ref{fig:pointmaps}). In all cases the maps are produced on a
-grid of azimuth (horizontal) and elevation (vertical) offsets from the
-position of Uranus (the origin), using 2\,arcsec pixels.
+in astronomically-useful units.  In this example we produce a 450\,\micron\ map
+of Uranus (observation 26 on 20111017), which is a nearly point-like
+source for \scuba\ that is commonly used as a primary flux
+calibrator. The CV daisy pattern was used, with a scan speed of 155
+arcsec\,sec$^{-1}$. We perform several different reductions of the
+data to illustrate the purpose of various model components and the
+convergence properties of the solution (Fig.~\ref{fig:pointmaps}). In
+all cases the maps are produced on a grid of azimuth (horizontal) and
+elevation (vertical) offsets from the position of Uranus (the origin),
+using 2\,arcsec pixels.
 
@@
 and final iteration. The resulting map after these two iterations is
-shown in Fig.~\ref{fig:pointmaps}a. While the peak \snr\ of Uranus is
-clearly large ($\sim$160), enabling us to see the faint sidelobes
-(circle and cross pattern), the map also has obvious circular streaks
-and other large-scale ripples. The circular streaks are due to the
-fact that \model{COM} does not account for all of the low-frequency
-(i.e., uncorrelated) noise (see Fig.~\ref{fig:pspec}), and therefore
+shown in Fig.~\ref{fig:pointmaps}a. While the \snr\ of Uranus is
+clearly large ($\sim$250 as estimated by SMURF), enabling us to see
+the faint sidelobes (circle and cross pattern), the map also has
+obvious circular streaks and other large-scale ripples. The circular
+streaks are due to the fact that \model{COM} does not account for all
+of the low-frequency noise (see Fig.~\ref{fig:pspec}), and therefore
 each bolometer leaves a trace of the circular scan pattern in the map,
-as their baselines slowly drift independently. A significant
+as each of their baselines slowly drift independently. A significant
 contribution to the larger-scale ripples in the map, however, can be
@@
 bolometer after 2 (black) and 100 (grey) iterations, which are nearly
-identical, yet the change in the estimated
-\model{COM}\footnote{\model{COM} has been multiplied by the
-time-varying extinction correction to enable direct comparison with
-\model{AST}.} (green) and \model{AST} (red) signals between 2 and 100
-iterations are large However, it is also clear that the estimated
-\model{COM} and \model{AST} signals are complimentary. In other words,
-the large change in the \model{AST} signal is cancelled by freedom in
-the \model{COM} signal to grow with opposite sign. For comparison, the
+identical, yet the change in the estimated \model{COM} (green) and
+\model{AST} (red) signals between 2 and 100 iterations are
+large. However, it is also clear that the estimated \model{COM} and
+\model{AST} signals are complimentary. In other words, the large
+change in the \model{AST} signal is cancelled by freedom in the
+\model{COM} signal to grow with opposite sign. For comparison, the
 bottom panel of Fig.~\ref{fig:degeneracy} shows the telescope pointing
@@
 
-Attacking the problem of streaks, a simple method of removing residual
-(un-correlated) sources of low-frequency noise is to apply a high-pass
-filter after the common-mode removal. A third reduction of the data
-uses the baseline map-making parameters, as described in
+One improvement to remove residual low-frequency noise after
+common-mode removal in this Uranus image, is to simply apply a
+high-pass filter to the data. A third reduction of the data uses the
+default map-making parameters, as described in
 Section~\ref{sec:algorithm}, which adds the \model{FLT} model to
@@
 scan speed of 155\,arcsec\,sec$^{-1}$, corresponds to a frequency of
-0.78\,Hz (note in Fig.~\ref{fig:pspec} that this frequency is slightly
-above the $1/f$ knee at 450\,\micron\ after common-mode removal). Now
-using the automatic map-based convergence test
-(Section~\ref{sec:converge}), the solution converges after 10
+0.78\,Hz. Now using the automatic map-based convergence test
+(Section~\ref{sec:converge}), the resulting map converges after 10
 iterations (Fig.~\ref{fig:pointmaps}c). Both the circular streaks and
@@
 centre, the result is an azimuthally-symmetric $\sinc$ function-like
-pattern in the map. The peak flux has also been attenuated to
-0.353\,pW in this reduction, compared with 0.358\,pW in the first two
-reductions. The gradient resulting from the degeneracy between the
-common-mode and map has been removed because the filter scale of
-200\,arcsec (3.3\,arcmin) is smaller than the array field-of-view
-($\sim$5\,arcmin). If it were greater, degeneracies would again begin
-to appear on those larger scales.
-
-This example illustrates the need for constraints in the map solution
+pattern in the map.
+
+This example illustrate the need for constraints in the map solution
 in many situations. For calibrators (and other previously known
 bright, compact sources), a good, simple prior is to constrain the map
-to a value of zero away from the known locations of emission (part of
-the \model{AST} model calculation, Section~\ref{sec:ast}).  In
+to a value of zero away from the known locations of emission (inside
+the \model{AST} model, Section~\ref{sec:ast}). In
 Fig.~\ref{fig:pointmaps}d, a solution is produced in an identical
@@
 final iteration. In this case, the map converges after 6 iterations,
-and the ringing has been effectively removed. The attenuation from the
-previous reduction is now removed, and Uranus again has a peak value
-of 0.358\,pW. For reference, the Uranus model distributed with
-Starlink predicts a peak brightness of 176\,Jy on this date, yielding
-an FCF of 491 Jy\,pW$^{-1}$, which is well within the typical range
-for \scuba\ of $(492 \pm 67)$\,Jy\,pW$^{-1}$ \citep[see][for
-details]{dempsey2012}. Since the map is now flat away from the source,
-and constrained to a value of zero, it is appropriate for performing
-aperture photometry directly, with no need for an additional reference
-annulus.
-
-The way this prior works can be understood from the point of view of
-differential measurements. Bolometer data contain information up to
-scales corresponding to the filter edge (or the scale of the array
-footprint, whichever is smallest). In this example, the relevant scale
-is 200\,arcsec. Since the map is constrained to zero within 60\,arcsec
-of the peak of Uranus (well within 200\,arcsec), the solution is able
-to accurately reconstruct the differential peak intensity of Uranus
-with respect to this constrained background. This approach to
-map-making is similar to that employed for poorly cross-linked scans
-of compact (though resolved) sources by \citet{wiebe2009} using BLAST
-data.
+and all of the previous artifacts have been effectively removed. Since
+the map is now flat away from the source, and constrained to a value
+of zero, it is appropriate for performing aperture photometry
+directly, with no need for an additional reference annulus. This
+approach to constrained map-making is similar to that employed for
+poorly cross-linked scans of compact (though resolved) sources by
+\citet{wiebe2009} using BLAST data.
+
+\begin{figure*}
+\centering
+\includegraphics[width=\linewidth]{pointmaps}
+\caption{Multiple reductions of a 450\,\micron\ CV daisy scan of
+  Uranus, all scaled between $-$0.002\,pW (white) and +0.002\,pW
+  (black). (a) Reduction in which only common-mode subtraction is used
+  to suppress low-frequency noise, and the reduction is forced to use
+  2 iterations (after the first iteration an estimate of the source
+  flux is removed, and the noise is measured in the residuals to
+  obtain appropriate weighting for the second and final
+  iteration). Circular streaks are caused by independent low-frequency
+  noise that is not removed by the common-mode (Uranus peak
+  0.15\,pW). (b) Same as (a) but using 50 iterations, illustrating the
+  degeneracy between large-scale structure and the common-mode (the
+  footprint of working bolometers is also shown for reference, Uranus
+  peak 0.15\,pW). (c) Reduction in which high-pass filtering above
+  0.775\,Hz (corresponding to a spatial scale of 200\,arcsec, as
+  indicated) is applied after common-mode removal, but before the map
+  estimate. The map-based convergence test is activated and the
+  solution halts after 10 iterations, but leaving large-scale ringing
+  due to the filter (Uranus peak 0.27\,pW). (d) Same as (c), but now
+  the region of the map beyond the white dot-dashed circle is
+  constrained to a value of zero until all but the final iteration.
+  The map is now extremely flat, there is little attenuation of the
+  source flux, and the diffraction pattern is clearly seen (Uranus
+  peak 0.27\,pW).}
+\label{fig:pointmaps}
+\end{figure*}
 
@@
 response, without prior knowledge of the location of sources; (ii)
-apply a Fourier-space ``whitening filter'' to suppress residual
-large-scale noise; and (iii) detecting point sources using a ``matched
-filter''. Note that variations on this general procedure have been
-used extensively in the submm cosmology community using previous
-instruments
+whitening (essentially flattening) the map to make the noise
+properties of the map easier to deal with; and (iii) detecting point
+sources using a ``matched filter''. Note that variations on this
+general procedure have been used extensively in the submm cosmology
+community using previous instruments
 \citep[e.g.,][]{scott2002,borys2003,laurent2005,coppin2006,scott2008,perera2008,devlin2009}.
@@
 $\sim8.5$\,hours of data taken in 36 separate scans (average
-15\,min. each) spread over February and March 2010. Each scan is a
-360\,arcsec rotating PONG pattern, with a scan speed of
-240\,arcsec\,s$^{-1}$, covering an area of about 50\,arcmin$^2$. The
-full list of dates and observation numbers includes: 20100218, 63, 64,
-65, 70, 71, 72, 90, 91, 92, 97, 98, 99; 20100220, 111, 112, 113, 118,
-128, 129, 130; 20100303, 61, 62, 64, 69, 70, 72, 73, 74, 75; 20100309,
-87, 88, 90, 91; and 20100311, 59, 64, 65, 72. These observations
-represent about 80 per cent of the total data taken as part of the
-project; the remaining observations were dropped due to problems
-keeping the arrays properly tuned during S2SRO, and were easily
-identified by their highly variable and erratic bolometer time-series
-(which resulted in maps full of large streaks).
+15\,min. each) spread over February and March 2010, covering an area
+of about 50\,arcmin$^2$. The full list of dates and observation
+numbers includes: 20100218, 63, 64, 65, 70, 71, 72, 90, 91, 92, 97,
+98, 99; 20100220, 111, 112, 113, 118, 128, 129, 130; 20100303, 61, 62,
+64, 69, 70, 72, 73, 74, 75; 20100309, 87, 88, 90, 91; and 20100311,
+59, 64, 65, 72. These observations represent about 80 per cent of the
+total data taken as part of the project; the remaining observations
+were dropped due to problems keeping the arrays properly tuned during
+S2SRO, and were easily identified by their highly variable and erratic
+bolometer time-series (which resulted in maps full of large streaks).
 
@@
   followed by 4 iterations using only the \model{COM} model to remove
-  residual correlated noise. White contours correspond to estimated
+  residual low-frequency noise. White contours correspond to estimated
   noise levels of 1.25, 2.5 and 5.0 times the minimum noise at the
@@
   from \citet{owen2008} with $S\gsim15$\,$\mu$Jy. Of the 10 submm
-  peaks, 9 are within a search radius 8\,arcsec of at least one radio
+  peaks, 9 are within this search radius of at least one radio
   source. (d) Jackknife map produced from the difference of two maps,
@@
 employed. Large-scale diverging structures in the map must be
-mitigated, and the method used in this example (the baseline
-processing in SMURF) is to apply a high-pass filter to the data once,
-as a pre-processing step. The iterative solution is then run using
-only \model{COM}, \model{EXT}, \model{AST}, and \model{NOI}. In other
+mitigated, and the method used in this example (the default processing
+in SMURF) is to apply a high-pass filter to the data once, as a
+pre-processing step. The iterative solution is then run using only
+\model{COM}, \model{EXT}, \model{AST}, and \model{NOI}. In other
 words, there is no information in the bolometer signals below some
@@
 the cutoff is only removed through iterative common-mode
-subtraction. In this case, the filter edge has been chosen to remove
-scales larger than 200\,arcsec, or a high-pass filter above 1.2\,Hz
-given the scan speed. Since the data are high-pass filtered prior to
-the iterative solution, \model{GAI} (fitting an independent amplitude
-of \model{COM} to each bolometer) has been de-activated, since there
-is very little structure in the common-mode with which to fit an
-accurate gain. The map is shown in Fig.~\ref{fig:lockman_maps}a. The
-map-maker has been tested in two ways: (i) large numbers of iterations
-are used to verify that the maps converge without the growth of large
-structure; and (ii) adding synthetic sources to the real time-series
-data (a built-in feature of SMURF) at a range of brightnesses verify
-that the map-maker response to them is linear (i.e., the relative
-shape and amplitude compared to the input source is independent of
-brightness). The response to a synthetic point source (solid line)
-after map-making (dotted line) is shown in
-Fig.~\ref{fig:lockman_psf}. Clearly the use of a high-pass filter as a
-pre-processing step, and having no other map-constraints, has the
-down-side of introducing sidelobes around the main peak. Furthermore,
-the details of this shape depends on the high-pass filter edge that
-has been chosen (the higher the frequency, the greater the attenuation
-of the central peak, and the larger the negative side-lobes).
-However, the way this filter affects point-sources is measurable
-(using the synthetic source injection facility of SMURF), and linear.
+subtraction. Since the data are high-pass filtered prior to the
+iterative solution, \model{GAI} (fitting an independent amplitude of
+\model{COM} to each bolometer) has been de-activated, since there is
+very little structure in the common-mode with which to fit an accurate
+gain. The map is shown in Fig.~\ref{fig:lockman_maps}a. The map-maker
+has been tested in two ways: (i) large numbers of iterations are used
+to verify that the maps converge without the growth of large
+structure; and (ii) adding synthetic sources to the time-series data
+at a range of brightnesses verify that the map-maker response to them
+is linear (i.e., the relative shape and amplitude compared to the
+input source is independent of brightness). The response to a
+synthetic point source (solid line) after map-making (dotted line) is
+shown in Fig.~\ref{fig:lockman_psf}. Clearly the use of a high-pass
+filter as a pre-processing step, and having no other map-constraints,
+has the down-side of introducing sidelobes around the main
+peak. However, the way this filter affects point-sources is both
+well-known, and linear.
 
@@
   source (solid line) along the R.A. axis following map-making using
-  the blank-field processing configuration with SMURF (dotted line),
-  and upon application of the whitening filter (dashed line). The
-  ``map filtered'' response is produced by adding the ideal Gaussian
-  to the real data near the centre of the Lockman Hole map, and
-  re-reducing the data. The resulting dotted line gives the expected
-  shape of a source in panel (a) from
-  Fig.~\ref{fig:lockman_maps}. Applying the whitening filter
-  (reciprocal of the solid orange line in
+  the default blank-field processing with SMURF (dotted line), and
+  upon application of the whitening filter (dashed line). The ``map
+  filtered'' response is produced by adding the ideal Gaussian to the
+  real data near the centre of the Lockman Hole map, and re-reducing
+  the data. The resulting dotted line gives the expected shape of a
+  source in panel (a) from Fig.~\ref{fig:lockman_maps}. Applying the
+  whitening filter (reciprocal of the solid orange line in
   Fig.~\ref{fig:lockman_pspec}) to the dotted line gives the whitened
-  line (dashed), which is the expected shape of a source in panel (b)
+  line (dashed) , which is the expected shape of a source in panel (b)
   from Fig.~\ref{fig:lockman_maps}. Finally, cross-correlating the
@@
 map (Fig.~\ref{fig:lockman_maps}b). To identify sources, we next apply
-a ``matched filter'' to the whitened maps.
+a matched filter to the whitened maps.
 
@@
 be un-resolved by the \scuba\ 7.5--14.5\,arcsec FWHM beams. Under this
-assumption, and also assuming that the map noise is white,
-cross-correlation between the map and the known PSF, or matched
-filtering, yields the maximum-likelihood flux density of supposed
-point-sources centred over every location in the resulting map
-\citep[an extremely well-known result throughout astronomy,
-see][]{stetson1987}. Peak identification in such smoothed maps have
-been used extensively in the submillimetre community, as both an
-efficient source-detection and photometric measurement strategy. In
-terms of the angular power spectrum, a matched filter may also be
-thought of as an optimal low-pass filter, that suppresses noise on
-scales smaller (and therefore at higher angular frequencies) than the
-beam. For the case at hand, we may use this cross-correlation
-technique since the map has been whitened. However, we must first
-establish the effective shape of point sources in this map due both to
-map-making itself, and the whitening filter. We determine the effect
-of map-making by adding a synthetic (and high-\snr) point source to
-the real data, and measure its resulting shape in the map (solid and
-dotted lines in Fig.~\ref{fig:lockman_psf}, respectively). Next, the
-Fourier transform of the map filtered PSF is divided by the jackknife
-noise power spectrum to calculate the final whitened PSF (dashed line
-in Fig.~\ref{fig:lockman_psf}). Maps smoothed by this shape are shown
-for the whitened signal, and jackknife maps in
+assumption, cross-correlation between the map and the known PSF yields
+the maximum-likelihood flux density of supposed point-sources centred
+over every location in the resulting map \citep[an extremely
+well-known result throughout astronomy, see][]{stetson1987}. Peak
+identification in such smoothed maps have been used extensively in the
+submillimetre community, as both an efficient source-detection and
+photometric measurement strategy. In terms of the angular power
+spectrum, a matched filter may also be thought of as an optimal
+low-pass filter, that suppresses noise on scales smaller (and
+therefore at higher angular frequencies) than the beam. For the case
+at hand, the effective PSF for the whitened maps is given by the
+dotted curve in Fig.~\ref{fig:lockman_psf}. Maps smoothed by this
+shape are shown for the whitened signal, and jackknife maps in
 Figs.~\ref{fig:lockman_maps}c and f, respectively. Note that these
@@
 proven invaluable for the precise identifications of high-redshift
-submillimetre galaxies due to their low surface densities (compared
-with optical catalogues, for example), and a strong correlation
-between the radio and submillimetre emission mechanisms
+submillimetre galaxies due to their relatively low surface densities
+(compared with optical catalogues, for example), and a strong
+correlation between the radio and submillimetre emission mechanisms
 \citep[e.g.,][]{smail2000,pope2006,ivison2007,chapin2009b}. Taking a
@@
 \caption{An 850\,\micron\ rotating PONG map of M17. Intensity is
-logarithmically scaled between $-$0.0003\,pW (white) and +0.01\,pW
+logarithmically scaled between $-$0.0003\,pw (white) and +0.01\,pW
 (black). Iteration numbers are given in the corner of each
 panel. Panels (a) and (b) show the results for a reduction using the
-baseline parameters (the solution halted after reaching the map-based
-convergence criterion in 17 iterations). Panel (a) also depicts the
-array footprint (position angle indicative of the start of the
-observation), and a 300\,arcsec line shows the spatial scale
-corresponding to the \model{FLT} high-pass filter. Similar to
-Fig.~\ref{fig:pointmaps}c, the high-pass filtering introduces ringing
-around bright sources. Panels (c) and (d) show the ``bright extended''
-reduction, in which a zero-mask is created iteratively from all of the
-pixels that lie below a \snr\ of 5. While this region (outside the red
-contour) only avoids the brightest peaks early in the solution, in the
-final iteration it skirts most of the bright, extended emission, and
-significantly helps with negative ringing.}
+default parameters (the solution halted after reaching the default
+map-based convergence criterion in 17 iterations). Panel (a) also
+depicts the array footprint (position angle indicative of the start of
+the observation), and a 300\,arcsec line shows the spatial scale
+corresponding to the default \model{FLT} high-pass filter. Similar to
+Fig.~\ref{fig:pointmaps}c, the lack of prior constraints on the map
+leads to degeneracies, and therefore ripples on scales larger than the
+array footprint (due to common-mode subtraction) and filter scale
+(especially evident around the brightest peak at the centre of the
+map). Panels (c) and (d) show the ``bright extended'' reduction, in
+which a zero-mask is created iteratively from all of the pixels that
+exceed an \snr\ of 5. While this region (red contour) only encompasses
+the brightest peaks early in the solution, in the final iteration it
+identifies most of the bright, extended emission, and significantly
+helps with negative ringing.}
 \label{fig:m17}
@@
 
-The baseline reduction of these data is shown in the top panels of
+The default reduction of these data is shown in the top panels of
 Fig.~\ref{fig:m17}, after 2 iterations (the first map estimated after
 the noise weights have been measured) and 17 iterations (when the map
-has converged), which uses iterative common-mode subtraction and
-high-pass filtering. The first panel also depicts the array footprint,
-and the angular scale (300\,arcsec) corresponding to the high-pass
-filter edge (0.6\,Hz). Much like the reduction of a point source
-without any prior constraints (Fig.~\ref{fig:pointmaps}c), there are
-ripples around bright sources due to the filtering.
+has converged). The first panel also depicts the array footprint, and
+the angular scale corresponding to the high-pass filter edge. Much
+like the reduction of a point source without any prior constraints
+(Fig.~\ref{fig:pointmaps}c), degeneracies on scales larger than the
+array footprint and filter produce ripples in the map.
 
@@
 has been added to SMURF to generate one automatically by flagging
-pixels below some \snr\ threshold to be set to zero after each
-iteration as part of the ``bright extended'' configuration.
+pixels above some \snr\ threshold after each iteration as part of the
+``bright extended'' configuration.
 
 The results of this automatic masking are shown in the bottom panels
-of Fig.~\ref{fig:m17}. After the second iteration, everything but the
-brightest peaks are set to zero (outside the red contours). However,
-as the solution progresses, the negative bowls around the bright
-sources are slowly reduced and the mask ``grows'' out from the
-brightest areas. By the final iteration, most of the obvious
-structures in the data are excluded in the the mask, negative bowling
-is significantly reduced, and the brightest regions are more extended.
+of Fig.~\ref{fig:m17}. After the second iteration, only narrow regions
+around the brightest peaks are identified. However, as the solution
+progresses, the negative bowls around the bright sources are slowly
+reduced and the mask ``grows'' out from the brightest areas. By the
+final iteration, most of the obvious structures in the data are
+encompassed by the mask, negative bowling is significantly reduced,
+and the brightest regions are more extended.
 
@@
 average of two halves of the data analysed independently, using the
-baseline configuration. The data have had added to them synthetic
+default configuration. The data have had added to them synthetic
 signal within a 600\,arcsec diameter circle (south of M17) created as
@@
 first row. Third row: average of the two halves of the data using the
-bright extended reduction. The regions outside the blue and red
-contours are constrained to zero for each half of the data (note for
-the 300\,arcsec case that these regions about M17 closely match the
-mask in Fig.~\ref{fig:m17}d using a full reduction). Clearly as the
-filter scale is increased, due to the high \snr\ of the data, larger
-emission regions are (correctly) detected and reproduced in the
-map. Fourth (bottom) row: Jackknife maps for the bright extended
-reductions.}
+bright extended reduction. The blue and red contours indicate the
+zero-masked regions for each half of the data (note for the
+300\,arcsec case that the region about M17 fairly closely matches the
+mask in Fig.~\ref{fig:m17}d). Clearly as the filter scale is
+increased, due to the high \snr\ of the data, larger emission regions
+are (correctly) detected and reproduced in the map. Fourth (bottom)
+row: Jackknife maps for the bright extended reductions.}
 \label{fig:m17_jk}
@@
 \centering
-\begin{minipage}[h]{0.495\linewidth}
-\textbf{(a) baseline reduction PSD} \\
-
-\includegraphics[width=\linewidth]{pspec_m17_default}
-\end{minipage}
-\begin{minipage}[h]{0.495\linewidth}
-\textbf{(b) baseline reduction transfer function} \\
-
-\includegraphics[width=\linewidth]{cor_pspec_m17_default}
-\end{minipage}
+\includegraphics[width=0.49\linewidth]{pspec_m17_default}
+\includegraphics[width=0.49\linewidth]{cor_pspec_m17_default}
 \caption{Angular PSDs for the region of the M17 map in
-Fig.~\ref{fig:m17_jk} containing synthetic data, using the baseline
-configuration. (a) Raw PSDs for the input (noiseless) simulated data
+Fig.~\ref{fig:m17_jk} containing synthetic data, using the default
+configuration. Left: raw PSDs for the input (noiseless) simulated data
 (thick black line), the signal (average of each half) PSDs (thin solid
@@
 150\,arcsec (red); 300\,arcsec (orange); 600\,arcsec (green); and
-900\,arcsec (blue). PSDs are also colour-coded by filter scale. (b)
-Since the input PSD is known, it is possible to measure the transfer
+900\,arcsec (blue). PSDs are also colour-coded by filter scale. Right:
+since the input PSD is known, it is possible to measure the transfer
 function of the map-maker as the ratio between the difference of the
@@
 the thin coloured lines (linear vertical axis shown on right of
-plot). The remaining lines are as in (a), but now corrected by the
-transfer function. This plot shows that increasing the filter scale
-improves the \snr\ at intermediate scales ($\sim$200--600\,arcsec),
-although the \snr\ worsens at smaller scales ($\lsim$200\,arcsec).}
+plot). The remaining lines are as in the left panel, but now corrected
+by the transfer function. This plot shows that increasing the filter
+scale improves the \snr\ at intermediates scales
+($\sim$200--600\,arcsec), although the \snr\ worsens at smaller scales
+($\lsim$200\,arcsec).}
 \label{fig:m17_def_ps}
@@
 \centering
-\begin{minipage}[h]{0.495\linewidth}
-\textbf{(a) bright extended reduction PSD} \\
-
-\includegraphics[width=\linewidth]{pspec_m17_bright_extended}
-\end{minipage}
-\begin{minipage}[h]{0.495\linewidth}
-\textbf{(b) bright extended reduction transfer function} \\
-
-\includegraphics[width=\linewidth]{cor_pspec_m17_bright_extended}
-\end{minipage}
+\includegraphics[width=0.49\linewidth]{pspec_m17_bright_extended}
+\includegraphics[width=0.49\linewidth]{cor_pspec_m17_bright_extended}
 \caption{Lines have same meaning as in Fig.~\ref{fig:m17_def_ps},
@@
 angular scales. Also note that increasing the filter scale does not
-have as large an impact on the small-scale noise as in the baseline
+have as large an impact on the small-scale noise as in the default
 configuration, although it is still noticeably larger up to scales of
@@
 
-While the reduction in the bottom panels of Fig.~\ref{fig:m17} is (in
-a cosmetic sense) superior to those in the top panels, it is important
-to quantify both the noise properties of the maps, and the response to
-real structures (the transfer function). We would also like to know
-how each are affected by our choice of filter scale.  Similar to the
-previous section, we will use a jackknife test to estimate the noise,
-as well as injecting known sources into the data to observe how they
-are attenuated.
+While the reduction in the bottom panels of Fig.~\ref{fig:m17} is (at
+least a cosmetic sense) superior to those in the top panels, it is
+important to quantify both the noise properties of the maps, and the
+response to real structures (the transfer function). We would also
+like to know how each are affected by our choice of filter scale.
+Similar to the previous section, we will use a jackknife test to
+estimate the noise, as well as injecting known sources into the data
+to observe how they are attenuated.
 
@@
 averaging the maps made of each independent half of the data, for the
-baseline configuration (inverse-variance weighting has been used). The
+default configuration (inverse-variance weighting has been used). The
 columns show reductions using 150, 300, 600, and 900\,arcsec filter
-edges. The synthetic data are clearly seen as the circular region
-south of M17. As the filter scale is increased, the size of the
-ripples increases accordingly. While larger astronomical structures do
-seem to appear, negative bowls are a major problem without any other
-map constraints. Since the largest scale that is completely inscribed
+edges. For reference, the largest scale that is completely inscribed
 by the array footprint is about 400\,arcsec, and the diagonal of the
-array is about 600\,arcsec, scales ranging from $\sim$400--600\,arcsec
-up to the scale of the filter will be unconstrained due to the
-degeneracy between the common-mode and map (see discussion in
-Section~\ref{sec:point}). This degeneracy therefore causes some of the
-large-scale patchiness in the 600 and 900\,arcsec filtered maps that
-lie away from the central bright sources.  The second row of
-Fig.~\ref{fig:m17_jk} shows the jackknife maps. The astronomical
-emission is almost perfectly removed, except for a slight impression
-of M17 near the centre of the map, that is probably due to some
-combination of detector gain and pointing drifts. Otherwise the
-jackknife appears to be a plausible realisation of noise from the same
-parent distribution as for the averaged maps.
+array is about 600\,arcsec (meaning that scales beyond 600\,arcsec are
+completely unconstrained when common-mode rejection is used).  The
+synthetic data are clearly seen as the circular region south of
+M17. As the filter scale is increased, the size of the ripples
+increases accordingly. While larger structures do seem to appear,
+negative bowls are a major problem without any other map
+constraints.The second row of Fig.~\ref{fig:m17_jk} shows the
+jackknife maps. The astronomical emission is almost perfectly removed,
+except for a slight impression of M17 near the centre of the map, that
+is probably due to some combination of detector gain and pointing
+drifts. Otherwise the jackknife appears to be a plausible realisation
+of noise from the same parent distribution as for the averaged maps.
 
@@
 exercise using the bright extended configuration, in which the \snr\
-threshold of 5 is again used to identify low-significance pixels that
-are set to zero after each iteration. As the filter scale is
-increased, more of the extended structure in M17 is reproduced in the
-map, as evidenced by the blue and red contours (masks generated from
-the first and second halves of the data, respectively). The masking
-does a generally good job of suppressing the largest-scale ripples
-that are produced by the baseline reduction. However, the noise away
-from regions of bright emission does increase noticeably (mottled
-appearance) --- due to residual $1/f$ noise following common-mode
-removal that has a knee at a frequency above the filter edge. With
-filter scales $\gsim600$\,arcsec, virtually the entire region of
-synthetic emission is correctly identified by the \snr\ mask and
-allowed to vary freely in the solution.
+threshold of 5 is again used to constrain the map. As the filter scale
+is increased, more of the extended structure in M17 is reproduced in
+the map, as evidenced by the blue and red contours (masks generated
+from the first and second halves of the data, respectively). The
+masking does a generally good job of suppressing the largest-scale
+ripples that are produced by the default reduction. However, the noise
+away from regions of bright emission does increase noticeably (mottled
+appearance). With filter scales $\gsim600$\,arcsec, virtually the
+entire region of synthetic emission is correctly identified by the \snr\
+mask and allowed to vary freely in the solution.
 
@@
 maps to understand the signal and noise properties of the map-maker in
-the region of synthetic sources, as a function of filter scale. In
-Fig.~\ref{fig:m17_def_ps}a we show the raw PSDs for the input
-synthetic signal (thick black line), the output map signals (thin
-solid lines), and the jackknife maps (dashed lines). Colours encode
-the filter scales used: 150\,arcsec (red); 300\,arcsec (orange);
-600\,arcsec (green); and 900\,arcsec (blue). Note that, with the
-exception of the synthetic data, we only plot the PSDs down to the
+the region of synthetic sources, as a function of filter scale. The
+left panel of Fig.~\ref{fig:m17_def_ps} shows the raw PSDs for the
+input synthetic signal (thick black line), the output map signals
+(thin solid lines), and the jackknife maps (dashed lines). Colours
+encode the filter scales used: 150\,arcsec (red); 300\,arcsec
+(orange); 600\,arcsec (green); and 900\,arcsec (blue). Note that, with
+the exception of the synthetic data, we only plot the PSDs down to the
 second-lowest spatial frequency bin of
-$1.875\times10^{-3}$\,arcsec$^{-1}$, or 533\,arcsec, since the lowest
-is very poorly sampled, and therefore noisy. At small scales
-($\lsim$20\,arcsec) the map and jackknife PSDs flatten. However,
-unlike Fig.~\ref{fig:lockman_pspec}, these white noise levels are
-slightly larger in the jackknife maps, suggesting that the differences
-have not completely removed the astronomical signals. A strong
-possibility is that the two half-maps that go into the differences
-have not converged equally, probably due to the lack of prior
-constraints.  As the filter edge is increased, more power is measured
-in the map PSDs at larger scales. However, much of this power is
-clearly produced by noise which appears in the jackknife PSDs. We
-estimate the map-maker transfer function as the ratio between the
-portion of the signal PSDs not produced by noise, to the input PSD, or
+$1.875\times10^{-3}$\,arcsec$^{-1}$, or 533\,arcsec, due to the fact
+that the smallest is very poorly sampled, and therefore noisy. As the
+filter edge is increased, more power is measured in the map PSDs at
+larger scales. However, much of this power is clearly produced by
+noise which appears in the jackknife PSDs. We estimate the map-maker
+transfer function as the ratio between the portion of the signal PSDs
+not produced by noise, to the input PSD, or
 %
@@
 jackknife PSDs by dividing by $G(k)$ to produce the thick solid, and
-dashed lines in Fig.~\ref{fig:m17_def_ps}b. This shows us that, even
-though the raw noise in Fig.~\ref{fig:m17_def_ps}a is lower at small
-scales when a smaller-scale filter is used (e.g., the red dashed
+dashed lines in the right panel of Fig.~\ref{fig:m17_def_ps}. This
+shows us that, even though the raw noise in the left panel is lower at
+small scales when a smaller-scale filter is used (e.g., the red dashed
 line), once we correct for the transfer function, we actually win in a
@@
 Fig.~\ref{fig:m17_be_ps}. The most obvious improvement with this
-reduction over the baseline reduction is that the transfer functions
+reduction over the default reduction is that the transfer functions
 fall more slowly at large angular scales, accompanied by a slower
 increase in noise; in other words, there is greatly improved \snr\ at
-large angular scales (an obvious conclusion given the appearance of
+large angular scales (an obvious conclusion given the appearence of
 the maps in Fig.~\ref{fig:m17_jk}). In fact, using the 900\,arcsec
@@
 in small-scale noise as the filter edge is increased is not as drastic
-as in the baseline reduction, and the \snr\ improves at scales
-$\gsim$60\,arcsec. Finally, note that both the map and jackknife white
-noise levels (at scales $\lsim$20\,arcsec) are in excellent agreement,
-unlike the previous example.
+as in the default reduction, and the \snr\ improves at scales
+$\gsim$60\,arcsec.
 
 One case in which the \snr\ is worse using the bright extended
-reduction is when using a 150\,arcsec filter. Here the noise is
-considerably larger in the bright extended reduction, as evidenced by
-the ``kink'' near 150\,arcsec. Referring to the mask contours in the
-left panel of the third row in Fig,~\ref{fig:m17_jk}, it is clear
-that the map-maker has failed to identify much of the bright, extended
-emission in the region of the synthetic source. Each area that is not
-within the contours is constrained to zero throughout the solution,
-therefore suppressing power (and lowering the transfer function), and
-subsequently reducing the \snr\ of the final result. This measurement
-serves as a warning: the map-maker response is non-linear when using
-\snr\ masking. Harsh filtering can provide misleading results, as in
-this example. Maps of faint extended emission will also suffer
-considerably, as the structures of interest may not clear the \snr\
-threshold for the mask.
+reduction is when using a 150\,arcsec filter. In this case the noise
+is considerably larger in the bright extended reduction, as evidenced
+by the ``kink'' near 150\,arcsec. Referring to the mask contours in
+the left panel of the third row, it is clear that the map-maker has
+failed to identify much of the bright, extended emission in the region
+of the synthetic source. Each area that is not within the contours is
+constrained to zero throughout the solution, therefore suppressing
+power (and lowering the transfer function), and subsequently reducing
+the \snr\ of the final result. This measurement serves as a warning:
+the map-maker response is non-linear when using \snr\ masking. Harsh
+filtering can provide misleading results, as in this example. Maps of
+faint extended emission will also suffer considerably, as the
+structures of interest may not clear the \snr\ threshold for the mask.
 
@@
 restricts the solution to a small fixed number of iterations
-($\lsim$10), so that there is probably little opportunity for
-degeneracies between the map and common-mode to appear in the map.
-Experimentation with the model order is also advocated to gain an
-impression of the convergence properties. This approach is perhaps
-more relevant to their SHARC-2 data for which a more complicated model
-is developed; the degeneracies we have observed for \scuba\ maps using
-our more brute-force approach with a high-pass filter tend to look
-similar regardless of the order (only the convergence time is
-affected).  For Bolocam data of the Galactic Plane,
-\citet{aguirre2011} used a maximum-entropy filtering step to suppress
-large-scales. Regardless of the method used, simulations are always
-required to establish the transfer function and noise properties as a
-function of angular scale.
+($\lsim$10), so that there is probably little chance that such
+structures have the chance to grow. Experimentation with the model
+order is also advocated to gain an impression of the convergence
+properties. This approach is perhaps more relevant to their SHARC-2
+data for which a more complicated model is developed; the degeneracies
+we have observed for \scuba\ maps using our more brute-force approach
+with a high-pass filter tend to look similar regardless of the order
+(only the convergence time is affected).  For Bolocam data of the
+Galactic Plane, \citet{aguirre2011} used a maximum-entropy filtering
+step to suppress large-scales. Regardless of the method used,
+simulations are always required to establish the transfer function and
+noise properties as a function of angular scale.
 
@@
 correlated noise components from the data, successively improving the
-map \citep[e.g.,][]{kovacs2008,aguirre2011,schuller2012}, we have
-shown that a fairly simple approach provides good results for \scuba\
+map \citep[e.g.,][]{kovacs2008,siringo2009,aguirre2011}, we have shown
+that a relatively simpler approach provides good results for \scuba\
 data, with reasonable computational requirements. This conclusion is
@@
 angular scales due to the degeneracy between the map, and the
-low-frequency signal components that are removed (namely the
-common-mode). In addition, the high-pass filtering produces
-significant ringing around bright sources. A simple strategy of
+low-frequency signal components that are removed. A simple strategy of
 constraining empty regions of the map to zero (using either a
 user-supplied mask for known sources, or an iterative determination of
-signal below some \snr\ threshold) provides good constraints for both
+signal above some \snr\ threshold) provides good constraints for both
 compact objects, and bright/extended structures. Particularly in the
@@
 
-For maps of faint point-sources, a single (non-iterative) high-pass
+For maps of faint point sources, a single (non-iterative) high-pass
 filter at the start of the reduction produces maps that are nearly
-white-noise limited and linear (i.e., the response does not depend on
-\snr). Residual large-scale noise can be removed with a whitening
-filter (also established from jackknife estimates of the noise) based
-on the Fourier Transform of the maps, and sources detected using a
-matched-filter (smoothing with the effective filtered point spread
-function).
+white-noise limited. Residual large-scale noise can be removed with a
+whitening filter (also established from jackknife estimates of the
+noise), and sources detected using a matched-filter (smoothing with
+the effective filtered point spread function).
 
@@
 for a wide variety of observations. Furthermore, the execution times
-are typically shorter than the observation lengths, and memory
-requirements for even the longest \scuba\ observations are within the
-capabilities of single, high-end desktop computers. SMURF can
-therefore provide real-time feedback at the telescope to observers.
+are shorter than the observation lengths, and memory requirements for
+even the longest \scuba\ observations are within the capabilities of
+single, high-end desktop computers. SMURF can therefore provide real-time
+feedback at the telescope to observers, and it also produces nearly
+science-grade data products for the JCMT Science Archive hosted by the
+Canadian Astronomy Data Centre.
+
+%SMURF is highly configurable, making it easy to vary the order
+%and types of models that are fit to the data. It is also easy to add
+%new models to the existing infrastructure.
 
@@
 Research, and the National Research Council of Canada. Additional
-funds for the construction of \scuba\ were provided by the Canada
+funds for the contruction of \scuba\ were provided by the Canada
 Foundation for Innovation.  This research was supported in part by the
@@
 members of the SCUBA-2 commissioning team for testing the map-maker
-and reporting anomalies; in particular Antonio Chrysostomou and
-Jessica Dempsey.  We also thank Mark Halpern, Matthew Hasselfield, and
-Gaelen Marsden for many useful discussions; observers who provided
-helpful feedback, especially David Nutter and Todd McKenzie; and
-Mandana Amiri and Dan Bintley for consultations regarding the SCUBA-2
-Multi-Channel Electronics.  We acknowledge the contributions of Dennis
-Kelly, Alex van Engelen and Jennifer Balfour for early investigations
-related to SMURF; and Mark Thompson, Craig Walther and S\'{e}verin
-Gaudet for being on the Critical Design Review panel. We thank Per
-Friberg and Gary Davis for their helpful comments on the
-manuscript. Finally, we thank the anonymous referee for a very helpful
-review that helped to clarify several key areas of the paper.
+and reporting anomalies; in particular Antonio Chrysostomou and Jessica
+Dempsey.  We also thank Mark Halpern, Matthew
+Hasselfield, and Gaelen Marsden for many useful discussions, and
+observers who provided helpful feedback; especially David Nutter and
+Todd McKenzie.  We acknowledge the contributions of Dennis Kelly, Alex
+van Engelen and Jennifer Balfour for early investigations related to
+SMURF; and Mark Thompson, Craig Walther and S\'{e}verin Gaudet for
+being on the Critical Design Review panel. Finally, we thank Per
+Friberg and Gary Davis for their helpful comments on the manuscript.
 

### 2012-12-11 resubmit_mnras/response.txt
ready for resubmission

@@
-Draft of my answers preceded by "-":
+To whom it may concern,
+
+We would like to thank the thorough reviewer for identifying several
+areas that needed clarification, and perhaps more importantly, three
+careless errors (which led to incorrect units in Figs.1 & 2, and
+confusing results in Figs.11 & 20), and an omission (impact /
+treatment of the response of the bolometers and data acquisition
+system).
+
+We have provided a version of the manuscript processed with
+"latexdiff" to highlight all of the differences between the original
+submitted (removed text in red) and the newly resubmitted version
+(added text in blue) with the supplementary material
+(diff_submission_resubmission.pdf). In addition, the new versions of
+the companion papers, Holland et al. (scuba2_general_paper.pdf), and
+Dempsey et al. (sc2calib_final.pdf), have also been provided.
+
+Detailed responses to the editor's and referee's comments are
+provided below, preceded by "-".
 
@@
 
-- Fig.4 had been modified, using a smaller number of sub-windows and a
+- Fig.4 has been modified, using a smaller number of sub-windows and a
   larger font it is now more readable.
@@
 
-- Fig.9 is now Fig. 10, and now has axis labels and an improved
-  caption - however, we elected to leave in the "broken" bolometers as
-  this is a fairly typical portion of the focal plane (remembering
-  that there is a ~70% yield); this gives the reader a good visual
-  sense of that.
+- Fig.9 is now Fig. 10, and has axis labels and an improved caption -
+  however, we elected to leave in the "broken" bolometers as this is a
+  fairly typical portion of the focal plane (remembering that there is
+  a ~70% yield); this gives the reader a good visual sense of that.
 
@@
 
-- We would like to thank the thorough reviewer for identfying several
-  areas that needed clarification, and perhaps more importantly, three
-  careless errors (which led to incorrect units in Figs.1 & 2, and
-  confusing results in Figs.11 & 20), and an ommission (impact /
-  treatment of the response of the bolometers and data acquisition
-  system). Detailed responses follow.
-
 Overall: The paper will be a useful reference for SMURF users and clearly represents a lot of work.
@@
   Section 3 is of the "baseline" parameters for the reduction, which
-  are then added-to for specific reduction, as in the examples from
+  are then added-to for specific reductions, as in the examples from
   section 4. The word "default" has been replaced with "baseline"
@@
 - Agreed, and re-worded so that SCUBA-2 is now "the largest of a new
-  generation of multiplexed kilopixel bolo cmaers"
+  generation of multiplexed kilopixel bolo cameras"
 
@@
 
-- Done (although used earlier Fowler 2010 as a reference).
+- Done (although used earlier Fowler et al. 2010 paper as a reference).
 
@@
 
-- added a setence describing the machine used for analysis in the
+- added a sentence describing the machine used for analysis in the
   paper: 64-bit, 8 cores, 2.7GHz, 48 Gb of RAM.
@@
 
-- This was a fairly major ommission in the first version. This is an
+- This was a fairly major omission in the first version. This is an
   effect that we considered long ago, and had been ignoring due to

### 2012-12-11 resubmit_mnras/smurf_paper.tex
ready for resubmission

--- 2012-12-06 smurf_paper.tex
@@
 
-\usepackage[pdftex,pdfpagemode={UseOutlines},bookmarks,bookmarksopen,colorlinks,linkcolor={blue},citecolor={green},urlcolor={red}]{hyperref}
-
-\usepackage[pdftex]{graphicx}
+%\usepackage[pdftex,pdfpagemode={UseOutlines},bookmarks,bookmarksopen,colorlinks,linkcolor={blue},citecolor={green},urlcolor={red}]{hyperref}
+
+%\usepackage[pdftex]{graphicx}
 
@@
 
-%\usepackage[bookmarks,bookmarksopen,colorlinks,linkcolor={blue},citecolor={green},urlcolor={red}]{hyperref}
+\usepackage[bookmarks,bookmarksopen,colorlinks,linkcolor={blue},citecolor={green},urlcolor={red}]{hyperref}
 
@@
   common-mode signals.  (a) Data taken during the S2SRO period,
-  observation 29 on 2010-03-13. There is a strong correlation between
-  the bolometers and the roughly $\sim$25\,s oscillation in the
-  fridge, but only a minor correlation with the opacity and telescope
-  motion. The nearly flat common-mode subtracted signals show: (i)
-  that most of the low-frequency signal is common to all of the
-  bolometers; and (ii) the non-correlated, and predominantly white
+  observation 29 on 2010 March 13. There is a strong correlation
+  between the bolometers and the roughly $\sim$25\,s oscillation in
+  the fridge, but only a minor correlation with the opacity and
+  telescope motion. The nearly flat common-mode subtracted signals
+  show: (i) that most of the low-frequency signal is common to all of
+  the bolometers; and (ii) the non-correlated, and predominantly white
   noise at 450\,\micron\ is significantly larger than at
   850\,\micron. (b) Data taken with the fully-commissioned instrument,
-  observation 38 on 2011-11-12. Unlike the S2SRO data, there is no
-  strong signal produced by variations in the fridge temperature. The
-  450\,\micron\ data have minimal correlated low-frequency noise (as
-  evidenced by the lack of a difference between the raw and
-  common-mode subtracted signals), although the 850\,\micron\ data
-  show a common signal correlated with the opacity and telescope
-  motion (note that this scan has a significantly larger amplitude
-  than the S2SRO data set), mostly likely caused by a combination of
-  changing airmass (anti-correlated with opacity), and magnetic field
-  pickup (see Section~\ref{sec:magpickup} and
+  observation 38 on 2011 November 12. Unlike the S2SRO data, there is
+  no strong signal produced by variations in the fridge
+  temperature. The 450\,\micron\ data have minimal correlated
+  low-frequency noise (as evidenced by the lack of a difference
+  between the raw and common-mode subtracted signals), although the
+  850\,\micron\ data show a common signal correlated with the opacity
+  and telescope motion (note that this scan has a significantly larger
+  amplitude than the S2SRO data set), mostly likely caused by a
+  combination of changing airmass (anti-correlated with opacity), and
+  magnetic field pickup (see Section~\ref{sec:magpickup} and
   Fig.~\ref{fig:magpickup}).}
@@
 
-In the S2SRO data (Fig.~\ref{fig:bolos_mix}a), observation 29 on
-2010-03-13, both bolometers share significant long-timescale structure
+In the S2SRO data (Fig.~\ref{fig:bolos_mix}a), observation 29 on 2010
+March 13, both bolometers share significant long-timescale structure
 ($\gsim10$\,s) that appears to be related to variations in the fridge
@@
 
-Data from the fully commissioned instrument, observation 38 on
-2011-11-12, are shown in Fig.~\ref{fig:bolos_mix}b. Having solved the
+Data from the fully commissioned instrument, observation 38 on 2011
+November 12, are shown in Fig.~\ref{fig:bolos_mix}b. Having solved the
 fridge oscillation problem, these data no longer exhibit a correlation
@@
 \caption{Evidence for significant magnetic field pickup for
-  observation 16 on 2010-02-28 at 450\,\micron\ (s4a subarray).  The top
-  panel shows two un-flatfielded (but mean-subtracted and step
+  observation 16 on 2010 February 28 at 450\,\micron\ (s4a subarray).
+  The top panel shows two un-flatfielded (but mean-subtracted and step
   corrected) bolometer time-series from the same column, with a 200
@@
 Example data from the 450\,\micron\ subarray s4a where pickup appears
-to be significant (observation 16 on 2010-02-28) are shown in
+to be significant (observation 16 on 2010 February 28) are shown in
 Fig.~\ref{fig:magpickup}. The time-series for two bolometers in the
@@
   3.6 to 7.7) from the fully-commissioned instrument, observation 18
-  on 2011-03-24. Red and blue lines show the 850 and 450\,\micron\
+  on 2011 March 24. Red and blue lines show the 850 and 450\,\micron\
   common-mode signals, corresponding to the left- and right- vertical
@@
 450\,\micron\ line-of-sight opacities $\sim$0.9 and 6.5 in the S2SRO
-and post-upgrade data sets, respectively), and are dominated at
-low frequencies ($\lsim$2\,Hz) by scan-synchronous noise, rather than
+and post-upgrade data sets, respectively), and are dominated at low
+frequencies ($\lsim$2\,Hz) by scan-synchronous noise, rather than
 underlying variations in the atmosphere, or ``sky noise''. However,
@@
 an order-of-magnitude longer (fully-commissioned) data set is plotted
-from observation 18 on 2011-03-24, during highly variable weather
+from observation 18 on 2011 March 24, during highly variable weather
 conditions (450\,\micron\ line-of-sight opacities ranging from 3.6 to
@@
 therefore the emission detected at 450\,\micron\ comes from water
-vapour that is closer to the telescope.
+vapour that is closer to the telescope than that observed at the
+longer wavelengths.
 
@@
 in astronomically-useful units.  In this example we reduce a
-450\,\micron\ map of Uranus (observation 26 on 2011-10-17), which is a
-nearly point-like source for \scuba\ that is commonly used as a
+450\,\micron\ map of Uranus (observation 26 on 2011 October 17), which
+is a nearly point-like source for \scuba\ that is commonly used as a
 primary flux calibrator. The CV daisy pattern was used, with a scan
@@
 240\,arcsec\,s$^{-1}$, covering an area of about 50\,arcmin$^2$. The
-full list of dates and observation numbers includes: 2010-02-18, 63, 64,
-65, 70, 71, 72, 90, 91, 92, 97, 98, 99; 2010-02-20, 111, 112, 113, 118,
-128, 129, 130; 2010-03-03, 61, 62, 64, 69, 70, 72, 73, 74, 75; 2010-03-09,
-87, 88, 90, 91; and 2010-03-11, 59, 64, 65, 72. These observations
-represent about 80 per cent of the total data taken as part of the
-project; the remaining observations were dropped due to problems
-keeping the arrays properly tuned during S2SRO, and were easily
-identified by their highly variable and erratic bolometer time-series
-(which resulted in maps full of large streaks).
+full list of dates and observation numbers includes: 2010 February 18,
+63, 64, 65, 70, 71, 72, 90, 91, 92, 97, 98, 99; 2010 February 20, 111,
+112, 113, 118, 128, 129, 130; 2010 March 3, 61, 62, 64, 69, 70, 72,
+73, 74, 75; 2010 March 9, 87, 88, 90, 91; and 2010 March 11, 59, 64, 65,
+72. These observations represent about 80 per cent of the total data
+taken as part of the project; the remaining observations were dropped
+due to problems keeping the arrays properly tuned during S2SRO, and
+were easily identified by their highly variable and erratic bolometer
+time-series (which resulted in maps full of large streaks).
 
@@
 In this final example, we analyse a map of M17 which contains bright,
-extended emission. The data are from observation 11 on 2011-05-31 using
-the 850\,\micron\ array. It is a rotating PONG scan covering a
+extended emission. The data are from observation 11 on 2011 May 31
+using the 850\,\micron\ array. It is a rotating PONG scan covering a
 diameter of 0.375\,deg, with a scan speed of 300\,arcsec\,s$^{-1}$,
@@
 Friberg and Gary Davis for their helpful comments on the
-manuscript. Finally, we thank the anonymous referee for a very helpful
+manuscript. Finally, we thank the anonymous referee for their thorough
 review that helped to clarify several key areas of the paper.

### 2013-01-07 smurf_paper.tex
address minor comments from second referee report

@@
     \model{EXT} & extinction correction \\
-    \model{FLT} & Fourier transform filter \\
+    \model{FLT} & Fourier Transform filter \\
     \model{AST} & map estimate of astronomical signal \\
@@
 
-\subsubsection{\model{FLT}: Fourier transform filter}
+\subsubsection{\model{FLT}: Fourier Transform filter}
 \label{sec:flt}
@@
   shape of a source in panel (a) from
-  Fig.~\ref{fig:lockman_maps}. Applying the whitening filter
-  (reciprocal of the solid orange line in
-  Fig.~\ref{fig:lockman_pspec}) to the dotted line gives the whitened
-  line (dashed), which is the expected shape of a source in panel (b)
-  from Fig.~\ref{fig:lockman_maps}. Finally, cross-correlating the
-  whitened map with this whitened PSF is an effective ``matched
-  filter'' for identifying point-like sources, and this smoothed map
-  is shown in panel (c) of Fig.~\ref{fig:lockman_maps}.}
+  Fig.~\ref{fig:lockman_maps}. Applying the whitening filter to the
+  dotted line (dividing the Fourier Transform of the map filtered
+  response by the square root of the solid orange line in
+  Fig.~\ref{fig:lockman_pspec}, and then transforming back to real
+  space) gives the whitened line (dashed), which is the expected shape
+  of a source in panel (b) from Fig.~\ref{fig:lockman_maps}. Finally,
+  cross-correlating the whitened map with this whitened PSF is an
+  effective ``matched filter'' for identifying point-like sources, and
+  this smoothed map is shown in panel (c) of
+  Fig.~\ref{fig:lockman_maps}.}
 \label{fig:lockman_psf}
@@
 
-To suppress noise in the map, we construct a whitening filter from the
-reciprocal of the jackknife angular power spectrum (orange line in
+To suppress noise in the map, we construct a whitening filter whereby
+the Fourier Transform of the map is divided by the square root of the
+jackknife power spectrum (orange line in
 Fig.~\ref{fig:lockman_pspec}), normalised by the white-noise level
 estimated from the RMS power at angular frequencies $>
-0.1$\,arcsec$^{-1}$. The filter is applied by scaling the power
-spectrum of the map by this function, and then transforming back into
-real space. The whitened map is shown in Fig.~\ref{fig:lockman_maps}b,
-and for comparison, the jackknife map has also been whitened in
+0.1$\,arcsec$^{-1}$, and transforming back into real space. The
+whitened map is shown in Fig.~\ref{fig:lockman_maps}b, and for
+comparison, the jackknife map has also been whitened in
 Fig.~\ref{fig:lockman_maps}e. In both cases, the maps are visibly
@@
 been used extensively in the submillimetre community, as both an
-efficient source-detection and photometric measurement strategy. In
-terms of the angular power spectrum, a matched filter may also be
-thought of as an optimal low-pass filter, that suppresses noise on
-scales smaller (and therefore at higher angular frequencies) than the
-beam. For the case at hand, we may use this cross-correlation
-technique since the map has been whitened. However, we must first
-establish the effective shape of point sources in this map due both to
-map-making itself, and the whitening filter. We determine the effect
-of map-making by adding a synthetic (and high-\snr) point source to
-the real data, and measure its resulting shape in the map (solid and
+efficient source-detection and photometric measurement strategy.  For
+the case at hand, we may use this cross-correlation technique since
+the map has been whitened. However, we must first establish the
+effective shape of point sources in this map due both to map-making
+itself, and the whitening filter. We determine the effect of
+map-making by adding a synthetic (and high-\snr) point source to the
+real data, and measure its resulting shape in the map (solid and
 dotted lines in Fig.~\ref{fig:lockman_psf}, respectively). Next, the
-Fourier transform of the map filtered PSF is divided by the jackknife
-noise power spectrum to calculate the final whitened PSF (dashed line
-in Fig.~\ref{fig:lockman_psf}). Maps smoothed by this shape are shown
-for the whitened signal, and jackknife maps in
-Figs.~\ref{fig:lockman_maps}c and f, respectively. Note that these
-images are plotted in \snr\ units, where the smooth noise maps have
-been calculated by propagating the original noise maps output by SMURF
-both through the whitening and matched filters (each of which are
-linear operations).
+Fourier Transform of the map filtered PSF is divided by the square
+root of the jackknife noise power spectrum to calculate the final
+whitened PSF (dashed line in Fig.~\ref{fig:lockman_psf}). Maps
+smoothed by this shape are shown for the whitened signal, and
+jackknife maps in Figs.~\ref{fig:lockman_maps}c and f,
+respectively. Note that these images are plotted in \snr\ units, where
+the smooth noise maps have been calculated by propagating the original
+noise maps output by SMURF both through the whitening and matched
+filters (each of which are linear operations). In terms of the angular
+power spectra, this complete process can be thought of as an optimal
+band-pass filter that has both suppressed low-frequency noise, and
+information on scales that are smaller (higher frequencies) than the
+beam.
 

### 2013-01-07 smurf_paper.tex
a couple of typos

@@
 anisotropic filtering in its maps can be accounted for during analysis
-to achieve SPTs focussed objectives \citep[in contrast, the similar
+to achieve SPT's focussed objectives \citep[in contrast, the similar
 ACT experiment uses cross-linking to improve its response to
@@
 (including a small region about the steps, Section~\ref{sec:steps});
-and 0.83\% if the data are flagged because the telescope was
+and 0.83\% of the data are flagged because the telescope was
 stationary (Section~\ref{sec:flagbadbol}). Once the iterative solution

### 2013-01-07 resubmit_mnras2/response.txt
resubmission 2

To whom it may concern,

The reviewer only requested minor changes to one section of the
paper. We agree with their suggestions and have implemented them.

We have provided a version of the manuscript processed with
"latexdiff" to highlight the differences between the previous version
(removed text in red) and the newly resubmitted version (added text in
blue) with the supplementary material
(diff_resubmission_resubmission2.pdf). As the reviewer did not request to
see this new version, we have not uploaded new versions of the
companion papers (Holland et al., and Dempsey et al.), although this
can easily be done if requested.

Detailed responses to the referee's comments are
provided below, preceded by "*".

Dear Dr Chapin

I attach the reviewer's comments on your revised manuscript entitled "SCUBA-2: iterative map-making with the Sub-Millimetre User Reduction Facility", ref. MN-12-2293-MJ.R1, which you submitted to Monthly Notices of the Royal Astronomical Society.

Some further, minor revision of your manuscript is requested before it is reconsidered for publication.

You should submit your further revised version, together with your response to the reviewer's comments via the Monthly Notices ScholarOne Manuscripts site http://mc.manuscriptcentral.com/mnras .
Enter your Author Centre, where you will find your manuscript title listed under "Manuscripts with Decisions." Under "Actions," click on "Create a Revision." Your manuscript reference will be appended to denote a revision.

IMPORTANT: do not submit your revised manuscript as a new paper!

You will not be able to make your revisions to the originally submitted files of the manuscript on ScholarOne Manuscripts. Instead, you must delete the original files and abstract and replace them with your revised files. Carefully proof read the resulting PDF and HTML files that are generated. If you have used a .bib file to generate your bibliography in Latex, please include this in your .tar archive along with the .bbl and .tex files; this will aid the editing and typesetting process.

When submitting your revised manuscript, you will be able to respond to the comments made by the reviewer in the space provided. You should also use this space to document any changes you make to the original manuscript. In order to expedite the processing of the revised manuscript, please be as specific as possible in your response to the reviewer. It would also be very helpful if you could highlight the changed sections, e.g. by the use of colour, colour highlighting or bold typeface - this will not delay in any way the subsequent processing of your paper.

Because we are trying to facilitate timely publication of manuscripts submitted to MNRAS, your revised manuscript should be uploaded promptly. If you do not submit your revision within six months, we may consider it withdrawn and request it be resubmitted as a new submission.

Please note that, due to the tight schedule, any post-acceptance changes notified after the paper has gone into production (i.e. the day after the acceptance email is sent) cannot be incorporated into the paper before it is typeset. Such changes will therefore need to be made as part of the proof corrections. To avoid excessive proof corrections and the delay that these can cause, you are strongly encouraged to ensure that each version of your paper submitted to MNRAS is completely ready for publication!

I look forward to receiving your revised manuscript.

Regards,

Keith

-- 
Dr Keith T. Smith

Assistant Editor (MNRAS)
Royal Astronomical Society
Burlington House
Piccadilly
London W1J 0BQ

cc: all S1M-listed co-authors.

Reviewer's Comments:

I commend the authors on a careful revision. I am happy the first review was useful. Reading through the corrections and the new revision, nearly everything makes sense to me now. The new section 3.4 gives a useful snapshot of reality to the SMURF user. The new atmosphere descriptions are also illuminating. Sorry to hear about the less than useful DKS (added text appreciated). The confusion in section 4 over the masking, filtering and degeneracies is all cleared up nicely.

I had only one hang-up in the matched filter section (the history of which is reproduced and expounded on below). Given your careful attention to the first review, I'm happy to leave this to your good judgement and do not require another look (unless you request it).

+ Sec 4.2, Para 8: "For the case at hand, the effective PSF for the whitened maps is given by the dotted curve in Fig. 13."

This is confusing language. Fig 13 list the dashed line as "whitened". It is true though that the dotted curve gives the correct PSF for your matched filter, since it's the one that corresponds to the pre-filtered beam. The matched filter in Fourier space is B(k)N^-1(k), where the N^-1(k) is your pre-whitening step (dividing by the power spectrum of the noise). The B(k) is the PSF before the whitening. It seems like you say this here, but it's a bit confusing with the figure labeling. Maybe it would be useful to include a simple equation like \Phi(k) = B(k)N^-1(k) or similar to describe your total operation?

- Yes, we are essentially breaking down the typical B(k)/N^-1(k)
matched-filter expression into two steps. We've clarified the text
here.

-- I follow this, though there may still be a confusing bit. In the added text: "Next, the Fourier transform of map filtered PSF is divided by the jackknife noise power spectrum to calculate the final whitened PSF (dashed line in Fig. 14)." In whitening the map you say "The filter is applied by scaling the power spectrum of the map by this function and then transforming back into real space." In order to retain phase information in the map, I'm guessing this sentence means you take the FT of the map and divide by the sqrt of the power spectrum of the jack knife and the transform back? The PS of such a map would in theory be white if it were all noise. If this is the case then I would have guessed that the "map filtered PSF is divided by the [sqrt of the] jackknife noise power spectrum to calculate the final whitened PSF". Taking N(k) as the power spectrum of the noise (or equivalently jackknife estimate of noise), this gives the right form of a matched filter B(k)/N(k) where a factor of 1/sqrt(N) comes from the whitening and another factor of B(k)/sqrt(N) comes from the "PSF" filtering. The only other qualm that comes to mind is that by including the 1/sqrt(N) in the PSF "matched filter" step, the claim in the text that it is an optimal *low-pass* filter is not really true. Since 1/sqrt(N) suppresses large scales for a "red" N, the matched filter is a *band-pass* filter. Said in another way, you have to start with a map with white noise to have the matched filter be a simple PSF that low-pass filters the map (presumably as would be the case in most optical datasets). One can't tag on a pre-whitening step to data with non-white noise and describe the match filter of the "whitened" map as a simple smoothing filter.

* Yes, as you point our we are in fact dividing the Fourier Transform
  coefficients by the square-root of the jackknife estimate of the
  noise power spectrum. This is now written correctly in the caption
  of Fig.14, as well as paragraphs 6 and 7 of Section 4.2.

* We have also changed and moved the text about "optimal low pass
  filtering" to the end of paragraph 7, and instead refer to it as a
  "band pass" filter to summarize the entire operation, as you have
  correctly suggested.

### 2013-01-15 arXiv_submission/smurf_paper.tex
minor fiddles for arXiv submission

--- 2013-01-07 smurf_paper.tex
@@
 %------------------------------------------------------------------------------
-
+\pdfoutput=1
 \documentclass[useAMS,usenatbib,nofootinbib]{mn2e}
@@
 The Submillimetre Common User Bolometer Array 2
-\citep[\scuba,][]{holland2012} is a new instrument for the 15-m James
+\citep[\scuba,][]{holland2013} is a new instrument for the 15-m James
 Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
@@
 instrument design, performance, and calibration are given in two
-companion papers: \citet{holland2012} and \citet{dempsey2012}.
+companion papers: \citet{holland2013} and \citet{dempsey2013}.
 
@@
 formation.  With $10,240$ nominal detectors \citep[of which $\sim$70\%
-work and are typically useful,][]{holland2012}, \scuba\ is presently
+work and are typically useful,][]{holland2013}, \scuba\ is presently
 the largest of a new generation of multiplexed kilopixel
@@
 While the details of how \scuba\ works are described in
-\citet{holland2012}, and its calibration in \citet{dempsey2012}, we
+\citet{holland2013}, and its calibration in \citet{dempsey2013}, we
 summarise the basic concepts relevant to map-making here.
@@
 of the pixel heaters (which provide a known input power); see
-Section~2.1 in \citet{dempsey2012} and Section~5.3 in
-\citet{holland2012} for more details. Finally, the conversion to
+Section~2.1 in \citet{dempsey2013} and Section~5.3 in
+\citet{holland2013} for more details. Finally, the conversion to
 astronomical flux units from power involves a correction for
@@
 JCMT Water Vapour Monitor to track line-of-sight opacity variations,
-see Section~3 in][]{dempsey2012}, and the application of a flux
+see Section~3 in][]{dempsey2013}, and the application of a flux
 conversion factor (FCF) which is established from regular measurements
-of calibrators such as Uranus \citep[Section~5 in][]{dempsey2012}.
+of calibrators such as Uranus \citep[Section~5 in][]{dempsey2013}.
 
@@
 thermometers in the focal planes has effectively mitigated this
-problem \citep[Section~2.5 in][]{holland2012}. In addition, there were
+problem \citep[Section~2.5 in][]{holland2013}. In addition, there were
 improvements to the magnetic shielding (reducing magnetic field
@@
 small circle. For a more complete description of the \scuba\ observing
-modes, see Section~5 in \citet{holland2012} and also
+modes, see Section~5 in \citet{holland2013} and also
 \citet{2010SPIE.7740E..66K}.
@@
 they are multiplied by the flatfield correction \citep[see Section~2.1
-in][]{dempsey2012}. Next, a series of configurable data cleaning and
+in][]{dempsey2013}. Next, a series of configurable data cleaning and
 filtering procedures are applied; these include the removal of large
@@
 with the most likely cause being cosmic ray events \citep[see
-Section~3.5.3 in][]{holland2012}. The black curves in
+Section~3.5.3 in][]{holland2013}. The black curves in
 Fig.~\ref{fig:steps1} show examples of such steps in the time-series
@@
 final map.  For details on how it is calculated see
-\citet{dempsey2012}. Note that its numerical value is calculated only
+\citet{dempsey2013}. Note that its numerical value is calculated only
 once, and simply applied as a scale factor in the iterative
@@
 the \rms\ time-series noise in a 1\,s integration \citep[see section
-3.5.2 in][]{holland2012}. The NEPs are then converted into more useful
+3.5.2 in][]{holland2013}. The NEPs are then converted into more useful
 noise equivalent flux densities (NEFDs, with units Jy\,s$^{1/2}$)
@@
 respectively, at an average airmass of 1.2, from table 3
-in][]{holland2012}, the mapping speed is approximately
+in][]{holland2013}, the mapping speed is approximately
 $6\times10^{-5}$ and $3\times10^{-3}$\,deg$^2$\,hr$^{-1}$\,mJy$^{-2}$,
@@
 for \scuba\ of $(492 \pm 67)$\,Jy\,pW$^{-1}$ \citep[see][for
-details]{dempsey2012}. Since the map is now flat away from the source,
+details]{dempsey2013}. Since the map is now flat away from the source,
 and constrained to a value of zero, it is appropriate for performing

### 2013-03-25 arXiv_submission/smurf_paper.tex
Paper published. Updated and new version sent to arXiv.

@@
 % common acronyms etc.
-\newcommand{\snr}{SNR}
+\newcommand{\snr}{S/N}
 \newcommand{\scuba}{SCUBA-2}
-\newcommand{\rms}{RMS}
+\newcommand{\rms}{rms}
 
@@
 filters, providing simultaneous illumination and wavelength definition
-of both the 450 and 850\,\micron\ focal planes. Each focal plan is
-populated by four ``subarrays'' (labelled s4a--s4d at 450\,\micron, and
-s8a--s8d at 850\,\micron), each consisting of 32 columns and 40 rows
-of bolometers. The bolometers are thermal absorbers (with a response
-time constant of $\tau \sim 1$\,ms) coupled to superconducting
-transition-edge sensors (TESs) for thermometry. Temperature variations
-in the TESs produce changing currents, and therefore varying magnetic
-fields, which are detected and amplified using chains of
-superconducting quantum interference devices (SQUIDs), before the
-larger output currents are digitised. Each detector has its own SQUID
-for the first-stage of the amplification, but the remaining stages
-occur within a common chain of SQUIDs for each of the 32 columns. All
-40 rows are read out in sequence at a row-visit frequency of about
-12\,kHz. Such a high sample rate is unnecessary to produce maps, so
-the data are re-sampled to approximately 200\,Hz before writing to
-disk. This rate provides a sample every 3.0\,arcsec, or approximately
-one third of the 450\,\micron\ diffraction-limited full-width at
-half-maximum (FWHM) -- a typical rule-of-thumb for adequately sampling
-a Gaussian point spread function (PSF) -- at the maximum scanning
-speed of 600\,arcsec\,s$^{-1}$. There is an additional 41st row of
-SQUID readouts that are not connected to TESs. These ``dark SQUIDs''
-track non-thermal noise sources that are common to each column's
-amplifier chain. The relationship between the output digitised
-current, $I$, and the input power, $P$, or $dI/dP$ is established
-using flatfield observations immediately prior to science
-observations, in which the output signal is measured throughout a ramp
-of the pixel heaters (which provide a known input power); see
-Section~2.1 in \citet{dempsey2013} and Section~5.3 in
+of both the 450 and 850\,\micron\ focal planes. Each focal plane is
+populated by four ``subarrays'' (labelled s4a--s4d at 450\,\micron,
+and s8a--s8d at 850\,\micron), each consisting of 32 columns and 40
+rows of bolometers. The bolometers are thermal absorbers (with a
+response time constant of $\tau \sim 1$\,ms) coupled to
+superconducting transition-edge sensors (TESs) for
+thermometry. Temperature variations in the TESs produce changing
+currents, and therefore varying magnetic fields, which are detected
+and amplified using chains of superconducting quantum interference
+devices (SQUIDs), before the larger output currents are
+digitised. Each detector has its own SQUID for the first-stage of the
+amplification, but the remaining stages occur within a common chain of
+SQUIDs for each of the 32 columns. All 40 rows are read out in
+sequence at a row-visit frequency of about 12\,kHz. Such a high sample
+rate is unnecessary to produce maps, so the data are re-sampled to
+approximately 200\,Hz before writing to disk. This rate provides a
+sample every 3.0\,arcsec, or approximately one third of the
+450\,\micron\ diffraction-limited full-width at half-maximum (FWHM) --
+a typical rule-of-thumb for adequately sampling a Gaussian point
+spread function (PSF) -- at the maximum scanning speed of
+600\,arcsec\,s$^{-1}$. There is an additional 41st row of SQUID
+readouts that are not connected to TESs. These ``dark SQUIDs'' track
+non-thermal noise sources that are common to each column's amplifier
+chain. The relationship between the output digitised current, $I$, and
+the input power, $P$, or $dI/dP$ is established using flatfield
+observations immediately prior to science observations, in which the
+output signal is measured throughout a ramp of the pixel heaters
+(which provide a known input power); see Section~2.1 in
+\citet{dempsey2013} and Section~5.3 in
 \citet{holland2013} for more details. Finally, the conversion to
@@
 (and therefore a mapping speed reduced by a factor of $\sim$4),
-instabilities in the fridge led to a large periodic signal with a
+instabilities in the fridge led to a large oscillating signal with a
 period of about 25\,s that was correlated amongst the detectors. After
@@
   corrected) bolometer time-series from the same column, with a 200
-  sample boxcar smooth (approximately 1\,s), illustrating that they
+  sample boxcar smoothing (approximately 1\,s), illustrating that they
   are dominated by a similar signal with opposite signs. The second
   panel shows the dark squid signal for the column, also
-  mean-subtracted and with the same boxcar smooth. The bottom panel
+  mean-subtracted and with the same boxcar smoothing. The bottom panel
   shows the azimuthal and elevation offsets from the map centre (mean
@@
 value decomposition identifies a new set of statistically uncorrelated
-eigenvectors, $\mathbf{\xi}_i(t)$ (i.e., whose covariance matrix is
+basis vectors, $\mathbf{\xi}_i(t)$ (i.e., whose covariance matrix is
 diagonal), such that each of the bolometer time-series is a linear
-combination of the eigenvectors, or components i.e., $\mathbf{b}_i(t)
+combination of the basis vectors, or components i.e., $\mathbf{b}_i(t)
 = \bar{\mathbf{\xi}} \mathbf{\lambda}_i^\mathrm{T}$, where each row of
-the matrix $\bar{\mathbf{\xi}}$ is an eigenvector, and
+the matrix $\bar{\mathbf{\xi}}$ is a basis vector, and
 $\mathbf{\lambda}_i^\mathrm{T}$ is a column vector containing the
-corresponding eigenvalues. The $\mathbf{\xi}_i(t)$ are normalised by
+corresponding amplitudes. The $\mathbf{\xi}_i(t)$ are normalised by
 $[\sum_t \mathbf{\xi}_i^2(t)]^{1/2}$, such that the amplitude of each
-component in a given bolometer signal is stored in the eigenvalues. In
-the earlier analyses mentioned, the low-frequency noise is assumed to
-be encapsulated in those components with the largest
-eigenvalues. Removing the projection of the time series along those
-components then significantly reduces $1/f$ noise while retaining most
-of the (higher-frequency) signal in point-like sources
+component is proportional to its \rms. In the earlier analyses
+mentioned, the low-frequency noise is assumed to be encapsulated in
+those components with the largest eigenvalues. Removing the projection
+of the time series along those components then significantly reduces
+$1/f$ noise while retaining most of the (higher-frequency) signal in
+point-like sources
 
@@
 \ref{fig:pspec}b. The nine most significant components are shown
-(ranked by the mean eigenvalues across all working bolometers in both
+(ranked by the mean amplitudes across all working bolometers in both
 bands). The top and middle panels for each component show the
-time-series and PSDs of the normalised eigenvectors. The bottom panels
-are maps indicating the eigenvalues (amplitudes) of the component
-across the focal plane (with the mean $\bar{\lambda}$ and and \rms,
-$\lambda_\mathrm{rms}$, of the eigenvalues for all of the bolometers
-calculated separately at each wavelength also shown).
+time-series and PSDs of the normalised basis vectors. The bottom
+panels are maps indicating the significance (amplitudes) of the
+component across the focal plane (with the mean $\bar{\lambda}$ and
+and \rms, $\lambda_\mathrm{rms}$, of the amplitudes for all of the
+bolometers calculated separately at each wavelength also shown).
 
 The majority of the correlated signal at both wavelengths in
-Fig.~\ref{fig:pca} is produced by Component~1, in which the
-eigenvector time-series exhibits a roughly periodic signal that
-resembles the scan pattern in Fig.~\ref{fig:bolos_mix}b. While this
-correlated signal appears at both wavelengths, referring to the maps
-of eigenvalues, they are considerably stronger at 850\,\micron\
+Fig.~\ref{fig:pca} is produced by Component~1, in which the basis
+vector time-series exhibits a roughly periodic signal that resembles
+the scan pattern in Fig.~\ref{fig:bolos_mix}b. While this correlated
+signal appears at both wavelengths, referring to the maps of
+amplitudes, they are considerably stronger at 850\,\micron\
 (consistent with the visual appearance of the bolometer signals in
-Fig.~\ref{fig:pca}). Also note that the PSD for this eigenvector
+Fig.~\ref{fig:pca}). Also note that the PSD for this basis vector
 exhibits nearly a pure $1/f$ signature with minimal line features, and
@@
 \caption{The first nine components from a principal component
-  analysis, ranked by the mean eigenvalues of combined 450 (s4a
+  analysis, ranked by the mean amplitudes of combined 450 (s4a
   subarray) and 850\,\micron\ (s8b subarray) time-series bolometer
@@
   Figs.~\ref{fig:bolos_mix}b and \ref{fig:pspec}b. For each component
-  the top plot shows the time-series of the eigenvectors normalised by
-  their \rms, the middle plot its PSD, and the bottom coloured panels
-  the eigenvalues for the bolometers at each wavelength (the
-  amplitudes of the eigenvectors in the time-series). For reference,
-  both the mean, $\bar{\lambda}$, and \rms, $\lambda_\mathrm{rms}$,
-  eigenvalues for the bolometers in each subarray are also shown. Most
-  of the correlated $1/f$ signal is encapsulated in Component~1, and
-  appears to be dominated by scan-synchronous noise (compare with the
-  scan pattern in Fig.~\ref{fig:bolos_mix}b). Some of this signal also
-  appears in Component~3. Component~2 is another $1/f$ signal that is
-  dominant in the 850\,\micron\ subarray. Most of the remaining
-  components consist of weaker $1/f$ noise and line features
-  predominantly above $\sim$30\,Hz. The eigenvalue maps show
-  correlated patterns along columns suggesting pickup in their common
-  amplifier chains.}
+  the top plot shows the time-series of the basis vectors normalised
+  by their \rms, the middle plot its PSD, and the bottom coloured
+  panels the ampliudes for the bolometers at each wavelength (the
+  contribution of the basis vectors to the time-series). For
+  reference, both the mean, $\bar{\lambda}$, and \rms,
+  $\lambda_\mathrm{rms}$, amplitudes for the bolometers in each
+  subarray are also shown. Most of the correlated $1/f$ signal is
+  encapsulated in Component~1, and appears to be dominated by
+  scan-synchronous noise (compare with the scan pattern in
+  Fig.~\ref{fig:bolos_mix}b). Some of this signal also appears in
+  Component~3. Component~2 is another $1/f$ signal that is dominant in
+  the 850\,\micron\ subarray. Most of the remaining components consist
+  of weaker $1/f$ noise and line features predominantly above
+  $\sim$30\,Hz. The amplitude maps show correlated patterns along
+  columns suggesting pickup in their common amplifier chains.}
 \label{fig:pca}
@@
 Why are there two components that seem to be dominated by
-scan-synchronous noise (which is presumably a mixture of
-magnetic field pickup and brightness variations in the atmosphere)?
-First, the smoothly-varying gradients in the eigenvalue maps
-(primarily at 850\,\micron) suggests that this noise has a different
-response across the focal plane; in this case the PCA has identified
-two orthogonal shapes that, when mixed in different quantities, can
-reproduce most of the signal in each bolometer. It is also likely that
-the atmospheric contribution is stronger at 450\,\micron\ (i.e.,
-predominantly Component~3).
+scan-synchronous noise (which is presumably a mixture of magnetic
+field pickup and brightness variations in the atmosphere)?  First, the
+smoothly-varying gradients in the amplitude maps (primarily at
+850\,\micron) suggests that this noise has a different response across
+the focal plane; in this case the PCA has identified two orthogonal
+shapes that, when mixed in different quantities, can reproduce most of
+the signal in each bolometer. It is also likely that the atmospheric
+contribution is stronger at 450\,\micron\ (i.e., predominantly
+Component~3).
 
@@
 block, and this fit is extrapolated to predict the data value expected
-at the centre of the block on the basis of the preceding data;
+at the centre of the block;
 
@@
 bolometer data just after the block, and this fit is extrapolated to
-predict the data value expected at the centre of the block on the
-basis of the following data;
+predict the data value expected at the centre of the block;
 
@@
 \item Padding with artificial data: instead of padding with zeros,
-each time-series is padded with artificial data that connects the two
+each time-series is padded with artificial data which connects the two
 ends of the data stream smoothly and includes Gaussian noise. No
@@
 significant components in the analysis are identified, and the
-projection of each bolometer time-series along their eigenvectors are
+projection of each bolometer time-series along their basis vectors are
 removed. A single parameter specifies the threshold on the amplitude
-of the eigenvalues to be removed, as a number of standard deviations
+of the components to be removed, as a number of standard deviations
 away from the mean value. Subarrays are cleaned independently, once,
@@
 converts this effective power to the digitised units recorded on disk
-(DAC) -- the flatfield multiplied by a digitisation constant (applied
-once during pre-processing) -- which in this formulation is assumed to
-be constant in time.
+-- the flatfield multiplied by a digitisation constant (applied once
+during pre-processing) -- which in this formulation is assumed to be
+constant in time.
 
@@
 $\mathbf{n}^\mathrm{f}_i(t)$ is (predominantly low-frequency) noise in
-excess of the white noise level, that is either uncorrelated from
+excess of the white noise level, which is either uncorrelated from
 bolometer-to-bolometer, or at least does not have a simple correlation
@@
 constraints to the map to improve convergence, which presently include
-setting user-specified or high-\snr\ regions to a value of zero for
+setting user-specified or low-\snr\ regions to a value of zero for
 all but the final iteration. See the discussion in
@@
 iterations are large However, it is also clear that the estimated
-\model{COM} and \model{AST} signals are complimentary. In other words,
+\model{COM} and \model{AST} signals are complementary. In other words,
 the large change in the \model{AST} signal is cancelled by freedom in
@@
 point-sources involves three major steps: (i) generating a map that
-removes most low-frequency noise sources with approximately linear
+removes most large-scale noise sources with approximately linear
 response, without prior knowledge of the location of sources; (ii)
@@
 In this section we reduce scans of the Lockman Hole taken during S2SRO
-as a pilot project for the Cosmology Legacy Survey. It consists of
-$\sim8.5$\,hours of data taken in 36 separate scans (average
-15\,min. each) spread over February and March 2010. Each scan is a
-360\,arcsec rotating PONG pattern, with a scan speed of
+as a pilot project for the SCUBA-2 Cosmology Legacy Survey. It
+consists of $\sim8.5$\,hours of data taken in 36 separate scans
+(average 15\,min. each) spread over February and March 2010. Each scan
+is a 360\,arcsec rotating PONG pattern, with a scan speed of
 240\,arcsec\,s$^{-1}$, covering an area of about 50\,arcmin$^2$. The
@@
 112, 113, 118, 128, 129, 130; 2010 March 3, 61, 62, 64, 69, 70, 72,
-73, 74, 75; 2010 March 9, 87, 88, 90, 91; and 2010 March 11, 59, 64, 65,
-72. These observations represent about 80 per cent of the total data
-taken as part of the project; the remaining observations were dropped
-due to problems keeping the arrays properly tuned during S2SRO, and
-were easily identified by their highly variable and erratic bolometer
-time-series (which resulted in maps full of large streaks).
+73, 74, 75; 2010 March 9, 87, 88, 90, 91; and 2010 March 11, 59, 64,
+65, 72. These observations represent about 80 per cent of the total
+data taken as part of the project; the remaining observations were
+dropped due to problems in keeping the arrays properly tuned during
+S2SRO, and were easily identified by their highly variable and erratic
+bolometer time-series (which resulted in maps full of large streaks).
 
@@
 root of the jackknife noise power spectrum to calculate the final
-whitened PSF (dashed line in Fig.~\ref{fig:lockman_psf}). Maps
-smoothed by this shape are shown for the whitened signal, and
-jackknife maps in Figs.~\ref{fig:lockman_maps}c and f,
-respectively. Note that these images are plotted in \snr\ units, where
-the smooth noise maps have been calculated by propagating the original
-noise maps output by SMURF both through the whitening and matched
-filters (each of which are linear operations). In terms of the angular
-power spectra, this complete process can be thought of as an optimal
-band-pass filter that has both suppressed low-frequency noise, and
-information on scales that are smaller (higher frequencies) than the
-beam.
+whitened PSF (dashed line in Fig.~\ref{fig:lockman_psf}).  Both the
+whitened signal and jackknife maps are smoothed by this shape and
+shown in Figs.~\ref{fig:lockman_maps}c and f, respectively. Note that
+these images are plotted in \snr\ units, where the smoothed noise maps
+have been calculated by propagating the original noise maps output by
+SMURF through both the whitening and matched filters (each of which is
+a linear operations). In terms of the angular power spectra, this
+complete process can be thought of as an optimal band-pass filter that
+has both suppressed low-frequency noise, and information on scales
+that are smaller (higher frequencies) than the beam.
 
@@
 required to produce a robust source lists. Additional tests, along
-with a careful consideration of completeness and false-positive rates
-usual require a series of Monte Carlo simulations that are beyond the
-scope of this work. We direct the interested reader to a selection of
-papers on the subject:
+with a careful consideration of completeness and false-positive rates,
+usually require a series of Monte Carlo simulations that are beyond
+the scope of this work. We direct the interested reader to a selection
+of papers on the subject:
 \citet{scott2002,coppin2006,perera2008,2009ApJ...707.1201W} and
@@
   good agreement indicates that these maps are indeed dominated by
-  white noise. The lower panel shows the results for matched-filtered
+  white noise. The lower panel shows the results for matched filtered
   signal and jackknife maps (Figs.~\ref{fig:lockman_maps}c,f). The
@@
 zero-masking depends on the \snr\ of the map, it will be restricted to
-regions approximately $\sqrt{2}$ shallower in these maps than for the
-full data set. Finally, the cross-linking (position angles sampled) is
-similar, though not identical in the two halves. Ideally one would
-have many full maps, as in the case of the Lockman Hole data in the
-previous section, from which alternating maps could be combined.
+regions approximately $\sqrt{2}$ times shallower in these maps than
+for the full data set. Finally, the cross-linking (position angles
+sampled) is similar, though not identical in the two halves. Ideally
+one would have many full maps, as in the case of the Lockman Hole data
+in the previous section, from which alternating maps could be
+combined.
 
@@
 emission is almost perfectly removed, except for a slight impression
-of M17 near the centre of the map, that is probably due to some
+of M17 near the centre of the map, which is probably due to some
 combination of detector gain and pointing drifts. Otherwise the
@@
 funds for the construction of \scuba\ were provided by the Canada
-Foundation for Innovation.  This research was supported in part by the
-Natural Sciences and Engineering Research Council of Canada.  EC
-thanks CANARIE/CANFAR for additional funding.  The authors thank the
-members of the SCUBA-2 commissioning team for testing the map-maker
-and reporting anomalies; in particular Antonio Chrysostomou and
-Jessica Dempsey.  We also thank Mark Halpern, Matthew Hasselfield, and
-Gaelen Marsden for many useful discussions; observers who provided
-helpful feedback, especially David Nutter and Todd McKenzie; and
-Mandana Amiri and Dan Bintley for consultations regarding the SCUBA-2
-Multi-Channel Electronics.  We acknowledge the contributions of Dennis
-Kelly, Alex van Engelen and Jennifer Balfour for early investigations
-related to SMURF; and Mark Thompson, Craig Walther and S\'{e}verin
-Gaudet for being on the Critical Design Review panel. We thank Per
-Friberg and Gary Davis for their helpful comments on the
-manuscript. Finally, we thank the anonymous referee for their thorough
-review that helped to clarify several key areas of the paper.
+Foundation for Innovation. This research used the facilities of the
+Canadian Astronomy Data Centre operated by the National Research
+Council of Canada with the support of the Canadian Space Agency.  This
+research was supported in part by the Natural Sciences and Engineering
+Research Council of Canada.  EC thanks CANARIE/CANFAR for additional
+funding.  The authors thank the members of the SCUBA-2 commissioning
+team for testing the map-maker and reporting anomalies; in particular
+Antonio Chrysostomou and Jessica Dempsey.  We also thank Mark Halpern,
+Matthew Hasselfield, and Gaelen Marsden for many useful discussions;
+observers who provided helpful feedback, especially David Nutter and
+Todd McKenzie; and Mandana Amiri and Dan Bintley for consultations
+regarding the SCUBA-2 Multi-Channel Electronics.  We acknowledge the
+contributions of Dennis Kelly, Alex van Engelen and Jennifer Balfour
+for early investigations related to SMURF; and Mark Thompson, Craig
+Walther and S\'{e}verin Gaudet for being on the Critical Design Review
+panel. We thank Per Friberg and Gary Davis for their helpful comments
+on the manuscript. Finally, we thank the anonymous referee for their
+thorough review that helped to clarify several key areas of the paper.
 
